### 2024-05-03 main.tex
Lit review additions

%% Package and Class "uiucthesis2021" for use with LaTeX2e.
\documentclass[edeposit,fullpage,11pt]{uiucthesis2021}
\doublespacing

\usepackage[acronym,toc]{glossaries}
\makeglossaries
\include{acros}

\usepackage{xspace}
\usepackage{graphicx}
\graphicspath{{images/}}
\usepackage[title,titletoc]{appendix}

\usepackage{placeins}
\usepackage{booktabs} % nice rules (thick lines) for tables
\usepackage{array}
\usepackage{microtype} % improves typography for PDF

\usepackage[hyphens]{url}
\usepackage[hidelinks]{hyperref}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{hhline}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{mathtools}
\allowdisplaybreaks
\usepackage{color}
\usepackage{multirow}
\usepackage{siunitx}
\usepackage{xfrac}
\usepackage{bm}
\usepackage{stmaryrd}

\usepackage{threeparttable, tablefootnote}

\usepackage{environ}
\makeatletter

\usepackage{tabularx}
\usepackage{float}
\usepackage{enumitem}
\usepackage{diagbox}
\usepackage{courier}
\usepackage{pdflscape}

\usepackage{cleveref}
\usepackage{datatool}
\usepackage[numbers]{natbib}
\usepackage{notoccite}

\usepackage{tikz}
\usepackage{tkz-euclide}
\usetikzlibrary{positioning, arrows, decorations, shapes}
\usetikzlibrary{shapes.geometric,arrows}
\definecolor{illiniblue}{HTML}{B1C6E2}
\definecolor{illiniorange}{HTML}{f8c2a2}
\definecolor{green}{HTML}{c2e2b1}
\tikzstyle{process} = [rectangle, rounded corners, thick, minimum width=7cm, minimum height=1cm, text centered, draw=black, fill=illiniblue, text width=18em]
\tikzstyle{object} = [ellipse, minimum width=2cm, thick, minimum height=2.2cm, text centered, draw=black, fill=illiniorange, text width=12em]
\tikzstyle{decision} = [diamond, thick, aspect=2, minimum width=2cm, minimum height=2cm, text centered, draw=black, fill=green, text width=6em]
\tikzstyle{arrow} = [thick,->,>=stealth]

\usepackage[ruled,linesnumbered,lined]{algorithm2e}
\SetArgSty{textrm}

\title{A Hybrid Neutronics Method for Time-Dependent Control Rod Modeling in Molten Salt Reactors}
\author{Sun Myung Park}
\department{Nuclear, Plasma \& Radiological Engineering}
\prelim
\degreeyear{2024}
\committee{Dr Madicken Munk, Chair \\
           Professor Rizwan Uddin \\
           Associate Professor Tomasz Kozlowski \\
           Professor Paul Fischer}

\begin{document}
\maketitle

\frontmatter
%% Create an abstract that can also be used for the ProQuest abstract.
%% Note that ProQuest truncates their abstracts at 350 words.
\begin{abstract}

%\input{abstract}

\end{abstract}
\addcontentsline{toc}{chapter}{Abstract}

%\begin{dedication}
%
%\end{dedication}

%\begin{acknowledgments}
%
%%\input{acks}
%
%\end{acknowledgments}
%\addcontentsline{toc}{chapter}{Acknowledgments}

%% The thesis format requires the Table of Contents to come
%% before any other major sections, all of these sections after
%% the Table of Contents must be listed therein (i.e., use \chapter,
%% not \chapter*).  Common sections to have between the Table of
%% Contents and the main text are:
%%
%% List of Tables
%% List of Figures
%% List Symbols and/or Abbreviations
%% etc.

{
  \hypersetup{linkcolor=black}
  \tableofcontents
}

%% Create a List of Abbreviations. The left column
%% is 1 inch wide and left-justified
\printglossary[title=List of Abbreviations,type=\acronymtype,nonumberlist,
nogroupskip=true]
%% Create a List of Symbols. The left column
%% is 0.7 inch wide and centered

\pagebreak
\mainmatter
\glsresetall

\chapter{Introduction}
\label{chap:intro}
%\input{introduction}
%\glsresetall

\chapter{Molten Salt Reactor Modeling and Simulation}
\label{chap:lit}
\input{lit}
\glsresetall
%
%\chapter{Description of Moltres and Assessment of its Current Capabilities}
%\label{chap:moltres}
%\input{moltres}
%\glsresetall
%
%\chapter{Verification of Moltres for Multiphysics Molten Salt
%Reactor Modeling}
%\label{chap:benchmark}
%\input{benchmark}
%\glsresetall
%
%\chapter{Hybrid S$_\text{N}$-Diffusion Method for Improved Control Rod Modeling
%in Neutron Diffusion Methods}
%\label{chap:hybrid}
%\input{hybrid}
%\glsresetall
%
%\chapter{Proposed Work}
%\label{chap:proposedwork}
%\input{proposed-work}
%\glsresetall

\backmatter

\bibliographystyle{IEEEtran}
\bibliography{bibliography}

\end{document}
\endinput
%%

### 2024-05-03 acros.tex

%\newacronym{<++>}{<++>}{<++>}
\newacronym[longplural={metric tons of heavy metal}]{MTHM}{MTHM}{metric ton of heavy metal}
\newacronym{ABM}{ABM}{agent-based modeling}
\newacronym{ACDIS}{ACDIS}{Program in Arms Control \& Domestic and International Security}
\newacronym{AEF}{AEF}{Averaged Eddington Factor}
\newacronym{AHTR}{AHTR}{Advanced High Temperature Reactor}
\newacronym{ANDRA}{ANDRA}{Agence Nationale pour la gestion des D\'echets RAdioactifs, the French National Agency for Radioactive Waste Management}
\newacronym{ANL}{ANL}{Argonne National Laboratory}
\newacronym{API}{API}{application programming interface}
\newacronym{ARDP}{ARDP}{Advanced Reactor Demonstration Program}
\newacronym{ARE}{ARE}{Aircraft Reactor Experiment}
\newacronym{ASME}{ASME}{American Society of Mechanical Engineers}
\newacronym{ATWS}{ATWS}{Anticipated Transient Without Scram}
\newacronym{BDBE}{BDBE}{Beyond Design Basis Event}
\newacronym{BIDS}{BIDS}{Berkeley Institute for Data Science}
\newacronym{BTE}{BTE}{Boltzmann transport equation}
\newacronym{BWR}{BWR}{boiling water reactor}
\newacronym{CAFCA}{CAFCA}{ Code for Advanced Fuel Cycles Assessment }
\newacronym{CAS}{CAS}{Chinese Academy of Sciences}
\newacronym{CDTN}{CDTN}{Centro de Desenvolvimento da Tecnologia Nuclear}
\newacronym{CEA}{CEA}{Commissariat \`a l'\'Energie Atomique et aux \'Energies Alternatives}
\newacronym{CFD}{CFD}{Computational Fluid Dynamics}
\newacronym{CI}{CI}{continuous integration}
\newacronym{CNEN}{CNEN}{Comiss\~{a}o Nacional de Energia Nuclear}
\newacronym{CNERG}{CNERG}{Computational Nuclear Engineering Research Group}
\newacronym{CNRS}{CNRS}{Centre National de la Recherche Scientifique}
\newacronym{COMSOL}{COMSOL}{COMmon SOLution}
\newacronym{COSI}{COSI}{Commelini-Sicard}
\newacronym{COTS}{COTS}{commercial, off-the-shelf}
\newacronym{CPM}{CPM}{collision probability method}
\newacronym{CSNF}{CSNF}{commercial spent nuclear fuel}
\newacronym{CTAH}{CTAHs}{Coiled Tube Air Heaters}
\newacronym{CUBIT}{CUBIT}{CUBIT Geometry and Mesh Generation Toolkit}
\newacronym{CURIE}{CURIE}{Centralized Used Fuel Resource for Information Exchange}
\newacronym{DAG}{DAG}{directed acyclic graph}
\newacronym{DANESS}{DANESS}{Dynamic Analysis of Nuclear Energy System Strategies}
\newacronym{DBE}{DBE}{Design Basis Event}
\newacronym{DES}{DES}{Detached eddy simulation}
\newacronym{DESAE}{DESAE}{Dynamic Analysis of Nuclear Energy Systems Strategies}
\newacronym{DF}{DF}{discontinuity factors}
\newacronym{DFEM}{DFEM}{discontinuous finite element method}
\newacronym{DG}{DG}{Discontinuous Galerkin}
\newacronym{DHS}{DHS}{Department of Homeland Security}
\newacronym{DNF}{DNF}{delayed neutron fraction}
\newacronym{DNP}{DNP}{delayed neutron precursor}
\newacronym{DNS}{DNS}{Direct numerical simulation}
\newacronym{DOE}{DOE}{Department of Energy}
\newacronym{DMSR}{DMSR}{Denatured Molten Salt Reactor}
\newacronym{DRACS}{DRACS}{Direct Reactor Auxiliary Cooling System}
\newacronym{DRE}{DRE}{dynamic resource exchange}
\newacronym{DSNF}{DSNF}{DOE spent nuclear fuel}
\newacronym{DYMOND}{DYMOND}{Dynamic Model of Nuclear Development }
\newacronym{EBS}{EBS}{Engineered Barrier System}
\newacronym{EDZ}{EDZ}{Excavation Disturbed Zone}
\newacronym{EPA}{EPA}{Environmental Protection Agency}
\newacronym{EP}{EP}{Engineering Physics}
\newacronym{EVOL}{EVOL}{Evaluation and Viability of Liquid Fuel Fast Reactor System}
\newacronym{FCO}{FCO}{Fuel Cycle Options}
\newacronym{FCT}{FCT}{Fuel Cycle Technology}
\newacronym{FEHM}{FEHM}{Finite Element Heat and Mass Transfer}
\newacronym{FEM}{FEM}{finite element method}
\newacronym{FEPs}{FEPs}{Features, Events, and Processes}
\newacronym{FHR}{FHR}{Fluoride-Salt-Cooled High-Temperature Reactor}
\newacronym{FLiBe}{FLiBe}{Fluoride-Lithium-Beryllium}
\newacronym{FP}{FP}{fission product}
\newacronym{FVM}{FVM}{finite volume method}
\newacronym{GDSE}{GDSE}{Generic Disposal System Environment}
\newacronym{GDSM}{GDSM}{Generic Disposal System Model}
\newacronym{GENIUSv1}{GENIUSv1}{Global Evaluation of Nuclear Infrastructure Utilization Scenarios, Version 1}
\newacronym{GENIUSv2}{GENIUSv2}{Global Evaluation of Nuclear Infrastructure Utilization Scenarios, Version 2}
\newacronym{GENIUS}{GENIUS}{Global Evaluation of Nuclear Infrastructure Utilization Scenarios}
\newacronym{GET}{GET}{general equivalence theory}
\newacronym{GIF}{GIF}{Generation IV International Forum}
\newacronym{GPAM}{GPAM}{Generic Performance Assessment Model}
\newacronym{GRSAC}{GRSAC}{Graphite Reactor Severe Accident Code}
\newacronym{GUI}{GUI}{graphical user interface}
\newacronym{HALEU}{HALEU}{high-assay low-enriched uranium}
\newacronym{HLW}{HLW}{high level waste}
\newacronym{HPC}{HPC}{high-performance computing}
\newacronym{HTC}{HTC}{high-throughput computing}
\newacronym{HTGR}{HTGR}{High-Temperature Gas-Cooled Reactor}
\newacronym{IAEA}{IAEA}{International Atomic Energy Agency}
\newacronym{IDT}{IDT}{integrated diffusion/transport}
\newacronym{IEA}{IEA}{International Energy Agency}
\newacronym{IEMA}{IEMA}{Illinois Emergency Mangament Agency}
\newacronym{IHLRWM}{IHLRWM}{International High Level Radioactive Waste Management}
\newacronym{IMSBR}{IMSBR}{Indian Molten Salt Breeder Reactor}
\newacronym{IMSR}{IMSR}{Integral Molten Salt Reactor}
\newacronym{INL}{INL}{Idaho National Laboratory}
\newacronym{INS}{INS}{incompressible Navier-Stokes}
\newacronym{IO}{I/O}{input/output}
\newacronym{IPRR1}{IRP-R1}{Instituto de Pesquisas Radioativas Reator 1}
\newacronym{IRP}{IRP}{Integrated Research Project}
\newacronym{IRPhEP}{IRPhEP}{International Reactor Physics Benchmark Experiment Evaluation Project}
\newacronym{ISFSI}{ISFSI}{Independent Spent Fuel Storage Installation}
\newacronym{ISRG}{ISRG}{Independent Student Research Group}
\newacronym{JFNK}{JFNK}{Jacobian-Free Newton Krylov}
\newacronym{LANL}{LANL}{Los Alamos National Laboratory}
\newacronym{LBNL}{LBNL}{Lawrence Berkeley National Laboratory}
\newacronym{LCOE}{LCOE}{levelized cost of electricity}
\newacronym{LDRD}{LDRD}{laboratory directed research and development}
\newacronym{LES}{LES}{Large eddy simulation}
\newacronym{LEU}{LEU}{low-enriched uranium}
\newacronym{LFR}{LFR}{Lead-Cooled Fast Reactor}
\newacronym{LFMSR}{LFMSR}{Liquid-fueled Molten Salt Reactor}
\newacronym{LGPL}{LGPL}{GNU Lesser General Public License}
\newacronym{LLNL}{LLNL}{Lawrence Livermore National Laboratory}
\newacronym{LMFBR}{LMFBR}{Liquid Metal Fast Breeder Reactor}
\newacronym{LOFC}{LOFC}{Loss of Forced Cooling}
\newacronym{LOHS}{LOHS}{Loss of Heat Sink}
\newacronym{LOLA}{LOLA}{Loss of Large Area}
\newacronym{LP}{LP}{linear program}
\newacronym{LWR}{LWR}{Light Water Reactor}
\newacronym{MAGNOX}{MAGNOX}{Magnesium Alloy Graphie Moderated Gas Cooled Uranium Oxide Reactor}
\newacronym{MA}{MA}{minor actinide}
\newacronym{MCFR}{MCFR}{Molten Chloride Fast Reactor}
\newacronym{MCNP}{MCNP}{Monte Carlo N-Particle code}
\newacronym{MCRE}{MCRE}{Molten Chloride Reactor Experiment}
\newacronym{MECS}{MECS}{Method of Equivalent Cross Sections}
\newacronym{MILP}{MILP}{mixed-integer linear program}
\newacronym{MIT}{MIT}{the Massachusetts Institute of Technology}
\newacronym{MOAB}{MOAB}{Mesh-Oriented datABase}
\newacronym{MOOSE}{MOOSE}{Multiphysics Object-Oriented Simulation Environment}
\newacronym{MOSART}{MOSART}{Molten Salt Actinide Recycler and Transmuter}
\newacronym{MOX}{MOX}{mixed oxide}
\newacronym{MPI}{MPI}{Message Passing Interface}
\newacronym{MPM}{MPM}{Multi-Physics Modelling}
\newacronym{MRPP}{MRPP}{Multiregion Processing Plant}
\newacronym{MSBR}{MSBR}{Molten Salt Breeder Reactor}
\newacronym{MSFR}{MSFR}{Molten Salt Fast Reactor}
\newacronym{MSRE}{MSRE}{Molten Salt Reactor Experiment}
\newacronym{MSR}{MSR}{Molten Salt Reactor}
\newacronym{NAGRA}{NAGRA}{National Cooperative for the Disposal of Radioactive Waste}
\newacronym{NEAMS}{NEAMS}{Nuclear Energy Advanced Modeling and Simulation}
\newacronym{NEUP}{NEUP}{Nuclear Energy University Programs}
\newacronym{NFCSim}{NFCSim}{Nuclear Fuel Cycle Simulator}
\newacronym{NGNP}{NGNP}{Next Generation Nuclear Plant}
\newacronym{NMWPC}{NMWPC}{Nuclear MW Per Capita}
\newacronym{NNSA}{NNSA}{National Nuclear Security Administration}
\newacronym{NPP}{NPP}{Nuclear Power Plant}
\newacronym{NPRE}{NPRE}{Department of Nuclear, Plasma, and Radiological Engineering}
\newacronym{NQA1}{NQA-1}{Nuclear Quality Assurance - 1}
\newacronym{NRC}{NRC}{Nuclear Regulatory Commission}
\newacronym{NSF}{NSF}{National Science Foundation}
\newacronym{NSSC}{NSSC}{Nuclear Science and Security Consortium}
\newacronym{NUWASTE}{NUWASTE}{Nuclear Waste Assessment System for Technical Evaluation}
\newacronym{NWF}{NWF}{Nuclear Waste Fund}
\newacronym{NWTRB}{NWTRB}{Nuclear Waste Technical Review Board}
\newacronym{NZE}{NZE}{Net-Zero Emissions by 2050 Scenario}
\newacronym{OCRWM}{OCRWM}{Office of Civilian Radioactive Waste Management}
\newacronym{OOP}{OOP}{Object-oriented programming}
\newacronym{ORION}{ORION}{ORION}
\newacronym{ORNL}{ORNL}{Oak Ridge National Laboratory}
\newacronym{PARCS}{PARCS}{Purdue Advanced Reactor Core Simulator}
\newacronym{PBAHTR}{PB-AHTR}{Pebble Bed Advanced High Temperature Reactor}
\newacronym{PBFHR}{PB-FHR}{Pebble-Bed Fluoride-Salt-Cooled High-Temperature Reactor}
\newacronym{PDE}{PDE}{partial differential equation}
\newacronym{PEI}{PEI}{Peak Environmental Impact}
\newacronym{PH}{PRONGHORN}{PRONGHORN}
\newacronym{PoliMi}{PoliMi}{Politecnico di Milano}
\newacronym{PRIS}{PRIS}{Power Reactor Information System}
\newacronym{PRKE}{PRKE}{Point Reactor Kinetics Equations}
\newacronym{PSI}{PSI}{Paul Scherrer Institute}
\newacronym{PSPG}{PSPG}{Pressure-Stabilizing/Petrov-Galerkin}
\newacronym{PV}{PV}{photovoltaic}
\newacronym{PWAR}{PWAR}{Pratt and Whitney Aircraft Reactor}
\newacronym{PWR}{PWR}{Pressurized Water Reactor}
\newacronym{PyNE}{PyNE}{Python toolkit for Nuclear Engineering}
\newacronym{PyRK}{PyRK}{Python for Reactor Kinetics}
\newacronym{QA}{QA}{quality assurance}
\newacronym{QD}{QD}{quasi-diffusion}
\newacronym{RANS}{RANS}{Reynolds-averaged Navier-Stokes}
\newacronym{RDD}{RD\&D}{Research Development and Demonstration}
\newacronym{RD}{R\&D}{Research and Development}
\newacronym{REE}{REE}{rare earth element}
\newacronym{RELAP}{RELAP}{Reactor Excursion and Leak Analysis Program}
\newacronym{RIA}{RIA}{Reactivity Insertion Accident}
\newacronym{RIF}{RIF}{Region-Institution-Facility}
\newacronym{RMM}{RMM}{Response Matrix Method}
\newacronym{ROD}{ROD}{Reactor Optimum Design}
\newacronym{RSM}{RSM}{Reynolds stress model}
\newacronym{SAM}{SAM}{System Analysis Module}
\newacronym{SAMOFAR}{SAMOFAR}{Safety Assessment of the Molten Salt Fast Reactor}
\newacronym{SAMOSAFER}{SAMOSAFER}{Severe Accident Modeling and Safety Assessment for Fluid-fuel Energy Reactors}
\newacronym{SD}{SD}{standard deviation}
\newacronym{SFR}{SFR}{Sodium-Cooled Fast Reactor}
\newacronym{SINAP}{SINAP}{Shanghai Institute of Applied Physics}
\newacronym{SINDAG}{SINDA{\textbackslash}G}{Systems Improved Numerical Differencing Analyzer $\backslash$ Gaski}
\newacronym{SKB}{SKB}{Svensk K\"{a}rnbr\"{a}nslehantering AB}
\newacronym{SNF}{SNF}{spent nuclear fuel}
\newacronym{SNL}{SNL}{Sandia National Laboratory}
\newacronym{SPH}{SPH}{superhomogenization}
\newacronym{STC}{STC}{specific temperature change}
\newacronym{SUPG}{SUPG}{Streamline-Upwind/Petrov-Galerkin}
\newacronym{SVDC}{SVDC}{spatially varying diffusion coefficient}
\newacronym{SWF}{SWF}{Separations and Waste Forms}
\newacronym{SWU}{SWU}{Separative Work Unit}
\newacronym{TFM}{TFM}{Transient Fission Matrix}
\newacronym{TH}{TH}{thermal-hydraulics}
\newacronym{TMSR}{TMSR}{Thorium Molten Salt Reactor}
\newacronym{TRACE}{TRACE}{TRAC/RELAP Advanced Computational Engine}
\newacronym{TRIGA}{TRIGA}{Training Research Isotope General Atomic}
\newacronym{TRISO}{TRISO}{Tristructural Isotropic}
\newacronym{TRU}{TRU}{transuranic}
\newacronym{TSM}{TSM}{Total System Model}
\newacronym{TSPA}{TSPA}{Total System Performance Assessment for the Yucca Mountain License Application}
\newacronym{ThOX}{ThOX}{thorium oxide}
\newacronym{TUD}{TU Delft}{Technische Universiteit Delft}
\newacronym{UFD}{UFD}{Used Fuel Disposition}
\newacronym{UML}{UML}{Unified Modeling Language}
\newacronym{UOX}{UOX}{uranium oxide}
\newacronym{UQ}{UQ}{uncertainty quantification}
\newacronym{US}{US}{United States}
\newacronym{UW}{UW}{University of Wisconsin}
\newacronym{VISION}{VISION}{the Verifiable Fuel Cycle Simulation Model}
\newacronym{VSOP}{VSOP}{Very Superior Old Programs}
\newacronym{VVER}{VVER}{Voda-Vodyanoi Energetichesky Reaktor (Russian Pressurized Water Reactor)}
\newacronym{VV}{V\&V}{verification and validation}
\newacronym{WIPP}{WIPP}{Waste Isolation Pilot Plant}
\newacronym{YMR}{YMR}{Yucca Mountain Repository Site}
\newacronym{BOL}{BOL}{Beginning-of-Life}
\newacronym{ULOF}{ULOF}{Unprotected Loss of Flow}
\newacronym{LOSCA}{LOSCA}{Loss of Secondary Cooling Accident}
\newacronym{ULOHS}{ULOHS}{Unprotected Loss of Heat Sink}

### 2024-05-03 lit.tex

\glspl{MSR} possess unique characteristics which render existing \gls{LWR}
analysis software inappropriate for \gls{MSR} analysis. Legacy \gls{LWR}
software typically scale poorly on modern high-performance computing
clusters and do not support complex geometries beyond the regular square or hexagonal shapes of
\gls{LWR} fuel assembly lattices. Furthermore, \glspl{MSR} feature strong coupling between the
neutronic and thermal-hydraulics physics, forcing segregated solvers into taking smaller timesteps
to maintain solver accuracy and stability. This chapter presents a literature
review of existing multiphysics simulation software developed for \gls{MSR}
analysis. This work focuses on software for analyzing short-term reactor
dynamics, which requires accurately simulating various transient
scenarios such as reactor start-up and coast-down, load-following operations,
steady-state operation, and accident analysis. Long-term dynamics such as fuel
burnup and structural corrosion fall outside the scope of this work.
This chapter also provides a literature review of turbulence modeling and transport-correction
techniques for the neutron diffusion method relevant to control rod modeling.

\section{MSR Multiphysics Modeling} \label{sec:msr-multiphysics}

The existence of multiple, disparate physics in a single system constitutes a multiphysics system.
Consequently, we need multiphysics software to accurately model a multiphysics system. Employing
well-verified single-physics software for coupled multiphysics simulations does not guarantee
stability, accuracy, or robustness \cite{keyes_multiphysics_2013}. Nevertheless, coupling separate
single-physics software for multiphysics simulations is a viable and popular option with proper
implementation. With strong coupling between different physics, extra care must be
taken to capture various multiphysics interactions accurately. For instance, the strong coupling
between neutron flux and temperature in \glspl{MSR} in \glspl{MSR} necessitates robust coupling
techniques. Na\"ive multiphysics coupling algorithms may result in longer computation times
at best and a lack of convergence in simulations at worst.

Whether one couples separate single-physics software together or develops a multiphysics software
under a single numerical solver framework, multiphysics software for modeling \glspl{MSR} must
employ \textit{tight coupling schemes} to couple the neutronics and thermal-hydraulics governing
equations to capture the strong multiphysics interactions in transient \gls{MSR} simulations
accurately without resorting to very small time-steps \cite{aufiero_development_2014}. Tightly
coupled numerical models handle multiphysics interactions by either updating all state
variables simultaneously in one monolithic solve (\textit{full coupling}) or
iteratively updating all state variables (\textit{fixed point iterations})
until the solution converges in every timestep \cite{keyes_multiphysics_2013}.
Full coupling tends to be more computationally expensive because it combines
all physics equations into a single large system of equations to be solved
simultaneously. In contrast, fixed point iterations involve operator splitting to
separate the system of equations into smaller systems based on their associated
physics, solving smaller systems separately, and iteratively updating the state
variables until convergence. Fixed point iterative methods can be less
stable, less accurate, and have poorer
convergence rates since these methods make iterative corrections
without regard to potentially destabilizing modes introduced by the
multiphysics coupling \cite{keyes_multiphysics_2013}. In particular, Lindsay et al.
\cite{lindsay_introduction_2018} demonstrated the non-convergence of a multiphysics \gls{MSR}
model without delayed neutron precursors when using fixed point iterations. While fully coupled
schemes result in large systems of equations to solve, they can still outperform fixed point
iterative methods in some multiphysics problems through superior stability and convergence rates.
Conversely, proven techniques exist for improving the performance of fixed point iteration coupling
schemes for many relevant computational multiphysics research fields, including reactor analysis
\cite{ragusa_consistent_2009}. Fixed point iterative methods also allow for different temporal and
spatial discretization schemes to meet the requisite resolution of each
individual physics component.

In contrast to tight coupling schemes, \textit{loose coupling schemes}
solve each set of single-physics equations using state variable
data from the previous timestep without iterative corrections within every
timestep. Loosely coupled schemes are inappropriate for modeling \glspl{MSR},
given the strong coupling between neutronics and thermal-hydraulics.
Aufiero et al. \cite{aufiero_development_2014} demonstrated a loose coupling
approach that failed to reproduce the expected increase in reactor power
in an \gls{MSR} in response to a 150 pcm reactivity insertion. One rare use case for loose
coupling is pseudo-transient problems for converging a multiphysics system to steady state.

Consequently, MSR multiphysics simulation tools must employ, at the very least, tight coupling
schemes through fixed-point iterations for most applications. The optimal coupling scheme depends
on the multiphysics model and the underlying numerical solver implementations.
The next subsection reviews existing \gls{MSR} multiphysics software and their applications.

\subsection{Review of MSR Multiphysics Simulations and Results} \label{sec:msr-tools}

Several simulation tools have been developed for \gls{MSR} modeling in the last two decades. Given
the focus on developing high-fidelity \gls{MSR} software to support reactor design optimization,
this literature review omits coupled neutronics/thermal-hydraulics MSR simulation work with
zero-dimensional point reactor kinetics in favor of those with spatially-resolved neutronics
methods.

\subsubsection{Coupling Existing Single-Physics Codes for MSR Modeling}

Coupling single-physics software to form integrated multiphysics tools allows researchers to
leverage existing well-validated, single-physics software. These single-physics software are also
highly optimized for solving specific \glspl{PDE} relevant to the investigated system.
Many \gls{MSR} simulation tools employ tight coupling to couple separate neutronics and
thermal-hydraulics solvers.

In 2007, Krepel et al. \cite{krepel_dyn3d-msr_2007} extended the
\gls{LWR} nodal diffusion code DYN3D to account for \gls{DNP} drift and fission energy deposition
in the coolant for \gls{MSR} modeling. The extended DYN3D-MSR code also contains the FLOCAL
thermal-hydraulics model for coolant flow and conjugate heat transfer modeling. Their \gls{MSRE}
and \gls{MSBR} models consisted of homogeneous hexagonal nodes (mesh elements) for the nodal
diffusion calculation coupled to individual hexagonal unit cells comprising of graphite with
circular fuel channels in the center for the thermal-hydraulics calculations. They simulated
steady-state operation; start-up, coast-down, natural circulation transients; and an overcooling
transient in their \gls{MSRE} model. They also modeled steady-state operation and a single-channel
blockage transient in their \gls{MSBR} model.

At the \gls{TUD}, Kophazi et al.
\cite{kophazi_development_2009} coupled their in-house 3-D neutron diffusion software DALTON
\cite{boer_validation_2010} and thermal-hydraulics
software THERM to develop a 3-D \gls{MSR} simulation tool. They homogenized the heterogeneous fuel
and graphite region of the \gls{MSRE} for the neutronics calculation. They discretized the
homogeneous region into a Cartesian mesh and modeled a more accurate representation of the oblong
fuel channels in the graphite core matrix instead of the circular shape by Krepel et al.
\cite{krepel_dyn3d-msr_2007}.

Both efforts by Krepel et al. and Kophazi et al. showed good
agreement with \gls{MSRE} experimental data in their respective validation tests.
The researchers from \gls{TUD} continued in their approach of coupling separate single-physics
software to create \gls{MSR} simulation tools, as noted by the DALTON-HEAT
\cite{de_zwaan_static_2007} coupling to model the \gls{MSFR} \cite{fiorina_modelling_2014}. The
fast-spectrum \gls{MSFR} features a large pool of fuel salt in its single-channel design instead of
the \gls{MSRE}'s multi-channel design. As a result, the authors had to incorporate
multidimensional turbulent flow modeling to accurately model the complex flow profiles and
turbulence effects in the \gls{MSFR} during steady-state operation and transient scenarios. In a
later effort from the same institute, Tiberga et al. \cite{tiberga_discontinuous_2019} coupled
PHANTOM-$S_N$ and DGFlows for their participation in the CNRS benchmark study
\cite{tiberga_results_2020}. Developed at the \gls{CNRS}, the CNRS benchmark facilitates
code-to-code verification of \gls{MSR} multiphysics software \cite{aufiero_testing_2018}. Tiberga
et al. showed that their coupled PHANTOM-$S_N$/DGFlows software was consistent with
OpenFOAM-based \gls{MSR} simulation tools. Further details on their study and the CNRS benchmark
is available in Section \ref{sec:msr-vv}.

Another multiphysics package was developed at the \gls{PSI} by coupling the \gls{LWR}
thermal-hydraulics system software \gls{TRACE} \cite{nrc_trace_2007}
with the nodal neutron diffusion software \gls{PARCS} \cite{downar_parcs_2010} for safety
analysis of the \gls{MSFR} \cite{pettersen_coupled_2016}. The right circular cylinder \gls{MSFR}
design was approximated using hexagonal nodes in \gls{PARCS} while \gls{TRACE} simulated
Reynolds-averaged, inviscid coolant flow. While modeling inviscid flow reduces computational costs
by eliminating the need for expensive turbulence modeling, this approach neglects turbulent flow
and turbulent mixing effects.

More recently, Jaradat et al. \cite{jaradat_development_2021}
extended the 3-D $P_1$ and $SP_3$ nodal transport code PROTEUS-NODAL to support \gls{DNP} drift.
Yang et al. \cite{yang_development_2022} then developed a coupled simulation tool of PROTEUS-NODAL
and \gls{SAM}. They modeled the \gls{MSFR} on a 2-D axisymmetric geometry
and the \gls{MSRE} on a 3-D R-$\theta$-Z geometry with twelve 1-D flow channels. The \gls{SAM} flow
model relied on the algebraic mixing length model for turbulence modeling. Both results
reproduced the expected trends observed in published experimental and simulation data. However,
their 1-D flow models neglected complex turbulent flows expected in the \gls{MSFR}.

With modern advancements in computing hardware and growing access to
high-performance computing systems, others have developed multiphysics tools by
coupling the \gls{CFD} software OpenFOAM
\cite{the_openfoam_foundation_ltd_openfoam_2021} with the Monte Carlo particle
transport software Serpent \cite{leppanen_serpent_2014}, thus achieving
high-fidelity neutronics calculations in transient reactor analyses \cite{laureau_transient_2017,
blanco_neutronic_2020}. To that end, Serpent has a multiphysics interface supporting coupling
with other physics software \cite{leppanen_development_2013}.

Laureau et al.
\cite{laureau_transient_2017} developed a different technique called the
\gls{TFM} method by introducing additional time-dependence
operators to conventional fission matrices typically used to accelerate source
convergence in Monte Carlo neutronics calculations. The \gls{TFM} method
pre-calculates three \glspl{TFM} of the reactor system in Serpent. It
interpolates the matrix values during the actual transient calculations to
incorporate temperature-induced salt expansion and Doppler effects on the neutron cross sections
and, ultimately, the neutron flux. Laureau et al. opted to subcycle the neutronics calculations by
setting smaller timesteps than the thermal-hydraulics calculations
to take advantage of the differing physics timescales. They demonstrated load-following capability
and parametric studies of overcooling and reactivity insertion transients in the \gls{MSFR}.

Blanco \cite{blanco_neutronic_2020} took a more direct approach by
compiling Serpent as an internal \texttt{C}-based function within OpenFOAM's
\texttt{C++}-based framework. This approach reduced the amount of external data
transfers between Serpent and OpenFOAM as both software have access to shared
memory during runtime. Their integrated tool employs the Quasi-Static
method for transient neutronics calculations and runs Serpent Monte Carlo
calculations several times per timestep until convergence is reached.
Blanco demonstrated the Serpent-OpenFOAM tool on the CNRS benchmark specified by Tiberga et al.
\cite{tiberga_results_2020}.

\subsubsection{Integrated Multiphysics Modeling Frameworks for MSR Modeling}

Another \gls{MSR} simulation approach involves developing inherently multiphysics software that
handle all multiphysics calculations and data transfer internally. Among earlier efforts, Nicolino
et al. \cite{nicolino_coupled_2008} and Zhang et al. \cite{zhang_development_2009} recognized the
need for more robust multiphysics coupling, acceleration techniques and internal data transfers
to reduce execution times. They each independently developed unnamed multiphysics simulation tools
and demonstrated their tools with non-moderated \gls{MSR} designs. Nicolino et al. presented one of
the first \gls{MSR} models with multi-dimensional flow, governed by the Navier-Stokes equation with
the Boussinesq approximation. They performed steady-state and transient calculations on a 2-D
axisymmetric model of the \gls{MOSART} reactor \cite{ignatiev_molten_2014}. Zhang et al. also
developed a similar 2-D axisymmetric solver, but they incorporated a $k$-$\epsilon$ turbulence
model and performed calculations on an unnamed pool-type \gls{MSR} design with graphite reflectors.
Both studies implemented the neutron diffusion model for neutronics modeling.
Later, Li et al. \cite{li_transient_2015} demonstrated the
steady-state and transient analysis capabilities of COUPLE, a neutronics and
thermal-hydraulics software, on a 2-D axisymmetric model of the \gls{MSFR}. COUPLE also solves the
neutron diffusion equations and the Navier-Stokes equations with the $k$-$\epsilon$ turbulence
model.

Others adopted extensible software frameworks for developing numerical solvers
to develop multiphysics reactor analysis software. Examples of these software
frameworks include the commercial COMSOL
Multiphysics\textsuperscript{\textregistered} software
\cite{comsol_ab_comsol_nodate}, the aforementioned open-source CFD toolbox
OpenFOAM, and the open-source finite-element
framework \gls{MOOSE} \cite{gaston_physics-based_2015}. Researchers at
\gls{PoliMi} developed an \gls{MSR} simulation tool in COMSOL and
modeled the \gls{MSBR} as a single axisymmetric fuel channel with a uniform
flow profile \cite{cammi_multi-physics_2011}, followed by the \gls{MSRE} core
also as a single axisymmetric fuel channel with parabola-shaped laminar flow
\cite{cammi_dimensional_2012}. They later expanded on their approach by
modeling the \gls{MSRE} upper plenum, downcomer and lower plenum, primary heat
exchanger, and secondary heat exchanger as 0D systems (lumped-parameter model)
and substituting the 2-D fuel channel with a 3-D fuel channel which more closely
resembled the actual fuel channels in the \gls{MSRE}
\cite{zanetti_geometric_2015}. Other than the \gls{MSRE}, they also modeled the
\gls{MSFR} in a later publication \cite{fiorina_modelling_2014}, which also featured \gls{TUD}'s
DALTON-HEAT coupled multiphysics tool.

Other institutes have dedicated significant development work towards
OpenFOAM-based \gls{MSR} simulation tools. Aufiero et al.
\cite{aufiero_development_2014} first introduced an OpenFOAM model developed
at \gls{PoliMi}. Their model implemented a neutron diffusion model and a
\gls{RANS}-based turbulence model under the incompressible flow assumption to demonstrate 2-D
and 3-D transient analyses of the \gls{MSFR}. Later advancements in the
\gls{PoliMi} solver include a fuel compressibility model with helium bubble
tracking to study fuel compressibility effects
\cite{cervi_development_2019} and an $SP_3$ neutron transport
model for improved neutronics calculations \cite{cervi_development_2019-1} in
the \gls{MSFR}. GeN-Foam is another OpenFOAM-based tool developed by Fiorina
et al. \cite{fiorina_gen-foam_2015} as a general reactor multiphysics solver
applicable to \glspl{MSR} and other reactor types. GeN-Foam features neutron
diffusion, $SP_3$, and $S_N$ neutronics models
\cite{fiorina_development_2016,fiorina_gen-foam_2015,fiorina_detailed_2019},
and thermo-mechanical modeling for reactor expansion effects. Using GeN-Foam,
Altahhan et al. \cite{altahhan_preliminary_2020} developed and optimized a
liquid-fuel \gls{MSR} design, while Shi \& Fratoni \cite{shi_gen-foam_2021}
simulated precursor drift effects in a homogenized \gls{MSRE} model. However, none of the
OpenFOAM-based tools have demonstrated the capability for time-dependent \gls{MSR} simulations
with explicit control rod modeling.

Finally, within the \gls{MOOSE} framework, simulation tools capable of modeling
\glspl{MSR} include Griffin \cite{abou-jaoude_coupled_2020}; and Moltres
\cite{lindsay_moltres_2017}\textemdash the subject of this work.
Griffin primarily tackles radiation transport problems, but the \gls{MOOSE}
framework facilitates multiphysics coupling with \gls{MOOSE}-based applications for other physics
such that all applications share the same data structure. This feature eliminates
computational costs from external data transfers and optionally allows for
\textit{fully coupled} calculations in which the application solves all physics
simultaneously. The mutual compatibility among different physics applications within the
\gls{MOOSE} framework simplifies the work required to couple
different physics solvers together to solve novel multiphysics problems. Similarly,
Moltres benefits from the highly-integrated cross-compatibility
within the ecosystem of \gls{MOOSE}-based applications. Abou-Jaoude et al.
\cite{abou-jaoude_coupled_2020} coupled Griffin with Pronghorn, another
\gls{MOOSE}-based application for advanced reactor thermal-hydraulics modeling, to
demonstrate several steady-state \gls{MSR} simulation capabilities defined in
the CNRS benchmark. Lindsay et al.
\cite{lindsay_introduction_2018} first demonstrated Moltres' \gls{MSR} modeling
capabilities on 2-D axisymmetric and 3-D Cartesian models of the \gls{MSRE} with
fixed velocity flow in a fully coupled neutronics and thermal-hydraulics simulations.
I later demonstrated some of Moltres' more recent developments through
modeling a 2-D axisymmetric model of the \gls{MSFR} for steady-state operation
and transient accident analysis \cite{park_advancement_2020}. The latter study
introduced looped \gls{DNP} flow, coupling the \gls{DNP} drift and temperature 
advection-diffusion to incompressible flow, and decay heat modeling
capabilities.

\subsection{MSR Multiphysics Modeling Verification \& Validation} \label{sec:msr-vv}

\Gls{VV} of simulation models are important components of simulation software development
\cite{sargent_verification_2010}. Verification involves checking whether a model's implementation
accurately represents the conceptual description and specifications. Model
validation is the process of checking whether a model is an accurate representation of the real
world within the range of its intended uses. For reactor software, model verification is commonly
performed by comparing them to other reactor software designed to run the same type of reactor
simulations. On the other hand, model validation is performed by comparing numerical results from
a simulation model to experimental data from the corresponding live test. The validity of a model
depends on the outcome of both model verification and validation.

The important multiphysics phenomena in \glspl{MSR} for model \gls{VV} are salt flow-induced
\gls{DNP} drift and the strong coupling between neutron flux and temperature advection-diffusion.
Delpech et al. \cite{delpech_benchmark_2003} published one of the first collaborative \gls{VV} work
for \gls{MSR} multiphysics modeling. Collaborators
from six institutions modeled \gls{MSRE} pump start-up, pump coast-down, and natural
circulation transients to assess and validate their models and codes for studying the effects of
salt flow on the reactivity and power. Given the wide range of neutronics methods, from
multidimensional Monte Carlo methods to zero-dimensional point reactor kinetics, some deviations
were observed between different codes. Nevertheless, all results generally agreed
with \gls{MSRE} experimental data from \gls{ORNL}.

As mentioned in Subsection \ref{sec:msr-tools}, Tiberga et al. \cite{tiberga_results_2020}
published the CNRS benchmark to verify \gls{MSR} simulation tools designed for
fast-spectrum \gls{MSR} modeling. In contrast with the multi-channel \gls{MSRE} and its derivative
designs, the CNRS benchmark has a 2 m$\times$2 m problem domain of homogeneous fuel salt mimicking
the large salt pool in fast-spectrum \gls{MSR} designs. The CNRS benchmark consists of three
phases: single-physics calculations in Phase 0, problems that gradually
introduce multiphysics coupling in Phase 1, and time-dependent perturbation problems in Phase
2. Thus, the benchmark provides a systematic approach to help code developers
identify sources of discrepancies that may otherwise be masked by error cancellation or other
dominant sources of discrepancies. The final steady-state and time-dependent problems involve
studying the effects of natural circulation and lid-driven flow on the reactivity and power output.
Aside from the problem specifications, Tiberga et al. also published the associated group
constant cross section data required by deterministic neutronics solvers to perform neutronics
calculations. Four institutions participated in the benchmarking exercise with neutron diffusion,
$SP_N$, and $S_N$-based solvers for the neutronics calculations. Their measured neutronics and
\gls{TH} parameters showed excellent agreement within up to 2.5\% discrepancy from their combined
average.

Neutronic benchmark studies of the \gls{MSRE} and the \gls{MSFR} by Fratoni et al.
\cite{fratoni_molten_2020} and Brovchenko et al. \cite{brovchenko_neutronic_2019} measured the
delayed neutron losses due to the decay of \glspl{DNP} flowing out of the active core region.
Fratoni et al. sought to establish a standard validation platform for \gls{MSR} neutronics
simulation tools with \gls{MSRE} experimental data for inclusion in the \gls{IRPhEP} handbook.
They characterized and validated a model of the \gls{MSRE} in the Monte Carlo particle transport
code Serpent. The \gls{MSFR} neutronics benchmark by Brovchenko et al. featured results
from multiple \gls{MSR} simulation tools by several collaborators. Their assessment found that
the choice of nuclear database for the cross sections and decay data has the most significant
impact on the neutronics results.

While these publications have addressed significant technical gaps, there is room to develop
further \gls{VV} benchmarks for multiphysics \gls{MSR} modeling. For instance, the
CNRS benchmark does not assess the loss of delayed neutrons due to the decay of \glspl{DNP} flowing
out of the active core region. This phenomenon is important as the delayed neutron fraction in the
core directly impacts the transient power response in unprotected accident scenarios. Meanwhile,
the neutronic benchmark studies by Fratoni et al. \cite{fratoni_molten_2020} and Brovchenko et al.
\cite{brovchenko_neutronic_2019} did not provide the standardized group constant data required by
most deterministic multiphysics \gls{MSR} simulation tools. Therefore, it is challenging to isolate
discrepancies arising from code implementations instead of discrepancies from using different
nuclear databases or stochastic uncertainties in Monte Carlo simulations. A suitable model
verification procedure for delayed neutron loss should ideally provide well-defined problems and
the necessary input data. It is also helpful to perform model verification studies on simpler
problems like the bare homogeneous problem domain in the CNRS benchmark before embarking on more
detailed validation studies that require accurate models of experimental setups.

\section{Turbulence Modeling in MSRs}

In fluid dynamics, turbulent flow is characterized by unsteady, irregular, and
chaotic fluid motion as opposed to neat, parallel flow layers in laminar flow
\cite{pope_turbulent_2000}. The transition from laminar to turbulent flow
typically occurs at Reynolds numbers between 2000 and 4000, depending on the
system \cite{pope_turbulent_2000}. Turbulent flows are expected in \glspl{MSR}.
Kedl \cite{kedl_fluid_1970} reports expected Reynolds numbers in the \gls{MSRE}
ranging from 1000 in the regular fuel coolant channels to over 10000 in the
flow distributor volute and core wall cooling annulus regions. For the
\gls{MSFR}, salt flow in the central core region is highly turbulent and
reaches Reynolds numbers on the order of $10^5$.

\subsection{Turbulence Models}

Numerous types of turbulence models exist for various turbulent flow
applications. The most common turbulence models can be classified into the
following categories by order of increasing fidelity and computational complexity:
%
\begin{itemize}
    \item \gls{RANS}-based models
    \begin{itemize}
        \item Eddy viscosity models
        \begin{itemize}
            \item Algebraic models
            \item One- and two-equation models
        \end{itemize}
        \item \gls{RSM}
    \end{itemize}
    \item \gls{DES}
    \item \gls{LES}
    \item \gls{DNS}
\end{itemize}

\gls{RANS}-based models are based on the \gls{RANS} equations obtained from
applying time-averaging to the fluid flow equations. The \gls{RANS}
equations separate flow into time-averaged $U$ and fluctuating $u$ components
and can be written in Einstein notation and Cartesian coordinates as:
%
\begin{align}
    \frac{\partial U_i}{\partial t} + U_j \frac{\partial u_i}{\partial x_j} =&
    -\frac{1}{\rho} \frac{\partial P}{\partial x_i} + \nu \nabla^2 U_i -
    \frac{\partial \langle u_i u_j \rangle}{x_j}
    \shortintertext{where}
    \langle \cdot \rangle =& \mbox{ time-averaging operator,} \nonumber \\
    \rho =& \mbox{ fluid density,} \nonumber \\
    P =& \mbox{ time-averaged pressure field,} \nonumber \\
    \nu =& \mbox{ kinematic viscosity.} \nonumber
\end{align}

Eddy viscosity models, which comprise the most widely used turbulence models
in use today \cite{rodi_turbulence_2017}, operate on the eddy viscosity
hypothesis which states that the Reynolds stresses in the \gls{RANS} equations
are given by:
%
\begin{align}
    \langle u_iu_j \rangle =& \frac{2}{3}k \delta_{ij} - \nu_T \left(
    \frac{\partial U_i}{\partial x_j} + \frac{\partial U_j}{\partial x_i}
    \right)
    \shortintertext{where}
    k =& \mbox{ mean turbulent kinetic energy,} \nonumber \\
    \delta_{ij} =& \mbox{ Kronecker delta,} \nonumber \\
    \nu_T =& \mbox{ eddy viscosity.} \nonumber \\
\end{align}

The various eddy viscosity models mainly differ in their approach to
the closure problem of calculating the eddy viscosity. As the name suggests,
algebraic models rely on algebraic equations to directly calculate the eddy viscosity
distribution from flow variables. As a result, algebraic models are
the least computationally intensive models for turbulence. Algebraic models
can be further categorized into uniform eddy viscosity models
and mixing length models. Uniform eddy viscosity models apply a uniform eddy
viscosity throughout the problem domain. The uniform eddy viscosity is
calculated from flow parameters such as the characteristic velocity, the
characteristic flow width, and empirically determined turbulent Reynolds
number.

Given that eddy viscosities usually vary significantly in most types of
flow, uniform eddy viscosity models have a limited range of applicability
\cite{pope_turbulent_2000}. Mixing length models add a level of complexity by
relating the eddy viscosity to spatially-varying flow parameters such as the
mean velocity gradient (Prandtl \cite{prandtl_7_1925} and Cebeci-Smith
\cite{smith_numerical_1967} models) or the mean rate of strain (Baldwin-Lomax
\cite{baldwin_thin-layer_1978} model) and an empirical mixing length parameter.
Combined with empirical data for the mixing length parameter, these
models provide better approximations of free shear flows but still
underperform for more complex flows involving flow separation and significant streamline curvature.

One- and two-equation turbulence models introduce differential equations to
describe turbulence quantities such as the turbulence kinetic energy and the
turbulence rate of dissipation to obtain the eddy viscosity distribution. The
most common and best-performing one-equation model is the Spalart-Allmaras
model, which directly provides an equation for the eddy viscosity with several
closure coefficients and functions \cite{wilcox_turbulence_2006}. The
Spalart-Allmaras model is considered ``complete'' as it does not involve
adjustable coefficients or functions. Calibrated for free shear flows in
aeronautical applications, the model performs modestly better than algebraic
models in these applications \cite{pope_turbulent_2000}, but it still deviates
significantly from experimental data
for separated flows \cite{wilcox_turbulence_2006}.

Investigations with
one-equation models reveal the need for an extra equation to account for
turbulent length scales separately from turbulent velocity. Thus, two-equation
models became the most widely adopted turbulence model since the late 20th century
\cite{pope_turbulent_2000}. Two-equation models include the $k$-$\epsilon$,
$k$-$\omega$, and $k$-$\tau$ models. The variables $k$, $\epsilon$, $\omega$,
and $\tau$ correspond to turbulent kinetic energy, turbulent dissipation,
specific turbulent dissipation rate, and turbulent time scale, respectively.
While none of these models perform universally well across all types of
turbulent flow problems, they are generally more
accurate than the algebraic and one-equation models. Successive contributions
and modifications to the two-equation models through the years have also
improved their performance in predicting various types of turbulent flow. Compared to expensive,
high-fidelity turbulence models, their moderate computational expense favors their adoption in most
commercial \gls{CFD} software for engineering applications \cite{pope_turbulent_2000}.

\glspl{RSM} directly compute the individual components $\langle u_i u_j
\rangle$ of the Reynolds stress tensor instead of approximating it with a
single, isotropic eddy viscosity term. Consequently, \glspl{RSM} provide
more realistic predictions for flows with significant rotational motion and
sudden changes in the mean strain rate, albeit at greater computational
expense, compared to the one- and two-equation models
\cite{wilcox_turbulence_2006}. More minor improvements are observed in modeling
free shear and backward-facing step flows \cite{wilcox_turbulence_2006}.

Due to the much higher computational cost for \gls{DES}, \gls{LES}, and
\gls{DNS}, these models have limited applicability in routine, high-Reynolds
number engineering problems today. However, given their high accuracy, these
models are useful for flow problems with relatively simple geometries and at
low Reynolds numbers and validating the lower-fidelity turbulence models
\cite{zhiyin_large-eddy_2015}.

\subsection{Turbulence Modeling in MSR Simulation Tools}

Podila et al. \cite{podila_cfd_2019} performed \gls{CFD} simulations of the
\gls{MSRE} core with six different turbulence models, namely a Spalart-Allmaras
model, two variants of the $k$-$\epsilon$ model, a $k$-$\omega$ model, and two
variants of \glspl{RSM}. While the turbulence models predicted widely differing near-wall
turbulence intensity distributions, their results showed small differences
in graphite and fuel temperatures among the different turbulence models. Ultimately, given
the lack of experimental data for model validation, the authors could compare between the
models, but they could not provide a definitive
assessment of the models' accuracies.

Amongst other \gls{MSR} simulation tools, the $k$-$\epsilon$ and $k$-$\omega$
turbulence models are the most commonly used models, as shown in published work
with COMSOL \cite{fiorina_modelling_2014}, OpenFOAM
\cite{aufiero_development_2014}, and \gls{TUD}'s in-house codes
\cite{fiorina_modelling_2014,tiberga_results_2020}. Fiorina et al.
\cite{fiorina_modelling_2014} compared the flow distribution from both models
in a 2-D axisymmetric \gls{MSFR} geometry. They observed that the $k$-$\omega$
model produced a more expansive recirculation zone near the outer wall, an additional
recirculation zone near the top wall, and significantly higher maximum
temperatures within the former recirculation zone.

In a recent study by
Laureau et al. \cite{laureau_unmoderated_2022}, the authors applied a \gls{DES} model to model
eddies in the \gls{MSFR}. Given the high computational costs required to obtain reasonably small
stochastic uncertainties from Monte Carlo calculations on arbitrarily varying temperature fields,
they developed a reduced-order model relating local reactivity feedback estimates to the nominal
temperature field to obtain the effective temperature field.
Additionally, they fixed the total
reactor power output to 3 GW to further reduce costs. With this multiphysics model, they found
that the pre-existing \gls{MSFR} model was prone to large power fluctuations of up to 7.5\% under
steady-state operation. Previous studies with lower-fidelity models, including the two-equation
models, failed to capture this effect due to the time-averaging assumption required in those
models. After replacing the large inlet pipes with multiple blade-like inlet channels in the
\gls{MSFR} model, they managed to reduce power fluctuations to 1.2\%.

In essence, the existing literature on turbulence flows in \glspl{MSR} highlight
the importance of modeling turbulent flows and calls for extra care towards the choice of
turbulence model depending on the \glspl{MSR} design and the phenomenon being studied.

\section{Control Rod Modeling in MSRs}

\subsection{Control Rods in MSRs}

The number and locations of control rods in a reactor vary significantly for different reactor
types, depending on their size, refueling frequency, and reactivity control systems among other
factors. \glspl{MSR} contain comparatively fewer control rods than most other reactor types due to
the stability and reactivity control capability provided by the large negative temperature
reactivity feedback. For instance, the \gls{MSRE} contained three rods and the \gls{MSBR}
developers recommended two to four rods for the \gls{MSBR} \cite{robertson_msre_1965,
robertson_conceptual_1971}. While these reactors contain control rods, the rods are used in
\glspl{MSR} differently than other reactor designs. The
strong temperature feedback allows operators to control reactivity by varying salt pump
speeds or heat removal rates. Many \gls{MSR} designs include control rods to facilitate reactor
start-up or shut-down rather than reactivity control during power operation. Using
control rods to achieve even fuel burnup is not a concern for \glspl{MSR} due to the liquid fuel
form. Nevertheless, it is important to characterize and accurately model control rod effects to
ensure safe operation of the reactor. Introducing control rod modeling capability to multiphysics
tools allows us to model control rod effects in time-dependent simulations.

There is a lack of research interest in methods for improving control rod modeling in
\glspl{MSR}.
The self-regulating nature and low excess reactivity of \glspl{MSR} reduce or even eliminate the
need for control rods \cite{dolan_1_2017}. This characteristic arises from the strongly negative
temperature reactivity feedback and the strong coupling to flow and heat removal rates. \gls{MSR}
reactor designers rely on these passive mechanisms for reactor safety. In addition, they propose
load-following operation by adjusting pump speeds and the temperature of the intermediate coolant
loop. Laureau et al. \cite{laureau_transient_2017} demonstrated simple power ramp-up and ramp-down
with the \gls{MSFR} by varying the temperature of the secondary coolant. \gls{MSR}
designs with no control rods include the fast-spectrum \gls{MSFR} and TerraPower's \gls{MCFR}
\cite{terrapower_terrapower_2021}. However, \gls{MSR} power control schemes which rely heavily on
these mechanisms may not be suitable for all accident scenarios. Control rod insertion provides
an important additional means of reactor shut-down without draining the fuel salt, which
complicates reactor restart procedures. We also note that thermal \gls{MSR} designs like the
\gls{MSRE}, \gls{MSBR}, and TMSR-LF1 designs still include control rods as essential components,
partly due to their less negative temperature reactivity feedback coefficients
compared to their fast-spectrum counterparts. Thus, the existing literature on control rod modeling
in \glspl{MSR} are based on thermal-spectrum \glspl{MSR}.

Few \gls{MSR} multiphysics studies explicitly include control rods in their models. For instance,
for the \gls{MSRE} pump start-up and coast-down experiments at \gls{ORNL} which involved control
rod movement, most numerical studies simulate the reactivity effects of the control rods by scaling
the neutron source term by the neutron multiplication factor to keep their reactor model at
criticality \cite{delpech_benchmark_2003, krepel_dyn3d-msr_2007}. Some studies include control rod
models in their steady-state calculations. However, they resort to neutron source term scaling for
the transient calculations due to the inaccuracy of neutron diffusion, $P_1$, and $SP_N$ methods in
highly neutron absorbing regions \cite{kophazi_development_2009, jaradat_development_2021,
yang_development_2022}. Kophazi et al. \cite{kophazi_development_2009} modeled annular control rods
as homogenized hexahedrons in their 3-D Cartesian geometry of the \gls{MSRE}. They
homogenized fuel and graphite regions in the rest of the reactor model, as mentioned in Subsection
\ref{sec:msr-tools}. To reduce errors introduced by the neutron diffusion method, they imposed
simple albedo boundary conditions for the thermal neutron group. Jaradat et al.
\cite{jaradat_development_2021} and Yang et al. \cite{yang_development_2022} modeled the \gls{MSRE}
control rods as homogenized wedges in their R-$\theta$-Z mesh due to the constraints of the nodal
neutronics solver. Cui et al. \cite{cui_development_2021} also homogenized the control rods as
regular hexahedral nodes in accordance with their nodal solver in Cartesian geometry. A disadvantage
of homogenization-based methods is that it removes some of the heterogeneity in the flux and
temperature. The heterogeneous temperature distribution is especially important in \glspl{MSR} due
to the combined fuel-coolant and the positive temperature reactivity feedback observed
in graphite under certain conditions \cite{mathieu_thorium_2006}.

In theory, multiphysics simulation tools with Monte Carlo neutron transport, such as the
Serpent-OpenFOAM tool developed by Blanco \cite{blanco_neutronic_2020} could boast excellent
control rod modeling performance. However, the computational cost would be extremely high for
reactor geometries much more complex than the 2 m$\times$2 m homogeneous domain of the CNRS
benchmark (Subsection \ref{sec:msr-tools}). Other high-fidelity neutronics methods, such as the
$S_N$ and $P_N$ methods are also typically too computationally expensive for time-dependent
multiphysics simulations. Conversely, less taxing methods, such as the neutron diffusion
method, cannot model control rods accurately. The remainder of this section describes the
challenges of control rod modeling with neutron diffusion and reviews existing transport
correction techniques for improving neutron diffusion method performance in control rod modeling.

\subsection{Summary of Neutronics Methods} \label{sec:summary-nts-mtds}

High-fidelity neutron transport methods fall under two general categories: stochastic Monte Carlo
methods and deterministic methods. Monte Carlo methods involve simulating a finite number of
neutron histories. Randomly generated numbers determine the outcome of various probabilistic events
such as travel distances between particle collisions, types of collisions, and scattering angles in
each history until a neutron capture event terminates it. Deterministic methods for neutron
transport, such as the discrete ordinates $S_N$ and spherical harmonics $P_N$ methods, solve the
\gls{BTE} for neutron transport with some approximations for handling the angular and energy
dependence. The \gls{BTE} for neutron transport is given as:
%
\begin{align}
  \frac{1}{v(E)} \frac{\partial}{\partial t} \Psi(\vec{r},E,\hat{\Omega},&t) + \hat{\Omega}\cdot
  \nabla\Psi(\vec{r},E,\hat{\Omega},t) + \Sigma_t(\vec{r},E,t)\Psi(\vec{r},E,\hat{\Omega},t) 
  \nonumber \\
  - \int^\infty_0 dE'& \int_{4\pi} d\hat{\Omega}' \Sigma_s(\vec{r},E'\rightarrow E,\hat{\Omega}'
  \rightarrow \hat{\Omega},t) \Psi(\vec{r},E',\hat{\Omega}',t) \nonumber \\
  =& \frac{\chi(E)}{4\pi}
  \int^\infty_0 dE' \int_{4\pi} d\hat{\Omega}' \nu\Sigma_f(\vec{r},E',t) \Psi(\vec{r},E',
  \hat{\Omega}',t)+S(\vec{r},E,\hat{\Omega}',t) \label{eq:bte}
  \shortintertext{where}
  v(E) =& \mbox{ neutron velocity,} \nonumber \\
  \Psi(\vec{r},E,\hat{\Omega},t) =& \mbox{ neutron angular flux,} \nonumber \\
  \vec{r} =& \mbox{ spatial coordinates,} \nonumber \\
  E =& \mbox{ neutron energy,} \nonumber \\
  \hat{\Omega} =& \mbox{ direction of neutron travel,} \nonumber \\
  t =& \mbox{ time,} \nonumber \\
  \Sigma_t(\vec{r},E,t) =& \mbox{ macroscopic total cross section,} \nonumber \\
  \Sigma_s(\vec{r},E'\rightarrow E,\hat{\Omega}'\rightarrow \hat{\Omega},t) =&
  \mbox{ macroscopic scattering cross section,} \nonumber \\
  \chi(E) =& \mbox{ fission neutron spectrum,} \nonumber \\
  \nu =& \mbox{ number of neutrons produced per fission reaction,} \nonumber \\
  \Sigma_f(\vec{r},E',t) =& \mbox{ macroscopic fission cross section,} \nonumber \\
  S(\vec{r},E,\hat{\Omega}',t) =& \mbox{ external neutron source.} \nonumber
\end{align}

The $S_N$ method discretizes the continuous angular directional phase space into a few discrete
angular directions (ordinates). It uses quadrature rules to replace the integrals over
$\hat{\Omega}$ with summations over the ordinates. On the other hand, the $P_N$ method introduces
Legendre polynomial expansions of the angular flux to approximate the angular dependence. Both
methods require higher-order approximations to produce more accurate flux solutions through more
discrete ordinates or higher-order Legendre expansions. Both methods also discretize the
continuous energy dependence into discrete neutron energy groups, which cover non-overlapping,
finite energy ranges across the entire energy spectrum to form multigroup equations as follows:
%
\begin{align}
  \frac{1}{v_g} \frac{\partial}{\partial t} \Psi_g(\vec{r},\hat{\Omega},&t) + \hat{\Omega}\cdot
  \nabla\Psi_g(\vec{r},\hat{\Omega},t) + \Sigma_{t,g}(\vec{r},t)\Psi_g(\vec{r},\hat{\Omega},t)
  \nonumber \\
  - \sum^G_{g'=1}& \int_{4\pi} d\hat{\Omega}' \Sigma_s^{g'\rightarrow g}(\vec{r},\hat{\Omega}'
  \rightarrow \hat{\Omega},t) \Psi_{g'}(\vec{r},\hat{\Omega}',t) \nonumber \\
  =& \frac{\chi_g}{4\pi}
  \sum^G_{g'=1} \int_{4\pi} d\hat{\Omega}' \nu\Sigma_{f,g'}(\vec{r},t) \Psi_{g'}(\vec{r},
  \hat{\Omega}',t)+S_g(\vec{r},\hat{\Omega}',t) \label{eq:mg-bte}
  \shortintertext{where}
  G =& \mbox{ total number of energy groups,} \nonumber \\
  g =& \mbox{ neutron energy group index (in decreasing energy order)} = 1,2,...,G. \nonumber
\end{align}

The subscript $g$ denotes the corresponding quantity for neutrons in energy group $g$.
Neutron transport methods are very computationally expensive and, thus, are mainly used for
time-independent neutronic analyses. For time-dependent multiphysics simulations coupling
neutronics to \gls{TH} and other physics present in nuclear reactors, most reactor codes
rely on the neutron diffusion equation for modeling neutronics. The multigroup neutron diffusion
equations are \glspl{PDE} derived from Eq. \ref{eq:mg-bte} by making
simplifying assumptions on the angular dependence in the scattering cross section and integrating
the equations over all solid angles to eliminate angular dependence as follows:
%
\begin{align}
  \frac{1}{v_g} \frac{\partial}{\partial t} \phi_g&(\vec{r},t) + \nabla\cdot J_g(\vec{r},t)
  +\Sigma_{t,g}(\vec{r},t) \phi_g(\vec{r},t) \nonumber \\
  =& \sum^G_{g'=1}\left[\Sigma_s^{g'\rightarrow g}
  \phi_{g'}(\vec{r},t) + \chi_g \nu\Sigma_{f,g'} \phi_{g'}(\vec{r},t)\right] + S_g(\vec{r},t)
  \label{eq:mg-diff} \\
  \shortintertext{where}
  J_g(\vec{r},t) =& \mbox{ neutron current for neutron group $g$.} \nonumber
\end{align}

Fick's first law of diffusion provides a closure relation for Eq. \ref{eq:mg-diff} by relating the
current to the flux:
%
\begin{align}
  J_g(\vec{r},t) =& -D_g(\vec{r},t)\nabla\phi_g(\vec{r},t) \label{eq:fick}
  \shortintertext{where}
  D_g(\vec{r},t) =& \mbox{ neutron diffusion coefficient for neutron group $g$.} \nonumber
\end{align}

The diffusion coefficient itself is estimated from the total or transport cross sections, depending
on whether we take scattering to be isotropic or linearly anisotropic in the $P_1$ approximation of
the \gls{BTE} \cite{lamarsh_introduction_1975}:
%
\begin{align}
  D(\vec{r},t) =& \frac{1}{3\Sigma_t(\vec{r},t)} \quad \mbox{(isotropic)} \\
  D(\vec{r},t) =& \frac{1}{3\Sigma_{tr}(\vec{r},t)} = \frac{1}{3\left(\Sigma_t(\vec{r},t)-
  \bar{\mu}\Sigma_s(\vec{r},t)\right)}
  \quad \mbox{(linearly anisotropic)} \label{eq:p1-diffcoef}
  \shortintertext{where}
  \Sigma_{tr} =& \mbox{ macroscopic transport cross section,} \nonumber \\
  \bar{\mu} =& \mbox{ average cosine of the scattering angles.} \nonumber
\end{align}

This treatment significantly reduces the number of coupled \glspl{PDE} to be solved and the
computational costs of modeling neutronics in reactors. However, the simplifications limit the
validity of the neutron diffusion equation to regions of high scattering-to-absorption ratios at
least a few mean free paths away from shared interfaces to neighboring media with highly dissimilar
neutronic properties \cite{shultis_chapter_2016}. Neutron diffusion theory cannot capture the
strongly anisotropic flux within and near control rod regions.

Many legacy diffusion solvers are based on coarse-mesh or nodal methods. These methods
primarily involve replacing heterogeneous lattices of materials of differing properties with
equivalent homogeneous mixtures of the same materials in each coarse mesh (referred to as nodes in
nodal methods) \cite{stacey_nuclear_2007}. Reducing the heterogeneity of the geometry reduces
computational complexity and circumvents poor diffusion performance in highly heterogeneous
interfaces. Each coarse mesh typically corresponds to a subregion comprising of repeating
substructures, e.g. a single fuel assembly or randomly distributed spherical fuel pebbles. The
homogenization procedure consists of two main steps: a transport calculation to obtain detailed
heterogeneous flux distribution within each subregion, and the calculation of homogenized
\textit{group constants} from the detailed flux distribution. \textit{Group constants} refer to
macroscopic neutron cross section values for various neutron interactions such as
scattering, absorption, and fission. Macroscopic cross sections represent the probability that a
neutron, in a given energy range, will undergo the associated interaction per unit distance
traveled in the material. Group constants also broadly include neutron diffusion
coefficients/constants and \gls{DNP} data. While advanced coarse-mesh and nodal
methods provide reasonable flux solutions for calculating global and intermediate-scale quantities,
they do not capture detailed heterogeneous flux distribution within each coarse-mesh subregion.

\subsection{Transport-Correction Techniques With Neutron Diffusion-Based Solvers}

Transport-correction techniques for diffusion-based solvers require additional information beyond
conventional homogenized group constants from transport-based solvers to ameliorate diffusion
solution accuracy in highly absorbing and near-interface regions.
Therefore, this literature review focuses on hybrid diffusion-transport methods developed to relay
transport-corrections to diffusion-based solvers for more accurate neutron flux and current
estimates. The hybrid methods differ mainly in how they incorporate corrections into a
diffusion-based solver. While this work focuses on improving diffusion-based control rod
modeling in Moltres for \gls{MSR}-like designs, this literature review explores existing methods
developed for other reactor types, as no advanced techniques have been applied to \glspl{MSR}.

\subsubsection{Absorber Blackness and Linear Extrapolation Length}

Methods based on absorber blackness \cite{davison_influence_1951, spinks_extrapolation_1965,
pellaud_extrapolation_1968, mendelson_two-dimensional_1969} encompass a
broad class of procedures for generating boundary conditions to match approximate solutions of
low-order methods (e.g., diffusion) to more accurate solutions from high-order methods (e.g.,
transport). The internal boundary conditions replace absorber regions and mimic their presence in
the problem domain. The boundary conditions are generalizations of the Marshak boundary condition
\cite{marshak_note_1947}, which in 1-D are of the form:
%
\begin{gather}
  \frac{\phi(x)}{d\phi(x)/dx} = \lambda \label{eq:marshak}
  \shortintertext{where}
  \begin{align}
    \lambda =& \mbox{ linear extrapolation length.} & \nonumber
  \end{align}
\end{gather}
%
The flux gradient in the denominator is taken in the outward direction at the surface of the
absorber or black body. In Cartesian coordinates, it is common to approximate Eq. \ref{eq:marshak}
in the form:
%
\begin{align}
  \phi(x+\lambda) =& 0
\end{align}
because Dirichlet boundary conditions are easier to work with analytically and numerically.
The corresponding relations in cylindrical and spherical coordinates depend on the radius of the
absorber rod. Various forms for $\lambda$ exist for specific absorber geometries such as slabs
\cite{maynard_blackness_1959} and cylinders \cite{spinks_extrapolation_1965,
pellaud_extrapolation_1968} which were dependent
on the size of the absorber region and coefficients quantifying the escape probability of neutrons
entering the absorber region. The $\lambda$ and the associated coefficients were derived
analytically from the \gls{BTE} with some simplifying assumptions, such as having uniform or
cosine-shaped incident neutron currents, isotropic scattering within the absorber, and mathematical
approximations in ignoring higher-order terms in intermediate steps. Blackness theory emerged in
the mid-20th century when computational resources were limited. Later, as computational
resources became more readily available and powerful enough for more complicated transport
calculations, appropriate boundary conditions could be calculated numerically
\cite{bretscher_computing_1997}. Alternatively, effective group constants can be calculated for
diffusion codes not programmed to handle internal boundary conditions. For instance,
Bretscher \cite{bretscher_computing_1997} provided formulae for effective diffusion coefficient and
absorption cross section for thin absorber slabs as a function of blackness coefficients, absorber
thickness, and mesh size.

Diffusion calculations with corrections from blackness theory can generally provide control rod
worths that reasonably agree with more accurate transport calculations. However, the
diffusion flux distributions still deviate significantly from transport calculations because no
corrections are introduced in the adjacent regions near the absorber where the angular flux can be
strongly anisotropic.

\subsubsection{Method of Equivalent Cross Sections}

Scherer \& Neef developed the \gls{MECS} \cite{scherer_determination_1976} to improve control rod
modeling in \gls{HTGR} with mesh-centered nodal diffusion methods. The \gls{MECS} involves running
a 1-D neutron transport calculation on a representative \textit{supercell} of the heterogeneous
absorber region and its vicinity. The supercell is a volume-preserved model of the absorber region
approximated from its cylindrical or Cartesian geometry in the diffusion calculation. The transport
calculation is typically performed using a fine group structure, fine spatial discretization, and
high-order angular discretization and scattering moments (e.g., $S_8$ method with $P_3$ scattering
matrix) \cite{fen_modelling_1992}. The calculated net neutron leakage rates from the transport
solver are matched with the calculated leakage rates from the diffusion solver according to an
analytic calculation to obtain equivalent diffusion coefficients or cross sections for the absorber
region. The diffusion coefficient formula for this analytic calculation depends on the solution
method of the diffusion solver. In the most widely used implementation of \gls{MECS} used in
conjunction with the mesh-centered diffusion code CITATION \cite{teuchert_vsop94_1994}, the
diffusion coefficient formula assumes there is only one inner mesh point in the absorber region as
using more inner mesh points dramatically increases the difficulty of finding the proper analytic
formula \cite{fen_modelling_1992}. Other macroscopic cross sections for the absorber region are
calculated using conventional group constant generation techniques from the transport solution
using the supercell average flux and reaction rates. To facilitate further discussion of the
\gls{MECS}, here is an example of applying the \gls{MECS} on a control rod in x-y geometry (Figure
\ref{fig:mecs-geometry}) in the CITATION code as provided by Fen et al. \cite{fen_modelling_1992}.
%
\begin{figure}[htb]
    \centering
    \includegraphics[width=.4\columnwidth]{mecs-geometry}
    \caption{Geometries of the absorber region and its vicinity in the supercell
      (top) and the diffusion solver mesh (bottom).
        The shaded and unshaded regions represent the homogenized absorber
        region and the adjacent non-absorber region, respectively.
        Retrieved from Fen et al. \cite{fen_modelling_1992}.}
    \label{fig:mecs-geometry}
\end{figure}

Figure \ref{fig:mecs-geometry} shows the 1-D cylindrical super cell for the
transport solver on the top and the corresponding 2-D Cartesian geometry for the
diffusion solver on the bottom. The four adjacent nodes in the Cartesian
geometry and their corresponding flux values are pairwise identical.

According to the mesh-centered geometry in CITATION, the leakage $L$ is
calculated as:
%
\begin{gather}
  L = \frac{F_y\left(\phi_x-\phi_a\right)}{\frac{\delta_x}{D_a}+
    \frac{\Delta_x}{D_o}} + \frac{F_x\left(\phi_y-\phi_a\right)}{
  \frac{\delta_y}{D_a}+\frac{\Delta_y}{D_o}} \label{eq:mecs}
  \shortintertext{where}
  \begin{align}
    \phi_x =& \mbox{ flux in the $x$-neighbor nodes,} \nonumber \\
    \phi_y =& \mbox{ flux in the $y$-neighbor nodes,} \nonumber \\
    \phi_a =& \mbox{ flux in the absorber node,} \nonumber \\
    F_y =& \mbox{ surface area to $y$-neighbor nodes} = 2pa, \nonumber \\
    F_x =& \mbox{ surface area to $x$-neighbor nodes} = 2a, \nonumber \\
    \delta_x =& \mbox{ distance from the absorber node center to the
      $x$-surface} = a/2, \nonumber \\
    \delta_y =& \mbox{ distance from the absorber node center to the
      $y$-surface} = pa/2, \nonumber \\
    \Delta_x =& \mbox{ distance from the $x$-surface to the $x$-neighbor node
      center} = p_x a/2, \nonumber \\
    \Delta_y =& \mbox{ distance from the $y$-surface to the $y$-neighbor node
      center} = p p_y a/2, \nonumber \\
    a, p, p_x, p_y =& \mbox{ geometric parameters of the CITATION mesh geometry
      (Figure \ref{fig:mecs-geometry}),} \nonumber \\
    D_a =& \mbox{ diffusion coefficient in the absorber node,} \nonumber \\
    D_o =& \mbox{ diffusion coefficient in the neighboring nodes.} \nonumber
  \end{align}
\end{gather}

In the original formulation by Scherer \& Neef \cite{scherer_determination_1976}, they let
$D_a=D_o$. Taking leakage and flux values at $R_x$ and $R_y$ (figure \ref{fig:mecs-geometry}) from
the transport solver to be equal to the corresponding values in equation \ref{eq:mecs} yields a
value for $\phi_a$ representing the average flux in the absorber region. The equivalent
cross sections for each reaction type $i$ are then determined by matching reaction rates from the
transport calculation to the reaction rate as governed by diffusion theory as follows:
%
\begin{gather}
  \Sigma_i = \frac{A_i}{\phi_a V}
  \shortintertext{where}
  \begin{align}
    \Sigma_i =& \mbox{ macroscopic cross section of reaction type $i$,} \nonumber \\
    A_i =& \mbox{ reaction rate of reaction type $i$ from the transport calculation,} \nonumber \\
    V =& \mbox{ volume of the absorber region} = a^2 p. \nonumber
  \end{align}
\end{gather}

Fen et al. \cite{fen_modelling_1992} later provided an updated formulation by assuming that the
cross sections are accurate and instead determined the equivalent diffusion coefficient from Eq.
\ref{eq:mecs} and the transport leakage and flux values as follows:
%
\begin{gather}
  \frac{1}{D_a} = 2p\frac{\phi_x-\phi_a}{L}+\frac{2}{p}\frac{\phi_y-\phi_a}{L}
    -\frac{p_x+p_y}{2D_o}+\sqrt{R}
  \shortintertext{where}
  \begin{align}
    R =& \left(2p\frac{\phi_x-\phi_a}{L}+\frac{2}{p}\frac{\phi_y-\phi_a}{L}+
    \frac{p_x+p_y}{2D_o}\right)^2+4p\left(p_y-p_x\right)\frac{\phi_x-\phi_a}{L}
      \frac{1}{D_o} & \nonumber
  \end{align}
\end{gather}

As illustrated by the implementation in CITATION, \gls{MECS} requires considerable user input on
the supercell configuration to ensure that the leakage rates of the transport solution are
equivalent to the leakage rates of the subsequent diffusion solution. This procedure also places
some constraints on the location and size of the absorber node that can be inconsistent with the
nodalization of the rest of the reactor geometry \cite{ougouag_transport_2010}. \gls{MECS} is
incompatible with reactor geometries which contain control rods that are too close to each other
such that there is not enough distance in between to define neighboring nodes where
flux-equivalence is assumed. Lastly, \gls{MECS} is only applicable to coarse-mesh solvers and
homogenizes the control rod along with the adjacent material in its vicinity. Nevertheless,
\gls{MECS} has been widely used within the \gls{VSOP} suite of codes (which contains CITATION) for
\gls{HTGR} neutronic safety analyses
\cite{fen_modelling_1992, reitsma_evaluating_2003, mulder_neutronics_2020}.

\subsubsection{Response-Based Methods}

Like the \gls{MECS}, \textit{response-based methods} rely on response-function-based
transport methods to resolve the flux around absorber regions. In this context, coarse mesh/nodal
response functions relate quantities of interest of an individual node to input values from its
neighboring nodes. For instance, a response function may provide a node's average nodal flux and
outgoing partial currents in response to a given set of incident
partial fluxes from its neighboring nodes \cite{ougouag_transport_2010}. Transport solutions are
used to generate sets of response functions characterizing individual coarse meshes which contain
absorber regions. These response functions can be used directly or indirectly as modified boundary
conditions to accurately capture the effects of control rods on the global flux solution.

Fen et al. \cite{fen_modelling_1992} developed the \gls{RMM} which generates modified boundary
conditions from response functions to treat absorber nodes in whole-core diffusion calculations.
The response functions relate the incident partial current on one face of the node to the resultant
outgoing partial currents on all four faces of the same node. To be more precise, each incident
partial current of each neutron energy group on each face may contribute to the outgoing partial
current of any energy group on any face of the same node. The following equation for the response
matrix $A$ encapsulates the response values to be generated from the \gls{RMM}:
%
\begin{gather}
  A^{jk}_{nm} = \frac{J^{+j}_n}{J^{-k}_m} \mbox{ for } j,k=1...G \mbox{ and } n,m=1...N
  \label{eq:rmm} \\
  \shortintertext{where}
  \begin{align}
    A =& \mbox{ response matrix,} \nonumber \\
    N =& \mbox{ number of spatial intervals along the perimeter of the absorber node,} \nonumber \\
    J^{-k}_m =& \mbox{ incident partial current in energy group $k$ at spatial interval $m$,}
      \nonumber \\
    J^{+j}_n =& \mbox{ outgoing partial current in energy group $j$ at spatial interval $n$}
      \nonumber \\
    &\mbox{ in response to $J^{-k}_m$.} \nonumber
  \end{align}
\end{gather}

The \gls{RMM} compares favorably against the \gls{MECS} because it captures non-isotropic flux
effects from the non-central control rod location within the reactor core
\cite{fen_modelling_1992}. The \gls{RMM} also does not require the
meticulous tuning of adjacent node thicknesses to obtain equivalent fluxes to apply the \gls{MECS}.
However, both \gls{RMM} and \gls{MECS} involve precalculation procedures that must be rerun if the
absorber region is subjected to significant changes such as control rod withdrawal.

Rahnema et al. \cite{rahnema_advanced_2011} later developed an \gls{IDT} method that embeds the
transport correction for absorber regions in the full-core nodal diffusion calculation. Instead of
modified boundary conditions, the \gls{IDT} method generates coupling coefficients that are
morphologically identical to those used in nodal diffusion methods. The response region, which
contains the absorber region, is subdivided into several nodes. The transport correction
relies on higher-order spatial and angular response moments to maintain detailed responses between
adjacent response nodes. Changes in the response region can be modeled by swapping out response
nodes without having to rerun the transport solver to generate new coupling coefficients. The
intra-response region calculations were iteratively coupled to the full-core diffusion calculations
to avoid introducing additional off-diagonal terms, which would increase the solve time of an
otherwise tri-diagonal system of nodal diffusion equations. In several verification studies of
static \gls{HTGR} core configurations, the \gls{IDT} method produced similar eigenvalue and flux
distribution results \cite{rahnema_advanced_2011} as the \gls{RMM}. However, they did not
demonstrate the response region swapping that the \gls{IDT} was designed for.

\subsubsection{Ronen Method}

Ronen \cite{ronen_accurate_2004} postulated an alternative formulation for diffusion coefficients
based on neutron currents from transport calculations.
Fick's law of diffusion is valid under three assumptions: the neutron flux gradient is small, the
absorption-to-scattering ratio is small, and scattering sources are isotropic. Therefore, diffusion
theory fails for anisotropic fluxes in and near absorber regions. The Ronen method proposes using
the integral form of the transport equation to derive transport operators for the neutron current
and substituting the values into Fick's first law of diffusion (Eq. \ref{eq:fick}) to obtain
space-dependent diffusion coefficients as follows:
%
\begin{align}
  D(\vec{r},E) =& -\frac{J_{tr}(\vec{r},E)}{\nabla \phi(\vec{r},E)}
  \label{eq:ronen}
  \shortintertext{where}
    J_{tr} =& \mbox{ transport-derived neutron current.} \nonumber
\end{align}

In doing so, the Ronen method provides pointwise corrections to the diffusion equation, overcoming
the small flux gradient limitation. Tomatis \& Dall'Osso \cite{tomatis_application_2011}
numerically implemented the Ronen method for a 1-D homogeneous slab with two energy groups and
isotropic scattering. Instead of using Eq. \ref{eq:ronen}, which showed instabilities near flat
flux regions where the denominator approaches zero, they calculated corrections to the diffusion
coefficients at cell interfaces using the difference between the transport- and diffusion-derived
currents as follows:
%
\begin{align}
  \delta D(x_{i+1/2},E) =& -\delta J(x_{i+1/2},E) \frac{(\Delta x_{i+1}+\Delta x_i)/2}{
  \phi(x_{i+1},E)+\phi(x_i,E)} \label{eq:tomatis}
  \shortintertext{where}
    x_i =& \mbox{ $i$-th spatial interval,} \nonumber \\
    \delta J(x,E) =& J_{tr}(x,E) - J_D(x,E), \nonumber \\
    \Delta x_i =& \mbox{ size of $i$-th spatial interval.} \nonumber
\end{align}

They derived expressions for the transport operators for $J_{tr}$ in terms of exponential integral
functions and Legendre expansions of the angular flux.
Gross et al. \cite{gross_high-accuracy_2020} extended the derivation of the transport operators to
handle 1-D heterogeneous problems in the form of fuel assemblies with fuel, water, and fuel+absorber
regions. Tomatis et al. \cite{tomatis_ronen_2021} developed new numerical implementations for 1-D
slab, cylindrical, and spherical geometries by employing probabilistic treatments from the
\gls{CPM} \cite{lewis_computational_1984} for the transport operators. The authors also implemented
a solver acceleration scheme which helped with the poor convergence rate observed in the earlier
studies \cite{tomatis_application_2011, gross_high-accuracy_2020}.

All three studies showed improvements in flux solutions with the Ronen method over
pure diffusion solvers in all of the test cases, particularly for a 1-D heterogeneous
\gls{BWR}-based problem with isotropic scattering and strong absorption cross sections
\cite{gross_high-accuracy_2020}. The error in reactivity values for three different \gls{BWR}
configurations differed from the reference $S_{16}$ transport calculations by at most 62 pcm after
one hundred iterations ($1$ pcm $=10^{-5}$) . However, the demonstrations were limited to simple
1-D geometries. Although the authors provided derivations for anisotropic scattering, their test
cases incorporated only isotropic scattering. The derivation of semi-analytic transport operators
from the integral form of neutron transport equations would be much more complicated
for more complex reactor geometries in 2-D or 3-D similar to how analytical solution forms do not
exist beyond those for simple geometries.

\subsubsection{Averaged Eddington Factors and High-Order Empirical Diffusion Coefficients}

Similarly, Pounders \& Rahnema developed two different methods for generating space-dependent
diffusion coefficients \cite{pounders_diffusion_2009}. Both methods possess the same caveat in
requiring \textit{a priori} knowledge of the flux and current. The first method, called the
\gls{AEF} method, relies on the Eddington factor, which is defined as the second angular moment of
the angular flux normalized by its zeroth moment. In 1-D, the Eddington factor is given as:
%
\begin{align}
  E_g(z) =& \frac{\int^1_{-1} \mu^2\psi(z,\mu)d\mu}{\int^1_{-1} \psi(z,\mu)d\mu}
\end{align}
%
The $g$ subscript denotes the discrete neutron group index of the multigroup diffusion equations
obtained from discretizing the continuous energy variable, as described in Section
\ref{sec:summary-nts-mtds}. The Eddington factor features in the first angular moment of the
1-D multigroup transport equations, obtained by multiplying the transport equation throughout by
the cosine of the scattering angle $\mu=\cos\theta$ and integrating over $\mu=-1$ to $\mu=1$:
%
\begin{gather}
  \frac{d\left[E_g(z)\phi_g(z)\right]}{dz} + \Sigma_{t,g}(z)J_g(z) = \sum^G_{g'=1}
  \Sigma^{g'\rightarrow g}_{s1}(z)J_{g'}(z) \label{eq:bte-1st-mom}
  \shortintertext{where}
    \Sigma^{g'\rightarrow g}_{s1} = \int^1_{-1} \mu \Sigma_s d\mu \nonumber
\end{gather}
%
Assuming the Eddington factor varies slowly in space, we can approximate Eq. \ref{eq:bte-1st-mom}
as:
%
\begin{gather}
  E_g(z)\frac{d\bar{\phi}_g(z)}{dz} + \Sigma_{t,g}(z)\bar{J}_g(z) = \sum^G_{g'=1}
  \Sigma^{g'\rightarrow g}_{s1}(z)\bar{J}_{g'}(z) \mbox{ for } z \in V_i \label{eq:bte-1st-est}
  \shortintertext{where}
  \begin{align}
    \bar{\phi}_g =& \mbox{ approximate flux solution of Eq. \ref{eq:bte-1st-est},} & \nonumber \\
    \bar{J}_g =& \mbox{ approximate current solution of Eq. \ref{eq:bte-1st-est},} & \nonumber \\
    V_i =& \mbox{ a subvolume of the system domain.} & \nonumber
  \end{align}
\end{gather}
The overbars distinguish the approximate solutions of Eq. \ref{eq:bte-1st-est} from the true
solution of Eq. \ref{eq:bte-1st-mom}. From Eq. \ref{eq:bte-1st-est}, the diffusion coefficient can
be defined as:
%
\begin{align}
  D_g(z) = E_g(z)\left[\Sigma_{t,g}(z)-\sum^g_{g'=1}\Sigma^{g'\rightarrow g}_{s1}(z)
  \frac{\bar{J}_{g'}(z)}{\bar{J}_g(z)}\right]^{-1} \label{eq:diffcoef-edd}
\end{align}
By further assuming that the Eddington factors are piecewise constant in each subvolume $V_i$, the
averaged Eddington factor $E^i_g$ can be evaluated as:
%
\begin{align}
  E^i_g =& \frac{E_g(z_{i+1})\phi_g(z_{i+1})-E_g(z_i)\phi_g(z_i)}{\bar{\phi}_g(z_{i+1})-
  \bar{\phi}_g(z_i)}
  \shortintertext{where}
  z_{i+1} =& \mbox{ upper bound of subvolume $V_i$,} \nonumber \\
  z_i =& \mbox{ lower bound of subvolume $V_i$.} \nonumber
\end{align}
%
The diffusion coefficient from Eq. \ref{eq:diffcoef-edd} can then be calculated as:
%
\begin{align}
  D^{AEF}_g(z) =& E^i_g\left[\hat{\Sigma}_{t,g}-\sum^G_{g'=1}\hat{\Sigma}^{g'\rightarrow g}_{s1}
  \frac{\hat{J}_{g'}}{\hat{J}_g}\right]^{-1}
  \shortintertext{where}
  \hat{\Sigma}_{t,g} =& \frac{\int_{V_i}\Sigma_{t,g}(z)J_g(z)dz}{\int_{V_i}\bar{J}_g(z)dz},
  \nonumber \\
  \hat{\Sigma}^{g'\rightarrow g}_{s1} =& \frac{\int_{V_i}\Sigma^{g'\rightarrow g}_{s1}(z)J_{g'}(z)
  dz}{\int_{V_i}\bar{J}_{g'}(z)dz}, \nonumber \\
  \hat{J}_g =& \int_{V_i} \bar{J}_g(z)dz. \nonumber
\end{align}

The second method employs a much more straightforward premise: Fick's law is assumed to be
accurate, with
high-order empirical diffusion coefficients to be determined. Given a known transport
solution, the following integration holds for a small homogeneous volume $V_i$:
%
\begin{align}
  \frac{1}{V_i}\int_{V_i}J_g(z)dV =& -\frac{1}{V_i}\int_{V_i}D_g(z)\frac{d\phi_g(z)}{dz}dV.
\end{align}
%
Taking $D_g$ to be constant in $V_i$, we may apply the divergence theorem to obtain
%
\begin{align}
  \bar{J}_gV_i =& -D^i_g\int_{\partial V_i} \phi_g(z)dA
  \shortintertext{where}
  \bar{J}_g =& \mbox{ average current in $V_i$,} \nonumber \\
  \partial V_i =& \mbox{ bounding surface of $V_i$.} \nonumber
\end{align}
%
Rearranging the terms, we may obtain the empirical diffusion coefficient formulation from the
transport-derived flux and current as follows:
%
\begin{align}
  D^i_g =& -\frac{\left(z_{i+1}-z_i\right) \bar{J}_g}{\left[\phi_g(z_{i+1})-\phi_g(z_i)\right]}
  \label{eq:emp}
\end{align}
%
Both \gls{AEF} and empirical methods require the respective subvolumes $V_i$ to be small enough
for the assumptions of constant Eddington factors and diffusion coefficients to hold within $V_i$.

Both methods performed much better than conventional diffusion coefficients derived using the
$P_1$ approximation method. For the same 1-D heterogeneous \gls{BWR} problem demonstrating the prior
Ronen method \cite{gross_high-accuracy_2020} but with anisotropic scattering, the absolute errors
in $k_{eff}$ are around 100 pcm after eliminating errors associated with energy group
condensation. The error values compare favorably with the 159-683 pcm error with the $P_1$ method.
The \gls{AEF} and empirical methods also reproduced the flux distributions better with a maximum
pointwise error of 2\% as opposed to 5\% from the $P_1$ method. Similar to the Ronen method, the
\gls{AEF} and empirical methods introduce pointwise corrections with information from transport
methods. However, both \gls{AEF} and empirical methods require a priori knowledge of the true
solution or otherwise
accurate estimates from transport methods. In contrast, the Ronen method relies on analytical
transport operators, which use diffusion flux estimates from the previous iteration to update the
flux solution. Nevertheless, the \gls{AEF} and empirical methods provide the foundation for further
development of practical, self-closing transport-correction techniques.

\subsubsection{General Equivalence Theory and Superhomogenization Method}

As mentioned in Section \ref{sec:summary-nts-mtds}, coarse-mesh and nodal methods homogenize
heterogeneities in reactor geometries to reduce the computational costs of running full-core
simulations. Equivalence techniques which reduce spatial homogenization error are also effective
for accurately modeling the integral worths of control rods within fuel assemblies. The \gls{GET}
\cite{koebke_new_1980, smith_nodal_1983} and \gls{SPH} \cite{kavenoky_sph_1978,
hebert_consistent_1991} methods represent the most widely used equivalence methods for improving
the performance of diffusion calculations in homogenized \gls{LWR} models. Both methods involve
deriving additional homogenization parameters from single-assembly transport calculations with
reflective boundary conditions. Since the transport calculation step is already a prerequisite
for generating homogenized group constants, the equivalence methods are simple to implement and
impose reasonably small additional computational costs.

Koebke \cite{koebke_new_1980} first proposed abandoning continuous surface fluxes to preserve net
surface currents through discontinuity factors. Smith \cite{smith_assembly_1986}
later extended this concept to assembly-homogenized calculations. The discontinuity factors are
calculated for each face of the homogenized region to preserve volumetric reaction rates and
surface neutron currents. The discontinuity factors are calculated as follows:
%
\begin{gather}
  DF = \frac{\phi^{het,sur}}{\phi^{hom,sur}}
  \shortintertext{where}
  \begin{align}
    DF =& \mbox{ discontinuity factor,} \nonumber \\
    \phi^{het,sur} =& \mbox{ surface flux of the region from the heterogeneous calculation,}
    \nonumber \\
    \phi^{hom,sur} =& \mbox{ surface flux of the region from the homogeneous calculation.}
    \nonumber
  \end{align}
\end{gather}
%
\gls{GET} was later extended for fine mesh calculations \cite{yamamoto_cell_2004}, which refers to
cell-level homogenization; each fuel cell in an assembly, consisting of the fuel pellet, cladding,
and moderator, is individually homogenized as opposed to lumping together the entire assembly.

In contrast, the \gls{SPH} method, first proposed by Kavenoky \cite{kavenoky_sph_1978} for
irregular lattices and later applied as a cell-homogenization technique by Hebert
\cite{hebert_consistent_1991}, introduces correction factors to homogeneous cross sections as
follows:
%
\begin{gather}
  \tilde{\Sigma}^{hom}_k = \mu_k \Sigma^{hom}_k
  \shortintertext{where}
  \begin{align}
    \tilde{\Sigma}^{hom}_k =& \mbox{ \gls{SPH}-corrected homogeneous cross section for region $k$,}
    \nonumber \\
    \mu_k =& \mbox{ \gls{SPH} correction factor for region $k$,} \nonumber \\
    \Sigma^{hom}_k =& \mbox{ uncorrected homogeneous cross section for region $k$.} \nonumber
  \end{align}
\end{gather}
%
The \gls{SPH} factor is calculated as follows:
%
\begin{gather}
  \mu_k = \frac{\bar{\phi}^{het}_k}{\phi^{hom}_k}
  \shortintertext{where}
  \begin{align}
    \bar{\phi}^{het}_k =& \mbox{ average flux in region $k$ from the heterogeneous calculation,}
    \nonumber \\
    \phi^{hom}_k =& \mbox{ average flux in region $k$ from the homogeneous calculation}
    \nonumber \\
                  & \mbox{ with \gls{SPH}-corrected cross sections.} \nonumber
  \end{align}
\end{gather}

The \gls{GET} and the \gls{SPH} methods are designed to preserve reaction rates at the assembly
level \cite{yamamoto_cell_2004}. Given that the \gls{SPH} method introduces only one correction
factor per cell, it only preserves the average net current across all surfaces as opposed
to the net current at each surface with \gls{GET}. Similarly, the isotropic nature of \gls{SPH}
factors leads to worse pin-power estimates near control rods and reflectors compared to \gls{GET}.
However, the \gls{SPH} method is simpler to implement in the diffusion solver as the \gls{SPH}
factors can be precombined with the cross sections to generate \gls{SPH}-corrected cross sections.
\gls{GET} requires diffusion solvers, which allow for flux discontinuities at the interfaces.
Furthermore, \gls{GET} requires more memory to store up to six discontinuity factors, one for each
mesh surface, in 3-D calculations.

Modeling control rods in full-core calculations with the \gls{SPH} method or \gls{GET} provides
better estimates of the multiplication factor and the flux distribution than reference diffusion
solutions. In this regard, the correction factors do not distinguish between errors arising from
homogenization or the diffusion approximation. However, the improved solutions, especially with the
\gls{SPH} method, can still deviate significantly from reference transport calculations. A
supercell, similar to the one adopted by \gls{MECS}, can be used to reduce errors within and near
control rod cells \cite{ortensi_newton_2018}. Another disadvantage, shared by other schemes
dependent on homogenization, is that homogenization removes some of the heterogeneity in the flux
and temperature.

\subsubsection{Multilevel Quasi-Diffusion Method}

In 1964, Gol'din \cite{goldin_quasi-diffusion_1964} originally developed the \gls{QD} method
as an acceleration scheme for the solution of the neutron transport equation. Consider the
steady-state, single-group neutron transport equation:
%
\begin{gather}
  \hat{\Omega}\cdot\nabla\psi(\vec{r},\hat{\Omega}) + \Sigma_t(\vec{r})\psi(\vec{r},\hat{\Omega}) =
  \int_{4\pi}d\hat{\Omega}'\ \Sigma_s(\vec{r},\hat{\Omega}'\rightarrow\hat{\Omega})
  \psi(\hat{\Omega}') + Q(\vec{r},\hat{\Omega}) \label{eq:qd-high}
  \shortintertext{where}
  Q = \mbox{ external source term.} \nonumber
\end{gather}
%
Take the zeroth and first angular moments to obtain the following two equations, respectively:
%
\begin{gather}
  \nabla\cdot J(\vec{r}) + \Sigma_t(\vec{r})\phi(\vec{r}) = \Sigma_{s,0}(\vec{r})\phi(\vec{r}) +
  Q_0(\vec{r}) \label{eq:qd-low-0th} \\
  \nabla\cdot\left(\vec{\vec{E}}(\vec{r})\phi(\vec{r})\right) + \Sigma_t(\vec{r})J(\vec{r}) =
  \Sigma_{s,1}(\vec{r})J(\vec{r}) + Q_1(\vec{r}) \label{eq:qd-low-1st}
  \shortintertext{where}
  \begin{align*}
    \phi(\vec{r}) =& \int_{4\pi}d\hat{\Omega}\ \psi(\vec{r}), & \\
    J(\vec{r}) =& \int_{4\pi}d\hat{\Omega}\ \hat{\Omega}\psi(\vec{r}), & \\
    Q_0(\vec{r}) =& \int_{4\pi}d\hat{\Omega}\ Q(\vec{r},\hat{\Omega}), & \\
    Q_1(\vec{r}) =& \int_{4\pi}d\hat{\Omega}\ \hat{\Omega}Q(\vec{r},\hat{\Omega}), & \\
    \Sigma_{s,l} =& \mbox{ $l$th Legendre moment of the scattering cross section $\Sigma_s$,} & \\
    \vec{\vec{E}} =& \frac{\int_{4\pi}d\hat{\Omega}\ \hat{\Omega}\hat{\Omega}\psi(\vec{r})}{
    \int_{4\pi}d\hat{\Omega}\ \psi(\vec{r})}. &
  \end{align*}
\end{gather}
%
The Eddington tensor $\vec{\vec{E}}$ captures flux anisotropies from the high-order transport
solution of Eq. \ref{eq:qd-high} and enables the low-order Eqs. \ref{eq:qd-low-0th} \&
\ref{eq:qd-low-1st} to accurately reproduce the high-order flux solution. These equations form a
two-level acceleration scheme. The solution algorithm is as follows:
%
\begin{enumerate}
  \item Initialize $\phi^l$ with an initial scalar flux distribution $\phi^0$
  \item Solve Eq. \ref{eq:qd-high} for $\psi^{l+1/2}$ using $\phi^l$
  \item Generate Eddington factors $\vec{\vec{E}}$ using $\psi^{l+1/2}$
  \item Solve Eqs. \ref{eq:qd-low-0th} \& \ref{eq:qd-low-1st} for $\phi^{l+1}$
  \item Repeat Steps 2-4 until flux convergence is reached
\end{enumerate}

The high- and low-order formulations produce the same scalar flux solution only if they share the
same spatial discretization scheme. Otherwise, the fluxes will differ by a spatial truncation
error \cite{adams_fast_2002}. The Eddington tensor converges significantly faster than the angular
fluxes, thus accelerating the neutron transport method by effectively replacing some of the
high-order iterations with the low-order iterations. Gol'din \cite{anistratov_solution_1986} later
extended the \gls{QD} method for multigroup neutron transport problems with anisotropic scattering.
\gls{QD} methods for multigroup neutron transport problems typically adopt three levels instead of
the two levels described by Eqs. \ref{eq:qd-high}, \ref{eq:qd-low-0th}, and \ref{eq:qd-low-1st}.
The three levels consist of the multigroup high-order neutron transport equations, the multigroup
low-order \gls{QD} equations, and the effective gray (one-group) low-order \gls{QD} equations
\cite{anistratov_solution_1986}. The gray low-order equations are derived through flux-averaging of
the multigroup neutron group constants and a closure relation across all energy groups.

Tamang \& Anistratov \cite{tamang_multilevel_2014} applied the multilevel \gls{QD} method to
time-dependent multiphysics neutron transport and heat transfer problems by iteratively coupling
the gray low-order equations to the heat transfer equations. Their coupling algorithm takes
advantage of the lower computational cost of the low-order equations to quickly resolve
multiphysics interactions during inner iterations, thereby requiring fewer outer iterations and
high-order transport calculations. Reynolds \& Palmer \cite{reynolds_analysis_2023} implemented the
multilevel \gls{QD} method in RZ cylindrical coordinates and the capability to model \gls{DNP} flow
for simulating molten salt reactor kinetics.

The multilevel \gls{QD} method may accurately model control rod provided that the method utilizes
sufficient fidelity in the high-order transport calculations. In particular, Reynolds \& Palmer
demonstrated this through a steady-state simulation of a 2-D axisymmetric, two-group \gls{MSRE}
model with a partially-inserted control rod. Their \gls{QD} method implementation with the $S_{12}$
method for the high-order transport iterations allowed for more accurate neutron flux estimates
near the control rod compared to a $P_1$ method. However, multilevel \gls{QD} methods still have
computational costs comparable to accelerated neutron transport methods, which remain too
computationally intensive for time-dependent 3-D full-core calculations.

\subsubsection{Hybrid Methods via Spatial Domain Decomposition}

Given the high computational costs of neutron transport calculations, hybrid methods have been
developed by decomposing the problem domain into \textit{transport} and \textit{diffusive}
subdomains. The neutron flux in diffusive subdomains may be resolved sufficiently accurately using
low-order methods such as neutron diffusion, $P_1$, or $S_2$ to provide significant computational
cost savings.

One type of approach for hybrid methods involves making predetermined domain decompositions prior
to the neutron flux calculation. These methods would require interface conditions at the
transport-diffusion interfaces to couple adjoining subdomains. Griffin
\cite{prince_neutron_2024}, a \gls{MOOSE}-based reactor multiphysics software, employs this
approach for its $S_N$-$P_N$ \textit{multischeme} transport method. Griffin provides two options
for the interface
conditions: the Lagrange multiplier and the upwinding methods \cite{wang_hybrid_2017}. The Lagrange
multiplier method enforces the continuity of angular flux moments at the interface through Lagrange
multipliers minimizing the differences in the moments at the interface. The angular flux moments on
$S_N$ side typically require smoothing due to the $S_N$ method's discrete treatment of the angular
dependence. The upwinding method is a more conventional scheme that maintains
causality by generating surface source boundaries on
either side of the interface based on the outgoing neutron currents from the opposing neutron
transport scheme. The upwinding method does not require smoothing on the $S_N$ side of the
interface and is applicable to discontinuous numerical solvers, but the flux solutions are
generally discontinuous across the interface. The weak formulation of the Lagrange multiplier
interface condition is:
%
\begin{gather}
  -\int_S \left[\left(\llbracket\Psi\rrbracket,\Lambda^*\right)_{\Gamma} +
  \left(\llbracket\Psi^*\rrbracket,\Lambda\right)_{\Gamma}\right] d\hat{\Omega}
  \shortintertext{where}
  \begin{align}
    \llbracket\Psi\rrbracket &= \Psi^+ - \Psi^-, \nonumber \\
    \Psi^\pm &= \lim_{s\rightarrow 0^+} \Psi(\vec{x}+s\hat{n}), \nonumber \\
    \Lambda &= \mbox{ Lagrange multiplier}, \nonumber \\
    \Gamma &= \mbox{ interface boundary}, \nonumber \\
    \hat{n} &= \mbox{ normal vector at interface boundary}, \nonumber
  \end{align}
\end{gather}
%
and the weak formulation of the upwinding method is:
%
\begin{gather}
  \int_S \left(|\hat{\Omega}\cdot\hat{n}|\llbracket\Psi^*\rrbracket,\Psi^-\right)_{\Gamma^+}
  d\hat{\Omega} -
  \int_S\left(|\hat{\Omega}\cdot\hat{n}|\llbracket\Psi^*\rrbracket,\Psi^+\right)_{\Gamma^-}
  d\hat{\Omega}
\end{gather}
%
The terms with asterisks denote test functions used in the \gls{FEM} framework.

Two recent works by Anistratov \& Stehle \cite{anistratov_computational_2012} and Stehle et al.
\cite{stehle_hybrid_2014} describe another approach that is distinct from the prior approach in
two ways. Firstly, they developed an on-the-fly scheme for estimating a transport metric which can
indicate whether a subregion is diffusive. The solver uses this information to apply adaptive
domain decomposition and decide which subregions require transport calculations for solution
accuracy. Secondly, they employ a two-level neutronics method similar to the multilevel \gls{QD}
method to obtain the neutron flux solution. The hybrid method evaluates the low-order neutron
diffusion problem on the entire problem domain while limiting high-order transport calculations to
\textit{transport} subregions. The low-order formulation also serves as an acceleration scheme to
the high-order formulation to provide greater computational cost savings. The authors showed the
viability of using Eddington tensor estimates as the metric for quantifying transport effects in
a subregion. The Eddington tensor is identical to one defined for the \gls{QD} scheme. The
two-level neutronics method is based on the \textit{Second-Moment} method developed by Lewis \&
Miller \cite{lewis_comparison_1976}. The low-order second-moment equations are:
%
\begin{gather}
  \nabla\cdot J^{(s+1)}+\Sigma_a\phi^{(s+1)} = q, \\
  \frac{1}{3}\nabla\cdot\phi^{(s+1)}+\Sigma_t J^{(s+1)} = \nabla\cdot F^{(s+1/2)}
  \shortintertext{where}
  F^{(s+1/2)}_{\alpha\beta} = \int_{4\pi}
  \left(\frac{1}{3}\delta_{\alpha\beta}-\Omega_\alpha\Omega_\beta\right)
  \Psi^{(s+1/2)} d\Omega
\end{gather}
%
$(s)$ denotes the iteration index. $F$ is a transport correction tensor calculated from the
high-order neutron transport equations in \textit{transport} subregions. $F$ is simply set to zero
in \textit{diffusive} subregions. The scalar flux from the low-order equations is also used to
evaluate the 0th-order scattering source term and accelerate the high-order calculations for
$\Psi^{(s+1/2)}$. Stehle et al. \cite{stehle_hybrid_2014} demonstrated their hybrid method on two
2-D, single-group test problems. Both problems consisted of two regions with idealized macroscopic
cross sections representing \textit{transport} and \textit{diffusive} regions with high absorption
and scattering ratios, respectively.

\section{Summary}

In the last two decades, many \gls{MSR} multiphysics software have been developed by
extending existing reactor software such as DALTON and \gls{TRACE} for \gls{MSR} modeling or
building on general multiphysics software frameworks such as OpenFOAM and \gls{MOOSE}. The
resultant software have a wide range of capabilities arising from their calculation schemes,
constraints from legacy code, and simplifying assumptions to balance computational costs. Several
research efforts have been targeted at establishing benchmarks for \gls{MSR} simulation tool
\gls{VV}, notably with the CNRS benchmark \cite{tiberga_results_2020} and neutronics benchmarks of
the \gls{MSRE} \cite{fratoni_molten_2020} and the \gls{MSFR} \cite{brovchenko_neutronic_2019}.
While these efforts have plugged significant technical gaps, more can be done to develop open
\gls{VV} benchmarks for multiphysics \gls{MSR} modeling, particularly for model verification of
\gls{DNP} drift out of the active core region. Among other considerations, the proposed
verification model should provide the necessary group constant input data for other multiphysics
simulation tools to replicate the results without discrepancies propagating from using different
nuclear databases or stochastic uncertainties from Monte Carlo simulations.

On the \gls{TH}-modeling side of \gls{MSR} multiphysics, turbulent flow is a complex, yet important
phenomenon to include in our simulation models to obtain accurate predictions of \gls{MSR} behavior
under various operating conditions. Most \gls{MSR} multiphysics studies with turbulence models rely
on the two-equation turbulence models, which provide acceptable estimates of time-averaged flow
characteristics. One study \cite{podila_cfd_2019} showed close agreement in the\gls{MSRE}
temperature distribution with various one-equation, two-equation, and \gls{RSM} turbulence
models despite relatively larger differences observed in turbulence intensities near the walls.
This observation suggests one-equation models may be adequate for \gls{MSR} designs and simulations
with weaker turbulent effects, although other analyses may require even more accurate turbulence
models like \gls{DES} to capture turbulent effects without time averaging adequately.

Lastly, there is a lack of research focus into robust techniques for accurate control rod modeling
for time-dependent \gls{MSR} analyses. Control rods induce highly anisotropic fluxes within
themselves and their vicinities.
Unfortunately, higher-fidelity neutron transport methods are typically too computationally
expensive for multiphysics \gls{MSR} simulations, while faster methods like neutron
diffusion perform poorly in regions with highly anisotropic fluxes. However, transport-correction
techniques can be applied to augment neutron diffusion calculations in these regions. While few
have been applied to \gls{MSR} multiphysics simulations other than albedo boundary conditions,
various techniques have been proposed and implemented for other reactor types. Helpful information
may be gleaned from these techniques to improve control rod modeling in \glspl{MSR} with the
neutron diffusion method.

### 2024-05-28 lit.tex
saaf sn method

@@
 spatial discretization schemes to meet the requisite resolution of each
-individual physics component.
+individual physics model.
 
@@
   \int^\infty_0 dE' \int_{4\pi} d\hat{\Omega}' \nu\Sigma_f(\vec{r},E',t) \Psi(\vec{r},E',
-  \hat{\Omega}',t)+S(\vec{r},E,\hat{\Omega}',t) \label{eq:bte}
+  \hat{\Omega}',t)+S(\vec{r},E,\hat{\Omega},t) \label{eq:bte}
   \shortintertext{where}
@@
   \sum^G_{g'=1} \int_{4\pi} d\hat{\Omega}' \nu\Sigma_{f,g'}(\vec{r},t) \Psi_{g'}(\vec{r},
-  \hat{\Omega}',t)+S_g(\vec{r},\hat{\Omega}',t) \label{eq:mg-bte}
+  \hat{\Omega}',t)+S_g(\vec{r},\hat{\Omega},t) \label{eq:mg-bte}
   \shortintertext{where}
@@
 on the two-equation turbulence models, which provide acceptable estimates of time-averaged flow
-characteristics. One study \cite{podila_cfd_2019} showed close agreement in the\gls{MSRE}
+characteristics. One study \cite{podila_cfd_2019} showed close agreement in the \gls{MSRE}
 temperature distribution with various one-equation, two-equation, and \gls{RSM} turbulence

### 2024-06-10 hybrid.tex
hybrid chapter additions

Section \ref{sec:summary-nts-mtds} highlighted the poor performance of neutron diffusion
methods for calculating neutron fluxes near control rods. Strong neutron absorption in the control
rod region produces a highly anisotropic neutron flux extending some distance outside the control
rod. Neutron transport methods, which retain angular dependence of the neutron flux to various
extents, generally fare better than neutron diffusion methods, which have isotropic diffusion
coefficients. However, neutron transport methods are also generally more computationally expensive,
given the increased dimensionality of the problem from the angular component. Adding an angular
dimension to the existing geometric and neutron energy group dimensions dramatically
increases the problem size and the computational resources necessary to solve the system. Many past
efforts have tried introducing
transport correction techniques to improve neutron flux and multiplication factor estimates in
diffusion-based methods. Other than control rod regions, these techniques may also correct
homogenization errors introduced by spatial homogenization of fuel assemblies and other
structures within a reactor core. They invariably rely on neutron transport methods to generate
transport corrections in the form of corrected diffusion coefficients
\cite{bretscher_computing_1997, scherer_determination_1976, ronen_accurate_2004,
pounders_diffusion_2009, kavenoky_sph_1978}, boundary conditions \cite{davison_influence_1951,
pellaud_extrapolation_1968, fen_modelling_1992}, Eddington factors, or discontinuity factors
\cite{koebke_new_1980}.

In this chapter, I propose a hybrid neutronics method for improving control rod modeling in neutron
diffusion solvers without spatial homogenization. In essence, the hybrid method involves applying
the $S_N$ discrete ordinates neutron transport method on subregions containing the control rod to
obtain pointwise transport corrections for the diffusion method on the subregions.

Section \ref{sec:theory} discusses the theoretical background for the hybrid $S_N$-diffusion
method. Section \ref{sec:implementation} provides numerical implementation details of
the hybrid method and its components. Sections \ref{sec:test-case} and \ref{sec:sim-param} describe
the 1-D test cases and several simulation parameters in that context. Section
\ref{sec:prelim-results} discusses the results of the hybrid method applied to the 1-D test cases
with comparisons to higher-fidelity Monte Carlo and $S_N$ neutron transport methods. Lastly,
Section \ref{sec:hybrid-summary} summarizes the key findings in this chapter.

\section{Theory} \label{sec:theory}

The proposed hybrid $S_N$-diffusion method is an iterative method to improve the
accuracy of neutron diffusion solutions in reactor systems with highly anisotropic flux regions
which I will use to model control rods. This section presents the mathematical derivation for the
\gls{SAAF} formulation of the $S_N$ equations, the neutron diffusion equations, the drift
transport correction term, and the computational algorithm for the hybrid method.

\subsection{\gls{SAAF} Formulation of the $S_N$ Method}

Continuing from the introduction to neutronics methods in Section \ref{sec:summary-nts-mtds},
the time-dependent, multigroup neutron transport equation defined on the 3-D spatial domain
$\mathcal{D}$ and 2-D unit sphere angular domain $\mathcal{S}$ is:
%
\begin{multline}
  \frac{\partial}{\partial t}\left[\frac{\Psi_g(\vec{r},\hat{\Omega},t)}{v_g}\right] +
  \hat{\Omega}\cdot\nabla\Psi_g(\vec{r},\hat{\Omega},t) + \Sigma_{t,g}
  \Psi_g(\vec{r},\hat{\Omega},t) =
  \sum^G_{g'=1}\int_\mathcal{S} \Sigma_{s,g'\rightarrow g}(\hat{\Omega}'\rightarrow\hat{\Omega})
  \Psi_{g'}(\vec{r},\hat{\Omega}',t)d\hat{\Omega}' \\
  + \frac{1}{4\pi}\chi_{p,g}(1-\beta)\sum^G_{g'=1} \nu\Sigma_{f,g'} \phi_{g'}(\vec{r},t)
  + \frac{1}{4\pi}\sum^I_{i=1}\chi_{d,g}
  \lambda_i C_i(\vec{r},t) + Q^{\text{ext}}_g(\vec{r},\hat{\Omega})
  \label{eq:mg-nte}
\end{multline}
%
with the boundary conditions
%
\begin{gather}
  \Psi_g(\vec{r},\hat{\Omega}) = \Psi^\text{inc}_g(\vec{r},\hat{\Omega}) +
  \alpha^s_g\Psi_g(\vec{r},\hat{\Omega}_r)
  \mbox{ on } \vec{r} \in \partial\mathcal{D} \mbox{ and } \hat{\Omega}\cdot\hat{n}_b < 0,
  \shortintertext{where}
  \begin{align*}
    \chi_{p,g} &= \mbox{prompt fission neutron spectrum in group $g$,} \\
    \beta &= \sum^I_{i=1} \beta_i = \mbox{total delayed neutron fraction,} \\
    \chi_{d,g} &= \mbox{delayed fission neutron spectrum in group $g$,} \\
    \lambda_i &= \mbox{decay constant of precursor group $i$,} \\
    C_i &= \mbox{delayed neutron precursor concentration for group $i$,} \\
    \Psi^\text{inc}_g &= \mbox{incident surface source in group $g$,} \\
    \alpha^s_g &= \mbox{specular reflectivity on }\partial \mathcal{D} \mbox{ for group }g, \\
    \hat{\Omega}_r &= \hat{\Omega}-2(\hat{\Omega}\cdot \hat{n}_b)\hat{n}_b, \\
    \hat{n}_b &= \mbox{outward unit normal vector on the boundary.}
  \end{align*}
\end{gather}
%
In order to introduce operators and facilitate subsequent mathematical derivations, we will define
the following vector forms:
%
\begin{gather}
  \bm{\Psi} =
  \begin{bmatrix}
    \Psi_1 \\
    \Psi_2 \\
    \vdots \\
    \Psi_G
  \end{bmatrix},
  \bm{\phi} = \int_S \bm{\Psi}d\hat{\Omega} =
  \begin{bmatrix}
    \phi_1 \\
    \phi_2 \\
    \vdots \\
    \phi_G
  \end{bmatrix},
  \bm{\frac{\Psi}{v}}
  \begin{bmatrix}
    \frac{\Psi_1}{v_1} \\
    \frac{\Psi_2}{v_2} \\
    \vdots \\
    \frac{\Psi_G}{v_G}
  \end{bmatrix},
  \bm{Q}^{\text{ext}} =
  \begin{bmatrix}
    Q^\text{ext}_1 \\
    Q^\text{ext}_2 \\
    \vdots \\
    Q^\text{ext}_G
  \end{bmatrix}, \nonumber \\
  \bm{Q}_f = \frac{1}{4\pi}\bm{Q}_{f,0} =
  \begin{bmatrix}
    \frac{1}{4\pi}\chi_{p,1}(1-\beta)\sum^G_{g'=1} \nu\Sigma_{f,g'} \phi_{g'} +
    \frac{1}{4\pi}\sum^I_{i=1}\chi_{d,1} \lambda_i C_i \\
    \frac{1}{4\pi}\chi_{p,2}(1-\beta)\sum^G_{g'=1} \nu\Sigma_{f,g'} \phi_{g'} +
    \frac{1}{4\pi}\sum^I_{i=1}\chi_{d,2} \lambda_i C_i \\
    \vdots \\
    \frac{1}{4\pi}\chi_{p,G}(1-\beta)\sum^G_{g'=1} \nu\Sigma_{f,g'} \phi_{g'} +
    \frac{1}{4\pi}\sum^I_{i=1}\chi_{d,G} \lambda_i C_i
  \end{bmatrix},
  \bm{Q} = \bm{Q}_f + \bm{Q}^\text{ext}. \nonumber
\end{gather}
%
We also define the following operators:
%
\begin{gather}
  \mathbb{L}_1\bm{\Psi} =
  \begin{bmatrix}
    \hat{\Omega}\cdot\nabla\Psi_1 \\
    \hat{\Omega}\cdot\nabla\Psi_2 \\
    \vdots \\
    \hat{\Omega}\cdot\nabla\Psi_G \\
  \end{bmatrix},
  \mathbb{L}_2\bm{\Psi} =
  \begin{bmatrix}
    \Sigma_{t,1}\Psi_1 \\
    \Sigma_{t,2}\Psi_2 \\
    \vdots \\
    \Sigma_{t,G}\Psi_G
  \end{bmatrix},
  \mathbb{L}\bm{\Psi} = \mathbb{L}_1\bm{\Psi} + \mathbb{L}_2\bm{\Psi}, \nonumber \\
  \mathbb{S}\bm{\Psi} =
  \begin{bmatrix}
    \sum^G_{g'=1}\int_S \Sigma_{s,g'\rightarrow 1}\Psi_{g'}d\hat{\Omega} \\
    \sum^G_{g'=2}\int_S \Sigma_{s,g'\rightarrow 2}\Psi_{g'}d\hat{\Omega} \\
    \vdots \\
    \sum^G_{g'=G}\int_S \Sigma_{s,g'\rightarrow G}\Psi_{g'}d\hat{\Omega}
  \end{bmatrix},
%  \mathbb{F}_p\bm{\Psi} = \frac{1}{4\pi} \mathbb{F}_{p,0}\bm{\phi} =
%  \begin{bmatrix}
%    \frac{1}{4\pi}\chi_{p,1}(1-\beta)\sum^G_{g'=1}\nu\Sigma_{f,g'}\phi_{g'} \\
%    \frac{1}{4\pi}\chi_{p,2}(1-\beta)\sum^G_{g'=1}\nu\Sigma_{f,g'}\phi_{g'} \\
%    \vdots \\
%    \frac{1}{4\pi}\chi_{p,G}(1-\beta)\sum^G_{g'=1}\nu\Sigma_{f,g'}\phi_{g'}
%  \end{bmatrix},
  \mathbb{B}\bm{\Psi} =
  \begin{bmatrix}
    \alpha^s_1\Psi_1(\hat{\Omega}_r) \\
    \alpha^s_2\Psi_2(\hat{\Omega}_r) \\
    \vdots \\
    \alpha^s_G\Psi_G(\hat{\Omega}_r)
  \end{bmatrix}. \nonumber
\end{gather}
%
Note that the operators apply element-wise multiplication as opposed to the more conventional
matrix multiplication. Eq.\ \ref{eq:mg-nte} can be reexpressed as:
%
\begin{gather}
  \frac{\partial}{\partial t} \left(\frac{\bm{\Psi}}{\bm{v}}\right)+\mathbb{L}_1\bm{\Psi}
  +\mathbb{L}_2\bm{\Psi} = \mathbb{S}\bm{\Psi} + \bm{Q}_f + \bm{Q}^\text{ext} \label{eq:nte-vec}
\end{gather}
%
with the boundary conditions on $\partial\mathcal{D}$
%
\begin{gather}
  \bm{\Psi} = \bm{\Psi}^\text{inc} + \mathbb{B}\bm{\Psi}
\end{gather}
%
We begin the derivation for the \gls{SAAF} formulation by rearranging Eq.\ \ref{eq:nte-vec} as
%
\begin{gather}
  \bm{\Psi} = \mathbb{L}^{-1}_2\left[\mathbb{S}\bm{\Psi}+\bm{Q}
  -\frac{\partial}{\partial t}\left(\frac{\bm{\Psi}}{\bm{v}}\right)-\mathbb{L}_1\bm{\Psi}\right]
  \label{eq:afe}
  \shortintertext{where}
  \mathbb{L}^{-1}_2 =
  \begin{bmatrix}
    \frac{1}{\Sigma_{t,1}} \\
    \frac{1}{\Sigma_{t,2}} \\
    \vdots \\
    \frac{1}{\Sigma_{t,G}}
  \end{bmatrix}. \nonumber
\end{gather}
%
We substitute Eq.\ \ref{eq:afe} back into the streaming term $\left(\mathbb{L}_1\bm{\Psi}\right)$
in Eq.\ \ref{eq:nte-vec} and rearrange the terms to obtain
%
\begin{gather}
  \frac{\partial}{\partial t}\left(\frac{\bm{\Psi}}{\bm{v}}\right) + \mathbb{L}^{-1}_2\mathbb{L}_1
  \mathbb{S}\bm{\Psi} + \mathbb{L}^{-1}_2\mathbb{L}_1\bm{Q} - \mathbb{L}^{-1}_2\mathbb{L}_1
  \frac{\partial}{\partial t}\left(\frac{\bm{\Psi}}{\bm{v}}\right) - \mathbb{L}^{-1}\mathbb{L}^2_1
  \bm{\Psi} + \mathbb{L}_2\bm{\Psi} = \mathbb{S}\bm{\Psi}+\bm{Q} \\
  \left(\mathbb{I} - \mathbb{L}^{-1}_2\mathbb{L}_1\right) \frac{\partial}{\partial t}\left(
  \frac{\bm{\Psi}}{\bm{v}}\right) - \mathbb{L}^{-1}_2\mathbb{L}^2_1\bm{\Psi} +
  \mathbb{L}_2\bm{\Psi} = \left(\mathbb{I} - \mathbb{L}^{-1}_2\mathbb{L}_1\right)
  \mathbb{S}\bm{\Psi} + \left(\mathbb{I} - \mathbb{L}^{-1}_2\mathbb{L}_1\right) \bm{Q}
%  \mathbb{L}^{-1}_2\mathbb{L}\frac{\partial}{\partial t}\frac{\bm{\Psi}}{\bm{v}} -
%  \mathbb{L}^{-1}_2\mathbb{L}^2_1 \bm{\Psi} + \mathbb{L}_2\bm{\Psi} = \mathbb{L}^{-1}_2\mathbb{L}
%  \mathbb{S}\bm{\Psi} + \mathbb{L}^{-1}_2\mathbb{L}\bm{Q}
  \shortintertext{where}
  \begin{align*}
    \mathbb{I} &= \mbox{identity operator,} \\
    \mathbb{I}-\mathbb{L}^{-1}_2\mathbb{L}_1 &=
    \mathbb{L}^{-1}_2\mathbb{L}_2-\mathbb{L}^{-1}_2\mathbb{L}_1 = 
  \end{align*}
\end{gather}
%

The discrete ordinates ($S_N$) method for solving the multigroup neutron transport equation
(Eq.\ \ref{eq:mg-bte}) discretizes the continuous angular directional phase space into a finite
number of discrete angular intervals (ordinates). The set of distinct direction variables
$\hat{\Omega}_d$ representing the discrete ordinates is typically chosen in conjunction with a
compatible quadrature set to replace integrals across continuous solid angles
$\hat{\Omega}$ with quadrature approximations. The multigroup, discrete ordinates ($S_N$
approximation) form of the 3-D time-dependent neutron transport equation with Legendre polynomial
expansion of the scattering cross section and spherical harmonic expansion of the flux in the
scattering source term is:
%
\begin{gather}
  \frac{\partial}{\partial t} \left(\frac{\Psi_{g,d}(\vec{r},t)}{v_g}\right) +
  \hat{\Omega}_d\cdot\nabla\Psi_{g,d}(\vec{r},t) + \Sigma_{t,g}(\vec{r},t)\Psi_{g,d}(\vec{r},t) =
  \nonumber \\
  \sum^G_{g'=1}\sum^L_{l=0} \frac{2l+1}{4\pi}\Sigma^{g'\rightarrow g}_{s,l}(\vec{r},t)\sum^l_{m=-l}
  \phi_{g',l,m}(\vec{r},t)Y_{l,m}(\hat{\Omega}_d) + \frac{\chi_g(\vec{r},t)}{4\pi}\sum^G_{g'=1}
  \nu\Sigma_{f,g'}(\vec{r},t) \phi_g(\vec{r},t) + S_{g,d}(\vec{r},t)
  \label{eq:3d-sn}
  \shortintertext{where}
  \begin{align*}
    \Psi_{g,d}(\vec{r},t) &= \Psi_g(\vec{r},\hat{\Omega}_d,t) \\
%    g &= \mbox{ neutron energy group index, ranging from 1 to G,} \nonumber \\
%    \Sigma_{t,g}(x) &= \mbox{ macroscopic total cross section of neutrons in group $g$,} \nonumber \\
%    G &= \mbox{ total number of energy groups,} \nonumber \\
%    n &= \mbox{ discrete ordinate index,} \nonumber \\
    N &= \mbox{ total number of discrete ordinates,} \\
    l &= \mbox{ Legendre expansion order,} \\
    L &= \mbox{ highest Legendre expansion order,} \\
    \Sigma^{g'\rightarrow g}_{s,l}(x) &= \mbox{ $l$-th Legendre moment of the macroscopic
scattering} \nonumber \\
    &\ \ \ \ \mbox{ cross section from group $g'$ to $g$,} \\
    \phi_{g',l,m} &= \mbox{ spherical harmonic expansion of the angular flux of degree $l$ and
    order $m$,} \\
    Y_{l,m} &= \mbox{ real basis functions of spherical harmonics of degree $l$ and order $m$.}
%    \chi_g &= \mbox{ fission neutron spectrum in group $g$,} \nonumber \\
%    \nu &= \mbox{ neutrons produced per fission reaction,} \nonumber \\
%    \Sigma_{f,g}(x) &= \mbox{ macroscopic fission cross section of neutrons in group $g$,}
%    \nonumber \\
%    k &= \mbox{ neutron multiplication factor.} \nonumber
%    \phi_{g'}(x) &= \mbox{ scalar neutron flux in group $g$,} \nonumber \\
%    S_g(x,\mu_n) &= \mbox{ external neutron source in group $g$.} \nonumber
  \end{align*}
\end{gather}
%
We can retrieve the neutron scalar flux and current by calculating the 0th and 1st Legendre
moments, respectively, as follows:
%
\begin{align}
  \phi_g(x) &= \frac{1}{2} \int^1_{-1} d\mu\ \Psi_g(x,\mu) = \frac{1}{2} \sum^N_{n=1} w_n
\Psi_g(x,\mu_n)
  \shortintertext{and}
  J_g(x) &= \frac{1}{2} \int^1_{-1} d\mu\ \mu\Psi_g(x,\mu) = \frac{1}{2} \sum^N_{n=1} w_n
\mu_n \Psi_g(x,\mu_n)
\end{align}

The 1-D form of the multigroup $k$-eigenvalue neutron diffusion equations (Eq.\ \ref{eq:mg-diff})
is:
%
\begin{align}
  -\frac{d}{dx} D_g(x) \frac{d}{dx} \phi_g(x) + \Sigma_{t,g}(x) \phi_g(x) &= \sum^G_{g'=1}\left[
  \Sigma_s^{g'\rightarrow g}(x)\phi_{g'}(x) + \chi_g\frac{\nu\Sigma_{f,g'}(x)}{k}
  \phi_{g'}(x)\right] + S_g(x)
  \label{eq:1d-diff}
  \shortintertext{where}
    D_g(x) &= \frac{1}{3 \Sigma_{tr}(x)} \nonumber \\
           &= \mbox{ isotropic neutron diffusion coefficient for group }g. \nonumber
\end{align}

\subsection{Spatially Varying Diffusion Coefficients} \label{sec:svdc}

The conventional approach for determining diffusion coefficients for each subregion involves
running a high-fidelity neutron transport simulation to tally region-wide estimates of the neutron
transport cross section. The transport cross section formulation is derived from the $P_1$
approximation of the neutron transport equation with isotropic sources \cite{bell_nuclear_1970}.
Essentially, each defined subregion has a constant diffusion coefficient value. However, as
discussed in Section \ref{sec:summary-nts-mtds}, the neutron diffusion
equation is only valid in regions of high scattering-to-removal ratios with, at most, linearly
anisotropic scattering and small flux gradients. These conditions do not hold near or within
control rods, near interfaces of neighboring materials with highly dissimilar neutronic properties,
and in materials with significant scattering contributions from light nuclei.

In the hybrid $S_N$-diffusion method, I propose replacing the conventional $P_1$-based
diffusion coefficients with an alternate formulation that incorporates pointwise corrections
to the neutron diffusion flux solution from the $S_N$-derived flux solution as follows:
%
\begin{align}
  D^s_g(x) &= -J^{tr}_g(x)\bigg/\frac{d\phi^{tr}_g(x)}{dx}. \label{eq:svdc}
\end{align}
%
where $D^s$ is the \glspl{SVDC}, and the $tr$ superscript denotes the transport-computed neutron
current and scalar flux solutions from the $S_N$ method.
The basic form of this equation is Fick's first law of diffusion. The result is a diffusion
coefficient variable that continuously varies in space from point to point except at dissimilar
material interfaces. \glspl{SVDC} provide pointwise corrections to closely match the diffusion flux
solution to the $S_N$ flux solution.

Eq.\ \ref{eq:svdc} is identical to Ronen's \cite{ronen_accurate_2004} and Pounders \& Rahnema's
\cite{pounders_diffusion_2009} formulations for space-dependent diffusion coefficients in Eq.\
\ref{eq:ronen} and Eq.\ \ref{eq:emp}, respectively. In comparison with
the Ronen method, our approaches differ in how $\phi_{tr}$ and $J_{tr}$ are obtained. Starting with
a standard neutron diffusion calculation, the Ronen method applies an analytically derived
transport operator on every iteration to calculate a new estimate of $J_{tr}$ using the $\phi$
solution from the previous iteration. The main difficulty for the Ronen method lies in deriving the
transport operators which has been achieved for only simple 1-D geometries thus far.

For the \textit{empirical method} by Pounders \& Rahnema, their diffusion coefficients are
piecewise-constant and they assumed a priori knowledge of the reference flux solution.
Pounders \& Rahnema \cite{pounders_diffusion_2009} demonstrated the effectiveness of applying
pointwise corrections derived from analytical or Monte Carlo reference flux solutions. Compared
with conventional $P_1$-based out-scatter and flux-limited approximations of the diffusion
coefficient, their diffusion coefficients showed superior agreement
with the reference flux solutions. They recognized that volume averaging within each mesh element
for piecewise-constant diffusion coefficients introduces some truncation error if the flux is
non-linear within the mesh element. This design choice may be due to an intention to retain the
diffusion coefficient as a constant in the $\frac{d}{dx}D\frac{d\phi}{dx}$ term of the neutron
diffusion equation (Eq.\ \ref{eq:1d-diff}).

Unlike their approach, the \gls{SVDC} formulation in Eq.\ \ref{eq:svdc} allows for continuously
varying diffusion coefficients to reduce truncation error. In practice, the discretization order of
\gls{SVDC} variables in a numerical calculation would be the same as the discretization order of
the reference flux solution. This formulation introduces a minor change to the finite difference
implementation of the 2nd-order diffusion term to handle spatial derivatives of the
diffusion coefficient. However, the finite element implementation sees no change as the 2nd-order
diffusion term is commonly treated with integration-by-parts to reduce it to a 1st-order
differential term with no derivatives of the diffusion coefficient. Refer to Section
\ref{sec:implementation} for key numerical implementation details of a finite difference neutron
diffusion solver with support for handling \glspl{SVDC}.

For the rest of Section \ref{sec:svdc}, I will show how the \glspl{SVDC} may differ from the
conventional $P_1$-based diffusion coefficient before presenting further details of the hybrid
$S_N$-diffusion method in Section \ref{sec:theory}.

\begin{figure}[htb!]
  \centering
  \includegraphics[width=.7\columnwidth]{case-0-geometry}
  \caption{Case 0: A 1-D, two-region system containing a 0.5-cm thick control rod and a 24.5-cm
    thick homogeneous mixture of molten fuel salt and graphite moderator. Reflective boundary
    conditions apply on both ends.}
  \label{fig:case-0-geom}
\end{figure}

\subsubsection{\gls{SVDC} Demonstration with Case 0}

To facilitate the following demonstration of a neutron diffusion calculation with \glspl{SVDC},
consider a 1-D, two-region system (which I will refer to as Case 0 moving forward) consisting of a
highly neutron-absorbing material and a
neutron-multiplying region with reflective boundary conditions on both ends (Figure
\ref{fig:case-0-geom}). I defined the material specifications of the control rod, molten fuel salt,
and graphite moderator based on the material compositions of the \gls{MSRE}
\cite{robertson_msre_1965}. The
homogenized molten fuel salt and graphite regions minimize discrepancies arising from geometrical
heterogeneity. The group constant input data for the diffusion and $S_N$ solvers (e.g., $P_1$-based
diffusion coefficients, cross sections, fission spectra) were sampled at 900 K and generated using
the OpenMC Monte Carlo neutronics software \cite{romano_openmc:_2015}. OpenMC uses the flux-limited
approximation method, developed by Pomraning \cite{pomraning_flux-limited_1984}, to calculate
diffusion coefficients. The condensed
neutron energy spectrum forms two energy groups bounded between $10^{-5}$, $10^0$, and $10^8$ eV.

I solved for the neutron flux in this system using the following set of numerical solvers:
%
\begin{enumerate}
  \item Neutron diffusion solver with $P_1$-based diffusion coefficients generated directly from
    the group constant generation step with OpenMC.
  \item $S_N$ neutron transport solver with $N=8$ and up to 2nd-order Legendre
    moments of the group-to-group neutron scattering cross sections.
  \item Neutron diffusion solver with \glspl{SVDC} generated from the prior $S_8$
    flux solution.
\end{enumerate}
%
I implemented the diffusion solvers using the finite difference method and the $S_N$
solver using the diamond-difference transport sweep method in Python. For
brevity, I defer further numerical implementation details of these solvers to Section
\ref{sec:implementation}.

\begin{figure}[htb!]
  \centering
  \begin{subfigure}[b]{.49\textwidth}
    \centering
    \includegraphics[width=\textwidth]{case-0-group-1-flux}
    \caption{Group 1}
    \label{fig:c0g1flux}
  \end{subfigure}
  \hfill
  \begin{subfigure}[b]{.49\textwidth}
    \centering
    \includegraphics[width=\textwidth]{case-0-group-2-flux}
    \caption{Group 2}
    \label{fig:c0g2flux}
  \end{subfigure}
  \caption{Neutron group 1 and 2 flux distributions from the diffusion, $S_8$, and
  diffusion-\gls{SVDC} solvers for Case 0.}
  \label{fig:c0flux}
\end{figure}
%
\begin{figure}[htb!]
  \centering
  \begin{subfigure}[b]{.49\textwidth}
    \centering
    \includegraphics[width=\textwidth]{case-0-group-1-diffcoef}
    \caption{Group 1}
    \label{fig:c0g1diffcoef}
  \end{subfigure}
  \hfill
  \begin{subfigure}[b]{.49\textwidth}
    \centering
    \includegraphics[width=\textwidth]{case-0-group-2-diffcoef}
    \caption{Group 2}
    \label{fig:c0g2diffcoef}
  \end{subfigure}
  \caption{$P_1$-based diffusion coefficient and \gls{SVDC} spatial distributions
  for Case 0.}
  \label{fig:c0diffcoef}
\end{figure}

Figure \ref{fig:c0flux} shows the group 1 and 2 neutron fluxes from the diffusion-$P_1$, $S_8$, and
diffusion-\gls{SVDC} solvers with a uniform mesh size of 0.005 cm. I omitted the OpenMC flux
solution because this exercise aims to demonstrate the effectiveness of \glspl{SVDC}
in reproducing an $S_N$-derived flux solution. As expected, the diffusion-$P_1$ flux solution
deviates from the $S_8$ flux solution, while the diffusion-\gls{SVDC} flux solution shows
significantly better agreement with the $S_8$ flux solution. As shown in Table \ref{table:c0k}, the
$k$ estimate from the diffusion-\gls{SVDC} solver is also closer to the $k$ estimate from the
$S_8$ solver than the diffusion-$P_1$ solver.

\begin{table}[tb!]
  \centering
  \caption{Multiplication factor $k$ estimates from the diffusion-$P_1$, $S_8$, and
  diffusion-\gls{SVDC} solvers and the absolute difference relative to the $S_8$ estimate.}
  \begin{tabular}{l S S}
    \toprule
    Solver type & {$k$} & {$k-k_{S_8}$} \\
    \midrule
    $S_8$ & 0.62736 & {-} \\
    Diffusion-$P_1$ & 0.60890 & -0.01846 \\
    Diffusion-\gls{SVDC} & 0.62750 & +0.00013 \\
    \bottomrule
  \end{tabular}
  \label{table:c0k}
\end{table}

The $P_1$-based diffusion coefficients and \glspl{SVDC} in Figure \ref{fig:c0diffcoef} show good
agreement in the bulk fuel salt region far from the control rod region.
This observation supports the validity of the neutron diffusion method under the conditions
listed at the start of Section \ref{sec:svdc}. Closer to the control rod
region, the \glspl{SVDC} deviate from the $P_1$-based diffusion coefficients because the
prerequisites for diffusion theory are violated.

Another key finding from this study is that the neutron diffusion method with $P_1$-based
diffusion coefficients fails to reproduce the steep $S_8$ flux gradient in the control rod region
for group 1 and in the homogeneous fuel-graphite region adjacent to the control rod region for both
neutron groups. From a purely mathematical perspective, the streaming term in the neutron
diffusion equation requires the lower
\gls{SVDC} values in the absorber-adjacent region to induce steeper flux gradients and match the
$S_8$ flux solution. Given that highly neutron-absorbing materials induce steep flux gradients
around them, we can expect similar trends of \glspl{SVDC} being suppressed relative to
$P_1$-based diffusion coefficients in other systems with control rods. The \gls{SVDC} values within
the highly-absorbing region exhibit significant spatial variation but do not necessarily fall
below the $P_1$-based diffusion coefficients.

Thus far, this method is redundant because it requires a priori knowledge of a reference
analytical flux solution or a highly accurate numerical solution calculated using neutron
transport methods. While existing workflows for diffusion-based methods already require
computationally intensive neutron transport simulations for group constant generation, this
preprocessing step incurs a fixed computational cost through
a series of neutron transport simulations. Compared to \glspl{SVDC}, $P_1$-based diffusion
coefficients behave much more like intrinsic material properties as their definitions are largely
tied to material cross sections. Variations in $P_1$-based diffusion coefficients primarily arise
from changes in the neutron energy spectrum with no direct contribution from proximity to material
interfaces (geometrical heterogeneity). $P_1$-based diffusion coefficients may be generated at
various reactor temperatures and interpolated to model the temperature dependence. Conversely, the
nature of \glspl{SVDC} as empirical, pointwise corrections for the diffusion equation
makes it highly dependent on the neutron flux gradient and susceptible to more significant
variations than region-wide estimates for $P_1$-based diffusion coefficients. \glspl{SVDC} will
likely need to be dynamically generated at every timestep, such as with a two-level iterative
scheme consisting of a $S_N$ neutron transport calculation and a neutron
diffusion calculation. In this case, the required number of neutron transport simulations scales
with the number of reactor simulations. Therefore, \gls{SVDC} generation cannot be adopted as a
one-off preprocessing step.

Another significant challenge of \glspl{SVDC} and the high-order empirical diffusion coefficients
involves their resolution near neutron flux peaks. The neutron current and flux gradient values in
Eq.\ \ref{eq:svdc} generally do not reach zero at the same points in space,
resulting in diffusion coefficient values tending to positive or negative infinity when the
flux gradient is close to zero. Pounders \& Rahnema avoided this issue by using
larger mesh sizes to calculate their empirical diffusion coefficients. However, their
remedy contradicts mesh convergence requirements and would worsen flux accuracy in regions with
steep, non-linear fluxes, such as near control rods. I will discuss my approach for investigating
and finding an alternative solution for this issue in Section \ref{sec:devel-hybrid}.

\subsection{Hybrid $S_N$-Diffusion Method} \label{sec:hybrid-method}

In order to reduce the computational cost of the high-level $S_N$ calculation, I propose reducing
the problem domain of the $S_N$ method to a \textit{correction region} containing the control rod
and its vicinity. Consequently, the hybrid $S_N$-diffusion method can retain accurate neutron flux
and current estimates around the control rod region from the $S_N$ method while making significant
computational cost savings by treating most of the reactor geometry with the neutron diffusion
method alone. Henceforth, I will refer to the $S_N$ calculation on the correction
region as the $S_N$ \textit{subproblem} or \textit{sub-solver}. I define the full problem
domain and the correction region as $V_0$ and $V_1$, respectively, where
$V_1\subseteq V_0$. The algorithm for the hybrid $S_N$-diffusion method is as follows:
%
\begin{enumerate}
  \item Start with an initial neutron diffusion calculation in $V_0$ with conventional
    $P_1$-based diffusion coefficients and other standard group constants (e.g., neutron cross
    sections).
  \item Use the neutron diffusion flux estimates in $V_1$ and current estimates along
    $\partial V_1$ as initial and boundary conditions for the $S_N$ sub-solver.
  \item With the $S_N$ sub-solver, calculate an improved neutron flux solution in $V_1$,
    which contains the control rod region and its immediate vicinity.
  \item Calculate \glspl{SVDC} using the $S_N$ flux solution and Eq.\ \ref{eq:svdc}.
  \item Pass the \glspl{SVDC} to the neutron diffusion solver to replace the conventional
    $P_1$ diffusion coefficients within the $V_1$ while continuing to use $P_1$-based
    diffusion coefficients in the rest of $V_0$.
  \item Start the next iteration by running neutron diffusion calculation with the \glspl{SVDC}
    replacing $P_1$ diffusion coefficients in part or all of $V_1$.
  \item Repeat Steps 2-6 until convergence is reached by meeting pre-defined convergence tolerance
    values.
\end{enumerate}
%
Figure \ref{fig:algorithm} shows a visual flowchart of the same hybrid method algorithm.

\begin{figure}[b!]
  \tikzstyle{every node}=[font=\small]
  \centering
  \begin{tikzpicture}
    \node (1) [object] {\textbf{Perform initial neutron\\diffusion calculation in $\bm{V_0}$}};
    \node (2) [process, below of=1, yshift=-1.8cm]
      {\textbf{Calculate incident flux boundary conditions along\\$\bm{\partial V_1}$ for $S_N$ calculation}};
    \node (3) [process, below of=2, yshift=-1.5cm]
      {\textbf{Perform $\bm{S_N}$ calculation\\in $\bm{V_1}$ for $\bm{\phi^{tr}_i}$ \& $\bm{J^{tr}_i}$}};
    \node (4) [process, below of=3, yshift=-1.5cm]
      {\textbf{Calculate $\bm{D^s_i}$ using Eq.\ \ref{eq:svdc}}};
    \node (5) [process, below of=4, yshift = -1.5cm]
      {\textbf{Perform neutron diffusion\\calculation in $\bm{V_0}$ using $\bm{D^s_i}$ in $\bm{V_1}$}};
    \node (6) [decision, below of=5, yshift = -1.8cm]
      {\textbf{Convergence\\check}};
    \node (7) [object, below of=6, yshift = -2cm]
      {\textbf{Final $\bm{\phi}$ and $\bm{k}$\\solutions obtained}};
    \draw [arrow] (1) -- node[anchor=east] {$\bm{\phi_0, J_0}$} (2);
    \draw [arrow] (2) -- node[anchor=east] {$\bm{\Psi^-}$ \textbf{along} $\bm{\partial V_1}$} (3);
    \draw [arrow] (3) -- node[anchor=east] {$\bm{\phi^{tr}_i, J^{tr}_i}$} (4);
    \draw [arrow] (4) -- node[anchor=east] {$\bm{D^s_i}$} (5);
    \draw [arrow] (5) -- node[anchor=east] {$\bm{\phi_i, J_i$}} (6);
    \draw [arrow] (6) -- node[anchor=east] {\textbf{Yes}} (7);
    \draw [arrow] (6) -- ([shift={(2cm,0cm)}]6.east)-- node[anchor=west] {\textbf{No}} ([shift={(0.5cm,0cm)}]2.east)--(2);
  \end{tikzpicture}
  \caption{Algorithm flowchart for the hybrid $S_N$-diffusion method.}
  \label{fig:algorithm}
\end{figure}

\subsubsection{$S_N$ Subsolver Boundary Conditions \& Buffer Zone}

The main challenge lies in determining appropriate boundary conditions for the $S_N$ subproblem.
Given that we want to limit the coverage of $V_1$ to the control rod region and its
immediate vicinity, $V_1$ should be sufficiently smaller than $V_0$, but large enough to capture
anisotropies in the flux due to the control rod. As a consequence, the boundaries $\partial V_1$ 
lie well within $V_0$. Crucially, there is currently no feasible method of generating
accurate boundary fluxes for an $S_N$ solver from a neutron diffusion flux solution. In 1-D, the
standard one-group $S_N$ method requires N/2
incoming boundary flux parameters per boundary mesh point for the N/2 neutron angular fluxes
flowing into $V_1$. The neutron diffusion method can produce at most one independent
incoming flux parameter per mesh point; this parameter is the neutron forward/backward current in
the $P_1$ approximation defined as:
%
\begin{align}
  J_{g,\pm} &= \frac{\phi_g}{4} \mp \frac{D_g}{2}\frac{d\phi_g}{dx} \label{eq:p1-j}
  \shortintertext{where}
  J_{g,\pm} &= \mbox{ neutron forward/backward current of group }g. \nonumber
\end{align}
%
Without additional information, the next best estimate is a uniformly
isotropic transmission of angular flux, i.e., all N/2 incoming angular fluxes $\Psi$ are equal in
magnitude. This is a good assumption far from the control rod, but will not produce accurate
fluxes if the boundary is too close to the control rod. This type of boundary condition is similar
to the white boundary condition, which describes the uniformly
isotropic reflection of particles at a boundary, except this case involves transmission rather than
reflection. Isotropic flux transmission at $x$ can be expressed mathematically for forward angular
fluxes as:
%
\begin{align}
  \sum^N_{n=N/2+1}w_n\mu_n\Psi(x,\mu_n) =& J_{+}(x) && (\mu_n>0) \nonumber \\
  \Psi(x,\mu_n)\sum^N_{n=N/2+1}w_n\mu_n =& J_{+}(x) && (\because \mbox{isotropic transmission})
  \nonumber \\
  \Psi(x,\mu_n) =& J_{+}(x)\Bigg/\sum^N_{n=N/2+1}w_n\mu_n
\end{align}
%
and backward angular fluxes as:
%
\begin{align}
  \sum^{N/2}_{n=1}w_n\mu_n\Psi(x,\mu_n) =& J_{-}(x) && (\mu_n<0) \nonumber \\
  \Psi(x,\mu_n)\sum^{N/2}_{n=1}w_n\mu_n =& J_{-}(x) && (\because \mbox{isotropic transmission})
  \nonumber \\
  \Psi(x,\mu_n) =& J_{-}(x)\Bigg/\sum^{N/2}_{n=1}w_n\mu_n \label{eq:sn-psi-j}
\end{align}
%
These boundary conditions for the $S_N$ sub-solver will generally yield some deviations in the
$\phi$ distributions since neutron fluxes in realistic reactor systems are at least slightly
anisotropic in most of the domain. However, in my preliminary investigations, the influence of
boundary conditions on the ratio $J$ and $\frac{d\phi}{dx}$ does not extend far from the boundaries
in optically thick media due to the strong scattering (e.g., graphite or molten salt) or
absorption effects (e.g., control rod). In the bulk regions far from the boundaries, the
scattering and absorption effects have the most significant influence on the flux. I observed that
the \glspl{SVDC} generated with the $S_N$ sub-solver within $V_1$ were accurate everywhere except
near $\partial V_1$. Therefore, $V_1$ must be large enough to provide transport corrections
through the \glspl{SVDC} and accommodate inaccurate \glspl{SVDC} near
$\partial V_1$. The inaccurate \glspl{SVDC} should then be discarded in favor of using
$P_1$-based diffusion coefficients as in the rest of $V_0$ not covered by $V_0$. The
subsequent neutron diffusion calculation with this mix of \glspl{SVDC} and $P_1$-based diffusion
coefficients should provide more accurate $\phi$ and $k$ estimates than a conventional
neutron diffusion calculation.

%To resolve this issue, I posit the following hypothesis: \textit{If there exists a highly
%neutron-absorbing control rod region within the $S_N$ subproblem domain $V_1$ and the $S_N$
%subproblem boundary $\partial V_1$ lies several neutron mean free paths away from this
%region, the \gls{SVDC} values calculated near the control rod region (using the $S_N$ sub-solver
%with isotropic hemisphere boundary conditions) will tend to the actual flux gradient solution
%(using a reference $S_N$ calculation across the entire problem domain $V$).} In other words,
%suboptimal boundary conditions may induce inaccurate \gls{SVDC} values near the $\partial
%V_1$ subdomain boundaries, but the \gls{SVDC} values further from $\partial V_1$ are
%accurate and very weakly dependent on the boundary conditions. These inaccurate \gls{SVDC} values
%close to $V_1$ may be discarded in favor of the default $P_1$-based diffusion coefficients.
%I will refer to the region containing these discarded values as the \textit{dead zone} or
%$V_2$, where $V_2 \subset V_1 \subseteq V_0$.

\subsubsection{Hybrid Method Verification with Case 0}

In this subsection, I revisit Case 0
(Figure \ref{fig:case-0-geom}) to test my hypothesis and demonstrate the hybrid $S_N$-diffusion
method. Once more, I defer general numerical implementation details to
Section \ref{sec:implementation}.

In Section \ref{sec:svdc}, the \gls{SVDC} generated from the reference $S_8$ calculation exhibited
significant spatial variations attributed to the presence of the control rod, as shown in Fig
\ref{fig:c0diffcoef}. The group 1 and 2 \gls{SVDC} values approach asymptotic values near the
corresponding group 1 and 2 $P_1$-based diffusion coefficient values near $x=15$ cm and $x=5$ cm,
respectively. Therefore, the $V_1$ should, at minimum, cover the domain between $x=0$ cm and
$x=15$ cm. Allowing for a buffer region for inaccurate \glspl{SVDC} near $\partial V_1$, I
defined $V_1$ to span from $x=0$ cm to $x=17.5$ cm.

I set up the hybrid method to automatically determine the cutoff point, beyond which \glspl{SVDC}
in $V_1$ are discarded, based on the \gls{SVDC} distribution generated in every iteration.
Starting from the interface between the control rod and fuel-graphite region, the algorithm sweeps
rightward for the point at which the group-wise \gls{SVDC} distribution approaches within 1\% of
the $P_1$-based diffusion coefficient value. This point serves as the cutoff point. The hybrid
method converges rapidly within two outer iterations comprising three neutron diffusion and two
$S_8$ calculations in total as outlined by the hybrid method algorithm; the change in $k$ between
the second and third outer iterations is approximately $10^{-7}$.
%
\begin{figure}[htb!]
  \centering
  \begin{subfigure}[b]{.49\textwidth}
    \centering
    \includegraphics[width=\textwidth]{case-0-group-1-hybrid-diffcoef}
    \caption{Group 1}
    \label{fig:c0g1hd}
  \end{subfigure}
  \hfill
  \begin{subfigure}[b]{.49\textwidth}
    \centering
    \includegraphics[width=\textwidth]{case-0-group-2-hybrid-diffcoef}
    \caption{Group 2}
    \label{fig:c0g2hd}
  \end{subfigure}
  \caption{$P_1$-based flux-limited diffusion coefficient and \gls{SVDC} spatial distribution for
  Case 0. The \gls{SVDC} distributions were generated from the reference $S_8$ (``SVDC'') and the
  hybrid (``Hybrid'') calculations.}
  \label{fig:c0hd}
\end{figure}

Refer to Figures \ref{fig:c0g1hd} and \ref{fig:c0g2hd} for a qualitative comparison of group 1 and
2 $P_1$-based diffusion coefficients, \glspl{SVDC} derived from the reference $S_8$ calculation as
demonstrated in Section \ref{sec:svdc}, and \glspl{SVDC} derived from the hybrid $S_N$-diffusion
method discussed here. The \glspl{SVDC} from the hybrid method agree well with the reference
\glspl{SVDC} in the control rod region and for most of the fuel-graphite region up to around
$x=15$ cm. Both sets of \glspl{SVDC} approximately coincide with the $P_1$ diffusion coefficient
values within 1\% difference from $x=14.615$ cm and $x=3.390$ cm onwards for group 1 and
group 2, respectively. Accordingly, the buffer region spans from $x=14.615$ cm to $x=17.500$ cm for
group 1 and $x=3.390$ cm to $x=17.500$ cm for group 2, in which the hybrid method defaults to the
$P_1$-based diffusion coefficients. Given that the reference \gls{SVDC} values will generally not
be known in real-world problems, the hybrid method relies on finding the intersection point of the
\gls{SVDC} and $P_1$ diffusion coefficient values to determine the buffer region cutoff point. The
size of $V_1$ will also have to be based on where this intersection point lies.
Another important consideration is the fact that the neutron flux gradient will be discontinuous if
the \gls{SVDC} and $P_1$ diffusion coefficient values are too far apart in magnitude at the buffer
region cutoff point. A flux gradient discontinuity in a homogeneous bulk region is non-physical and
unacceptable. In Section \ref{sec:prelim-results}, I study \gls{SVDC} distribution trends
in more heterogeneous systems and their implications on determining the size of $V_1$ and
the cutoff point.
%
\begin{figure}[htb!]
  \centering
  \begin{subfigure}[b]{.49\textwidth}
    \centering
    \includegraphics[width=\textwidth]{case-0-group-1-hybrid-flux}
    \caption{Group 1}
    \label{fig:c0g1hf}
  \end{subfigure}
  \hfill
  \begin{subfigure}[b]{.49\textwidth}
    \centering
    \includegraphics[width=\textwidth]{case-0-group-2-hybrid-flux}
    \caption{Group 2}
    \label{fig:c0g2hf}
  \end{subfigure}
  \caption{Neutron group 1 and 2 flux distributions from the diffusion, $S_8$, reference
  \gls{SVDC}, and hybrid methods for Case 0.}
  \label{fig:c0hf}
  \centering
  \begin{subfigure}[b]{.49\textwidth}
    \centering
    \includegraphics[width=\textwidth]{case-0-group-1-hybrid-flux-diff}
    \caption{Group 1}
    \label{fig:c0g1hfdiff}
  \end{subfigure}
  \hfill
  \begin{subfigure}[b]{.49\textwidth}
    \centering
    \includegraphics[width=\textwidth]{case-0-group-2-hybrid-flux-diff}
    \caption{Group 2}
    \label{fig:c0g2hfdiff}
  \end{subfigure}
  \caption{Difference in neutron group 1 and 2 flux distributions from the diffusion,
  diffusion-\gls{SVDC}, and hybrid methods with respect to the $S_8$ flux distribution for Case 0.}
  \label{fig:c0hfdiff}
\end{figure}

Figures \ref{fig:c0g1hf} and \ref{fig:c0g2hf} show the group 1 and 2 neutron flux distributions
from the various methods, while Figures \ref{fig:c0g1hfdiff} and \ref{fig:c0g2hfdiff} show the
difference in flux with respect to the $S_8$ flux distribution. As
expected following the diffusion coefficient discussion, the hybrid method flux distribution
matches the $S_8$ and Diffusion-\gls{SVDC} flux distributions well. The $k$ estimate from the
hybrid method is 0.62774, which is only 0.00024 higher than the Diffusion-\gls{SVDC} method, and
0.00038 higher than the $S_8$ method.

In Case 0, $V_1$ covers more than half of $V_0$ due to the control rod region being
the only significant source of influence on the neutron flux distribution in the
otherwise homogeneous system with reflective boundary conditions. I chose Case 0 as a simple test
case to aid in introducing the hybrid $S_N$-diffusion method. In Section \ref{sec:prelim-results},
I present more results with the hybrid method on more complicated geometries containing neutron
reflector, air, and heterogeneous fuel-graphite lattice regions and vacuum boundary conditions. In
these other test cases, the hybrid method yields smaller $V_1$-to-$V_0$ ratios,
reducing computational costs from running $S_N$ calculations over a relatively smaller
region.

\section{Numerical Implementation} \label{sec:implementation}

In Section \ref{sec:theory}, I presented the theoretical basis for the hybrid
$S_N$-diffusion method and demonstrated the method on a simple problem. Now I will discuss
numerical implementation details of group
constant data processing and the neutron diffusion, $S_N$ neutron transport, and hybrid
$S_N$-diffusion solvers. I implemented all numerical solvers and analysis scripts in the Python
programming language. First, I will discuss the material group constant data generation and
postprocessing steps. Next, I present the implementation details of the neutron
diffusion and $S_N$ numerical methods. Lastly, I present how they are coupled to form the hybrid
$S_N$-diffusion method.

\subsection{Group Constant Data Generation}

The group constants required by either or both neutron diffusion and $S_N$ neutron transport
methods are:
%
\begin{itemize}
  \item $\Sigma_{t,g}$: Macroscopic total cross section in group $g$,
  \item $\Sigma_{r,g}$: Macroscopic removal cross section in group $g$,
  \item $\Sigma_s^{g'\rightarrow g}$: Macroscopic group-to-group scattering cross section matrix,
  \item $\Sigma_{s,l}^{g'\rightarrow g}$: $l$-th Legendre moment of the macroscopic
    group-to-group scattering cross section matrix,
  \item $\Sigma_{sp,l}^{g'\rightarrow g}$: $l$-th Legendre moment of the macroscopic
    group-to-group scattering production cross section matrix,
  \item $D_g$: $P_1$-based diffusion coefficient in group $g$,
  \item $\nu\Sigma_{f,g}$: Product of the average number of neutrons produced per fission and the
    macroscopic fission cross section in group $g$,
  \item $\chi_g$: Neutron fission spectrum in group $g$.
\end{itemize}
%
These group constants are generated using OpenMC's multigroup cross section generation capability
\cite{boyd_multigroup_2019} and postprocessed using a Python script into JSON format files.
$\Sigma_{r,g}$ is the only quantity that OpenMC does not directly provide. It is calculated as
%
\begin{align}
  \Sigma_{r,g} =& \sum^G_{g'\neq g}\Sigma_s^{g\rightarrow g'}+\Sigma_{a,g}-\left(\Sigma_{sp}^{g
    \rightarrow g} - \Sigma_s^{g\rightarrow g}\right)
  \shortintertext{where}
      \Sigma_{a,g} =& \mbox{ macroscopic absorption cross section in group $g$,} \nonumber \\
      \Sigma_{sp}^{g\rightarrow g} =& \mbox{ macroscopic scattering production cross section from
      group $g$ to $g$.} \nonumber
\end{align}
%
$\Sigma_{r,g}$ primarily represents the loss of neutrons from group $g$ through outscattering and
absorption. $\Sigma_{sp}^{g\rightarrow g}$ incorporates neutron multiplication effects from neutron
knockout reactions into the scattering cross section. Neutron knockout reactions are commonly
tallied as scattering reactions, and $\Sigma_{r,g}$ is a convenient term to
incorporate neutron knockout effects into the neutron diffusion equations. I ran all test cases for
$S_N$ and hybrid method calculations with up to the 2nd Legendre moments of the scattering
cross sections ($L=2$).

OpenMC uses the $P_1$ flux-limited formulation \cite{pomraning_flux-limited_1984} for calculating
$D_g$ as follows
%
\begin{align}
  D_g =& \frac{1}{3\Sigma_{tr,g}},
  \shortintertext{where}
  \Sigma_{tr,g} =& \frac{\langle\Sigma_{t,g}\phi_g\rangle-\langle\Sigma_{s1,g}\phi_g\rangle}
  {\langle\phi_g\rangle}, \nonumber \\
  \langle\Sigma_{t,g}\phi_g\rangle =& \int_{r\in V}dr \int_{4\pi}d\Omega\int^{E_{g-1}}_{E_g}dE\
  \Sigma_{t,g}(r,E)\Psi(r,E,\Omega), \nonumber \\
  \langle\Sigma_{s1,g}\phi_g\rangle =& \int_{r\in V}dr \int_{4\pi}d\Omega\int^{E_{g-1}}_{E_g}dE
  \int_{4\pi}d\Omega'\int^{\infty}_0dE'\int^1_{-1}d\mu\ \mu\Sigma_s(r,E'\rightarrow E,\Omega'\cdot
  \Omega)\phi(r,E',\Omega'), \nonumber \\
  \langle \phi \rangle =& \int_{r\in V}dr\int_{4\pi}d\Omega\int^{E_{g-1}}_{E_g}dE\ \Psi(r,E,\Omega)
  .\nonumber
\end{align}

\subsection{Neutron Diffusion Method}

On a 1-D uniform spatial grid with $I+1$ mesh points, the neutron scalar flux variables
$\phi_{g,i}$ are defined on the mesh points $x_i$. Theoretically, group constants are volumetric
material properties that should be defined on the cell-centered half-integer mesh points
$x_{i+\sfrac{1}{2}}$. In practice, all material properties except diffusion coefficients are
uniform in each subregion and sampled at $x_i$. To avoid ambiguity concerning diffusion
coefficient sampling, I formulated all test cases such that all material interfaces fall on $x_i$.

Discretizing the multigroup $k$-eigenvalue neutron diffusion equations in Eq.\ \ref{eq:1d-diff}
and reformulating the scattering term using neutron balance in the control volume bounded by
$x_{i-\sfrac{1}{2}}$ and $x_{i+\sfrac{1}{2}}$ yields
%
\begin{align}
  J_{g,i+\sfrac{1}{2}} - J_{g,i-\sfrac{1}{2}} + \Sigma_{t,g,i} \phi_{g,i} \Delta x = \sum^G_{g'=1}\left[
  \Sigma_{s,i}^{g'\rightarrow g}\phi_{g',i} + \chi_{g,i}\frac{\nu\Sigma_{f,g',i}}{k} \phi_{g',i}
\right]\Delta x. \label{eq:diff-j}
\end{align}
%
Using the diamond difference scheme to replace the $J$ terms with the discretized form of
Fick's first law of diffusion,
%
\begin{align}
  J_{g,i+\sfrac{1}{2}} = -D_{g,i+\sfrac{1}{2}}\frac{d\phi_{g,i+\sfrac{1}{2}}}{dx} =
  -D_{g,i+\sfrac{1}{2}} \frac{\phi_{g,i+1}-\phi_{g,i}}{\Delta x},
\end{align}
%
and rearranging the terms in Eq.\ \ref{eq:diff-j} yields
%
\begin{align}
  -\frac{D_{g,i-\sfrac{1}{2}}}{\Delta x} \phi_{g,i-1} + &\left[\frac{D_{g,i-\sfrac{1}{2}}+
  D_{g,i+\sfrac{1}{2}}}{\Delta x} + \Delta x\ \Sigma_{r,g,i} \right]\phi_{g,i} -
  \frac{D_{g,i+\sfrac{1}{2}}}{\Delta x}\phi_{g,i+1} -\Delta x\sum^G_{g'\neq g}
  \Sigma_{s,i}^{g'\rightarrow g}\phi_{g',i} \nonumber \\
  =& \Delta x\sum^G_{g'=1}
  \chi_{g,i} \frac{\nu\Sigma_{f,g',i}}{k} \phi_{g',i}, \label{eq:diff-fd}
  \shortintertext{where}
  \Sigma_{r,g} =& \Sigma_{t,g} - \Sigma_s^{g\rightarrow g} \nonumber \\
  =& \mbox{ macroscopic removal cross section for neutron group }g. \nonumber
\end{align}
%
The diamond difference scheme is 2nd-order accurate, and this form is
equivalent to applying 2nd-order finite differencing to the original neutron diffusion equation in
Eq.\ \ref{eq:1d-diff} with diamond differencing for the cell-centered group constants.
The fixed neutron source $S_g$ is ignored here since the test cases are all neutron-multiplying
systems with no fixed source.

I implemented two types of boundary conditions: vacuum and reflective boundary conditions. The
\textbf{vacuum boundary conditions} are imposed by setting the incoming flux in the $P_1$
approximation to zero and applying 2nd-order finite differencing as follows:
%
\begin{align}
  \mbox{Left boundary: } \frac{\phi_{g,0}}{4}-\frac{D_{g,\sfrac{1}{2}}}{2}
  \frac{\left(-\phi_{g,2}+4\phi_{g,1}-3\phi_{g,0}\right)}{2\Delta x} =& 0 \\
  \mbox{Right boundary: } \frac{\phi_{g,I}}{4}+\frac{D_{g,I-\sfrac{1}{2}}}{2}
  \frac{\left(\phi_{g,I-2}-4\phi_{g,I-1}+3\phi_{g,I}\right)}{2\Delta x} =& 0.
\end{align}
%
The \textbf{reflective boundary conditions} are imposed by setting the flux gradient to zero and
applying 2nd-order finite differencing as follows:
%
\begin{align}
  \mbox{Left boundary: } \frac{-3\phi_{g,0}+4\phi_{g,1}-\phi_{g,2}}{2\Delta x} =& 0 \\
  \mbox{Right boundary: } \frac{3\phi_{g,I}-4\phi_{g,I-1}+\phi_{g,I-2}}{2\Delta x} =& 0
\end{align}
%
At material interfaces, the continuity condition requires that the net neutron current on either
side of the interface be equal as follows:
%
\begin{align}
  -\frac{3D_{g,i-\sfrac{1}{2}} - D_{g,i-\sfrac{3}{2}} }{2}
  \frac{\phi_{g,i-2}-4\phi_{g,i-1}+3\phi_{g,i}}{2\Delta x} =&
  -\frac{3D_{g,i+\sfrac{1}{2}} - D_{g,i+\sfrac{3}{2}} }{2}
  \frac{-3\phi_{g,i}+4\phi_{g,i+1}-\phi_{g,i}}{2\Delta x} \label{eq:itf-bc}
\end{align}
%
for a material interface at $x_i$.

Altogether, they form a system of equations of the form $\bm{A\overline{\phi}}=\bm{\frac{1}{k}
B\overline{\phi}}$, where $\bm{\overline{\phi}}$ is a flattened vector representation of
$\phi_{g,i}$, and $\bm{A}$ and $\bm{B}$ are matrices of the coefficients of $\phi_{g,i}$ as given
by Eqs. \ref{eq:diff-fd} to \ref{eq:itf-bc}. I implemented the inverse power method to find $k$ and
$\overline{\phi}$. The inverse power method algorithm is as follows
%
\begin{align}
  \shortintertext{1. Initialize $k^0$ and $\bm{\overline{\phi}}^0$}
  \shortintertext{2. Update $\bm{\overline{\phi}}$ and $k$}
  \bm{\overline{\phi}}^m =& \frac{1}{k^{m-1}}\bm{A}^{-1}\bm{B\overline{\phi}}^{m-1} \\
  k^m =& k^{m-1}\frac{|\bm{B\overline{\phi}}^m|}{|\bm{B\overline{\phi}}^{m-1}|}
  \shortintertext{3. Check whether convergence is reached}
  \epsilon_\phi =
  \frac{|\bm{\overline{\phi}}^m-\bm{\overline{\phi}}^{m-1}|}{|\bm{\overline{\phi}}^m|} <& \
  tol_{\bm{\overline{\phi}}} \\
  \epsilon_k =
  \frac{|k^m-k^{m-1}|}{|k^m|} <& \ tol_k
  \shortintertext{4. Return to Step 2 if either expression is false, otherwise exit.} \nonumber
\end{align}
%
$k^m$ and $\overline{\phi}^m$ denote estimates of $k$ and $\overline{\phi}$ after the $m$-th
iteration. Matrix $\bm{A}$ is a primarily tridiagonal matrix with at most $G-1$ off-diagonal terms
from the fourth term in Eq.\ \ref{eq:diff-fd}. Thus, $\bm{A}$ is initialized as a sparse matrix to
take advantage of the computationally efficient sparse matrix solver functions from the
\texttt{sparse} class of the \texttt{SciPy} \cite{virtanen_scipy_2020} Python library for
scientific and technical computing. Matrix $\bm{B}$
is never initialized explicitly as a matrix. Instead, the vector $\bm{b}=\bm{B\overline{\phi}}$ is
updated directly in every iteration. In this system of equations, $|\bm{b}|$ corresponds to the
total number of fission neutrons produced in the system for a given $\bm{\overline{\phi}}$. This
quantity is calculated using the \texttt{trapezoid} numerical integration function from
\texttt{SciPy} to estimate the integral value of the source term, $\nu\Sigma_{g,f}\phi_{g}$, in $x$
from the discrete flux values in $\bm{x}$. Finally, the final $\phi_{g,i}$ is normalized by a
factor of $|\bm{b}|/k$ (no. of source neutrons) to obtain the neutron scalar flux per source
neutron.

\subsection{$S_N$ Neutron Transport Method}

For the 1-D $S_N$ neutron transport method on the same uniform spatial grid with $I+1$ mesh points,
the neutron angular flux variables $\Psi_{g,i,n}=\Psi_g(x_i,\mu_n)$ are defined on the mesh points
$x_i$ while neutron scalar flux variables $\phi_{g,i\pm\sfrac{1}{2}}=\phi_g(x_{i\pm\sfrac{1}{2}})$
are defined on the half-integer mesh points $x_{i\pm\sfrac{1}{2}}$. All group constants are sampled
at $x_{i\pm\sfrac{1}{2}}$.

The $S_N$ equations are solved using the transport sweep method in which the algorithm ``sweeps''
through the spatial grid and sequentially updates $\Psi_{g,i,n}$. The algorithm sweep
direction follows the direction of neutron travel, i.e. it sweeps in the positive direction for
$\Psi_{g,i,n}$ with $\mu_n>0$ and vice versa. Doing otherwise is unphysical and causes numerical
instabilities.

Discretizing the multigroup $S_N$ neutron transport equations in Eq.\ \ref{eq:1d-sn} about
$x_{i+\sfrac{1}{2}}$ yields
%
\begin{align}
  \mu_n\frac{\Psi_{g,i+1,n}-\Psi_{g,i,n}}{x_{i+1}-x_i} + \Sigma_{t,g,i+\sfrac{1}{2}}
  \Psi_{g,i+\sfrac{1}{2},n} = q_{g,i+\sfrac{1}{2},n}
\end{align}
%
where $q_{g,i+\sfrac{1}{2},n}$ represents the combined scattering and fission neutron source term.
After expressing $\Psi_{g,i+\sfrac{1}{2}}$ as the average of $\Psi_{g,i+1,n}$ and $\Psi_{g,i,n}$
in the diamond difference scheme and rearranging the terms, we obtain
%
\begin{align}
  \Psi_{g,i+1,n} =& \frac{1-\Sigma_{t,g}\Delta x/2\mu_n}{1+
    \Sigma_{t,g}\Delta x/2\mu_n}\Psi_{g,i,n} +
    q\frac{\Delta x}{\mu_n\left(1+\Sigma_{t,g}\Delta x/2\mu_n\right)} \label{eq:sweep-right} &&
    (\mbox{for } \mu_n > 0)
\shortintertext{and}
  \Psi_{g,i,n} =& \frac{1+\Sigma_{t,g}\Delta x/2\mu_n}{1-
    \Sigma_{t,g}\Delta x/2\mu_n}\Psi_{g,i+1,n} -
    q\frac{\Delta x}{\mu_n\left(1-\Sigma_{t,g}\Delta x/2\mu_n\right)}. \label{eq:sweep-left} &&
    (\mbox{for } \mu_n < 0)
\end{align}
%
The remaining $i+\sfrac{1}{2}$ indexes on the group constants are dropped to reduce visual clutter.
These expressions are used to update $\Psi_{g,i,n}$ in the forward and backward transport sweeps.

The discretized scattering and fission terms in $q_{g,i+\sfrac{1}{2},n}$ are given as
%
\begin{align}
  q_{g,i+\sfrac{1}{2},n} =& \sum^G_{g'=1}\sum^L_{l=0}\frac{\left(2l+1\right)}
  {2}\Sigma_{s,l}^{g'\rightarrow g}P_l(\mu_n)\phi_{l,g',i+\sfrac{1}{2}} \nonumber \\
  &+\frac{\chi_g}{2}\sum^G_{g'=1}\frac{\nu\Sigma_{f,g'}}{k}\phi_{0,g',i+\sfrac{1}{2}}
  \label{eq:sn-q}
  \shortintertext{where}
  \phi_{l,g',i+\sfrac{1}{2}} =& \sum^N_{n'=1}w_{n'}P_l(\mu_{n'})\frac{\Psi_{g',i,n'}+
  \Psi_{g',i+1,n'}}{2}. \label{eq:phi-l}
\end{align}
%
$\phi_{l,g,i}$ are $l$-th Legendre expansions of the neutron flux evaluated using Gauss-Legendre
quadrature over $\mu_{n'}=[-1,1]$. $\phi_{0,g,i}$ and $\phi_{1,g,i}$ also correspond to the neutron
scalar flux $\phi_{g,i}$ and net current $J_{g,i}$, respectively.

For vacuum boundary conditions, $\Psi_{g,0,n}$ is zero for all positive $\mu_n$ while
$\Psi_{g,I,n}$ is zero for all negative $\mu_n$ as follows
%
\begin{align}
  \mbox{Left boundary: } \Psi_{g,0,n} =& 0 && (\mbox{for } \mu_n > 0) \\
  \mbox{Right boundary: } \Psi_{g,I,n} =& 0 && (\mbox{for } \mu_n < 0)
\end{align}
%
Reflective boundary conditions are imposed by equating the incoming angular flux to the
outgoing angular flux in the opposite direction as follows
%
\begin{align}
  \mbox{Left boundary: } \Psi_{g,0,n} =& \Psi_{g,0,n'} && (\mbox{for } \mu_n > 0, \mu_n =
  -\mu_{n'}) \\
  \mbox{Right boundary: } \Psi_{g,I,n} =& \Psi_{g,I,n'} && (\mbox{for } \mu_n < 0, \mu_n =
  -\mu_{n'})
\end{align}

The power iteration algorithm for the $S_N$ method is similar to the inverse power method algorithm
applied in the neutron diffusion method. The transport sweep step replaces the matrix-solving step
for updating $\overline{\phi}$.
%
\begin{align}
  \shortintertext{1. Initialize $k^0$, $\phi_{l,g,i+\sfrac{1}{2}}^0$, and $q^0$}
  \shortintertext{2. Apply transport sweeps to solve for $\Psi^m$ using Eqs. \ref{eq:sweep-right},
  \ref{eq:sweep-left}, and \ref{eq:sn-q}}
  \shortintertext{3. Update $\phi^m$ and $k^m$ using Eqs. \ref{eq:phi-l} and \ref{eq:sn-k}}
  k^m =& k^{m-1}\frac{\sum^I_{i=0}\sum^G_{g=1}\nu\Sigma_{f,g}\phi^m_{g,i+\sfrac{1}{2}}}
  {\sum^I_{i=0}\sum^G_{g=1}\nu\Sigma_{f,g}\phi^{m-1}_{g,i+\sfrac{1}{2}}} \label{eq:sn-k}
  \shortintertext{4. Check whether convergence is reached}
  \epsilon_\phi =
  \frac{|\bm{\overline{\phi}}^m-\bm{\overline{\phi}}^{m-1}|}{|\bm{\overline{\phi}}^m|} <& \
  tol_{\bm{\overline{\phi}}} \\
  \epsilon_k =
  \frac{|k^m-k^{m-1}|}{|k^m|} <& \ tol_k
  \shortintertext{5. Return to Step 2 if either expression is false, otherwise exit.} \nonumber
\end{align}

The transport sweep and $\phi^m$-update algorithms are parallelized using the \texttt{joblib}
parallel computing Python library \cite{noauthor_joblib_nodate} across all available CPU threads.
The total computational work is subdivided into several smaller tasks classified by unique pairs
of $g$ and $n$ values; each thread computes all $\Psi^m$ or $\phi^m$ values on the mesh for a given
pair of indexes $g$ and $n$.

\subsection{Hybrid $S_N$-Diffusion Method}

Without loss of generality, consider a 1-D system symmetric about the left boundary at $x_0=0$ cm.
In this case, as with Case 0, the control rod region is the left-most region, followed by other
regions. The $S_N$ subproblem domain $V_1$, containing the control rod and some region beyond it,
is bounded by $x_0$ and $x_j$ for some $x_j$
located several mean free paths to the right of the control rod region as governed by the relevant
discussion in Section \ref{sec:hybrid-method}.

Excluding the initial neutron diffusion calculation to initialize $k^0$ and $\phi^0$, each outer
iteration in the hybrid method consists of one $S_N$ neutron transport calculation and one neutron
diffusion calculation. The $\phi$ estimates from these $S_N$ transport and neutron diffusion
calculations are labeled as $\phi^{m+\sfrac{1}{2}}$ and $\phi^{m+1}$ during the
$(m+1)$-th outer iteration. $\phi^{m+\sfrac{1}{2}}$ spans $V_1$ while $\phi^{m+1}$ spans
the entire domain $V_0$.

For the $S_N$ subproblem boundary conditions, we discretize Eqs. \ref{eq:p1-j} and
\ref{eq:sn-psi-j} as follows
%
\begin{align}
  \Psi_{g,j,n} =& \frac{J_{g,-}(x_j)}{\sum^{N/2}_{n'=1}w_{n'}\mu_{n'}} \nonumber \\
  =& \frac{\frac{\phi_g(x_j)}{4}+
  \frac{D_g(x_{j-\sfrac{1}{2}})}{2}\frac{d\phi_g(x_j)}{dx}}{\sum^{N/2}_{n'=1}w_{n'}\mu_{n'}}
  \nonumber \\
  =& \frac{\frac{\phi_{g,j}}{4}+
  \frac{D_{g,j-\sfrac{1}{2}}}{2}\frac{\phi_{g,j}-\phi_{g,j-1}}{\Delta x}}
    {\sum^{N/2}_{n'=1}w_{n'}\mu_{n'}} && (\mu_n < 0) \label{eq:sn-bc}
\end{align}
%
where $D_g$ is the $P_1$-based diffusion coefficient value.

The \gls{SVDC} formulation in Eq.\ \ref{eq:svdc} is discretized using the diamond difference scheme
as follows
%
\begin{align}
  D^s_{g,i+\sfrac{1}{2}} =& -\frac{J^{tr}_{g,i+1}+J^{tr}_{g,i}}{2} \left(
  \frac{\phi^{tr}_{g,i+1}-\phi^{tr}_{g,i}}{\Delta x} \right)^{-1} = -\frac{\Delta x}{2}
  \frac{J^{tr}_{g,i+1}+J^{tr}_{g,i}}{\phi^{tr}_{g,i+1}-\phi^{tr}_{g,i}}. \label{eq:svdc-num}
\end{align}
%
%From Eq.\ \ref{eq:svdc-num}, we note that numerical instabilities may occur near flux peaks where
%the flux gradient and thus the denominator of Eq.\ \ref{eq:svdc-num} approach zero. On the other
%hand, this observation also implies that diffusion coefficient values have negligible influence on
%the flux solution near flux peaks due to the small flux gradient values. Therefore, we can avoid
%numerical instabilities by applying the following logic after calculating $D^s_g$: If the ratio of
%$|\frac{d\phi_g}{dx}|$ to $|\phi_g|$ is sufficiently small and $|D^s_g-D_g| > D_g$, the hybrid
%method defaults to using $D_g$ instead of $|D^s_g-D_g|$. The first conditional statement locates
%regions of relatively flat flux, and of these regions, the second conditional statement excludes
%regions which naturally experience flat flux such as air-filled regions. In all test cases,
%applying the first conditional as $|\frac{1}{\phi_g}\frac{d\phi_g}{dx}|<5\times 10^{-2}$ was
%sufficient for this task.

The hybrid $S_N$-diffusion algorithm is as follows
%
\begin{align}
  \shortintertext{1. Initialize $k^0$ and $\phi^0$ with an initial neutron diffusion calculation on
  $V_0$}
  \shortintertext{2. Generate estimates for $\Psi_{g,j,n}$ at $x_j$ for the $S_N$ transport
  calculation boundary conditions using Eq.\ \ref{eq:sn-bc}}
  \shortintertext{3. Calculate $\phi^{m+\sfrac{1}{2}}$ with the $S_N$ transport sub-solver on
  $V_1$}
  \shortintertext{4. Generate \glspl{SVDC} with $\phi^{m+\sfrac{1}{2}}$ and $J^{m+\sfrac{1}{2}}$
  using Eq.\ \ref{eq:svdc-num}}
  \shortintertext{5. Calculate $\phi^{m+1}$ with the newly generated \glspl{SVDC} and the neutron
  diffusion solver on $V_0$}
  \shortintertext{6. Check whether convergence is reached}
  \frac{|\bm{\overline{\phi}}^m-\bm{\overline{\phi}}^{m-1}|}{|\bm{\overline{\phi}}^m|} <& \
  tol_{\bm{\overline{\phi}}} \\
  \frac{|k^m-k^{m-1}|}{|k^m|} <& \ tol_k
  \shortintertext{7. Return to Step 2 if either expression is false, otherwise exit.} \nonumber
\end{align}
%
Generally, the convergence tolerance values for the outer hybrid method iteration must be smaller
than the tolerance values for the neutron diffusion and $S_N$ transport inner iterations. Using
$\phi^m$ from the neutron diffusion calculation as initial conditions for the $S_N$ transport
calculation in the $(m+1)$-th iteration helps to reduce the number of transport
sweeps required significantly.

The $S_N$ sub-solver only covers $V_1$, so it does not calculate an independent $k$
estimate. Instead, the $k^m$ from the neutron diffusion calculation scales the fission neutron
source term in the $S_N$ sub-solver.

%\section{Description of 1-D Test Cases} \label{sec:test-case}
%
%I designed ten 1-D test cases with increasing complexity to test the performance of the hybrid
%$S_N$-diffusion method in response to various geometrical features. The last two cases resemble the
%reference \gls{MSRE} design \cite{robertson_msre_1965}, which has centrally located control rods
%and air-filled rod guide tubes. Figure \ref{fig:case-geom} shows the geometries of Cases 0 to
%5b. All geometries have reflective boundary conditions at $x=0$ cm to reduce computational costs by
%creating half-core or repeating unit cell models. Cases 1a, 2a, 3a, and 4a are repeating unit
%cell models with reflecting boundaries on the right-side boundaries. Cases 1b, 2b, 3b, 4b, 5a, and
%5b are half-core models with vacuum boundaries on the right-side boundaries.
%
%Starting with an infinite, homogeneous region in Case 1a, I systematically added geometric
%complexities to each successive test case to identify how each feature impacts the neutronics
%results. Case 1b introduces a reflector region and vacuum boundary conditions relative to Case 1a.
%Cases 2a and 2b introduce a control rod region between $x=0$ cm and $x=0.5$ cm relative to Cases 1a
%and 1b. Cases 3a and 3b introduce a 0.5 cm-thick air gap between the control rod and the
%fuel-graphite mixture regions relative to Cases 2a and 2b. For Cases 4a and 4b, I replaced the
%homogeneous fuel-graphite mixture in Cases 1a and 2b with explicit fuel-graphite lattices, as shown
%in Figure \ref{fig:case-geom}. Case 5a introduces an air gap between the control rod and the
%fuel-graphite lattice regions. Lastly, Case 5b replaces the control rod region with an extended
%air gap region to provide a base case for the control rod worth calculations relative to Case 5a.
%
%The cases that do not contain control rod regions (1a, 1b, 4a, 4b, and 5b) serve as reference cases
%to quantify other sources of discrepancies in the neutron diffusion solution (e.g., geometrical
%heterogeneities, vacuum boundary condition approximation for neutron diffusion) and highlight how
%much the neutron diffusion method underperforms when the problem includes control rod regions.
%
%\begin{figure}[htb!]
%  \centering
%  \includegraphics[width=\columnwidth]{case-geometry}
%  \caption{Geometries of the 1-D test cases. The material labeled ``mixture'' represents a
%    homogeneous mixture of fuel and graphite at a ratio of 22.5\%-77.5\% by volume. All geometries
%    have reflective boundary conditions on the boundary at $x=0$ cm. The right-side boundaries are
%    reflective for Cases 1a, 2a, 3a, and 4a, and vacuum for Cases 1b, 2b, 3b, 4b, 5a, and 5b.}
%  \label{fig:case-geom}
%\end{figure}
%%
%\begin{figure}[htb!]
%  \centering
%  \begin{subfigure}[t]{.49\textwidth}
%    \centering
%    \includegraphics[width=\textwidth]{spectrum}
%    \caption{Neutron flux energy spectrum per source neutron}
%    \label{fig:spectrum}
%  \end{subfigure}
%  \hfill
%  \begin{subfigure}[t]{.49\textwidth}
%    \centering
%    \includegraphics[width=\textwidth]{reaction}
%    \caption{Flux-normalized scattering and absorption reaction rates}
%    \label{fig:reaction}
%  \end{subfigure}
%  \caption{Neutron flux energy spectrum and reaction rates for Case 5a. The dotted vertical lines
%  correspond to the discrete neutron energy group boundaries.}
%  \label{fig:spec-reac}
%\end{figure}
%%
%%\begin{table}[tb!]
%%  \centering
%%  \caption{Description of the 1-D Test Case Geometries.}
%%  \begin{tabular}{l c c c}
%%    \toprule
%%    Cases & Ordered list of regions & Ordered list of interface x-coordinates & Left \& Right BCs\\
%%    \midrule
%%    Case 0 & Control rod, homogenized fuel-graphite lattice &
%%    \bottomrule
%%  \end{tabular}
%%  \label{table:c0k}
%%\end{table}
%
%I ran all cases with four neutron energy groups. The list of energy group bounds are:
%$E=[10^{-5}, 10^0, 10^2, 10^5, 10^8]$ eV. Figures \ref{fig:spectrum} and \ref{fig:reaction} show
%how this energy group structure
%partitions the neutron flux energy spectrum and neutron reaction rates. For this preliminary work,
%the four-group structure is sufficient for showing different scattering and absorption trends of
%slow, intermediate, and fast neutrons. I also ran all cases on OpenMC to generate the required
%group constants and to assess the accuracy of the deterministic methods against OpenMC in
%continuous-energy (OpenMC-CE) and multigroup (OpenMC-MG) modes. Table \ref{table:var} shows how the
%position, direction of travel, neutron energy, and angle-dependence in $\Sigma_s$ are handled by
%OpenMC and the deterministic methods. Comparing OpenMC-MG results with OpenMC-CE results allows us
%to quantify errors arising from neutron energy group discretization and the scattering cross
%section simplifications. Each Monte Carlo simulation ran on 25 inactive and 125 active cycles of
%20,000 neutrons each, which sums up to 2.5 million tallied neutron histories.
%
%\begin{table}[tb!]
%  \centering
%  \footnotesize
%  \caption{Variable handling in OpenMC under continuous-energy (OpenMC-CE) and multigroup
%  (OpenMC-MG) modes, and in the $S_N$ neutron transport, neutron diffusion, and hybrid
%  $S_N$-diffusion methods. }
%  \begin{tabular}{c c c c c c}
%    \toprule
%    Variable & OpenMC-CE & OpenMC-MG & $S_N$ Transport & Diffusion & Hybrid \\
%    \midrule
%    Position, $\bm{r}$ & Continuous & Continuous & Discrete & Discrete & Discrete \\
%    Direction of travel, $\bm{\hat{\Omega}}$ & Continuous & Continuous & Discrete & N/A & N/A \\
%    Energy, $E$ & Continuous & Discrete & Discrete & Discrete & Discrete \\
%    Angle-dependence in $\Sigma_s$ & Continuous & 2nd Legendre moment & 2nd Legendre moment
%    & N/A & N/A \\
%    \bottomrule
%  \end{tabular}
%  \label{table:var}
%\end{table}
%
%\section{Simulation Parameters for Convergence} \label{sec:sim-param}
%
%The convergence of the results depends mainly on the following simulation parameters: mesh size
%($\Delta x$), convergence tolerance ($tol$), number of discrete ordinates ($N$), and the highest
%order moment of the scattering cross sections ($L$). The last two parameters apply only to $S_N$
%calculations. Table \ref{table:param} shows the values of simulation parameters applied to each
%method for the 1-D test cases.
%
%\begin{table}[tb!]
%  \centering
%  \footnotesize
%  \caption{Simulation parameters applied to the neutron diffusion, $S_N$, and hybrid
%  $S_N$-diffusion methods.}
%  \begin{tabular}{c c c c}
%    \toprule
%    Parameter & $S_N$ Transport & Neutron Diffusion & Hybrid $S_N$-Diffusion \\
%    \midrule
%    Mesh size, $\Delta x$ [cm]        & 0.0125    & 0.0125    & 0.0125 \\[.2cm]
%    \multirow{3}{*}{Convergence tolerance, $tol$}
%                                      &           &           & Diffusion: $10^{-6}$ \\
%                                      & $10^{-7}$ & $10^{-6}$ & $S_N$: $10^{-5}$ \\
%                                      &           &           & Outer: $10^{-3}$ \\[.2cm]
%    Discrete ordinates, $N$           & 8         & N/A       & 8 \\
%    Highest moment of $\Sigma^{g'\rightarrow g}_s$, $L$
%                                      & 2         & N/A       & 2 \\
%    \bottomrule
%  \end{tabular}
%  \label{table:param}
%\end{table}
%
%The simulation parameters did not show any significant correlation in the convergence tests. All
%three methods achieve reasonable convergence with $\Delta x=0.0125$ cm; halving the mesh size
%causes less than $30\times10^{-5}$ change in $k$ for the neutron diffusion and hybrid methods and
%less than $10^{-5}$ for the $S_8$ method. Further refinement of the $tol$, $N$, and $L$ parameters
%result in less than $2\times10^{-5}$
%change in $k$. The same $tol$ value applies to convergence checks for both $\phi$ and $k$ in all
%three methods because the variables converge at similar rates. The hybrid method requires three
%separate $tol$ values due to the inner and outer iterations. Notably, the hybrid method can reach
%convergence with a larger $tol$ value for its $S_N$ sub-solver than the $tol$ value for the
%standalone $S_N$ solver, which implies that $J$ and $\sfrac{d\phi}{dx}$ converge faster than $\phi$
%because the \glspl{SVDC} depend on those variables instead of $\phi$. Consequently, the $S_N$
%sub-solver requires much fewer iterations than the standalone $S_N$ solver. In combination with
%restricting $S_N$ calculations to the correction region, the hybrid method saves significant
%computation time compared to the standalone $S_N$ method.
%
%The outer iterations
%converge rapidly, so a $tol$ value of $10^{-3}$ is sufficient for convergence. The $\phi$ and $k$
%relative error norms ($\epsilon$) in all test cases decreased by two orders of magnitude on
%every outer iteration. For instance, Table \ref{table:epsilon} shows the $\epsilon$ values from the
%hybrid method after every outer iteration for Case 5a. The $\epsilon$ values fall below
%$10^{-5}$ by the 3rd and 4th iterations.
%
%\begin{table}[tb!]
%  \centering
%  \footnotesize
%  \caption{Relative error norm, $\epsilon$, values for $k$ and $\phi$ after each outer iteration in
%    the hybrid $S_N$-diffusion method for Case 5a.}
%  \begin{tabular}{c S[table-format=1.1e+1] S[table-format=1.1e+1]}
%    \toprule
%    Iteration   & {$\epsilon_k$}    & {$\epsilon_\phi$} \\
%    \midrule
%    1           & 1.6e-2            & 1.5e-2 \\
%    2           & 5.0e-5            & 4.7e-5 \\
%    3           & 1.4e-7            & 1.6e-7 \\
%    4           & 3.5e-9            & 1.4e-8 \\
%    \bottomrule
%  \end{tabular}
%  \label{table:epsilon}
%\end{table}
%
%
%\section{Results \& Discussion} \label{sec:prelim-results}
%
%In this section, I compare neutronics parameters for all test cases calculated from
%the OpenMC and deterministic methods. I refer to the OpenMC calculations under continuous-energy
%and multigroup modes as OpenMC-CE and OpenMC-MG, respectively. I did not apply the hybrid method to
%Cases 1a, 1b, 4a, 4b, and 5b, which serve as reference cases for how much the neutron diffusion
%method underperforms when control rod regions are present in the other cases.
%
%\subsection{Comparison of Multiplication factors, $k$}
%
%Tables \ref{table:ck1} and \ref{table:ck2} show all test case $k$ estimates from the OpenMC and
%deterministic methods. The tables also include statistical uncertainties for the OpenMC
%$k$ estimates.
%
%All five methods generally show reasonable agreement within each test case. The differences between
%the OpenMC-CE and OpenMC-MG $k$ estimates vary significantly from 0.00018 in Case 1a to 0.01622 in
%Case 4b. These Monte Carlo methods exhibit more significant differences in cases
%with vacuum boundary conditions (e.g., Cases 1b, 2b, 3b, 4b, 5a, 5b). These differences imply that
%the neutron energy discretization scheme must be revised to capture neutron leakage rates
%accurately. Beyond this preliminary study, I will explore better discrete energy group structures
%for further neutronics analyses of graphite-moderated \glspl{MSR}. For the rest of this
%section, I compare the $k$ estimates of the deterministic methods with OpenMC-MG.
%
%\begin{table}[htb!]
%  \centering
%  \footnotesize
%  \caption{Multiplication factor $k$ estimates for Cases 1a, 1b, 2a, 2b, 3a, and 3b from the
%    OpenMC-CE, OpenMC-MG, $S_8$ neutron transport, neutron diffusion, and hybrid $S_N$-diffusion
%    methods.}
%  \begin{tabular}{c S[table-format=1.5(2)] S[table-format=1.5(2)] S[table-format=1.5(2)]
%  S[table-format=1.5(2)] S[table-format=1.5(2)] S[table-format=1.5(2)]}
%    \toprule
%    \multirow{2}{*}{\textbf{Method}} &
%    \multicolumn{6}{c}{\textbf{Multiplication factor,} $\bm{k}$} \\
%    \cmidrule{2-7}
%    & {\textbf{Case 1a}} & {\textbf{Case 1b}} & {\textbf{Case 2a}} &
%    {\textbf{Case 2b}} & {\textbf{Case 3a}} & {\textbf{Case 3b}} \\
%    \midrule
%    OpenMC-CE & 1.60033(56) & 1.16749(60) & 1.20666(66) & 0.69499(47) & 1.19908(63) & 0.68565(53)\\
%    OpenMC-MG & 1.60015(33) & 1.18104(54) & 1.20570(66) & 0.70539(50) & 1.19976(48) & 0.69730(50)\\
%    $S_8$     & 1.60035     & 1.18068     & 1.20596     & 0.70470     & 1.19968     & 0.69652    \\
%    Diffusion & 1.60036     & 1.17932     & 1.19709     & 0.69274     & 1.19064     & 0.68450    \\
%    Hybrid    & {N/A}       & {N/A}       & 1.20630     & 0.70387     & 1.20004     & 0.69568    \\
%    \bottomrule
%  \end{tabular}
%  \label{table:ck1}
%\end{table}
%%
%\begin{table}[htb!]
%  \centering
%  \footnotesize
%  \caption{Multiplication factor $k$ estimates for Cases 4a, 4b, 5a, and 5b from the OpenMC-CE,
%    OpenMC-MG, $S_8$ neutron transport, neutron diffusion, and hybrid $S_N$-diffusion methods.}
%  \begin{tabular}{c S[table-format=1.5(2)] S[table-format=1.5(2)] S[table-format=1.5(2)]
%    S[table-format=1.5(2)]}
%    \toprule
%    \multirow{2}{*}{\textbf{Method}} &
%    \multicolumn{4}{c}{\textbf{Multiplication factor,} $\bm{k}$} \\
%    \cmidrule{2-5}
%    & {\textbf{Case 4a}} & {\textbf{Case 4b}} & {\textbf{Case 5a}} &
%    {\textbf{Case 5b}} \\
%    \midrule
%    OpenMC-CE & 1.64506(53) & 1.19116(63) & 0.70352(64) & 1.16832(58) \\
%    OpenMC-MG & 1.64355(37) & 1.20738(63) & 0.71396(50) & 1.18204(53) \\
%    $S_8$     & 1.64376     & 1.20744     & 0.71379     & 1.18218     \\
%    Diffusion & 1.64355     & 1.20811     & 0.70524     & 1.18274     \\
%    Hybrid    & {N/A}       & {N/A}       & 0.71673     & {N/A}       \\
%    \bottomrule
%  \end{tabular}
%  \label{table:ck2}
%\end{table}
%%
%\begin{table}[htb!]
%  \centering
%  \footnotesize
%  \caption{Differences in $k$ estimates for Cases 1a, 1b, 2a, 2b, 3a, and 3b for the $S_8$ neutron
%    transport, neutron diffusion, and hybrid $S_N$-diffusion methods relative to OpenMC-MG.}
%  \begin{tabular}{c S[table-format=+1.5(2)] S[table-format=+1.5(2)] S[table-format=+1.5(2)]
%      S[table-format=+1.5(2)] S[table-format=+1.5(2)] S[table-format=+1.5(2)]}
%    \toprule
%    \multirow{2}{*}{\textbf{Method}} & \multicolumn{6}{c}{$\bm{k-k_{MG}}$} \\
%    \cmidrule{2-7}
%    & {\textbf{Case 1a}} & {\textbf{Case 1b}} & {\textbf{Case 2a}} &
%    {\textbf{Case 2b}} & {\textbf{Case 3a}} & {\textbf{Case 3b}} \\
%    \midrule
%    $S_8$     & +0.00020(33) & -0.00036(54) & +0.00026(66) & -0.00069(50) & -0.00008(48) & -0.00078(50) \\
%    Diffusion & +0.00021(33) & -0.00172(54) & -0.00861(66) & -0.01265(50) & -0.00912(48) & -0.01280(50) \\
%    Hybrid    & {N/A}        & {N/A}        & +0.00060(66) & -0.00152(50) & +0.00028(48) & -0.00162(50) \\
%    \bottomrule
%  \end{tabular}
%  \label{table:ckdiff1}
%\end{table}
%%
%\begin{table}[htb!]
%  \centering
%  \footnotesize
%  \caption{Differences in $k$ estimates for Cases 4a, 4b, 5a, and 5b for the $S_8$ neutron
%    transport, neutron diffusion, and hybrid $S_N$-diffusion methods relative to OpenMC-MG.}
%  \begin{tabular}{c S[table-format=+1.5(2)] S[table-format=+1.5(2)] S[table-format=+1.5(2)]
%    S[table-format=+1.5(2)]}
%    \toprule
%    \multirow{2}{*}{\textbf{Method}} & \multicolumn{4}{c}{$\bm{k-k_{MG}}$} \\
%    \cmidrule{2-5}
%    & {\textbf{Case 4a}} & {\textbf{Case 4b}} & {\textbf{Case 5a}} &
%    {\textbf{Case 5b}} \\
%    \midrule
%    $S_8$     & +0.00021(37) & +0.00006(63) & -0.00017(50) & -0.00027(53) \\
%    Diffusion & +0.00000(37) & +0.00073(63) & -0.00872(50) & +0.00277(53) \\
%    Hybrid    & {N/A}        & {N/A}        & +0.00070(50) & {N/A}    \\
%    \bottomrule
%  \end{tabular}
%  \label{table:ckdiff2}
%\end{table}
%%
%\begin{table}[tb!]
%  \centering
%  \footnotesize
%  \caption{Control rod worths calculated as changes in the reactivity between Case 5a and 5b for
%    the OpenMC continuous-energy and multigroup mode calculations, and the $S_8$ neutron transport,
%  neutron diffusion, and hybrid $S_N$-diffusion methods.}
%  \begin{tabular}{c S}
%    \toprule
%    \multirow{2}{*}{\textbf{Method}} & {\textbf{Control rod worth}} \\
%                                     & {$\bm{\rho_{(5b)}-\rho_{(5a)}}$} \\
%    \midrule
%    OpenMC-CE & 0.56549(213) \\
%    OpenMC-MG & 0.55464(176) \\
%    $S_8$     & 0.55507 \\
%    Diffusion & 0.57246 \\
%    Hybrid    & 0.54974 \\
%    \bottomrule
%  \end{tabular}
%  \label{table:rod-worth}
%\end{table}
%
%Tables \ref{table:ckdiff1} and \ref{table:ckdiff2} show the differences in $k$ between OpenMC-MG
%and deterministic methods. The $k$ estimates from the $S_8$ neutron transport method show
%good agreement, with the estimates from OpenMC-MG with all pairs of $k$ values falling within
%1-2 standard deviations of each other. The only significant difference in implementation between
%the OpenMC-MG and $S_8$ methods is the discretization of the direction of travel variable
%$\hat{\Omega}$. The flux distributions exhibit more apparent differences, will be discussed in the
%following subsection. Nevertheless, the $S_8$ method is sufficiently accurate and consistent for
%modeling this graphite-moderated system with control rods.
%
%As expected, the neutron diffusion method performs significantly worse than the $S_8$ method in
%Cases 2, 3, and 5a, containing control rod regions. It deviates from the OpenMC-MG $k$
%estimate by up to -0.01280 in Case 3b compared to -0.00078 for the $S_8$ method. The hybrid method
%effectively improves the $k$ estimate to -0.00162 of the
%OpenMC-MG method for Case 3b. The hybrid method also appreciably improved the accuracy in all
%other applied cases.
%
%In a time-dependent multiphysics simulation involving rod
%insertion or withdrawal, the control rod worth determines the change in reactivity and the
%subsequent effects on the neutron flux and temperature distribution. I calculated the control rod
%worth by taking the change in reactivity $\rho$ between Case 5a and 5b for all five methods, as
%shown in Table \ref{table:rod-worth}. The formula for $\rho$ is given as
%%
%\begin{align}
%  \rho =& \frac{k-1}{k}.
%\end{align}
%%
%I took the $k$ estimate from the neutron diffusion method in Case 5b as the reference non-inserted
%case for the control rod worth calculation for the hybrid method. Compared with OpenMC-MG, the
%control rod worth from the neutron diffusion method deviates the most by 1782 pcm (percent mille)
%as opposed to 43 pcm and -490 pcm by the $S_8$ and hybrid methods, respectively.
%
%\subsection{Comparison of Neutron Flux Distributions, $\phi$}
%
%This section focuses on the neutron flux distributions in Cases 3b and 5a, representing the most
%complex geometries with homogeneous and lattice fuel-graphite regions. I sampled the flux
%distributions in OpenMC-CE and OpenMC-MG across coarser $0.1$-cm intervals to reduce the
%statistical uncertainties of each flux reading.
%
%\begin{figure}[htb!]
%  \centering
%  \begin{subfigure}[t]{.49\textwidth}
%    \centering
%    \includegraphics[width=\textwidth]{case-3b-group-1-flux}
%    \caption{Group 1}
%    \label{fig:c3bg1}
%  \end{subfigure}
%  \hfill
%  \begin{subfigure}[t]{.49\textwidth}
%    \centering
%    \includegraphics[width=\textwidth]{case-3b-group-2-flux}
%    \caption{Group 2}
%    \label{fig:c3bg2}
%  \end{subfigure}
%  \begin{subfigure}[t]{.49\textwidth}
%    \centering
%    \includegraphics[width=\textwidth]{case-3b-group-3-flux}
%    \caption{Group 3}
%    \label{fig:c3bg3}
%  \end{subfigure}
%  \hfill
%  \begin{subfigure}[t]{.49\textwidth}
%    \centering
%    \includegraphics[width=\textwidth]{case-3b-group-4-flux}
%    \caption{Group 4}
%    \label{fig:c3bg4}
%  \end{subfigure}
%  \caption{Neutron group flux distributions for Case 3b from OpenMC-CE, OpenMC-MG, $S_8$, neutron
%  diffusion, and hybrid methods.}
%  \label{fig:c3bflux}
%\end{figure}
%
%Figure \ref{fig:c3bflux} shows the neutron flux distributions for groups 1 to 4. Notably, the
%OpenMC-CE flux distribution is the most dissimilar, likely due to
%the inadequate neutron energy discretization scheme. Consequently, I compared the
%deterministic methods with OpenMC-MG to eliminate energy discretization errors. Figure
%\ref{fig:c3bfluxe} shows the relative differences in the flux distributions from $S_8$, neutron
%diffusion, and hybrid methods with respect to OpenMC-MG. The neutron diffusion method
%performs worse than $S_8$, especially near the control rod region between $x=0$ cm and $x=20$ cm
%for the slower group 3 and 4 neutron flux. Although the control rod region spans between $x=0$ cm
%and $x=0.5$ cm only, it induces strong flux gradients in its vicinity, thus rendering the neutron
%diffusion method less valid. The hybrid method significantly improves all four
%group fluxes in this region. Beyond $x=40$ cm, the neutron diffusion and hybrid methods produce
%similar flux error distributions due to the influence of the material interface between the
%fuel-graphite and reflector regions.
%
%\begin{figure}[htb!]
%  \centering
%  \begin{subfigure}[t]{.49\textwidth}
%    \centering
%    \includegraphics[width=\textwidth]{case-3b-group-1-flux-error}
%    \caption{Group 1}
%    \label{fig:c3bg1e}
%  \end{subfigure}
%  \hfill
%  \begin{subfigure}[t]{.49\textwidth}
%    \centering
%    \includegraphics[width=\textwidth]{case-3b-group-2-flux-error}
%    \caption{Group 2}
%    \label{fig:c3bg2e}
%  \end{subfigure}
%  \begin{subfigure}[t]{.49\textwidth}
%    \centering
%    \includegraphics[width=\textwidth]{case-3b-group-3-flux-error}
%    \caption{Group 3}
%    \label{fig:c3bg3e}
%  \end{subfigure}
%  \hfill
%  \begin{subfigure}[t]{.49\textwidth}
%    \centering
%    \includegraphics[width=\textwidth]{case-3b-group-4-flux-error}
%    \caption{Group 4}
%    \label{fig:c3bg4e}
%  \end{subfigure}
%  \caption{Relative differences of the neutron group flux distributions for Case 3b from the $S_8$,
%    neutron diffusion, and hybrid methods with respect to OpenMC-MG.}
%  \label{fig:c3bfluxe}
%\end{figure}
%
%For a quantitative assessment, I defined the normalized flux error $\varepsilon$ (not to be
%confused with relative error norm $\epsilon$) for each energy group with respect to the OpenMC-MG
%flux distribution using a normalized Frobenius/Euclidean norm formula given as
%%
%\begin{align}
%  \varepsilon_g =& \frac{\lVert\sum^I_{i=1}\phi_{g,i} - \phi^{MG}_{g,i}\rVert_F}
%  {\lVert\sum^I_{i=1}\phi^{MG}_{g,i}\rVert_F} =
%  \frac{\left[\sum^I_{i=1}\lvert\phi_{g,i} - \phi^{MG}_{g,i}\rvert^2 \right]^{\sfrac{1}{2}}}
%  {\left[\sum^I_{i=1}\lvert\phi^{MG}_{g,i}\rvert^2\right]^{\sfrac{1}{2}}}
%\end{align}
%%
%I integrated on the fine flux distributions over 0.1-cm intervals from $S_8$, neutron diffusion,
%and hybrid methods to match the 0.1-cm intervals of the OpenMC-MG flux distribution. Table
%\ref{table:c3berror} shows the $\varepsilon$ values for the deterministic methods. As expected,
%the hybrid method produces smaller flux error norms than the neutron diffusion method. The
%improvements are more pronounced for group 3 and 4 fluxes than group 1 and 2 fluxes.
%%
%\begin{table}[tb!]
%  \centering
%  \footnotesize
%  \caption{Normalized flux error $\varepsilon$ for Case 3b from the $S_8$, neutron diffusion, and
%    hybrid methods with respect to OpenMC-MG.}
%  \begin{tabular}{c S S S S}
%    \toprule
%    {\multirow{2}{*}{\textbf{Method}}} &
%    \multicolumn{4}{c}{\textbf{Normalized flux error,} $\bm{\varepsilon_g}$} \\
%    \cmidrule{2-5}
%    & {\textbf{Group 1}} & {\textbf{Group 2}} & {\textbf{Group 3}} &
%    {\textbf{Group 4}} \\
%    \midrule
%    $S_8$     & 0.0066 & 0.0048 & 0.0058 & 0.0067 \\
%    Diffusion & 0.0144 & 0.0147 & 0.0258 & 0.0267 \\
%    Hybrid    & 0.0122 & 0.0099 & 0.0087 & 0.0084 \\
%    \bottomrule
%  \end{tabular}
%  \label{table:c3berror}
%\end{table}
%
%Case 5a introduces significant geometrical heterogeneity from the fuel-graphite lattice region.
%The heterogeneity is very apparent from the flux distributions in Figure \ref{fig:c5aflux}. The
%OpenMC Monte
%Carlo methods captured the fluctuations in the flux through their continuous angle variable as
%opposed to the $S_8$, neutron diffusion, and hybrid methods. The group 1 flux peaks are notably
%pronounced since all fission neutrons are born in the fuel region, with the majority in group 1.
%Figure \ref{fig:c5afluxe} shows the relative differences in flux for the $S_8$, neutron diffusion,
%and hybrid methods relative to OpenMC-MG. All three methods show regular fluctuations in the
%relative flux differences corresponding to the fuel-graphite lattice. The $S_8$ method generally
%deviates the least from OpenMC-MG, while the neutron diffusion method exhibits
%significant flux deviations near the control rod region prior to $x=20$ cm and the material
%interface between the fuel and reflector regions beyond $x=40$ cm.
%
%\begin{figure}[htb!]
%  \centering
%  \begin{subfigure}[t]{.49\textwidth}
%    \centering
%    \includegraphics[width=\textwidth]{case-5a-group-1-flux}
%    \caption{Group 1}
%    \label{fig:c5ag1}
%  \end{subfigure}
%  \hfill
%  \begin{subfigure}[t]{.49\textwidth}
%    \centering
%    \includegraphics[width=\textwidth]{case-5a-group-2-flux}
%    \caption{Group 2}
%    \label{fig:c5ag2}
%  \end{subfigure}
%  \begin{subfigure}[t]{.49\textwidth}
%    \centering
%    \includegraphics[width=\textwidth]{case-5a-group-3-flux}
%    \caption{Group 3}
%    \label{fig:c5ag3}
%  \end{subfigure}
%  \hfill
%  \begin{subfigure}[t]{.49\textwidth}
%    \centering
%    \includegraphics[width=\textwidth]{case-5a-group-4-flux}
%    \caption{Group 4}
%    \label{fig:c5ag4}
%  \end{subfigure}
%  \caption{Neutron group flux distributions for Case 5a from OpenMC-CE, OpenMC-MG, $S_8$, neutron
%  diffusion, and hybrid methods.}
%  \label{fig:c5aflux}
%\end{figure}
%%
%\begin{figure}[htb!]
%  \centering
%  \begin{subfigure}[t]{.49\textwidth}
%    \centering
%    \includegraphics[width=\textwidth]{case-5a-group-1-flux-error}
%    \caption{Group 1}
%    \label{fig:c5ag1e}
%  \end{subfigure}
%  \hfill
%  \begin{subfigure}[t]{.49\textwidth}
%    \centering
%    \includegraphics[width=\textwidth]{case-5a-group-2-flux-error}
%    \caption{Group 2}
%    \label{fig:c5ag2e}
%  \end{subfigure}
%  \begin{subfigure}[t]{.49\textwidth}
%    \centering
%    \includegraphics[width=\textwidth]{case-5a-group-3-flux-error}
%    \caption{Group 3}
%    \label{fig:c5ag3e}
%  \end{subfigure}
%  \hfill
%  \begin{subfigure}[t]{.49\textwidth}
%    \centering
%    \includegraphics[width=\textwidth]{case-5a-group-4-flux-error}
%    \caption{Group 4}
%    \label{fig:c5ag4e}
%  \end{subfigure}
%  \caption{Relative differences of the neutron group flux distributions for Case 5a from the $S_8$,
%    neutron diffusion, and hybrid methods with respect to OpenMC-MG.}
%  \label{fig:c5afluxe}
%\end{figure}
%%
%\begin{table}[tb!]
%  \centering
%  \footnotesize
%  \caption{Normalized flux error $\varepsilon$ for Case 5a from the $S_8$, neutron diffusion, and
%    hybrid methods with respect to OpenMC-MG.}
%  \begin{tabular}{c S S S S}
%    \toprule
%    {\multirow{2}{*}{\textbf{Method}}} &
%    \multicolumn{4}{c}{\textbf{Normalized flux error,} $\bm{\varepsilon_g}$} \\
%    \cmidrule{2-5}
%    & {\textbf{Group 1}} & {\textbf{Group 2}} & {\textbf{Group 3}} &
%    {\textbf{Group 4}} \\
%    \midrule
%    $S_8$     & 0.0078 & 0.0074 & 0.0083 & 0.0085 \\
%    Diffusion & 0.0305 & 0.0202 & 0.0313 & 0.0406 \\
%    Hybrid    & 0.0290 & 0.0147 & 0.0143 & 0.0216 \\
%    \bottomrule
%  \end{tabular}
%  \label{table:c5aerror}
%\end{table}
%
%\subsection{Correction Regions \& Diffusion Coefficients}
%
%In this section, I compare the $P_1$-based diffusion coefficients, reference \glspl{SVDC}, and
%hybrid \glspl{SVDC} for Cases 3b and 5a. As in Section \ref{sec:hybrid-method}, the reference
%\glspl{SVDC} are derived from the $S_8$ flux solution, while the hybrid \glspl{SVDC} are derived
%from the $S_8$ sub-solver flux solution within the hybrid $S_N$-diffusion method.
%
%Figure \ref{fig:c3bdiffcoef} shows the various diffusion coefficient distributions for Case 3b
%between $x=0$ cm and $x=30$ cm. The figures omit the large diffusion coefficient values in the
%air gap region between $x=0.5$ cm and $x=1$ cm since they do not show any significant features. The
%hybrid \glspl{SVDC} closely match
%the reference \gls{SVDC} within most of the correction region, which terminates at $x=20$ cm.
%
%\begin{figure}[htb!]
%  \centering
%  \begin{subfigure}[t]{.49\textwidth}
%    \centering
%    \includegraphics[width=\textwidth]{case-3b-group-1-diffcoef}
%    \caption{Group 1}
%    \label{fig:c3bg1dc}
%  \end{subfigure}
%  \hfill
%  \begin{subfigure}[t]{.49\textwidth}
%    \centering
%    \includegraphics[width=\textwidth]{case-3b-group-2-diffcoef}
%    \caption{Group 2}
%    \label{fig:c3bg2dc}
%  \end{subfigure}
%  \begin{subfigure}[t]{.49\textwidth}
%    \centering
%    \includegraphics[width=\textwidth]{case-3b-group-3-diffcoef}
%    \caption{Group 3}
%    \label{fig:c3bg3dc}
%  \end{subfigure}
%  \hfill
%  \begin{subfigure}[t]{.49\textwidth}
%    \centering
%    \includegraphics[width=\textwidth]{case-3b-group-4-diffcoef}
%    \caption{Group 4}
%    \label{fig:c3bg4dc}
%  \end{subfigure}
%  \caption{$P_1$-based diffusion coefficient, reference \gls{SVDC}, and hybrid \gls{SVDC}
%    distributions for Case 3b between $x=0$ cm and $x=30$ cm.}
%  \label{fig:c3bdiffcoef}
%\end{figure}
%%
%\begin{figure}[htb!]
%  \centering
%  \includegraphics[width=.6\textwidth]{case-3b-grad-j}
%  \caption{The gradient of current $J$ in the homogeneous fuel-graphite mixture from $x=1$ cm to
%  $x=45$ cm for Case 3b.}
%  \label{fig:c3bgradj}
%\end{figure}
%
%We recall that the hybrid method determines cutoff points within the correction region where the
%\glspl{SVDC} coincide with the $P_1$-based diffusion coefficients so that the flux gradients
%remain continuous. The hybrid method discards \glspl{SVDC} in the buffer region between this
%cutoff point and the correction region boundary in favor of the $P_1$-based diffusion coefficients.
%Notably, the group 1 reference \glspl{SVDC} (Figure \ref{fig:c3bg1dc}) only coincides
%with the $P_1$-based diffusion coefficient after the vertical asymptotes where the flux
%gradient approaches zero. The trends in group 1 reference \glspl{SVDC} before and after the
%vertical asymptote imply that the \gls{SVDC} would have gradually approached the $P_1$-based
%diffusion coefficient value if the asymptote did not exist. Determining the cutoff point for the
%\glspl{SVDC} would have been problematic due to the vertical asymptote. However, the hybrid
%\glspl{SVDC} coincide with the $P_1$-based diffusion coefficients before the vertical asymptote.
%This behavior can be reliably reproduced in other geometries with large homogeneous fissile regions
%(Cases 2a, 2b, and 3a) as long as the correction region terminates near the neutron flux peaks.
%Nevertheless, the issue of handling vertical asymptotes in the \glspl{SVDC} has to be addressed
%since flux peaks and troughs will inevitably occur in more realistic reactor geometries.
%
%In contrast with the flux distributions in Figure \ref{fig:c3bflux}, the \gls{SVDC} distributions
%suggest that the control rod contributes to an extended range of transport corrections on group 1
%neutrons than the slower neutron groups. This phenomenon can be largely attributed to the highly
%non-uniform fission neutron source distribution, which serves as the only source of group 1
%neutrons in the absence of neutron up-scattering. The group 4 flux distribution
%significantly influences on fission neutron source distribution, leading to more group 1 neutrons
%being born closer to the reflector on the right than the
%control rod on the left. The biased neutron source distribution induces a highly non-linear neutron
%current distribution, as shown in Figure \ref{fig:c3bgradj}, and a departure from Fick's first law,
%which relates the current to the flux gradient. Therefore, the \gls{SVDC} distributions show that
%transport corrections to the faster neutron groups are as necessary for overall accuracy as
%transport corrections to the slower neutron groups.
%
%Figure \ref{fig:c5adiffcoef} shows the various diffusion coefficient distributions for Case 5a
%between $x=0$ cm and $x=30$ cm. As with Case 3b, the figures omit the large diffusion coefficient
%values in the air gap region between $x=0.5$ cm and $x=1.5$ cm  The hybrid \glspl{SVDC} closely match
%the reference \gls{SVDC} within most of the correction region, which terminates at $x=10$ cm.
%
%\begin{figure}[htb!]
%  \centering
%  \begin{subfigure}[t]{.49\textwidth}
%    \centering
%    \includegraphics[width=\textwidth]{case-5a-group-1-diffcoef}
%    \caption{Group 1}
%    \label{fig:c5ag1dc}
%  \end{subfigure}
%  \hfill
%  \begin{subfigure}[t]{.49\textwidth}
%    \centering
%    \includegraphics[width=\textwidth]{case-5a-group-2-diffcoef}
%    \caption{Group 2}
%    \label{fig:c5ag2dc}
%  \end{subfigure}
%  \begin{subfigure}[t]{.49\textwidth}
%    \centering
%    \includegraphics[width=\textwidth]{case-5a-group-3-diffcoef}
%    \caption{Group 3}
%    \label{fig:c5ag3dc}
%  \end{subfigure}
%  \hfill
%  \begin{subfigure}[t]{.49\textwidth}
%    \centering
%    \includegraphics[width=\textwidth]{case-5a-group-4-diffcoef}
%    \caption{Group 4}
%    \label{fig:c5ag4dc}
%  \end{subfigure}
%  \caption{$P_1$-based diffusion coefficient, reference \gls{SVDC}, and hybrid \gls{SVDC}
%    distributions for Case 5a between $x=0$ cm and $x=30$ cm.}
%  \label{fig:c5adiffcoef}
%\end{figure}
%
%Unlike Case 3b, the heterogeneous fuel-graphite lattice structure in Case 5a gives rise to
%extreme fluctuations in the \gls{SVDC} distributions. The numerous asymptotes correspond to the
%flux peaks and troughs throughout the lattice. Regardless, the presence of the control rod
%suppresses the \gls{SVDC} fluctuations in the lattice close to the control rod, and the cutoff
%points can be determined without manual intervention. The heterogeneous lattice induces competing
%effects on the neutron flux, allowing for a smaller correction region of 10 cm in width. However,
%the reduced correction region introduces less transport correction as a consequence.
%Further investigation is needed to observe how the \gls{SVDC} distributions change with
%increases in dimensionality for 2-D and 3-D systems.
%
%\section{Summary} \label{sec:hybrid-summary}
%
%In this chapter, I presented the theoretical basis for a hybrid $S_N$-diffusion method, a novel
%method for improving the neutron diffusion method for neutronics simulations of reactor
%systems containing highly neutron-absorbing regions. The hybrid method relies on \glspl{SVDC}
%derived from reference flux solutions of $S_N$ calculations to provide transport
%corrections to the neutron diffusion method. The hybrid method is less computationally intensive
%than the standalone $S_N$ neutron transport method because 1) $S_N$ calculations are limited to
%correction regions which cover a fraction of the overall geometry, 2) and the \glspl{SVDC} are
%observed to converge faster than the neutron fluxes in $S_N$ calculations.
%
%I established a framework for the hybrid method for coupling an $S_N$ sub-solver and a neutron
%diffusion solver through outer iterations by taking advantage of the weak dependence of the $S_N$
%flux solution on the approximate boundary conditions determined from the neutron diffusion flux
%solution. This framework could be extended to improve multischeme methods, which
%divide a problem domain into subdomains where different neutronics methods are applied. For
%instance, in an $S_N$-diffusion multischeme method, the $S_N$ method is applied in subdomains with
%significant transport effects while other subdomains are treated with the neutron diffusion method
%to reduce the overall computational effort required \cite{wang_rattlesnake_2021}. 
%
%I also developed an algorithm for identifying and discarding inaccurate \gls{SVDC} values, which
%are expected to occur near the correction region boundary. I demonstrated the hybrid
%method on 1-D graphite-moderated test cases modeled after the \gls{MSRE}. The hybrid method showed
%good agreement with the reference Monte Carlo and $S_8$ methods after eliminating neutron energy
%group discretization errors in test cases that included highly neutron-absorbing regions. The
%hybrid method had better accuracy in more homogeneous systems largely due to the larger effective
%transport correction regions through \glspl{SVDC}. Further work
%is needed to extend and verify the hybrid method to 2-D and 3-D neutronics simulations and to
%create a robust solution for determining \glspl{SVDC} around neutron flux peaks and troughs.

### 2024-06-11 hybrid.tex
hybrid chapter edits

@@
 \subsection{\gls{SAAF} Formulation of the $S_N$ Method}
+
+\subsubsection{Multigroup Neutron Transport Equations}
 
@@
 \begin{gather}
-  \frac{\partial}{\partial t} \left(\frac{\bm{\Psi}}{\bm{v}}\right)+\mathbb{L}_1\bm{\Psi}
-  +\mathbb{L}_2\bm{\Psi} = \mathbb{S}\bm{\Psi} + \bm{Q}_f + \bm{Q}^\text{ext} \label{eq:nte-vec}
+  \frac{\partial}{\partial t} \left(\frac{\bm{\Psi}}{\bm{v}}\right)+\mathbb{L}\bm{\Psi}
+  = \mathbb{S}\bm{\Psi} + \bm{Q}_f + \bm{Q}^\text{ext} \label{eq:nte-vec}
 \end{gather}
@@
 \end{gather}
-%
-We begin the derivation for the \gls{SAAF} formulation by rearranging Eq.\ \ref{eq:nte-vec} as
+
+\subsubsection{Weak Form of the Multigroup Neutron Transport Equations}
+
+Before we begin the derivation for the \gls{SAAF} formulation, we introduce inner product notations
+to express Eq.\ \ref{eq:nte-vec} and the eventual \gls{SAAF} formulation in their weak forms to be
+solved using \gls{FEM}. We define the inner product consisting of volume integrals on
+$\mathcal{D}\otimes\mathcal{S}$ as
+%
+\begin{gather}
+  (\bm{a}, \bm{b})_{\mathcal{D}\otimes\mathcal{S}} \equiv
+  \sum^G_{g=1}\int_\mathcal{S}d\hat{\Omega}\int_\mathcal{D}d\vec{r}\
+  a_g(\vec{r},\hat{\Omega}) b_g(\vec{r},\hat{\Omega}), \label{eq:weak-domain}
+\end{gather}
+%
+where $\bm{a}$ and $\bm{b}$ represent multigroup function vectors such as $\bm{\Psi}$. We also
+define boundary integrals on $\partial\mathcal{D}\otimes\mathcal{S}$ as
+%
+\begin{gather}
+  \langle\bm{a},\bm{b}\rangle^\pm_{\partial\mathcal{D}\otimes\mathcal{S}} \equiv
+  \sum^G_{g=1}\int_{\partial\mathcal{D}}d\vec{r}
+  \int_{\mathcal{S}^{\pm}}d\hat{\Omega}\ |\hat{\Omega}\cdot\hat{n}_b|
+  a_g(\vec{r},\hat{\Omega}) b_g(\vec{r},\hat{\Omega}), \label{eq:weak-boundary}
+\end{gather}
+%
+where the $\pm$ sign depends on the signedness of $\hat{\Omega}\cdot\hat{n}_b$ at the boundary.
+We henceforth drop the phase space subscripts for brevity.
+%
+In accordance with standard weak form derivation procedure, we multiply Eq.\ \ref{eq:nte-vec}
+by a test function $\bm{\Psi}^*$ and integrate throughout over the $\mathcal{D}$ and $\mathcal{S}$
+domains
+%
+\begin{gather}
+  \left(\bm{\Psi}^*,\frac{\partial}{\partial t}\left(\frac{\bm{\Psi}}{\bm{v}}\right)\right) +
+  \left(\bm{\Psi}^*,\mathbb{L}\bm{\Psi}\right) = \left(\bm{\Psi}^*,\mathbb{S}\bm{\Psi}\right) +
+  \left(\bm{\Psi}^*,\bm{Q}\right).
+\end{gather}
+%
+We apply integration by parts on the streaming term to obtain
+%
+\begin{gather}
+  \left(\bm{\Psi}^*,\frac{\partial}{\partial t}\left(\frac{\bm{\Psi}}{\bm{v}}\right)\right) +
+  \left(\mathbb{L}^*\bm{\Psi}^*,\bm{\Psi}\right) + \langle\bm{\Psi}^*,\bm{\Psi}\rangle^+ -
+  \langle\bm{\Psi}^*,\bm{\Psi}\rangle^- = \left(\bm{\Psi}^*,\mathbb{S}\bm{\Psi}\right) +
+  \left(\bm{\Psi}^*,\bm{Q}\right), \label{eq:nte-weak}
+\end{gather}
+%
+where $\mathbb{L}^*$ is the adjoint of $\mathbb{L}$.
+%\begin{multline}
+%  \left(\bm{\Psi}^*,\left(\mathbb{I} - \mathbb{L}^{-1}_2\mathbb{L}_1\right)
+%  \frac{\partial}{\partial t}\left(\frac{\bm{\Psi}}{\bm{v}}\right)\right) -
+%  \left(\mathbb{L}^*_1\bm{\Psi}^*,\mathbb{L}^{-1}_2\mathbb{L}_1\bm{\Psi}\right) +
+%  \langle\bm{\Psi}^*,
+%  \left(\bm{\Psi}^*,\mathbb{L}_2\bm{\Psi}\right) = \\
+%  \left(\bm{\Psi}^*,\left(\mathbb{I} - \mathbb{L}^{-1}_2\mathbb{L}_1\right)\mathbb{S}\bm{\Psi}
+%  \right) +
+%  \left(\bm{\Psi}^*,\left(\mathbb{I} - \mathbb{L}^{-1}_2\mathbb{L}_1\right) \bm{Q}\right).
+%\end{multline}
+
+\subsubsection{\gls{SAAF} Formulation}
+
+We begin the derivation for the \gls{SAAF} formulation by first rearranging Eq.\ \ref{eq:nte-vec}
+as
 %
@@
 %
-We substitute Eq.\ \ref{eq:afe} back into the streaming term $\left(\mathbb{L}_1\bm{\Psi}\right)$
-in Eq.\ \ref{eq:nte-vec} and rearrange the terms to obtain
+We substitute Eq.\ \ref{eq:afe} back into the streaming term in Eq.\ \ref{eq:nte-weak} and
+rearrange the terms to obtain
+%
+\begin{multline}
+  \left(\mathbb{L}^{-1}_2\mathbb{L}\bm{\Psi}^*,\frac{\partial}{\partial t}\left(\frac{\bm{\Psi}}
+      {\bm{v}}\right)\right) + 
+  \left(\mathbb{L}_1\bm{\Psi}^*,\mathbb{L}^{-1}_2\mathbb{L}_1\bm{\Psi}\right) +
+  \langle\bm{\Psi}^*,\bm{\Psi}\rangle^+ - \langle\bm{\Psi}^*,\bm{\Psi}\rangle^- +
+  \left(\mathbb{L}_2\bm{\Psi}^*,\bm{\Psi}\right) = \\
+  \left(\mathbb{L}^{-1}_2\mathbb{L}\bm{\Psi}^*,
+  \mathbb{S}\bm{\Psi}\right) + \left(\mathbb{L}^{-1}_2\mathbb{L}\bm{\Psi}^*,\bm{Q}\right)
+  \label{eq:saaf}
+\end{multline}
+%
+using the following relations
 %
 \begin{gather}
-  \frac{\partial}{\partial t}\left(\frac{\bm{\Psi}}{\bm{v}}\right) + \mathbb{L}^{-1}_2\mathbb{L}_1
-  \mathbb{S}\bm{\Psi} + \mathbb{L}^{-1}_2\mathbb{L}_1\bm{Q} - \mathbb{L}^{-1}_2\mathbb{L}_1
-  \frac{\partial}{\partial t}\left(\frac{\bm{\Psi}}{\bm{v}}\right) - \mathbb{L}^{-1}\mathbb{L}^2_1
-  \bm{\Psi} + \mathbb{L}_2\bm{\Psi} = \mathbb{S}\bm{\Psi}+\bm{Q} \\
-  \left(\mathbb{I} - \mathbb{L}^{-1}_2\mathbb{L}_1\right) \frac{\partial}{\partial t}\left(
-  \frac{\bm{\Psi}}{\bm{v}}\right) - \mathbb{L}^{-1}_2\mathbb{L}^2_1\bm{\Psi} +
-  \mathbb{L}_2\bm{\Psi} = \left(\mathbb{I} - \mathbb{L}^{-1}_2\mathbb{L}_1\right)
-  \mathbb{S}\bm{\Psi} + \left(\mathbb{I} - \mathbb{L}^{-1}_2\mathbb{L}_1\right) \bm{Q}
-%  \mathbb{L}^{-1}_2\mathbb{L}\frac{\partial}{\partial t}\frac{\bm{\Psi}}{\bm{v}} -
-%  \mathbb{L}^{-1}_2\mathbb{L}^2_1 \bm{\Psi} + \mathbb{L}_2\bm{\Psi} = \mathbb{L}^{-1}_2\mathbb{L}
-%  \mathbb{S}\bm{\Psi} + \mathbb{L}^{-1}_2\mathbb{L}\bm{Q}
+  \mathbb{L}_1\mathbb{L}_2 = \mathbb{L}_2\mathbb{L}_1, \hspace{1cm}
+  \mathbb{L}^*_1 = -\mathbb{L}_1,\hspace{1cm}
+  \mathbb{L}^*_2 = \mathbb{L}_2, \nonumber \\
+  \mathbb{I}-\mathbb{L}^{-1}_2\mathbb{L}^*_1 =
+  \mathbb{L}^{-1}_2\mathbb{L}_2 + \mathbb{L}^{-1}_2\mathbb{L}_1 =
+  \mathbb{L}^{-1}_2\mathbb{L}. \nonumber
+\end{gather}
+
+Notably, the \gls{SAAF} formulation contains a second-order derivative streaming term
+(prior to applying integration by parts) which lends itself better to standard
+\gls{FEM} solver routines, such as the \gls{GMRES} method, than the first-order streaming
+term of the original formulation.
+However, the $\mathbb{L}^{-1}_2$ operators render the current \gls{SAAF} formulation
+unsolveable in void and near-void regions where the total cross sections $\Sigma_{t,g}$ approach
+zero \cite{wang_diffusion_2014}.
+
+\subsubsection{Spatial Discretization and Void Treatment}
+
+To overcome the issue of voids, Wang et al.\ \cite{wang_diffusion_2014} modified the standard
+\gls{SAAF} formulation by taking inspiration from the \gls{SUPG} stabilization technique
+\cite{brooks_streamline_1982}. We start by first applying spatial discretization on the \gls{SAAF}
+formulation. For \gls{FEM}, the spatial domain is discretized into finite, non-overlapping mesh
+elements $\mathcal{D}_e$ such that $\mathcal{D} = \cup_{e\in\mathbb{T}_h}\mathcal{D}_e$, where
+$\mathbb{T}_h$ denotes the index set of elements discretizing $\mathcal{D}$ and $\mathcal{D}_e$
+is the subdomain covered by the element of index $e$. Similarly, the outermost boundary
+$\partial\mathcal{D}$ is discretized by non-overlapping mesh sides indexed by $s$. Within each mesh
+element, any spatially varying function can be approximated with basis functions
+%
+\begin{gather}
+  f(\vec{r}) = \sum^N_{i=1}f_i b_i(\vec{r}),
   \shortintertext{where}
   \begin{align*}
-    \mathbb{I} &= \mbox{identity operator,} \\
-    \mathbb{I}-\mathbb{L}^{-1}_2\mathbb{L}_1 &=
-    \mathbb{L}^{-1}_2\mathbb{L}_2-\mathbb{L}^{-1}_2\mathbb{L}_1 = 
+    N &= \mbox{degrees of freedom in the mesh element,} \\
+    b_i &= \mbox{basis function,} \\
+    f_i &= \mbox{coefficient corresponding to $b_i$ in approximating $f$}.
   \end{align*}
 \end{gather}
-%
+
+Thus, we can define the inner products
+%
+\begin{gather}
+  \left(\bm{a},\bm{b}\right) \equiv \sum^G_{g=1}\int_\mathcal{S}d\hat{\Omega}
+  \sum_{e\in\mathbb{T}_h}\int_{\mathcal{D}_e}d\vec{r}\sum_{i\in E_e}a_{g,i}(\hat{\Omega})
+  b_i(\vec{r})\sum_{j\in E_e}b_{g,j}(\hat{\Omega})b_j(\vec{r}), \\
+  \langle\bm{a},\bm{b}\rangle^{\pm} \equiv \sum^G_{g=1}\sum_{s\in\partial\mathcal{D}}\int_s
+  d\vec{r} \int_{mathcal{S}^\pm_{\hat{n}_b}}d\hat{\Omega} |\hat{\Omega}\cdot\hat{n}_b|
+  \sum_{i\in E_s}a_{g,i}(\hat{\Omega})b_i(\vec{r})\sum_{j\in E_s}b_{g,j}(\hat{\Omega})b_j(\vec{r}),
+  \shortintertext{where}
+  \begin{align*}
+    E_e &= \mbox{index set of the local basis functions in element $e$,} \\
+    E_s &= \mbox{index set of the local basis functions in element side $s$,}
+  \end{align*}
+\end{gather}
+%
+to apply spatial discretization on Eq.\ \ref{eq:saaf} by replacing the preexisting inner product
+notations defined in Eqs.\ \ref{eq:weak-domain} and \ref{eq:weak-boundary}.
+
+We can now proceed with the void treatment for stabilization.
+We start by defining the following stabilization operator
+%
+\begin{gather}
+  \bm{\tau} \equiv
+  \begin{bmatrix}
+    \tau_1 \\
+    \tau_2 \\
+    \vdots \\
+    \tau_G
+  \end{bmatrix}
+  \shortintertext{where}
+  \begin{align*}
+    \tau_g &=
+    \begin{cases}
+      \frac{1}{\Sigma_{t,g}} \mbox{ for } h\Sigma_{t,g} \geq \varsigma \\
+      \frac{h}{\varsigma} \mbox{ for } h\Sigma_{t,g} < \varsigma
+    \end{cases}, & \\
+    h &= \mbox{mesh element size,} & \\
+    \varsigma &= \mbox{void constant.} &
+  \end{align*}
+\end{gather}
+%
+Next, we redefine Eq.\ \ref{eq:afe} as
+%
+\begin{gather}
+  \bm{\Psi} = \left(\mathbb{I}-\bm{\tau}\mathbb{L}_2\right)\bm{\Psi}+
+  \bm{\tau}\left[\mathbb{S}\bm{\Psi}+\bm{Q}
+  -\frac{\partial}{\partial t}\left(\frac{\bm{\Psi}}{\bm{v}}\right)-\mathbb{L}_1\bm{\Psi}\right]
+  \label{eq:afe-vt}
+\end{gather}
+%
+Substituting Eq.\ \ref{eq:afe-vt} into Eq.\ \ref{eq:nte-weak} yields the stabilized \gls{SAAF}
+formulation given as
+%
+\begin{multline}
+  \left(\left(\mathbb{I}+\bm{\tau}\mathbb{L}_1\right)\bm{\Psi}^*,
+  \frac{\partial}{\partial t}\left(\frac{\bm{\Psi}}{\bm{v}}\right)\right) + 
+  \left(\mathbb{L}_1\bm{\Psi}^*,
+  \left(\bm{\tau}\mathbb{L}_1-\mathbb{I}+\bm{\tau}\mathbb{L}_2\right)\bm{\Psi}\right) +
+  \langle\bm{\Psi}^*,\bm{\Psi}\rangle^+ - \langle\bm{\Psi}^*,\bm{\Psi}\rangle^- +
+  \left(\mathbb{L}_2\bm{\Psi}^*,\bm{\Psi}\right) \\
+  = \left(\left(\mathbb{I}+\bm{\tau}\mathbb{L}_1\right)\bm{\Psi}^*,\mathbb{S}\bm{\Psi}\right) +
+  \left(\left(\mathbb{I}+\bm{\tau}\mathbb{L}_1\right)\bm{\Psi}^*,\bm{Q}\right)
+  \label{eq:saaf-vt}
+\end{multline}
+%
+In non-void regions where $\Sigma_{t,g}$ is non-zero, we can disable the stabilization scheme by
+setting $\varsigma$ to zero. Then $\bm{\tau}\mathbb{L}=\mathbb{I}$
+and the stabilized \gls{SAAF} formulation reduces to the unstabilized formulation in Eq.\
+\ref{eq:saaf}.
+
+\subsubsection{Angular Discretization}
+
+The $S_N$ method applies a collocation method for the angular discretization by solving for the
+angular flux at $N_d$ discrete directions $\hat{\Omega}_d$ with weights $w_d$. Numerous options
+exist for the choice of the angular quadrature set \left{\hat{\Omega}_d,w_d | d=1,\dots,N_d\right}.
+For this work, we chose the commonly used level-symmetric quadrature set
+\cite{wang_rattlesnake_2018} for ease of implementation, particularly when applying reflective
+boundary conditions due to rotational symmetry about each of the three Cartesian axes.
 

### 2024-06-11 hybrid.tex
sn terms

@@
 The $S_N$ method applies a collocation method for the angular discretization by solving for the
-angular flux at $N_d$ discrete directions $\hat{\Omega}_d$ with weights $w_d$. Numerous options
-exist for the choice of the angular quadrature set \left{\hat{\Omega}_d,w_d | d=1,\dots,N_d\right}.
+angular flux along $N_d$ discrete directions $\hat{\Omega}_d$ with weights $w_d$. Numerous options
+exist for the choice of the angular quadrature set
+$\{\hat{\Omega}_d,w_d | d=1,\dots,N_d\}$.
 For this work, we chose the commonly used level-symmetric quadrature set
 \cite{wang_rattlesnake_2018} for ease of implementation, particularly when applying reflective
-boundary conditions due to rotational symmetry about each of the three Cartesian axes.
-
-The discrete ordinates ($S_N$) method for solving the multigroup neutron transport equation
-(Eq.\ \ref{eq:mg-bte}) discretizes the continuous angular directional phase space into a finite
-number of discrete angular intervals (ordinates). The set of distinct direction variables
-$\hat{\Omega}_d$ representing the discrete ordinates is typically chosen in conjunction with a
-compatible quadrature set to replace integrals across continuous solid angles
-$\hat{\Omega}$ with quadrature approximations. The multigroup, discrete ordinates ($S_N$
-approximation) form of the 3-D time-dependent neutron transport equation with Legendre polynomial
-expansion of the scattering cross section and spherical harmonic expansion of the flux in the
-scattering source term is:
-%
-\begin{gather}
-  \frac{\partial}{\partial t} \left(\frac{\Psi_{g,d}(\vec{r},t)}{v_g}\right) +
-  \hat{\Omega}_d\cdot\nabla\Psi_{g,d}(\vec{r},t) + \Sigma_{t,g}(\vec{r},t)\Psi_{g,d}(\vec{r},t) =
-  \nonumber \\
-  \sum^G_{g'=1}\sum^L_{l=0} \frac{2l+1}{4\pi}\Sigma^{g'\rightarrow g}_{s,l}(\vec{r},t)\sum^l_{m=-l}
-  \phi_{g',l,m}(\vec{r},t)Y_{l,m}(\hat{\Omega}_d) + \frac{\chi_g(\vec{r},t)}{4\pi}\sum^G_{g'=1}
-  \nu\Sigma_{f,g'}(\vec{r},t) \phi_g(\vec{r},t) + S_{g,d}(\vec{r},t)
-  \label{eq:3d-sn}
+boundary conditions due to 90-degree rotational symmetry about each of the three Cartesian axes.
+
+We use the level-symmetric quadrature sets to approximate integrals of the angular flux over the
+angular domain
+%
+\begin{gather}
+  \phi_g = \int_\mathcal{S} \Psi_{g,d}\hat{\Omega} \approx \sum^{N_d}_{d=1}w_d
+  \Psi_{g,d}), \label{eq:0th-mmt}
+  \shortintertext{where}
+  \Psi_{g,d} = \Psi_g(\hat{\Omega}_d). \nonumber
+\end{gather}
+%
+We calculate the scalar flux $\phi_g$ by taking the zeroth moment with respect to $\hat{\Omega}$ as
+shown in Eq.\ \ref{eq:0th-mmt}. We can also calculate higher moments
+%
+\begin{gather}
+  \phi_{g,l,m} = \int_\mathcal{S} Y_{l,m}(\hat{\Omega})\Psi_g(\hat{\Omega})d\hat{\Omega} \approx
+  \sum^{N_d}_{d=1}w_d Y_{l,m}(\hat{\Omega}_d)\Psi_{g}(\hat{\Omega}_d) \label{eq:nth-mmt}
   \shortintertext{where}
   \begin{align*}
-    \Psi_{g,d}(\vec{r},t) &= \Psi_g(\vec{r},\hat{\Omega}_d,t) \\
-%    g &= \mbox{ neutron energy group index, ranging from 1 to G,} \nonumber \\
-%    \Sigma_{t,g}(x) &= \mbox{ macroscopic total cross section of neutrons in group $g$,} \nonumber \\
-%    G &= \mbox{ total number of energy groups,} \nonumber \\
-%    n &= \mbox{ discrete ordinate index,} \nonumber \\
-    N &= \mbox{ total number of discrete ordinates,} \\
-    l &= \mbox{ Legendre expansion order,} \\
-    L &= \mbox{ highest Legendre expansion order,} \\
-    \Sigma^{g'\rightarrow g}_{s,l}(x) &= \mbox{ $l$-th Legendre moment of the macroscopic
-scattering} \nonumber \\
-    &\ \ \ \ \mbox{ cross section from group $g'$ to $g$,} \\
-    \phi_{g',l,m} &= \mbox{ spherical harmonic expansion of the angular flux of degree $l$ and
-    order $m$,} \\
-    Y_{l,m} &= \mbox{ real basis functions of spherical harmonics of degree $l$ and order $m$.}
-%    \chi_g &= \mbox{ fission neutron spectrum in group $g$,} \nonumber \\
-%    \nu &= \mbox{ neutrons produced per fission reaction,} \nonumber \\
-%    \Sigma_{f,g}(x) &= \mbox{ macroscopic fission cross section of neutrons in group $g$,}
-%    \nonumber \\
-%    k &= \mbox{ neutron multiplication factor.} \nonumber
-%    \phi_{g'}(x) &= \mbox{ scalar neutron flux in group $g$,} \nonumber \\
-%    S_g(x,\mu_n) &= \mbox{ external neutron source in group $g$.} \nonumber
+    Y_{l,m}(\mu,\omega) &=
+    \begin{cases}
+      \sqrt{2}C^{|m|}_l P^{|m|}_l \cos(|m|\omega), & \mbox{ for } 0<m\leq l \\
+      C^0_l P^0_l(\mu), & \mbox{ for } m=0, 0\leq l < \infty \\
+      \sqrt{2} C^{|m|}_l P^{|m|}_l(\mu) \sin(|m|\omega), & -l \leq m < 0
+    \end{cases}, \\
+    \mu &= \mbox{cosine of the polar angle,} \\
+    \omega &= \mbox{the azimuthal angle,} \\
+    C^{|m|}_l &= \sqrt{\frac{\left(l-|m|\right)!}{\left(l+|m|\right)!}}, \\
+    P^m_l &= \mbox{associated Legendre polynomial of degree $l$ and order $m$.}
   \end{align*}
@@
 %
-We can retrieve the neutron scalar flux and current by calculating the 0th and 1st Legendre
-moments, respectively, as follows:
-%
-\begin{align}
-  \phi_g(x) &= \frac{1}{2} \int^1_{-1} d\mu\ \Psi_g(x,\mu) = \frac{1}{2} \sum^N_{n=1} w_n
-\Psi_g(x,\mu_n)
-  \shortintertext{and}
-  J_g(x) &= \frac{1}{2} \int^1_{-1} d\mu\ \mu\Psi_g(x,\mu) = \frac{1}{2} \sum^N_{n=1} w_n
-\mu_n \Psi_g(x,\mu_n)
-\end{align}
-
-The 1-D form of the multigroup $k$-eigenvalue neutron diffusion equations (Eq.\ \ref{eq:mg-diff})
-is:
-%
-\begin{align}
-  -\frac{d}{dx} D_g(x) \frac{d}{dx} \phi_g(x) + \Sigma_{t,g}(x) \phi_g(x) &= \sum^G_{g'=1}\left[
-  \Sigma_s^{g'\rightarrow g}(x)\phi_{g'}(x) + \chi_g\frac{\nu\Sigma_{f,g'}(x)}{k}
-  \phi_{g'}(x)\right] + S_g(x)
-  \label{eq:1d-diff}
-  \shortintertext{where}
-    D_g(x) &= \frac{1}{3 \Sigma_{tr}(x)} \nonumber \\
-           &= \mbox{ isotropic neutron diffusion coefficient for group }g. \nonumber
-\end{align}
+for anisotropic scattering treatment.
+
+\subsubsection{Final Forms of the \gls{SAAF} $S_N$ Equations}
+
+Finally, we present the expanded forms of each term in the \gls{SAAF} formulation of the multigroup
+$S_N$ neutron transport equations. With the angular discretization scheme laid out, we define the
+inner products
+%
+\begin{gather}
+  \left(\bm{a},\bm{b}\right)_\mathcal{D} \equiv \sum^G_{g=1} \int_\mathcal{D}d\vec{r}
+  \sum_{i\in E_e}a_{g,i}b_i(\vec{r})\sum_{j\in E_e}b_{g,j}b_j(\vec{r}), \\
+  \left(\bm{a},\bm{b}\right)_{\partial\mathcal{D}} \equiv \sum^G_{g=1}
+  \sum_{s\in\partial\mathcal{D}}\int_s d\vec{r}\sum_{i\in E_s}a_{g,i}b_i(\vec{r})\sum_{j\in E_s}
+  b_{g,j}b_j(\vec{r}),
+\end{gather}
+%
+involving integrals over the spatial domain to explicitly show the $S_N$ angular discretization.
+
+Time derivative term:
+%
+\begin{gather}
+  \left(\left(\mathbb{I}+\bm{\tau}\mathbb{L}_1\right)\bm{\Psi}^*,
+  \frac{\partial}{\partial t}\left(\frac{\bm{\Psi}}{\bm{v}}\right)\right) =
+  \sum^G_{g=1}\sum^{N_d}_{d=1}w_d\left(\Psi^*_{g,d}+\tau_g\hat{\Omega}_d\cdot\Psi^*_{g,d},
+  \frac{\Psi_{g,d}}{v_g}\right)_\mathcal{D}
+\end{gather}
+%
+Streaming term:
+%
+\begin{gather}
+  \left(\mathbb{L}_1\bm{\Psi}^*,
+  \left(\bm{\tau}\mathbb{L}_1-\mathbb{I}+\bm{\tau}\mathbb{L}_2\right)\bm{\Psi}\right) =
+  \sum^G_{g=1}\sum^{N_d}_{d=1}w_d\left(\hat{\Omega}_d\cdot\Psi^*_{g,d},\tau_g\hat{\Omega}\cdot
+  \nabla\Psi_{g,d}-(1-\tau_g\Sigma_{t,g})\Psi_{g,d}\right)_\mathcal{D}
+\end{gather}
+%
+Collision term:
+%
+\begin{gather}
+  \left(\mathbb{L}_2\bm{\Psi}^*,\bm{\Psi}\right) =
+  \sum^G_{g=1}\sum^{N_d}_{d=1}w_d\left(\Psi^*_{g,d},\Sigma_{t,g}\Psi_{g,d}\right)_\mathcal{D}
+\end{gather}
+%
+Scattering term:
+%
+\begin{gather}
+  \left(\left(\mathbb{I}+\bm{\tau}\mathbb{L}_1\right)\bm{\Psi}^*,\mathbb{S}\bm{\Psi}\right) =
+  \sum^G_{g=1}\sum^{N_d}_{d=1}w_d\left(\Psi^*_{g,d}+\tau_g\hat{\Omega}_d\cdot\Psi^*_{g,d},
+    \sum^G_{g'=1}\sum^L_{l=0}\Sigma^{g'\rightarrow g}_{s,l}\sum^l_{m=-l}
 

### 2024-06-12 lit.tex
Remove large spacing after al.

@@
 without regard to potentially destabilizing modes introduced by the
-multiphysics coupling \cite{keyes_multiphysics_2013}. In particular, Lindsay et al.
+multiphysics coupling \cite{keyes_multiphysics_2013}. In particular, Lindsay et al.\
 \cite{lindsay_introduction_2018} demonstrated the non-convergence of a multiphysics \gls{MSR}
@@
 given the strong coupling between neutronics and thermal-hydraulics.
-Aufiero et al. \cite{aufiero_development_2014} demonstrated a loose coupling
+Aufiero et al.\ \cite{aufiero_development_2014} demonstrated a loose coupling
 approach that failed to reproduce the expected increase in reactor power
@@
 
-In 2007, Krepel et al. \cite{krepel_dyn3d-msr_2007} extended the
+In 2007, Krepel et al.\ \cite{krepel_dyn3d-msr_2007} extended the
 \gls{LWR} nodal diffusion code DYN3D to account for \gls{DNP} drift and fission energy deposition
@@
 
-At the \gls{TUD}, Kophazi et al.
+At the \gls{TUD}, Kophazi et al.\
 \cite{kophazi_development_2009} coupled their in-house 3-D neutron diffusion software DALTON
@@
 homogeneous region into a Cartesian mesh and modeled a more accurate representation of the oblong
-fuel channels in the graphite core matrix instead of the circular shape by Krepel et al.
+fuel channels in the graphite core matrix instead of the circular shape by Krepel et al.\
 \cite{krepel_dyn3d-msr_2007}.
 
-Both efforts by Krepel et al. and Kophazi et al. showed good
+Both efforts by Krepel et al.\ and Kophazi et al.\ showed good
 agreement with \gls{MSRE} experimental data in their respective validation tests.
@@
 turbulence effects in the \gls{MSFR} during steady-state operation and transient scenarios. In a
-later effort from the same institute, Tiberga et al. \cite{tiberga_discontinuous_2019} coupled
+later effort from the same institute, Tiberga et al.\ \cite{tiberga_discontinuous_2019} coupled
 PHANTOM-$S_N$ and DGFlows for their participation in the CNRS benchmark study
@@
 code-to-code verification of \gls{MSR} multiphysics software \cite{aufiero_testing_2018}. Tiberga
-et al. showed that their coupled PHANTOM-$S_N$/DGFlows software was consistent with
+et al.\ showed that their coupled PHANTOM-$S_N$/DGFlows software was consistent with
 OpenFOAM-based \gls{MSR} simulation tools. Further details on their study and the CNRS benchmark
@@
 
-More recently, Jaradat et al. \cite{jaradat_development_2021}
+More recently, Jaradat et al.\ \cite{jaradat_development_2021}
 extended the 3-D $P_1$ and $SP_3$ nodal transport code PROTEUS-NODAL to support \gls{DNP} drift.
-Yang et al. \cite{yang_development_2022} then developed a coupled simulation tool of PROTEUS-NODAL
+Yang et al.\ \cite{yang_development_2022} then developed a coupled simulation tool of PROTEUS-NODAL
 and \gls{SAM}. They modeled the \gls{MSFR} on a 2-D axisymmetric geometry
@@
 
-Laureau et al.
+Laureau et al.\
 \cite{laureau_transient_2017} developed a different technique called the
@@
 incorporate temperature-induced salt expansion and Doppler effects on the neutron cross sections
-and, ultimately, the neutron flux. Laureau et al. opted to subcycle the neutronics calculations by
+and, ultimately, the neutron flux. Laureau et al.\ opted to subcycle the neutronics calculations by
 setting smaller timesteps than the thermal-hydraulics calculations
@@
 calculations several times per timestep until convergence is reached.
-Blanco demonstrated the Serpent-OpenFOAM tool on the CNRS benchmark specified by Tiberga et al.
+Blanco demonstrated the Serpent-OpenFOAM tool on the CNRS benchmark specified by Tiberga et al.\
 \cite{tiberga_results_2020}.
@@
 handle all multiphysics calculations and data transfer internally. Among earlier efforts, Nicolino
-et al. \cite{nicolino_coupled_2008} and Zhang et al. \cite{zhang_development_2009} recognized the
+et al.\ \cite{nicolino_coupled_2008} and Zhang et al.\ \cite{zhang_development_2009} recognized the
 need for more robust multiphysics coupling, acceleration techniques and internal data transfers
 to reduce execution times. They each independently developed unnamed multiphysics simulation tools
-and demonstrated their tools with non-moderated \gls{MSR} designs. Nicolino et al. presented one of
-the first \gls{MSR} models with multi-dimensional flow, governed by the Navier-Stokes equation with
-the Boussinesq approximation. They performed steady-state and transient calculations on a 2-D
-axisymmetric model of the \gls{MOSART} reactor \cite{ignatiev_molten_2014}. Zhang et al. also
+and demonstrated their tools with non-moderated \gls{MSR} designs. Nicolino et al.\ presented one
+of the first \gls{MSR} models with multi-dimensional flow, governed by the Navier-Stokes equation
+with the Boussinesq approximation. They performed steady-state and transient calculations on a 2-D
+axisymmetric model of the \gls{MOSART} reactor \cite{ignatiev_molten_2014}. Zhang et al.\ also
 developed a similar 2-D axisymmetric solver, but they incorporated a $k$-$\epsilon$ turbulence
@@
 Both studies implemented the neutron diffusion model for neutronics modeling.
-Later, Li et al. \cite{li_transient_2015} demonstrated the
+Later, Li et al.\ \cite{li_transient_2015} demonstrated the
 steady-state and transient analysis capabilities of COUPLE, a neutronics and
@@
 Other institutes have dedicated significant development work towards
-OpenFOAM-based \gls{MSR} simulation tools. Aufiero et al.
+OpenFOAM-based \gls{MSR} simulation tools. Aufiero et al.\
 \cite{aufiero_development_2014} first introduced an OpenFOAM model developed
@@
 the \gls{MSFR}. GeN-Foam is another OpenFOAM-based tool developed by Fiorina
-et al. \cite{fiorina_gen-foam_2015} as a general reactor multiphysics solver
+et al.\ \cite{fiorina_gen-foam_2015} as a general reactor multiphysics solver
 applicable to \glspl{MSR} and other reactor types. GeN-Foam features neutron
@@
 and thermo-mechanical modeling for reactor expansion effects. Using GeN-Foam,
-Altahhan et al. \cite{altahhan_preliminary_2020} developed and optimized a
+Altahhan et al.\ \cite{altahhan_preliminary_2020} developed and optimized a
 liquid-fuel \gls{MSR} design, while Shi \& Fratoni \cite{shi_gen-foam_2021}
@@
 Moltres benefits from the highly-integrated cross-compatibility
-within the ecosystem of \gls{MOOSE}-based applications. Abou-Jaoude et al.
+within the ecosystem of \gls{MOOSE}-based applications. Abou-Jaoude et al.\
 \cite{abou-jaoude_coupled_2020} coupled Griffin with Pronghorn, another
@@
 demonstrate several steady-state \gls{MSR} simulation capabilities defined in
-the CNRS benchmark. Lindsay et al.
+the CNRS benchmark. Lindsay et al.\
 \cite{lindsay_introduction_2018} first demonstrated Moltres' \gls{MSR} modeling
@@
 \gls{DNP} drift and the strong coupling between neutron flux and temperature advection-diffusion.
-Delpech et al. \cite{delpech_benchmark_2003} published one of the first collaborative \gls{VV} work
+Delpech et al.\ \cite{delpech_benchmark_2003} published one of the first collaborative \gls{VV} work
 for \gls{MSR} multiphysics modeling. Collaborators
@@
 
-As mentioned in Subsection \ref{sec:msr-tools}, Tiberga et al. \cite{tiberga_results_2020}
+As mentioned in Subsection \ref{sec:msr-tools}, Tiberga et al.\ \cite{tiberga_results_2020}
 published the CNRS benchmark to verify \gls{MSR} simulation tools designed for
@@
 studying the effects of natural circulation and lid-driven flow on the reactivity and power output.
-Aside from the problem specifications, Tiberga et al. also published the associated group
+Aside from the problem specifications, Tiberga et al.\ also published the associated group
 constant cross section data required by deterministic neutronics solvers to perform neutronics
@@
 
-Neutronic benchmark studies of the \gls{MSRE} and the \gls{MSFR} by Fratoni et al.
-\cite{fratoni_molten_2020} and Brovchenko et al. \cite{brovchenko_neutronic_2019} measured the
+Neutronic benchmark studies of the \gls{MSRE} and the \gls{MSFR} by Fratoni et al.\
+\cite{fratoni_molten_2020} and Brovchenko et al.\ \cite{brovchenko_neutronic_2019} measured the
 delayed neutron losses due to the decay of \glspl{DNP} flowing out of the active core region.
-Fratoni et al. sought to establish a standard validation platform for \gls{MSR} neutronics
+Fratoni et al.\ sought to establish a standard validation platform for \gls{MSR} neutronics
 simulation tools with \gls{MSRE} experimental data for inclusion in the \gls{IRPhEP} handbook.
 They characterized and validated a model of the \gls{MSRE} in the Monte Carlo particle transport
-code Serpent. The \gls{MSFR} neutronics benchmark by Brovchenko et al. featured results
+code Serpent. The \gls{MSFR} neutronics benchmark by Brovchenko et al.\ featured results
 from multiple \gls{MSR} simulation tools by several collaborators. Their assessment found that
@@
 While these publications have addressed significant technical gaps, there is room to develop
-further \gls{VV} benchmarks for multiphysics \gls{MSR} modeling. For instance, the
-CNRS benchmark does not assess the loss of delayed neutrons due to the decay of \glspl{DNP} flowing
-out of the active core region. This phenomenon is important as the delayed neutron fraction in the
-core directly impacts the transient power response in unprotected accident scenarios. Meanwhile,
-the neutronic benchmark studies by Fratoni et al. \cite{fratoni_molten_2020} and Brovchenko et al.
+further \gls{VV} benchmarks for multiphysics \gls{MSR} modeling. For instance, the CNRS benchmark
+does not assess the loss of delayed neutrons due to the decay of \glspl{DNP} flowing out of the
+active core region. This phenomenon is important as the delayed neutron fraction in the core
+directly impacts the transient power response in unprotected accident scenarios. Meanwhile, the
+neutronic benchmark studies by Fratoni et al.\ \cite{fratoni_molten_2020} and Brovchenko et al.\
 \cite{brovchenko_neutronic_2019} did not provide the standardized group constant data required by
@@
 
-Podila et al. \cite{podila_cfd_2019} performed \gls{CFD} simulations of the
+Podila et al.\ \cite{podila_cfd_2019} performed \gls{CFD} simulations of the
 \gls{MSRE} core with six different turbulence models, namely a Spalart-Allmaras
@@
 \cite{aufiero_development_2014}, and \gls{TUD}'s in-house codes
-\cite{fiorina_modelling_2014,tiberga_results_2020}. Fiorina et al.
+\cite{fiorina_modelling_2014,tiberga_results_2020}. Fiorina et al.\
 \cite{fiorina_modelling_2014} compared the flow distribution from both models
@@
 In a recent study by
-Laureau et al. \cite{laureau_unmoderated_2022}, the authors applied a \gls{DES} model to model
+Laureau et al.\ \cite{laureau_unmoderated_2022}, the authors applied a \gls{DES} model to model
 eddies in the \gls{MSFR}. Given the high computational costs required to obtain reasonably small
@@
 load-following operation by adjusting pump speeds and the temperature of the intermediate coolant
-loop. Laureau et al. \cite{laureau_transient_2017} demonstrated simple power ramp-up and ramp-down
+loop. Laureau et al.\ \cite{laureau_transient_2017} demonstrated simple power ramp-up and ramp-down
 with the \gls{MSFR} by varying the temperature of the secondary coolant. \gls{MSR}
@@
 highly neutron absorbing regions \cite{kophazi_development_2009, jaradat_development_2021,
-yang_development_2022}. Kophazi et al. \cite{kophazi_development_2009} modeled annular control rods
+yang_development_2022}. Kophazi et al.\ \cite{kophazi_development_2009} modeled annular control rods
 as homogenized hexahedrons in their 3-D Cartesian geometry of the \gls{MSRE}. They
@@
 \ref{sec:msr-tools}. To reduce errors introduced by the neutron diffusion method, they imposed
-simple albedo boundary conditions for the thermal neutron group. Jaradat et al.
-\cite{jaradat_development_2021} and Yang et al. \cite{yang_development_2022} modeled the \gls{MSRE}
-control rods as homogenized wedges in their R-$\theta$-Z mesh due to the constraints of the nodal
-neutronics solver. Cui et al. \cite{cui_development_2021} also homogenized the control rods as
-regular hexahedral nodes in accordance with their nodal solver in Cartesian geometry. A disadvantage
-of homogenization-based methods is that it removes some of the heterogeneity in the flux and
-temperature. The heterogeneous temperature distribution is especially important in \glspl{MSR} due
-to the combined fuel-coolant and the positive temperature reactivity feedback observed
-in graphite under certain conditions \cite{mathieu_thorium_2006}.
+simple albedo boundary conditions for the thermal neutron group. Jaradat et al.\
+\cite{jaradat_development_2021} and Yang et al.\ \cite{yang_development_2022} modeled the
+\gls{MSRE} control rods as homogenized wedges in their R-$\theta$-Z mesh due to the constraints of
+the nodal neutronics solver. Cui et al.\ \cite{cui_development_2021} also homogenized the control
+rods as regular hexahedral nodes in accordance with their nodal solver in Cartesian geometry. A
+disadvantage of homogenization-based methods is that it removes some of the heterogeneity in the
+flux and temperature. The heterogeneous temperature distribution is especially important in
+\glspl{MSR} due to the combined fuel-coolant and the positive temperature reactivity feedback
+observed in graphite under certain conditions \cite{mathieu_thorium_2006}.
 
@@
 neutron, in a given energy range, will undergo the associated interaction per unit distance
-traveled in the material. Group constants also broadly include neutron diffusion
+traveled in the material.\ Group constants also broadly include neutron diffusion
 coefficients/constants and \gls{DNP} data. While advanced coarse-mesh and nodal
@@
 \gls{MECS}, here is an example of applying the \gls{MECS} on a control rod in x-y geometry (Figure
-\ref{fig:mecs-geometry}) in the CITATION code as provided by Fen et al. \cite{fen_modelling_1992}.
+\ref{fig:mecs-geometry}) in the CITATION code as provided by Fen et al.\ \cite{fen_modelling_1992}.
 %
@@
         region and the adjacent non-absorber region, respectively.
-        Retrieved from Fen et al. \cite{fen_modelling_1992}.}
+        Retrieved from Fen et al.\ \cite{fen_modelling_1992}.}
     \label{fig:mecs-geometry}
@@
 diffusion solver on the bottom. The four adjacent nodes in the Cartesian
-geometry and their corresponding flux values are pairwise identical.
+geometry and their corresponding flux values are pairwise identical.\
 
@@
 
-Fen et al. \cite{fen_modelling_1992} later provided an updated formulation by assuming that the
+Fen et al.\ \cite{fen_modelling_1992} later provided an updated formulation by assuming that the
 cross sections are accurate and instead determined the equivalent diffusion coefficient from Eq.
@@
 
-Fen et al. \cite{fen_modelling_1992} developed the \gls{RMM} which generates modified boundary
+Fen et al.\ \cite{fen_modelling_1992} developed the \gls{RMM} which generates modified boundary
 conditions from response functions to treat absorber nodes in whole-core diffusion calculations.
@@
 
-Rahnema et al. \cite{rahnema_advanced_2011} later developed an \gls{IDT} method that embeds the
+Rahnema et al.\ \cite{rahnema_advanced_2011} later developed an \gls{IDT} method that embeds the
 transport correction for absorber regions in the full-core nodal diffusion calculation. Instead of
@@
 functions and Legendre expansions of the angular flux.
-Gross et al. \cite{gross_high-accuracy_2020} extended the derivation of the transport operators to
+Gross et al.\ \cite{gross_high-accuracy_2020} extended the derivation of the transport operators to
 handle 1-D heterogeneous problems in the form of fuel assemblies with fuel, water, and fuel+absorber
-regions. Tomatis et al. \cite{tomatis_ronen_2021} developed new numerical implementations for 1-D
+regions. Tomatis et al.\ \cite{tomatis_ronen_2021} developed new numerical implementations for 1-D
 slab, cylindrical, and spherical geometries by employing probabilistic treatments from the
@@
 
-Two recent works by Anistratov \& Stehle \cite{anistratov_computational_2012} and Stehle et al.
+Two recent works by Anistratov \& Stehle \cite{anistratov_computational_2012} and Stehle et al.\
 \cite{stehle_hybrid_2014} describe another approach that is distinct from the prior approach in
@@
 evaluate the 0th-order scattering source term and accelerate the high-order calculations for
-$\Psi^{(s+1/2)}$. Stehle et al. \cite{stehle_hybrid_2014} demonstrated their hybrid method on two
+$\Psi^{(s+1/2)}$. Stehle et al.\ \cite{stehle_hybrid_2014} demonstrated their hybrid method on two
 2-D, single-group test problems. Both problems consisted of two regions with idealized macroscopic

### 2024-06-15 hybrid.tex
hybrid method edits

@@
-Section \ref{sec:summary-nts-mtds} highlighted the poor performance of neutron diffusion
+Section \ref{sec:summary-nts-mtds} highlights the poor performance of neutron diffusion
 methods for calculating neutron fluxes near control rods. Strong neutron absorption in the control
@@
 
-In this chapter, I propose a hybrid neutronics method for improving control rod modeling in neutron
-diffusion solvers without spatial homogenization. In essence, the hybrid method involves applying
+In this chapter, I propose a hybrid $S_N$-diffusion neutronics method to improve control rod
+modeling in neutron diffusion solvers without spatial homogenization. In essence, the hybrid
+method is an iterative method that applies
 the $S_N$ discrete ordinates neutron transport method on subregions containing the control rod to
 obtain pointwise transport corrections for the diffusion method on the subregions.
-
-Section \ref{sec:theory} discusses the theoretical background for the hybrid $S_N$-diffusion
-method. Section \ref{sec:implementation} provides numerical implementation details of
-the hybrid method and its components. Sections \ref{sec:test-case} and \ref{sec:sim-param} describe
-the 1-D test cases and several simulation parameters in that context. Section
-\ref{sec:prelim-results} discusses the results of the hybrid method applied to the 1-D test cases
-with comparisons to higher-fidelity Monte Carlo and $S_N$ neutron transport methods. Lastly,
-Section \ref{sec:hybrid-summary} summarizes the key findings in this chapter.
-
-\section{Theory} \label{sec:theory}
-
-The proposed hybrid $S_N$-diffusion method is an iterative method to improve the
-accuracy of neutron diffusion solutions in reactor systems with highly anisotropic flux regions
-which I will use to model control rods. This section presents the mathematical derivation for the
-\gls{SAAF} formulation of the $S_N$ equations, the neutron diffusion equations, the drift
-transport correction term, and the computational algorithm for the hybrid method.
-
-\subsection{\gls{SAAF} Formulation of the $S_N$ Method}
-
-\subsubsection{Multigroup Neutron Transport Equations}
+This chapter presents the mathematical derivation for the
+\gls{SAAF} formulation of the $S_N$ equations and the drift transport correction term for the
+neutron diffusion equations, and the computational algorithm for the hybrid method.
+
+% Section \ref{sec:theory} discusses the theoretical background for the hybrid $S_N$-diffusion
+% method. Section \ref{sec:implementation} provides numerical implementation details of
+% the hybrid method and its components. Sections \ref{sec:test-case} and \ref{sec:sim-param} describe
+% the 1-D test cases and several simulation parameters in that context. Section
+% \ref{sec:prelim-results} discusses the results of the hybrid method applied to the 1-D test cases
+% with comparisons to higher-fidelity Monte Carlo and $S_N$ neutron transport methods. Lastly,
+% Section \ref{sec:hybrid-summary} summarizes the key findings in this chapter.
+
+% The proposed hybrid $S_N$-diffusion method is an iterative method to improve the
+% accuracy of neutron diffusion solutions in reactor systems with highly anisotropic flux regions
+% which I will use to model control rods.
+
+\section{\gls{SAAF} Formulation of the $S_N$ Method}
+
+\subsection{Multigroup Neutron Transport Equations}
 
@@
 \begin{gather}
-  \bm{\Psi} =
+  \bm{\Psi} \equiv
   \begin{bmatrix}
@@
   \end{bmatrix},
-  \bm{\phi} = \int_S \bm{\Psi}d\hat{\Omega} =
+  \bm{\phi} \equiv \int_S \bm{\Psi}d\hat{\Omega} \equiv
   \begin{bmatrix}
@@
   \end{bmatrix},
-  \bm{\frac{\Psi}{v}}
+  \bm{\frac{\Psi}{v}} \equiv
   \begin{bmatrix}
@@
   \end{bmatrix},
-  \bm{Q}^{\text{ext}} =
+  \bm{C} \equiv
+  \begin{bmatrix}
+    C_1 \\
+    C_2 \\
+    \vdots \\
+    C_G
+  \end{bmatrix},
+  \bm{Q}^{\text{ext}} \equiv
   \begin{bmatrix}
@@
     Q^\text{ext}_G
-  \end{bmatrix}, \nonumber \\
-  \bm{Q}_f = \frac{1}{4\pi}\bm{Q}_{f,0} =
-  \begin{bmatrix}
-    \frac{1}{4\pi}\chi_{p,1}(1-\beta)\sum^G_{g'=1} \nu\Sigma_{f,g'} \phi_{g'} +
-    \frac{1}{4\pi}\sum^I_{i=1}\chi_{d,1} \lambda_i C_i \\
-    \frac{1}{4\pi}\chi_{p,2}(1-\beta)\sum^G_{g'=1} \nu\Sigma_{f,g'} \phi_{g'} +
-    \frac{1}{4\pi}\sum^I_{i=1}\chi_{d,2} \lambda_i C_i \\
-    \vdots \\
-    \frac{1}{4\pi}\chi_{p,G}(1-\beta)\sum^G_{g'=1} \nu\Sigma_{f,g'} \phi_{g'} +
-    \frac{1}{4\pi}\sum^I_{i=1}\chi_{d,G} \lambda_i C_i
-  \end{bmatrix},
-  \bm{Q} = \bm{Q}_f + \bm{Q}^\text{ext}. \nonumber
+  \end{bmatrix}, \nonumber
+%  \bm{Q}_f = \frac{1}{4\pi}\bm{Q}_{f,0} =
+%  \begin{bmatrix}
+%    \frac{1}{4\pi}\chi_{p,1}(1-\beta)\sum^G_{g'=1} \nu\Sigma_{f,g'} \phi_{g'} +
+%    \frac{1}{4\pi}\sum^I_{i=1}\chi_{d,1} \lambda_i C_i \\
+%    \frac{1}{4\pi}\chi_{p,2}(1-\beta)\sum^G_{g'=1} \nu\Sigma_{f,g'} \phi_{g'} +
+%    \frac{1}{4\pi}\sum^I_{i=1}\chi_{d,2} \lambda_i C_i \\
+%    \vdots \\
+%    \frac{1}{4\pi}\chi_{p,G}(1-\beta)\sum^G_{g'=1} \nu\Sigma_{f,g'} \phi_{g'} +
+%    \frac{1}{4\pi}\sum^I_{i=1}\chi_{d,G} \lambda_i C_i
+%  \end{bmatrix},
+%  \bm{Q} = \bm{Q}_f + \bm{Q}^\text{ext}. \nonumber
 \end{gather}
@@
 \begin{gather}
-  \mathbb{L}_1\bm{\Psi} =
+  \mathbb{L}_1\bm{\Psi} \equiv
   \begin{bmatrix}
@@
   \end{bmatrix},
-  \mathbb{L}_2\bm{\Psi} =
+  \mathbb{L}_2\bm{\Psi} \equiv
   \begin{bmatrix}
@@
   \end{bmatrix},
-  \mathbb{L}\bm{\Psi} = \mathbb{L}_1\bm{\Psi} + \mathbb{L}_2\bm{\Psi}, \nonumber \\
-  \mathbb{S}\bm{\Psi} =
+  \mathbb{L}\bm{\Psi} \equiv \mathbb{L}_1\bm{\Psi} + \mathbb{L}_2\bm{\Psi}, \nonumber \\
+  \mathbb{S}\bm{\Psi} \equiv
   \begin{bmatrix}
@@
   \end{bmatrix},
-%  \mathbb{F}_p\bm{\Psi} = \frac{1}{4\pi} \mathbb{F}_{p,0}\bm{\phi} =
-%  \begin{bmatrix}
-%    \frac{1}{4\pi}\chi_{p,1}(1-\beta)\sum^G_{g'=1}\nu\Sigma_{f,g'}\phi_{g'} \\
-%    \frac{1}{4\pi}\chi_{p,2}(1-\beta)\sum^G_{g'=1}\nu\Sigma_{f,g'}\phi_{g'} \\
-%    \vdots \\
-%    \frac{1}{4\pi}\chi_{p,G}(1-\beta)\sum^G_{g'=1}\nu\Sigma_{f,g'}\phi_{g'}
-%  \end{bmatrix},
-  \mathbb{B}\bm{\Psi} =
+  \mathbb{B}\bm{\Psi} \equiv
   \begin{bmatrix}
@@
     \alpha^s_G\Psi_G(\hat{\Omega}_r)
+  \end{bmatrix}, \nonumber \\
+  \mathbb{F}\bm{\Psi} \equiv \frac{1}{4\pi}\mathbb{F}_0\bm{\Psi} \equiv
+  \begin{bmatrix}
+    \frac{1}{4\pi}\chi_{p,1}(1-\beta)\sum^G_{g'=1}\nu\Sigma_{f,g'}\phi_{g'} \\
+    \frac{1}{4\pi}\chi_{p,2}(1-\beta)\sum^G_{g'=1}\nu\Sigma_{f,g'}\phi_{g'} \\
+    \vdots \\
+    \frac{1}{4\pi}\chi_{p,G}(1-\beta)\sum^G_{g'=1}\nu\Sigma_{f,g'}\phi_{g'}
+  \end{bmatrix},
+  \mathbb{C}\bm{C} \equiv \frac{1}{4\pi}\mathbb{C}_0\bm{C} \equiv
+  \begin{bmatrix}
+    \frac{1}{4\pi}\sum^I_{i=1}\chi_{d,1} \lambda_i C_i \\
+    \frac{1}{4\pi}\sum^I_{i=1}\chi_{d,2} \lambda_i C_i \\
+    \vdots \\
+    \frac{1}{4\pi}\sum^I_{i=1}\chi_{d,G} \lambda_i C_i
   \end{bmatrix}. \nonumber
@@
   \frac{\partial}{\partial t} \left(\frac{\bm{\Psi}}{\bm{v}}\right)+\mathbb{L}\bm{\Psi}
-  = \mathbb{S}\bm{\Psi} + \bm{Q}_f + \bm{Q}^\text{ext} \label{eq:nte-vec}
+  = \mathbb{S}\bm{\Psi} + \bm{Q}, \label{eq:nte-vec}
 \end{gather}
@@
 \begin{gather}
-  \bm{\Psi} = \bm{\Psi}^\text{inc} + \mathbb{B}\bm{\Psi}
-\end{gather}
-
-\subsubsection{Weak Form of the Multigroup Neutron Transport Equations}
+  \bm{\Psi} = \bm{\Psi}^\text{inc} + \mathbb{B}\bm{\Psi},
+  \shortintertext{where}
+  \begin{align*}
+    \bm{Q} &= \bm{Q}_f + \bm{Q}^\text{ext}, & \\
+    \bm{Q}_f &= \mathbb{F}\bm{\Psi} + \mathbb{C}\bm{C}. &
+  \end{align*}
+\end{gather}
+
+\subsection{Weak Form of the Multigroup Neutron Transport Equations}
 
@@
 
-\subsubsection{\gls{SAAF} Formulation}
+\subsection{\gls{SAAF} Formulation}
 
@@
 
-\subsubsection{Spatial Discretization and Void Treatment}
+\subsection{Spatial Discretization and Void Treatment}
 
@@
 
-\subsubsection{Angular Discretization}
+\subsection{Angular Discretization}
 
@@
 exist for the choice of the angular quadrature set
-$\{\hat{\Omega}_d,w_d | d=1,\dots,N_d\}$.
+$\{\hat{\Omega}_d,w_d\ |\ d=1,\dots,N_d\}$.
 For this work, we chose the commonly used level-symmetric quadrature set
@@
   \phi_g = \int_\mathcal{S} \Psi_{g,d}\hat{\Omega} \approx \sum^{N_d}_{d=1}w_d
-  \Psi_{g,d}), \label{eq:0th-mmt}
+  \Psi_{g,d}, \label{eq:0th-mmt}
   \shortintertext{where}
-  \Psi_{g,d} = \Psi_g(\hat{\Omega}_d). \nonumber
+  \Psi_{g,d} = \Psi_g(\hat{\Omega}_d), \nonumber \\
+  w = \sum^{N_d}_{d=1} w_d\ (= 8 \mbox{ for 3-D level-symmetric set}). \nonumber
 \end{gather}
@@
     \end{cases}, \\
-    \mu &= \mbox{cosine of the polar angle,} \\
-    \omega &= \mbox{the azimuthal angle,} \\
+    \mu &= \mbox{cosine of polar angle,} \\
+    \omega &= \mbox{azimuthal angle,} \\
     C^{|m|}_l &= \sqrt{\frac{\left(l-|m|\right)!}{\left(l+|m|\right)!}}, \\
@@
 
-\subsubsection{Final Forms of the \gls{SAAF} $S_N$ Equations}
+\subsection{Final Forms of the \gls{SAAF} $S_N$ Equations}
 
@@
 
-Time derivative term:
+\noindent Time derivative term:
 %
@@
   \left(\bm{\tau}\mathbb{L}_1-\mathbb{I}+\bm{\tau}\mathbb{L}_2\right)\bm{\Psi}\right) =
-  \sum^G_{g=1}\sum^{N_d}_{d=1}w_d\left(\hat{\Omega}_d\cdot\Psi^*_{g,d},\tau_g\hat{\Omega}\cdot
-  \nabla\Psi_{g,d}-(1-\tau_g\Sigma_{t,g})\Psi_{g,d}\right)_\mathcal{D}
+  \sum^G_{g=1}\sum^{N_d}_{d=1}w_d\left(\hat{\Omega}_d\cdot\nabla\Psi^*_{g,d},\tau_g\hat{\Omega}
+  \cdot\nabla\Psi_{g,d}-(1-\tau_g\Sigma_{t,g})\Psi_{g,d}\right)_\mathcal{D}
 \end{gather}
@@
   \left(\left(\mathbb{I}+\bm{\tau}\mathbb{L}_1\right)\bm{\Psi}^*,\mathbb{S}\bm{\Psi}\right) =
-  \sum^G_{g=1}\sum^{N_d}_{d=1}w_d\left(\Psi^*_{g,d}+\tau_g\hat{\Omega}_d\cdot\Psi^*_{g,d},
+  \sum^G_{g=1}\sum^{N_d}_{d=1}w_d\left(\Psi^*_{g,d}+\tau_g\hat{\Omega}_d\cdot\nabla\Psi^*_{g,d},
   \sum^G_{g'=1}\sum^L_{l=0}\Sigma^{g'\rightarrow g}_{s,l}\sum^l_{m=-l}
+  \frac{2l+1}{w}Y_{l,m}(\hat{\Omega}_d)\phi_{g',l,m}\right)_\mathcal{D}
+\end{gather}
+%
+Prompt fission source term:
+%
+\begin{gather}
+  \left(\left(\mathbb{I}+\bm{\tau}\mathbb{L}_1\right)\bm{\Psi}^*,\mathbb{F}\bm{\Psi}\right) =
+  \sum^G_{g=1}\sum^{N_d}_{d=1}w_d\left(\Psi^*_{g,d}+\tau_g\hat{\Omega}_d\cdot\nabla\Psi^*_{g,d},
+  \frac{1}{w}\bar{\chi}_g\sum^G_{g'=1}\nu\Sigma_{f,g'}\phi_{g'}\right)_\mathcal{D}
+\end{gather}
+%
+where $\bar{\chi}_g=\chi_{p,g}\left(1-\beta\right)$ for time-dependent problems with \glspl{DNP}
+and $\bar{\chi}_g=\chi_{g}$ for steady-state problems without \glspl{DNP}.
+
+\noindent Delayed neutron source term:
+%
+\begin{gather}
+  \left(\left(\mathbb{I}+\bm{\tau}\mathbb{L}_1\right)\bm{\Psi}^*,\mathbb{C}\bm{C}\right) =
+  \sum^G_{g=1}\sum^{N_d}_{d=1}w_d\left(\Psi^*_{g,d}+\tau_g\hat{\Omega}_d\cdot\nabla\Psi^*_{g,d},
+  \frac{1}{w}\sum ^I_{i=1}\chi_{d,g}\lambda_i C_i\right)_\mathcal{D}
+\end{gather}
+%
+We scale both prompt fission and delayed neutron source terms by the inverse of the
+multiplication factor $\frac{1}{k}$ for $k$-eigenvalue problems.
+
+\noindent Vacuum boundary term:
+%
+\begin{gather}
+  \langle\bm{\Psi}^*,\bm{\Psi}\rangle^+ - \langle\bm{\Psi}^*,\bm{\Psi}^\text{inc}\rangle^- =
+  \begin{cases}
+    \sum^G_{g=1}\sum^{N_d}_{d=1}w_d\left(\Psi^*_{g,d},
+    \hat{\Omega}_d\cdot\hat{n}_b\Psi_{g,d}\right),
+    & \hat{\Omega}\cdot\hat{n}_b>0,\vec{r}\in\partial\mathcal{D}_s \\
+    \sum^G_{g=1}\sum^{N_d}_{d=1}w_d\left(\Psi^*_{g,d},
+    \hat{\Omega}_d\cdot\hat{n}_b\Psi^\text{inc}_{g,d}\right),
+    & \hat{\Omega}\cdot\hat{n}_b<0,\vec{r}\in\partial\mathcal{D}_s
+  \end{cases},
+\end{gather}
+%
+where $\Psi^\text{inc}_{g,d}=0$ for a vacuum boundary and positive for a boundary source.
+
+\noindent Reflecting boundary term:
+%
+\begin{gather}
+  \langle\bm{\Psi}^*,\bm{\Psi}\rangle^+ - \langle\bm{\Psi}^*,\mathbb{B}\bm{\Psi}\rangle^- =
+  \begin{cases}
+    \sum^G_{g=1}\sum^{N_d}_{d=1}w_d\left(\Psi^*_{g,d},
+    \hat{\Omega}_d\cdot\hat{n}_b\Psi_{g,d}\right),
+    & \hat{\Omega}\cdot\hat{n}_b>0,\vec{r}\in\partial\mathcal{D}_s \\
+    \sum^G_{g=1}\sum^{N_d}_{d=1}w_d\left(\Psi^*_{g,d},
+    \hat{\Omega}_d\cdot\hat{n}_b\Psi_{g,d_r}\right),
+    & \hat{\Omega}\cdot\hat{n}_b<0,\vec{r}\in\partial\mathcal{D}_s
+  \end{cases},
+  \shortintertext{where}
+  \hat{\Omega}_{d_r} = \hat{\Omega}_d - 2(\hat{\Omega}_d\cdot\hat{n}_b)\hat{n}_b. \nonumber
+\end{gather}
+
+This concludes the derivation of the \gls{SAAF} formulation of the $S_N$ neutron transport method
+for generating transport corrections in the hybrid $S_N$-diffusion method.
+
+\section{Transport Correction Formulation}
+
+The conventional approach for determining diffusion coefficients for each subregion involves
+running a high-fidelity neutron transport simulation to tally region-wide estimates of the neutron
+transport cross section. The transport cross section formulation is derived from the $P_1$
+approximation of the neutron transport equation with isotropic sources \cite{bell_nuclear_1970} as
+shown in Eq.\ \ref{eq:p1-diffcoef}.
+Essentially, each defined subregion has a constant diffusion coefficient value. However, as
+discussed in Section \ref{sec:summary-nts-mtds}, the neutron diffusion
+equation is only valid in regions of high scattering-to-removal ratios with, at most, linearly
+anisotropic scattering and small flux gradients. These conditions do not hold near or within
+control rods, near interfaces of neighboring materials with highly dissimilar neutronic properties,
+and in materials with significant scattering contributions from light nuclei.
+
+Three main options exist in the literature for applying
+transport corrections to the neutron diffusion equations: applying
+diffusion coefficient corrections, adding a drift term, or adding a
+correction term.
+
+\subsection{Diffusion Coefficient Correction}
+
+Diffusion coefficient corrections involve replacing the diffusion coefficient $D_g$ in the
+diffusion term with ``optimal'' diffusion coefficients based on
+pointwise transport corrections. The literature review in Chapter \ref{chap:lit} covers two such
+examples in the Ronen method by Ronen \cite{ronen_accurate_2004} and the space-dependent diffusion
+coefficients by Pounders \& Rahnema \cite{pounders_diffusion_2009}.
+
+For the hybrid $S_N$-diffusion method, I used a formulation that incorporates pointwise
+corrections to the neutron diffusion flux solution from the $S_N$-derived flux solution as follows:
+%
+\begin{align}
+  D^s_g(x) &= -J^{tr}_g(x)\bigg/\frac{d\phi^{tr}_g(x)}{dx}. \label{eq:svdc}
+\end{align}
+%
+where $D^s_g$ is the \gls{SVDC}, and the $tr$ superscript denotes the transport-derived neutron
+current and scalar flux solutions from the $S_N$ method.
+The basic form of this equation is Fick's first law of diffusion. The result is a diffusion
+coefficient variable that varies in space from point to point. We assume that it is continuous and
+at least once differentiable except at dissimilar material interfaces. \glspl{SVDC} provide
+pointwise corrections to closely match the diffusion flux
+solution to the $S_N$ flux solution.
+
+Eq.\ \ref{eq:svdc} is identical to Ronen's \cite{ronen_accurate_2004} and Pounders \& Rahnema's
+\cite{pounders_diffusion_2009} formulations for space-dependent diffusion coefficients in Eq.\
+\ref{eq:ronen} and Eq.\ \ref{eq:emp}, respectively. In comparison with
+the Ronen method, our approaches differ in how $\phi_{tr}$ and $J_{tr}$ are obtained. Starting with
+a standard neutron diffusion calculation, the Ronen method applies an analytically derived
+transport operator on every iteration to calculate a new estimate of $J_{tr}$ using the $\phi$
+solution from the previous iteration. The main difficulty for the Ronen method lies in deriving the
+transport operators which has been demonstrated for only 1-D geometries thus far.
+
+For the empirical method developed by Pounders \& Rahnema, they assumed a priori knowledge of the
+reference flux solution and used it to generate piecewise-constant, space-dependent diffusion
+coefficients.
+Pounders \& Rahnema \cite{pounders_diffusion_2009} demonstrated the effectiveness of applying
+pointwise corrections derived from analytical or Monte Carlo reference flux solutions. Compared
+with conventional $P_1$-based out-scatter and flux-limited approximations of the diffusion
+coefficient, their diffusion coefficients showed superior agreement
+with the reference flux solutions. They recognized that volume averaging within each mesh element
+for piecewise-constant diffusion coefficients introduces some truncation error if the flux is
+non-linear within the mesh element. This design choice may be due to an intention to retain the
+diffusion coefficient as a constant in the $\frac{d}{dx}D\frac{d\phi}{dx}$ term of the neutron
+diffusion equation (Eq.\ \ref{eq:1d-diff}).
+
+Unlike their approach, the \gls{SVDC} formulation in Eq.\ \ref{eq:svdc} allows for continuously
+varying diffusion coefficients to reduce truncation error. In practice, the discretization order of
+\gls{SVDC} variables in a numerical calculation would be the same as the discretization order of
+the reference flux solution. This formulation introduces a minor change to the finite difference
+implementation of the 2nd-order diffusion term to handle spatial derivatives of the
+diffusion coefficient as $D^s_g$ is not uniform in space.
+
+Regardless of the differences in implementations amongst the methods discussed here,
+diffusion coefficient corrections applied through Eq.\ \ref{eq:saaf} were
+found to be effective in enabling the neutron diffusion method to accurately reproduce reference
+flux solutions \cite{tomatis_application_2011, gross_high-accuracy_2020, tomatis_ronen_2021,
+gross_comprehensive_2023, cufe_study_2024, pounders_diffusion_2009}.
+However, a significant challenge for space-dependent diffusion coefficients
+involves their resolution near neutron flux peaks. The neutron current and flux gradient values in
+Eq.\ \ref{eq:svdc} generally do not reach zero at the same points in space,
+resulting in diffusion coefficient values tending to positive or negative infinity when the
+flux gradient is close to zero. Pounders \& Rahnema avoided this issue by using
+larger mesh sizes to calculate their empirical diffusion coefficients. However, their
+remedy contradicts mesh convergence requirements and would worsen flux accuracy in regions with
+steep, non-linear fluxes, such as near control rods. For the Ronen method, Gross et al.\
+\cite{gross_comprehensive_2023} applied a numerical fix by avoiding correction
+calculations wherever the flux gradient values fall below a certain threshold. 
+Tomatis \& Dall'Osso
+\cite{tomatis_application_2011} developed an alternative formulation in the form of drift terms
+while
 

### 2024-06-17 hybrid.tex
transport correction formulation

@@
   \Psi_g(\vec{r},\hat{\Omega},t) =
-  \sum^G_{g'=1}\int_\mathcal{S} \Sigma_{s,g'\rightarrow g}(\hat{\Omega}'\rightarrow\hat{\Omega})
+  \sum^G_{g'=1}\int_\mathcal{S} \Sigma_s^{g'\rightarrow g}(\hat{\Omega}'\rightarrow\hat{\Omega})
   \Psi_{g'}(\vec{r},\hat{\Omega}',t)d\hat{\Omega}' \\
@@
   \end{bmatrix},
-  \bm{\phi} \equiv \int_S \bm{\Psi}d\hat{\Omega} \equiv
+  \bm{\Phi} \equiv \int_S \bm{\Psi}d\hat{\Omega} \equiv
   \begin{bmatrix}
@@
   \begin{bmatrix}
-    \sum^G_{g'=1}\int_S \Sigma_{s,g'\rightarrow 1}\Psi_{g'}d\hat{\Omega} \\
-    \sum^G_{g'=2}\int_S \Sigma_{s,g'\rightarrow 2}\Psi_{g'}d\hat{\Omega} \\
+    \sum^G_{g'=1}\int_S \Sigma_s^{g'\rightarrow 1}\Psi_{g'}d\hat{\Omega} \\
+    \sum^G_{g'=2}\int_S \Sigma_s^{g'\rightarrow 2}\Psi_{g'}d\hat{\Omega} \\
     \vdots \\
-    \sum^G_{g'=G}\int_S \Sigma_{s,g'\rightarrow G}\Psi_{g'}d\hat{\Omega}
+    \sum^G_{g'=G}\int_S \Sigma_s^{g'\rightarrow G}\Psi_{g'}d\hat{\Omega}
   \end{bmatrix},
@@
 %
-where $\Psi^\text{inc}_{g,d}=0$ for a vacuum boundary and positive for a boundary source.
+where $\Psi^\text{inc}_{g,d}$ is zero on a vacuum boundary or positive for a boundary source.
 
@@
 
-\section{Transport Correction Formulation}
+\section{Transport Correction Formulations}
 
@@
 
-Three main options exist in the literature for applying
+In this work, we explored two options for applying
 transport corrections to the neutron diffusion equations: applying
-diffusion coefficient corrections, adding a drift term, or adding a
-correction term.
-
-\subsection{Diffusion Coefficient Correction}
-
-Diffusion coefficient corrections involve replacing the diffusion coefficient $D_g$ in the
+diffusion corrections or adding a drift correction term.
+
+\subsection{Diffusion Correction}
+
+Diffusion corrections involve replacing the diffusion coefficient $D_g$ in the
 diffusion term with ``optimal'' diffusion coefficients based on
-pointwise transport corrections. The literature review in Chapter \ref{chap:lit} covers two such
+pointwise transport corrections. The literature review in Section \ref{sec:summary-nts-mtds} covers
+two such
 examples in the Ronen method by Ronen \cite{ronen_accurate_2004} and the space-dependent diffusion
@@
 
-For the hybrid $S_N$-diffusion method, I used a formulation that incorporates pointwise
+For the hybrid $S_N$-diffusion method, we used a formulation that incorporates pointwise
 corrections to the neutron diffusion flux solution from the $S_N$-derived flux solution as follows:
@@
 where $D^s_g$ is the \gls{SVDC}, and the $tr$ superscript denotes the transport-derived neutron
-current and scalar flux solutions from the $S_N$ method.
-The basic form of this equation is Fick's first law of diffusion. The result is a diffusion
-coefficient variable that varies in space from point to point. We assume that it is continuous and
+current and scalar flux solutions from the $S_N$ method. Transport corrections introduced through
+$J_{tr}$ are scaled by the flux gradient. We assumed that it varies continuously and
 at least once differentiable except at dissimilar material interfaces. \glspl{SVDC} provide
-pointwise corrections to closely match the diffusion flux
-solution to the $S_N$ flux solution.
+pointwise corrections to closely match the diffusion flux solution to the $S_N$ flux solution.
+By replacing $D_g$ with $D^s_g$, we are effectively adding the following transport correction term
+%
+\begin{gather}
+  -\frac{\partial}{\partial x}(D^s_g-D_g)\frac{\partial\phi_g}{\partial x}
+\end{gather}
+to the neutron diffusion equations. Alternatively, we may define
+$\partial D\equiv\left(D^s_g-D_g)\right)$. This alternative form shows that diffusion
+correction applies a multiplicative closure that scales with the flux gradient.
 
@@
 diffusion coefficient as a constant in the $\frac{d}{dx}D\frac{d\phi}{dx}$ term of the neutron
-diffusion equation (Eq.\ \ref{eq:1d-diff}).
+diffusion equation (Eq.\ \ref{eq:mg-diff}).
 
@@
 Regardless of the differences in implementations amongst the methods discussed here,
-diffusion coefficient corrections applied through Eq.\ \ref{eq:saaf} were
+diffusion corrections applied through Eq.\ \ref{eq:saaf} were
 found to be effective in enabling the neutron diffusion method to accurately reproduce reference
-flux solutions \cite{tomatis_application_2011, gross_high-accuracy_2020, tomatis_ronen_2021,
-gross_comprehensive_2023, cufe_study_2024, pounders_diffusion_2009}.
-However, a significant challenge for space-dependent diffusion coefficients
-involves their resolution near neutron flux peaks. The neutron current and flux gradient values in
-Eq.\ \ref{eq:svdc} generally do not reach zero at the same points in space,
+flux solutions \cite{gross_comprehensive_2023, pounders_diffusion_2009}.
+Howver, a significant challenge for diffusion corrections
+involves their resolution near neutron flux peaks. As observed in Eq.\ \ref{eq:svdc}, the neutron
+current and flux gradient values do not necessarily reach zero at the same points in space,
 resulting in diffusion coefficient values tending to positive or negative infinity when the
@@
 calculations wherever the flux gradient values fall below a certain threshold.
-Tomatis \& Dall'Osso
-\cite{tomatis_application_2011} developed an alternative formulation in the form of drift terms
-while
-
-\subsection{Spatially Varying Diffusion Coefficients} \label{sec:svdc}
-
-The conventional approach for determining diffusion coefficients for each subregion involves
-running a high-fidelity neutron transport simulation to tally region-wide estimates of the neutron
-transport cross section. The transport cross section formulation is derived from the $P_1$
-approximation of the neutron transport equation with isotropic sources \cite{bell_nuclear_1970}.
-Essentially, each defined subregion has a constant diffusion coefficient value. However, as
-discussed in Section \ref{sec:summary-nts-mtds}, the neutron diffusion
-equation is only valid in regions of high scattering-to-removal ratios with, at most, linearly
-anisotropic scattering and small flux gradients. These conditions do not hold near or within
-control rods, near interfaces of neighboring materials with highly dissimilar neutronic properties,
-and in materials with significant scattering contributions from light nuclei.
-
-In the hybrid $S_N$-diffusion method, I propose replacing the conventional $P_1$-based
-diffusion coefficients with an alternate formulation that incorporates pointwise corrections
-to the neutron diffusion flux solution from the $S_N$-derived flux solution as follows:
-%
-\begin{align}
-  D^s_g(x) &= -J^{tr}_g(x)\bigg/\frac{d\phi^{tr}_g(x)}{dx}. \label{eq:svdc}
-\end{align}
-%
-where $D^s$ is the \glspl{SVDC}, and the $tr$ superscript denotes the transport-computed neutron
-current and scalar flux solutions from the $S_N$ method.
-The basic form of this equation is Fick's first law of diffusion. The result is a diffusion
-coefficient variable that continuously varies in space from point to point except at dissimilar
-material interfaces. \glspl{SVDC} provide pointwise corrections to closely match the diffusion flux
-solution to the $S_N$ flux solution.
-
-Eq.\ \ref{eq:svdc} is identical to Ronen's \cite{ronen_accurate_2004} and Pounders \& Rahnema's
-\cite{pounders_diffusion_2009} formulations for space-dependent diffusion coefficients in Eq.\
-\ref{eq:ronen} and Eq.\ \ref{eq:emp}, respectively. In comparison with
-the Ronen method, our approaches differ in how $\phi_{tr}$ and $J_{tr}$ are obtained. Starting with
-a standard neutron diffusion calculation, the Ronen method applies an analytically derived
-transport operator on every iteration to calculate a new estimate of $J_{tr}$ using the $\phi$
-solution from the previous iteration. The main difficulty for the Ronen method lies in deriving the
-transport operators which has been achieved for only simple 1-D geometries thus far.
-
-For the \textit{empirical method} by Pounders \& Rahnema, their diffusion coefficients are
-piecewise-constant and they assumed a priori knowledge of the reference flux solution.
-Pounders \& Rahnema \cite{pounders_diffusion_2009} demonstrated the effectiveness of applying
-pointwise corrections derived from analytical or Monte Carlo reference flux solutions. Compared
-with conventional $P_1$-based out-scatter and flux-limited approximations of the diffusion
-coefficient, their diffusion coefficients showed superior agreement
-with the reference flux solutions. They recognized that volume averaging within each mesh element
-for piecewise-constant diffusion coefficients introduces some truncation error if the flux is
-non-linear within the mesh element. This design choice may be due to an intention to retain the
-diffusion coefficient as a constant in the $\frac{d}{dx}D\frac{d\phi}{dx}$ term of the neutron
-diffusion equation (Eq.\ \ref{eq:1d-diff}).
-
-Unlike their approach, the \gls{SVDC} formulation in Eq.\ \ref{eq:svdc} allows for continuously
-varying diffusion coefficients to reduce truncation error. In practice, the discretization order of
-\gls{SVDC} variables in a numerical calculation would be the same as the discretization order of
-the reference flux solution. This formulation introduces a minor change to the finite difference
-implementation of the 2nd-order diffusion term to handle spatial derivatives of the
-diffusion coefficient. However, the finite element implementation sees no change as the 2nd-order
-diffusion term is commonly treated with integration-by-parts to reduce it to a 1st-order
-differential term with no derivatives of the diffusion coefficient. Refer to Section
-\ref{sec:implementation} for key numerical implementation details of a finite difference neutron
-diffusion solver with support for handling \glspl{SVDC}.
-
-For the rest of Section \ref{sec:svdc}, I will show how the \glspl{SVDC} may differ from the
-conventional $P_1$-based diffusion coefficient before presenting further details of the hybrid
-$S_N$-diffusion method in Section \ref{sec:theory}.
-
-\begin{figure}[htb!]
-  \centering
-  \includegraphics[width=.7\columnwidth]{case-0-geometry}
-  \caption{Case 0: A 1-D, two-region system containing a 0.5-cm thick control rod and a 24.5-cm
-    thick homogeneous mixture of molten fuel salt and graphite moderator. Reflective boundary
-    conditions apply on both ends.}
-  \label{fig:case-0-geom}
-\end{figure}
-
-\subsubsection{\gls{SVDC} Demonstration with Case 0}
-
-To facilitate the following demonstration of a neutron diffusion calculation with \glspl{SVDC},
-consider a 1-D, two-region system (which I will refer to as Case 0 moving forward) consisting of a
-highly neutron-absorbing material and a
-neutron-multiplying region with reflective boundary conditions on both ends (Figure
-\ref{fig:case-0-geom}). I defined the material specifications of the control rod, molten fuel salt,
-and graphite moderator based on the material compositions of the \gls{MSRE}
-\cite{robertson_msre_1965}. The
-homogenized molten fuel salt and graphite regions minimize discrepancies arising from geometrical
-heterogeneity. The group constant input data for the diffusion and $S_N$ solvers (e.g., $P_1$-based
-diffusion coefficients, cross sections, fission spectra) were sampled at 900 K and generated using
-the OpenMC Monte Carlo neutronics software \cite{romano_openmc:_2015}. OpenMC uses the flux-limited
-approximation method, developed by Pomraning \cite{pomraning_flux-limited_1984}, to calculate
-diffusion coefficients. The condensed
-neutron energy spectrum forms two energy groups bounded between $10^{-5}$, $10^0$, and $10^8$ eV.
-
-I solved for the neutron flux in this system using the following set of numerical solvers:
-%
-\begin{enumerate}
-  \item Neutron diffusion solver with $P_1$-based diffusion coefficients generated directly from
-    the group constant generation step with OpenMC.
-  \item $S_N$ neutron transport solver with $N=8$ and up to 2nd-order Legendre
-    moments of the group-to-group neutron scattering cross sections.
-  \item Neutron diffusion solver with \glspl{SVDC} generated from the prior $S_8$
-    flux solution.
-\end{enumerate}
-%
-I implemented the diffusion solvers using the finite difference method and the $S_N$
-solver using the diamond-difference transport sweep method in Python. For
-brevity, I defer further numerical implementation details of these solvers to Section
-\ref{sec:implementation}.
-
-\begin{figure}[htb!]
-  \centering
-  \begin{subfigure}[b]{.49\textwidth}
-    \centering
-    \includegraphics[width=\textwidth]{case-0-group-1-flux}
-    \caption{Group 1}
-    \label{fig:c0g1flux}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}[b]{.49\textwidth}
-    \centering
-    \includegraphics[width=\textwidth]{case-0-group-2-flux}
-    \caption{Group 2}
-    \label{fig:c0g2flux}
-  \end{subfigure}
-  \caption{Neutron group 1 and 2 flux distributions from the diffusion, $S_8$, and
-  diffusion-\gls{SVDC} solvers for Case 0.}
-  \label{fig:c0flux}
-\end{figure}
-%
-\begin{figure}[htb!]
-  \centering
-  \begin{subfigure}[b]{.49\textwidth}
-    \centering
-    \includegraphics[width=\textwidth]{case-0-group-1-diffcoef}
-    \caption{Group 1}
-    \label{fig:c0g1diffcoef}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}[b]{.49\textwidth}
-    \centering
-    \includegraphics[width=\textwidth]{case-0-group-2-diffcoef}
-    \caption{Group 2}
-    \label{fig:c0g2diffcoef}
-  \end{subfigure}
-  \caption{$P_1$-based diffusion coefficient and \gls{SVDC} spatial distributions
-  for Case 0.}
-  \label{fig:c0diffcoef}
-\end{figure}
-
-Figure \ref{fig:c0flux} shows the group 1 and 2 neutron fluxes from the diffusion-$P_1$, $S_8$, and
-diffusion-\gls{SVDC} solvers with a uniform mesh size of 0.005 cm. I omitted the OpenMC flux
-solution because this exercise aims to demonstrate the effectiveness of \glspl{SVDC}
-in reproducing an $S_N$-derived flux solution. As expected, the diffusion-$P_1$ flux solution
-deviates from the $S_8$ flux solution, while the diffusion-\gls{SVDC} flux solution shows
-significantly better agreement with the $S_8$ flux solution. As shown in Table \ref{table:c0k}, the
-$k$ estimate from the diffusion-\gls{SVDC} solver is also closer to the $k$ estimate from the
-$S_8$ solver than the diffusion-$P_1$ solver.
-
-\begin{table}[tb!]
-  \centering
-  \caption{Multiplication factor $k$ estimates from the diffusion-$P_1$, $S_8$, and
-  diffusion-\gls{SVDC} solvers and the absolute difference relative to the $S_8$ estimate.}
-  \begin{tabular}{l S S}
-    \toprule
-    Solver type & {$k$} & {$k-k_{S_8}$} \\
-    \midrule
-    $S_8$ & 0.62736 & {-} \\
-    Diffusion-$P_1$ & 0.60890 & -0.01846 \\
-    Diffusion-\gls{SVDC} & 0.62750 & +0.00013 \\
-    \bottomrule
-  \end{tabular}
-  \label{table:c0k}
-\end{table}
-
-The $P_1$-based diffusion coefficients and \glspl{SVDC} in Figure \ref{fig:c0diffcoef} show good
-agreement in the bulk fuel salt region far from the control rod region.
-This observation supports the validity of the neutron diffusion method under the conditions
-listed at the start of Section \ref{sec:svdc}. Closer to the control rod
-region, the \glspl{SVDC} deviate from the $P_1$-based diffusion coefficients because the
-prerequisites for diffusion theory are violated.
-
-Another key finding from this study is that the neutron diffusion method with $P_1$-based
-diffusion coefficients fails to reproduce the steep $S_8$ flux gradient in the control rod region
-for group 1 and in the homogeneous fuel-graphite region adjacent to the control rod region for both
-neutron groups. From a purely mathematical perspective, the streaming term in the neutron
-diffusion equation requires the lower
-\gls{SVDC} values in the absorber-adjacent region to induce steeper flux gradients and match the
-$S_8$ flux solution. Given that highly neutron-absorbing materials induce steep flux gradients
-around them, we can expect similar trends of \glspl{SVDC} being suppressed relative to
-$P_1$-based diffusion coefficients in other systems with control rods. The \gls{SVDC} values within
-the highly-absorbing region exhibit significant spatial variation but do not necessarily fall
-below the $P_1$-based diffusion coefficients.
-
-Thus far, this method is redundant because it requires a priori knowledge of a reference
-analytical flux solution or a highly accurate numerical solution calculated using neutron
-transport methods. While existing workflows for diffusion-based methods already require
-computationally intensive neutron transport simulations for group constant generation, this
-preprocessing step incurs a fixed computational cost through
-a series of neutron transport simulations. Compared to \glspl{SVDC}, $P_1$-based diffusion
-coefficients behave much more like intrinsic material properties as their definitions are largely
-tied to material cross sections. Variations in $P_1$-based diffusion coefficients primarily arise
-from changes in the neutron energy spectrum with no direct contribution from proximity to material
-interfaces (geometrical heterogeneity). $P_1$-based diffusion coefficients may be generated at
-various reactor temperatures and interpolated to model the temperature dependence. Conversely, the
-nature of \glspl{SVDC} as empirical, pointwise corrections for the diffusion equation
-makes it highly dependent on the neutron flux gradient and susceptible to more significant
-variations than region-wide estimates for $P_1$-based diffusion coefficients. \glspl{SVDC} will
-likely need to be dynamically generated at every timestep, such as with a two-level iterative
-scheme consisting of a $S_N$ neutron transport calculation and a neutron
-diffusion calculation. In this case, the required number of neutron transport simulations scales
-with the number of reactor simulations. Therefore, \gls{SVDC} generation cannot be adopted as a
-one-off preprocessing step.
-
-Another significant challenge of \glspl{SVDC} and the high-order empirical diffusion coefficients
-involves their resolution near neutron flux peaks. The neutron current and flux gradient values in
-Eq.\ \ref{eq:svdc} generally do not reach zero at the same points in space,
-resulting in diffusion coefficient values tending to positive or negative infinity when the
-flux gradient is close to zero. Pounders \& Rahnema avoided this issue by using
-larger mesh sizes to calculate their empirical diffusion coefficients. However, their
-remedy contradicts mesh convergence requirements and would worsen flux accuracy in regions with
-steep, non-linear fluxes, such as near control rods. I will discuss my approach for investigating
-and finding an alternative solution for this issue in Section \ref{sec:devel-hybrid}.
-
-\subsection{Hybrid $S_N$-Diffusion Method} \label{sec:hybrid-method}
+
+\subsection{Drift Correction Term}
+
+Drift correction terms feature prominently in existing literature as transport correction
+terms for \gls{NDA} schemes or nodal diffusion methods
+\cite{smith_nodal_1983, smith_assembly_1986, adams_fast_2002, wang_diffusion_2014}. \gls{NDA}
+schemes are \gls{HOLO} methods \cite{chacon_multiscale_2017} similar to the \gls{QD} method for
+accelerating a neutron transport
+method (high-order) with a modified neutron diffusion method (low-order). Drift terms
+are added to the low-order neutron diffusion equations to correct them such that the modified
+equations can reproduce the transport solution upon reaching iterative convergence. In nodal
+diffusion methods, the drift terms are used to rectify the incoming and outgoing neutron
+currents between adjacent nodes.
+
+The general form of drift correction terms to be added to the neutron diffusion equations is a
+first-order derivative term $\nabla\cdot \vec{D}_g\phi_g$ (note the vector notation). As such,
+drift terms apply
+multiplicative closures that scale with the flux. This contrasts with diffusion
+corrections which scale with the flux gradient instead. Therefore, drift terms do not suffer from
+the division-by-zero errors encountered with diffusion corrections.
+
+Derivations for the drift terms depend on the discretization schemes of the high- and low-order
+equations. For this work, we will adopt drift term derivations developed by Wang et al.\
+\cite{wang_diffusion_2014, wang_rattlesnake_2018}. We start by integrating Eq.\ \ref{eq:saaf-vt}
+over the 2-D unit sphere angular domain $\mathcal{S}$ and replacing $\bm{\Psi}^*$ with isotropic
+test functions $\bm{\Phi}^*$ to obtain the neutron balance equation:
+%
+\begin{multline}
+  \left(\bm{\Phi}^*,\frac{\partial}{\partial t}\left(\frac{\bm{\Phi}}{\bm{v}}\right)\right)_\mathcal{D}
+  + \left(\mathbb{L}_2\bm{\Phi}^*,\bm{\Phi}\right)_\mathcal{D}
+  - \left(\nabla\bm{\Phi}^*,\vec{\bm{J}}\right)_\mathcal{D}
+  + \left(\bm{\Phi}^*,\bm{J}^\text{out}\right)_{\partial\mathcal{D}_v} + \\
+  \left(\bm{\tau}\nabla\bm{\Phi}^*, \frac{\partial}{\partial t}\left(\frac{\vec{\bm{J}}}{\bm{v}}\right)
+    + \nabla\cdot(\vec{\vec{\bm{E}}}\bm{\Phi}) +\mathbb{L}_2\vec{\bm{J}} -
+    \mathbb{S}_1\vec{\bm{J}}\right)_\mathcal{D}
+  = \left(\bm{\Phi}^*,\mathbb{S}_0\bm{\Phi}\right)_\mathcal{D}
+  + \left(\bm{\Phi}^*,\bm{Q}_0\right)_\mathcal{D},
+\end{multline}
+%
+\begin{gather}
+  \shortintertext{where}
+  \vec{\bm{J}} \equiv
+  \begin{bmatrix}
+    \vec{J}_1 \\
+    \vec{J}_2 \\
+    \vdots \\
+    \vec{J}_G
+  \end{bmatrix},
+  \bm{J}^\text{out} \equiv
+  \begin{bmatrix}
+    \int_{|\hat{\Omega}\cdot\hat{n}_b|>0}|\hat{\Omega}\cdot\hat{n}_b|\Psi_1 d\hat{\Omega} \\
+    \int_{|\hat{\Omega}\cdot\hat{n}_b|>0}|\hat{\Omega}\cdot\hat{n}_b|\Psi_2 d\hat{\Omega} \\
+    \vdots \\
+    \int_{|\hat{\Omega}\cdot\hat{n}_b|>0}|\hat{\Omega}\cdot\hat{n}_b|\Psi_G d\hat{\Omega}
+  \end{bmatrix},
+  \bm{J}^\text{inc} \equiv
+  \begin{bmatrix}
+    J^\text{inc}_1 \\
+    J^\text{inc}_2 \\
+    \vdots \\
+    J^\text{inc}_G
+  \end{bmatrix},
+  \vec{\vec{\bm{E}}} \equiv
+  \begin{bmatrix}
+    \frac{\int_\mathcal{S} \hat{\Omega}\hat{\Omega}\Psi_1 d\hat{\Omega}}{\int_\mathcal{S}\Psi_1 d\hat{\Omega}} \\
+    \frac{\int_\mathcal{S} \hat{\Omega}\hat{\Omega}\Psi_2 d\hat{\Omega}}{\int_\mathcal{S}\Psi_2 d\hat{\Omega}} \\
+    \vdots \\
+    \frac{\int_\mathcal{S} \hat{\Omega}\hat{\Omega}\Psi_G d\hat{\Omega}}{\int_\mathcal{S}\Psi_G d\hat{\Omega}}
+  \end{bmatrix}, \nonumber \\
+  \mathbb{S}_0 \bm{\Phi} \equiv
+  \begin{bmatrix}
+    \sum^G_{g'=1}\Sigma_s^{g'\rightarrow 1}\phi_{g'} \\
+    \sum^G_{g'=1}\Sigma_s^{g'\rightarrow 2}\phi_{g'} \\
+    \vdots \\
+    \sum^G_{g'=1}\Sigma_s^{g'\rightarrow G}\phi_{g'}
+  \end{bmatrix},
+  \mathbb{S}_1 \vec{\bm{J}} \equiv
+  \begin{bmatrix}
+    \sum^G_{g'=1}\Sigma_{s,1}^{g'\rightarrow 1}\vec{J}_{g'} \\
+    \sum^G_{g'=1}\Sigma_{s,1}^{g'\rightarrow 2}\vec{J}_{g'} \\
+    \vdots \\
+    \sum^G_{g'=1}\Sigma_{s,1}^{g'\rightarrow G}\vec{J}_{g'}
+  \end{bmatrix},
+  \bm{Q}_0 \equiv
+  \begin{bmatrix}
+    \int_\mathcal{S} Q_1 d\hat{\Omega} \\
+    \int_\mathcal{S} Q_2 d\hat{\Omega} \\
+    \vdots \\
+    \int_\mathcal{S} Q_G d\hat{\Omega}
+  \end{bmatrix}, \nonumber
+\end{gather}
+%
+and $\partial\mathcal{D}_v$ represents sections of $\partial\mathcal{D}$ which are vacuum boundaries. We
+assumed that $\bm{Q}$ is an isotropic source because we do not deal with anisotropic sources in
+this work outside of scattering.
+We define the following expression
+%
+\begin{gather}
+  \vec{\bm{r}}_1 \equiv \frac{\partial}{\partial t}\left(\frac{\vec{\bm{J}}}{\bm{v}}\right)
+  + \nabla\cdot(\vec{\vec{\bm{E}}}\bm{\Phi}) + \mathbb{L}_2\vec{\bm{J}} -
+  \mathbb{S}_1\vec{\bm{J}} \label{eq:1st-moment}
+\end{gather}
+%
+to simplify the balance equation. We note that Eq.\ \ref{eq:1st-moment} is identical to the second
+low-order equation of the \gls{QD} method (Eq.\ \ref{eq:qd-low-1st}).
+We also introduce the diffusion operator
+%
+\begin{gather}
+  \mathbb{D}\nabla\bm{\Phi} \equiv
+  \begin{bmatrix}
+    D_1\nabla\phi_1 \\
+    D_2\nabla\phi_2 \\
+    \vdots \\
+    D_G\nabla\phi_G
+  \end{bmatrix},
+\end{gather}
+%
+and recognize that under the diffusion approximation, the vacuum boundary condition is given as
+%
+\begin{gather}
+  -\mathbb{D}\nabla\bm{\Phi}\cdot\hat{n}_b = \frac{\bm{\Phi}}{2}, \hspace{1cm} \vec{r}\in\partial
+  \mathcal{D}_v.
+\end{gather}
+%
+We can rewrite the balance equation as
+%
+\begin{multline}
+  \left(\bm{\Phi}^*,\frac{\partial}{\partial t}\left(\frac{\bm{\Phi}}{\bm{v}}\right)\right)_\mathcal{D}
+  + \left(\nabla\bm{\Phi}^*, \mathbb{D}\nabla\bm{\Phi}\right)_\mathcal{D}
+  + \left(\mathbb{L}_2\bm{\Phi}^*,\bm{\Phi}\right)_\mathcal{D}
+  + \left(\bm{\Phi}^*,\frac{\bm{\Phi}}{2}\right)_{\mathcal{D}_v}
+  + \left(\nabla\bm{\Phi}^*,\bm{\tau}\vec{\bm{r}}_1-\vec{\bm{J}}-\mathcal{D}\bm{\Phi}\right)_\mathcal{D}
+  + \\
+  \left(\bm{\Phi}^*,\bm{J}^\text{out}-\frac{\bm{\Phi}}{2}\right)_{\partial\mathcal{D}_v}
+  = \left(\bm{\Phi}^*,\mathbb{S}_0\bm{\Phi}\right)_\mathcal{D}
+  + \left(\bm{\Phi}^*,\bm{Q}_0\right)_\mathcal{D}. \label{eq:modified-diff}
+\end{multline}
+%
+We can observe that Eq.\ \ref{eq:modified-diff} is a modified neutron diffusion equation with
+transport corrections provided by the fifth and sixth terms. The fifth term is a drift term with
+the drift vector defined as
+%
+\begin{gather}
+  \vec{\bm{D}} \equiv \frac{\bm{\tau}\vec{\bm{r}}_1-\vec{\bm{J}}-\mathbb{D}\nabla\bm{\Phi}}{\bm{\Phi}}
+\end{gather}
+%
+and the sixth term is a boundary correction term with the boundary coefficient vector defined as
+%
+\begin{gather}
+  \bm{\gamma} \equiv \frac{\bm{J}^\text{out}}{\bm{\Phi}}-\frac{1}{2}\bm{I}.
+\end{gather}
+%
+With the $S_N$ angular discretization scheme, the drift and boundary correction vector components
+can be evaluated as
+%
+\begin{gather}
+  \vec{D}_g = \frac{\sum^{N_d}_{d=1}w_d\left(\tau_g\hat{\Omega}_d\hat{\Omega}_d\cdot\nabla\Psi_{g,d}
+  + \left(\tau_g\Sigma_{t,g}-1\right)\hat{\Omega}_d\Psi_{g,d}
+  - \tau_g\sum^G_{g'=1}\Sigma^{g'\rightarrow g}_{s,1}\hat{\Omega}_d\Psi_{g',d}
+  - D_g\nabla\Psi_{g,d}\right)}{\sum^{N_d}_{d=1}w_d\Psi_{g,d}}, \\
+  \gamma_g =
+  \frac{\sum_{\hat{\Omega}_d\cdot\hat{n}_b > 0}w_d |\hat{\Omega}_d\cdot\hat{n}_b |
+  \Psi_{g,d}}{\sum^{N_d}_{d=1}w_d\Psi_{g,d}}.
+\end{gather}
+
+%\subsection{Additive Correction Term}
+%
+%Lastly, we can also incorporate transport corrections into the neutron diffusion equation through
+%additional additive terms.
+%We can obtain an additive correction term formulation from the \textit{Second-Moment}
+%method by Lewis \& Miller \cite{lewis_comparison_1976}. The literature review in Section
+%\ref{sec:summary-nts-mtds} covers how Anistratov \& Stehle \cite{anistratov_computational_2012}
+%used this formulation for their hybrid method with domain decomposition. Refer to Eqs.\
+%\ref{eq:2nd-moment-1}, \ref{eq:2nd-moment-2}, and \ref{eq:2nd-moment-3} for mathematical
+%formulations. After substituting Eq.\ \ref{eq:2nd-moment-2} into Eq.\ \ref{eq:2nd-moment-1} to
+%eliminate $J^{(s+1)}$, we have a modified neutron diffusion equation for $\phi^{(s+1)}$ with an
+%additional correction term $\nabla\cdot\nabla\cdot F^{(s+1/2)}$, where $(s)$ denotes the
+%iteration index for iterating between the high- and low-order equations.
+%
+%Unlike the previous two transport correction formulations, the additive correction term is an
+%additive closure that does not scale with the flux or the flux gradient in the low-order equation.
+%Therefore, for implementations of \gls{HOLO} or transport correction methods in which the high- and
+%low-order scalar flux solutions differ, additive correction terms may not provide the correction.
+
+\section{Hybrid $S_N$-Diffusion Method} \label{sec:hybrid-method}
 

### 2024-06-18 hybrid.tex
Move old implementation to appendix

@@
 
-\section{\gls{SAAF} Formulation of the $S_N$ Method}
+\section{\Gls{SAAF} Formulation of the $S_N$ Method}
 
@@
 
-The conventional approach for determining diffusion coefficients for each subregion involves
+The conventional approach for determining diffusion coefficients for the neutron diffusion
+equations in each subregion involves
 running a high-fidelity neutron transport simulation to tally region-wide estimates of the neutron
@@
 
+Transport corrections to allow neutron diffusion methods to reproduce high-fidelity flux solutions
+can be introduced by modifying the neutron diffusion equations.
 In this work, we explored two options for applying
@@
 current and scalar flux solutions from the $S_N$ method. Transport corrections introduced through
-$J_{tr}$ are scaled by the flux gradient. We assumed that it varies continuously and
+$J_{tr}$ are scaled by the flux gradient. We assumed that it varies continuously and is
 at least once differentiable except at dissimilar material interfaces. \glspl{SVDC} provide
@@
   - \tau_g\sum^G_{g'=1}\Sigma^{g'\rightarrow g}_{s,1}\hat{\Omega}_d\Psi_{g',d}
-  - D_g\nabla\Psi_{g,d}\right)}{\sum^{N_d}_{d=1}w_d\Psi_{g,d}}, \\
+  - D_g\nabla\Psi_{g,d}\right)}{\sum^{N_d}_{d=1}w_d\Psi_{g,d}}, \label{eq:drift} \\
   \gamma_g =
   \frac{\sum_{\hat{\Omega}_d\cdot\hat{n}_b > 0}w_d |\hat{\Omega}_d\cdot\hat{n}_b |
-  \Psi_{g,d}}{\sum^{N_d}_{d=1}w_d\Psi_{g,d}}.
-\end{gather}
-
-%\subsection{Additive Correction Term}
-%
-%Lastly, we can also incorporate transport corrections into the neutron diffusion equation through
-%additional additive terms.
-%We can obtain an additive correction term formulation from the \textit{Second-Moment}
-%method by Lewis \& Miller \cite{lewis_comparison_1976}. The literature review in Section
-%\ref{sec:summary-nts-mtds} covers how Anistratov \& Stehle \cite{anistratov_computational_2012}
-%used this formulation for their hybrid method with domain decomposition. Refer to Eqs.\
-%\ref{eq:2nd-moment-1}, \ref{eq:2nd-moment-2}, and \ref{eq:2nd-moment-3} for mathematical
-%formulations. After substituting Eq.\ \ref{eq:2nd-moment-2} into Eq.\ \ref{eq:2nd-moment-1} to
-%eliminate $J^{(s+1)}$, we have a modified neutron diffusion equation for $\phi^{(s+1)}$ with an
-%additional correction term $\nabla\cdot\nabla\cdot F^{(s+1/2)}$, where $(s)$ denotes the
-%iteration index for iterating between the high- and low-order equations.
-%
-%Unlike the previous two transport correction formulations, the additive correction term is an
-%additive closure that does not scale with the flux or the flux gradient in the low-order equation.
-%Therefore, for implementations of \gls{HOLO} or transport correction methods in which the high- and
-%low-order scalar flux solutions differ, additive correction terms may not provide the correction.
+  \Psi_{g,d}}{\sum^{N_d}_{d=1}w_d\Psi_{g,d}}. \label{eq:bound-coef}
+\end{gather}
 
@@
 
-In order to reduce the computational cost of the high-level $S_N$ calculation, I propose reducing
+As described in the previous section, transport correction formulations are commonly used in
+iterative acceleration schemes or hybrid methods to couple high-order neutron transport
+calculations with low-order neutron diffusion calculations. However, the high-order calculations
+remain computationally intensive and limit their applicability to time-independent neutronics
+simulations on \gls{HPC} clusters.
+In order to reduce the computational cost of the high-level $S_N$ calculation in a reactor
+simulation, I propose reducing
 the problem domain of the $S_N$ method to a \textit{correction region} containing the control rod
-and its vicinity. Consequently, the hybrid $S_N$-diffusion method can retain accurate neutron flux
+and its vicinity. Consequently, the hybrid $S_N$-diffusion method may retain accurate neutron flux
 and current estimates around the control rod region from the $S_N$ method while making significant
@@
 domain and the correction region as $V_0$ and $V_1$, respectively, where
-$V_1\subseteq V_0$. The algorithm for the hybrid $S_N$-diffusion method is as follows:
+$V_1\subseteq V_0$. The iterative algorithm for the hybrid $S_N$-diffusion method is as follows:
 %
 \begin{enumerate}
-  \item Start with an initial neutron diffusion calculation in $V_0$ with conventional
-    $P_1$-based diffusion coefficients and other standard group constants (e.g., neutron cross
-    sections).
-  \item Use the neutron diffusion flux estimates in $V_1$ and current estimates along
-    $\partial V_1$ as initial and boundary conditions for the $S_N$ sub-solver.
-  \item With the $S_N$ sub-solver, calculate an improved neutron flux solution in $V_1$,
-    which contains the control rod region and its immediate vicinity.
-  \item Calculate \glspl{SVDC} using the $S_N$ flux solution and Eq.\ \ref{eq:svdc}.
-  \item Pass the \glspl{SVDC} to the neutron diffusion solver to replace the conventional
-    $P_1$ diffusion coefficients within the $V_1$ while continuing to use $P_1$-based
-    diffusion coefficients in the rest of $V_0$.
-  \item Start the next iteration by running neutron diffusion calculation with the \glspl{SVDC}
-    replacing $P_1$ diffusion coefficients in part or all of $V_1$.
-  \item Repeat Steps 2-6 until convergence is reached by meeting pre-defined convergence tolerance
+  \item Start with an initial neutron diffusion calculation in $V_0$ using the standard neutron
+    diffusion method.
+  \item Pass the neutron diffusion current estimates along
+    $\partial V_1$ to the $S_N$ sub-solver to evaluate the boundary conditions for the $S_N$
+    subproblem.
+  \item With the $S_N$ sub-solver, evaluate transport correction terms in $V_1$ using Eq.
+    \ref{eq:svdc} or \ref{eq:drift}.
+  \item Pass the transport correction terms to the neutron diffusion solver to apply corrections
+    within $V_1$ while continuing to apply the standard neutron diffusion solver
+    in the rest of $V_0$.
+  \item Start the next iteration by running a neutron diffusion calculation with transport
+    corrections in $V_1$.
+  \item Repeat Steps 2-6 until convergence is reached by meeting desired convergence tolerance
     values.
@@
     \node (2) [process, below of=1, yshift=-1.8cm]
-      {\textbf{Calculate incident flux boundary conditions along\\$\bm{\partial V_1}$ for $S_N$ calculation}};
+      {\textbf{Calculate incident flux boundary conditions along\\$\bm{\partial V_1}$ for $\bm{S_N}$ calculation}};
     \node (3) [process, below of=2, yshift=-1.5cm]
-      {\textbf{Perform $\bm{S_N}$ calculation\\in $\bm{V_1}$ for $\bm{\phi^{tr}_i}$ \& $\bm{J^{tr}_i}$}};
+      {\textbf{Perform $\bm{S_N}$ calculation\\in $\bm{V_1}$ for $\bm{\phi^{tr,(i)}_g}$ \& $\bm{J^{tr,(i)}_g}$}};
     \node (4) [process, below of=3, yshift=-1.5cm]
-      {\textbf{Calculate $\bm{D^s_i}$ using Eq.\ \ref{eq:svdc}}};
+      {\textbf{Evaluate transport corrections, $\bm{D^{s,(i)}_g}$ or $\bm{\vec{D}^{(i)}_g}$}};
     \node (5) [process, below of=4, yshift = -1.5cm]
-      {\textbf{Perform neutron diffusion\\calculation in $\bm{V_0}$ using $\bm{D^s_i}$ in $\bm{V_1}$}};
+      {\textbf{Perform neutron diffusion\\calculation in $\bm{V_0}$ with transport\\corrections in $\bm{V_1}$}};
     \node (6) [decision, below of=5, yshift = -1.8cm]
@@
       {\textbf{Final $\bm{\phi}$ and $\bm{k}$\\solutions obtained}};
-    \draw [arrow] (1) -- node[anchor=east] {$\bm{\phi_0, J_0}$} (2);
-    \draw [arrow] (2) -- node[anchor=east] {$\bm{\Psi^-}$ \textbf{along} $\bm{\partial V_1}$} (3);
-    \draw [arrow] (3) -- node[anchor=east] {$\bm{\phi^{tr}_i, J^{tr}_i}$} (4);
-    \draw [arrow] (4) -- node[anchor=east] {$\bm{D^s_i}$} (5);
-    \draw [arrow] (5) -- node[anchor=east] {$\bm{\phi_i, J_i$}} (6);
+    \draw [arrow] (1) -- node[anchor=east] {$\bm{\phi^{(0)}_g, J^{(0)}_g}$} (2);
+    \draw [arrow] (2) -- node[anchor=east] {$\bm{\Psi^-_g}$ \textbf{along} $\bm{\partial V_1}$} (3);
+    \draw [arrow] (3) -- node[anchor=east] {$\bm{\phi^{tr,(i)}_g, J^{tr,(i)}_g}$} (4);
+    \draw [arrow] (4) -- node[anchor=east] {$\bm{D^{s,(i)}_g}$ or $\bm{\vec{D}^{(i)}_g}$} (5);
+    \draw [arrow] (5) -- node[anchor=east] {$\bm{\phi^{(i)}_g, J^{(i)}_g}$} (6);
     \draw [arrow] (6) -- node[anchor=east] {\textbf{Yes}} (7);
@@
   \end{tikzpicture}
-  \caption{Algorithm flowchart for the hybrid $S_N$-diffusion method.}
+  \caption{Algorithm flowchart for the hybrid $S_N$-diffusion method. $(i)$ denotes the iteration
+  index.}
   \label{fig:algorithm}
@@
 
-%To resolve this issue, I posit the following hypothesis: \textit{If there exists a highly
-%neutron-absorbing control rod region within the $S_N$ subproblem domain $V_1$ and the $S_N$
-%subproblem boundary $\partial V_1$ lies several neutron mean free paths away from this
-%region, the \gls{SVDC} values calculated near the control rod region (using the $S_N$ sub-solver
-%with isotropic hemisphere boundary conditions) will tend to the actual flux gradient solution
-%(using a reference $S_N$ calculation across the entire problem domain $V$).} In other words,
-%suboptimal boundary conditions may induce inaccurate \gls{SVDC} values near the $\partial
-%V_1$ subdomain boundaries, but the \gls{SVDC} values further from $\partial V_1$ are
-%accurate and very weakly dependent on the boundary conditions. These inaccurate \gls{SVDC} values
-%close to $V_1$ may be discarded in favor of the default $P_1$-based diffusion coefficients.
-%I will refer to the region containing these discarded values as the \textit{dead zone} or
-%$V_2$, where $V_2 \subset V_1 \subseteq V_0$.
-
-\subsubsection{Hybrid Method Verification with Case 0}
-
-In this subsection, I revisit Case 0
-(Figure \ref{fig:case-0-geom}) to test my hypothesis and demonstrate the hybrid $S_N$-diffusion
-method. Once more, I defer general numerical implementation details to
-Section \ref{sec:implementation}.
-
-In Section \ref{sec:svdc}, the \gls{SVDC} generated from the reference $S_8$ calculation exhibited
-significant spatial variations attributed to the presence of the control rod, as shown in Fig
-\ref{fig:c0diffcoef}. The group 1 and 2 \gls{SVDC} values approach asymptotic values near the
-corresponding group 1 and 2 $P_1$-based diffusion coefficient values near $x=15$ cm and $x=5$ cm,
-respectively. Therefore, the $V_1$ should, at minimum, cover the domain between $x=0$ cm and
-$x=15$ cm. Allowing for a buffer region for inaccurate \glspl{SVDC} near $\partial V_1$, I
-defined $V_1$ to span from $x=0$ cm to $x=17.5$ cm.
-
-I set up the hybrid method to automatically determine the cutoff point, beyond which \glspl{SVDC}
-in $V_1$ are discarded, based on the \gls{SVDC} distribution generated in every iteration.
-Starting from the interface between the control rod and fuel-graphite region, the algorithm sweeps
-rightward for the point at which the group-wise \gls{SVDC} distribution approaches within 1\% of
-the $P_1$-based diffusion coefficient value. This point serves as the cutoff point. The hybrid
-method converges rapidly within two outer iterations comprising three neutron diffusion and two
-$S_8$ calculations in total as outlined by the hybrid method algorithm; the change in $k$ between
-the second and third outer iterations is approximately $10^{-7}$.
-%
-\begin{figure}[htb!]
-  \centering
-  \begin{subfigure}[b]{.49\textwidth}
-    \centering
-    \includegraphics[width=\textwidth]{case-0-group-1-hybrid-diffcoef}
-    \caption{Group 1}
-    \label{fig:c0g1hd}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}[b]{.49\textwidth}
-    \centering
-    \includegraphics[width=\textwidth]{case-0-group-2-hybrid-diffcoef}
-    \caption{Group 2}
-    \label{fig:c0g2hd}
-  \end{subfigure}
-  \caption{$P_1$-based flux-limited diffusion coefficient and \gls{SVDC} spatial distribution for
-  Case 0. The \gls{SVDC} distributions were generated from the reference $S_8$ (``SVDC'') and the
-  hybrid (``Hybrid'') calculations.}
-  \label{fig:c0hd}
-\end{figure}
-
-Refer to Figures \ref{fig:c0g1hd} and \ref{fig:c0g2hd} for a qualitative comparison of group 1 and
-2 $P_1$-based diffusion coefficients, \glspl{SVDC} derived from the reference $S_8$ calculation as
-demonstrated in Section \ref{sec:svdc}, and \glspl{SVDC} derived from the hybrid $S_N$-diffusion
-method discussed here. The \glspl{SVDC} from the hybrid method agree well with the reference
-\glspl{SVDC} in the control rod region and for most of the fuel-graphite region up to around
-$x=15$ cm. Both sets of \glspl{SVDC} approximately coincide with the $P_1$ diffusion coefficient
-values within 1\% difference from $x=14.615$ cm and $x=3.390$ cm onwards for group 1 and
-group 2, respectively. Accordingly, the buffer region spans from $x=14.615$ cm to $x=17.500$ cm for
-group 1 and $x=3.390$ cm to $x=17.500$ cm for group 2, in which the hybrid method defaults to the
-$P_1$-based diffusion coefficients. Given that the reference \gls{SVDC} values will generally not
-be known in real-world problems, the hybrid method relies on finding the intersection point of the
-\gls{SVDC} and $P_1$ diffusion coefficient values to determine the buffer region cutoff point. The
-size of $V_1$ will also have to be based on where this intersection point lies.
-Another important consideration is the fact that the neutron flux gradient will be discontinuous if
-the \gls{SVDC} and $P_1$ diffusion coefficient values are too far apart in magnitude at the buffer
-region cutoff point. A flux gradient discontinuity in a homogeneous bulk region is non-physical and
-unacceptable. In Section \ref{sec:prelim-results}, I study \gls{SVDC} distribution trends
-in more heterogeneous systems and their implications on determining the size of $V_1$ and
-the cutoff point.
-%
-\begin{figure}[htb!]
-  \centering
-  \begin{subfigure}[b]{.49\textwidth}
-    \centering
-    \includegraphics[width=\textwidth]{case-0-group-1-hybrid-flux}
-    \caption{Group 1}
-    \label{fig:c0g1hf}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}[b]{.49\textwidth}
-    \centering
-    \includegraphics[width=\textwidth]{case-0-group-2-hybrid-flux}
-    \caption{Group 2}
-    \label{fig:c0g2hf}
-  \end{subfigure}
-  \caption{Neutron group 1 and 2 flux distributions from the diffusion, $S_8$, reference
-  \gls{SVDC}, and hybrid methods for Case 0.}
-  \label{fig:c0hf}
-  \centering
-  \begin{subfigure}[b]{.49\textwidth}
-    \centering
-    \includegraphics[width=\textwidth]{case-0-group-1-hybrid-flux-diff}
-    \caption{Group 1}
-    \label{fig:c0g1hfdiff}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}[b]{.49\textwidth}
-    \centering
-    \includegraphics[width=\textwidth]{case-0-group-2-hybrid-flux-diff}
-    \caption{Group 2}
-    \label{fig:c0g2hfdiff}
-  \end{subfigure}
-  \caption{Difference in neutron group 1 and 2 flux distributions from the diffusion,
-  diffusion-\gls{SVDC}, and hybrid methods with respect to the $S_8$ flux distribution for Case 0.}
-  \label{fig:c0hfdiff}
-\end{figure}
-
-Figures \ref{fig:c0g1hf} and \ref{fig:c0g2hf} show the group 1 and 2 neutron flux distributions
-from the various methods, while Figures \ref{fig:c0g1hfdiff} and \ref{fig:c0g2hfdiff} show the
-difference in flux with respect to the $S_8$ flux distribution. As
-expected following the diffusion coefficient discussion, the hybrid method flux distribution
-matches the $S_8$ and Diffusion-\gls{SVDC} flux distributions well. The $k$ estimate from the
-hybrid method is 0.62774, which is only 0.00024 higher than the Diffusion-\gls{SVDC} method, and
-0.00038 higher than the $S_8$ method.
-
-In Case 0, $V_1$ covers more than half of $V_0$ due to the control rod region being
-the only significant source of influence on the neutron flux distribution in the
-otherwise homogeneous system with reflective boundary conditions. I chose Case 0 as a simple test
-case to aid in introducing the hybrid $S_N$-diffusion method. In Section \ref{sec:prelim-results},
-I present more results with the hybrid method on more complicated geometries containing neutron
-reflector, air, and heterogeneous fuel-graphite lattice regions and vacuum boundary conditions. In
-these other test cases, the hybrid method yields smaller $V_1$-to-$V_0$ ratios,
-reducing computational costs from running $S_N$ calculations over a relatively smaller
-region.
-
-\section{Numerical Implementation} \label{sec:implementation}
-
-In Section \ref{sec:theory}, I presented the theoretical basis for the hybrid
-$S_N$-diffusion method and demonstrated the method on a simple problem. Now I will discuss
-numerical implementation details of group
-constant data processing and the neutron diffusion, $S_N$ neutron transport, and hybrid
-$S_N$-diffusion solvers. I implemented all numerical solvers and analysis scripts in the Python
-programming language. First, I will discuss the material group constant data generation and
-postprocessing steps. Next, I present the implementation details of the neutron
-diffusion and $S_N$ numerical methods. Lastly, I present how they are coupled to form the hybrid
-$S_N$-diffusion method.
-
-\subsection{Group Constant Data Generation}
-
-The group constants required by either or both neutron diffusion and $S_N$ neutron transport
-methods are:
-%
-\begin{itemize}
-  \item $\Sigma_{t,g}$: Macroscopic total cross section in group $g$,
-  \item $\Sigma_{r,g}$: Macroscopic removal cross section in group $g$,
-  \item $\Sigma_s^{g'\rightarrow g}$: Macroscopic group-to-group scattering cross section matrix,
-  \item $\Sigma_{s,l}^{g'\rightarrow g}$: $l$-th Legendre moment of the macroscopic
-    group-to-group scattering cross section matrix,
-  \item $\Sigma_{sp,l}^{g'\rightarrow g}$: $l$-th Legendre moment of the macroscopic
-    group-to-group scattering production cross section matrix,
-  \item $D_g$: $P_1$-based diffusion coefficient in group $g$,
-  \item $\nu\Sigma_{f,g}$: Product of the average number of neutrons produced per fission and the
-    macroscopic fission cross section in group $g$,
-  \item $\chi_g$: Neutron fission spectrum in group $g$.
-\end{itemize}
-%
-These group constants are generated using OpenMC's multigroup cross section generation capability
-\cite{boyd_multigroup_2019} and postprocessed using a Python script into JSON format files.
-$\Sigma_{r,g}$ is the only quantity that OpenMC does not directly provide. It is calculated as
-%
-\begin{align}
-  \Sigma_{r,g} =& \sum^G_{g'\neq g}\Sigma_s^{g\rightarrow g'}+\Sigma_{a,g}-\left(\Sigma_{sp}^{g
-    \rightarrow g} - \Sigma_s^{g\rightarrow g}\right)
-  \shortintertext{where}
-      \Sigma_{a,g} =& \mbox{ macroscopic absorption cross section in group $g$,} \nonumber \\
-      \Sigma_{sp}^{g\rightarrow g} =& \mbox{ macroscopic scattering production cross section from
-      group $g$ to $g$.} \nonumber
-\end{align}
-%
-$\Sigma_{r,g}$ primarily represents the loss of neutrons from group $g$ through outscattering and
-absorption. $\Sigma_{sp}^{g\rightarrow g}$ incorporates neutron multiplication effects from neutron
-knockout reactions into the scattering cross section. Neutron knockout reactions are commonly
-tallied as scattering reactions, and $\Sigma_{r,g}$ is a convenient term to
-incorporate neutron knockout effects into the neutron diffusion equations. I ran all test cases for
-$S_N$ and hybrid method calculations with up to the 2nd Legendre moments of the scattering
-cross sections ($L=2$).
-
-OpenMC uses the $P_1$ flux-limited formulation \cite{pomraning_flux-limited_1984} for calculating
-$D_g$ as follows
-%
-\begin{align}
-  D_g =& \frac{1}{3\Sigma_{tr,g}},
-  \shortintertext{where}
-  \Sigma_{tr,g} =& \frac{\langle\Sigma_{t,g}\phi_g\rangle-\langle\Sigma_{s1,g}\phi_g\rangle}
-  {\langle\phi_g\rangle}, \nonumber \\
-  \langle\Sigma_{t,g}\phi_g\rangle =& \int_{r\in V}dr \int_{4\pi}d\Omega\int^{E_{g-1}}_{E_g}dE\
-  \Sigma_{t,g}(r,E)\Psi(r,E,\Omega), \nonumber \\
-  \langle\Sigma_{s1,g}\phi_g\rangle =& \int_{r\in V}dr \int_{4\pi}d\Omega\int^{E_{g-1}}_{E_g}dE
-  \int_{4\pi}d\Omega'\int^{\infty}_0dE'\int^1_{-1}d\mu\ \mu\Sigma_s(r,E'\rightarrow E,\Omega'\cdot
-  \Omega)\phi(r,E',\Omega'), \nonumber \\
-  \langle \phi \rangle =& \int_{r\in V}dr\int_{4\pi}d\Omega\int^{E_{g-1}}_{E_g}dE\ \Psi(r,E,\Omega)
-  .\nonumber
-\end{align}
-
-\subsection{Neutron Diffusion Method}
-
-On a 1-D uniform spatial grid with $I+1$ mesh points, the neutron scalar flux variables
-$\phi_{g,i}$ are defined on the mesh points $x_i$. Theoretically, group constants are volumetric
-material properties that should be defined on the cell-centered half-integer mesh points
-$x_{i+\sfrac{1}{2}}$. In practice, all material properties except diffusion coefficients are
-uniform in each subregion and sampled at $x_i$. To avoid ambiguity concerning diffusion
-coefficient sampling, I formulated all test cases such that all material interfaces fall on $x_i$.
-
-Discretizing the multigroup $k$-eigenvalue neutron diffusion equations in Eq.\ \ref{eq:1d-diff}
-and reformulating the scattering term using neutron balance in the control volume bounded by
-$x_{i-\sfrac{1}{2}}$ and $x_{i+\sfrac{1}{2}}$ yields
-%
-\begin{align}
-  J_{g,i+\sfrac{1}{2}} - J_{g,i-\sfrac{1}{2}} + \Sigma_{t,g,i} \phi_{g,i} \Delta x = \sum^G_{g'=1}\left[
-  \Sigma_{s,i}^{g'\rightarrow g}\phi_{g',i} + \chi_{g,i}\frac{\nu\Sigma_{f,g',i}}{k} \phi_{g',i}
-\right]\Delta x. \label{eq:diff-j}
-\end{align}
-%
-Using the diamond difference scheme to replace the $J$ terms with the discretized form of
-Fick's first law of diffusion,
-%
-\begin{align}
-  J_{g,i+\sfrac{1}{2}} = -D_{g,i+\sfrac{1}{2}}\frac{d\phi_{g,i+\sfrac{1}{2}}}{dx} =
-  -D_{g,i+\sfrac{1}{2}} \frac{\phi_{g,i+1}-\phi_{g,i}}{\Delta x},
-\end{align}
-%
-and rearranging the terms in Eq.\ \ref{eq:diff-j} yields
-%
-\begin{align}
-  -\frac{D_{g,i-\sfrac{1}{2}}}{\Delta x} \phi_{g,i-1} + &\left[\frac{D_{g,i-\sfrac{1}{2}}+
-  D_{g,i+\sfrac{1}{2}}}{\Delta x} + \Delta x\ \Sigma_{r,g,i} \right]\phi_{g,i} -
-  \frac{D_{g,i+\sfrac{1}{2}}}{\Delta x}\phi_{g,i+1} -\Delta x\sum^G_{g'\neq g}
-  \Sigma_{s,i}^{g'\rightarrow g}\phi_{g',i} \nonumber \\
-  =& \Delta x\sum^G_{g'=1}
-  \chi_{g,i} \frac{\nu\Sigma_{f,g',i}}{k} \phi_{g',i}, \label{eq:diff-fd}
-  \shortintertext{where}
-  \Sigma_{r,g} =& \Sigma_{t,g} - \Sigma_s^{g\rightarrow g} \nonumber \\
-  =& \mbox{ macroscopic removal cross section for neutron group }g. \nonumber
-\end{align}
-%
-The diamond difference scheme is 2nd-order accurate, and this form is
-equivalent to applying 2nd-order finite differencing to the original neutron diffusion equation in
-Eq.\ \ref{eq:1d-diff} with diamond differencing for the cell-centered group constants.
-The fixed neutron source $S_g$ is ignored here since the test cases are all neutron-multiplying
-systems with no fixed source.
-
-I implemented two types of boundary conditions: vacuum and reflective boundary conditions. The
-\textbf{vacuum boundary conditions} are imposed by setting the incoming flux in the $P_1$
-approximation to zero and applying 2nd-order finite differencing as follows:
-%
-\begin{align}
-  \mbox{Left boundary: } \frac{\phi_{g,0}}{4}-\frac{D_{g,\sfrac{1}{2}}}{2}
-  \frac{\left(-\phi_{g,2}+4\phi_{g,1}-3\phi_{g,0}\right)}{2\Delta x} =& 0 \\
-  \mbox{Right boundary: } \frac{\phi_{g,I}}{4}+\frac{D_{g,I-\sfrac{1}{2}}}{2}
-  \frac{\left(\phi_{g,I-2}-4\phi_{g,I-1}+3\phi_{g,I}\right)}{2\Delta x} =& 0.
-\end{align}
-%
-The \textbf{reflective boundary conditions} are imposed by setting the flux gradient to zero and
-applying 2nd-order finite differencing as follows:
-%
-\begin{align}
-  \mbox{Left boundary: } \frac{-3\phi_{g,0}+4\phi_{g,1}-\phi_{g,2}}{2\Delta x} =& 0 \\
-  \mbox{Right boundary: } \frac{3\phi_{g,I}-4\phi_{g,I-1}+\phi_{g,I-2}}{2\Delta x} =& 0
-\end{align}
-%
-At material interfaces, the continuity condition requires that the net neutron current on either
-side of the interface be equal as follows:
-%
-\begin{align}
-  -\frac{3D_{g,i-\sfrac{1}{2}} - D_{g,i-\sfrac{3}{2}} }{2}
-  \frac{\phi_{g,i-2}-4\phi_{g,i-1}+3\phi_{g,i}}{2\Delta x} =&
-  -\frac{3D_{g,i+\sfrac{1}{2}} - D_{g,i+\sfrac{3}{2}} }{2}
-  \frac{-3\phi_{g,i}+4\phi_{g,i+1}-\phi_{g,i}}{2\Delta x} \label{eq:itf-bc}
-\end{align}
-%
-for a material interface at $x_i$.
-
-Altogether, they form a system of equations of the form $\bm{A\overline{\phi}}=\bm{\frac{1}{k}
-B\overline{\phi}}$, where $\bm{\overline{\phi}}$ is a flattened vector representation of
-$\phi_{g,i}$, and $\bm{A}$ and $\bm{B}$ are matrices of the coefficients of $\phi_{g,i}$ as given
-by Eqs. \ref{eq:diff-fd} to \ref{eq:itf-bc}. I implemented the inverse power method to find $k$ and
-$\overline{\phi}$. The inverse power method algorithm is as follows
-%
-\begin{align}
-  \shortintertext{1. Initialize $k^0$ and $\bm{\overline{\phi}}^0$}
-  \shortintertext{2. Update $\bm{\overline{\phi}}$ and $k$}
-  \bm{\overline{\phi}}^m =& \frac{1}{k^{m-1}}\bm{A}^{-1}\bm{B\overline{\phi}}^{m-1} \\
-  k^m =& k^{m-1}\frac{|\bm{B\overline{\phi}}^m|}{|\bm{B\overline{\phi}}^{m-1}|}
-  \shortintertext{3. Check whether convergence is reached}
-  \epsilon_\phi =
-  \frac{|\bm{\overline{\phi}}^m-\bm{\overline{\phi}}^{m-1}|}{|\bm{\overline{\phi}}^m|} <& \
-  tol_{\bm{\overline{\phi}}} \\
-  \epsilon_k =
-  \frac{|k^m-k^{m-1}|}{|k^m|} <& \ tol_k
-  \shortintertext{4. Return to Step 2 if either expression is false, otherwise exit.} \nonumber
-\end{align}
-%
-$k^m$ and $\overline{\phi}^m$ denote estimates of $k$ and $\overline{\phi}$ after the $m$-th
-iteration. Matrix $\bm{A}$ is a primarily tridiagonal matrix with at most $G-1$ off-diagonal terms
-from the fourth term in Eq.\ \ref{eq:diff-fd}. Thus, $\bm{A}$ is initialized as a sparse matrix to
-take advantage of the computationally efficient sparse matrix solver functions from the
-\texttt{sparse} class of the \texttt{SciPy} \cite{virtanen_scipy_2020} Python library for
-scientific and technical computing. Matrix $\bm{B}$
-is never initialized explicitly as a matrix. Instead, the vector $\bm{b}=\bm{B\overline{\phi}}$ is
-updated directly in every iteration. In this system of equations, $|\bm{b}|$ corresponds to the
-total number of fission neutrons produced in the system for a given $\bm{\overline{\phi}}$. This
-quantity is calculated using the \texttt{trapezoid} numerical integration function from
-\texttt{SciPy} to estimate the integral value of the source term, $\nu\Sigma_{g,f}\phi_{g}$, in $x$
-from the discrete flux values in $\bm{x}$. Finally, the final $\phi_{g,i}$ is normalized by a
-factor of $|\bm{b}|/k$ (no. of source neutrons) to obtain the neutron scalar flux per source
-neutron.
-
-\subsection{$S_N$ Neutron Transport Method}
-
-For the 1-D $S_N$ neutron transport method on the same uniform spatial grid with $I+1$ mesh points,
-the neutron angular flux variables $\Psi_{g,i,n}=\Psi_g(x_i,\mu_n)$ are defined on the mesh points
-$x_i$ while neutron scalar flux variables $\phi_{g,i\pm\sfrac{1}{2}}=\phi_g(x_{i\pm\sfrac{1}{2}})$
-are defined on the half-integer mesh points $x_{i\pm\sfrac{1}{2}}$. All group constants are sampled
-at $x_{i\pm\sfrac{1}{2}}$.
-
-The $S_N$ equations are solved using the transport sweep method in which the algorithm ``sweeps''
-through the spatial grid and sequentially updates $\Psi_{g,i,n}$. The algorithm sweep
-direction follows the direction of neutron travel, i.e. it sweeps in the positive direction for
-$\Psi_{g,i,n}$ with $\mu_n>0$ and vice versa. Doing otherwise is unphysical and causes numerical
-instabilities.
-
-Discretizing the multigroup $S_N$ neutron transport equations in Eq.\ \ref{eq:1d-sn} about
-$x_{i+\sfrac{1}{2}}$ yields
-%
-\begin{align}
-  \mu_n\frac{\Psi_{g,i+1,n}-\Psi_{g,i,n}}{x_{i+1}-x_i} + \Sigma_{t,g,i+\sfrac{1}{2}}
-  \Psi_{g,i+\sfrac{1}{2},n} = q_{g,i+\sfrac{1}{2},n}
-\end{align}
-%
-where $q_{g,i+\sfrac{1}{2},n}$ represents the combined scattering and fission neutron source term.
-After expressing $\Psi_{g,i+\sfrac{1}{2}}$ as the average of $\Psi_{g,i+1,n}$ and $\Psi_{g,i,n}$
-in the diamond difference scheme and rearranging the terms, we obtain
-%
-\begin{align}
-  \Psi_{g,i+1,n} =& \frac{1-\Sigma_{t,g}\Delta x/2\mu_n}{1+
-    \Sigma_{t,g}\Delta x/2\mu_n}\Psi_{g,i,n} +
-    q\frac{\Delta x}{\mu_n\left(1+\Sigma_{t,g}\Delta x/2\mu_n\right)} \label{eq:sweep-right} &&
-    (\mbox{for } \mu_n > 0)
-\shortintertext{and}
-  \Psi_{g,i,n} =& \frac{1+\Sigma_{t,g}\Delta x/2\mu_n}{1-
-    \Sigma_{t,g}\Delta x/2\mu_n}\Psi_{g,i+1,n} -
-    q\frac{\Delta x}{\mu_n\left(1-\Sigma_{t,g}\Delta x/2\mu_n\right)}. \label{eq:sweep-left} &&
-    (\mbox{for } \mu_n < 0)
-\end{align}
-%
-The remaining $i+\sfrac{1}{2}$ indexes on the group constants are dropped to reduce visual clutter.
-These expressions are used to update $\Psi_{g,i,n}$ in the forward and backward transport sweeps.
-
-The discretized scattering and fission terms in $q_{g,i+\sfrac{1}{2},n}$ are given as
-%
-\begin{align}
-  q_{g,i+\sfrac{1}{2},n} =& \sum^G_{g'=1}\sum^L_{l=0}\frac{\left(2l+1\right)}
-  {2}\Sigma_{s,l}^{g'\rightarrow g}P_l(\mu_n)\phi_{l,g',i+\sfrac{1}{2}} \nonumber \\
-  &+\frac{\chi_g}{2}\sum^G_{g'=1}\frac{\nu\Sigma_{f,g'}}{k}\phi_{0,g',i+\sfrac{1}{2}}
-  \label{eq:sn-q}
-  \shortintertext{where}
-  \phi_{l,g',i+\sfrac{1}{2}} =& \sum^N_{n'=1}w_{n'}P_l(\mu_{n'})\frac{\Psi_{g',i,n'}+
-  \Psi_{g',i+1,n'}}{2}. \label{eq:phi-l}
-\end{align}
-%
-$\phi_{l,g,i}$ are $l$-th Legendre expansions of the neutron flux evaluated using Gauss-Legendre
-quadrature over $\mu_{n'}=[-1,1]$. $\phi_{0,g,i}$ and $\phi_{1,g,i}$ also correspond to the neutron
-scalar flux $\phi_{g,i}$ and net current $J_{g,i}$, respectively.
-
-For vacuum boundary conditions, $\Psi_{g,0,n}$ is zero for all positive $\mu_n$ while
-$\Psi_{g,I,n}$ is zero for all negative $\mu_n$ as follows
-%
-\begin{align}
-  \mbox{Left boundary: } \Psi_{g,0,n} =& 0 && (\mbox{for } \mu_n > 0) \\
-  \mbox{Right boundary: } \Psi_{g,I,n} =& 0 && (\mbox{for } \mu_n < 0)
-\end{align}
-%
-Reflective boundary conditions are imposed by equating the incoming angular flux to the
-outgoing angular flux in the opposite direction as follows
-%
-\begin{align}
-  \mbox{Left boundary: } \Psi_{g,0,n} =& \Psi_{g,0,n'} && (\mbox{for } \mu_n > 0, \mu_n =
-  -\mu_{n'}) \\
-  \mbox{Right boundary: } \Psi_{g,I,n} =& \Psi_{g,I,n'} && (\mbox{for } \mu_n < 0, \mu_n =
-  -\mu_{n'})
-\end{align}
-
-The power iteration algorithm for the $S_N$ method is similar to the inverse power method algorithm
-applied in the neutron diffusion method. The transport sweep step replaces the matrix-solving step
-for updating $\overline{\phi}$.
-%
-\begin{align}
-  \shortintertext{1. Initialize $k^0$, $\phi_{l,g,i+\sfrac{1}{2}}^0$, and $q^0$}
-  \shortintertext{2. Apply transport sweeps to solve for $\Psi^m$ using Eqs. \ref{eq:sweep-right},
-  \ref{eq:sweep-left}, and \ref{eq:sn-q}}
-  \shortintertext{3. Update $\phi^m$ and $k^m$ using Eqs. \ref{eq:phi-l} and \ref{eq:sn-k}}
-  k^m =& k^{m-1}\frac{\sum^I_{i=0}\sum^G_{g=1}\nu\Sigma_{f,g}\phi^m_{g,i+\sfrac{1}{2}}}
-  {\sum^I_{i=0}\sum^G_{g=1}\nu\Sigma_{f,g}\phi^{m-1}_{g,i+\sfrac{1}{2}}} \label{eq:sn-k}
-  \shortintertext{4. Check whether convergence is reached}
-  \epsilon_\phi =
-  \frac{|\bm{\overline{\phi}}^m-\bm{\overline{\phi}}^{m-1}|}{|\bm{\overline{\phi}}^m|} <& \
-  tol_{\bm{\overline{\phi}}} \\
-  \epsilon_k =
-  \frac{|k^m-k^{m-1}|}{|k^m|} <& \ tol_k
-  \shortintertext{5. Return to Step 2 if either expression is false, otherwise exit.} \nonumber
-\end{align}
-
-The transport sweep and $\phi^m$-update algorithms are parallelized using the \texttt{joblib}
-parallel computing Python library \cite{noauthor_joblib_nodate} across all available CPU threads.
-The total computational work is subdivided into several smaller tasks classified by unique pairs
-of $g$ and $n$ values; each thread computes all $\Psi^m$ or $\phi^m$ values on the mesh for a given
-pair of indexes $g$ and $n$.
-
-\subsection{Hybrid $S_N$-Diffusion Method}
-
-Without loss of generality, consider a 1-D system symmetric about the left boundary at $x_0=0$ cm.
-In this case, as with Case 0, the control rod region is the left-most region, followed by other
-regions. The $S_N$ subproblem domain $V_1$, containing the control rod and some region beyond it,
-is bounded by $x_0$ and $x_j$ for some $x_j$
-located several mean free paths to the right of the control rod region as governed by the relevant
-discussion in Section \ref{sec:hybrid-method}.
-
-Excluding the initial neutron diffusion calculation to initialize $k^0$ and $\phi^0$, each outer
-iteration in the hybrid method consists of one $S_N$ neutron transport calculation and one neutron
-diffusion calculation. The $\phi$ estimates from these $S_N$ transport and neutron diffusion
-calculations are labeled as $\phi^{m+\sfrac{1}{2}}$ and $\phi^{m+1}$ during the
-$(m+1)$-th outer iteration. $\phi^{m+\sfrac{1}{2}}$ spans $V_1$ while $\phi^{m+1}$ spans
-the entire domain $V_0$.
-
-For the $S_N$ subproblem boundary conditions, we discretize Eqs. \ref{eq:p1-j} and
-\ref{eq:sn-psi-j} as follows
-%
-\begin{align}
-  \Psi_{g,j,n} =& \frac{J_{g,-}(x_j)}{\sum^{N/2}_{n'=1}w_{n'}\mu_{n'}} \nonumber \\
-  =& \frac{\frac{\phi_g(x_j)}{4}+
-  \frac{D_g(x_{j-\sfrac{1}{2}})}{2}\frac{d\phi_g(x_j)}{dx}}{\sum^{N/2}_{n'=1}w_{n'}\mu_{n'}}
-  \nonumber \\
-  =& \frac{\frac{\phi_{g,j}}{4}+
-  \frac{D_{g,j-\sfrac{1}{2}}}{2}\frac{\phi_{g,j}-\phi_{g,j-1}}{\Delta x}}
-    {\sum^{N/2}_{n'=1}w_{n'}\mu_{n'}} && (\mu_n < 0) \label{eq:sn-bc}
-\end{align}
-%
-where $D_g$ is the $P_1$-based diffusion coefficient value.
-
-The \gls{SVDC} formulation in Eq.\ \ref{eq:svdc} is discretized using the diamond difference scheme
-as follows
-%
-\begin{align}
-  D^s_{g,i+\sfrac{1}{2}} =& -\frac{J^{tr}_{g,i+1}+J^{tr}_{g,i}}{2} \left(
-  \frac{\phi^{tr}_{g,i+1}-\phi^{tr}_{g,i}}{\Delta x} \right)^{-1} = -\frac{\Delta x}{2}
-  \frac{J^{tr}_{g,i+1}+J^{tr}_{g,i}}{\phi^{tr}_{g,i+1}-\phi^{tr}_{g,i}}. \label{eq:svdc-num}
-\end{align}
-%
-%From Eq.\ \ref{eq:svdc-num}, we note that numerical instabilities may occur near flux peaks where
-%the flux gradient and thus the denominator of Eq.\ \ref{eq:svdc-num} approach zero. On the other
-%hand, this observation also implies that diffusion coefficient values have negligible influence on
-%the flux solution near flux peaks due to the small flux gradient values. Therefore, we can avoid
-%numerical instabilities by applying the following logic after calculating $D^s_g$: If the ratio of
-%$|\frac{d\phi_g}{dx}|$ to $|\phi_g|$ is sufficiently small and $|D^s_g-D_g| > D_g$, the hybrid
-%method defaults to using $D_g$ instead of $|D^s_g-D_g|$. The first conditional statement locates
-%regions of relatively flat flux, and of these regions, the second conditional statement excludes
-%regions which naturally experience flat flux such as air-filled regions. In all test cases,
-%applying the first conditional as $|\frac{1}{\phi_g}\frac{d\phi_g}{dx}|<5\times 10^{-2}$ was
-%sufficient for this task.
-
-The hybrid $S_N$-diffusion algorithm is as follows
-%
-\begin{align}
-  \shortintertext{1. Initialize $k^0$ and $\phi^0$ with an initial neutron diffusion calculation on
-  $V_0$}
-  \shortintertext{2. Generate estimates for $\Psi_{g,j,n}$ at $x_j$ for the $S_N$ transport
-  calculation boundary conditions using Eq.\ \ref{eq:sn-bc}}
-  \shortintertext{3. Calculate $\phi^{m+\sfrac{1}{2}}$ with the $S_N$ transport sub-solver on
-  $V_1$}
-  \shortintertext{4. Generate \glspl{SVDC} with $\phi^{m+\sfrac{1}{2}}$ and $J^{m+\sfrac{1}{2}}$
-  using Eq.\ \ref{eq:svdc-num}}
-  \shortintertext{5. Calculate $\phi^{m+1}$ with the newly generated \glspl{SVDC} and the neutron
-  diffusion solver on $V_0$}
-  \shortintertext{6. Check whether convergence is reached}
-  \frac{|\bm{\overline{\phi}}^m-\bm{\overline{\phi}}^{m-1}|}{|\bm{\overline{\phi}}^m|} <& \
-  tol_{\bm{\overline{\phi}}} \\
-  \frac{|k^m-k^{m-1}|}{|k^m|} <& \ tol_k
-  \shortintertext{7. Return to Step 2 if either expression is false, otherwise exit.} \nonumber
-\end{align}
-%
-Generally, the convergence tolerance values for the outer hybrid method iteration must be smaller
-than the tolerance values for the neutron diffusion and $S_N$ transport inner iterations. Using
-$\phi^m$ from the neutron diffusion calculation as initial conditions for the $S_N$ transport
-calculation in the $(m+1)$-th iteration helps to reduce the number of transport
-sweeps required significantly.
-
-The $S_N$ sub-solver only covers $V_1$, so it does not calculate an independent $k$
-estimate. Instead, the $k^m$ from the neutron diffusion calculation scales the fission neutron
-source term in the $S_N$ sub-solver.
-
-%\section{Description of 1-D Test Cases} \label{sec:test-case}
-%
-%I designed ten 1-D test cases with increasing complexity to test the performance of the hybrid
-%$S_N$-diffusion method in response to various geometrical features. The last two cases resemble the
-%reference \gls{MSRE} design \cite{robertson_msre_1965}, which has centrally located control rods
-%and air-filled rod guide tubes. Figure \ref{fig:case-geom} shows the geometries of Cases 0 to
-%5b. All geometries have reflective boundary conditions at $x=0$ cm to reduce computational costs by
-%creating half-core or repeating unit cell models. Cases 1a, 2a, 3a, and 4a are repeating unit
-%cell models with reflecting boundaries on the right-side boundaries. Cases 1b, 2b, 3b, 4b, 5a, and
-%5b are half-core models with vacuum boundaries on the right-side boundaries.
-%
-%Starting with an infinite, homogeneous region in Case 1a, I systematically added geometric
-%complexities to each successive test case to identify how each feature impacts the neutronics
-%results. Case 1b introduces a reflector region and vacuum boundary conditions relative to Case 1a.
-%Cases 2a and 2b introduce a control rod region between $x=0$ cm and $x=0.5$ cm relative to Cases 1a
-%and 1b. Cases 3a and 3b introduce a 0.5 cm-thick air gap between the control rod and the
-%fuel-graphite mixture regions relative to Cases 2a and 2b. For Cases 4a and 4b, I replaced the
-%homogeneous fuel-graphite mixture in Cases 1a and 2b with explicit fuel-graphite lattices, as shown
-%in Figure \ref{fig:case-geom}. Case 5a introduces an air gap between the control rod and the
-%fuel-graphite lattice regions. Lastly, Case 5b replaces the control rod region with an extended
-%air gap region to provide a base case for the control rod worth calculations relative to Case 5a.
-%
-%The cases that do not contain control rod regions (1a, 1b, 4a, 4b, and 5b) serve as reference cases
-%to quantify other sources of discrepancies in the neutron diffusion solution (e.g., geometrical
-%heterogeneities, vacuum boundary condition approximation for neutron diffusion) and highlight how
-%much the neutron diffusion method underperforms when the problem includes control rod regions.
-%
-%\begin{figure}[htb!]
-%  \centering
-%  \includegraphics[width=\columnwidth]{case-geometry}
-%  \caption{Geometries of the 1-D test cases. The material labeled ``mixture'' represents a
-%    homogeneous mixture of fuel and graphite at a ratio of 22.5\%-77.5\% by volume. All geometries
-%    have reflective boundary conditions on the boundary at $x=0$ cm. The right-side boundaries are
-%    reflective for Cases 1a, 2a, 3a, and 4a, and vacuum for Cases 1b, 2b, 3b, 4b, 5a, and 5b.}
-%  \label{fig:case-geom}
-%\end{figure}
-%%
-%\begin{figure}[htb!]
-%  \centering
-%  \begin{subfigure}[t]{.49\textwidth}
-%    \centering
-%    \includegraphics[width=\textwidth]{spectrum}
-%    \caption{Neutron flux energy spectrum per source neutron}
-%    \label{fig:spectrum}
-%  \end{subfigure}
-%  \hfill
-%  \begin{subfigure}[t]{.49\textwidth}
-%    \centering
-%    \includegraphics[width=\textwidth]{reaction}
-%    \caption{Flux-normalized scattering and absorption reaction rates}
-%    \label{fig:reaction}
-%  \end{subfigure}
-%  \caption{Neutron flux energy spectrum and reaction rates for Case 5a. The dotted vertical lines
-%  correspond to the discrete neutron energy group boundaries.}
-%  \label{fig:spec-reac}
-%\end{figure}
-%%
-%%\begin{table}[tb!]
-%%  \centering
-%%  \caption{Description of the 1-D Test Case Geometries.}
-%%  \begin{tabular}{l c c c}
-%%    \toprule
-%%    Cases & Ordered list of regions & Ordered list of interface x-coordinates & Left \& Right BCs\\
-%%    \midrule
-%%    Case 0 & Control rod, homogenized fuel-graphite lattice &
-%%    \bottomrule
-%%  \end{tabular}
-%%  \label{table:c0k}
-%%\end{table}
-%
-%I ran all cases with four neutron energy groups. The list of energy group bounds are:
-%$E=[10^{-5}, 10^0, 10^2, 10^5, 10^8]$ eV. Figures \ref{fig:spectrum} and \ref{fig:reaction} show
-%how this energy group structure
-%partitions the neutron flux energy spectrum and neutron reaction rates. For this preliminary work,
-%the four-group structure is sufficient for showing different scattering and absorption trends of
-%slow, intermediate, and fast neutrons. I also ran all cases on OpenMC to generate the required
-%group constants and to assess the accuracy of the deterministic methods against OpenMC in
-%continuous-energy (OpenMC-CE) and multigroup (OpenMC-MG) modes. Table \ref{table:var} shows how the
-%position, direction of travel, neutron energy, and angle-dependence in $\Sigma_s$ are handled by
-%OpenMC and the deterministic methods. Comparing OpenMC-MG results with OpenMC-CE results allows us
-%to quantify errors arising from neutron energy group discretization and the scattering cross
-%section simplifications. Each Monte Carlo simulation ran on 25 inactive and 125 active cycles of
-%20,000 neutrons each, which sums up to 2.5 million tallied neutron histories.
-%
-%\begin{table}[tb!]
-%  \centering
-%  \footnotesize
-%  \caption{Variable handling in OpenMC under continuous-energy (OpenMC-CE) and multigroup
-%  (OpenMC-MG) modes, and in the $S_N$ neutron transport, neutron diffusion, and hybrid
-%  $S_N$-diffusion methods. }
-%  \begin{tabular}{c c c c c c}
-%    \toprule
-%    Variable & OpenMC-CE & OpenMC-MG & $S_N$ Transport & Diffusion & Hybrid \\
-%    \midrule
-%    Position, $\bm{r}$ & Continuous & Continuous & Discrete & Discrete & Discrete \\
-%    Direction of travel, $\bm{\hat{\Omega}}$ & Continuous & Continuous & Discrete & N/A & N/A \\
-%    Energy, $E$ & Continuous & Discrete & Discrete & Discrete & Discrete \\
-%    Angle-dependence in $\Sigma_s$ & Continuous & 2nd Legendre moment & 2nd Legendre moment
-%    & N/A & N/A \\
-%    \bottomrule
-%  \end{tabular}
-%  \label{table:var}
-%\end{table}
-%
-%\section{Simulation Parameters for Convergence} \label{sec:sim-param}
-%
-%The convergence of the results depends mainly on the following simulation parameters: mesh size
-%($\Delta x$), convergence tolerance ($tol$), number of discrete ordinates ($N$), and the highest
-%order moment of the scattering cross sections ($L$). The last two parameters apply only to $S_N$
-%calculations. Table \ref{table:param} shows the values of simulation parameters applied to each
-%method for the 1-D test cases.
-%
-%\begin{table}[tb!]
-%  \centering
-%  \footnotesize
-%  \caption{Simulation parameters applied to the neutron diffusion, $S_N$, and hybrid
-%  $S_N$-diffusion methods.}
-%  \begin{tabular}{c c c c}
-%    \toprule
-%    Parameter & $S_N$ Transport & Neutron Diffusion & Hybrid $S_N$-Diffusion \\
-%    \midrule
-%    Mesh size, $\Delta x$ [cm]        & 0.0125    & 0.0125    & 0.0125 \\[.2cm]
-%    \multirow{3}{*}{Convergence tolerance, $tol$}
-%                                      &           &           & Diffusion: $10^{-6}$ \\
-%                                      & $10^{-7}$ & $10^{-6}$ & $S_N$: $10^{-5}$ \\
-%                                      &           &           & Outer: $10^{-3}$ \\[.2cm]
-%    Discrete ordinates, $N$           & 8         & N/A       & 8 \\
-%    Highest moment of $\Sigma^{g'\rightarrow g}_s$, $L$
-%                                      & 2         & N/A       & 2 \\
-%    \bottomrule
-%  \end{tabular}
-%  \label{table:param}
-%\end{table}
-%
-%The simulation parameters did not show any significant correlation in the convergence tests. All
-%three methods achieve reasonable convergence with $\Delta x=0.0125$ cm; halving the mesh size
-%causes less than $30\times10^{-5}$ change in $k$ for the neutron diffusion and hybrid methods and
-%less than $10^{-5}$ for the $S_8$ method. Further refinement of the $tol$, $N$, and $L$ parameters
-%result in less than $2\times10^{-5}$
-%change in $k$. The same $tol$ value applies to convergence checks for both $\phi$ and $k$ in all
-%three methods because the variables converge at similar rates. The hybrid method requires three
-%separate $tol$ values due to the inner and outer iterations. Notably, the hybrid method can reach
-%convergence with a larger $tol$ value for its $S_N$ sub-solver than the $tol$ value for the
-%standalone $S_N$ solver, which implies that $J$ and $\sfrac{d\phi}{dx}$ converge faster than $\phi$
-%because the \glspl{SVDC} depend on those variables instead of $\phi$. Consequently, the $S_N$
-%sub-solver requires much fewer iterations than the standalone $S_N$ solver. In combination with
-%restricting $S_N$ calculations to the correction region, the hybrid method saves significant
-%computation time compared to the standalone $S_N$ method.
-%
-%The outer iterations
-%converge rapidly, so a $tol$ value of $10^{-3}$ is sufficient for convergence. The $\phi$ and $k$
-%relative error norms ($\epsilon$) in all test cases decreased by two orders of magnitude on
-%every outer iteration. For instance, Table \ref{table:epsilon} shows the $\epsilon$ values from the
-%hybrid method after every outer iteration for Case 5a. The $\epsilon$ values fall below
-%$10^{-5}$ by the 3rd and 4th iterations.
-%
-%\begin{table}[tb!]
-%  \centering
-%  \footnotesize
-%  \caption{Relative error norm, $\epsilon$, values for $k$ and $\phi$ after each outer iteration in
-%    the hybrid $S_N$-diffusion method for Case 5a.}
-%  \begin{tabular}{c S[table-format=1.1e+1] S[table-format=1.1e+1]}
-%    \toprule
-%    Iteration   & {$\epsilon_k$}    & {$\epsilon_\phi$} \\
-%    \midrule
-%    1           & 1.6e-2            & 1.5e-2 \\
-%    2           & 5.0e-5            & 4.7e-5 \\
-%    3           & 1.4e-7            & 1.6e-7 \\
-%    4           & 3.5e-9            & 1.4e-8 \\
-%    \bottomrule
-%  \end{tabular}
-%  \label{table:epsilon}
-%\end{table}
-%
-%
-%\section{Results \& Discussion} \label{sec:prelim-results}
-%
-%In this section, I compare neutronics parameters for all test cases calculated from
-%the OpenMC and deterministic methods. I refer to the OpenMC calculations under continuous-energy
-%and multigroup modes as OpenMC-CE and OpenMC-MG, respectively. I did not apply the hybrid method to
-%Cases 1a, 1b, 4a, 4b, and 5b, which serve as reference cases for how much the neutron diffusion
-%method underperforms when control rod regions are present in the other cases.
-%
-%\subsection{Comparison of Multiplication factors, $k$}
-%
-%Tables \ref{table:ck1} and \ref{table:ck2} show all test case $k$ estimates from the OpenMC and
-%deterministic methods. The tables also include statistical uncertainties for the OpenMC
-%$k$ estimates.
-%
-%All five methods generally show reasonable agreement within each test case. The differences between
-%the OpenMC-CE and OpenMC-MG $k$ estimates vary significantly from 0.00018 in Case 1a to 0.01622 in
-%Case 4b. These Monte Carlo methods exhibit more significant differences in cases
-%with vacuum boundary conditions (e.g., Cases 1b, 2b, 3b, 4b, 5a, 5b). These differences imply that
-%the neutron energy discretization scheme must be revised to capture neutron leakage rates
-%accurately. Beyond this preliminary study, I will explore better discrete energy group structures
-%for further neutronics analyses of graphite-moderated \glspl{MSR}. For the rest of this
-%section, I compare the $k$ estimates of the deterministic methods with OpenMC-MG.
-%
-%\begin{table}[htb!]
-%  \centering
-%  \footnotesize
-%  \caption{Multiplication factor $k$ estimates for Cases 1a, 1b, 2a, 2b, 3a, and 3b from the
-%    OpenMC-CE, OpenMC-MG, $S_8$ neutron transport, neutron diffusion, and hybrid $S_N$-diffusion
-%    methods.}
-%  \begin{tabular}{c S[table-format=1.5(2)] S[table-format=1.5(2)] S[table-format=1.5(2)]
-%  S[table-format=1.5(2)] S[table-format=1.5(2)] S[table-format=1.5(2)]}
-%    \toprule
-%    \multirow{2}{*}{\textbf{Method}} &
-%    \multicolumn{6}{c}{\textbf{Multiplication factor,} $\bm{k}$} \\
-%    \cmidrule{2-7}
-%    & {\textbf{Case 1a}} & {\textbf{Case 1b}} & {\textbf{Case 2a}} &
-%    {\textbf{Case 2b}} & {\textbf{Case 3a}} & {\textbf{Case 3b}} \\
-%    \midrule
-%    OpenMC-CE & 1.60033(56) & 1.16749(60) & 1.20666(66) & 0.69499(47) & 1.19908(63) & 0.68565(53)\\
-%    OpenMC-MG & 1.60015(33) & 1.18104(54) & 1.20570(66) & 0.70539(50) & 1.19976(48) & 0.69730(50)\\
-%    $S_8$     & 1.60035     & 1.18068     & 1.20596     & 0.70470     & 1.19968     & 0.69652    \\
-%    Diffusion & 1.60036     & 1.17932     & 1.19709     & 0.69274     & 1.19064     & 0.68450    \\
-%    Hybrid    & {N/A}       & {N/A}       & 1.20630     & 0.70387     & 1.20004     & 0.69568    \\
-%    \bottomrule
-%  \end{tabular}
-%  \label{table:ck1}
-%\end{table}
-%%
-%\begin{table}[htb!]
-%  \centering
-%  \footnotesize
-%  \caption{Multiplication factor $k$ estimates for Cases 4a, 4b, 5a, and 5b from the OpenMC-CE,
-%    OpenMC-MG, $S_8$ neutron transport, neutron diffusion, and hybrid $S_N$-diffusion methods.}
-%  \begin{tabular}{c S[table-format=1.5(2)] S[table-format=1.5(2)] S[table-format=1.5(2)]
-%    S[table-format=1.5(2)]}
-%    \toprule
-%    \multirow{2}{*}{\textbf{Method}} &
-%    \multicolumn{4}{c}{\textbf{Multiplication factor,} $\bm{k}$} \\
-%    \cmidrule{2-5}
-%    & {\textbf{Case 4a}} & {\textbf{Case 4b}} & {\textbf{Case 5a}} &
-%    {\textbf{Case 5b}} \\
-%    \midrule
-%    OpenMC-CE & 1.64506(53) & 1.19116(63) & 0.70352(64) & 1.16832(58) \\
-%    OpenMC-MG & 1.64355(37) & 1.20738(63) & 0.71396(50) & 1.18204(53) \\
-%    $S_8$     & 1.64376     & 1.20744     & 0.71379     & 1.18218     \\
-%    Diffusion & 1.64355     & 1.20811     & 0.70524     & 1.18274     \\
-%    Hybrid    & {N/A}       & {N/A}       & 0.71673     & {N/A}       \\
-%    \bottomrule
-%  \end{tabular}
-%  \label{table:ck2}
-%\end{table}
-%%
-%\begin{table}[htb!]
-%  \centering
-%  \footnotesize
-%  \caption{Differences in $k$ estimates for Cases 1a, 1b, 2a, 2b, 3a, and 3b for the $S_8$ neutron
-%    transport, neutron diffusion, and hybrid $S_N$-diffusion methods relative to OpenMC-MG.}
-%  \begin{tabular}{c S[table-format=+1.5(2)] S[table-format=+1.5(2)] S[table-format=+1.5(2)]
-%      S[table-format=+1.5(2)] S[table-format=+1.5(2)] S[table-format=+1.5(2)]}
-%    \toprule
-%    \multirow{2}{*}{\textbf{Method}} & \multicolumn{6}{c}{$\bm{k-k_{MG}}$} \\
-%    \cmidrule{2-7}
-%    & {\textbf{Case 1a}} & {\textbf{Case 1b}} & {\textbf{Case 2a}} &
-%    {\textbf{Case 2b}} & {\textbf{Case 3a}} & {\textbf{Case 3b}} \\
-%    \midrule
-%    $S_8$     & +0.00020(33) & -0.00036(54) & +0.00026(66) & -0.00069(50) & -0.00008(48) & -0.00078(50) \\
-%    Diffusion & +0.00021(33) & -0.00172(54) & -0.00861(66) & -0.01265(50) & -0.00912(48) & -0.01280(50) \\
-%    Hybrid    & {N/A}        & {N/A}        & +0.00060(66) & -0.00152(50) & +0.00028(48) & -0.00162(50) \\
-%    \bottomrule
-%  \end{tabular}
-%  \label{table:ckdiff1}
-%\end{table}
-%%
-%\begin{table}[htb!]
-%  \centering
-%  \footnotesize
-%  \caption{Differences in $k$ estimates for Cases 4a, 4b, 5a, and 5b for the $S_8$ neutron
-%    transport, neutron diffusion, and hybrid $S_N$-diffusion methods relative to OpenMC-MG.}
-%  \begin{tabular}{c S[table-format=+1.5(2)] S[table-format=+1.5(2)] S[table-format=+1.5(2)]
-%    S[table-format=+1.5(2)]}
-%    \toprule
-%    \multirow{2}{*}{\textbf{Method}} & \multicolumn{4}{c}{$\bm{k-k_{MG}}$} \\
-%    \cmidrule{2-5}
-%    & {\textbf{Case 4a}} & {\textbf{Case 4b}} & {\textbf{Case 5a}} &
-%    {\textbf{Case 5b}} \\
-%    \midrule
-%    $S_8$     & +0.00021(37) & +0.00006(63) & -0.00017(50) & -0.00027(53) \\
-%    Diffusion & +0.00000(37) & +0.00073(63) & -0.00872(50) & +0.00277(53) \\
-%    Hybrid    & {N/A}        & {N/A}        & +0.00070(50) & {N/A}    \\
-%    \bottomrule
-%  \end{tabular}
-%  \label{table:ckdiff2}
-%\end{table}
-%%
-%\begin{table}[tb!]
-%  \centering
-%  \footnotesize
-%  \caption{Control rod worths calculated as changes in the reactivity between Case 5a and 5b for
-%    the OpenMC continuous-energy and multigroup mode calculations, and the $S_8$ neutron transport,
-%  neutron diffusion, and hybrid $S_N$-diffusion methods.}
-%  \begin{tabular}{c S}
-%    \toprule
-%    \multirow{2}{*}{\textbf{Method}} & {\textbf{Control rod worth}} \\
-%                                     & {$\bm{\rho_{(5b)}-\rho_{(5a)}}$} \\
-%    \midrule
-%    OpenMC-CE & 0.56549(213) \\
-%    OpenMC-MG & 0.55464(176) \\
-%    $S_8$     & 0.55507 \\
-%    Diffusion & 0.57246 \\
-%    Hybrid    & 0.54974 \\
-%    \bottomrule
-%  \end{tabular}
-%  \label{table:rod-worth}
-%\end{table}
-%
-%Tables \ref{table:ckdiff1} and \ref{table:ckdiff2} show the differences in $k$ between OpenMC-MG
-%and deterministic methods. The $k$ estimates from the $S_8$ neutron transport method show
-%good agreement, with the estimates from OpenMC-MG with all pairs of $k$ values falling within
-%1-2 standard deviations of each other. The only significant difference in implementation between
-%the OpenMC-MG and $S_8$ methods is the discretization of the direction of travel variable
-%$\hat{\Omega}$. The flux distributions exhibit more apparent differences, will be discussed in the
-%following subsection. Nevertheless, the $S_8$ method is sufficiently accurate and consistent for
-%modeling this graphite-moderated system with control rods.
-%
-%As expected, the neutron diffusion method performs significantly worse than the $S_8$ method in
-%Cases 2, 3, and 5a, containing control rod regions. It deviates from the OpenMC-MG $k$
-%estimate by up to -0.01280 in Case 3b compared to -0.00078 for the $S_8$ method. The hybrid method
-%effectively improves the $k$ estimate to -0.00162 of the
-%OpenMC-MG method for Case 3b. The hybrid method also appreciably improved the accuracy in all
-%other applied cases.
-%
-%In a time-dependent multiphysics simulation involving rod
-%insertion or withdrawal, the control rod worth determines the change in reactivity and the
-%subsequent effects on the neutron flux and temperature distribution. I calculated the control rod
-%worth by taking the change in reactivity $\rho$ between Case 5a and 5b for all five methods, as
-%shown in Table \ref{table:rod-worth}. The formula for $\rho$ is given as
-%%
-%\begin{align}
-%  \rho =& \frac{k-1}{k}.
-%\end{align}
-%%
-%I took the $k$ estimate from the neutron diffusion method in Case 5b as the reference non-inserted
-%case for the control rod worth calculation for the hybrid method. Compared with OpenMC-MG, the
-%control rod worth from the neutron diffusion method deviates the most by 1782 pcm (percent mille)
-%as opposed to 43 pcm and -490 pcm by the $S_8$ and hybrid methods, respectively.
-%
-%\subsection{Comparison of Neutron Flux Distributions, $\phi$}
-%
-%This section focuses on the neutron flux distributions in Cases 3b and 5a, representing the most
-%complex geometries with homogeneous and lattice fuel-graphite regions. I sampled the flux
-%distributions in OpenMC-CE and OpenMC-MG across coarser $0.1$-cm intervals to reduce the
-%statistical uncertainties of each flux reading.
-%
-%\begin{figure}[htb!]
-%  \centering
-%  \begin{subfigure}[t]{.49\textwidth}
-%    \centering
-%    \includegraphics[width=\textwidth]{case-3b-group-1-flux}
-%    \caption{Group 1}
-%    \label{fig:c3bg1}
-%  \end{subfigure}
-%  \hfill
-%  \begin{subfigure}[t]{.49\textwidth}
-%    \centering
-%    \includegraphics[width=\textwidth]{case-3b-group-2-flux}
-%    \caption{Group 2}
-%    \label{fig:c3bg2}
-%  \end{subfigure}
-%  \begin{subfigure}[t]{.49\textwidth}
-%    \centering
-%    \includegraphics[width=\textwidth]{case-3b-group-3-flux}
-%    \caption{Group 3}
-%    \label{fig:c3bg3}
-%  \end{subfigure}
-%  \hfill
-%  \begin{subfigure}[t]{.49\textwidth}
-%    \centering
-%    \includegraphics[width=\textwidth]{case-3b-group-4-flux}
-%    \caption{Group 4}
-%    \label{fig:c3bg4}
-%  \end{subfigure}
-%  \caption{Neutron group flux distributions for Case 3b from OpenMC-CE, OpenMC-MG, $S_8$, neutron
-%  diffusion, and hybrid methods.}
-%  \label{fig:c3bflux}
-%\end{figure}
-%
-%Figure \ref{fig:c3bflux} shows the neutron flux distributions for groups 1 to 4. Notably, the
-%OpenMC-CE flux distribution is the most dissimilar, likely due to
-%the inadequate neutron energy discretization scheme. Consequently, I compared the
-%deterministic methods with OpenMC-MG to eliminate energy discretization errors. Figure
-%\ref{fig:c3bfluxe} shows the relative differences in the flux distributions from $S_8$, neutron
-%diffusion, and hybrid methods with respect to OpenMC-MG. The neutron diffusion method
-%performs worse than $S_8$, especially near the control rod region between $x=0$ cm and $x=20$ cm
-%for the slower group 3 and 4 neutron flux. Although the control rod region spans between $x=0$ cm
-%and $x=0.5$ cm only, it induces strong flux gradients in its vicinity, thus rendering the neutron
-%diffusion method less valid. The hybrid method significantly improves all four
-%group fluxes in this region. Beyond $x=40$ cm, the neutron diffusion and hybrid methods produce
-%similar flux error distributions due to the influence of the material interface between the
-%fuel-graphite and reflector regions.
-%
-%\begin{figure}[htb!]
-%  \centering
-%  \begin{subfigure}[t]{.49\textwidth}
-%    \centering
-%    \includegraphics[width=\textwidth]{case-3b-group-1-flux-error}
-%    \caption{Group 1}
-%    \label{fig:c3bg1e}
-%  \end{subfigure}
-%  \hfill
-%  \begin{subfigure}[t]{.49\textwidth}
-%    \centering
-%    \includegraphics[width=\textwidth]{case-3b-group-2-flux-error}
-%    \caption{Group 2}
-%    \label{fig:c3bg2e}
-%  \end{subfigure}
-%  \begin{subfigure}[t]{.49\textwidth}
-%    \centering
-%    \includegraphics[width=\textwidth]{case-3b-group-3-flux-error}
-%    \caption{Group 3}
-%    \label{fig:c3bg3e}
-%  \end{subfigure}
-%  \hfill
-%  \begin{subfigure}[t]{.49\textwidth}
-%    \centering
-%    \includegraphics[width=\textwidth]{case-3b-group-4-flux-error}
-%    \caption{Group 4}
-%    \label{fig:c3bg4e}
-%  \end{subfigure}
-%  \caption{Relative differences of the neutron group flux distributions for Case 3b from the $S_8$,
-%    neutron diffusion, and hybrid methods with respect to OpenMC-MG.}
-%  \label{fig:c3bfluxe}
-%\end{figure}
-%
-%For a quantitative assessment, I defined the normalized flux error $\varepsilon$ (not to be
-%confused with relative error norm $\epsilon$) for each energy group with respect to the OpenMC-MG
-%flux distribution using a normalized Frobenius/Euclidean norm formula given as
-%%
-%\begin{align}
-%  \varepsilon_g =& \frac{\lVert\sum^I_{i=1}\phi_{g,i} - \phi^{MG}_{g,i}\rVert_F}
-%  {\lVert\sum^I_{i=1}\phi^{MG}_{g,i}\rVert_F} =
-%  \frac{\left[\sum^I_{i=1}\lvert\phi_{g,i} - \phi^{MG}_{g,i}\rvert^2 \right]^{\sfrac{1}{2}}}
-%  {\left[\sum^I_{i=1}\lvert\phi^{MG}_{g,i}\rvert^2\right]^{\sfrac{1}{2}}}
-%\end{align}
-%%
-%I integrated on the fine flux distributions over 0.1-cm intervals from $S_8$, neutron diffusion,
-%and hybrid methods to match the 0.1-cm intervals of the OpenMC-MG flux distribution. Table
-%\ref{table:c3berror} shows the $\varepsilon$ values for the deterministic methods. As expected,
-%the hybrid method produces smaller flux error norms than the neutron diffusion method. The
-%improvements are more pronounced for group 3 and 4 fluxes than group 1 and 2 fluxes.
-%%
-%\begin{table}[tb!]
-%  \centering
-%  \footnotesize
-%  \caption{Normalized flux error $\varepsilon$ for Case 3b from the $S_8$, neutron diffusion, and
-%    hybrid methods with respect to OpenMC-MG.}
-%  \begin{tabular}{c S S S S}
-%    \toprule
-%    {\multirow{2}{*}{\textbf{Method}}} &
-%    \multicolumn{4}{c}{\textbf{Normalized flux error,} $\bm{\varepsilon_g}$} \\
-%    \cmidrule{2-5}
-%    & {\textbf{Group 1}} & {\textbf{Group 2}} & {\textbf{Group 3}} &
-%    {\textbf{Group 4}} \\
-%    \midrule
-%    $S_8$     & 0.0066 & 0.0048 & 0.0058 & 0.0067 \\
-%    Diffusion & 0.0144 & 0.0147 & 0.0258 & 0.0267 \\
-%    Hybrid    & 0.0122 & 0.0099 & 0.0087 & 0.0084 \\
-%    \bottomrule
-%  \end{tabular}
-%  \label{table:c3berror}
-%\end{table}
-%
-%Case 5a introduces significant geometrical heterogeneity from the fuel-graphite lattice region.
-%The heterogeneity is very apparent from the flux distributions in Figure \ref{fig:c5aflux}. The
-%OpenMC Monte
-%Carlo methods captured the fluctuations in the flux through their continuous angle variable as
-%opposed to the $S_8$, neutron diffusion, and hybrid methods. The group 1 flux peaks are notably
-%pronounced since all fission neutrons are born in the fuel region, with the majority in group 1.
-%Figure \ref{fig:c5afluxe} shows the relative differences in flux for the $S_8$, neutron diffusion,
-%and hybrid methods relative to OpenMC-MG. All three methods show regular fluctuations in the
-%relative flux differences corresponding to the fuel-graphite lattice. The $S_8$ method generally
-%deviates the least from OpenMC-MG, while the neutron diffusion method exhibits
-%significant flux deviations near the control rod region prior to $x=20$ cm and the material
-%interface between the fuel and reflector regions beyond $x=40$ cm.
-%
-%\begin{figure}[htb!]
-%  \centering
-%  \begin{subfigure}[t]{.49\textwidth}
-%    \centering
-%    \includegraphics[width=\textwidth]{case-5a-group-1-flux}
-%    \caption{Group 1}
-%    \label{fig:c5ag1}
-%  \end{subfigure}
-%  \hfill
-%  \begin{subfigure}[t]{.49\textwidth}
-%    \centering
-%    \includegraphics[width=\textwidth]{case-5a-group-2-flux}
-%    \caption{Group 2}
-%    \label{fig:c5ag2}
-%  \end{subfigure}
-%  \begin{subfigure}[t]{.49\textwidth}
-%    \centering
-%    \includegraphics[width=\textwidth]{case-5a-group-3-flux}
-%    \caption{Group 3}
-%    \label{fig:c5ag3}
-%  \end{subfigure}
-%  \hfill
-%  \begin{subfigure}[t]{.49\textwidth}
-%    \centering
-%    \includegraphics[width=\textwidth]{case-5a-group-4-flux}
-%    \caption{Group 4}
-%    \label{fig:c5ag4}
-%  \end{subfigure}
-%  \caption{Neutron group flux distributions for Case 5a from OpenMC-CE, OpenMC-MG, $S_8$, neutron
-%  diffusion, and hybrid methods.}
-%  \label{fig:c5aflux}
-%\end{figure}
-%%
-%\begin{figure}[htb!]
-%  \centering
-%  \begin{subfigure}[t]{.49\textwidth}
-%    \centering
-%    \includegraphics[width=\textwidth]{case-5a-group-1-flux-error}
-%    \caption{Group 1}
-%    \label{fig:c5ag1e}
-%  \end{subfigure}
-%  \hfill
-%  \begin{subfigure}[t]{.49\textwidth}
-%    \centering
-%    \includegraphics[width=\textwidth]{case-5a-group-2-flux-error}
-%    \caption{Group 2}
-%    \label{fig:c5ag2e}
-%  \end{subfigure}
-%  \begin{subfigure}[t]{.49\textwidth}
-%    \centering
-%    \includegraphics[width=\textwidth]{case-5a-group-3-flux-error}
-%    \caption{Group 3}
-%    \label{fig:c5ag3e}
-%  \end{subfigure}
-%  \hfill
-%  \begin{subfigure}[t]{.49\textwidth}
-%    \centering
-%    \includegraphics[width=\textwidth]{case-5a-group-4-flux-error}
-%    \caption{Group 4}
-%    \label{fig:c5ag4e}
-%  \end{subfigure}
-%  \caption{Relative differences of the neutron group flux distributions for Case 5a from the $S_8$,
-%    neutron diffusion, and hybrid methods with respect to OpenMC-MG.}
-%  \label{fig:c5afluxe}
-%\end{figure}
-%%
-%\begin{table}[tb!]
-%  \centering
-%  \footnotesize
-%  \caption{Normalized flux error $\varepsilon$ for Case 5a from the $S_8$, neutron diffusion, and
-%    hybrid methods with respect to OpenMC-MG.}
-%  \begin{tabular}{c S S S S}
-%    \toprule
-%    {\multirow{2}{*}{\textbf{Method}}} &
-%    \multicolumn{4}{c}{\textbf{Normalized flux error,} $\bm{\varepsilon_g}$} \\
-%    \cmidrule{2-5}
-%    & {\textbf{Group 1}} & {\textbf{Group 2}} & {\textbf{Group 3}} &
-%    {\textbf{Group 4}} \\
-%    \midrule
-%    $S_8$     & 0.0078 & 0.0074 & 0.0083 & 0.0085 \\
-%    Diffusion & 0.0305 & 0.0202 & 0.0313 & 0.0406 \\
-%    Hybrid    & 0.0290 & 0.0147 & 0.0143 & 0.0216 \\
-%    \bottomrule
-%  \end{tabular}
-%  \label{table:c5aerror}
-%\end{table}
-%
-%\subsection{Correction Regions \& Diffusion Coefficients}
-%
-%In this section, I compare the $P_1$-based diffusion coefficients, reference \glspl{SVDC}, and
-%hybrid \glspl{SVDC} for Cases 3b and 5a. As in Section \ref{sec:hybrid-method}, the reference
-%\glspl{SVDC} are derived from the $S_8$ flux solution, while the hybrid \glspl{SVDC} are derived
-%from the $S_8$ sub-solver flux solution within the hybrid $S_N$-diffusion method.
-%
-%Figure \ref{fig:c3bdiffcoef} shows the various diffusion coefficient distributions for Case 3b
-%between $x=0$ cm and $x=30$ cm. The figures omit the large diffusion coefficient values in the
-%air gap region between $x=0.5$ cm and $x=1$ cm since they do not show any significant features. The
-%hybrid \glspl{SVDC} closely match
-%the reference \gls{SVDC} within most of the correction region, which terminates at $x=20$ cm.
-%
-%\begin{figure}[htb!]
-%  \centering
-%  \begin{subfigure}[t]{.49\textwidth}
-%    \centering
-%    \includegraphics[width=\textwidth]{case-3b-group-1-diffcoef}
-%    \caption{Group 1}
-%    \label{fig:c3bg1dc}
-%  \end{subfigure}
-%  \hfill
-%  \begin{subfigure}[t]{.49\textwidth}
-%    \centering
-%    \includegraphics[width=\textwidth]{case-3b-group-2-diffcoef}
-%    \caption{Group 2}
-%    \label{fig:c3bg2dc}
-%  \end{subfigure}
-%  \begin{subfigure}[t]{.49\textwidth}
-%    \centering
-%    \includegraphics[width=\textwidth]{case-3b-group-3-diffcoef}
-%    \caption{Group 3}
-%    \label{fig:c3bg3dc}
-%  \end{subfigure}
-%  \hfill
-%  \begin{subfigure}[t]{.49\textwidth}
-%    \centering
-%    \includegraphics[width=\textwidth]{case-3b-group-4-diffcoef}
-%    \caption{Group 4}
-%    \label{fig:c3bg4dc}
-%  \end{subfigure}
-%  \caption{$P_1$-based diffusion coefficient, reference \gls{SVDC}, and hybrid \gls{SVDC}
-%    distributions for Case 3b between $x=0$ cm and $x=30$ cm.}
-%  \label{fig:c3bdiffcoef}
-%\end{figure}
-%%
-%\begin{figure}[htb!]
-%  \centering
-%  \includegraphics[width=.6\textwidth]{case-3b-grad-j}
-%  \caption{The gradient of current $J$ in the homogeneous fuel-graphite mixture from $x=1$ cm to
-%  $x=45$ cm for Case 3b.}
-%  \label{fig:c3bgradj}
-%\end{figure}
-%
-%We recall that the hybrid method determines cutoff points within the correction region where the
-%\glspl{SVDC} coincide with the $P_1$-based diffusion coefficients so that the flux gradients
-%remain continuous. The hybrid method discards \glspl{SVDC} in the buffer region between this
-%cutoff point and the correction region boundary in favor of the $P_1$-based diffusion coefficients.
-%Notably, the group 1 reference \glspl{SVDC} (Figure \ref{fig:c3bg1dc}) only coincides
-%with the $P_1$-based diffusion coefficient after the vertical asymptotes where the flux
-%gradient approaches zero. The trends in group 1 reference \glspl{SVDC} before and after the
-%vertical asymptote imply that the \gls{SVDC} would have gradually approached the $P_1$-based
-%diffusion coefficient value if the asymptote did not exist. Determining the cutoff point for the
-%\glspl{SVDC} would have been problematic due to the vertical asymptote. However, the hybrid
-%\glspl{SVDC} coincide with the $P_1$-based diffusion coefficients before the vertical asymptote.
-%This behavior can be reliably reproduced in other geometries with large homogeneous fissile regions
-%(Cases 2a, 2b, and 3a) as long as the correction region terminates near the neutron flux peaks.
-%Nevertheless, the issue of handling vertical asymptotes in the \glspl{SVDC} has to be addressed
-%since flux peaks and troughs will inevitably occur in more realistic reactor geometries.
-%
-%In contrast with the flux distributions in Figure \ref{fig:c3bflux}, the \gls{SVDC} distributions
-%suggest that the control rod contributes to an extended range of transport corrections on group 1
-%neutrons than the slower neutron groups. This phenomenon can be largely attributed to the highly
-%non-uniform fission neutron source distribution, which serves as the only source of group 1
-%neutrons in the absence of neutron up-scattering. The group 4 flux distribution
-%significantly influences on fission neutron source distribution, leading to more group 1 neutrons
-%being born closer to the reflector on the right than the
-%control rod on the left. The biased neutron source distribution induces a highly non-linear neutron
-%current distribution, as shown in Figure \ref{fig:c3bgradj}, and a departure from Fick's first law,
-%which relates the current to the flux gradient. Therefore, the \gls{SVDC} distributions show that
-%transport corrections to the faster neutron groups are as necessary for overall accuracy as
-%transport corrections to the slower neutron groups.
-%
-%Figure \ref{fig:c5adiffcoef} shows the various diffusion coefficient distributions for Case 5a
-%between $x=0$ cm and $x=30$ cm. As with Case 3b, the figures omit the large diffusion coefficient
-%values in the air gap region between $x=0.5$ cm and $x=1.5$ cm  The hybrid \glspl{SVDC} closely match
-%the reference \gls{SVDC} within most of the correction region, which terminates at $x=10$ cm.
-%
-%\begin{figure}[htb!]
-%  \centering
-%  \begin{subfigure}[t]{.49\textwidth}
-%    \centering
-%    \includegraphics[width=\textwidth]{case-5a-group-1-diffcoef}
-%    \caption{Group 1}
-%    \label{fig:c5ag1dc}
-%  \end{subfigure}
-%  \hfill
-%  \begin{subfigure}[t]{.49\textwidth}
-%    \centering
-%    \includegraphics[width=\textwidth]{case-5a-group-2-diffcoef}
-%    \caption{Group 2}
-%    \label{fig:c5ag2dc}
-%  \end{subfigure}
-%  \begin{subfigure}[t]{.49\textwidth}
-%    \centering
-%    \includegraphics[width=\textwidth]{case-5a-group-3-diffcoef}
-%    \caption{Group 3}
-%    \label{fig:c5ag3dc}
-%  \end{subfigure}
-%  \hfill
-%  \begin{subfigure}[t]{.49\textwidth}
-%    \centering
-%    \includegraphics[width=\textwidth]{case-5a-group-4-diffcoef}
-%    \caption{Group 4}
-%    \label{fig:c5ag4dc}
-%  \end{subfigure}
-%  \caption{$P_1$-based diffusion coefficient, reference \gls{SVDC}, and hybrid \gls{SVDC}
-%    distributions for Case 5a between $x=0$ cm and $x=30$ cm.}
-%  \label{fig:c5adiffcoef}
-%\end{figure}
-%
-%Unlike Case 3b, the heterogeneous fuel-graphite lattice structure in Case 5a gives rise to
-%extreme fluctuations in the \gls{SVDC} distributions. The numerous asymptotes correspond to the
-%flux peaks and troughs throughout the lattice. Regardless, the presence of the control rod
-%suppresses the \gls{SVDC} fluctuations in the lattice close to the control rod, and the cutoff
-%points can be determined without manual intervention. The heterogeneous lattice induces competing
-%effects on the neutron flux, allowing for a smaller correction region of 10 cm in width. However,
-%the reduced correction region introduces less transport correction as a consequence.
-%Further investigation is needed to observe how the \gls{SVDC} distributions change with
-%increases in dimensionality for 2-D and 3-D systems.
-%
-%\section{Summary} \label{sec:hybrid-summary}
+\section{Summary} \label{sec:hybrid-summary}
 %

### 2024-06-19 cnrs-benchmark/conclusion.tex
add verification chapter for referencing, hybrid chapter edits

\subsection{Summary}

\glspl{MSR} feature significant multiphysics interactions, presenting
computational challenges for many existing multiphysics reactor analysis
software. This chapter presents code-to-code verification of Moltres'
capabilities in modeling such multiphysics phenomena in fast-spectrum
\glspl{MSR} based on the CNRS benchmark \cite{tiberga_results_2020}.
The CNRS benchmark assesses multiphysics \gls{MSR} simulation
software through several steps involving single-physics and coupled
neutronics/thermal-hydraulics problems.

The results showed that Moltres is consistent with the participating software
presented in the CNRS benchmark paper for modeling important phenomena
in fast-spectrum \glspl{MSR}. The percentage discrepancies in the various
neutronics, velocity, and temperature quantities mostly fall below or within
one standard deviation of the average of the benchmark participants.
Minor deviations in the temperature in Steps 0.3 and 1.2 
stem from the discontinuous velocity
boundaries on the top corners in the lid-driven cavity flow. We have shown that
these deviations are limited to the top boundary of the domain and do not
affect the rest of the physical parameters. The results from
Moltres agree closest with the TUD-S$_2$ software package, which implements the
$S_2$ discrete ordinates method for
neutron transport on a uniform structured mesh with a \gls{DFEM}-based solver.
These features make Moltres the most similar to the TUD-$S_2$ model as compared
to the other models, which employ different neutron transport models,
non-uniform meshes, or finite volume-based solvers.

This work verifies Moltres' capabilities for future work involving the modeling and
simulation of fast-spectrum \glspl{MSR} such as the European \gls{MSFR} and
TerraPower's \gls{MCFR} \cite{terrapower_terrapower_2021}. Notably, the CNRS
benchmark does not assess modeling capabilities for complex physics phenomena
such as turbulent flow in \glspl{MSR}. However, we expect coolant loops in many \gls{MSR} designs
will experience turbulent flow under regular operation or accident scenarios.
These expectations, alongside the subpar results of pump-initiated accidents
reported in Section \ref{sec:msfr}, call for the implementation and
verification of a turbulence model in Moltres to accurately model \glspl{MSR}.

\FloatBarrier

### 2024-06-19 cnrs-benchmark/moltres.tex
add verification chapter for referencing, hybrid chapter edits

\subsection{Modeling Approach with Moltres} \label{sec:model}

This section describes the specific modeling approach for
simulating the CNRS Benchmark cases in Moltres.

For this work\footnote{The input files for
all benchmark
cases are available on the Moltres GitHub repository at 
\url{https://github.com/arfc/moltres/tree/devel/problems/2021-cnrs-benchmark}.
}, I ran the benchmark cases on a uniformly-spaced mesh
of 200$\times$200 elements. Thus, the dimensions of each mesh element are
0.01m$\times$0.01m. I adopted the group constant data
provided by Tiberga et al. \cite{tiberga_results_2020}. Next, I
discretized most variables, i.e., neutron fluxes, velocity
components, pressure, and temperature, using continuous, first-order, Lagrange
shape functions. The only exception is the precursor concentration variables,
which I discretized using zeroth-order monomial shape functions and solved
using a \gls{DFEM}. I interpolated the resulting discontinuous,
cell-centered precursor values to obtain the nodal values for results
analysis.

The
\texttt{Navier-Stokes} and \texttt{Heat} \texttt{Conduction} modules from
\gls{MOOSE} provide some of the capabilities for
modeling incompressible flow and heat transfer. In particular, I stabilized
the incompressible flow and temperature governing equations using the
\gls{SUPG} stabilization method implemented in \gls{MOOSE}
\cite{peterson_overview_2018}. Without \gls{SUPG} stabilization, I
observed spurious numerical oscillations in the velocity and temperature near
the top boundary due to the singularity on the top left corner where different
velocity boundary conditions meet. I also applied the \gls{PSPG} stabilization
scheme \cite{hughes_new_1986} from the Navier-Stokes module
\cite{peterson_overview_2018},
which enables equal-order discretizations in the velocity and pressure
variables. Equal-order discretizations with \gls{PSPG} are computationally
cheaper and more convenient than implementing higher-order
velocity discretizations for stability without \gls{PSPG}
\cite{chapelle_inf-sup_1993}.

Using the inverse power method solver in \gls{MOOSE}, I ran all eigenvalue calculations in
Steps 0.2, 1.1, 1.2, 1.3, and 1.4. I ran all other steps
using the Preconditioned Newton-Krylov solver
\cite{gaston_physics-based_2015}. The coupled steady-state problems in
Steps 1.2, 1.3, and 1.4 required segregated solvers for the neutronics
and the thermal-hydraulics due to the unique problem setups involving a
criticality search problem for the neutron multiplication factor
and a steady-state problem in thermal-hydraulics simultaneously.

\begin{table}[tb]
    \caption{Timestep sizes used for the time-dependent cases in
    Step 2.1, corresponding to 1/200th of the perturbation period.}
	\centering
	\setlength\tabcolsep{2.5pt}
	\begin{tabular}{l l l l l l l l}
	    \toprule
	    Frequency [Hz] & 0.0125 & 0.025 & 0.05 & 0.1 & 0.2 & 0.4 & 0.8 \\
	    \midrule
	    Timestep size [s] & 0.2 & 0.2 & 0.1 & 0.05 & 0.025 & 0.0125 & 0.00625
	    \\
	    \bottomrule
	\end{tabular}
	\label{table:timestep}
\end{table}

For the time-dependent cases in Step 2.1, I employed full coupling with
a second-order implicit Backward Differential Formula (BDF2) time-stepping
scheme. I set the timestep sizes for each driving frequency in the heat transfer
coefficient to 1/200th of the perturbation period. Table
\ref{table:timestep} shows the timestep sizes. I assumed the
systems reached asymptotic behavior when the magnitudes of neighboring power
peaks differed by less than 0.001\% for at least ten wavelengths. Under this
assumption, the phase shift measurements between adjacent waves always
converged before the magnitude measurements of the power peaks.

Table \ref{table:software} compares the numerical methods, meshing schemes, and
neutronics models of Moltres and the four participating software packages in
the CNRS benchmark paper \cite{tiberga_results_2020}. The $SP_N$ and
$S_N$ neutronics models refer to the simplified $P_N$ spherical harmonics and
$S_N$ discrete ordinates neutron transport models, respectively. Based on the
solvers and methods of solution, Moltres is most similar to the
PHANTOM-$S_N$ + DGFlows \cite{tiberga_discontinuous_2019} multiphysics package
from \gls{TUD} with the $S_2$ neutron transport model. Participants from
\gls{CNRS} and \gls{PSI}
employed non-uniform meshes which were refined near the boundaries. In contrast,
we and the \gls{PoliMi} and \gls{TUD} participants employed uniform meshes.

\FloatBarrier

\begin{landscape}
\begin{table*}[p]
    \caption{List of software packages and their corresponding model
    specifications for the CNRS Benchmark simulations
    \cite{tiberga_results_2020}.}
    \centering
    \begin{tabular}{p{4.2cm} p{7cm} p{3.3cm} p{2cm} p{2.7cm}}
        \toprule
        Software & Institute & Numerical method & Mesh & Neutronics model \\
        \midrule
        OpenFOAM & Centre national de la recherche scientifique (CNRS) & Finite volume & 200$\times$200 \newline Non-uniform & $SP_1$ \& $SP_3$ \\
        OpenFOAM & Politecnico di Milano (PoliMi) & Finite volume & 400$\times$400 \newline Uniform & Neutron diffusion \\
        GeN-Foam & Paul Scherrer Institute (PSI) & Finite volume & 200$\times$200 \newline Non-uniform & Neutron diffusion \\
        PHANTOM-$S_N$+DGFlows & Delft University of Technology (TUD) & Discontinuous finite \newline element & 50$\times$50 \newline Uniform & $S_2$ \& $S_6$ \\
        Moltres (This work) & University of Illinois at Urbana-Champaign (UIUC) & Continuous \& discontinuous finite element & 200$\times$200 \newline Uniform & Neutron diffusion \\
        \bottomrule
    \end{tabular}
    \label{table:software}
\end{table*}
\end{landscape}

\FloatBarrier

### 2024-06-19 hybrid.tex
add verification chapter for referencing, hybrid chapter edits

@@
 
+Section \ref{sec:saaf} provides the mathematical derivation for the discretized weak form of the
+\gls{SAAF} $S_N$ neutron transport equations. Section \ref{sec:transport-correction} provides the
+derivations for the transport correction formulations I investigated for use in the hybrid method.
+Section \ref{sec:hybrid-algorithm} details the iteration algorithm for the hybrid method. Section
+\ref{sec:sn-bc} provides the boundary conditions formulated for the $S_N$ calculations in the
+hybrid method. In Section \ref{sec:buffer-region}, I discuss how transport corrections are handled
+in the hybrid method. Lastly, Section \ref{sec:hybrid-summary} summarizes the hybrid method and its
+implementation.
+
 % Section \ref{sec:theory} discusses the theoretical background for the hybrid $S_N$-diffusion
@@
 
-% The proposed hybrid $S_N$-diffusion method is an iterative method to improve the
-% accuracy of neutron diffusion solutions in reactor systems with highly anisotropic flux regions
-% which I will use to model control rods.
+\section{\Gls{SAAF} $S_N$ Method} \label{sec:saaf}
 
-\section{\Gls{SAAF} Formulation of the $S_N$ Method}
+This chapter starts with the $S_N$ method implementation required for the hybrid $S_N$-diffusion
+method. I implemented both $S_N$ and hybrid methods using \gls{FEM} numerical solver capabilities
+available through Moltres \cite{lindsay_moltres_2017} and the \gls{MOOSE} framework. In this
+section, the derivation for the \gls{SAAF} formulation of the $S_N$ method and its
+implementation in Moltres follows closely the derivation developed by Wang et al.
+\cite{wang_diffusion_2014, wang_rattlesnake_2018}.
 
@@
 %
+$w$ replaces all instances of the $4\pi$ normalization constant in the \gls{SAAF} equations.
 We calculate the scalar flux $\phi_g$ by taking the zeroth moment with respect to $\hat{\Omega}$ as
@@
     \sum^G_{g=1}\sum^{N_d}_{d=1}w_d\left(\Psi^*_{g,d},
-    \hat{\Omega}_d\cdot\hat{n}_b\Psi_{g,d}\right),
-    & \hat{\Omega}\cdot\hat{n}_b>0,\vec{r}\in\partial\mathcal{D}_s \\
-    \sum^G_{g=1}\sum^{N_d}_{d=1}w_d\left(\Psi^*_{g,d},
-    \hat{\Omega}_d\cdot\hat{n}_b\Psi^\text{inc}_{g,d}\right),
-    & \hat{\Omega}\cdot\hat{n}_b<0,\vec{r}\in\partial\mathcal{D}_s
+    \hat{\Omega}_d\cdot\hat{n}_b\Psi_{g,d}\right)_{\partial\mathcal{D}},
+    & \hat{\Omega}\cdot\hat{n}_b>0,\vec{r}\in\partial\mathcal{D} \\
+    0,
+    & \hat{\Omega}\cdot\hat{n}_b<0,\vec{r}\in\partial\mathcal{D}
   \end{cases},
@@
 %
-where $\Psi^\text{inc}_{g,d}$ is zero on a vacuum boundary or positive for a boundary source.
-
-\noindent Reflecting boundary term:
+Boundary source term:
+%
+\begin{gather}
+  \langle\bm{\Psi}^*,\bm{\Psi}\rangle^+ - \langle\bm{\Psi}^*,\bm{\Psi}^\text{inc}\rangle^- =
+  \begin{cases}
+    \sum^G_{g=1}\sum^{N_d}_{d=1}w_d\left(\Psi^*_{g,d},
+    \hat{\Omega}_d\cdot\hat{n}_b\Psi_{g,d}\right)_{\partial\mathcal{D}},
+    & \hat{\Omega}\cdot\hat{n}_b>0,\vec{r}\in\partial\mathcal{D} \\
+    \sum^G_{g=1}\sum^{N_d}_{d=1}w_d\left(\Psi^*_{g,d},
+    \hat{\Omega}_d\cdot\hat{n}_b\Psi^\text{inc}_{g,d}\right)_{\partial\mathcal{D}},
+    & \hat{\Omega}\cdot\hat{n}_b<0,\vec{r}\in\partial\mathcal{D}
+  \end{cases}, \label{eq:boundary-source}
+\end{gather}
+%
+Reflecting boundary term:
 %
@@
 
-\section{Transport Correction Formulations}
+\section{Transport Correction Formulations} \label{sec:transport-correction}
 
@@
 
-\section{Hybrid $S_N$-Diffusion Method} \label{sec:hybrid-method}
+\section{Iteration Algorithm} \label{sec:hybrid-algorithm}
 
 As described in the previous section, transport correction formulations are commonly used in
-iterative acceleration schemes or hybrid methods to couple high-order neutron transport
-calculations with low-order neutron diffusion calculations. However, the high-order calculations
-remain computationally intensive and limit their applicability to time-independent neutronics
-simulations on \gls{HPC} clusters.
+iterative acceleration schemes or other hybrid methods to couple high-order neutron transport
+calculations with low-order neutron diffusion calculations. However, the high-order calculations of
+full reactor core models remain computationally intensive and limited in their applicability to
+time-independent neutronics simulations on \gls{HPC} clusters.
 In order to reduce the computational cost of the high-level $S_N$ calculation in a reactor
@@
 method alone. Henceforth, I will refer to the $S_N$ calculation on the correction
-region as the $S_N$ \textit{subproblem} or \textit{sub-solver}. I define the full problem
+region as the $S_N$ \textit{subproblem} or \textit{subsolver}. I define the full problem
 domain and the correction region as $V_0$ and $V_1$, respectively, where
@@
   \item Pass the neutron diffusion current estimates along
-    $\partial V_1$ to the $S_N$ sub-solver to evaluate the boundary conditions for the $S_N$
+    $\partial V_1$ to the $S_N$ subsolver to evaluate the boundary conditions for the $S_N$
     subproblem.
-  \item With the $S_N$ sub-solver, evaluate transport correction terms in $V_1$ using Eq.
+  \item With the $S_N$ subsolver, evaluate transport correction terms in $V_1$ using Eq.
     \ref{eq:svdc} or \ref{eq:drift}.
@@
     \draw [arrow] (3) -- node[anchor=east] {$\bm{\phi^{tr,(i)}_g, J^{tr,(i)}_g}$} (4);
-    \draw [arrow] (4) -- node[anchor=east] {$\bm{D^{s,(i)}_g}$ or $\bm{\vec{D}^{(i)}_g}$} (5);
+    \draw [arrow] (4) -- node[anchor=east] {$\bm{D^{s,(i)}_g}$ \textbf{or} $\bm{\vec{D}^{(i)}_g}$} (5);
     \draw [arrow] (5) -- node[anchor=east] {$\bm{\phi^{(i)}_g, J^{(i)}_g}$} (6);
@@
 
-\subsubsection{$S_N$ Subsolver Boundary Conditions \& Buffer Zone}
+\section{$S_N$ Subsolver Boundary Conditions} \label{sec:sn-bc}
 
-The main challenge lies in determining appropriate boundary conditions for the $S_N$ subproblem.
+For the hybrid $S_N$-diffusion method to converge, it requires appropriate boundary conditions for
+the $S_N$ subproblem.
 Given that we want to limit the coverage of $V_1$ to the control rod region and its
@@
 anisotropies in the flux due to the control rod. As a consequence, the boundaries $\partial V_1$ 
-lie well within $V_0$. Crucially, there is currently no feasible method of generating
-accurate boundary fluxes for an $S_N$ solver from a neutron diffusion flux solution. In 1-D, the
-standard one-group $S_N$ method requires N/2
-incoming boundary flux parameters per boundary mesh point for the N/2 neutron angular fluxes
-flowing into $V_1$. The neutron diffusion method can produce at most one independent
-incoming flux parameter per mesh point; this parameter is the neutron forward/backward current in
-the $P_1$ approximation defined as:
+lie well within $V_0$. Crucially, there is no feasible method of generating
+accurate boundary fluxes for an $S_N$ solver from a neutron diffusion flux solution. We may obtain
+the best estimate by applying the $P_1$ approximation for evaluating the neutron angular flux along
+the discrete ordinates $\hat{\Omega}_d$ of the $S_N$ angular quadrature set as follows
 %
 \begin{align}
-  J_{g,\pm} &= \frac{\phi_g}{4} \mp \frac{D_g}{2}\frac{d\phi_g}{dx} \label{eq:p1-j}
-  \shortintertext{where}
-  J_{g,\pm} &= \mbox{ neutron forward/backward current of group }g. \nonumber
+  \Psi_{g,d} &\approx \frac{1}{4\pi}\left(\phi^\text{diff}_g+3\hat{\Omega}_d\cdot
+  \vec{J}^\text{diff}_g\right) \nonumber \\
+  &=\frac{1}{4\pi}\left(\phi^\text{diff}_g-3\hat{\Omega}_d\cdot D_g\nabla\phi^\text{diff}_g\right)
 \end{align}
 %
-Without additional information, the next best estimate is a uniformly
-isotropic transmission of angular flux, i.e., all N/2 incoming angular fluxes $\Psi$ are equal in
-magnitude. This is a good assumption far from the control rod, but will not produce accurate
-fluxes if the boundary is too close to the control rod. This type of boundary condition is similar
-to the white boundary condition, which describes the uniformly
-isotropic reflection of particles at a boundary, except this case involves transmission rather than
-reflection. Isotropic flux transmission at $x$ can be expressed mathematically for forward angular
-fluxes as:
+where the diff superscript denotes flux estimates from the latest neutron diffusion calculation.
+With this relation, we can formulate the boundary source term for the $S_N$ subsolver by setting
+$\Psi^\text{inc}_{g,d}$ in Eq.\ \ref{eq:boundary-source} to
 %
-\begin{align}
-  \sum^N_{n=N/2+1}w_n\mu_n\Psi(x,\mu_n) =& J_{+}(x) && (\mu_n>0) \nonumber \\
-  \Psi(x,\mu_n)\sum^N_{n=N/2+1}w_n\mu_n =& J_{+}(x) && (\because \mbox{isotropic transmission})
-  \nonumber \\
-  \Psi(x,\mu_n) =& J_{+}(x)\Bigg/\sum^N_{n=N/2+1}w_n\mu_n
-\end{align}
-%
-and backward angular fluxes as:
-%
-\begin{align}
-  \sum^{N/2}_{n=1}w_n\mu_n\Psi(x,\mu_n) =& J_{-}(x) && (\mu_n<0) \nonumber \\
-  \Psi(x,\mu_n)\sum^{N/2}_{n=1}w_n\mu_n =& J_{-}(x) && (\because \mbox{isotropic transmission})
-  \nonumber \\
-  \Psi(x,\mu_n) =& J_{-}(x)\Bigg/\sum^{N/2}_{n=1}w_n\mu_n \label{eq:sn-psi-j}
-\end{align}
-%
-These boundary conditions for the $S_N$ sub-solver will generally yield some deviations in the
-$\phi$ distributions since neutron fluxes in realistic reactor systems are at least slightly
-anisotropic in most of the domain. However, in my preliminary investigations, the influence of
-boundary conditions on the ratio $J$ and $\frac{d\phi}{dx}$ does not extend far from the boundaries
-in optically thick media due to the strong scattering (e.g., graphite or molten salt) or
-absorption effects (e.g., control rod). In the bulk regions far from the boundaries, the
-scattering and absorption effects have the most significant influence on the flux. I observed that
-the \glspl{SVDC} generated with the $S_N$ sub-solver within $V_1$ were accurate everywhere except
-near $\partial V_1$. Therefore, $V_1$ must be large enough to provide transport corrections
-through the \glspl{SVDC} and accommodate inaccurate \glspl{SVDC} near
-$\partial V_1$. The inaccurate \glspl{SVDC} should then be discarded in favor of using
-$P_1$-based diffusion coefficients as in the rest of $V_0$ not covered by $V_0$. The
-subsequent neutron diffusion calculation with this mix of \glspl{SVDC} and $P_1$-based diffusion
-coefficients should provide more accurate $\phi$ and $k$ estimates than a conventional
-neutron diffusion calculation.
+\begin{gather}
+  \Psi^\text{inc}_{g,d} = \frac{1}{w}
+  \left(\phi^\text{diff}_g-3\hat{\Omega}_d\cdot D_g\nabla\phi^\text{diff}_g\right)
+\end{gather}
+
+\section{Buffer Region} \label{sec:buffer-region}
+
+This boundary condition for the $S_N$ subsolver will generally cause deviations
+in the flux and transport correction values since the $P_1$ approximation is accurate for up to
+linear anisotropies in the flux. However, as I will show in the next chapter, the influence of
+boundary conditions on the accuracy of
+the transport correction terms does not extend far from the boundary $\partial V_1$ in optically
+thick media due to competing influences from the bulk scattering and absorption effects in $V_1$.
+Results in the next chapter will show that transport corrections generated with the $S_N$
+subsolver within $V_1$ are accurate everywhere up to a few centimeters near $\partial V_1$.
+
+When applying the hybrid method, $V_1$ must be large enough to provide accurate transport
+corrections near the highly-absorbing control rod while simultaneously accommodating for inaccurate
+transport corrections near $\partial V_1$. I will henceforth refer to the region in $V_1$
+near $\partial V_1$ with inaccurate transport corrections as the \textit{buffer region} ($V_1'$).
+We will discard inaccurate transport correction values in $V_1'$. Therefore, the subsequent neutron
+diffusion calculation on $V_0$ will only apply transport corrections in $V_1$ excluding $V_1'$.
+
+\section{Numerical Implementation}
 
 \section{Summary} \label{sec:hybrid-summary}
-%
-%In this chapter, I presented the theoretical basis for a hybrid $S_N$-diffusion method, a novel
-%method for improving the neutron diffusion method for neutronics simulations of reactor
-%systems containing highly neutron-absorbing regions. The hybrid method relies on \glspl{SVDC}
-%derived from reference flux solutions of $S_N$ calculations to provide transport
-%corrections to the neutron diffusion method. The hybrid method is less computationally intensive
-%than the standalone $S_N$ neutron transport method because 1) $S_N$ calculations are limited to
-%correction regions which cover a fraction of the overall geometry, 2) and the \glspl{SVDC} are
-%observed to converge faster than the neutron fluxes in $S_N$ calculations.
-%
-%I established a framework for the hybrid method for coupling an $S_N$ sub-solver and a neutron
+
+In this chapter, I presented the mathematical derivations for the \gls{SAAF} formulation of the
+$S_N$ method, transport correction formulations, and boundary conditions relevant to the hybrid
+$S_N$-diffusion method. I also presented details on the iteration algorithm and the handling of
+transport corrections in the hybrid $S_N$-diffusion method to improve on the standard neutron
+diffusion method for neutronics simulations of reactor systems containing highly neutron-absorbing
+regions (control rods). The hybrid method relies on transport corrections derived from the flux
+solutions from the $S_N$ subsolver to fix inaccurate flux estimates from using the neutron
+diffusion method alone. The hybrid method saves on computational expense at the cost of some
+accuracy in reactor regions not treated with transport corrections. In the next chapter, I will
+also show that I can further reduce computational expense by leaving the flux in the $S_N$
+subsolver because the transport correction parameters converge faster than the flux in the $S_N$
+subsolver.
+
+%I established a framework for the hybrid method for coupling an $S_N$ subsolver and a neutron
 %diffusion solver through outer iterations by taking advantage of the weak dependence of the $S_N$

### 2024-06-19 moltres.tex
add verification chapter for referencing, hybrid chapter edits

\section{Description of Moltres}

This chapter provides an overview of Moltres as a multiphysics
simulation software for molten salt reactors. 
Section \ref{sec:moltres-features}
describes the general software features of Moltres, Section
\ref{sec:moltres-physics} expands on the physics models in Moltres, and Section
\ref{sec:moltres-previous} describes previous work on \gls{MSR} analysis with Moltres to illustrate
its current capabilities and limitations.

\subsection{General Features} \label{sec:moltres-features}

This section discusses the general software features of Moltres. The
discussion focuses on robustness, extensibility, and ease of use
since these characteristics represent the hallmarks of good multiphysics
software \cite{keyes_multiphysics_2013}. These criteria also apply to
assessing \gls{MSR} simulation tools since the nonlinearity of \gls{MSR}
multiphysics analysis and the complexity of advanced reactor designs
necessitate robust, scalable, and flexible computational tools.

Moltres draws many advantages from being developed on MOOSE
\cite{permann_moose_2020}. MOOSE is an open-source, finite-element,
multiphysics framework developed at \gls{INL}. The framework provides a
user-friendly interface for developing multiphysics software through
\gls{OOP} in \texttt{C++} to modularize various
functions relevant to finite-element multiphysics solvers. In this approach,
MOOSE and MOOSE-based applications break down \glspl{PDE} into individual terms
and store them as individual \texttt{C++ objects} referred to as
\texttt{Kernels}. These \texttt{Kernels} contain functions for calculating
their weak form residual and Jacobian
contributions and other relevant functions required to solve a given
\gls{PDE}. \gls{OOP} in MOOSE simplifies software development
since developers can write new \texttt{Kernels} as child classes in
\texttt{C++} derived from existing \texttt{Kernels} (base classes), which share
similar physics to inherit common functions.
The same philosophy applies to all other systems in MOOSE, such as
the \texttt{BCs}, \texttt{Materials}, and \texttt{Postprocessor}
systems for handling relevant boundary conditions, material properties, and
postprocessing calculations, respectively. Overall, this approach also saves
researchers time and effort since they are unencumbered by the technical details
and complexities of programming efficient computational tools for
numerical analysis.

\begin{figure}[htb!]
	\centering
	\includegraphics[width=.7\columnwidth]{moose}
	\caption{Structure of MOOSE and its dependencies.}
	\label{fig:moose}
\end{figure}

MOOSE relies on two other open-source libraries: libMesh
\cite{kirk_libmesh_2006} for its \gls{FEM} capabilities on unstructured mesh
and PETSc \cite{satish_petsc_2019} for its nonlinear solvers and
preconditioning routines. By extension, Moltres gains access to these
sophisticated numerical analysis tools and benefits from their continuous
development. Figure
\ref{fig:moose} shows how MOOSE serves as an interface between physics
applications and libMesh/PETSc. MOOSE supports
modeling on up to three-dimensional (3-D) unstructured meshes for a wide range
of mesh file formats, including the commonly used Exodus II file format. For
2-D meshes, users can opt between Cartesian and polar RZ coordinates. MOOSE
also supports parallel computing through the \gls{MPI} library to leverage
modern high-performance computing for large multiphysics simulations.

Moltres benefits from the highly-integrated cross compatibility within the
ecosystem of MOOSE-based applications. MOOSE facilitates multiphysics coupling
among all MOOSE-based applications by providing a common framework for shared
data access and file input/output, thus eliminating computational costs from
data transfers and allowing for fully coupled simulations. For example, Moltres
can couple with the \texttt{Navier-Stokes} module \cite{peterson_overview_2018}
from MOOSE for fully coupled reactor simulations modeling neutronics and
thermal-hydraulics with incompressible flow modeling. Section
\ref{sec:msr-multiphysics} highlighted the advantages of fully coupled schemes
for modeling strongly coupled systems, such as the coupled
neutronics and thermal-hydraulics in \glspl{MSR}. Moltres can also easily
couple to other MOOSE-based applications in a similar fashion. In
addition, MOOSE
provides the option for either tight or loose coupling through the
\texttt{MultiApp} system \cite{gaston_physics-based_2015}. Tight coupling
schemes can outperform fully coupled schemes in weakly coupled systems in which
the computational expenses of fully coupled schemes outweigh the savings from
running fewer Newton iterations due to the superior convergence rate. Loose
coupling schemes help accelerate time-dependent simulations of
stable systems towards a steady state if only the steady-state
configuration interests the user. This situation may occur in the later
stages of \gls{MSR} simulations when the \gls{DNP}
concentrations converge slowly due to their relatively large decay half-lives.
Furthermore, segregated solvers through the \texttt{MultiApp} system enable
Moltres to introduce \gls{DNP} drift and non-uniform
temperature distributions into criticality search simulations. Regardless of
which coupling scheme works best, MOOSE-based applications provide the flexibility
to switch among the schemes as users see fit. For time-dependent simulations,
MOOSE provides more than ten implicit and explicit timestepping
schemes. The default scheme is the first-order backward Euler
method which offers excellent solver stability for stiff \glspl{PDE}.

Lastly, Moltres is an open-source \gls{LGPL} software hosted on
GitHub \cite{github_build_2017}. Open-sourcing software provides ease of access
and expands the user base. These characteristics promote software quality
through increased feedback on users' needs and transparency for peer review.
Open-source software accelerate research progress by supporting
collaboration and sharing best software practices. Supporting Moltres'
continued development, Moltres relies on GitHub for online version control with
continuous integration testing to protect its existing capabilities.

In summary, Moltres provides robust and flexible coupling capabilities to
model strongly coupled neutronics and thermal-hydraulics in \glspl{MSR}. As a
MOOSE-based application, Moltres is highly extensible through coupling with
other MOOSE-based applications and benefits from MOOSE's user-friendly
interface for software development and general ease of use.

\subsection{Physics Models} \label{sec:moltres-physics}

This section describes the various physics models available in Moltres to model
coupled neutronics and thermal-hydraulics in \glspl{MSR}. Section \ref{sec:nts}
discusses the neutronics model in Moltres, Section \ref{sec:th} discusses
the thermal-hydraulics model, and Section \ref{sec:moltres-loop} discusses the
external loop model for \gls{DNP} looping and heat removal via external heat
exchangers.

\subsubsection{Multigroup Neutron Diffusion Model} \label{sec:nts}

Moltres solves the multigroup neutron diffusion equations for the neutron
flux solution within the problem domain. These equations are derived from the
neutron transport equation in the diffusion-dominated limit with Fick's law of
diffusion. They are further simplified by discretizing the continuous neutron energy
variable into a finite number of energy groups \cite{bell_nuclear_1970,
duderstadt_nuclear_1976}. The time-dependent multigroup neutron
diffusion equations with $G$ energy groups and $I$ \gls{DNP}
groups are given by:
%
\begin{align}
    \frac{1}{v_g} \frac{\partial \phi_g}{\partial t} =& \nabla \cdot D_g
    \nabla \phi_g - \Sigma^r_g \phi_g +
    \sum^G_{g' \neq g} \Sigma^s_{g' \rightarrow g} \phi_{g'} \nonumber \\
    &+ \chi^p_g \sum^G_{g'=1} \left( 1-\beta \right) \nu \Sigma^f_{g'}
    \phi_{g'} + \chi^d_g \sum^I_i \lambda_i C_i \label{eq:neutron} %\\
%    %
%    \shortintertext{where}
%    v_g =& \text{ average speed of neutrons in group $g$,} 
%    \nonumber \\
%    \phi_g =& \text{ neutron flux in group $g$,}
%    \nonumber \\
%    t =& \text{ time,} \nonumber \\
%    D_g =& \text{ diffusion coefficient of neutrons in group $g$,} \nonumber \\
%    \Sigma^r_g =& \text{ macroscopic removal cross section for} \nonumber \\
%    &\text{ neutrons from group $g$,} \nonumber \\
%    \Sigma^s_{g' \rightarrow g} =& \text{ macroscopic scattering cross section
%    for neutrons from} \nonumber \\
%    &\text{ groups $g'$ to $g$,} \nonumber \\
%    \chi^p_g =& \text{ prompt fission spectrum for neutrons in group $g$,} \nonumber \\
%    G =& \text{ total number of discrete neutron groups,} \nonumber \\
%    \nu_g =& \text{ average number of neutrons produced per fission,} \nonumber
%    \\
%    \Sigma^f_{g} =& \text{ macroscopic fission cross section for neutron}
%    \nonumber \\
%    &\text{ in group $g$,} \nonumber \\
%    \chi^d_g =& \text{ delayed fission spectrum for neutrons in group $g$,} \nonumber \\
%    I =& \text{ total number of \gls{DNP} groups,} \nonumber \\
%    \beta =& \text{ total delayed neutron fraction.} \nonumber
\end{align}

Despite forming only around 0.7\% of all neutrons emitted, delayed neutrons
play outsized roles in reactor kinetics. The relatively long half-lives of
\glspl{DNP} give reactor operators ample time in adequately designed reactors
to control reactor power output and intervene in case of power excursions.
The precursor concentration balance equations for $I$ precursor
groups are given by:
%
\begin{align}
    \frac{\partial C_i}{\partial t} =& \beta_i \sum^G_{g'=1} \nu \Sigma^f_{g'}
    \phi_{g'} - \lambda_i C_i - \vec{u} \cdot \nabla C_i + \nabla \cdot
    D_{\text{P}} \nabla C_i \label{eq:precursor} %\\
    %
%    \shortintertext{where}
%    \beta_i =& \text{ delayed neutron fraction of precursor group $i$,}
%    \nonumber \\
%    \lambda_i =& \text{ average decay constant of delayed neutron} \nonumber \\
%    &\text{ precursors in precursor group $i$,} \nonumber \\
%    C_i =& \text{ concentration of \gls{DNP}s in}
%    \nonumber \\
%    &\text{ precursor group $i$,} \nonumber \\
%    \vec{u} =& \text{ molten salt flow velocity vector,}
%    \nonumber \\
%    D_{\text{P}} =& \text{ effective diffusion coefficient of the delayed}
%    \nonumber \\
%    &\text{ neutron precursors.} \nonumber
\end{align}

These two equations are largely similar to conventional formulations of the
multigroup neutron diffusion equations with delayed neutrons for most reactor
types. The only differences are in the last two terms in Equation
\ref{eq:precursor}
which represent the advection and diffusion terms, respectively, to model the
movement of \glspl{DNP} in liquid-fuel \glspl{MSR}.

As shown in Equations \ref{eq:neutron} and \ref{eq:precursor}, Moltres requires
group constant data from dedicated high-fidelity neutronics software such as
the NEWT module in SCALE \cite{dehart_reactor_2011}, Serpent
\cite{leppanen_serpent_2014}, or OpenMC \cite{romano_openmc:_2015}. These group
constant data are the neutron energy group $g$ values for $v_g$, $D_g$,
$\Sigma^r_g$, $\Sigma^s_{g' \rightarrow g}$, $\chi^p_g$, $\chi^d_g$,
$\Sigma^f_{g}$, and $\nu\Sigma^f_{g}$, and precursor group $i$ values for
$\beta_i$ and $\lambda_i$. Users
can run a Python script in Moltres' Github repository, which automatically reads
user-provided SCALE/Serpent/OpenMC output data files and creates
Moltres-compatible JSON format files containing all required group constant
data. Moltres allows for an arbitrary number of neutron energy groups $G$ and
precursor groups $I$ as long as the user provides the necessary group constant
data. In practice, $I$ depends on the nuclear data library used to generate
group constants---the JEFF \cite{plompen_joint_2020} and ENDF
\cite{brown_endfb-viii0_2018} data libraries define eight precursor groups
and six precursor groups, respectively.

In multiphysics reactor simulations, we model the coupling between neutronics
and thermodynamics through temperature-dependent group constants. To sample
group constants at different temperatures in Moltres, users must provide group
constant data measured at more than one temperature (e.g., 800K--1500K at 100K
intervals). Users can then choose from linear spline, cubic spline, or monotone
cubic interpolation methods available in Moltres to interpolate the group
constant data for values falling within the provided temperature range.

Moltres provides two types of boundary conditions for neutron fluxes; these are
conventionally known as the vacuum and reflective boundary conditions given,
respectively, as:

\begin{align}
  D_g \nabla \phi_g \cdot \hat{n} + \frac{\phi}{2} =& 0 \label{eq:vacuum}
    \shortintertext{and}
  \nabla \phi \cdot \hat{n} =& 0
    \shortintertext{where}
  \hat{n} =& \mbox{ outward unit normal vector to the boundary.} \nonumber
\end{align}

The vacuum boundary condition typically applies to the external boundaries of
the reactor beyond which lies low-interaction media such as air. The
reflective boundary condition is useful for exploiting symmetries in the
model geometry, such as along the axial boundary in axisymmetric geometries. The
reflective boundary condition is equivalent to the more generally known
homogeneous Neumann boundary condition. Relevant boundary conditions for
\glspl{DNP} include the homogeneous Neumann boundary condition
along fuel salt-structural interfaces and outflow/inflow boundary
conditions along the outlet/inlet boundaries through which the precursors
flow as they circulate the fuel salt loop.

\subsubsection{Incompressible Flow Model} \label{sec:th}

Moltres relies on MOOSE's \texttt{Heat} \texttt{Conduction} and
\texttt{Navier-Stokes} physics modules for its thermal-hydraulics modeling
capabilities. While the \texttt{Navier-Stokes} module supports
compressible and incompressible flow modeling, this work focuses on
multiphysics coupling in \glspl{MSR} with the latter. The
time-dependent incompressible Navier-Stokes equations for velocity $\vec{u}$
with the Boussinesq approximation for buoyancy-driven flow are given as:

\begin{align}
    \text{Momentum equation: } \rho \frac{\partial \vec{u}}{\partial t} =&
    -\rho (\vec{u}
    \cdot \nabla) \vec{u} - \nabla p + \mu \nabla^2 \vec{u}
    + \rho \alpha \vec{g} \left(T - T_{\text{ref}} \right)
    \label{eq:momemtum}
    \shortintertext{and}
    \text{Mass equation: } \nabla \cdot \vec{u} =& 0
    \label{eq:divergence}
%    \shortintertext{where}
%    \rho =& \text{ fluid density,} \nonumber \\
%    p =& \text{ pressure,} \nonumber \\
%    \mu =& \text{ dynamic viscosity,} \nonumber \\
%    \alpha =& \text{ coefficient of thermal expansion,} \nonumber \\
%    \vec{g} =& \text{ gravitational force vector,} \nonumber
%    \\
%    T =& \text{ fluid temperature,} \nonumber \\
%    T_{\text{ref}} =& \text{ reference temperature at which the nominal}
%    \nonumber \\
%    &\text{ density is provided.} \nonumber
%    \nonumber
\end{align}

Velocity variables and advected quantities such as temperature are susceptible
to numerical node-to-node oscillations
commonly observed when resolving advection-dominated flows using continuous
Galerkin methods \cite{kuhlmann_lid-driven_2018}. The \texttt{Navier-Stokes}
module provides the \gls{SUPG} stabilization scheme
\cite{brooks_streamline_1982} for the velocity and temperature variables to
minimize these oscillations. The module also provides the \gls{PSPG}
stabilization scheme \cite{hughes_new_1986}, enabling equal-order
discretizations of pressure and velocity. Peterson et al. \cite{peterson_overview_2018}
provide further detail on the implementation of these stabilization schemes in
the \texttt{Navier-Stokes} module.

\subsubsection{Temperature Advection-Diffusion Model}

Lastly, Moltres solves for the temperature distribution through the temperature
advection-diffusion equation given by:

\begin{align}
    \rho c_{p} \frac{\partial T}{\partial t} =& - \rho c_p \vec{u}
    \cdot \nabla T + \nabla \cdot \left(k \nabla T \right) + Q_f - Q_s
    \label{eq:temp}
    \shortintertext{and}
    Q_f =& \sum^G_{g=1} \epsilon_g \Sigma_g^f \phi_g \label{eq:heat-source}
    \shortintertext{where}
%    c_p =& \text{ specific heat capacity of molten salt,} \nonumber \\
%    k =& \text{ effective thermal conductivity of molten salt,} \nonumber \\
%    Q_f =& \text{ fission heat source,} \nonumber \\
%    \epsilon_g =& \text{ average fission energy released by neutrons in group
%    $g$,} \nonumber \\
    Q_s =& \text{ heat sink/removal.} \nonumber
\end{align}

$Q_f$ represents the fission heat source term and is calculated by taking the
sum of neutron group fluxes multiplied by their respective macroscopic fission
cross sections and the average fission energy released per fission.

The \texttt{Navier-Stokes} module provides the following types of boundary
conditions for the velocity and temperature variables:

\begin{align}
    \text{Dirichlet: }& & u \ \left(\text{or } T\right) =& c & \\
    \text{Homogeneous Neumann: }& & \frac{\partial u}{\partial x_i} \
    \left(\text{or } \frac{\partial T}{\partial x_i}\right) =& 0 & \\
    \text{``No boundary condition'' outflow: }& &
    \left[ \nabla \vec{u} + \left(\nabla \vec{u} \right)^T \right] \cdot
    \hat{n} =& 0 \ \text{ (velocity)} & \\
    \text{``No boundary condition'' outflow: }& &
    k \nabla T \cdot\hat{n} =& 0 \ \text{ (temperature)} &
    \shortintertext{where}
    & & c =& \text{ user-defined constant value,} & \nonumber \\
    & & \hat{n} =& \text{ unit normal vector to the boundary.} & \nonumber
\end{align}

The Dirichlet boundary condition can be used to set the inlet velocities and
temperatures and no-slip conditions along solid boundaries. The
homogeneous Neumann boundary condition is commonly
imposed along the outlet boundary. However, the latter approach
artificially influences upstream behavior, especially in developing flow. The
``no boundary condition'' outflow boundary condition by Griffiths
\cite{griffiths_no_1997} has been shown to reduce such upstream errors.

\subsubsection{External Loop Model} \label{sec:moltres-loop}

Moltres also accounts for the decay of
\glspl{DNP} outside the active core region by simulating its flow in a
separate 1-D pipe geometry. This external loop pipe calculation is tightly
coupled to the active core simulation through Picard iterations in MOOSE's
MultiApp functionality and inlet/outlet boundary values.
The external loop region is assumed to be subcritical to minimize neutron
irradiation upon heat exchangers, pumps, and other equipment. Therefore, the
only significant neutronic-related phenomena are the drift and decay of
\glspl{DNP}. The governing equation for the \glspl{DNP} is:
%
\begin{align}
    \frac{\partial C_i}{\partial t} =& - \lambda_i C_i - u
    \frac{\partial C_i}{\partial x}.
    \label{eq:dnploop}
\end{align}
%
Equation \ref{eq:dnploop} is derived from equation \ref{eq:precursor} by
removing the fission \gls{DNP} source and diffusion terms and reducing the
dimensionality from 3-D to 1-D.

Moltres also simulates the temperature distribution in the external loop to model heat removal via
heat exchangers. The governing equation
for temperature, derived from equation \ref{eq:temp}, is:
%
\begin{align}
    \rho c_{p} \frac{\partial T}{\partial t} =& - \rho c_p u
    \frac{\partial T}{\partial x} - Q_{hx} \label{eq:temploop}
    \shortintertext{where}
    Q_{hx} =& \text{heat removal rate through the heat exchanger.} 
    \nonumber
\end{align}
%
The fission heat source term is replaced with a heat
exchanger sink term $Q_{hx}$.

Table \ref{table:loopbc} lists the boundary conditions for all variables on the inlet and outlet of
the 1-D external loop region. The inlet boundary conditions are all Dirichlet boundary conditions. The
prescribed value for the inlet boundary conditions are set to match the average outflow from the
active core region. The outlet boundary conditions are all outflow boundary conditions, as shown in
Table \ref{table:loopbc}.

\begin{table}[htbp!]
    \small
	\caption{Boundary conditions in the 1-D external loop geometry. $u$
	represents the 1-D velocity in this region.}
	\centering
	\begin{tabular}{ l l c}
		\toprule
		Variable & Boundary & Boundary Condition \\
        \midrule
        \multirow{2}{*}{Delayed neutron precursor concentration $C_i$} &
        Inlet (Core) & $C_i = c$ \\
        & Outlet (Core) & $u \cdot C_i = 0$ \\
        \midrule
        \multirow{2}{*}{Temperature $T$} &
        Inlet (Core) & $T = c$ \\
        & Outlet (Core) & $u \cdot T = 0$ \\
		\bottomrule
	\end{tabular}
	\label{table:loopbc}
\end{table}

\subsubsection{Core and External Loop Coupling Model}

This subsection details the \gls{DNP} and
temperature coupling between the core and external loop regions.

At every timestep, Moltres calculates weighted averages of the
temperature and the precursors at the outlet. These values are weighted by the
outflow velocity values at the outlet according to the following equation:
%
\begin{align}
    \overline{\psi} =& \frac{\int_\mathcal{C} Y(x_j) u(x_j) dx_j}{
    \int_\mathcal{C} u(x_j) dx_j} \\
    \shortintertext{where}
    Y =& \text{ variable to be weighted} \nonumber \\
    \mathcal{C} =& \text{ outlet boundary area} \nonumber \\
    u =& \text{ outflow velocity perpendicular to the outlet boundary,} \nonumber \\
    x_j =& \text{ spatial coordinate parallel to the outlet boundary.}
    \nonumber
\end{align}

Moltres transfers this outflow value from the core region to the 1-D
external loop region to be used as the boundary value for the inhomogeneous
Dirichlet boundary
condition at the inlet. Likewise, the outflow value from the external
loop region is used for the inflow value in the central core region. No
averaging is required for this step as the external loop region is a 1-D system.
This approach results in uniform inflow temperature and \gls{DNP} at the
inlet. The Picard iterations within every timestep ensure the two systems
are tightly coupled.

\subsection{Previous \gls{MSR} Analyses with Moltres} \label{sec:moltres-previous}

This section discusses some previous work with Moltres to illustrate its
various capabilities and coupling approaches for multiphysics \gls{MSR}
modeling and simulation. Section \ref{sec:msre} summarizes the work by Lindsay et
al. \cite{lindsay_introduction_2018} in modeling the \gls{MSRE}, and Section
\cite{park_advancement_2020} summarizes my previous work in modeling the
\gls{MSFR}.

\subsubsection{Introduction to Moltres and Modeling the MSRE} \label{sec:msre}

This section follows the work by Lindsay et al. in \textit{Introduction to Moltres:
An Application for Simulation of Molten Salt Reactors}
\cite{lindsay_introduction_2018}.

In 2017, Lindsay et al. introduced
Moltres to the \gls{MSR} community for multiphysics simulations of \glspl{MSR}.
Their work showcased neutron diffusion and thermal-hydraulics coupling
capabilities in Moltres. The authors
demonstrated these capabilities by running time-dependent simulations of 2-D
axisymmetric and 3-D \gls{MSRE} models until the flux, precursor, and
temperature distributions reached a steady state.

\begin{figure}[htb!]
	\centering
	\includegraphics[width=.45\columnwidth]{msre-geometry}
	\caption{Schematic diagram of the 2-D axisymmetric \gls{MSRE} geometry
	adopted by Lindsay et al. \cite{lindsay_introduction_2018}.}
	\label{fig:msre-geometry}
\end{figure}

Figure \ref{fig:msre-geometry} shows the fuel channels and moderator regions of
the 2-D \gls{MSRE} geometry that Lindsay et al. adopted for their study.
They ran a two-group neutron diffusion model with six precursor groups and
vacuum boundary conditions on the outer boundaries governed by Equations
\ref{eq:neutron}, \ref{eq:precursor}, and \ref{eq:vacuum} shown in Section
\ref{sec:nts}. They modeled precursor drift due to fuel salt flow by imposing
fixed uniform flow upwards through the fuel channels shown in Figure
\ref{fig:msre-geometry}. Their thermal-hydraulics model employed a
governing equation for temperature in the fuel salt equivalent to Equation
\ref{eq:temp} with fixed uniform flows while imposing a cosine-shaped heat
source term representing heat dissipation from gamma and neutron irradiation in
the graphite moderator region. In addition, all governing equations were fully
coupled and solved simultaneously as a single system of equations with implicit
Euler timestepping to accurately and efficiently resolve the strong coupling
expected between the neutronics and temperature.

\begin{figure}[htb!]
	\centering
	\includegraphics[width=.45\columnwidth]{2d_gamma_heating_group1}
	\includegraphics[width=.45\columnwidth]{2d_gamma_heating_group2}
	\caption{Neutron group 1 and 2 fluxes in the 2-D axisymmetric \gls{MSRE}
	model from Lindsay et al. \cite{lindsay_introduction_2018}.}
	\label{fig:msre-flux}
\end{figure}

\begin{figure}[htb!]
	\centering
	\includegraphics[width=.45\columnwidth]{2d_gamma_heating_pre1_scaled}
	\includegraphics[width=.45\columnwidth]{2d_gamma_heating_pre6_scaled}
	\caption{Longest- and shortest-lived precursor concentrations ($\lambda =
	1.24\times 10^{-2}$s$^{-1}$ and $3.07$s${-1}$, respectively) in the 2-D
	axisymmetric \gls{MSRE} model from Lindsay et al.
	\cite{lindsay_introduction_2018}.}
	\label{fig:msre-precursor}
\end{figure}

\begin{figure}[htb!]
	\centering
	\begin{minipage}[b]{0.45\columnwidth}
	    \includegraphics[width=\columnwidth]{2d_gamma_heating_temp}
	    \caption{Temperature distribution in the 2-D
	    axisymmetric \gls{MSRE} model from Lindsay et al.
	    \cite{lindsay_introduction_2018}.}
	    \label{fig:msre-temp}
	\end{minipage}
	\hfill
	\begin{minipage}[b]{0.45\columnwidth}
	    \includegraphics[width=\columnwidth]{combined_msre_moltres_axial_temps}
	    \caption{Moltres \cite{lindsay_introduction_2018} and \gls{ORNL}
	    \gls{MSRE} \cite{briggs_molten-salt_1964} axial temperature
	    distributions in the hottest fuel channel and adjacent graphite.}
	    \label{fig:msre-temp-plot}
	\end{minipage}
\end{figure}

Figure \ref{fig:msre-flux} shows the fast and thermal neutron fluxes
corresponding to group 1 and 2 in the 2-D \gls{MSRE} model. As expected, the
fluxes exhibit general cosine shapes in the axial and radial directions. We
also observe minor oscillations in the radial direction coinciding with the
regular fuel and moderator lattice. The fuel regions favor the fast flux, while
the moderator regions favor the thermal flux.

Figure \ref{fig:msre-precursor}
shows the longest- and shortest-lived precursor concentrations in the fuel
channels. With a long half-life of 55.9 s relative to the 6.91 s it takes for
salt to flow from bottom to top, the longest-lived precursor concentration peaks
outside the model domain. By contrast, the shortest-lived precursor
concentration closely follows the neutron fluxes' cosine shape, which
dictate where the precursors are born.

Finally, Figure \ref{fig:msre-temp}
shows the temperature distribution in the 2-D \gls{MSRE} model. The temperature
naturally peaks near the outlet due to upward advection. The moderator regions
experience hotter temperatures than the fuel regions due to radiative heating
and the relative inefficiency of heat conduction in the graphite compared to
advection in the fuel salt.

\begin{figure}[htb!]
	\centering
	\begin{subfigure}[h]{0.45\columnwidth}
	    \includegraphics[width=\columnwidth]{combined_msre_moltres_radial}
	    \caption{Radial fluxes at reactor half-height.}
	    \label{fig:msre-flux-radial}
	\end{subfigure}
	\hfill
	\begin{subfigure}[h]{0.45\columnwidth}
	    \includegraphics[width=\columnwidth]{combined_msre_moltres_axial}
	    \caption{Axial fluxes along the core centerline ($r=0$ cm).}
	    \label{fig:msre-flux-axial}
	\end{subfigure}
	\caption{The fast and thermal fluxes from
	Moltres \cite{lindsay_introduction_2018} and the \gls{ORNL} \gls{MSRE}
	design calculations \cite{briggs_molten-salt_1964}.}
\end{figure}

The corresponding neutronics and thermal-hydraulics results from their 3-D model
show good qualitative agreement with the 2-D results. Given the lack of
\gls{MSR} experimental data, they compared their 2-D model results with
\gls{MSRE} design calculations performed using legacy software in 1963-1964 at
\gls{ORNL}. Their neutron flux and temperature distribution results showed good
qualitative agreement with \gls{ORNL} data. The authors attributed
discrepancies to the following differences in the two modeling approaches: the
absence of axial heat conduction and the use of 32 neutron groups in the
\gls{ORNL} calculations, and the exclusion of control rod thimbles in the
Moltres calculations.

\paragraph{Critical Assessment} \label{sec:msre-critique}

Lindsay et al.'s work illustrated early development efforts for neutronics and thermal-hydraulics
models in Moltres. It demonstrated, along with advanced capabilities from the \gls{MOOSE}
framework, fully-coupled simulations of the \gls{MSRE} with implicit timestepping. The
decent qualitative agreement observed between Moltres and the \gls{ORNL}
\gls{MSRE} calculations proved that Moltres could simulate
\glspl{MSR} under some simplifying assumptions.

On the other hand, significant improvements could be made to Moltres to better
model multiphysics phenomena in \glspl{MSR}. For instance, replacing the
fixed uniform salt flow with a proper flow profile governed by fluid flow
equations would accurately capture precursor and temperature advection.
Temperature advection has a particularly large impact on the temperature
distribution in the fuel salt since molten salts generally have large Prandtl
numbers, which measure the ratio of convective to conductive heat transfer.
The flow-modeling feature would be of even greater importance when modeling
pool-type \glspl{MSR}, which consist of a single large fuel salt region in the
reactor core.

Another essential feature for modeling \glspl{MSR}, which was already under
development at the time of publication, is a precursor loop
system to recirculate precursors back into the core. While some precursors
decay outside the core, others survive long enough to recirculate back into the
core. The loop system would provide a more accurate estimate of the delayed
neutron fraction instead of discarding all precursors that flowed out of
the core. In transient simulations involving sudden increases in the neutron
flux, precursors recirculating into the core can induce observable jumps and
dips in the power output due to the associated reactivity insertion from the
delayed neutrons.

Moltres would also benefit from a decay heat model which Lindsay
et al. mentioned in their work \cite{lindsay_introduction_2018},
especially for transient accident analyses. While decay heat from fission
product decays represents a small fraction ($\sim5\%$) of total power output,
this heat source can be significant in unprotected loss of flow or loss of
secondary cooling accidents. Therefore, understanding residual heat generation
from fission product decays in \gls{MSR} is essential in preventing further
structural failure after an accident.

Specific to reactor physics modeling, Moltres also requires an accurate method for modeling
control rods. As observed in Figure \ref{fig:msre-flux}, the omission of control rods led to
significant discrepancies in the thermal flux distribution. However, developing control rod
modeling capability in Moltres is nontrivial given that neutron diffusion theory performs poorly
within highly neutron-absorbing regions.

Lastly, the authors recognized two other avenues for future work: further \gls{VV} of Moltres'
capabilities by comparison
with more detailed experimental data and other modern \gls{MSR} modeling
efforts; and the study of transient simulation cases such as control rod
ejection, single channel blockage, loss of flow, and loss of secondary cooling.

\subsubsection{Advancements in Moltres and Transient Simulations of the MSFR}
\label{sec:msfr}

This section follows my previous work for my Master's degree in \textit{Advancement and
Verification of Moltres for Molten Salt Reactor Safety Analysis} \cite{park_advancement_2020},
which I shall refer to as ``this work'' in this subsection.

In many ways, this work represents a continuation of the work by Lindsay et al.
\cite{lindsay_introduction_2018} in developing Moltres as an \gls{MSR}
simulation tool. New features introduced include coupling the existing
neutronics and temperature equations to incompressible Navier-Stokes equations
to model flow dynamics, the precursor loop system, and the decay heat model.
I later demonstrated these features through steady-state and transient
simulations of a 2-D axisymmetric \gls{MSFR} model.

Moltres couples natively with the \texttt{Navier-Stokes} module in \gls{MOOSE} for
incompressible flow modeling capabilities. Section \ref{sec:th} describes the governing equations
for incompressible flow and temperature advection-diffusion.
The precursor loop model was developed to model
precursor recirculation into the core. The precursor loop system leverages
the \texttt{MultiApp} system in \gls{MOOSE} to couple a 1-D pipe model to the
core to simulate precursor flow outside the core. The core and the pipe models
are coupled via inlet and outlet boundary conditions, as detailed in Section
\ref{sec:moltres-loop}. The loop system can also accommodate a pointwise heat
exchanger model with a heat removal rate as a function of time
or auxiliary variables in the Moltres simulation.

In addition, I developed a decay heat model to simulate delayed heat release
from the decay of fission products, similar to delayed neutron emission
from precursors. The decay heat model introduces decay heat groups of different
decay constants and variables representing the decay heat groups and modifies the
heat source term given by Equation \ref{eq:heat-source} to account for
the delayed heat generation.

The number of decay heat groups depends on
the decay heat profile of the reactor after a specified period of operation.
Determining the decay heat profile requires fuel depletion calculations to
obtain the fuel composition during/after reactor operation and subsequent
calculations for the cumulative decay heat released from individual isotopes.
In this work, I cited results by Aufiero et al. \cite{aufiero_extended_2013},
who determined that three decay heat groups ($J=3$) were sufficient to
reproduce the decay heat profile of the \gls{MSFR} equilibrium fuel composition
for up to 300 seconds after shutdown with a relative error of less than 2\%.

\begin{figure}[htb!]
	\centering
	\includegraphics[width=.55\columnwidth]{central-core-legend}
	\caption{Schematic diagram of the 2-D axisymmetric \gls{MSFR} model with
	the core region enclosed by the red box \cite{park_advancement_2020}.}
	\label{fig:msfr-geometry}
\end{figure}

For the verification and demonstration of the capabilities introduced in this
work, I ran coupled steady-state and unprotected accident transient
simulations of the \gls{MSFR} and compared the results with published results
by Fiorina et al. \cite{fiorina_modelling_2014} and Aufiero et al.
\cite{aufiero_development_2014}. In this section, the PoliMi and TU Delft models
refer to two sets of results published by Fiorina et al. I compensated for
the lack of a turbulence model in Moltres by imposing a fixed turbulent
viscosity of 40 Pa$\cdot$s in addition to the molecular viscosity $\mu$ in
Equation \ref{eq:momemtum}. Figure \ref{fig:msfr-geometry} shows the 2-D
axisymmetric \gls{MSFR} model with the core region indicated with the red box,
while the rest of the fuel region was modeled with the 1-D loop system. Fuel
salt flows into the core region from the inlet at the bottom-right corner and
out through the outlet at the top-right corner.

\begin{figure}[htb!]
    \centering
    \begin{subfigure}[t]{.35\textwidth}
        \centering
        \includegraphics[width=\textwidth]{flow-temp-plasma}
    \end{subfigure}
    \hfill
    \begin{subfigure}[t]{.625\textwidth}
        \centering
        \includegraphics[width=\textwidth]{flow-temp-fiorina}
    \end{subfigure}
    \caption{Temperature and velocity fields in the core from Moltres
    (left), PoliMi (center), and TU Delft (right) models. The colors represent
    temperature according to the respective color bars and the arrows
    represent velocity fields. \cite{park_advancement_2020}}
    \label{fig:flow-temp}
\end{figure}

Figure \ref{fig:flow-temp} shows the core steady-state temperature and velocity
fields from the Moltres, PoliMi, and TU Delft models. The Moltres
model showed
good qualitative agreement with the PoliMi and TU Delft models since the plots
show similar flow and hotspot features in all three models. The salt flow
follows a parabolic path from the inlet to the outlet. A large
recirculation region formed near the right wall while the top and bottom
regions along the central axis experience relatively stagnant flow compared to
the main salt flow stream. The temperature hotspots coincide with the regions of
recirculation and stagnation because convection is the dominant heat transfer
mechanism. Maximum temperatures are observed near the bottom of the
recirculation zones; the Moltres model reports 1275 K, which is closer to the
PoliMi model ($\sim1300$ K) than the TU Delft model ($\sim1200$ K). Thus, the
incompressible flow model in Moltres performed well in reproducing the
thermal-hydraulic profiles of the \gls{MSFR} under steady-state conditions
despite the fixed turbulent viscosity assumption.

\begin{figure}[b!]
    \centering
    \begin{subfigure}[t]{.30\textwidth}
        \centering
        \vspace{.9cm}
        \includegraphics[width=\textwidth]{pre}
    \end{subfigure}
    \begin{subfigure}[t]{.69\textwidth}
        \centering
        \vspace{0pt}
        \includegraphics[width=\textwidth]{fiorina-pre}
    \end{subfigure}
    \caption{Total delayed neutron source distribution in the core from the
    Moltres (left), PoliMi (center), and TU Delft (right) models.}
    \label{fig:pre}
\end{figure}

Moltres reported a peak total neutron flux of $9.80 \times 10^{15}$
cm$^{-2}\cdot$s$^{-1}$, which is close to the 8.6 and 9.0 $\times 10^{15}$
cm$^{-2}\cdot$s$^{-1}$ values reported by Fiorina et al.
\cite{fiorina_molten_2013} and \cite{aufiero_development_2014}. However, the
delayed neutron source distribution from the Moltres model exhibits some
differences compared to the PoliMi and TU Delft models, as shown in Figure
\ref{fig:pre}. The Moltres model retains fewer precursors than the other two
models, resulting in a higher 44.16\% out-of-core delayed neutron emissions
compared with the 34.80\% and 34.85\% from the other models
\cite{park_advancement_2020}. Therefore, the incompressible flow model with
constant turbulent viscosity fell short of accurately capturing precursor
drift in the \gls{MSFR}.

\begin{figure}[htb]
    \centering
    \includegraphics[width=.5\textwidth]{decay-heat-temp}
    \caption{Difference in core temperatures at steady-state with decay heat
    relative to the result without decay heat (Figure \ref{fig:flow-temp}).}
    \label{fig:decayheattemp}
\end{figure}

This work also compared steady-state temperature distributions with and without
the decay heat model. As expected, the decay heat model effectively flattens
the heat source distribution, reducing the maximum
temperatures and increasing the minimum temperatures observed in the model.
Figure \ref{fig:decayheattemp} shows a decrease in temperature in the hotspot
regions and an increase in temperature near the inlet, the coolest
region in the core.

For the transient studies, this work considered unprotected reactivity
insertion, loss-of-heat-sink, loss-of-flow, and pump-overspeed accidents. The
term ``unprotected'' signifies accident scenarios without reactor SCRAM.
Moltres reproduced similar trends observed in the PoliMi and TU Delft
models for the reactivity insertion and loss-of-heat-sink scenarios. For
instance, Figure \ref{fig:200pcmheat} shows the power output response of the
three models following a 200 pcm step reactivity insertion. The lower peak from
the Moltres model is attributed to the stronger negative fuel temperature
reactivity feedback coefficient of $-7.184$ pcm$\cdot$K$^{-1}$ compared to
approximately $-6.5$ pcm$\cdot$K$^{-1}$ for the PoliMi and TU Delft models.

\begin{figure}[htb!]
    \centering
    \begin{minipage}[b]{.49\textwidth}
      \centering
      \includegraphics[width=\textwidth]{200pcm-heat}
      \caption{Power output following a 200 pcm step-wise unprotected reactivity
        insertion in the Moltres, PoliMi, and
        TU Delft models \cite{fiorina_modelling_2014}.}
      \label{fig:200pcmheat}
    \end{minipage}
    \hfill
    \begin{minipage}[b]{.49\textwidth}
      \centering
      \includegraphics[width=\textwidth]{lohs-heat}
      \caption{Power output during
        an unprotected loss of heat sink transient in the Moltres, PoliMi, and
        TU Delft models \cite{fiorina_modelling_2014} without decay heat.}
      \label{fig:lohsheat}
    \end{minipage}
\end{figure}

The loss-of-heat-sink transients were performed with and without the decay heat
model because the TU Delft model did not possess this capability. As shown in
Figure \ref{fig:lohsheat}, without decay heat, all three models reported
similar power output responses following the loss of cooling through the heat
exchanger modeled as an exponential decrease in heat removal rate with a time
constant of 1 s. As prompt heat generation decreases due to rising core
temperatures and the negative temperature reactivity feedback, decay heat
becomes a significant heat source. We observe this in Figure
\ref{fig:moltresdecaypower}, which shows decay heat output overtaking prompt
heat output 34 s into the accident scenario. With the decay heat model, the
Moltres model shows good agreement with the PoliMi model. The temperature
increase averaged over the fuel salt loop falls within 10\% of the PoliMi
model (Figure \ref{fig:polimidecaytemp}).

\begin{figure}[htb!]
	\centering
	\begin{minipage}[t]{0.485\columnwidth}
	    \includegraphics[width=\columnwidth]{moltres-decay-power}
	    \caption{Power output during
    an unprotected loss of heat sink transient in the Moltres model with
    decay heat.}
	    \label{fig:moltresdecaypower}
	\end{minipage}
	\hfill
	\begin{minipage}[t]{0.485\columnwidth}
	    \includegraphics[width=\columnwidth]{decay-temp}
	    \caption{Loop-averaged temperature increase during
    an unprotected loss of heat sink transient in the Moltres and PoliMi
    models \cite{fiorina_modelling_2014} with decay heat.}
	    \label{fig:polimidecaytemp}
	\end{minipage}
\end{figure}

On the other hand, for the pump-initiated accident scenarios, significant
changes in the flow affected the validity of the uniform turbulent viscosity
assumption. These transients required ad hoc adjustments to the uniform
turbulent viscosity assumption as a function of volumetric flow to reproduce
the trends observed in the PoliMi and TU Delft models. Furthermore, unlike the
other two models, the Moltres model did not apply the Boussinesq approximation
for buoyancy-driven flow. The Moltres model could replicate expected trends in
the pump overspeed scenario (Figure \ref{fig:poshort}), but performed
poorly in the loss of flow scenario (Figure \ref{fig:lof}). In the latter
scenario, buoyancy effects become significant as the model loses forced flow.

\begin{figure}[htb!]
    \centering
    \begin{subfigure}[t]{.485\textwidth}
        \centering
        \includegraphics[width=\textwidth]{po-heat-short}
    \end{subfigure}
    \hfill
    \begin{subfigure}[t]{.485\textwidth}
        \centering
        \includegraphics[width=\textwidth]{po-temp-short}
    \end{subfigure}
    \caption{Average core temperature increase during
    an unprotected pump overspeed transient in the Moltres, PoliMi, and
    TU Delft models \cite{fiorina_modelling_2014}.}
    \label{fig:poshort}
\end{figure}

\begin{figure}[htb!]
    \centering
    \begin{subfigure}[t]{.485\textwidth}
        \centering
        \includegraphics[width=\textwidth]{lof-heat}
    \end{subfigure}
    \hfill
    \begin{subfigure}[t]{.485\textwidth}
        \centering
        \includegraphics[width=\textwidth]{lof-temp}
    \end{subfigure}
    \caption{Power output and average core temperature increase during
    an unprotected loss of flow transient in the Moltres, PoliMi, and
    TU Delft models \cite{fiorina_modelling_2014}.}
    \label{fig:lof}
\end{figure}

\paragraph{Critical Assessment} \label{sec:msfr-critique}

Given the inherent and unique characteristics of \glspl{MSR}, the new
capabilities introduced in this work---coupling to incompressible Navier-Stokes
equations, precursor loop system, and decay heat model---are essential for
accurate \gls{MSR} modeling and simulation. This work demonstrated these
capabilities through steady-state and transient studies of a 2-D axisymmetric
\gls{MSFR} model similar to models by Fiorina et al.
\cite{fiorina_modelling_2014} and Aufiero et al.
\cite{aufiero_development_2014}. The Moltres model showed good agreement with
the other models in
most steady-state and transient cases with respect to the neutron flux, reactor
power, temperature, and velocity distributions. Most crucially, the
incompressible flow model reproduced characteristic regions of recirculating
flow and near-stagnant flow observed in the PoliMi and TU Delft \gls{MSFR}
models, which led to the formation of temperature hotspots and precursor
accumulation in the core.

However, significant discrepancies observed during pump-initiated transient
scenarios highlight the need for a proper turbulence model to capture
turbulent flow effects in some \gls{MSR} designs. Unlike the \gls{MSRE}, the
\gls{MSFR} experiences highly turbulent flow, whose Reynolds number is on the
order of $10^6$ under normal operating conditions. This level of turbulent flow
produces eddies of a wide range of length scales, and the computational demands
of the fine mesh and time resolution render it numerically unsolvable with
today's computational resources. Turbulence models allow for cheaper turbulence
simulations on coarser meshes by approximating the turbulent effects through
statistical analysis. Fiorina et al. \cite{fiorina_modelling_2014} employed the
$k$-$\epsilon$ turbulence model in their PoliMi and TU Delft models. Moltres
will benefit from coupling to similar intermediate-fidelity turbulence models
for turbulence simulations at reasonable computational costs.

While the Moltres model demonstrated good agreement with published data in the
steady-state, reactivity, and loss of heat sink scenarios, this work does not
thoroughly verify Moltres' capabilities for \gls{MSR} modeling.
The \gls{MSFR} simulations involve various physics models which combine to form
a complex multiphysics model. Therefore, it is difficult to pinpoint sources of
discrepancy with a high degree of certainty. Differences in the modeling
approaches also introduce discrepancies that cannot be reliably identified.
Moltres will benefit from code-to-code verification of
individual components responsible for modeling various physics present in
\glspl{MSR} and Moltres' multiphysics coupling approach.

### 2024-06-20 hybrid.tex
Fix typos and add numerical implementation section

@@
   \sum^G_{g=1}\sum^{N_d}_{d=1}w_d\left(\Psi^*_{g,d}+\tau_g\hat{\Omega}_d\cdot\Psi^*_{g,d},
-  \frac{\Psi_{g,d}}{v_g}\right)_\mathcal{D}
+  \frac{\Psi_{g,d}}{v_g}\right)_\mathcal{D} \label{eq:time-derivative}
 \end{gather}
@@
     & \hat{\Omega}\cdot\hat{n}_b<0,\vec{r}\in\partial\mathcal{D}_s
-  \end{cases},
+  \end{cases}, \label{eq:reflecting-bc}
   \shortintertext{where}
@@
 can be introduced by modifying the neutron diffusion equations.
-In this work, we explored two options for applying
+In this work, I explored two options for applying
 transport corrections to the neutron diffusion equations: applying
 diffusion corrections or adding a drift correction term.
+I performed preliminary investigations on 1-D test cases with both types of transport correction. I
+eventually elected to use the drift correction term for the remaining test cases and reactor model
+demonstrations due to superior numerical properties. I will make clear the reasons for this
+selection through 1-D test case results in Chapter 5.
 
@@
 
+In this section, I detail the numerical implementation of the hybrid $S_N$-diffusion method with
+drift correction terms. As mentioned in Section \ref{sec:transport-correction}, I opted for the
+drift correction terms over diffusion correction for investigations beyond 1-D test cases.
+Therefore, for brevity, I moved numerical implementation details for the hybrid method with
+diffusion correction to Appendix A.
+
+In Section \ref{sec:moltres-description}, I detailed the code structure of Moltres
+\cite{lindsay_moltres_2017} in relation to the underlying \gls{MOOSE} finite-element framework
+\cite{giudicelli_30_2024} and the pre-existing multigroup neutron diffusion and thermal-hydraulics
+models. I implemented the hybrid $S_N$-diffusion model in Moltres to
+leverage existing capabilities and functions that are also relevant to the new $S_N$ neutron
+transport model and the $S_N$-to-diffusion coupling features. I also leveraged several basic
+functions from \gls{MOOSE} that are useful for setting up physics models.
+
+All source code for the hybrid $S_N$-diffusion model can be found in my Moltres GitHub fork
+repository\footnote{\url{https://github.com/smpark7/moltres/tree/sn-method}} as of the time of
+writing. This branch may be merged into the original upstream Moltres GitHub repository in future.
+
+Moltres presently comes with a Python script for parsing and extracting neutron group constant data
+from OpenMC \cite{boyd_multigroup_2019} or Serpent 2 \cite{leppanen_serpent_2014} simulation
+output files. The script reorganizes the group constant data into a
+JSON format file. I expanded the Python script to extract
+the macroscopic total cross sections $\Sigma_{t,g}$ and the higher-order macroscopic Legendre
+scattering moments $\Sigma^{g'\rightarrow g}_{s,l}$ for the $S_N$ neutron transport model. A new
+\texttt{Material} class called \texttt{MoltresSNMaterial} creates material
+property variables for each group constant for the $S_N$ model to read during a neutronics
+simulation. If the user provides group constant data at multiple reactor temperature states,
+\texttt{MoltresSNMaterial} can interpolate the data to simulate temperature-dependent reactivity
+changes.
+
+The new \texttt{MoltresUtils} class contains ``utility'' functions required by the $S_N$ model.
+These functions include: a function to retrieve directions and weights defined by a
+level-symmetric quadrature set of order $N$, a function to retrieve reflecting directions at
+reflecting boundaries, and a function to compute spherical harmonic flux moments for anisotropic
+scattering. I pre-computed the direction and weight values for the level-symmetric quadrature sets
+using a Python script and hard-coded them in Moltres to reduce computational burden during
+simulations.
+
+The \gls{SAAF} $S_N$ method defines $G$ energy groups and $N_d$ discrete directions for a total
+of $G\times N_d$ angular flux variables $\Psi_{g,d}$. The Moltres $S_N$ model groups the
+$\Psi_{g,d}$ variables by energy group into $G$ \textit{array variables}. An array variable is a
+set of standard field variables to help with dealing with setting up simulations with high
+dimensionality. Each standard variable of an array variable is referred to as a component in the
+array variable. In this work, I opted for
+2nd-order Lagrange shape functions to approximate the $\Psi_{g,d}$ variables in the finite-element
+analysis. The 2nd-order approximation showed better convergence rates in mesh refinement studies.
+
+The terms in the \gls{SAAF} $S_N$ equations (Eqs.\ \ref{eq:time-derivative}-\ref{eq:reflecting-bc})
+are represented by kernel and BC classes inheriting from \texttt{ArrayKernel} and
+\texttt{ArrayIntegratedBC} template classes. Table \ref{tab:saaf-sn} lists the class names
+corresponding to each term in Eq.\ \ref{eq:saaf-vt}.
+%
+\begin{table}[htb]
+  \centering
+  \caption{Names of the Kernels and BCs representing the \gls{SAAF} $S_N$ equations.}
+  \begin{tabular}{l l l}
+    \toprule
+    Term name & Class name \\
+    \midrule
+    Time derivative & \texttt{SNTimeDerivative} \\
+    Streaming & \texttt{SNStreaming} \\
+    Collision & \texttt{SNCollision} \\
+    Scattering & \texttt{SNScattering} \\
+    Prompt fission source & \texttt{SNFission} \\
+    Delayed neutron source & \texttt{SNDelayed} \\
+    Vacuum boundary & \texttt{SNVacuumBC} \\
+    Reflecting boundary & \texttt{SNReflectingBC} \\
+    Hybrid method boundary source & \texttt{SNDiffusionBC} \\
+    \bottomrule
+  \end{tabular}
+  \label{tab:saaf-sn}
+\end{table}
+
+The \gls{MOOSE} framework leverages the Eigen \cite{guennebaud_eigen_2010} C++ template library for
+linear algebra operations defined in ArrayKernels and ArrayBCs for array variables. The \gls{SAAF}
+$S_N$ kernels and BCs contain functions for computing the residual and Jacobian contributions of
+each term. The \gls{FEM} numerical solver deems to have reached a converged solution if the sum of
+residual contributions falls below a defined tolerance value. All $S_N$ simulations in this work
+employ the \gls{PJFNK} method with the HYPRE-BoomerAMG method for preconditioning. BoomerAMG is an
+algebraic multigrid method from the HYPRE \cite{hypre_hypre_2022} linear solver library.
+
+The drift correction terms $\vec{D}_g$ are initialized as auxiliary array variables
+with three components corresponding to their $x$, $y$, and $z$ components. Auxiliary variables
+encompass all field variables which are not directly determined during the \gls{PJFNK}
+calculation stage. The drift and boundary correction variables are computed during the
+post-calculation stage from $\Psi_{g,d}$.
+
+I employed the \texttt{MultiApp} system \cite{gaston_physics-based_2015} to implement the fixed point
+iterative coupling between the $S_N$ and neutron diffusion solvers for the hybrid $S_N$-diffusion method.
+The \texttt{MultiApp} and \texttt{Transfers} systems from \gls{MOOSE} provide flexible interfaces
+for setting up iterative solvers and transferring data between the solvers. A typical hybrid
+method simulation on Moltres starts with the neutron diffusion solver for the first estimate of the
+neutron flux distribution. Moltres transfers the neutron flux distribution to the $S_N$ solver to
+compute the boundary source distribution. The $S_N$ solver then generates its own flux estimate and
+computes drift and boundary correction variables for the neutron diffusion solver. Moltres
+transfers the correction variables to the neutron diffusion solver to compute its next flux
+estimate. This process continues until the neutron flux distribution converges to within tolerance
+values provided in the input file.
+
 \section{Summary} \label{sec:hybrid-summary}

### 2024-06-20 lit.tex
add verification chapter for referencing, hybrid chapter edits
Fix typos and add numerical implementation section

@@
 OpenFOAM, and the open-source finite-element
-framework \gls{MOOSE} \cite{gaston_physics-based_2015}. Researchers at
+framework \gls{MOOSE} \cite{giudicelli_30_2024}. Researchers at
 \gls{PoliMi} developed an \gls{MSR} simulation tool in COMSOL and
@@
 \begin{gather}
-  \nabla\cdot J^{(s+1)}+\Sigma_a\phi^{(s+1)} = q, \\
+  \nabla\cdot J^{(s+1)}+\Sigma_a\phi^{(s+1)} = q, \label{eq:2nd-moment-1} \\
   \frac{1}{3}\nabla\cdot\phi^{(s+1)}+\Sigma_t J^{(s+1)} = \nabla\cdot F^{(s+1/2)}
+  \label{eq:2nd-moment-2}
   \shortintertext{where}
@@
   \left(\frac{1}{3}\delta_{\alpha\beta}-\Omega_\alpha\Omega_\beta\right)
-  \Psi^{(s+1/2)} d\Omega
+  \Psi^{(s+1/2)} d\Omega \label{eq:2nd-moment-3}
 \end{gather}

### 2024-06-20 moltres.tex
Fix typos and add numerical implementation section

@@
-\section{Description of Moltres}
+\section{Description of Moltres} \label{sec:moltres-description}
 
@@
 Moltres draws many advantages from being developed on MOOSE
-\cite{permann_moose_2020}. MOOSE is an open-source, finite-element,
+\cite{giudicelli_30_2024}. MOOSE is an open-source, finite-element,
 multiphysics framework developed at \gls{INL}. The framework provides a
@@
     \phi_{g'} + \chi^d_g \sum^I_i \lambda_i C_i \label{eq:neutron} %\\
-%    %
-%    \shortintertext{where}
-%    v_g =& \text{ average speed of neutrons in group $g$,} 
-%    \nonumber \\
-%    \phi_g =& \text{ neutron flux in group $g$,}
-%    \nonumber \\
-%    t =& \text{ time,} \nonumber \\
-%    D_g =& \text{ diffusion coefficient of neutrons in group $g$,} \nonumber \\
-%    \Sigma^r_g =& \text{ macroscopic removal cross section for} \nonumber \\
-%    &\text{ neutrons from group $g$,} \nonumber \\
-%    \Sigma^s_{g' \rightarrow g} =& \text{ macroscopic scattering cross section
-%    for neutrons from} \nonumber \\
-%    &\text{ groups $g'$ to $g$,} \nonumber \\
-%    \chi^p_g =& \text{ prompt fission spectrum for neutrons in group $g$,} \nonumber \\
-%    G =& \text{ total number of discrete neutron groups,} \nonumber \\
-%    \nu_g =& \text{ average number of neutrons produced per fission,} \nonumber
-%    \\
-%    \Sigma^f_{g} =& \text{ macroscopic fission cross section for neutron}
-%    \nonumber \\
-%    &\text{ in group $g$,} \nonumber \\
-%    \chi^d_g =& \text{ delayed fission spectrum for neutrons in group $g$,} \nonumber \\
-%    I =& \text{ total number of \gls{DNP} groups,} \nonumber \\
-%    \beta =& \text{ total delayed neutron fraction.} \nonumber
+    %
+    \shortintertext{where}
+    v_g =& \text{ average speed of neutrons in group $g$,} 
+    \nonumber \\
+    \phi_g =& \text{ neutron flux in group $g$,}
+    \nonumber \\
+    t =& \text{ time,} \nonumber \\
+    D_g =& \text{ diffusion coefficient of neutrons in group $g$,} \nonumber \\
+    \Sigma^r_g =& \text{ macroscopic removal cross section for} \nonumber \\
+    &\text{ neutrons from group $g$,} \nonumber \\
+    \Sigma^s_{g' \rightarrow g} =& \text{ macroscopic scattering cross section
+    for neutrons from} \nonumber \\
+    &\text{ groups $g'$ to $g$,} \nonumber \\
+    \chi^p_g =& \text{ prompt fission spectrum for neutrons in group $g$,} \nonumber \\
+    G =& \text{ total number of discrete neutron groups,} \nonumber \\
+    \nu_g =& \text{ average number of neutrons produced per fission,} \nonumber
+    \\
+    \Sigma^f_{g} =& \text{ macroscopic fission cross section for neutron}
+    \nonumber \\
+    &\text{ in group $g$,} \nonumber \\
+    \chi^d_g =& \text{ delayed fission spectrum for neutrons in group $g$,} \nonumber \\
+    I =& \text{ total number of \gls{DNP} groups,} \nonumber \\
+    \beta =& \text{ total delayed neutron fraction.} \nonumber
 \end{align}
@@
     %
-%    \shortintertext{where}
-%    \beta_i =& \text{ delayed neutron fraction of precursor group $i$,}
-%    \nonumber \\
-%    \lambda_i =& \text{ average decay constant of delayed neutron} \nonumber \\
-%    &\text{ precursors in precursor group $i$,} \nonumber \\
-%    C_i =& \text{ concentration of \gls{DNP}s in}
-%    \nonumber \\
-%    &\text{ precursor group $i$,} \nonumber \\
-%    \vec{u} =& \text{ molten salt flow velocity vector,}
-%    \nonumber \\
-%    D_{\text{P}} =& \text{ effective diffusion coefficient of the delayed}
-%    \nonumber \\
-%    &\text{ neutron precursors.} \nonumber
+    \shortintertext{where}
+    \beta_i =& \text{ delayed neutron fraction of precursor group $i$,}
+    \nonumber \\
+    \lambda_i =& \text{ average decay constant of delayed neutron} \nonumber \\
+    &\text{ precursors in precursor group $i$,} \nonumber \\
+    C_i =& \text{ concentration of \gls{DNP}s in}
+    \nonumber \\
+    &\text{ precursor group $i$,} \nonumber \\
+    \vec{u} =& \text{ molten salt flow velocity vector,}
+    \nonumber \\
+    D_{\text{P}} =& \text{ effective diffusion coefficient of the delayed}
+    \nonumber \\
+    &\text{ neutron precursors.} \nonumber
 \end{align}
@@
     \label{eq:divergence}
-%    \shortintertext{where}
-%    \rho =& \text{ fluid density,} \nonumber \\
-%    p =& \text{ pressure,} \nonumber \\
-%    \mu =& \text{ dynamic viscosity,} \nonumber \\
-%    \alpha =& \text{ coefficient of thermal expansion,} \nonumber \\
-%    \vec{g} =& \text{ gravitational force vector,} \nonumber
-%    \\
-%    T =& \text{ fluid temperature,} \nonumber \\
-%    T_{\text{ref}} =& \text{ reference temperature at which the nominal}
-%    \nonumber \\
-%    &\text{ density is provided.} \nonumber
-%    \nonumber
+    \shortintertext{where}
+    \rho =& \text{ fluid density,} \nonumber \\
+    p =& \text{ pressure,} \nonumber \\
+    \mu =& \text{ dynamic viscosity,} \nonumber \\
+    \alpha =& \text{ coefficient of thermal expansion,} \nonumber \\
+    \vec{g} =& \text{ gravitational force vector,} \nonumber
+    \\
+    T =& \text{ fluid temperature,} \nonumber \\
+    T_{\text{ref}} =& \text{ reference temperature at which the nominal}
+    \nonumber \\
+    &\text{ density is provided.} \nonumber
+    \nonumber
 \end{align}
@@
     \shortintertext{where}
-%    c_p =& \text{ specific heat capacity of molten salt,} \nonumber \\
-%    k =& \text{ effective thermal conductivity of molten salt,} \nonumber \\
-%    Q_f =& \text{ fission heat source,} \nonumber \\
-%    \epsilon_g =& \text{ average fission energy released by neutrons in group
-%    $g$,} \nonumber \\
+    c_p =& \text{ specific heat capacity of molten salt,} \nonumber \\
+    k =& \text{ effective thermal conductivity of molten salt,} \nonumber \\
+    Q_f =& \text{ fission heat source,} \nonumber \\
+    \epsilon_g =& \text{ average fission energy released by neutrons in group
+    $g$,} \nonumber \\
     Q_s =& \text{ heat sink/removal.} \nonumber
@@
 coupled to the active core simulation through Picard iterations in MOOSE's
-MultiApp functionality and inlet/outlet boundary values.
+MultiApp \cite{gaston_physics-based_2015} functionality and inlet/outlet boundary values.
 The external loop region is assumed to be subcritical to minimize neutron

### 2024-06-24 hybrid.tex
final edits for first draft of chapter 4; add image files for chapter 2,3,4

@@
 demonstrations due to superior numerical properties. I will make clear the reasons for this
-selection through 1-D test case results in Chapter 5.
+selection through 1-D test case results in Chapter \ref{chap:hybrid-verification}.
 
@@
 
-This boundary condition for the $S_N$ subsolver will generally cause deviations
+The $S_N$ subsolver boundary conditions will generally cause deviations
 in the flux and transport correction values since the $P_1$ approximation is accurate for up to
-linear anisotropies in the flux. However, as I will show in the next chapter, the influence of
-boundary conditions on the accuracy of
+linear anisotropies in the flux. However, as I will show in Chapter \ref{chap:hybrid-verification},
+the influence of boundary conditions on the accuracy of
 the transport correction terms does not extend far from the boundary $\partial V_1$ in optically
 thick media due to competing influences from the bulk scattering and absorption effects in $V_1$.
-Results in the next chapter will show that transport corrections generated with the $S_N$
-subsolver within $V_1$ are accurate everywhere up to a few centimeters near $\partial V_1$.
-
-When applying the hybrid method, $V_1$ must be large enough to provide accurate transport
-corrections near the highly-absorbing control rod while simultaneously accommodating for inaccurate
-transport corrections near $\partial V_1$. I will henceforth refer to the region in $V_1$
-near $\partial V_1$ with inaccurate transport corrections as the \textit{buffer region} ($V_1'$).
-We will discard inaccurate transport correction values in $V_1'$. Therefore, the subsequent neutron
+Results in Chapter \ref{chap:hybrid-verification} will show that transport correction values
+($D^s_g$ or $\vec{D}_g$) generated with the $S_N$ subsolver within $V_1$ are accurate everywhere up
+to a few centimeters away from $\partial V_1$.
+
+For accurate flux estimates with the hybrid method, I propose that $V_1$ must be large enough to
+provide accurate transport corrections near the highly-absorbing control rod while simultaneously
+accommodating for inaccurate transport correction values near $\partial V_1$. Right before each
+neutron diffusion calculation, the solver must determine where the inaccurate transport correction
+values occur and discard them. I will refer to the region near $\partial V_1$ in which the
+transport correction values are discarded as the \textit{buffer region} ($V_1'$).
+$V_1'$ shares the same outer boundary $\partial V_1$ as $V_1$. Consequently, the neutron
 diffusion calculation on $V_0$ will only apply transport corrections in $V_1$ excluding $V_1'$.
+
+In this discussion thus far, I have yet to establish the cutoff boundary for $V_1'$ within $V_1$
+opposite to $\partial V_1$. A natural/intuitive criterion for the location of the cutoff boundary
+would be wherever the components of the drift correction variable $\vec{D}_g$ is zero, i.e.,
+wherever the components change signs. The reasons for this criterion are twofold. Firstly, at
+points where the $\vec{D}_g$ components are zero, the flux is approximately isotropic along the
+axes corresponding to the components because value of $\vec{D}_g$ represents how much correction a
+standard neutron diffusion model requires. Secondly, this choice preserves the smoothness of the
+neutron flux and flux gradient.
+%
+\begin{figure}[htb!]
+	\centering
+	\includegraphics[width=.75\columnwidth]{case-3b-geometry}
+	\caption{1-D geometry for Case 3b.}
+	\label{fig:3b-geometry}
+\end{figure}
+
+Here, I use Case 3b of the 1-D test cases to be covered in Chapter \ref{chap:hybrid-verification}
+to illustrate the buffer region and the behavior of $\vec{D}_g$. Case 3b is a 1-D slab problem
+consisting of control rod, air, fuel-graphite repeating lattice, and core vessel regions as shown
+in Figure \ref{fig:3b-geometry}. The fuel regions are neutron-multiplying media containing
+$^\text{235}$U fissile material. I took material compositions of each region from the \gls{MSRE}
+reference design and the start-up fuel composition \cite{fratoni_molten_2020}.
+Case 3b has reflecting boundary conditions on the left boundary and vacuum boundary conditions on
+the right boundary.
+%
+\begin{figure}[htb!]
+    \centering
+    \begin{subfigure}[t]{.49\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{case-3b-group-1-drift}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{.49\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{case-3b-group-8-drift}
+    \end{subfigure}
+    \caption{The reference and hybrid drift ($\vec{D}_g$) distributions for group 1 and 8 calculated
+      from $S_8$ and $S_8$-diffusion simulations. The correction region $V_1$ spans $x=0$ cm to
+      $x=10$ cm.}
+    \label{fig:3b-drift}
+\end{figure}
+
+I ran an eight energy group $S_8$ $k$-eigenvalue simulation with the newly implemented $S_N$ solver
+on Moltres to generate reference $\vec{D}_g$ distributions. I also ran an eight-group hybrid
+$S_8$-diffusion simulation with the $S_8$ method for generating $\vec{D}_g$ in $V_1$. I set $V_1$
+to encompass the control rod region and its vicinity from $x=0$ to $10$ cm based on the geometry in
+Figure \ref{fig:3b-geometry}. Figure \ref{fig:3b-drift} shows the $\vec{D}_1$ and $\vec{D}_8$
+distributions from $x=0$ to $15$ from the reference $S_8$ and hybrid simulations. The hybrid
+$\vec{D}_g$ distribution is zero beyond $x=10$ cm because no corrections are generated outside of
+$V_1$. Comparing the hybrid distributions to the reference distributions in both energy groups, the
+hybrid distributions are accurate up to around $x=8$ cm. Working backward towards the control rod
+at $x=0$-$0.5$ cm, the $\vec{D}_g$ values go to zero at $x=7.5$ cm. Therefore, the buffer
+region $V_1'$ lies between $x=7.5$ cm and $x=10$ cm. While $V_1$ is pre-determined at the start of
+the simulation and fixed for all energy groups, $V_1'$ does not necessarily coincide for all
+energy groups and coordinate axes.
 
@@
 
-All source code for the hybrid $S_N$-diffusion model can be found in my Moltres GitHub fork
+All source code for the hybrid $S_N$-diffusion model is hosted on my Moltres GitHub fork
 repository\footnote{\url{https://github.com/smpark7/moltres/tree/sn-method}} as of the time of
-writing. This branch may be merged into the original upstream Moltres GitHub repository in future.
+writing. This branch may be merged into the original upstream Moltres GitHub repository in the near
+future.
 
@@
 of $G\times N_d$ angular flux variables $\Psi_{g,d}$. The Moltres $S_N$ model groups the
-$\Psi_{g,d}$ variables by energy group into $G$ \textit{array variables}. An array variable is a
+$\Psi_{g,d}$ variables by energy group into $G$ array variables. An array variable is a
 set of standard field variables to help with dealing with setting up simulations with high
@@
 residual contributions falls below a defined tolerance value. All $S_N$ simulations in this work
-employ the \gls{PJFNK} method with the HYPRE-BoomerAMG method for preconditioning. BoomerAMG is an
-algebraic multigrid method from the HYPRE \cite{hypre_hypre_2022} linear solver library.
-
-The drift correction terms $\vec{D}_g$ are initialized as auxiliary array variables
+employ the \gls{PJFNK} method \cite{knoll_jacobian-free_2004} with HYPRE-BoomerAMG
+\cite{hypre_hypre_2022} for preconditioning. BoomerAMG is a parallel implementation of the
+algebraic multigrid method from the HYPRE linear solver library.
+
+Moltres initializes the drift correction variables $\vec{D}_g$ as auxiliary array variables
 with three components corresponding to their $x$, $y$, and $z$ components. Auxiliary variables
@@
 calculation stage. The drift and boundary correction variables are computed during the
-post-calculation stage from $\Psi_{g,d}$.
+post-calculation stages using $\Psi_{g,d}$.
+I created new kernel classes, \texttt{GroupDrift} and \texttt{VacuumCorrectionBC}, to add the
+drift and boundary correction terms to the transport-corrected neutron diffusion solver.
 
@@
 transfers the correction variables to the neutron diffusion solver to compute its next flux
-estimate. This process continues until the neutron flux distribution converges to within tolerance
+estimate. This process iterates until the neutron flux distribution converges to within tolerance
 values provided in the input file.
@@
 
-In this chapter, I presented the mathematical derivations for the \gls{SAAF} formulation of the
-$S_N$ method, transport correction formulations, and boundary conditions relevant to the hybrid
-$S_N$-diffusion method. I also presented details on the iteration algorithm and the handling of
-transport corrections in the hybrid $S_N$-diffusion method to improve on the standard neutron
-diffusion method for neutronics simulations of reactor systems containing highly neutron-absorbing
-regions (control rods). The hybrid method relies on transport corrections derived from the flux
-solutions from the $S_N$ subsolver to fix inaccurate flux estimates from using the neutron
-diffusion method alone. The hybrid method saves on computational expense at the cost of some
-accuracy in reactor regions not treated with transport corrections. In the next chapter, I will
-also show that I can further reduce computational expense by leaving the flux in the $S_N$
-subsolver because the transport correction parameters converge faster than the flux in the $S_N$
-subsolver.
-
-%I established a framework for the hybrid method for coupling an $S_N$ subsolver and a neutron
-%diffusion solver through outer iterations by taking advantage of the weak dependence of the $S_N$
-%flux solution on the approximate boundary conditions determined from the neutron diffusion flux
-%solution. This framework could be extended to improve multischeme methods, which
-%divide a problem domain into subdomains where different neutronics methods are applied. For
-%instance, in an $S_N$-diffusion multischeme method, the $S_N$ method is applied in subdomains with
-%significant transport effects while other subdomains are treated with the neutron diffusion method
-%to reduce the overall computational effort required \cite{wang_rattlesnake_2021}. 
-%
-%I also developed an algorithm for identifying and discarding inaccurate \gls{SVDC} values, which
-%are expected to occur near the correction region boundary. I demonstrated the hybrid
-%method on 1-D graphite-moderated test cases modeled after the \gls{MSRE}. The hybrid method showed
-%good agreement with the reference Monte Carlo and $S_8$ methods after eliminating neutron energy
-%group discretization errors in test cases that included highly neutron-absorbing regions. The
-%hybrid method had better accuracy in more homogeneous systems largely due to the larger effective
-%transport correction regions through \glspl{SVDC}. Further work
-%is needed to extend and verify the hybrid method to 2-D and 3-D neutronics simulations and to
-%create a robust solution for determining \glspl{SVDC} around neutron flux peaks and troughs.
+The hybrid $S_N$-diffusion method improves on the standard neutron
+diffusion method by iteratively applying transport corrections generated from solving the $S_N$
+neutron transport
+method in subdomains containing highly neutron-absorbing regions (e.g., control rods). By limiting
+the size of the subdomain in which $S_N$ subsolver is active, the hybrid method saves on
+computational expense at the cost of some accuracy in the regions not treated with transport
+corrections. In turn, the neutron diffusion subsolver provides approximate boundary source terms
+for the $S_N$ subsolver boundary conditions.
+
+I presented two different formulations of transport correction that I investigated in this work for
+implementation in the hybrid method: diffusion correction and drift correction. Diffusion
+correction involves replacing the diffusion coefficient in the neutron diffusion equations with
+corrected coefficients derived from neutron transport methods. Drift correction involves
+introducing additional drift terms to the neutron diffusion equations to provide transport
+corrections. I found drift correction to be more promising due to issues with division by zero when
+obtaining diffusion corrections.
+
+In my investigations with the hybrid method, I found that the $S_N$ subsolver produces
+accurate transport correction values in most of the subdomain. However, the transport correction
+values near the $S_N$ subsolver boundary are inaccurate due to the approximate boundary conditions
+provided by the diffusion subsolver. Thus, I developed and presented a numerical fix for discarding
+inaccurate transport correction values and preserving the smoothness of the neutron flux and flux
+gradient solution.
+
+By implementing the hybrid method in Moltres, I leveraged existing neutronics modelling
+capabilities and parallel numerical solver libraries available through Moltres and \gls{MOOSE} to
+enable full-core, time-dependent reactor simulations on \gls{HPC} clusters.
+In the next chapter, I verify the hybrid method against Monte Carlo and $S_N$ neutron transport
+methods through various neutronics test cases and discuss some significant observations made in my
+investigations.

### 2024-06-24 appendix.tex

\section{Numerical Implementation} \label{sec:implementation}

In Section \ref{sec:theory}, I presented the theoretical basis for the hybrid
$S_N$-diffusion method and demonstrated the method on a simple problem. Now I will discuss
numerical implementation details of group
constant data processing and the neutron diffusion, $S_N$ neutron transport, and hybrid
$S_N$-diffusion solvers. I implemented all numerical solvers and analysis scripts in the Python
programming language. First, I will discuss the material group constant data generation and
postprocessing steps. Next, I present the implementation details of the neutron
diffusion and $S_N$ numerical methods. Lastly, I present how they are coupled to form the hybrid
$S_N$-diffusion method.

\subsection{Group Constant Data Generation}

The group constants required by either or both neutron diffusion and $S_N$ neutron transport
methods are:
%
\begin{itemize}
  \item $\Sigma_{t,g}$: Macroscopic total cross section in group $g$,
  \item $\Sigma_{r,g}$: Macroscopic removal cross section in group $g$,
  \item $\Sigma_s^{g'\rightarrow g}$: Macroscopic group-to-group scattering cross section matrix,
  \item $\Sigma_{s,l}^{g'\rightarrow g}$: $l$-th Legendre moment of the macroscopic
    group-to-group scattering cross section matrix,
  \item $\Sigma_{sp,l}^{g'\rightarrow g}$: $l$-th Legendre moment of the macroscopic
    group-to-group scattering production cross section matrix,
  \item $D_g$: $P_1$-based diffusion coefficient in group $g$,
  \item $\nu\Sigma_{f,g}$: Product of the average number of neutrons produced per fission and the
    macroscopic fission cross section in group $g$,
  \item $\chi_g$: Neutron fission spectrum in group $g$.
\end{itemize}
%
These group constants are generated using OpenMC's multigroup cross section generation capability
\cite{boyd_multigroup_2019} and postprocessed using a Python script into JSON format files.
$\Sigma_{r,g}$ is the only quantity that OpenMC does not directly provide. It is calculated as
%
\begin{align}
  \Sigma_{r,g} =& \sum^G_{g'\neq g}\Sigma_s^{g\rightarrow g'}+\Sigma_{a,g}-\left(\Sigma_{sp}^{g
    \rightarrow g} - \Sigma_s^{g\rightarrow g}\right)
  \shortintertext{where}
      \Sigma_{a,g} =& \mbox{ macroscopic absorption cross section in group $g$,} \nonumber \\
      \Sigma_{sp}^{g\rightarrow g} =& \mbox{ macroscopic scattering production cross section from
      group $g$ to $g$.} \nonumber
\end{align}
%
$\Sigma_{r,g}$ primarily represents the loss of neutrons from group $g$ through outscattering and
absorption. $\Sigma_{sp}^{g\rightarrow g}$ incorporates neutron multiplication effects from neutron
knockout reactions into the scattering cross section. Neutron knockout reactions are commonly
tallied as scattering reactions, and $\Sigma_{r,g}$ is a convenient term to
incorporate neutron knockout effects into the neutron diffusion equations. I ran all test cases for
$S_N$ and hybrid method calculations with up to the 2nd Legendre moments of the scattering
cross sections ($L=2$).

OpenMC uses the $P_1$ flux-limited formulation \cite{pomraning_flux-limited_1984} for calculating
$D_g$ as follows
%
\begin{align}
  D_g =& \frac{1}{3\Sigma_{tr,g}},
  \shortintertext{where}
  \Sigma_{tr,g} =& \frac{\langle\Sigma_{t,g}\phi_g\rangle-\langle\Sigma_{s1,g}\phi_g\rangle}
  {\langle\phi_g\rangle}, \nonumber \\
  \langle\Sigma_{t,g}\phi_g\rangle =& \int_{r\in V}dr \int_{4\pi}d\Omega\int^{E_{g-1}}_{E_g}dE\
  \Sigma_{t,g}(r,E)\Psi(r,E,\Omega), \nonumber \\
  \langle\Sigma_{s1,g}\phi_g\rangle =& \int_{r\in V}dr \int_{4\pi}d\Omega\int^{E_{g-1}}_{E_g}dE
  \int_{4\pi}d\Omega'\int^{\infty}_0dE'\int^1_{-1}d\mu\ \mu\Sigma_s(r,E'\rightarrow E,\Omega'\cdot
  \Omega)\phi(r,E',\Omega'), \nonumber \\
  \langle \phi \rangle =& \int_{r\in V}dr\int_{4\pi}d\Omega\int^{E_{g-1}}_{E_g}dE\ \Psi(r,E,\Omega)
  .\nonumber
\end{align}

\subsection{Neutron Diffusion Method}

On a 1-D uniform spatial grid with $I+1$ mesh points, the neutron scalar flux variables
$\phi_{g,i}$ are defined on the mesh points $x_i$. Theoretically, group constants are volumetric
material properties that should be defined on the cell-centered half-integer mesh points
$x_{i+\sfrac{1}{2}}$. In practice, all material properties except diffusion coefficients are
uniform in each subregion and sampled at $x_i$. To avoid ambiguity concerning diffusion
coefficient sampling, I formulated all test cases such that all material interfaces fall on $x_i$.

Discretizing the multigroup $k$-eigenvalue neutron diffusion equations in Eq.\ \ref{eq:1d-diff}
and reformulating the scattering term using neutron balance in the control volume bounded by
$x_{i-\sfrac{1}{2}}$ and $x_{i+\sfrac{1}{2}}$ yields
%
\begin{align}
  J_{g,i+\sfrac{1}{2}} - J_{g,i-\sfrac{1}{2}} + \Sigma_{t,g,i} \phi_{g,i} \Delta x = \sum^G_{g'=1}\left[
  \Sigma_{s,i}^{g'\rightarrow g}\phi_{g',i} + \chi_{g,i}\frac{\nu\Sigma_{f,g',i}}{k} \phi_{g',i}
\right]\Delta x. \label{eq:diff-j}
\end{align}
%
Using the diamond difference scheme to replace the $J$ terms with the discretized form of
Fick's first law of diffusion,
%
\begin{align}
  J_{g,i+\sfrac{1}{2}} = -D_{g,i+\sfrac{1}{2}}\frac{d\phi_{g,i+\sfrac{1}{2}}}{dx} =
  -D_{g,i+\sfrac{1}{2}} \frac{\phi_{g,i+1}-\phi_{g,i}}{\Delta x},
\end{align}
%
and rearranging the terms in Eq.\ \ref{eq:diff-j} yields
%
\begin{align}
  -\frac{D_{g,i-\sfrac{1}{2}}}{\Delta x} \phi_{g,i-1} + &\left[\frac{D_{g,i-\sfrac{1}{2}}+
  D_{g,i+\sfrac{1}{2}}}{\Delta x} + \Delta x\ \Sigma_{r,g,i} \right]\phi_{g,i} -
  \frac{D_{g,i+\sfrac{1}{2}}}{\Delta x}\phi_{g,i+1} -\Delta x\sum^G_{g'\neq g}
  \Sigma_{s,i}^{g'\rightarrow g}\phi_{g',i} \nonumber \\
  =& \Delta x\sum^G_{g'=1}
  \chi_{g,i} \frac{\nu\Sigma_{f,g',i}}{k} \phi_{g',i}, \label{eq:diff-fd}
  \shortintertext{where}
  \Sigma_{r,g} =& \Sigma_{t,g} - \Sigma_s^{g\rightarrow g} \nonumber \\
  =& \mbox{ macroscopic removal cross section for neutron group }g. \nonumber
\end{align}
%
The diamond difference scheme is 2nd-order accurate, and this form is
equivalent to applying 2nd-order finite differencing to the original neutron diffusion equation in
Eq.\ \ref{eq:1d-diff} with diamond differencing for the cell-centered group constants.
The fixed neutron source $S_g$ is ignored here since the test cases are all neutron-multiplying
systems with no fixed source.

I implemented two types of boundary conditions: vacuum and reflective boundary conditions. The
\textbf{vacuum boundary conditions} are imposed by setting the incoming flux in the $P_1$
approximation to zero and applying 2nd-order finite differencing as follows:
%
\begin{align}
  \mbox{Left boundary: } \frac{\phi_{g,0}}{4}-\frac{D_{g,\sfrac{1}{2}}}{2}
  \frac{\left(-\phi_{g,2}+4\phi_{g,1}-3\phi_{g,0}\right)}{2\Delta x} =& 0 \\
  \mbox{Right boundary: } \frac{\phi_{g,I}}{4}+\frac{D_{g,I-\sfrac{1}{2}}}{2}
  \frac{\left(\phi_{g,I-2}-4\phi_{g,I-1}+3\phi_{g,I}\right)}{2\Delta x} =& 0.
\end{align}
%
The \textbf{reflective boundary conditions} are imposed by setting the flux gradient to zero and
applying 2nd-order finite differencing as follows:
%
\begin{align}
  \mbox{Left boundary: } \frac{-3\phi_{g,0}+4\phi_{g,1}-\phi_{g,2}}{2\Delta x} =& 0 \\
  \mbox{Right boundary: } \frac{3\phi_{g,I}-4\phi_{g,I-1}+\phi_{g,I-2}}{2\Delta x} =& 0
\end{align}
%
At material interfaces, the continuity condition requires that the net neutron current on either
side of the interface be equal as follows:
%
\begin{align}
  -\frac{3D_{g,i-\sfrac{1}{2}} - D_{g,i-\sfrac{3}{2}} }{2}
  \frac{\phi_{g,i-2}-4\phi_{g,i-1}+3\phi_{g,i}}{2\Delta x} =&
  -\frac{3D_{g,i+\sfrac{1}{2}} - D_{g,i+\sfrac{3}{2}} }{2}
  \frac{-3\phi_{g,i}+4\phi_{g,i+1}-\phi_{g,i}}{2\Delta x} \label{eq:itf-bc}
\end{align}
%
for a material interface at $x_i$.

Altogether, they form a system of equations of the form $\bm{A\overline{\phi}}=\bm{\frac{1}{k}
B\overline{\phi}}$, where $\bm{\overline{\phi}}$ is a flattened vector representation of
$\phi_{g,i}$, and $\bm{A}$ and $\bm{B}$ are matrices of the coefficients of $\phi_{g,i}$ as given
by Eqs. \ref{eq:diff-fd} to \ref{eq:itf-bc}. I implemented the inverse power method to find $k$ and
$\overline{\phi}$. The inverse power method algorithm is as follows
%
\begin{align}
  \shortintertext{1. Initialize $k^0$ and $\bm{\overline{\phi}}^0$}
  \shortintertext{2. Update $\bm{\overline{\phi}}$ and $k$}
  \bm{\overline{\phi}}^m =& \frac{1}{k^{m-1}}\bm{A}^{-1}\bm{B\overline{\phi}}^{m-1} \\
  k^m =& k^{m-1}\frac{|\bm{B\overline{\phi}}^m|}{|\bm{B\overline{\phi}}^{m-1}|}
  \shortintertext{3. Check whether convergence is reached}
  \epsilon_\phi =
  \frac{|\bm{\overline{\phi}}^m-\bm{\overline{\phi}}^{m-1}|}{|\bm{\overline{\phi}}^m|} <& \
  tol_{\bm{\overline{\phi}}} \\
  \epsilon_k =
  \frac{|k^m-k^{m-1}|}{|k^m|} <& \ tol_k
  \shortintertext{4. Return to Step 2 if either expression is false, otherwise exit.} \nonumber
\end{align}
%
$k^m$ and $\overline{\phi}^m$ denote estimates of $k$ and $\overline{\phi}$ after the $m$-th
iteration. Matrix $\bm{A}$ is a primarily tridiagonal matrix with at most $G-1$ off-diagonal terms
from the fourth term in Eq.\ \ref{eq:diff-fd}. Thus, $\bm{A}$ is initialized as a sparse matrix to
take advantage of the computationally efficient sparse matrix solver functions from the
\texttt{sparse} class of the \texttt{SciPy} \cite{virtanen_scipy_2020} Python library for
scientific and technical computing. Matrix $\bm{B}$
is never initialized explicitly as a matrix. Instead, the vector $\bm{b}=\bm{B\overline{\phi}}$ is
updated directly in every iteration. In this system of equations, $|\bm{b}|$ corresponds to the
total number of fission neutrons produced in the system for a given $\bm{\overline{\phi}}$. This
quantity is calculated using the \texttt{trapezoid} numerical integration function from
\texttt{SciPy} to estimate the integral value of the source term, $\nu\Sigma_{g,f}\phi_{g}$, in $x$
from the discrete flux values in $\bm{x}$. Finally, the final $\phi_{g,i}$ is normalized by a
factor of $|\bm{b}|/k$ (no. of source neutrons) to obtain the neutron scalar flux per source
neutron.

\subsection{$S_N$ Neutron Transport Method}

For the 1-D $S_N$ neutron transport method on the same uniform spatial grid with $I+1$ mesh points,
the neutron angular flux variables $\Psi_{g,i,n}=\Psi_g(x_i,\mu_n)$ are defined on the mesh points
$x_i$ while neutron scalar flux variables $\phi_{g,i\pm\sfrac{1}{2}}=\phi_g(x_{i\pm\sfrac{1}{2}})$
are defined on the half-integer mesh points $x_{i\pm\sfrac{1}{2}}$. All group constants are sampled
at $x_{i\pm\sfrac{1}{2}}$.

The $S_N$ equations are solved using the transport sweep method in which the algorithm ``sweeps''
through the spatial grid and sequentially updates $\Psi_{g,i,n}$. The algorithm sweep
direction follows the direction of neutron travel, i.e. it sweeps in the positive direction for
$\Psi_{g,i,n}$ with $\mu_n>0$ and vice versa. Doing otherwise is unphysical and causes numerical
instabilities.

Discretizing the multigroup $S_N$ neutron transport equations in Eq.\ \ref{eq:1d-sn} about
$x_{i+\sfrac{1}{2}}$ yields
%
\begin{align}
  \mu_n\frac{\Psi_{g,i+1,n}-\Psi_{g,i,n}}{x_{i+1}-x_i} + \Sigma_{t,g,i+\sfrac{1}{2}}
  \Psi_{g,i+\sfrac{1}{2},n} = q_{g,i+\sfrac{1}{2},n}
\end{align}
%
where $q_{g,i+\sfrac{1}{2},n}$ represents the combined scattering and fission neutron source term.
After expressing $\Psi_{g,i+\sfrac{1}{2}}$ as the average of $\Psi_{g,i+1,n}$ and $\Psi_{g,i,n}$
in the diamond difference scheme and rearranging the terms, we obtain
%
\begin{align}
  \Psi_{g,i+1,n} =& \frac{1-\Sigma_{t,g}\Delta x/2\mu_n}{1+
    \Sigma_{t,g}\Delta x/2\mu_n}\Psi_{g,i,n} +
    q\frac{\Delta x}{\mu_n\left(1+\Sigma_{t,g}\Delta x/2\mu_n\right)} \label{eq:sweep-right} &&
    (\mbox{for } \mu_n > 0)
\shortintertext{and}
  \Psi_{g,i,n} =& \frac{1+\Sigma_{t,g}\Delta x/2\mu_n}{1-
    \Sigma_{t,g}\Delta x/2\mu_n}\Psi_{g,i+1,n} -
    q\frac{\Delta x}{\mu_n\left(1-\Sigma_{t,g}\Delta x/2\mu_n\right)}. \label{eq:sweep-left} &&
    (\mbox{for } \mu_n < 0)
\end{align}
%
The remaining $i+\sfrac{1}{2}$ indexes on the group constants are dropped to reduce visual clutter.
These expressions are used to update $\Psi_{g,i,n}$ in the forward and backward transport sweeps.

The discretized scattering and fission terms in $q_{g,i+\sfrac{1}{2},n}$ are given as
%
\begin{align}
  q_{g,i+\sfrac{1}{2},n} =& \sum^G_{g'=1}\sum^L_{l=0}\frac{\left(2l+1\right)}
  {2}\Sigma_{s,l}^{g'\rightarrow g}P_l(\mu_n)\phi_{l,g',i+\sfrac{1}{2}} \nonumber \\
  &+\frac{\chi_g}{2}\sum^G_{g'=1}\frac{\nu\Sigma_{f,g'}}{k}\phi_{0,g',i+\sfrac{1}{2}}
  \label{eq:sn-q}
  \shortintertext{where}
  \phi_{l,g',i+\sfrac{1}{2}} =& \sum^N_{n'=1}w_{n'}P_l(\mu_{n'})\frac{\Psi_{g',i,n'}+
  \Psi_{g',i+1,n'}}{2}. \label{eq:phi-l}
\end{align}
%
$\phi_{l,g,i}$ are $l$-th Legendre expansions of the neutron flux evaluated using Gauss-Legendre
quadrature over $\mu_{n'}=[-1,1]$. $\phi_{0,g,i}$ and $\phi_{1,g,i}$ also correspond to the neutron
scalar flux $\phi_{g,i}$ and net current $J_{g,i}$, respectively.

For vacuum boundary conditions, $\Psi_{g,0,n}$ is zero for all positive $\mu_n$ while
$\Psi_{g,I,n}$ is zero for all negative $\mu_n$ as follows
%
\begin{align}
  \mbox{Left boundary: } \Psi_{g,0,n} =& 0 && (\mbox{for } \mu_n > 0) \\
  \mbox{Right boundary: } \Psi_{g,I,n} =& 0 && (\mbox{for } \mu_n < 0)
\end{align}
%
Reflective boundary conditions are imposed by equating the incoming angular flux to the
outgoing angular flux in the opposite direction as follows
%
\begin{align}
  \mbox{Left boundary: } \Psi_{g,0,n} =& \Psi_{g,0,n'} && (\mbox{for } \mu_n > 0, \mu_n =
  -\mu_{n'}) \\
  \mbox{Right boundary: } \Psi_{g,I,n} =& \Psi_{g,I,n'} && (\mbox{for } \mu_n < 0, \mu_n =
  -\mu_{n'})
\end{align}

The power iteration algorithm for the $S_N$ method is similar to the inverse power method algorithm
applied in the neutron diffusion method. The transport sweep step replaces the matrix-solving step
for updating $\overline{\phi}$.
%
\begin{align}
  \shortintertext{1. Initialize $k^0$, $\phi_{l,g,i+\sfrac{1}{2}}^0$, and $q^0$}
  \shortintertext{2. Apply transport sweeps to solve for $\Psi^m$ using Eqs. \ref{eq:sweep-right},
  \ref{eq:sweep-left}, and \ref{eq:sn-q}}
  \shortintertext{3. Update $\phi^m$ and $k^m$ using Eqs. \ref{eq:phi-l} and \ref{eq:sn-k}}
  k^m =& k^{m-1}\frac{\sum^I_{i=0}\sum^G_{g=1}\nu\Sigma_{f,g}\phi^m_{g,i+\sfrac{1}{2}}}
  {\sum^I_{i=0}\sum^G_{g=1}\nu\Sigma_{f,g}\phi^{m-1}_{g,i+\sfrac{1}{2}}} \label{eq:sn-k}
  \shortintertext{4. Check whether convergence is reached}
  \epsilon_\phi =
  \frac{|\bm{\overline{\phi}}^m-\bm{\overline{\phi}}^{m-1}|}{|\bm{\overline{\phi}}^m|} <& \
  tol_{\bm{\overline{\phi}}} \\
  \epsilon_k =
  \frac{|k^m-k^{m-1}|}{|k^m|} <& \ tol_k
  \shortintertext{5. Return to Step 2 if either expression is false, otherwise exit.} \nonumber
\end{align}

The transport sweep and $\phi^m$-update algorithms are parallelized using the \texttt{joblib}
parallel computing Python library \cite{noauthor_joblib_nodate} across all available CPU threads.
The total computational work is subdivided into several smaller tasks classified by unique pairs
of $g$ and $n$ values; each thread computes all $\Psi^m$ or $\phi^m$ values on the mesh for a given
pair of indexes $g$ and $n$.

\subsection{Hybrid $S_N$-Diffusion Method}

Without loss of generality, consider a 1-D system symmetric about the left boundary at $x_0=0$ cm.
In this case, as with Case 0, the control rod region is the left-most region, followed by other
regions. The $S_N$ subproblem domain $V_1$, containing the control rod and some region beyond it,
is bounded by $x_0$ and $x_j$ for some $x_j$
located several mean free paths to the right of the control rod region as governed by the relevant
discussion in Section \ref{sec:hybrid-method}.

Excluding the initial neutron diffusion calculation to initialize $k^0$ and $\phi^0$, each outer
iteration in the hybrid method consists of one $S_N$ neutron transport calculation and one neutron
diffusion calculation. The $\phi$ estimates from these $S_N$ transport and neutron diffusion
calculations are labeled as $\phi^{m+\sfrac{1}{2}}$ and $\phi^{m+1}$ during the
$(m+1)$-th outer iteration. $\phi^{m+\sfrac{1}{2}}$ spans $V_1$ while $\phi^{m+1}$ spans
the entire domain $V_0$.

For the $S_N$ subproblem boundary conditions, we discretize Eqs. \ref{eq:p1-j} and
\ref{eq:sn-psi-j} as follows
%
\begin{align}
  \Psi_{g,j,n} =& \frac{J_{g,-}(x_j)}{\sum^{N/2}_{n'=1}w_{n'}\mu_{n'}} \nonumber \\
  =& \frac{\frac{\phi_g(x_j)}{4}+
  \frac{D_g(x_{j-\sfrac{1}{2}})}{2}\frac{d\phi_g(x_j)}{dx}}{\sum^{N/2}_{n'=1}w_{n'}\mu_{n'}}
  \nonumber \\
  =& \frac{\frac{\phi_{g,j}}{4}+
  \frac{D_{g,j-\sfrac{1}{2}}}{2}\frac{\phi_{g,j}-\phi_{g,j-1}}{\Delta x}}
    {\sum^{N/2}_{n'=1}w_{n'}\mu_{n'}} && (\mu_n < 0) \label{eq:sn-bc}
\end{align}
%
where $D_g$ is the $P_1$-based diffusion coefficient value.

The \gls{SVDC} formulation in Eq.\ \ref{eq:svdc} is discretized using the diamond difference scheme
as follows
%
\begin{align}
  D^s_{g,i+\sfrac{1}{2}} =& -\frac{J^{tr}_{g,i+1}+J^{tr}_{g,i}}{2} \left(
  \frac{\phi^{tr}_{g,i+1}-\phi^{tr}_{g,i}}{\Delta x} \right)^{-1} = -\frac{\Delta x}{2}
  \frac{J^{tr}_{g,i+1}+J^{tr}_{g,i}}{\phi^{tr}_{g,i+1}-\phi^{tr}_{g,i}}. \label{eq:svdc-num}
\end{align}

The hybrid $S_N$-diffusion algorithm is as follows
%
\begin{align}
  \shortintertext{1. Initialize $k^0$ and $\phi^0$ with an initial neutron diffusion calculation on
  $V_0$}
  \shortintertext{2. Generate estimates for $\Psi_{g,j,n}$ at $x_j$ for the $S_N$ transport
  calculation boundary conditions using Eq.\ \ref{eq:sn-bc}}
  \shortintertext{3. Calculate $\phi^{m+\sfrac{1}{2}}$ with the $S_N$ transport sub-solver on
  $V_1$}
  \shortintertext{4. Generate \glspl{SVDC} with $\phi^{m+\sfrac{1}{2}}$ and $J^{m+\sfrac{1}{2}}$
  using Eq.\ \ref{eq:svdc-num}}
  \shortintertext{5. Calculate $\phi^{m+1}$ with the newly generated \glspl{SVDC} and the neutron
  diffusion solver on $V_0$}
  \shortintertext{6. Check whether convergence is reached}
  \frac{|\bm{\overline{\phi}}^m-\bm{\overline{\phi}}^{m-1}|}{|\bm{\overline{\phi}}^m|} <& \
  tol_{\bm{\overline{\phi}}} \\
  \frac{|k^m-k^{m-1}|}{|k^m|} <& \ tol_k
  \shortintertext{7. Return to Step 2 if either expression is false, otherwise exit.} \nonumber
\end{align}
%
Generally, the convergence tolerance values for the outer hybrid method iteration must be smaller
than the tolerance values for the neutron diffusion and $S_N$ transport inner iterations. Using
$\phi^m$ from the neutron diffusion calculation as initial conditions for the $S_N$ transport
calculation in the $(m+1)$-th iteration helps to reduce the number of transport
sweeps required significantly.

The $S_N$ sub-solver only covers $V_1$, so it does not calculate an independent $k$
estimate. Instead, the $k^m$ from the neutron diffusion calculation scales the fission neutron
source term in the $S_N$ sub-solver.

### 2024-06-24 benchmark.tex

\section{CNRS Benchmark}

As identified in Section \ref{sec:msfr-critique},
Moltres will benefit from further \gls{VV} of its
\gls{MSR} modeling capabilities. This chapter presents verification
results from Moltres for problems within the CNRS Benchmark
\cite{tiberga_results_2020}---a numerical benchmark specifically designed to
assess \gls{MSR} simulation tools on coupled multiphysics simulations of
fast-spectrum \glspl{MSR}. The benchmark consists of several steps in which
steady-state and transient simulations are prescribed. The benchmark starts with simple
single-physics cases and incrementally introduces various types of
multiphysics coupling until full coupling is simulated in the final steps. This
gradual approach helps participants identify sources of discrepancies
that can arise from differences in modeling assumptions or
cross-section libraries, or actual errors in the software. I adapted the
contents of this chapter from my journal article titled
``\textit{Verification of Moltres for Multiphysics Simulations of Fast-Spectrum
Molten Salt Reactors}'' and published in the Annals of Nuclear Energy journal
\cite{park_verification_2022}.

\input{cnrs-benchmark/benchmark}
\input{cnrs-benchmark/moltres}
\input{cnrs-benchmark/results}
\input{cnrs-benchmark/conclusion}

### 2024-06-24 cnrs-benchmark/benchmark.tex

\subsection{Description of the CNRS Benchmark} \label{sec:benchmark}

The CNRS Benchmark \cite{tiberga_results_2020} is a numerical
benchmark for multiphysics software dedicated to modeling \glspl{MSR}. It
consists of three phases and eight steps in total. Each
step is a well-defined subproblem for systematically assessing the
capabilities of \gls{MSR} software and pinpointing sources of discrepancies
between software. Phase 0 consists of three single-physics problems in fluid
dynamics, neutronics, and temperature. Phase 1 consists
of four coupled steady-state problems. Lastly, Phase 2 consists of one
coupled, time-dependent problem.

\begin{figure}[htb!]
	\centering
	\includegraphics[width=.6\columnwidth]{cnrs-geometry}
	\caption{2m$\times$2m 2D domain of the CNRS Benchmark. $U_{lid}$
	represents the velocity along the top boundary. For comparison, various quantities are
	measured along the centerlines AA' and BB'. From Tiberga et
	al. \cite{tiberga_results_2020}.}
	\label{fig:cnrs-geometry}
\end{figure}

As shown in Figure \ref{fig:cnrs-geometry}, the domain geometry is a
2m$\times$2m square cavity filled with LiF-BeF$_2$-UF$_4$ molten salt at an
initial temperature of 900K \cite{tiberga_results_2020}.
Standard vacuum boundary conditions apply for neutron flux along all
boundaries whereby outgoing neutrons are considered lost, while homogeneous
boundary conditions apply for delayed neutron precursors. No-slip boundary
conditions apply for velocity variables in the cavity, except along the top
boundary for Steps 0.1, 0.3, 1.1, 1.2, and 1.4, which impose forced flow in the
form of lid-driven
cavity flow. For the temperature variable, all boundaries are insulated, and we
simulate salt cooling with the following volumetric heat sink equation:
%
\begin{align}
    q'''(\vec{r}) &= \gamma \left(900 - T(\vec{r})\right) \label{eq:cnrs-heat}
    \shortintertext{where}
    q''' &= \mbox{volumetric heat sink [W$\cdot$m$^{-3}$],}
    \nonumber \\
    \gamma &= \mbox{heat transfer coefficient [W$\cdot$m$^{-3}\cdot$K$^{-1}$],}
    \nonumber \\
    T(\vec{r}) &= \mbox{temperature at point $\vec{r}$ [K].} \nonumber
\end{align}

Tiberga et al. \cite{tiberga_results_2020} used Serpent 2
\cite{leppanen_serpent_2014} with the JEFF-3.1 library
\cite{koning_jeff-31_2006} to generate multigroup neutronics data for the
LiF-BeF$_2$-UF$_4$ salt in the domain at 900K, which they condensed into six
energy groups and eight precursor groups. We direct readers to their paper for
the group constant data \cite{tiberga_results_2020}. In addition, the
benchmark prescribes the following equations to govern the temperature
dependence in the cross sections and the neutron diffusion coefficients:
%
\begin{align}
    \Sigma_i (T) &= \Sigma_i(T_{ref})
    \frac{\rho_{fuel}(T)}{\rho_{fuel}(T_{ref})}
    \shortintertext{and}
    D (T) &= D(T_{ref})
    \frac{\rho_{fuel}(T_{ref})}{\rho_{fuel}(T)}
    \shortintertext{where}
    \Sigma_i &= \mbox{relevant macroscopic cross section [cm${-1}$],}
    \nonumber \\
    D &= \mbox{neutron diffusion coefficient [cm$^2\cdot$s$^{-1}$],}   
    \nonumber \\
    \rho_{fuel} &= \mbox{density of the fuel salt [kg$\cdot$m$^{-3}$],}
    \nonumber \\
    T_{ref} &= \mbox{reference temperature} = 900\mbox{ K}. \nonumber
\end{align}

The benchmark also prescribes incompressible Navier-Stokes flow with the
Boussinesq approximation for evaluating the salt flow in the
domain but does not restrict the type of neutronics model.
Table \ref{table:benchmark} lists the relevant input parameters and observables.

\begin{table*}[tp!]
	\caption{Input parameters and observables of each benchmark step.}
	\centering
	\footnotesize
    \begin{tabular}{p{.05\textwidth} p{.1\textwidth} p{.3\textwidth} p{.45\textwidth}}
		\toprule
        \textbf{Step} & \textbf{Name} & \textbf{Input parameters} & \textbf{Observables} \\
		\midrule
        0.1 & Velocity field &
		\begin{itemize}[nosep,noitemsep,left=0pt,
		                before={\begin{minipage}[t]{\hsize}},
                        after ={\end{minipage}}]
		    \item $U_{lid} = 0.5$ m$\cdot$s$^{-1}$
		\end{itemize}\vspace*{-\baselineskip}\mbox{} &
		\begin{itemize}[nosep,noitemsep,left=0pt,
		                before={\begin{minipage}[t]{\hsize}},
                        after ={\end{minipage}}]
		    \item Velocity components $(u_x,u_y)$ along AA' and BB'
		\end{itemize}\vspace*{-\baselineskip}\mbox{} \\
        \midrule
        0.2 & Neutronics &
        \begin{itemize}[nosep,noitemsep,left=0pt,
		                before={\begin{minipage}[t]{\hsize}},
                        after ={\end{minipage}}]
		    \item $U_{lid} = 0$ m$\cdot$s$^{-1}$
		    \item $T = 900$ K
		    \item $P = 1$ GW
		\end{itemize} &
		\begin{itemize}[nosep,noitemsep,left=0pt,
		                before={\begin{minipage}[t]{\hsize}},
                        after ={\end{minipage}}]
		    \item Fission rate density $\sum^6_g \Sigma_{f,g} \phi_g(\vec{r})$ along AA'
            \item Reactivity $\rho$
		\end{itemize}\vspace*{-\baselineskip}\mbox{} \\
        \midrule
        0.3 & Temperature &
        \begin{itemize}[nosep,noitemsep,left=0pt,
		                before={\begin{minipage}[t]{\hsize}},
                        after ={\end{minipage}}]
		    \item Fixed flow field from Step 0.1 for
		    $U_{lid} = 0.5$ m$\cdot$s$^{-1}$
		    \item Fixed heat source distribution
		    $\sum^6_{g} \epsilon_g \Sigma_{f,g} \phi_g(\vec{r})$ from Step 0.2
		    \item $\gamma = 10^6$ W$\cdot$m$^{-3}\cdot$K$^{-1}$
		\end{itemize} &
		\begin{itemize}[nosep,noitemsep,left=0pt,
		                before={\begin{minipage}[t]{\hsize}},
                        after ={\end{minipage}}]
		    \item Temperature $T$ along AA' and BB'
		\end{itemize}\vspace*{-\baselineskip}\mbox{} \\
        \midrule
        1.1 & Circulating fuel &
        \begin{itemize}[nosep,noitemsep,left=0pt,
		                before={\begin{minipage}[t]{\hsize}},
                        after ={\end{minipage}}]
		    \item Fixed flow field from Step 0.1 for
		    $U_{lid} = 0.5$ m$\cdot$s$^{-1}$
		    \item $T = 900$ K
		    \item $P = 1$ GW
		\end{itemize} &
		\begin{itemize}[nosep,noitemsep,left=0pt,
		                before={\begin{minipage}[t]{\hsize}},
                        after ={\end{minipage}}]
		    \item Delayed neutron source $\sum^8_i \lambda_i C_i$ along AA' and BB'
		    \item Reactivity change between Step 1.1 and Step 0.2,
		    $\Delta \rho = \rho - \rho_{s_{0.2}}$
		\end{itemize}\vspace*{-\baselineskip}\mbox{} \\
        \midrule
        1.2 & Power coupling &
        \begin{itemize}[nosep,noitemsep,left=0pt,
		                before={\begin{minipage}[t]{\hsize}},
                        after ={\end{minipage}}]
		    \item Fixed flow field from Step 0.1 for
		    $U_{lid} = 0.5$ m$\cdot$s$^{-1}$
		    \item $P = 1$ GW
		    \item $\gamma = 10^6$ W$\cdot$m$^{-3}\cdot$K$^{-1}$
		\end{itemize}\vspace*{-\baselineskip}\mbox{} &
		\begin{itemize}[nosep,noitemsep,left=0pt,
		                before={\begin{minipage}[t]{\hsize}},
                        after ={\end{minipage}}]
		    \item Temperature $T$ along AA' and BB'
            \item Reactivity change between Step 1.2 and Step 1.1,
            $\Delta\rho = \rho - \rho_{s_{1.1}}$
            \item Change in fission rate density
            $\sum^6_g \Sigma_{f,g} \phi_g(\vec{r}) -
            \left[\sum^6_g \Sigma_{f,g} \phi_g(\vec{r})\right]_{s_{0.2}}$
		\end{itemize} \\
        \midrule
        1.3 & Buoyancy &
        \begin{itemize}[nosep,noitemsep,left=0pt,
		                before={\begin{minipage}[t]{\hsize}},
                        after ={\end{minipage}}]
		    \item $P = 1$ GW
		    \item $U_{lid} = 0$ m$\cdot$s$^{-1}$
		    \item $\gamma = 10^6$ W$\cdot$m$^{-3}\cdot$K$^{-1}$
		\end{itemize}\vspace*{-\baselineskip}\mbox{} &
		\begin{itemize}[nosep,noitemsep,left=0pt,
		                before={\begin{minipage}[t]{\hsize}},
                        after ={\end{minipage}}]
		    \item Velocity components $(u_x, u_y)$ along AA' and BB'
            \item Temperature $T$ along AA' and BB'
            \item Delayed neutron source $\sum^8_i \lambda_i C_i$ along AA' and BB'
            \item Reactivity change from Step 0.2
        $\Delta\rho = \rho - \rho_{s_{0.2}}$
		\end{itemize} \\
        \midrule
        1.4 & Full coupling &
        \begin{itemize}[nosep,noitemsep,left=0pt,
		                before={\begin{minipage}[t]{\hsize}},
                        after ={\end{minipage}}]
		    \item $\gamma = 10^6$ W$\cdot$m$^{-3}\cdot$K$^{-1}$
		    \item $P$ variable in the range $[0,1]$ GW with a step of 0.2 GW
		    \item $U_{lid}$ variable in the range $[0,0.5]$ m$\cdot$s$^{-1}$
		    with a step of 0.1 m$\cdot$s$^{-1}$
		\end{itemize} &
		\begin{itemize}[nosep,noitemsep,left=0pt,
		                before={\begin{minipage}[t]{\hsize}},
                        after ={\end{minipage}}]
		    \item Reactivity change between Step 1.4 and Step 0.2,
		    $\Delta\rho = \rho - \rho_{s_{0.2}}$, for all permutations of $P$
		    and $U_{lid}$ values
		\end{itemize}\vspace*{-\baselineskip}\mbox{} \\
        \midrule
        2.1 & Forced convection transient &
        \begin{itemize}[nosep,noitemsep,left=0pt,
		                before={\begin{minipage}[t]{\hsize}},
                        after ={\end{minipage}}]
		    \item $\gamma = 10^6$ W$\cdot$m$^{-3}\cdot$K$^{-1}$
            \item Steady-state solution from Step 1.4 for $U_{lid} = 0.5$
        m$\cdot$s$^{-1}$ and $P = 1.0$ GW
		\end{itemize} &
		\begin{itemize}[nosep,noitemsep,left=0pt,
		                before={\begin{minipage}[t]{\hsize}},
                        after ={\end{minipage}}]
		    \item Power gain and shift as a function of the perturbation frequency
		\end{itemize}\vspace*{-\baselineskip}\mbox{} \\
		\bottomrule
	\end{tabular}
	\label{table:benchmark}
\end{table*}

Step 2.1 studies the transient response of the fully coupled nonlinear system.
Linear perturbation analyses are performed by introducing periodic
perturbations to the heat transfer coefficient $\gamma$ and studying the gain
and phase shift of the response in the total power $P$. For the initial
conditions, the steady-state solution from Step 1.4 with
$U_{lid} = 0.5$ m$\cdot$s$^{-1}$ and $P = 1$ GW is used. This initial
configuration is made exactly critical by scaling the neutron source terms,
from fission and \gls{DNP} decay, by the inverse of the criticality eigenvalue
solution from Step 1.4.

$\gamma$ is uniformly perturbed according to small-amplitude sine waves given
as:
%
\begin{align}
    \gamma =& \gamma_0 \left[ 1 + 0.1\sin\left(2 \pi f \right) \right]
    \shortintertext{where}
    \gamma_0 =& 10^6 \mbox{ W$\cdot$m$^{-3}\cdot$K$^{-1}$}, \nonumber \\
    f \in& \left\lbrace 0.0125, 0.025, 0.05, 0.1, 0.2, 0.4, 0.8 \right\rbrace 
    \mbox{ Hz.} \nonumber
\end{align}

The benchmark defines power gain as:
%
\begin{align}
    \mbox{Power gain} =& \frac{\left(P_{max} - P_{avg}\right)/P_{avg}}{
    \left(\gamma_{max} - \gamma_{avg}\right)/\gamma_{avg}}
\end{align}
%
The subscripts denote the maximum and time-averaged values of $P$ and $\gamma$.

\FloatBarrier

### 2024-06-24 cnrs-benchmark/results.tex

\subsection{Results}

In this section, we compare the results from Moltres for each CNRS Benchmark
step to the results in the benchmark paper \cite{tiberga_results_2020}.
The software packages from \gls{CNRS} and \gls{TUD}
each report two sets of results from different angular discretizations
in their neutronics models for Steps 0.2, 1.1, 1.2, 1.3, 1.4, and 2.1. These
sets of results are labeled as CNRS-$SP_1$ and
CNRS-$SP_3$; and TUD-$S_2$ and TUD-$S_6$, respectively. The
authors performed code-to-code verification by sampling observable values at
201 equidistant points along the centerlines AA' and BB' and reporting the
discrepancy $\epsilon_c$ of each observable from each software
(indexed by $c$) for each measured observable $Q_c$ (not to be confused with
fission heat source $Q_f$), relative to the average of
that same observable $Q_{avg}$ from all participating software. Variables
$\epsilon_c$ and $Q_{avg}$ are calculated as:
%
\begin{align}
    \epsilon_c =& \sqrt{\frac{\sum^{N_p}_{i=1}\left[Q_c(\vec{r_i}) - Q_{avg}
    (\vec{r_i})\right]^2}{\sum^{N_p}_{i=1} Q^2_{avg}(\vec{r_i})}}
    \shortintertext{and}
    Q_{avg}(\vec{r_i}) =& \frac{1}{N_c} \sum^{N_c}_{c=1} Q_c(\vec{r_i})
    \shortintertext{where}
    Q_c(\vec{r_i}) =&
    \mbox{ value of observable $Q$ at location $\vec{r_i}$ from software $c$,}
    \nonumber \\
    N_p =& \mbox{ number of sampling points of quantity $Q$} = 201,
    \nonumber \\
    N_c =& \mbox{ number of participating software packages.} \nonumber
\end{align}

The average discrepancy $\epsilon$ across all software is calculated as:
%
\begin{align}
    \epsilon =& \frac{1}{N_c}\sum^{N_c}_{c=1} \epsilon_c
\end{align}

We adopted the averaged values $\epsilon$ and $Q_{avg}$ directly from the
reference work \cite{tiberga_results_2020} without including our results
in the calculations. We note that the benchmark does not provide a reference
solution, and a significantly erroneous value from one of the software packages
could heavily skew the discrepancy values. Nevertheless, the benchmark paper
reports good agreement among their software packages.

For observables measured along the centerlines AA' and BB', Tables
\ref{table:disc0} and \ref{table:disc1} report the discrepancy $\epsilon_c$ of
each observable from Moltres relative to the average of the benchmark
participants $Q_{avg}$ alongside the average discrepancy $\epsilon$ of
the benchmark participants. We also reproduce corresponding plots
in the benchmark paper for every observable along AA' or BB' in Figures
\ref{fig:0.1}, \ref{fig:0.2}, \ref{fig:0.3}, \ref{fig:1.1}, \ref{fig:1.2},
\ref{fig:1.3}, and \ref{fig:2.1} for a qualitative comparison of the results
from Moltres and the benchmark participants. Given the significant overlap in
the plot curves, these figures omit results from CNRS-$SP_1$ and TUD-$S_2$ to
reduce cluttering. The full dataset
of all observable results used in this results analysis is
available at \cite{park_results_2021}. Lastly, Table
\ref{table:rho} reports all reactivity and reactivity changes from
Steps 0.2, 1.1, 1.2, and 1.3.

\begin{figure}[h]
	\centering
    \includegraphics[width=.49\columnwidth]{0-1-vel-plot}
	\caption{Step 0.1 \textemdash\ Horizontal velocity component along BB'.}
	\label{fig:0.1}
\end{figure}
%
\begin{figure}[h]
	\centering
    \begin{subfigure}[b]{.49\textwidth}
      \centering
	  \includegraphics[width=\columnwidth]{0-2-fiss-plot}
	  \caption{Step 0.2 \textemdash\ Fission rate density along AA'.}
	  \label{fig:0.2}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{.49\textwidth}
      \centering
	  \includegraphics[width=\columnwidth]{0-3-temp-plot}
	  \caption{Step 0.3 \textemdash\ Temperature distribution along BB'.}
	  \label{fig:0.3}
    \end{subfigure}
\end{figure}
%
\FloatBarrier
%
\begin{table}[htb]
	\caption{Discrepancy values from Moltres alongside the average and standard
	deviation of the discrepancy values of the benchmark participants for Phase
	0.}
	\centering
	\small
	\begin{tabular}{l l c S S S}
		\toprule
		\multirow{2}{*}{\textbf{Step}} & \multirow{2}{*}{\textbf{Observable}} & \multirow{2}{*}{\textbf{Centerline}} & {\multirow{2}{*}{\textbf{Moltres [\%]}}} & \multicolumn{2}{c}{\textbf{Benchmark [\%]}} \\
		& & & & {Average} & {SD} \\
		\midrule
		\multirow{4}{*}{0.1} &
		\multirow{2}{*}{$u_x$} & AA' & 0.247 & 0.253 & 0.150 \\
		& & BB' & 0.266 & 0.318 & 0.102 \\
		\cmidrule{2-6}
		& \multirow{2}{*}{$u_y$} & AA' & 0.540 & 0.598 & 0.266 \\
		& & BB' & 0.468 & 0.795 & 0.421 \\
		\midrule
		{0.2} &
		{$\sum^6_g \Sigma_{f,g} \phi_g(\vec{r})$} & AA' & 0.313 & 0.285 & 0.153
		\\
		\midrule
		\multirow{2}{*}{0.3} &
		\multirow{2}{*}{$T$} & AA' & 0.090 & 0.085 & 0.031 \\
		& & BB' & 0.164 & 0.083 & 0.027\\
		\bottomrule
	\end{tabular}
	\label{table:disc0}
\end{table}

\subsubsection{Phase 0 results \& discussion}

Figures \ref{fig:0.1}, \ref{fig:0.2}, and \ref{fig:0.3} show that Moltres
accurately reproduced all three sets of results in Phase 0 for the velocity
field, fission rate density, and temperature. Table
\ref{table:disc0} reports the discrepancy values from Moltres for Phase 0 and
the corresponding average and \gls{SD} of the discrepancy values from
the benchmark participants
\cite{tiberga_results_2020}. Moltres performs very well as most discrepancy
values are either lower than or fall within one \gls{SD} of the benchmark
average discrepancies. The discrepancy value for $T$ along centerline BB' in
Step 0.3 is the only exception, with its value of 0.164\% being larger than
the benchmark average by 3 \gls{SD}.

Figure \ref{fig:0.3} shows that the $T$ distribution from Moltres is almost
identical to the corresponding distributions from CNRS-$SP_3$ and TUD-$S_6$
along most of centerline BB'. However, Figure \ref{fig:0.3-zoom} shows a
significant spread in the $T$ distributions along BB' from all software
packages near the top boundary. At $y = 2.0$ m, Moltres underpredicts the
temperature at 912.3 K compared to the benchmark participants' values which
range between 930.3 K and 948.1 K. This point on the top boundary lies directly downstream of
the velocity boundary condition discontinuity at the top-left corner.
Corner singularities are generally tricky to approximate with
continuous Galerkin methods \cite{kuhlmann_lid-driven_2018}.
The \gls{SUPG} stabilization scheme dampens numerical oscillations by
introducing pointwise artificial thermal diffusivity, which depends strongly on
the inverse of local velocity magnitude \cite{peterson_overview_2018}.
Therefore, while the \gls{SUPG} scheme effectively eliminates
spurious numerical oscillations everywhere else, it provides little damping
along the top boundary due to the relatively large non-zero velocity boundary
condition. On the other hand, the temperature values in the rest of the domain
and the average discrepancies of the other variables show that Moltres can
still accurately reproduce the expected results, and the temperature deviations
along the top boundary do not impact the overall integrity of our results.

\begin{figure}[htb]
	\centering
	\includegraphics[width=.49\columnwidth]{0-3-temp-plot-zoom}
	\caption{Step 0.3 \textemdash\ Temperature distribution along BB' for y = 1.94 m to
	y = 2.00 m.}
	\label{fig:0.3-zoom}
\end{figure}

Lastly, we observe in table \ref{table:rho} that the reactivity $\rho$ value of
465.6 pcm from Moltres falls well within the range of $\rho$ values from the
benchmark, which range from 353.7 pcm up to 578.1 pcm. Given that Moltres 
adopts the neutron diffusion model, our $\rho$ value agrees closest to the
results from the software packages which also adopt the neutron diffusion model
or theoretically-equivalent models such as the $SP_1$ and $S_2$ neutron
transport models, namely CNRS-$SP_1$, PoliMi, PSI, and TUD-$S_2$.

\begin{table}[htb]
    \caption{Reactivity $\rho$ and change in reactivity
    $\left(\rho_a - \rho_b\right)$ values from Steps 0.2, 1.1,
    1.2, and 1.3. All units are in pcm.}
    \centering
    \small
    \setlength\tabcolsep{2pt}
    \begin{tabular}{l S S S S}
        \toprule
        \multirow{2}{*}{\textbf{Software}} & {\textbf{Step 0.2}} &
        {\textbf{Step 1.1}} & {\textbf{Step 1.2}} & {\textbf{Step 1.3}} \\
        & {$\rho_{s_{0.2}}$}
        & {$\rho_{s_{1.1}} - \rho_{s_{0.2}}$}
        & {$\rho_{s_{1.2}} - \rho_{s_{1.1}}$}
        & {$\rho_{s_{1.3}} - \rho_{s_{0.2}}$} \\
        \midrule
        Moltres     & 465.6 & -62.7 & -1142.2 & -1207.7 \\
        CNRS-$SP_1$ & 411.3 & -62.5 & -1152.0 & -1220.5 \\
        CNRS-$SP_3$ & 353.7 & -62.6 & -1152.7 & -1220.7 \\
        PoliMi      & 421.2 & -62.0 & -1161.0 & -1227.0 \\
        PSI         & 411.7 & -63.0 & -1154.8 & -1219.6 \\
        TUD-$S_2$   & 482.6 & -62.0 & -1145.2 & -1208.5 \\
        TUD-$S_6$   & 578.1 & -60.7 & -1122.0 & -1184.4 \\
        \bottomrule
    \end{tabular}
    \label{table:rho}
\end{table}

\FloatBarrier

\subsubsection{Phase 1 results \& discussion}

Table \ref{table:disc1} shows the discrepancy values from Moltres relative to
the average and \gls{SD} of the benchmark participants for Steps 1.1, 1.2, and
1.3 and the corresponding average discrepancy values from the benchmark
\cite{tiberga_results_2020}. The subsequent subsections discuss the results
for each benchmark step in Phase 1.
%
\begin{table}[htb]
	\caption{Discrepancy values from Moltres alongside the average and standard
	deviation of the discrepancy values of the benchmark participants for Phase
	1.}
	\centering
	\small
	\begin{tabular}{l l c S S S}
		\toprule
		\multirow{2}{*}{\textbf{Step}} & \multirow{2}{*}{\textbf{Observable}} & \multirow{2}{*}{\textbf{Centerline}} & {\multirow{2}{*}{\textbf{Moltres [\%]}}} & \multicolumn{2}{c}{\textbf{Benchmark [\%]}} \\
		& & & & {Average} & {SD} \\
		\midrule
		\multirow{2}{*}{1.1} &
		\multirow{2}{*}{$\sum_i \lambda_i C_i$} & AA' & 0.603 & 0.346 & 0.166
		\\
		& & BB' & 0.327 & 0.294 & 0.153 \\
		\midrule
		\multirow{4}{*}{1.2} &
		\multirow{2}{*}{$T$} & AA' & 0.076 & 0.095 & 0.015 \\
		& & BB' & 0.179 & 0.089 & 0.012 \\
		\cmidrule{2-6}
		& \multirow{2}{*}{\footnotesize $\Delta\left[\sum^6_g \Sigma_{f,g} \phi_g(\vec{r})
		\right]_{s_{1.2}-s_{0.2}}$} & AA' & 1.110 & 1.576 & 0.564 \\
		& & BB' & 1.089 & 1.133 & 0.392 \\
		\midrule
		\multirow{7}{*}{1.3} &
		{$u_x$} & AA' & 0.123 & 0.691 & 0.566 \\
		\cmidrule{2-6}
		& \multirow{2}{*}{$u_y$} & AA' & 0.237 & 0.329 & 0.131 \\
		& & BB' & 0.238 & 0.356 & 0.217 \\
		\cmidrule{2-6}
		& \multirow{2}{*}{$T$} & AA' & 0.064 & 0.057 & 0.023 \\
		& & BB' & 0.070 & 0.080 & 0.024 \\
		\cmidrule{2-6}
		& \multirow{2}{*}{$\sum_i \lambda_i C_i$} & AA' & 1.043 & 0.460 & 0.190
		\\
		& & BB' & 0.462 & 1.194 & 0.178 \\
		\bottomrule
	\end{tabular}
	\label{table:disc1}
\end{table}

\paragraph{Step 1.1: Circulating fuel}

Figure \ref{fig:1.1} shows good qualitative agreement in the delayed neutron
source distribution along BB' among Moltres and the benchmark participants.
From Table \ref{table:disc1}, Moltres reports discrepancies of 0.603\% and
0.327\% along the centerlines AA' and BB', respectively. Both values are
within two and one \gls{SD}, respectively, of the average discrepancies of the
benchmark participants (0.346\% and 0.294\%).
In Table \ref{table:rho}, we observe that the change in
$\rho$ relative to Step 0.2 is $-62.7$ pcm for Moltres, and this value is
consistent with the $-63.0$ to $-62.0$ pcm range that most of the benchmark
participants' values fall in.
%
\begin{figure}[htb]
	\centering
    \includegraphics[width=.49\columnwidth]{1-1-dnp-x-plot}
    \includegraphics[width=.49\columnwidth]{1-1-dnp-y-plot}
	\caption{Step 1.1 \textemdash\ Delayed neutron source along AA' (top) and BB'
	(bottom).}
	\label{fig:1.1}
\end{figure}
%
\begin{figure}[htb]
	\centering
	\includegraphics[width=.49\columnwidth]{1-2-temp-plot}
	\includegraphics[width=.49\columnwidth]{1-2-fiss-plot}
	\caption{Step 1.2 \textemdash\ Temperature distribution and change in fission rate
	density along AA'.}
	\label{fig:1.2}
\end{figure}

\FloatBarrier

\paragraph{Step 1.2: Power coupling}

Figure \ref{fig:1.2} shows the temperature distribution and the change in
fission rate density along AA' from Step 1.2. Similar to Step 0.3, the
temperature distribution from Moltres agrees closest with CNRS-$SP_3$ and
TUD-$S_2$. Table \ref{table:disc1} reports the same trends we observed in Phase
0; the average discrepancy in temperature along BB' from Moltres for Step 1.2
is marginally higher than the benchmark, while the average discrepancy in the
fission rate density is within one \gls{SD} range to the benchmark average.
From Table \ref{table:rho}, Moltres also reports a change in $\rho$
relative to Step 1.1 of $-1142.2$ pcm, which
falls within the range of reported benchmark participants' values.

\paragraph{Step 1.3: Buoyancy}

Figure \ref{fig:1.3} shows the vertical velocity component, temperature, and
delayed neutron source distributions along AA'.
Moltres reproduces the correct trend in all three physical
observables. Step 1.3 has a relatively slower buoyancy-driven flow profile with
no forced flow from the top boundary. Consequently, there are no temperature
deviations near the top boundary, and we observe in Table \ref{table:disc1} that
the average discrepancy value of 0.070\% for the temperature distribution along
BB' is in much closer agreement to the benchmark value of 0.080\% compared to
the corresponding temperature discrepancy values for Step 0.3 and 1.2.

Table \ref{table:rho} shows that the change in $\rho$ from
Moltres (-1207.7 pcm) falls within the range of reported benchmark values and
matches the software from TUD-$S_2$ (-1208.5 pcm) the closest. This observation is likely
due to the similar solvers (diffusion and $S_2$ neutronics models) and
methods of solution (finite element method on uniform meshes) employed by our
respective software.
%
\begin{figure}[htb]
	\centering
	\includegraphics[width=.49\columnwidth]{1-3-vel-plot}
	\includegraphics[width=.49\columnwidth]{1-3-temp-plot}
	\includegraphics[width=.49\columnwidth]{1-3-dnp-plot}
	\caption{Step 1.3 \textemdash\ Vertical velocity component, temperature distribution,
	and delayed neutron source along AA'.}
	\label{fig:1.3}
\end{figure}

\FloatBarrier

\paragraph{Step 1.4: Full coupling}

Figure \ref{fig:cnrs-color} shows 2D temperature distribution and velocity
streamlines from Moltres for Step 1.4 with $U_{lid} = 0.5$ m$\cdot$s$^{-1}$ and
$P = 1$ GW. Table \ref{table:full} shows the change in $\rho$ under the various
$U_{lid}$ and $P$ values. Refer to Tiberga et al.'s paper
\cite{tiberga_results_2020} for the benchmark participants' corresponding
values. The change in $\rho$ values from Moltres all fall within the range of
benchmark values
for all cases. Furthermore, the $\Delta\rho$ values are all within 1.1 pcm of
the corresponding values from the TUD-S$_2$ model in the benchmark paper. Given
that the $S_2$ discrete ordinates method with isotropic sources is theoretically equivalent to the
multigroup neutron diffusion method, Moltres is largely
consistent with the benchmark participants outside of differences from the
neutronics models.

\begin{figure}[htb]
  \centering
  \includegraphics[width=.7\columnwidth]{full-coupled}
  \caption{Temperature distribution from Moltres for the fully coupled
  system (Step 1.4) with buoyancy effects, $P = 1$ GW, and $U_{lid} = 0.5$
  m$\cdot$s$^{-1}$. The lines correspond to the streamlines of the velocity
  field.}
  \label{fig:cnrs-color}
\end{figure}
%
\begin{table}[htb]
	\caption{Reactivity change in Step 1.4, relative to Step 0.2 under various
	$U_{lid}$ and $P$ values.}
	\centering
	\small
	\setlength\tabcolsep{1.5pt}
	\begin{tabular}{c c c c c c}
		\toprule
		& \multicolumn{5}{c}{$\rho_{s1.4} - \rho_{s0.2}$ [pcm]} \\
		\midrule
		{\backslashbox{$U_{lid}$ [m$\cdot$s$^{-1}$]}{$P$ [GW]}} & 0.2 & 0.4 & 0.6 & 0.8 & 1.0 \\
		\midrule
		0.0 & -263.7 & -498.3 & -730.9 & -966.7 & -1207.7 \\
		0.1 & -265.9 & -498.7 & -730.6 & -966.0 & -1206.7 \\
		0.2 & -268.1 & -498.8 & -729.4 & -963.7 & -1203.6 \\
		0.3 & -269.9 & -498.5 & -727.8 & -960.8 & -1199.5 \\
		0.4 & -271.9 & -498.5 & -726.5 & -958.3 & -1195.7 \\
		0.5 & -274.2 & -498.7 & -725.6 & -956.4 & -1192.7 \\
		\bottomrule
	\end{tabular}
	\label{table:full}
\end{table}

\FloatBarrier

\begin{table}[htb]
	\caption{Discrepancy values from Moltres alongside the average and standard
	deviation of the discrepancy values of the benchmark participants for Step
	2.1.}
	\centering
	\small
	\begin{tabular}{l l S S S}
		\toprule
		\multirow{2}{*}{\textbf{Step}} & \multirow{2}{*}{\textbf{Observable}} & {\multirow{2}{*}{\textbf{Moltres [\%]}}} & \multicolumn{2}{c}{\textbf{Benchmark [\%]}} \\
		& & & {Average} & {SD} \\
		\midrule
		\multirow{2}{*}{2.1} & Gain & 0.493 & 0.587 & 0.244 \\
		\cmidrule{2-5}
		& Phase shift & 1.741 & 2.176 & 0.554 \\
		\bottomrule
	\end{tabular}
	\label{table:disc2}
\end{table}
%
\begin{figure}[htb]
	\centering
	\includegraphics[width=.49\columnwidth]{2-1-gain-plot}
	\includegraphics[width=.49\columnwidth]{2-1-phase-plot}
	\caption{Step 2.1 \textemdash\ Bode gain and phase plots of the frequency response of
	the fully coupled system.}
	\label{fig:2.1}
\end{figure}

\subsubsection{Phase 2 results \& discussion}

Lastly, the following subsection discusses the results for the transient cases
in Step 2.1, which involve measuring the response in power output to periodic
perturbations in the heat transfer coefficient.

\paragraph{Step 2.1: Forced convection transient}

Figure \ref{fig:2.1} shows the Bode gain and phase shift plots of the response
in power output in the fully coupled system. Along with the average discrepancy
values from Table \ref{table:disc2}, the results show that Moltres is
consistent with the benchmark. The gain data points from all \gls{MSR} software
agree closely with one another. Moltres reports an average discrepancy value of
0.496\%, slightly lower than the benchmark average of 0.587\%. On the other
hand, the phase shift data points show a greater spread over the various driving
frequencies. We note the different timestepping schemes and timestep
sizes among the different software packages, which is likely responsible for
the variations in the phase shift. Even with a precision of
$\pm0.9^\circ$ for each phase shift value, Moltres accurately reproduces the
correct trend with a lower average discrepancy (1.741\%) than the benchmark
participants' average (2.176\%).

\FloatBarrier

### 2024-07-22 msre.tex
MSRE chapter edits

\section{Group Constant Data Generation on OpenMC}

The group constants required by either or both neutron diffusion and $S_N$ neutron transport
methods are:
%
\begin{itemize}
  \item $\Sigma_{t,g}$: Macroscopic total cross section in group $g$,
  \item $\Sigma_{r,g}$: Macroscopic removal cross section in group $g$,
  \item $\Sigma_s^{g'\rightarrow g}$: Macroscopic group-to-group scattering cross section matrix,
  \item $\Sigma_{s,l}^{g'\rightarrow g}$: $l$-th Legendre moment of the macroscopic
    group-to-group scattering cross section matrix,
  \item $\Sigma_{sp,l}^{g'\rightarrow g}$: $l$-th Legendre moment of the macroscopic
    group-to-group scattering production cross section matrix,
  \item $D_g$: $P_1$-based diffusion coefficient in group $g$,
  \item $\nu\Sigma_{f,g}$: Product of the average number of neutrons produced per fission and the
    macroscopic fission cross section in group $g$,
  \item $\chi_g$: Neutron fission spectrum in group $g$.
\end{itemize}
%
I generated the group constants using OpenMC's multigroup cross section generation capability
\cite{boyd_multigroup_2019}. The group constant postprocessing script for Moltres
(\texttt{moltres\_xs.py}) calculates $\Sigma_{r,g}$ as
%
\begin{align}
  \Sigma_{r,g} =& \sum^G_{g'\neq g}\Sigma_s^{g\rightarrow g'}+\Sigma_{a,g}-\left(\Sigma_{sp}^{g
    \rightarrow g} - \Sigma_s^{g\rightarrow g}\right)
  \shortintertext{where}
      \Sigma_{a,g} =& \mbox{ macroscopic absorption cross section in group $g$,} \nonumber \\
      \Sigma_{sp}^{g\rightarrow g} =& \mbox{ macroscopic scattering production cross section from
      group $g$ to $g$.} \nonumber
\end{align}
%
$\Sigma_{r,g}$ primarily represents the loss of neutrons from group $g$ through outscattering and
absorption. $\Sigma_{sp}^{g\rightarrow g}$ incorporates neutron multiplication effects from neutron
knockout reactions into the scattering cross section. Neutron knockout reactions are commonly
tallied as scattering reactions, but $\Sigma_{r,g}$ is a convenient parameter in which to
incorporate neutron knockout effects in the neutron diffusion equations. I ran all test cases for
$S_N$ and hybrid method calculations with up to the 2nd Legendre moments of the scattering
cross sections ($L=2$).

OpenMC uses the $P_1$ flux-limited formulation \cite{pomraning_flux-limited_1984} for calculating
$D_g$ as follows
%
\begin{align}
  D_g =& \frac{1}{3\Sigma_{tr,g}},
  \shortintertext{where}
  \Sigma_{tr,g} =& \frac{\langle\Sigma_{t,g}\phi_g\rangle-\langle\Sigma_{s1,g}\phi_g\rangle}
  {\langle\phi_g\rangle}, \nonumber \\
  \langle\Sigma_{t,g}\phi_g\rangle =& \int_{r\in V}dr \int_{4\pi}d\Omega\int^{E_{g-1}}_{E_g}dE\
  \Sigma_{t,g}(r,E)\Psi(r,E,\Omega), \nonumber \\
  \langle\Sigma_{s1,g}\phi_g\rangle =& \int_{r\in V}dr \int_{4\pi}d\Omega\int^{E_{g-1}}_{E_g}dE
  \int_{4\pi}d\Omega'\int^{\infty}_0dE'\int^1_{-1}d\mu\ \mu\Sigma_s(r,E'\rightarrow E,\Omega'\cdot
  \Omega)\phi(r,E',\Omega'), \nonumber \\
  \langle \phi \rangle =& \int_{r\in V}dr\int_{4\pi}d\Omega\int^{E_{g-1}}_{E_g}dE\ \Psi(r,E,\Omega)
  .\nonumber
\end{align}

I ran all \gls{MSRE} neutronics simulations with the eight neutron energy group structure proposed
by Jaradat for \gls{MSRE} analysis \cite{jaradat_development_2021-1}.
Table \ref{table:energy-group} shows the upper neutron energy bounds of the eight-group structure.
I selected this energy group structure as a compromise between accuracy and computational expense
among the various group structures investigated by Jaradat ranging from four to sixteen groups.

\begin{table}[htb]
  \centering
  \caption{Neutron energy group structure in this work. Originally devised by Jaradat
  \cite{jaradat_development_2021-1}.}
  \begin{tabular}{c S}
    \toprule
    Group & {Upper energy bound [eV]} \\
    \midrule
    1 & 2.000$\times 10^7$ \\
    2 & 1.353$\times 10^6$ \\
    3 & 6.734$\times 10^4$ \\
    4 & 9.118$\times 10^3$ \\
    5 & 1.487$\times 10^2$ \\
    6 & 4.000$\times 10^0$ \\
    7 & 6.250$\times 10^{-1}$ \\
    8 & 8.000$\times 10^{-2}$ \\
    \bottomrule
  \end{tabular}
  \label{table:energy-group}
\end{table}

I ran all cases on OpenMC to generate the required group constants and reference $k_\text{eff}$
values to assess the accuracy of the deterministic methods against OpenMC in continuous-energy
(OpenMC-CE) and multigroup (OpenMC-MG) modes. Table \ref{table:var} shows how the
position, direction of travel, neutron energy, and angle-dependence in $\Sigma_s$ are represented
by OpenMC and the deterministic methods. The hybrid $S_N$-diffusion method features representations
from either $S_N$ or neutron diffusion depending on the subregion under consideration.
Comparing OpenMC-MG results with OpenMC-CE results allows us
to quantify errors arising from neutron energy group discretization and the scattering cross
section simplifications.

\begin{table}[tb!]
  \centering
  \footnotesize
  \caption{Variable representations in OpenMC under continuous-energy (OpenMC-CE) and multigroup
  (OpenMC-MG) modes, and in the $S_N$ neutron transport and neutron diffusion.}
  \begin{tabular}{c c c c c c}
    \toprule
    Variable & OpenMC-CE & OpenMC-MG & $S_N$ Transport & Diffusion \\
    \midrule
    Position, $\bm{r}$ & Continuous & Continuous & Discrete & Discrete \\
    Direction of travel, $\bm{\hat{\Omega}}$ & Continuous & Continuous & Discrete & N/A \\
    Energy, $E$ & Continuous & Discrete & Discrete & Discrete & Discrete \\
    Angle-dependence in $\Sigma_s$ & Continuous & $L=2$ & $L=2$
    & N/A \\
    \bottomrule
  \end{tabular}
  \label{table:var}
\end{table}

\section{1-D Models}

\subsection{1-D Model Setup}

\subsubsection{Model Geometry}

I designed six 1-D test cases with increasing complexity to verify the $S_N$ and hybrid
$S_N$-diffusion implementations and to test the performance of the hybrid
method in response to various geometrical features. The last two cases resemble the
reference \gls{MSRE} design \cite{robertson_msre_1965}, which has centrally located control rods
and air-filled rod guide tubes. Figure \ref{fig:case-geom} shows the geometries of Cases 1a to
3b. All geometries have reflective boundary conditions at $x=0$ cm to reduce computational costs by
creating half-core or repeating unit cell models. Cases 1a and 1b are repeating unit
cell models with reflecting boundaries on the right-side boundaries. Cases 2a, 2b, 3a, and 3b
are 1-D half-core models with vacuum boundaries on the right-side boundaries.

Cases 1a and 1b are simple test cases containing neutron multiplying regions. Cases 1a and 1b serve
to verify basic neutronics phenomena modeling (e.g., fission, scattering, absorption) of the newly
implemented $S_N$ and neutron diffusion methods in Python and $S_N$ method in Moltres. I did not
apply the hybrid method on these cases. Cases 2a and 2b represent 1-D models of the \gls{MSRE} with
air-filled control rod guide tube regions, a homogenized fuel-graphite lattice regions, and outer
vessel regions. Case 2b additionally contains a 0.5cm-thick control rod region to contrast with
Case 2a for control rod worth calculations. Cases 3a and 3b retain the heterogeneous fuel-graphite
lattice geometry present in the \gls{MSRE}.

\begin{figure}[htb!]
  \centering
  \includegraphics[width=\columnwidth]{case-geometry}
  \caption{Geometries of the 1-D test cases. The material labeled ``mixture'' represents a
    homogeneous mixture of fuel and graphite at a ratio of 22.5\%-77.5\% by volume. All geometries
    have reflective boundary conditions on the boundary at $x=0$ cm. The right-side boundaries are
    reflective for Cases 1a and 1b, and vacuum for Cases 2a, 2b, 3a, and 3b.}
  \label{fig:case-geom}
\end{figure}

\subsubsection{Mesh Convergence Tests}

\begin{figure}[htb!]
  \centering
  \includegraphics[width=\columnwidth]{rho}
  \caption{}
  \label{fig:1d-rho}
\end{figure}

\begin{figure}[htb!]
  \centering
  \includegraphics[width=\columnwidth]{rho}
  \caption{}
  \label{fig:1d-rho}
\end{figure}

\subsection{1-D Numerical Results}

\begin{figure}[htb!]
  \centering
  \includegraphics[width=\columnwidth]{rho}
  \caption{}
  \label{fig:1d-rho}
\end{figure}

\begin{figure}[htb!]
  \centering
  \includegraphics[width=\columnwidth]{worth}
  \caption{}
  \label{fig:1d-worth}
\end{figure}

\section{2-D Models}

\subsection{2-D Model Setup}

\subsection{2-D Numerical Results}

\section{3-D Models}

\subsection{3-D Model Setup}

\subsection{3-D Numerical Results}

### 2024-07-22 verification.tex
MSRE chapter edits

\input{moltres}
\input{benchmark}
\section{MSRE Zero-Power Pump Experiments}
\section{Turbulence Model Verification}

### 2024-08-11 appendix.tex
msre chapter edits

@@
-\section{Numerical Implementation} \label{sec:implementation}
+\section{Numerical Implementation of Neutronics Solvers on Python} \label{sec:implementation}
 
-In Section \ref{sec:theory}, I presented the theoretical basis for the hybrid
-$S_N$-diffusion method and demonstrated the method on a simple problem. Now I will discuss
-numerical implementation details of group
-constant data processing and the neutron diffusion, $S_N$ neutron transport, and hybrid
-$S_N$-diffusion solvers. I implemented all numerical solvers and analysis scripts in the Python
-programming language. First, I will discuss the material group constant data generation and
-postprocessing steps. Next, I present the implementation details of the neutron
-diffusion and $S_N$ numerical methods. Lastly, I present how they are coupled to form the hybrid
-$S_N$-diffusion method.
+In Section \ref{sec:diffusion-correction}, I presented the theoretical basis for the hybrid
+$S_N$-diffusion method with diffusion correction terms. This section presents numerical
+implementation details of the neutron diffusion, $S_N$ neutron transport, and hybrid
+$S_N$-diffusion solvers in the Python programming language for this work.
 
-\subsection{Group Constant Data Generation}
-
-The group constants required by either or both neutron diffusion and $S_N$ neutron transport
-methods are:
-%
-\begin{itemize}
-  \item $\Sigma_{t,g}$: Macroscopic total cross section in group $g$,
-  \item $\Sigma_{r,g}$: Macroscopic removal cross section in group $g$,
-  \item $\Sigma_s^{g'\rightarrow g}$: Macroscopic group-to-group scattering cross section matrix,
-  \item $\Sigma_{s,l}^{g'\rightarrow g}$: $l$-th Legendre moment of the macroscopic
-    group-to-group scattering cross section matrix,
-  \item $\Sigma_{sp,l}^{g'\rightarrow g}$: $l$-th Legendre moment of the macroscopic
-    group-to-group scattering production cross section matrix,
-  \item $D_g$: $P_1$-based diffusion coefficient in group $g$,
-  \item $\nu\Sigma_{f,g}$: Product of the average number of neutrons produced per fission and the
-    macroscopic fission cross section in group $g$,
-  \item $\chi_g$: Neutron fission spectrum in group $g$.
-\end{itemize}
-%
-These group constants are generated using OpenMC's multigroup cross section generation capability
-\cite{boyd_multigroup_2019} and postprocessed using a Python script into JSON format files.
-$\Sigma_{r,g}$ is the only quantity that OpenMC does not directly provide. It is calculated as
-%
-\begin{align}
-  \Sigma_{r,g} =& \sum^G_{g'\neq g}\Sigma_s^{g\rightarrow g'}+\Sigma_{a,g}-\left(\Sigma_{sp}^{g
-    \rightarrow g} - \Sigma_s^{g\rightarrow g}\right)
-  \shortintertext{where}
-      \Sigma_{a,g} =& \mbox{ macroscopic absorption cross section in group $g$,} \nonumber \\
-      \Sigma_{sp}^{g\rightarrow g} =& \mbox{ macroscopic scattering production cross section from
-      group $g$ to $g$.} \nonumber
-\end{align}
-%
-$\Sigma_{r,g}$ primarily represents the loss of neutrons from group $g$ through outscattering and
-absorption. $\Sigma_{sp}^{g\rightarrow g}$ incorporates neutron multiplication effects from neutron
-knockout reactions into the scattering cross section. Neutron knockout reactions are commonly
-tallied as scattering reactions, and $\Sigma_{r,g}$ is a convenient term to
-incorporate neutron knockout effects into the neutron diffusion equations. I ran all test cases for
-$S_N$ and hybrid method calculations with up to the 2nd Legendre moments of the scattering
-cross sections ($L=2$).
-
-OpenMC uses the $P_1$ flux-limited formulation \cite{pomraning_flux-limited_1984} for calculating
-$D_g$ as follows
-%
-\begin{align}
-  D_g =& \frac{1}{3\Sigma_{tr,g}},
-  \shortintertext{where}
-  \Sigma_{tr,g} =& \frac{\langle\Sigma_{t,g}\phi_g\rangle-\langle\Sigma_{s1,g}\phi_g\rangle}
-  {\langle\phi_g\rangle}, \nonumber \\
-  \langle\Sigma_{t,g}\phi_g\rangle =& \int_{r\in V}dr \int_{4\pi}d\Omega\int^{E_{g-1}}_{E_g}dE\
-  \Sigma_{t,g}(r,E)\Psi(r,E,\Omega), \nonumber \\
-  \langle\Sigma_{s1,g}\phi_g\rangle =& \int_{r\in V}dr \int_{4\pi}d\Omega\int^{E_{g-1}}_{E_g}dE
-  \int_{4\pi}d\Omega'\int^{\infty}_0dE'\int^1_{-1}d\mu\ \mu\Sigma_s(r,E'\rightarrow E,\Omega'\cdot
-  \Omega)\phi(r,E',\Omega'), \nonumber \\
-  \langle \phi \rangle =& \int_{r\in V}dr\int_{4\pi}d\Omega\int^{E_{g-1}}_{E_g}dE\ \Psi(r,E,\Omega)
-  .\nonumber
-\end{align}
-
-\subsection{Neutron Diffusion Method}
+\subsection{Neutron Diffusion Method} \label{sec:python-diffusion}
 
@@
 
-\subsection{$S_N$ Neutron Transport Method}
+\subsection{$S_N$ Neutron Transport Method} \label{python-sn}
 
@@
 
-\subsection{Hybrid $S_N$-Diffusion Method}
+\subsection{Hybrid $S_N$-Diffusion Method} \label{python-hybrid}
 

### 2024-08-11 hybrid.tex
msre chapter edits

@@
 demonstrations due to superior numerical properties. I will make clear the reasons for this
-selection through 1-D test case results in Chapter \ref{chap:hybrid-verification}.
-
-\subsection{Diffusion Correction}
+selection through 1-D test case results in Chapter \ref{chap:msre}.
+
+\subsection{Diffusion Correction} \label{sec:diffusion-correction}
 
@@
 %
-where $D^s_g$ is the \gls{SVDC}, and the $tr$ superscript denotes the transport-derived neutron
+where $D^s_g$ is the diffusion correction parameter, and the $tr$ superscript denotes the
+transport-derived neutron
 current and scalar flux solutions from the $S_N$ method. Transport corrections introduced through
 $J_{tr}$ are scaled by the flux gradient. We assumed that it varies continuously and is
-at least once differentiable except at dissimilar material interfaces. \glspl{SVDC} provide
+at least once differentiable except at dissimilar material interfaces. $D^s_g$ provides
 pointwise corrections to closely match the diffusion flux solution to the $S_N$ flux solution.
@@
 
-Unlike their approach, the \gls{SVDC} formulation in Eq.\ \ref{eq:svdc} allows for continuously
+Unlike their approach, my diffusion correction scheme (Eq. \ref{eq:svdc} allows for continuously
 varying diffusion coefficients to reduce truncation error. In practice, the discretization order of
-\gls{SVDC} variables in a numerical calculation would be the same as the discretization order of
+$D^s_g$ in a numerical calculation would be the same as the discretization order of
 the reference flux solution. This formulation introduces a minor change to the finite difference
@@
 
-\subsection{Drift Correction Term}
+\subsection{Drift Correction Term} \label{sec:drift-correction}
 
@@
 In order to reduce the computational cost of the high-level $S_N$ calculation in a reactor
-simulation, I propose reducing
-the problem domain of the $S_N$ method to a \textit{correction region} containing the control rod
+simulation, I propose reducing the problem domain of the $S_N$ method to a
+\textit{correction subregion} containing the control rod
 and its vicinity. Consequently, the hybrid $S_N$-diffusion method may retain accurate neutron flux
@@
 region as the $S_N$ \textit{subproblem} or \textit{subsolver}. I define the full problem
-domain and the correction region as $V_0$ and $V_1$, respectively, where
+domain and the correction subregion as $V_0$ and $V_1$, respectively, where
 $V_1\subseteq V_0$. The iterative algorithm for the hybrid $S_N$-diffusion method is as follows:
@@
 in the flux and transport correction values since the $P_1$ approximation is accurate for up to
-linear anisotropies in the flux. However, as I will show in Chapter \ref{chap:hybrid-verification},
+linear anisotropies in the flux. However, as I will show in Chapter \ref{chap:msre},
 the influence of boundary conditions on the accuracy of
@@
 thick media due to competing influences from the bulk scattering and absorption effects in $V_1$.
-Results in Chapter \ref{chap:hybrid-verification} will show that transport correction values
+Results in Chapter \ref{chap:msre} will show that transport correction values
 ($D^s_g$ or $\vec{D}_g$) generated with the $S_N$ subsolver within $V_1$ are accurate everywhere up
@@
 
-Here, I use Case 3b of the 1-D test cases to be covered in Chapter \ref{chap:hybrid-verification}
+Here, I use Case 3b of the 1-D test cases to be covered in Chapter \ref{chap:msre}
 to illustrate the buffer region and the behavior of $\vec{D}_g$. Case 3b is a 1-D slab problem
@@
     \caption{The reference and hybrid drift ($\vec{D}_g$) distributions for group 1 and 8 calculated
-      from $S_8$ and $S_8$-diffusion simulations. The correction region $V_1$ spans $x=0$ cm to
+      from $S_8$ and $S_8$-diffusion simulations. The correction subregion $V_1$ spans $x=0$ cm to
       $x=10$ cm.}

### 2024-08-11 msre.tex
msre chapter edits

@@
 
-The group constants required by either or both neutron diffusion and $S_N$ neutron transport
+The group constants required by the neutron diffusion or $S_N$ neutron transport
 methods are:
@@
   \cite{jaradat_development_2021-1}.}
-  \begin{tabular}{c S}
+  \begin{tabular}{r S}
     \toprule
@@
 
+I set the material compositions and densities of every component in the \gls{MSRE} models using
+reference data from \gls{MSRE} benchmark specifications collated by Fratoni et al.
+\cite{fratoni_molten_2020}. All OpenMC models utilize the ENDF/B-VII.1 nuclear data library
+\cite{chadwick_endf/b-vii.1_2011} and include thermal scattering cross section data for beryllium
+and carbon target nuclei in the fuel salt and graphite moderator. All models have a uniform
+temperature of 900 K unless otherwise stated. Detailed information pertaining specifically to the
+1-D, 2-D, and 3-D OpenMC models may be found in the respective subsections below.
+
+\section{Neutronics Methods}
+
 I ran all cases on OpenMC to generate the required group constants and reference $k_\text{eff}$
@@
 
-\begin{table}[tb!]
+\begin{table}[htb!]
   \centering
@@
   (OpenMC-MG) modes, and in the $S_N$ neutron transport and neutron diffusion.}
-  \begin{tabular}{c c c c c c}
+  \begin{tabular}{l l l l l}
     \toprule
@@
     Direction of travel, $\bm{\hat{\Omega}}$ & Continuous & Continuous & Discrete & N/A \\
-    Energy, $E$ & Continuous & Discrete & Discrete & Discrete & Discrete \\
-    Angle-dependence in $\Sigma_s$ & Continuous & $L=2$ & $L=2$
-    & N/A \\
+    Energy, $E$ & Continuous & Discrete & Discrete & Discrete \\
+    Angle-dependence in $\Sigma_s$ & Continuous & $L=2$ & $L=2$ & N/A \\
     \bottomrule
@@
 
-\section{1-D Models}
+The rest of this section pertains to the deterministic neutronics methods used in this work for
+modeling the \gls{MSRE}.
+For the 1-D models, I performed neutronics calculations with the $S_N$, neutron diffusion, and
+hybrid method implementations on Moltres and Python. Table \ref{table:nts-methods} highlights the
+differences between the two implementations. The $S_N$ neutron transport method implementation on
+Moltres follows the methodology for solving the \gls{SAAF} formulation of the $S_N$ equations as
+described in Section \label{sec:saaf} while the implementation on Python follows the methodology
+for solving the original first-order formulation as described in Section \label{sec:python-sn}.
 
-\subsection{1-D Model Setup}
+The implementations on Moltres are discretized in space using the finite element method
+while the implementations on Python use finite difference or diamond difference schemes. The most
+crucial difference lies in the transport correction formulations for the hybrid method. The Moltres
+implementation is based on the drift correction formulation described in Section
+\label{sec:drift-correction} while the Python implementation is based on the diffusion correction
+formulation described in Section \label{sec:diffusion-correction}.
 
-\subsubsection{Model Geometry}
+I will refer to the methods and results from either Molres or Python implementations with the
+Moltres- or Python- prefix to clearly set them apart, i.e. Moltres-diffusion, Moltres-$S_N$,
+Moltres-hybrid, Python-diffusion, Python-$S_N$, and Python-hybrid.
+
+For 2- and 3-D models, I abandoned the diffusion correction formulation in favor of the drift
+correction formulation due to superior numerical performance in the latter. The analysis
+supporting this decision may be found in the analysis of 1-D results in Section
+\ref{sec:1d-results}.
+
+\begin{table}[htb!]
+  \centering
+  \footnotesize
+  \caption{Implementation differences of neutronics methods on Moltres and Python.}
+  \begin{tabular}{p{0.14\linewidth} p{0.25\linewidth} p{0.15\linewidth} p{0.15\linewidth} p{0.17\linewidth}}
+    \toprule
+    Implementations & Spatial discretization & $S_N$ formulation & $S_N$ acceleration &
+    Hybrid transport\newline correction \\
+    \midrule
+    Moltres & Finite element method & SAAF & Nonlinear diffusion acceleration & Drift correction \\
+    Python & Finite difference (Diffusion)\newline Diamond difference ($S_N$) & First-order & None & Diffusion correction \\
+    \bottomrule
+  \end{tabular}
+  \label{table:nts-methods}
+\end{table}
+
+\section{Test Metrics} \label{sec:test-metrics}
+
+In this section, I define several metrics for comparing simulation results among the various
+neutronics methods.
+
+For steady-state eigenvalue calculations, all solvers directly provide the effective neutron
+multiplication factor ($k_\text{eff}$), the rate of neutron population increase between successive
+neutron generations. Reactivity is defined as
+%
+\begin{gather}
+  \rho \equiv \frac{k_\text{eff}-1}{k_\text{eff}}.
+\end{gather}
+%
+Differences in reactivity, for either comparisons between neutronics methods or control rod worth
+calculations, can be calculated as
+%
+\begin{gather}
+  \Delta\rho_{1,2} = \rho_1 - \rho_2 =
+  \frac{k_{\text{eff},1}-k_{\text{eff},2}}{k_{\text{eff},1}k_{\text{eff},2}}.
+\end{gather}
+
+Spatially dependent variables (e.g., neutron group flux, neutron drift) are sampled at fixed 0.1 cm
+intervals along the specified lines (e.g., horizontal, diagonal, etc.) regardless of the actual
+mesh resolution.
+
+\section{1-D Models \& Results} \label{sec:1d-results}
+
+\subsection{1-D Model Setup \& Geometries}
 
@@
 
+\begin{figure}[htb!]
+  \centering
+  \includegraphics[width=0.8\columnwidth]{case-geometry}
+  \caption{Geometries of the 1-D test cases. The material labeled ``mixture'' represents a
+    homogeneous mixture of fuel and graphite at a ratio of 22.5\%-77.5\% by volume. All geometries
+    have reflective boundary conditions on the boundary at $x=0$ cm. The right-side boundaries are
+    reflective for Cases 1a and 1b, and vacuum for Cases 2a, 2b, 3a, and 3b.}
+  \label{fig:case-geom}
+\end{figure}
+
 Cases 1a and 1b are simple test cases containing neutron multiplying regions. Cases 1a and 1b serve
@@
 
+In my preliminary investigations, I discovered that using the original control rod material
+composition (70 wt\% Gd$_2$O$_3$-30 wt\% Al$_2$O$_3$) in the 1-D models led to excessively large
+control rod worth estimates of approximately 55000 pcm. The worth estimates are ten times larger
+than the 5500 pcm expected from full rod insertion of all three rods in the 3-D models
+\cite{fratoni_molten_2020}. Therefore, I greatly reduced the proportion of Gd$_2$O$_3$ in the
+control rod to 0.35 wt\% for a fairer assessment of rod worth in the 1-D models. This change
+brought the new rod worth estimates down to below 20000 pcm.
+
+\subsection{1-D Model Mesh Convergence Tests}
+
+I ran mesh convergence studies based on Case 3b for the neutron diffusion and $S_N$ neutron
+transport methods. Given that the hybrid $S_N$-diffusion method comprises of a combination of those
+methods, hybrid simulations are fully spatially resolved on meshes that are appropriately resolved
+for both neutron diffusion and $S_N$ simulations.
+
 \begin{figure}[htb!]
   \centering
-  \includegraphics[width=\columnwidth]{case-geometry}
-  \caption{Geometries of the 1-D test cases. The material labeled ``mixture'' represents a
-    homogeneous mixture of fuel and graphite at a ratio of 22.5\%-77.5\% by volume. All geometries
-    have reflective boundary conditions on the boundary at $x=0$ cm. The right-side boundaries are
-    reflective for Cases 1a and 1b, and vacuum for Cases 2a, 2b, 3a, and 3b.}
-  \label{fig:case-geom}
+  \begin{subfigure}[b]{0.48\columnwidth}
+    \centering
+    \includegraphics[width=\columnwidth]{diffusion-mesh-convergence-k}
+    \caption{Neutron diffusion method}
+    \label{fig:diff-mesh-k}
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}[b]{0.48\columnwidth}
+    \centering
+    \includegraphics[width=\columnwidth]{sn-mesh-convergence-k}
+    \caption{$S_8$ neutron transport method}
+    \label{fig:sn-mesh-k}
+  \end{subfigure}
+  \caption{Convergence of multiplication factor ($k_\text{eff}$) estimates for Case 3b across five
+    levels of mesh refinement. Higher refinement levels correspond to smaller mesh element sizes.}
+  \label{fig:mesh-k}
 \end{figure}
 
-\subsubsection{Mesh Convergence Tests}
+Figures \ref{fig:diff-mesh-k} and \ref{fig:sn-mesh-k} show the convergence of $k_\text{eff}$
+estimates observed for Case 3b simulations with neutron diffusion and $S_8$ neutron transport
+methods. The $k_\text{eff}$ values plateau at 0.9713203 and 0.9813790 for the neutron diffusion and
+$S_8$ methods, respectively. By mesh refinement level 2, both methods are within $5\times 10^{-7}$
+range of the $k_\text{eff}$ value at level 4.
+
+Figures \ref{fig:diff-mesh-flux} and \ref{fig:sn-mesh-flux} show the absolute differences in the
+group flux distributions at mesh refinement levels 1 to 3 relative to the reference group flux
+distribution at level 4 (maximum refinement). I omitted data from level 0 to improve the fidelity
+of the plots as flux differences drop by one order of magnitude with each level of refinement.
+With these results, I concluded that the mesh resolutaion at level 2 is a sufficient guideline for
+the mesh resolutions used in the subsequent 1-D analyses. The absolute differences in both
+$k_\text{eff}$ and group flux distributions at level 2 remain below $5\times10^{-7}$ relative to
+level 4.
+
+\begin{figure}[htb!]
+  \centering
+  \includegraphics[width=\columnwidth]{diffusion-mesh-convergence-flux}
+  \caption{Absolute difference in group fluxes ($\phi_g$) from the neutron diffusion method for
+  mesh refinement levels 1 to 4 relative to the reference group flux distributions at mesh
+  refinement level 5.}
+  \label{fig:diff-mesh-flux}
+\end{figure}
+
+\begin{figure}[htb!]
+  \centering
+  \includegraphics[width=\columnwidth]{sn-mesh-convergence-flux}
+  \caption{Absolute difference in group fluxes ($\phi_g$) from the $S_8$ neutron transport method
+  for mesh refinement levels 1 to 4 relative to the reference group flux distributions at mesh
+  refinement level 5.}
+  \label{fig:sn-mesh-flux}
+\end{figure}
+
+\FloatBarrier
+
+\subsection{1-D Numerical Results}
+
+\subsubsection{Reactivity}
+
+Figure \ref{fig:1d-rho} shows the difference in reactivities of OpenMC-MG, $S_8$,
+neutron diffusion, and hybrid methods relative to OpenMC-CE. The standard deviations of
+reactivity values from OpenMC-CE are approximately 40 pcm for Cases 1a and 1b and 60 pcm for the
+rest as depicted by the blue highlighted area in Figure \ref{fig:1d-rho}. For Cases 1a and 1b, the
+OpenMC-MG, $S_8$, and neutron diffusion methods show excellent agreement with OpenMC-CE as the
+reactivity values fall within the one or two standard deviation range. The Case 1 results show that
+the Moltres and Python implementations are consistent with each other given that they are spatially
+well-resolved.
 
@@
   \includegraphics[width=\columnwidth]{rho}
-  \caption{}
+  \caption{Difference in reactivity $\rho$ of all neutronics methods investigated relative
+  to OpenMC-CE.}
   \label{fig:1d-rho}
+\end{figure}
+
+For all cases, the OpenMC-MG and the $S_8$ methods show consistent agreement with one another. 
+While they deviate from OpenMC-CE by approximately 350 pcm for Cases 2a and 3a, they fall within
+two standard deviations of OpenMC-CE for Cases 2b and 3b. When increasing the maximum Legendre
+polynomial order to approximate the angular dependence in $\Sigma_s^{g'\rightarrow g}$ from
+2nd-order to 3rd-order, the reactivity changed by only 1 pcm. Therefore, the reactivity
+discrepancies are mainly attributed to the eight neutron energy group structure which remains the
+only significant difference between OpenMC-CE and the multigroup methods.
+
+The neutron diffusion and hybrid methods agree closely with one another for Cases 2a and 3a which
+exclude the control rod region. This shows that the hybrid method provides similar $k_\text{eff}$
+estimates to the neutron diffusion method in 1-D simulations without highly neutron-absorbing
+regions. Compared to the OpenMC-MG and $S_8$ reactivity values, the neutron diffusion and hybrid
+method reactivity values agree closer with the reference OpenMC-CE value. However, this is likely
+due to favorable error cancellation; Figures \ref{fig:3a-flux} and \ref{fig:3a-flux-diff} show that
+OpenMC-MG and the $S_8$ methods reproduce
+the OpenMC-CE flux distribution more accurately than the neutron diffusion and hybrid methods.
+
+In Cases 2b and 3b, the neutron diffusion method largely fails to accurately capture the
+effect of introducing the control rod region as evidenced by the -1500 pcm and -1150 pcm
+discrepancies relative to OpenMC-CE. The hybrid method fares better with -400 pcm and -300 pcm
+discrepancies.
+
+\begin{figure}[htb!]
+  \centering
+  \includegraphics[width=\columnwidth]{worth}
+  \caption{Percentage difference in rod worth for Cases 2 and 3 of all neutronics methods
+  investigated relative to OpenMC-CE.}
+  \label{fig:1d-worth}
+\end{figure}
+
+\subsubsection{Control rod worth}
+
+Moving the discussion to control rod worths, figure \ref{fig:1d-worth} shows the percentage
+difference in rod worths for Cases 2 and 3 of OpenMC-MG, $S_8$, neutron diffusion, and hybrid
+methods relative to OpenMC-CE. I calculated the rod worth estimates from each method by taking the
+difference in reactivity between the ``rod out'' (a) and ``rod in'' (b) configurations of Cases 2
+and 3. All neutronics methods overestimate the rod worth relative to OpenMC-CE, resulting in
+positive percentage difference values observed in the figure.
+The neutron diffusion methods clearly show up as outliers with rod worth estimates in excess
+of 8\%. Most of the remaining methods cluster around 2.5\% and 3\% percentage difference for Cases
+2 and 3 with the exception of the Python-hybrid method for Case 3.
+
+Given that the hybrid methods rely on transport corrections derived from the $S_N$ method, the
+$S_8$ rod worth estimates serve as the
+reference point for hybrid method verification. Additionally, all hybrid and $S_N$ simulations are
+equally affected by errors arising from the multigroup approximation. While the Python-hybrid
+method produces the lowest percentage difference relative to OpenMC-CE for Case 3, the
+Python-hybrid method deviates from the $S_8$ ``multigroup'' reference rod worths more than the
+Moltres-hybrid method. Nevertheless, both hybrid method implementations provide significant
+improvements in rod worth estimates over the neutron diffusion method.
+
+\subsubsection{Neutron flux distribution}
+
+This discussion on the neutron flux distributions focuses on Cases 3a and 3b which most closely
+resemble the actual \gls{MSRE} reactor geometry. Cases 3a and 3b resemble the \gls{MSRE} reactor
+with the control rod withdrawn and inserted, respectively. Figures \ref{fig:3a-flux} and
+\ref{fig:3b-flux} show the neutron group flux distributions for Cases 3a and 3b from OpenMC-CE and
+OpenMC-MG. Differences between the two methods arise from the multigroup approximation, notably in
+group 2 for Case 3a and groups 2, 6, 7, 8 for Case 3b. Future work aimed at finding a better
+neutron energy group structure should at minimum focus on improving fidelity in the listed energy
+groups.
+
+For a fair evaluation of the deterministic multigroup methods without distortions from the
+multigroup approximation, I chose to compare their flux distributions with the OpenMC-MG flux
+distribution. Figures \ref{fig:3a-flux-diff} and \ref{fig:3b-flux-diff} show the absolute
+difference in neutron group flux distributions of the Moltres-$S_8$, Moltres-diffusion,
+Moltres-hybrid, and Python-hybrid methods relative to OpenMC-MG. The figures omit Python-$S_8$ and
+Python-diffusion as their distributions are nearly identical to their Moltres-based counterparts.
+The neutron diffusion and hybrid methods fare worse than the $S_8$ method at capturing the
+oscillatory pattern in groups 1, 2, 5, 7, and 8 arising from the fuel-graphite lattice geometry
+along $x=5$ cm to 10 cm. In both Cases 3a and 3b, the neutron diffusion method (yellow plot lines)
+exhibits significant flux deviations in groups 1, 2, 7, and 8 near $x=0$ cm.
+
+Both hybrid methods
+provide significant improvements in neutron flux distributions compared to the neutron diffusion
+method as shown by the generally smaller flux difference values particularly near $x=0$ cm. The
+figures also show that Moltres-hybrid (green) provides more accurate flux estimates than
+Python-hybrid (red) in all instances. The differences indicate deficiencies in the diffusion
+correction scheme in Python-hybrid that lead to poorer performance compared to the drift
+correction scheme in Moltres-hybrid.
+
+\begin{figure}[htb!]
+  \centering
+  \includegraphics[width=\columnwidth]{case-3a-flux}
+  \caption{Case 3a neutron group flux distributions from OpenMC-CE and OpenMC-MG.}
+  \label{fig:3a-flux}
 \end{figure}
@@
   \centering
-  \includegraphics[width=\columnwidth]{rho}
-  \caption{}
-  \label{fig:1d-rho}
-\end{figure}
-
-\subsection{1-D Numerical Results}
-
-\begin{figure}[htb!]
-  \centering
-  \includegraphics[width=\columnwidth]{rho}
-  \caption{}
-  \label{fig:1d-rho}
+  \includegraphics[width=\columnwidth]{case-3b-flux}
+  \caption{Case 3b neutron group flux distributions from OpenMC-CE and OpenMC-MG.}
+  \label{fig:3b-flux}
 \end{figure}
@@
   \centering
-  \includegraphics[width=\columnwidth]{worth}
-  \caption{}
-  \label{fig:1d-worth}
+  \includegraphics[width=\columnwidth]{case-3a-flux-diff}
+  \caption{Absolute difference in neutron group flux distributions for Case 3a from Moltres-$S_8$,
+  Moltres-diffusion, Moltres-hybrid, and Python-hybrid relative to OpenMC-MG.}
+  \label{fig:3a-flux-diff}
 \end{figure}
 
-\section{2-D Models}
+\begin{figure}[htb!]
+  \centering
+  \includegraphics[width=\columnwidth]{case-3b-flux-diff}
+  \caption{Absolute difference in neutron group flux distributions for Case 3b from Moltres-$S_8$,
+  Moltres-diffusion, Moltres-hybrid, and Python-hybrid relative to OpenMC-MG.}
+  \label{fig:3b-flux-diff}
+\end{figure}
+
+\FloatBarrier
+
+\subsubsection{Drift \& diffusion correction distribution}
+
+Figures \ref{fig:3b-drift} and \ref{fig:3b-svdc} show the distributions of the drift and diffusion
+correction parameters, respectively, in each neutron energy group for Case 3b. As a reminder to the
+reader, the hybrid $S_N$-diffusion methods generate these correction parameters in subregions
+containing highly neutron absorbing materials. I implemented the drift correction scheme for
+Moltres-hybrid and the diffusion correction scheme for Python-hybrid. The hybrid methods
+selectively apply the $S_N$ method to improve neutronic modeling within the subregions. I computed
+the reference distributions (yellow data points) in both plots from reference flux solutions of the
+standard $S_8$ method.
+
+As discussed in Section \ref{sec:buffer-region}, both sets of correction parameters generated by
+the two hybrid methods match their respective reference values within the correction subregion from
+$x=0$ cm to 10 cm except near the subregion boundary at $x=10$ cm. Therefore, both hybrid methods
+provide improved flux estimates in the subregion compared to the neutron diffusion method.
+The prior reactivity and flux distribution results indicate that Moltres-hybrid is more accurate
+than Python-hybrid relative to OpenMC-MG as the reference.
+
+In Figure \ref{fig:3b-drift}, the drift
+correction distributions (Moltres-hybrid) for all eight energy groups are well-defined throughout
+the subregion and are continuous except at material interfaces. Consequently, the drift correction
+scheme provides accurate flux corrections for about 75\% of the subregion based on the buffer
+region cutoff criteria (refer to Section \ref{sec:buffer-region}). By contrast, the diffusion
+correction distributions in Figure \ref{fig:3b-svdc} contains numerous discontinuities tending towards
+positive or negative infinity in all eight energy groups. The discontinuities arise when
+computing corrected diffusion coefficients (Eq. \ref{eq:svdc}) near flux peaks and troughs
+(stationary points) where the flux gradients approach zero. The location of stationary points from
+the diffusion and $S_N$ calculations in the hybrid method do not necessarily coincide, resulting in
+discontinuous flux gradients in the final flux solution. Discontinuities are generally unavoidable
+in most problems because the fluxes typically reach local minima in control rod regions.
+Furthermore, unlike negative drift coefficients, negative diffusion coefficients are unphysical.
+Taking everything into consideration, the diffusion correction scheme is less accurate and more
+numerically unstable than the drift correction scheme.
+
+The subsequent discussions on the correction region size and the number of outer
+iterations focus on the drift correction scheme of the hybrid method.
+
+\begin{figure}[htb!]
+  \centering
+  \includegraphics[width=\columnwidth]{case-3b-drift}
+  \caption{Multigroup drift correction ($\vec{D}_g$) $x$-component distributions from the
+  Moltres-hybrid and Moltres-$S_8$ solvers.}
+  \label{fig:3b-drift}
+\end{figure}
+
+\begin{figure}[htb!]
+  \centering
+  \includegraphics[width=\columnwidth]{case-3b-svdc}
+  \caption{Multigroup diffusion correction ($D^s_g$) $x$-component distributions from the
+    Python-hybrid and Python-$S_8$ solvers.}
+  \label{fig:3b-svdc}
+\end{figure}
+
+\FloatBarrier
+
+\subsubsection{Impact of correction subregion sizes}
+
+The hybrid $S_N$-diffusion method relies on $S_N$ calculations in the correction subregion to
+generate flux corrections. Minimizing the size of the correction subregion and the $S_N$ subproblem
+is essential for the hybrid method to be computationally competitive for time-dependent full-core
+simulations. I investigated the effect of the correction subregion size on the $k_\text{eff}$ and
+control rod worth estimates with Cases 3a and 3b by varying the subregion sizes from 10 cm to 40 cm
+at 5 cm-intervals.
+
+Figures \ref{fig:v1-size-a-k} and \ref{fig:v1-size-b-k} show the $k_\text{eff}$ estimates from the
+hybrid method for Cases 3a and 3b, respectively. In both cases, the $k_\text{eff}$ values initially
+decrease as the correction subregion sizes increase before reversing in trend when the subregion
+size reaches 35 cm and beyond. The $k_\text{eff}$ values vary by up to 164 pcm for Case 3a and 109
+pcm for Case 3b. An important observation here is that the $k_\text{eff}$ values do not
+monotonically converge towards the $k_\text{eff}$ estimate from the $S_8$ method. The data implies
+that other significant sources of discrepancies exist beyond those being corrected in the
+correction subregion by the hybrid method.
+
+% In general,
+% discrepancies arise between the neutron diffusion method and $S_8$ from the neutron diffusion 
+% method's failure to accurately model the neutron flux in the fuel-graphite lattice region. The
+% figures imply that those discrepancies result in a positive change in $k_\text{eff}$ since the
+% $k_\text{eff}$ decreases as larger fractions of the problem domain benefit from flux corrections.
+% The remaining discrepancy in $k_\text{eff}$ between the hybrid method with a correction subregion
+% size of 40 cm and the $S_8$ method arises from
+
+\begin{figure}[htb!]
+  \centering
+  \begin{subfigure}[b]{0.49\columnwidth}
+    \centering
+    \includegraphics[width=\columnwidth]{correction-size-a-k}
+    \caption{Case 3a}
+    \label{fig:v1-size-a-k}
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}[b]{0.49\columnwidth}
+    \centering
+    \includegraphics[width=\columnwidth]{correction-size-b-k}
+    \caption{Case 3b}
+    \label{fig:v1-size-b-k}
+  \end{subfigure}
+  \caption{$k_\text{eff}$ estimates from the hybrid method for Cases 3a and 3b with different
+  correction subregion sizes. The horizontal lines indicate $k_\text{eff}$ estimates from the
+  OpenMC-CE, OpenMC-MG, and $S_8$ methods.}
+  \label{fig:v1-size-k}
+\end{figure}
+
+Figure \ref{fig:v1-size-rho} shows the percentage difference in rod worth relative to OpenMC-CE. 
+Due to the identical trends observed in the $k_\text{eff}$ estimates of both cases, the control rod
+worth estimates do not change significantly when the correction subregion
+sizes change. This indicates that limiting the correction subregion size to save
+computational cost on the expensive $S_N$ calculations has a negligible impact on rod worth
+estimates. The rod worth estimates for all investigated correction subregion sizes remain within
+0.2\% of the $S_8$ method.
+
+\begin{figure}[htb!]
+  \centering
+  \includegraphics[width=0.8\columnwidth]{correction-size-rho}
+  \caption{Percentage difference in rod worth from the hybrid method relative to OpenMC-CE for
+    Cases 3a and 3b with different correction subregion sizes. The horizontal lines indicate
+    equivalent rod worth differences from the OpenMC-MG and $S_8$ methods.}
+  \label{fig:v1-size-rho}
+\end{figure}
+
+The $k_\text{eff}$ and rod worth results show that the hybrid method is able to accurately
+reproduce rod worth estimates as long as the correction subregion size remains consistent. The
+correction subregion size must also be sufficiently large to capture the highly neutron-absorptive
+rod's influence on the neutron flux around the rod at $x=0$ cm. The drift correction parameter
+distributions in Figure \ref{fig:3b-drift} indicate that the influence of the rod on the drift, and
+consequently the flux, extend to approximately $x=5$ cm before the drift distribution settles on a
+regular repeating pattern influenced by the fuel-graphite lattice structure. I applied a consistent
+correction subregion size of 10 cm for Cases 2 and 3 since it sufficiently meets the requirements
+discussed while also minimizing the computational cost of the $S_8$ subsolver. Other reactor
+designs consisting of different material components may require different minimum correction
+subregion sizes.
+
+\subsubsection{Relaxing the $S_N$ subsolver convergence threshold}
+
+As described in Section \ref{sec:hybrid-algorithm}, the hybrid $S_N$-diffusion method works by
+iteratively coupling a neutron diffusion solver and a $S_N$ subsolver with the latter generating
+drift correction parameters for the former. Excluding the initial neutron diffusion calculation,
+each outer (fixed point) iteration comprises of an inner $S_N$ calculation on the correction
+subregion and an inner neutron diffusion calculation on the entire domain. The \gls{FEM} algorithm
+in Moltres determines that an individual solver reaches convergence when the combined
+residual contributions fall below a specified convergence tolerance threshold. The fixed point
+iteration scheme under the \texttt{MultiApps} system in Moltres for coupling multiple individual
+solvers (referred to as apps) requires users to specify a main app for the final convergence check
+after each outer iteration to terminate the iterations. The scheme also requires users to set a
+fixed point convergence tolerance value independent of the tolerance values of each individual
+solver. Naturally, the fixed point tolerance value must be equal to or greater than the tolerance
+value of the main individual solver for the possibility of overall convergence. The hybrid method
+implementation in Moltres sets the neutron diffusion solver as the main app and the $S_N$ subsolver
+as the sub app.
+
+The hybrid method shares some similarities with diffusion-based \gls{HOLO} acceleration schemes
+(refer to Section \ref{sec:drift-correction}) for the neutron transport methods which do not
+require the neutron transport side of the calculations to fully converge
+\cite{reynolds_analysis_2023, wang_diffusion_2014}.
+The typical convergence threshold value required for converged neutron flux calculations in Moltres
+is $10^{-8}$; I set the convergence tolerance value to $10^{-8}$ for the neutron diffusion solver
+and the fixed point scheme. I varied the convergence tolerance value for the $S_N$ subsolver to 
+find the optimal value that maintains sufficient solution accuracy and minimizes computational
+costs; tightening the $S_N$ convergence tolerance value generally results in more outer iterations
+performed before the solver meets the fixed point convergence tolerance.
+Table \ref{table:sn-tol} lists the number of outer iterations required by the hybrid method for
+a given set of $S_8$ subsolver convergence tolerance values in Case 3b.
+
+\begin{table}[htb]
+  \centering
+  \caption{Number of outer iterations in hybrid method calculations of Case 3b for a given set of
+  convergence tolerance values imposed on the $S_8$ subsolver.}
+  \begin{tabular}{l S S S S S S}
+    \toprule
+    $S_8$ subsolver tolerance, $\epsilon_\text{tol}$ & {$10^{-8}$} & {$10^{-7}$} & {$10^{-6}$} & {$10^{-5}$} & {$10^{-4}$} & {$10^{-3}$} \\
+    \midrule
+    Number of outer iterations & 3 & 3 & 3 & 2 & 2 & 1 \\
+    \bottomrule
+  \end{tabular}
+  \label{table:sn-tol}
+\end{table}
+
+Figure \ref{fig:sn-tol} shows the $k_\text{eff}$ error estimates relative to the reference value
+when $\epsilon_\text{tol}=10^{-8}$. The hybrid method exhibits superlinear convergence ($q=1.333$)
+with respect to the $S_N$ subsolver convergence tolerance value. Based on the results, setting
+$\epsilon_\text{tol}$ to $10^{-5}$ is the best balance between accuracy and computational
+cost because the $k_\text{eff}$ error is less than 0.01 pcm and the number of outer iterations
+would increase from 2 to 3 if the tolerance value is further tightened.
+
+\begin{figure}[htb!]
+  \centering
+  \includegraphics[width=0.6\columnwidth]{sn-tol}
+  \caption{$k_\text{eff}$ error estimates of Case 3b for a range of convergence tolerance values
+  imposed on the $S_8$ subsolver relative to the reference $k_\text{eff}$ value when
+  $\epsilon_\text{tol}=10^{-8}$. The $q=1.333$ line represents the approximate rate of
+  convergence.}
+  \label{fig:sn-tol}
+\end{figure}
+
+\section{2-D Models \& Results}
 
@@
 
+Figures \ref{fig:1/4-geom} and \ref{fig:full-geom} show the 2-D quarter-core and full-core
+\gls{MSRE} models investigated. They are similar to the horizontal cross section of the base
+\gls{MSRE} benchmark model developed by \cite{fratoni_molten_2020} with the following differences
+in the 2-D \gls{MSRE} model for this work:
+
+\begin{itemize}
+  \item Ignored thermal expansion of all reactor components.
+  \item Modeled the control rods as perfectly annular cylinders instead of strings of beaded
+    control elements. Ignored the inconel shells encasing the control elements and the steel
+    helical cables linking the control elements together.
+  \item Set channel pitch, width, and length to 2, 0.2, and 1.2 inches, respectively.
+  \item Homogenized the INOR, graphite, and molten salt regions in the sample basket.
+  \item Modified the outer edge of the core graphite to be circular with no jagged edges.
+  \item Replaced partial channels with full-size channels near the edge of the core graphite.
+\end{itemize}
+
+The differences listed above, along with other changes in the 3-D model, result in negligible
+changes to the $k\text{eff}$ (See Section 5.6) relative to the \gls{MSRE} benchmark model
+\cite{fratoni_molten_2020}.
+
+\begin{figure}[htb!]
+  \centering
+  \includegraphics[width=\columnwidth]{quarter-core-geom}
+  \caption{2-D \gls{MSRE} quarter-core model based on the horizontal cross section of the actual
+  \gls{MSRE} geometry.}
+  \label{fig:1/4-geom}
+\end{figure}
+
+\begin{figure}[htb!]
+  \centering
+  \includegraphics[width=\columnwidth]{full-core-geom}
+  \caption{2-D \gls{MSRE} full-core model based on the horizontal cross section of the actual
+  \gls{MSRE} geometry.}
+  \label{fig:full-geom}
+\end{figure}
+
+\begin{figure}[htb!]
+  \centering
+  \includegraphics[width=0.7\columnwidth]{full-core-closeup}
+  \caption{Detailed view of the control rod thimbles and sample basket in 2-D \gls{MSRE} full-core
+    model.}
+  \label{fig:full-geom-closeup}
+\end{figure}
+
+\FloatBarrier
+
 \subsection{2-D Numerical Results}
 
-\section{3-D Models}
+\subsubsection{2-D quarter-core results}
 
-\subsection{3-D Model Setup}
+\begin{table}[htb]
+  \small
+  \centering
+  \caption{$k_\text{eff}$ and control rod worth estimates for the 2-D quarter-core \gls{MSRE}
+    model. Error estimates are relative to OpenMC-CE.}
+  \begin{tabular}{l S[table-format=1.5(2)] S S[table-format=1.5(2)] S S[table-format=4(2)] S}
+    \toprule
+    \multirow{2}{*}{Method} & \multicolumn{2}{c}{No Rod} & \multicolumn{2}{c}{Rod} & \multicolumn{2}{c}{Rod worth} \\
+                            & {$k_\text{eff}$} & {Error [pcm]} & {$k_\text{eff}$} & {Error [pcm]} & {$\Delta\rho_\text{worth}$ [pcm]} & {Error [pcm]} \\
+                            \cmidrule(r){1-1} \cmidrule(rl){2-3} \cmidrule(rl){4-5} \cmidrule(l){6-7}
+	  OpenMC-CE & 1.11209(43) & {-} & 1.01740(42) & {-} & 8370(53) & {-} \\
+	  OpenMC-MG & 1.11979(42) & 618 & 1.02204(41) & 446 & 8541(51) & 172 \\
+      Hybrid & 1.12174 & 773 & 1.02532 & 760 & 8383 & 13 \\
+      Diffusion & 1.12059 & 682 & 1.00903 & -816 & 9867 & -1484 \\
+    \bottomrule
+  \end{tabular}
+  \label{table:quarter-core}
+\end{table}
 
-\subsection{3-D Numerical Results}
+\begin{figure}[htb!]
+  \centering
+  \includegraphics[width=\columnwidth]{msre-quarter-no-rod-power}
+  \caption{Normalized channel fission rate distribution of the 2-D \gls{MSRE} quarter-core model.
+  This figure depicts the 1/8th-core distribution due to symmetry about the diagonal.}
+  \label{fig:1/4-no-rod}
+\end{figure}
+
+\begin{figure}[htb!]
+  \centering
+  \includegraphics[width=\columnwidth]{msre-quarter-rod-power}
+  \caption{Normalized channel fission rate distribution of the 2-D \gls{MSRE} quarter-core model.
+  This figure depicts the 1/8th-core distribution due to symmetry about the diagonal.}
+  \label{fig:1/4-rod}
+\end{figure}
+
+\begin{table}[htb]
+  \small
+  \centering
+  \caption{Absolute mean and maximum percentage errors in the normalized channel fission rates of
+  the 2-D \gls{MSRE} quarter-core models relative to OpenMC. The average standard deviation of
+  OpenMC normalized channel fission rates is 0.2\%.}
+  \begin{tabular}{l S S S S}
+    \toprule
+    \multirow{2}{*}{Method} & \multicolumn{2}{c}{No Rod} & \multicolumn{2}{c}{Rod} \\
+                            & {Mean [\%]} & {Maximum [\%]} & {Mean [\%]} & {Maximum [\%]} \\
+                            \cmidrule(r){1-1} \cmidrule(rl){2-3} \cmidrule(l){4-5}
+    Hybrid & 0.40 & 1.32 & 0.43 & 3.08 \\
+    Diffusion & 0.40 & 2.63 & 2.01 & 17.44 \\
+    \bottomrule
+  \end{tabular}
+  \label{table:quarter-core-power}
+\end{table}
+
+\FloatBarrier
+
+\subsubsection{2-D full-core results}
+
+\begin{table}[htb]
+  \small
+  \centering
+  \caption{$k_\text{eff}$ estimates for the 2-D full-core \gls{MSRE} model with the indicated rods
+    inserted. Error estimates are relative to OpenMC-CE.}
+  \setlength\tabcolsep{2pt}
+  \begin{tabular}{l S[table-format=1.5(2)] S S[table-format=1.5(2)] S S[table-format=1.5(2)] S S[table-format=1.5(2)] S}
+    \toprule
+    \multirow{3}{*}{Method} & \multicolumn{2}{c}{No Rod} & \multicolumn{2}{c}{Rod 1} & \multicolumn{2}{c}{Rod 1 \& 2} & \multicolumn{2}{c}{Rod 1, 2 \& 3} \\
+                            & {\multirow{2}{*}{$k_\text{eff}$}} & {Error} & {\multirow{2}{*}{$k_\text{eff}$}} & {Error} & {\multirow{2}{*}{$k_\text{eff}$}} & {Error} & {\multirow{2}{*}{$k_\text{eff}$}} & {Error} \\
+                            & & {[pcm]} & & {[pcm]} & & {[pcm]} & & {[pcm]} \\
+                            \cmidrule(r){1-1} \cmidrule(rl){2-3} \cmidrule(rl){4-5} \cmidrule(rl){6-7} \cmidrule(l){8-9}
+    OpenMC-CE & 1.09859(20) & {-} & 1.06980(21) & {-} & 1.04690(18) & {-} & 1.02688(18) & {-} \\
+    OpenMC-MG & 1.10637(19) & 640 & 1.07633(20) & 567 & 1.05235(20) & 495 & 1.03262(17) & 541 \\
+    Hybrid & 1.10843 & 808 & 1.07906 & 802 & 1.05553 & 781 & 1.03582 & 841 \\
+    Diffusion & 1.10701 & 692 & 1.07120 & 122 & 1.04415 & -252 & 1.02195 & -470 \\
+    \bottomrule
+  \end{tabular}
+  \label{table:full-core-k}
+\end{table}
+
+\begin{table}[htb]
+  \small
+  \centering
+  \caption{Control rod worth estimates for the 2-D full-core \gls{MSRE} with the
+  indicated rods inserted. Error estimates are relative to OpenMC-CE.}
+  \setlength\tabcolsep{5pt}
+  \begin{tabular}{l S[table-format=4(2)] S S[table-format=4(2)] S S[table-format=4(2)] S}
+    \toprule
+    \multirow{2}{*}{Method} & \multicolumn{2}{c}{Rod 1} & \multicolumn{2}{c}{Rod 1 \& 2} & \multicolumn{2}{c}{Rod 1, 2 \& 3} \\
+                            & {$\Delta\rho_\text{worth}$ [pcm]} & {Error [pcm]} & {$\Delta\rho_\text{worth}$ [pcm]} & {Error [pcm]} & {$\Delta\rho_\text{worth}$ [pcm]} & {Error [pcm]} \\
+                            \cmidrule(r){1-1} \cmidrule(rl){2-3} \cmidrule(rl){4-5} \cmidrule(l){6-7}
+    OpenMC-CE & 2450(25) & {-} & 4494(23) & {-} & 6357(24) & {-} \\
+    OpenMC-MG & 2523(23) & 73 & 4640(24) & 146 & 6455(22) & 98 \\
+    Hybrid & 2455 & 5 & 4521 & 27 & 6323 & -34 \\
+    Diffusion & 3019 & 569 & 5439 & 945 & 7519 & 1162 \\
+    \bottomrule
+  \end{tabular}
+  \label{table:full-core-worth}
+\end{table}
+
+\begin{table}[htb]
+  \footnotesize
+  \centering
+  \caption{Absolute mean and maximum percentage errors in the normalized channel fission rates of
+  the 2-D \gls{MSRE} full-core models relative to OpenMC. The average standard deviation of
+  OpenMC normalized channel fission rates is 0.27\%.}
+  \setlength\tabcolsep{2.5pt}
+  \begin{tabular}{l S S S S S S S S}
+    \toprule
+    \multirow{2}{*}{Method} & \multicolumn{2}{c}{No Rod} & \multicolumn{2}{c}{Rod 1} & \multicolumn{2}{c}{Rod 1 \& 2} & \multicolumn{2}{c}{Rod 1, 2 \& 3} \\
+                            & {Mean [\%]} & {Maximum [\%]} & {Mean [\%]} & {Maximum [\%]} & {Mean [\%]} & {Maximum [\%]} & {Mean [\%]} & {Maximum [\%]} \\
+                            \cmidrule(r){1-1} \cmidrule(rl){2-3} \cmidrule(rl){4-5} \cmidrule(rl){6-7} \cmidrule(l){8-9}
+    Hybrid & 0.43 & 1.45 & 0.43 & 1.82 & 0.43 & 2.26 & 0.43 & 2.52 \\
+    Diffusion & 0.45 & 2.95 & 0.94 & 12.61 & 1.35 & 15.34 & 1.67 & 17.09 \\
+    \bottomrule
+  \end{tabular}
+  \label{table:full-core-power}
+\end{table}
+
+\FloatBarrier
+
+\section{3-D Models \& Results}
+
+%\subsection{3-D Model Setup}
+
+%\subsection{3-D Numerical Results}
+
+\section{Summary}
+
+The hybrid method does not alter the $k_\text{eff}$ and flux estimates relative to the neutron
+diffusion method in cases where the neutron diffusion method provides

### 2024-08-23 msre.tex
chapter 5 draft

@@
-\section{Group Constant Data Generation on OpenMC}
+In Chapter \ref{chap:hybrid}, I laid out the mathematical background and numerical implementation
+of the hybrid $S_N$-diffusion method for improving neutronic accuracy in reactor systems containing
+highly neutron-absorbing regions. I developed and implemented the diffusion and drift correction
+schemes for the hybrid method on Python and Moltres, respectively. This chapter demonstrates
+the hybrid method through 1-D, 2-D, and 3-D $k$-eigenvalue simulations of the \gls{MSRE}.
+I employed the OpenMC Monte Carlo neutral particle transport software \cite{romano_openmc:_2015} to
+generate group constants for the hybrid method and reference solutions for method verification.
+The 1-D models serve to verify the newly-implemented $S_N$ and hybrid methods
+against OpenMC as well as to analyze the impact of key input parameters on the hybrid method. The
+2-D models test the hybrid method performance through increased spatial dimensionality and reduced
+geometric asymmetry relative to the 1-D models. Lastly, the 3-D models serve as neutronics model
+validation against the \gls{MSRE} benchmark model by Fratoni et al.\ \cite{fratoni_molten_2020}, a
+testbed for \gls{HPC} scaling tests, and preparation for the time-dependent multiphysics
+simulations of \gls{MSRE} reactivity insertion transients.
 
-The group constants required by the neutron diffusion or $S_N$ neutron transport
+Section \ref{sec:msre-gc} details the OpenMC model setup for group constant generation. Section
+\ref{sec:nts-methods} describes and labels the different neutronics methods employed for the
+$k$-eigenvalue simulations. Section \ref{sec:test-metrics} details basic test metrics relevant to
+neutronics verification and validation. Sections \ref{sec:1d-results}, \ref{sec:2d-results}, and
+\ref{sec:3d-results} discuss the model setup, results, and discussions for the 1-D, 2-D, and 3-D
+\gls{MSRE} models, respectively.
+
+\section{Group Constant Data Generation on OpenMC} \label{sec:msre-gc}
+
+Group constants relevant to the neutron diffusion, $S_N$ neutron transport, or both
 methods are:
@@
 I generated the group constants using OpenMC's multigroup cross section generation capability
-\cite{boyd_multigroup_2019}. The group constant postprocessing script for Moltres
-(\texttt{moltres\_xs.py}) calculates $\Sigma_{r,g}$ as
+\cite{boyd_multigroup_2019}. A group constant postprocessing script for Moltres
+(\texttt{moltres\_xs.py}) parses OpenMC simulation output files and rewrites the group constant
+data into a JSON format file suitable for Moltres simulations. While OpenMC provides most of the
+group constants directly, the postprocessing script calculates $\Sigma_{r,g}$ as
 %
@@
 absorption. $\Sigma_{sp}^{g\rightarrow g}$ incorporates neutron multiplication effects from neutron
-knockout reactions into the scattering cross section. Neutron knockout reactions are commonly
-tallied as scattering reactions, but $\Sigma_{r,g}$ is a convenient parameter in which to
-incorporate neutron knockout effects in the neutron diffusion equations. I ran all test cases for
-$S_N$ and hybrid method calculations with up to the 2nd Legendre moments of the scattering
-cross sections ($L=2$).
+knockout reactions into the scattering cross section. Neutronics codes commonly identify neutron
+knockout reactions as scattering reactions, but $\Sigma_{r,g}$ is a convenient parameter in which
+to incorporate neutron knockout effects in the neutron diffusion method. I ran all test cases
+for $S_N$ and hybrid method calculations with scattering cross sections approximated by up to 2nd
+Legendre moments ($L=2$).
 
-OpenMC uses the $P_1$ flux-limited formulation \cite{pomraning_flux-limited_1984} for calculating
-$D_g$ as follows
+OpenMC uses the $P_1$ flux-limited formulation \cite{pomraning_flux-limited_1984} by default for
+calculating $D_g$ as follows
 %
@@
 
-\section{Neutronics Methods}
+\section{Neutronics Methods} \label{sec:nts-methods}
 
 I ran all cases on OpenMC to generate the required group constants and reference $k_\text{eff}$
-values to assess the accuracy of the deterministic methods against OpenMC in continuous-energy
-(OpenMC-CE) and multigroup (OpenMC-MG) modes. Table \ref{table:var} shows how the
-position, direction of travel, neutron energy, and angle-dependence in $\Sigma_s$ are represented
-by OpenMC and the deterministic methods. The hybrid $S_N$-diffusion method features representations
-from either $S_N$ or neutron diffusion depending on the subregion under consideration.
-Comparing OpenMC-MG results with OpenMC-CE results allows us
+values to assess the accuracy of the deterministic methods ($S_N$, neutron diffusion, hybrid 
+methods) against OpenMC in continuous-energy (OpenMC-CE) and multigroup (OpenMC-MG) modes. Table
+\ref{table:var} details how OpenMC and the deterministic methods represent the position, direction
+of travel, neutron energy, and angle-dependence in $\Sigma_s$. The hybrid $S_N$-diffusion method
+features representations from either $S_N$ or neutron diffusion depending on the subregion under
+consideration. Comparing OpenMC-MG results with OpenMC-CE results allows us
 to quantify errors arising from neutron energy group discretization and the scattering cross
@@
 Moltres follows the methodology for solving the \gls{SAAF} formulation of the $S_N$ equations as
-described in Section \label{sec:saaf} while the implementation on Python follows the methodology
-for solving the original first-order formulation as described in Section \label{sec:python-sn}.
+described in Section \ref{sec:saaf} while the implementation on Python follows the methodology
+for solving the original first-order formulation as described in Section \ref{sec:python-sn}.
 
-The implementations on Moltres are discretized in space using the finite element method
-while the implementations on Python use finite difference or diamond difference schemes. The most
-crucial difference lies in the transport correction formulations for the hybrid method. The Moltres
-implementation is based on the drift correction formulation described in Section
-\label{sec:drift-correction} while the Python implementation is based on the diffusion correction
-formulation described in Section \label{sec:diffusion-correction}.
+The implementations on Moltres discretize the problem domain in space using the finite element
+method while the implementations on Python use finite difference or diamond difference schemes. The
+most crucial difference lies in the transport correction formulations for the hybrid method. The
+Moltres implementation uses the drift correction scheme described in Section
+\ref{sec:drift-correction} while the Python implementation uses the diffusion correction
+scheme described in Section \ref{sec:diffusion-correction}.
 
-I will refer to the methods and results from either Molres or Python implementations with the
-Moltres- or Python- prefix to clearly set them apart, i.e. Moltres-diffusion, Moltres-$S_N$,
-Moltres-hybrid, Python-diffusion, Python-$S_N$, and Python-hybrid.
+In this chapter, I refer to the methods and results from the Moltres and Python implementations
+with the Moltres- or Python- prefix, i.e., Moltres-$S_N$, Moltres-diffusion, Moltres-hybrid,
+Python-$S_N$, Python-diffusion, and Python-hybrid.
 
 For 2- and 3-D models, I abandoned the diffusion correction formulation in favor of the drift
-correction formulation due to superior numerical performance in the latter. The analysis
-supporting this decision may be found in the analysis of 1-D results in Section
-\ref{sec:1d-results}.
+correction formulation due to superior numerical performance in the latter. Readers may find the
+analysis supporting this decision in the analysis of 1-D results in Section \ref{sec:1d-results}.
 
@@
 
-\section{Test Metrics} \label{sec:test-metrics}
+\section{Basic Test Metrics} \label{sec:test-metrics}
 
@@
 
-\section{1-D Models \& Results} \label{sec:1d-results}
+\section{1-D Neutronics Models \& Results} \label{sec:1d-results}
 
-\subsection{1-D Model Setup \& Geometries}
+\subsection{1-D Neutronics Model Setup \& Geometries}
 
@@
 and air-filled rod guide tubes. Figure \ref{fig:case-geom} shows the geometries of Cases 1a to
-3b. All geometries have reflective boundary conditions at $x=0$ cm to reduce computational costs by
-creating half-core or repeating unit cell models. Cases 1a and 1b are repeating unit
+3b. All geometries have reflective boundary conditions at $x=0$ cm, thereby
+forming half-core or repeating unit cell models. Cases 1a and 1b are repeating unit
 cell models with reflecting boundaries on the right-side boundaries. Cases 2a, 2b, 3a, and 3b
@@
 
-In my preliminary investigations, I discovered that using the original control rod material
-composition (70 wt\% Gd$_2$O$_3$-30 wt\% Al$_2$O$_3$) in the 1-D models led to excessively large
-control rod worth estimates of approximately 55000 pcm. The worth estimates are ten times larger
-than the 5500 pcm expected from full rod insertion of all three rods in the 3-D models
+In my preliminary investigations, the original control rod material composition
+(70 wt\% Gd$_2$O$_3$-30 wt\% Al$_2$O$_3$) led to excessively large control rod worth estimates of
+approximately 55000 pcm in the 1-D models, effectively halving the $k_\text{eff}$. The worth
+estimates are ten times larger
+than the 5500 pcm expected from full rod insertion of all three rods in the actual \gls{MSRE}
 \cite{fratoni_molten_2020}. Therefore, I greatly reduced the proportion of Gd$_2$O$_3$ in the
@@
 
-\subsection{1-D Model Mesh Convergence Tests}
+\subsection{1-D Neutronics Model Mesh Convergence Tests}
 
@@
 transport methods. Given that the hybrid $S_N$-diffusion method comprises of a combination of those
-methods, hybrid simulations are fully spatially resolved on meshes that are appropriately resolved
-for both neutron diffusion and $S_N$ simulations.
+methods, I found hybrid simulations were fully spatially resolved on meshes that are
+appropriately resolved for both neutron diffusion and $S_N$ simulations.
 
@@
 methods. The $k_\text{eff}$ values plateau at 0.9713203 and 0.9813790 for the neutron diffusion and
-$S_8$ methods, respectively. By mesh refinement level 2, both methods are within $5\times 10^{-7}$
+$S_8$ methods, respectively. By mesh refinement level 2, both methods fall within $5\times 10^{-7}$
 range of the $k_\text{eff}$ value at level 4.
@@
 distribution at level 4 (maximum refinement). I omitted data from level 0 to improve the fidelity
-of the plots as flux differences drop by one order of magnitude with each level of refinement.
-With these results, I concluded that the mesh resolutaion at level 2 is a sufficient guideline for
+of the plots as flux differences drop significantly by one order of magnitude with each level of
+refinement.
+With these results, I concluded that the mesh resolution at level 2 is a sufficient guideline for
 the mesh resolutions used in the subsequent 1-D analyses. The absolute differences in both
@@
 
-\subsection{1-D Numerical Results}
+\subsection{1-D Neutronics Numerical Results}
 
@@
 OpenMC-MG, $S_8$, and neutron diffusion methods show excellent agreement with OpenMC-CE as the
-reactivity values fall within the one or two standard deviation range. The Case 1 results show that
+reactivity values fall within the one or two standard deviation range. The Case 1 results show
 the Moltres and Python implementations are consistent with each other given that they are spatially
@@
 polynomial order to approximate the angular dependence in $\Sigma_s^{g'\rightarrow g}$ from
-2nd-order to 3rd-order, the reactivity changed by only 1 pcm. Therefore, the reactivity
-discrepancies are mainly attributed to the eight neutron energy group structure which remains the
+2nd-order to 3rd-order, the reactivity changed by only 1 pcm. Therefore, I attribute the reactivity
+discrepancies to the eight neutron energy group structure which remains the
 only significant difference between OpenMC-CE and the multigroup methods.
@@
 method reactivity values agree closer with the reference OpenMC-CE value. However, this is likely
-due to favorable error cancellation; Figures \ref{fig:3a-flux} and \ref{fig:3a-flux-diff} show that
+due to favorable error cancellation; Figures \ref{fig:3a-flux} and \ref{fig:3a-flux-diff} show
 OpenMC-MG and the $S_8$ methods reproduce
@@
 $S_8$ rod worth estimates serve as the
-reference point for hybrid method verification. Additionally, all hybrid and $S_N$ simulations are
-equally affected by errors arising from the multigroup approximation. While the Python-hybrid
+reference point for hybrid method verification. Additionally, errors arising from the multigroup
+approximation affect both hybrid and $S_N$ simulations to similar degrees. While the Python-hybrid
 method produces the lowest percentage difference relative to OpenMC-CE for Case 3, the
@@
 resemble the actual \gls{MSRE} reactor geometry. Cases 3a and 3b resemble the \gls{MSRE} reactor
-with the control rod withdrawn and inserted, respectively. Figures \ref{fig:3a-flux} and
+with a control rod withdrawn and inserted, respectively. Figures \ref{fig:3a-flux} and
 \ref{fig:3b-flux} show the neutron group flux distributions for Cases 3a and 3b from OpenMC-CE and
@@
 method as shown by the generally smaller flux difference values particularly near $x=0$ cm. The
-figures also show that Moltres-hybrid (green) provides more accurate flux estimates than
+figures also show Moltres-hybrid (green) provides more accurate flux estimates than
 Python-hybrid (red) in all instances. The differences indicate deficiencies in the diffusion
-correction scheme in Python-hybrid that lead to poorer performance compared to the drift
+correction scheme in Python-hybrid which lead to poorer performance compared to the drift
 correction scheme in Moltres-hybrid.
@@
 As discussed in Section \ref{sec:buffer-region}, both sets of correction parameters generated by
-the two hybrid methods match their respective reference values within the correction subregion from
-$x=0$ cm to 10 cm except near the subregion boundary at $x=10$ cm. Therefore, both hybrid methods
+the two correction schemes match their respective reference values within the correction subregion
+from $x=0$ cm to 10 cm except near the subregion boundary at $x=10$ cm. Therefore, both schemes
 provide improved flux estimates in the subregion compared to the neutron diffusion method.
-The prior reactivity and flux distribution results indicate that Moltres-hybrid is more accurate
+The prior reactivity and flux distribution results indicate Moltres-hybrid is more accurate
 than Python-hybrid relative to OpenMC-MG as the reference.
@@
 scheme provides accurate flux corrections for about 75\% of the subregion based on the buffer
-region cutoff criteria (refer to Section \ref{sec:buffer-region}). By contrast, the diffusion
-correction distributions in Figure \ref{fig:3b-svdc} contains numerous discontinuities tending towards
+region cutoff criteria which occurs around $x=7.5$ cm where most of the drift distributions are
+zero (refer to Section \ref{sec:buffer-region}). By contrast, the diffusion correction
+distributions in Figure \ref{fig:3b-svdc} contains numerous discontinuities tending towards
 positive or negative infinity in all eight energy groups. The discontinuities arise when
@@
 monotonically converge towards the $k_\text{eff}$ estimate from the $S_8$ method. The data implies
-that other significant sources of discrepancies exist beyond those being corrected in the
+other significant sources of discrepancies exist beyond those being corrected in the
 correction subregion by the hybrid method.
@@
 worth estimates do not change significantly when the correction subregion
-sizes change. This indicates that limiting the correction subregion size to save
+sizes change. This indicates limiting the correction subregion size to save
 computational cost on the expensive $S_N$ calculations has a negligible impact on rod worth
@@
 
-The $k_\text{eff}$ and rod worth results show that the hybrid method is able to accurately
+The $k_\text{eff}$ and rod worth results show the hybrid method is able to accurately
 reproduce rod worth estimates as long as the correction subregion size remains consistent. The
@@
 subregion and an inner neutron diffusion calculation on the entire domain. The \gls{FEM} algorithm
-in Moltres determines that an individual solver reaches convergence when the combined
+in Moltres determines an individual solver reaches convergence when the combined
 residual contributions fall below a specified convergence tolerance threshold. The fixed point
@@
 and the fixed point scheme. I varied the convergence tolerance value for the $S_N$ subsolver to 
-find the optimal value that maintains sufficient solution accuracy and minimizes computational
+find the optimal value for maintaining sufficient solution accuracy and minimizing computational
 costs; tightening the $S_N$ convergence tolerance value generally results in more outer iterations
@@
 
-\section{2-D Models \& Results}
+\section{2-D Neutronics Models \& Results} \label{sec:2d-results}
 
-\subsection{2-D Model Setup}
+\subsection{2-D Neutronics Model Setup}
 
 Figures \ref{fig:1/4-geom} and \ref{fig:full-geom} show the 2-D quarter-core and full-core
-\gls{MSRE} models investigated. They are similar to the horizontal cross section of the base
+\gls{MSRE} models investigated. I adapted the models from the horizontal cross section of the base
 \gls{MSRE} benchmark model developed by \cite{fratoni_molten_2020} with the following differences
@@
   \item Ignored thermal expansion of all reactor components.
-  \item Modeled the control rods as perfectly annular cylinders instead of strings of beaded
-    control elements. Ignored the inconel shells encasing the control elements and the steel
-    helical cables linking the control elements together.
+  \item Modeled the control rods as perfectly annular cylinders instead of beaded control elements
+    strung on flexible steel cables. Ignored the inconel shells encasing the control elements and
+    the steel helical cables linking the control elements together.
   \item Set channel pitch, width, and length to 2, 0.2, and 1.2 inches, respectively.
@@
   \item Replaced partial channels with full-size channels near the edge of the core graphite.
+  \item Ignored the thermal insulation and shield structures outside the core.
 \end{itemize}
@@
 The differences listed above, along with other changes in the 3-D model, result in negligible
-changes to the $k\text{eff}$ (See Section 5.6) relative to the \gls{MSRE} benchmark model
+changes to the $k_\text{eff}$ (refer to Section 5.6) relative to the \gls{MSRE} benchmark model
 \cite{fratoni_molten_2020}.
@@
 
+\begin{table}[htb]
+  \small
+  \centering
+  \setlength\tabcolsep{4pt}
+  \caption{\gls{MSRE} molten salt composition when the $^{235}$U loading was at 65.25 kg.}
+  \begin{tabular}{l S S S S S S S}
+    \toprule
+    Element & {Li} & {Be} & {Zr} & {Hf} & {$^{234}$U} & {$^{235}$U} & {$^{236}$U} \\
+    \midrule
+    Composition [kg] & 507.27 & 293.96 & 513.97 & 0.0029 & 0.67 & 65.25 & 0.27 \\
+    \bottomrule
+  \end{tabular}
+  \begin{tabular}{l S S S S S S}
+    \toprule
+    Element & {$^{238}$U} & {Fe} & {Cr} & {Ni} & {O} & {F} \\
+    \midrule
+    Composition [kg] & 141.91 & 0.75 & 0.13 & 0.14 & 2.27 & 3103.22 \\
+    \bottomrule
+  \end{tabular}
+  \label{table:2d-salt-composition}
+\end{table}
+
+INOR-8, also referred to as Hastelloy N, is a nickel-based alloy used for structural components in
+the \gls{MSRE} \cite{robertson_msre_1965}.
+Table \ref{table:2d-salt-composition} lists the molten salt composition for the 2-D models. The
+$^{235}$U loading of 65.25 kg is the amount at which the \gls{MSRE} first achieved criticality. The
+control rods consist of the original \gls{MSRE} 70 wt\% Gd$_2$O$_3$-30 wt\% Al$_2$O$_3$ mixture.
+The models impose neutron vacuum boundary conditions on the outermost boundary of the INOR-8
+vessel structure. The quarter-core model imposes neutron reflecting boundary conditions along the
+left and bottom straight edges (Figure \ref{fig:1/4-geom}). Note that the quarter-core and
+full-core models are not identical because the full-core model contains a sample basket instead of
+a control rod in the lower-right segment of Figure \ref{fig:full-geom-closeup}.
+
 \begin{figure}[htb!]
@@
 
+I ran neutronics simulations on the quarter- and full-core models on OpenMC under both continuous
+energy (OpenMC-CE) and multigroup (OpenMC-MG) modes, and on Moltres with the hybrid $S_N$-diffusion
+and neutron diffusion methods. The hybrid method ran with the $S_8$ angular discretization scheme
+and up to 2nd-order scattering approximations for the $S_N$ subsolver. I simulated the
+quarter-core model with and without the control rod in the air-filled rod thimble; air replaces the
+control rod material when the rod is removed. For the full-core model, I considered four rod
+configurations: no rods present, rod 1 present, rods 1 \& 2 present, and rods 1, 2 \& present.
+Other than providing a close-up view of the core center, Figure \ref{fig:full-geom-closeup} also
+presents the correction subregion on which the $S_8$ subsolver of the hybrid method generates
+drift correction parameters for the coupled diffusion solver.
+
 \FloatBarrier
 
-\subsection{2-D Numerical Results}
+\subsection{2-D Neutronics Numerical Results} \label{sec:2d-nts-results}
 
 \subsubsection{2-D quarter-core results}
+
+Table \ref{table:quarter-core} details the $k_\text{eff}$ and control rod worth estimates and error
+values relative to OpenMC-CE. For the ``No Rod'' case, OpenMC-CE estimated the $k_\text{eff}$ to be
+1.11209(43). The OpenMC-MG, hybrid, and neutron diffusion methods performed
+similarly to one another with reactivity errors ranging from 618 pcm to 773 pcm relative to
+OpenMC-CE. Based on $k_\text{eff}$ alone, the hybrid method performed marginally worse than the
+neutron diffusion method. However, for the ``Rod'' case, then neutron diffusion method saw a
+change in sign of the error to -816 pcm. The errors for OpenMC-MG and the hybrid method
+remained positive and reduced to 446 pcm and 760 pcm, respectively. Consequently, when taking the
+difference in reactivities to estimate the rod worth, the neutron diffusion method performed
+significantly worse than OpenMC-MG and the hybrid method. OpenMC-CE estimated the rod worth to be
+8370(53) pcm. The neutron diffusion method overestimated the rod worth by 1484 pcm. The hybrid
+method improved significantly on the neutron diffusion method with a rod worth estimate that is
+only 13 pcm
+higher than OpenMC-CE. OpenMC-MG reported a larger error of 172 pcm likely due to favorable error
+cancellation with the hybrid method and statistical uncertainties with the Monte Carlo methods.
+The OpenMC-CE and OpenMC-MG rod worth estimates agree within two standard deviations of each other.
 
@@
   \caption{$k_\text{eff}$ and control rod worth estimates for the 2-D quarter-core \gls{MSRE}
-    model. Error estimates are relative to OpenMC-CE.}
+    model. Error values are relative to OpenMC-CE.}
   \begin{tabular}{l S[table-format=1.5(2)] S S[table-format=1.5(2)] S S[table-format=4(2)] S}
@@
 	  OpenMC-MG & 1.11979(42) & 618 & 1.02204(41) & 446 & 8541(51) & 172 \\
+      Diffusion & 1.12059 & 682 & 1.00903 & -816 & 9867 & 1484 \\
       Hybrid & 1.12174 & 773 & 1.02532 & 760 & 8383 & 13 \\
-      Diffusion & 1.12059 & 682 & 1.00903 & -816 & 9867 & -1484 \\
     \bottomrule
@@
 
+\FloatBarrier
+
+I also calculated the normalized channel fission rate distribution by integrating the fission rate
+within each fuel salt channel and normalizing the values such that the integrated fission rate in
+each channel averages to 1 as follows
+%
+\begin{align}
+  \text{Normalized fission rate of channel }i &= \frac{\text{Fission rate in channel }i}{
+  \text{Total fission rate}}\cdot\frac{\text{Total number of channels}}{
+    \text{Total fission rate in channels}} \nonumber \\
+                                              &= \frac{\sum^G_{g=1}\int_{V_i}\Sigma_{f,g}
+  \phi_g(\vec{r})}{\sum^G_{g=1}\int_V\Sigma_{f,g}\phi_g(\vec{r})}\cdot\frac{N_i}{
+  \sum^{N_i}_{i=1}\sum^G_{g=1}\int_{V_i}\Sigma_{f,g}\phi_g(\vec{r})}.
+\end{align}
+%
+The total fission rate includes fission rate contributions from peripheral fuel salt regions.
+Figures \ref{fig:1/4-no-rod} and \ref{fig:1/4-rod} show the normalized fission rate distributions
+of the quarter-core model for the ``No Rod'' and ``Rod'' cases, respectively. I presented 1/8th
+slices of the core in the figures because the distributions are symmetrical about the $y=x$ line
+outside of statistical variances for OpenMC-CE. I combined the midpoints between the centroids of
+neighboring fuel salt channels to form the diamond lattices shown in the figures.
+Each diamond element in the figures corresponds a horizontally or vertically oriented channel
+depicted in Figure \ref{fig:1/4-geom}. The lattice also provides a guide for subdividing the
+annular fuel regions around the control rod thimbles. The fission rate
+distributions omit four channels along the outer edge due to the proximity to peripheral annular
+fuel regions and the consequent difficulty in accurately tallying fission rates.
+
+In the ``No Rod'' case, the peak channel fission rate occured several channels away from the
+physical center of the reactor core. The presence of the empty (air-filled) rod thimble depressed
+fission rates near the center. The neutron diffusion and hybrid methods report similar percentage
+errors relative to OpenMC-CE except near the center. While the hybrid method benefits from drift
+corrections, the neutron diffusion method's reliance on scalar diffusion coefficients tends to
+artificially flatten neutron fluxes in near-void regions. Therefore, the largest percentage errors
+from the neutron diffusion method occurred in the second and third diamond elements
+diagonally up from the bottom which represent the annular fuel region surrounding the rod thimble.
+In the ``Rod'' case, the introduction of the control rod shifted the peak further away from the
+center as expected of a highly neutron-absorbing material. The rod further worsened fission rate
+errors for the neutron diffusion method with large errors of 1\% or more at the core center and
+periphery. On the other hand, the fission rate errors for the hybrid method largely remained below
+1\%.
+
 \begin{table}[htb]
@@
   \caption{Absolute mean and maximum percentage errors in the normalized channel fission rates of
-  the 2-D \gls{MSRE} quarter-core models relative to OpenMC. The average standard deviation of
-  OpenMC normalized channel fission rates is 0.2\%.}
+  the 2-D \gls{MSRE} quarter-core models relative to OpenMC. The mean relative standard deviation of
+  OpenMC normalized channel fission rates is 0.20\%.}
   \begin{tabular}{l S S S S}
@@
                             \cmidrule(r){1-1} \cmidrule(rl){2-3} \cmidrule(l){4-5}
+    Diffusion & 0.40 & 2.63 & 2.01 & 17.44 \\
     Hybrid & 0.40 & 1.32 & 0.43 & 3.08 \\
-    Diffusion & 0.40 & 2.63 & 2.01 & 17.44 \\
     \bottomrule
@@
 
+Table \ref{table:quarter-core-power} shows the absolute mean and maximum percentage errors of the
+normalized channel fission rates from the neutron diffusion and hybrid methods relative to
+OpenMC-CE. Overall, the hybrid method reported smaller percentage errors than the neutron diffusion
+method. The absolute maximum percentage error from the neutron diffusion method jumped to 17.44\%
+as compared to the 3.08\% for the hybrid method in the ``Rod'' case. Along with the figures, these
+results demonstrate the hybrid method provides improved fission rate distribution accuracy in
+the existence of near-void and highly neutron-absorbing regions.
+
 \FloatBarrier
@@
 \subsubsection{2-D full-core results}
+
+I performed the same $k_\text{eff}$ and fission rate distribution analyses with the 2-D full-core
+model under four configurations starting from no rod inserted to all three rods inserted. The
+$k_\text{eff}$ and error estimates in Table \ref{table:full-core-k} show the same trends observed
+with the 2-D quarter-core model. OpenMC-MG and the hybrid method exhibit moderate but consistent
+reactivity errors of approximately the 550 pcm and 800 pcm, respectively. The neutron diffusion
+method reported errors ranging from 692 pcm when no rod is inserted to -470 pcm when all three rods
+are inserted. Consequently, in Table \ref{table:full-core-worth}, OpenMC-MG and the hybrid method
+report accurate control rod worths within 1-3 standard deviations of OpenMC-CE while the neutron
+diffusion method increasingly overestimates the rod worth as more rods are inserted.
 
@@
   \caption{$k_\text{eff}$ estimates for the 2-D full-core \gls{MSRE} model with the indicated rods
-    inserted. Error estimates are relative to OpenMC-CE.}
+    inserted. Error values are relative to OpenMC-CE.}
   \setlength\tabcolsep{2pt}
@@
     OpenMC-MG & 1.10637(19) & 640 & 1.07633(20) & 567 & 1.05235(20) & 495 & 1.03262(17) & 541 \\
+    Diffusion & 1.10701 & 692 & 1.07120 & 122 & 1.04415 & -252 & 1.02195 & -470 \\
     Hybrid & 1.10843 & 808 & 1.07906 & 802 & 1.05553 & 781 & 1.03582 & 841 \\
-    Diffusion & 1.10701 & 692 & 1.07120 & 122 & 1.04415 & -252 & 1.02195 & -470 \\
     \bottomrule
@@
   \caption{Control rod worth estimates for the 2-D full-core \gls{MSRE} with the
-  indicated rods inserted. Error estimates are relative to OpenMC-CE.}
+  indicated rods inserted. Error values are relative to OpenMC-CE.}
   \setlength\tabcolsep{5pt}
@@
     OpenMC-MG & 2523(23) & 73 & 4640(24) & 146 & 6455(22) & 98 \\
+    Diffusion & 3019 & 569 & 5439 & 945 & 7519 & 1162 \\
     Hybrid & 2455 & 5 & 4521 & 27 & 6323 & -34 \\
-    Diffusion & 3019 & 569 & 5439 & 945 & 7519 & 1162 \\
     \bottomrule
@@
 
+Plots of the normalized channel fission rate distributions for all four rod configurations are
+available in Appendix \ref{chap:2d-fiss-rate}. They are best viewed in digital pdf format with
+magnification due to the numerous fuel salt channels and small fonts.
+
+Table \ref{table:full-core-power} shows the absolute mean and maximum percentage errors in the
+normalized channel fission rates from the neutron diffusion and hybrid methods relative to
+OpenMC-CE. Both mean and maximum percentage errors from the neutron diffusion method steadily
+increase with the number of rods inserted. The mean percentage error from the hybrid method remains
+constant at 0.43\% while the maximum percentage error increases marginally from 1.45\% to 2.52\%
+when all three rods are inserted. These results demonstrate the hybrid method remains
+effective at improving control rod worth and fission rate distribution estimates in less symmetric
+geometries.
+
 \begin{table}[htb]
@@
   \caption{Absolute mean and maximum percentage errors in the normalized channel fission rates of
-  the 2-D \gls{MSRE} full-core models relative to OpenMC. The average standard deviation of
+  the 2-D \gls{MSRE} full-core models relative to OpenMC. The mean relative standard deviation of
   OpenMC normalized channel fission rates is 0.27\%.}
@@
                             \cmidrule(r){1-1} \cmidrule(rl){2-3} \cmidrule(rl){4-5} \cmidrule(rl){6-7} \cmidrule(l){8-9}
+    Diffusion & 0.45 & 2.95 & 0.94 & 12.61 & 1.35 & 15.34 & 1.67 & 17.09 \\
     Hybrid & 0.43 & 1.45 & 0.43 & 1.82 & 0.43 & 2.26 & 0.43 & 2.52 \\
-    Diffusion & 0.45 & 2.95 & 0.94 & 12.61 & 1.35 & 15.34 & 1.67 & 17.09 \\
     \bottomrule
@@
 
-\section{3-D Models \& Results}
+\section{3-D Neutronics Models \& Results} \label{sec:3d-results}
+
+TBD
 
@@
 
-The hybrid method does not alter the $k_\text{eff}$ and flux estimates relative to the neutron
-diffusion method in cases where the neutron diffusion method provides
+TBD
+
+%The hybrid method does not alter the $k_\text{eff}$ and flux estimates relative to the neutron
+%diffusion method in cases where the neutron diffusion method provides

### 2024-08-23 appendix.tex

@@
-\section{Numerical Implementation of Neutronics Solvers on Python} \label{sec:implementation}
+\chapter{Numerical Implementation of Neutronics Solvers on Python} \label{chap:implementation}
 
@@
 
-\subsection{Neutron Diffusion Method} \label{sec:python-diffusion}
+\section{Neutron Diffusion Method} \label{sec:python-diffusion}
 
@@
 
-Discretizing the multigroup $k$-eigenvalue neutron diffusion equations in Eq.\ \ref{eq:1d-diff}
+The 1-D form of the multigroup $k$-eigenvalue neutron diffusion equations (Eq. \ref{eq:mg-diff})
+is:
+%
+\begin{align}
+  -\frac{d}{dx} D_g(x) \frac{d}{dx} \phi_g(x) + \Sigma_{t,g}(x) \phi_g(x) &= \sum^G_{g'=1}\left[
+  \Sigma_s^{g'\rightarrow g}(x)\phi_{g'}(x) + \chi_g\frac{\nu\Sigma_{f,g'}(x)}{k}
+  \phi_{g'}(x)\right] + S_g(x)
+  \label{eq:1d-diff}
+  \shortintertext{where}
+    D_g(x) &= \frac{1}{3 \Sigma_{tr}(x)} \nonumber \\
+           &= \mbox{ isotropic neutron diffusion coefficient for group }g. \nonumber
+\end{align}
+
+Discretizing Eq.\ \ref{eq:1d-diff}
 and reformulating the scattering term using neutron balance in the control volume bounded by
@@
 
-\subsection{$S_N$ Neutron Transport Method} \label{python-sn}
+\section{$S_N$ Neutron Transport Method} \label{sec:python-sn}
 
@@
 
-Discretizing the multigroup $S_N$ neutron transport equations in Eq.\ \ref{eq:1d-sn} about
-$x_{i+\sfrac{1}{2}}$ yields
+The multigroup, discrete ordinates ($S_N$ approximation) form of the Cartesian 1-D, $k$-eigenvalue
+neutron transport equation with the Gauss-Legendre quadrature set is:
+%
+\begin{align}
+  \mu_n \frac{d}{dx}\Psi_g(x, \mu_n) + \Sigma_{t,g}(x)&\Psi_g(x, \mu_n) -
+\sum^G_{g'=1} \sum^N_{n'=1} \sum^L_{l=0} \frac{\left(2l+1\right)}{2}
+\Sigma^{g'\rightarrow g}_{s,l}(x) P_l(\mu_{n'} - \mu_n)
+w_{n'}\Psi_{g'}(x,\mu_{n'}) \nonumber \\
+  &= \sum^G_{g'=1} \frac{\chi_g}{2} \frac{\nu\Sigma_{f,g'}(x)}{k} \phi_{g'}(x) + S_g(x,\mu_n)
+  \label{eq:1d-sn}
+  \shortintertext{where}
+  \mu_n &= \mbox{ cosine of $\hat{\Omega}_n$ relative to the $x$-axis,} \nonumber \\
+  \Psi_g(x,\mu_n) &= \mbox{ neutron angular flux along $\mu_n$ in group $g$,} \nonumber \\
+  G &= \mbox{ total number of energy groups,} \nonumber \\
+  N &= \mbox{ total number of discrete ordinates,} \nonumber \\
+  l &= \mbox{ Legendre expansion order,} \nonumber \\
+  L &= \mbox{ highest Legendre expansion order,} \nonumber \\
+  \Sigma^{g'\rightarrow g}_{s,l}(x) &= \mbox{ $l$-th Legendre moment of the macroscopic
+scattering} \nonumber \\
+  &\ \ \ \ \mbox{ cross section from group $g'$ to $g$,} \nonumber \\
+  P_l &= \mbox{ Legendre polynomial of degree $l$,} \nonumber \\
+  w_{n'} &= \mbox{ $n'$-th quadrature weight,} \nonumber \\
+  k &= \mbox{ neutron multiplication factor.} \nonumber
+\end{align}
+%
+Discretizing Eq.\ \ref{eq:1d-sn} about $x_{i+\sfrac{1}{2}}$ yields
 %
@@
 
-\subsection{Hybrid $S_N$-Diffusion Method} \label{python-hybrid}
+\section{Hybrid $S_N$-Diffusion Method with Diffusion Correction} \label{sec:python-hybrid}
 
-Without loss of generality, consider a 1-D system symmetric about the left boundary at $x_0=0$ cm.
-In this case, as with Case 0, the control rod region is the left-most region, followed by other
+Without loss of generality, consider a 1-D system symmetric about the left boundary at $x_0=0$ cm
+with the control rod region being the left-most region, followed by other
 regions. The $S_N$ subproblem domain $V_1$, containing the control rod and some region beyond it,
 is bounded by $x_0$ and $x_j$ for some $x_j$
-located several mean free paths to the right of the control rod region as governed by the relevant
-discussion in Section \ref{sec:hybrid-method}.
+located several mean free paths to the right of the control rod region.
 
@@
 
-For the $S_N$ subproblem boundary conditions, we discretize Eqs. \ref{eq:p1-j} and
-\ref{eq:sn-psi-j} as follows
+An option for deriving $S_N$ subproblem boundary conditions is to assume a uniformly
+isotropic transmission of angular flux, i.e., all N/2 incoming angular fluxes $\Psi$ are equal in
+magnitude. This is a good assumption far from the control rod, but will not produce accurate
+fluxes if the boundary is too close to the control rod. This type of boundary condition is similar
+to the white boundary condition, which describes the uniformly
+isotropic reflection of particles at a boundary, except this case involves transmission rather than
+reflection. Isotropic flux transmission at $x$ can be expressed mathematically for forward angular
+fluxes as:
+%
+\begin{align}
+  \sum^N_{n=N/2+1}w_n\mu_n\Psi(x,\mu_n) =& J_{+}(x) && (\mu_n>0) \nonumber \\
+  \Psi(x,\mu_n)\sum^N_{n=N/2+1}w_n\mu_n =& J_{+}(x) && (\because \mbox{isotropic transmission})
+  \nonumber \\
+  \Psi(x,\mu_n) =& J_{+}(x)\Bigg/\sum^N_{n=N/2+1}w_n\mu_n
+\end{align}
+%
+and backward angular fluxes as:
+%
+\begin{align}
+  \sum^{N/2}_{n=1}w_n\mu_n\Psi(x,\mu_n) =& J_{-}(x) && (\mu_n<0) \nonumber \\
+  \Psi(x,\mu_n)\sum^{N/2}_{n=1}w_n\mu_n =& J_{-}(x) && (\because \mbox{isotropic transmission})
+  \nonumber \\
+  \Psi(x,\mu_n) =& J_{-}(x)\Bigg/\sum^{N/2}_{n=1}w_n\mu_n \label{eq:sn-psi-j}
+\end{align}
+%
+Discretizing Eqs. \ref{eq:sn-psi-j} yields
 %
@@
 
-The \gls{SVDC} formulation in Eq.\ \ref{eq:svdc} is discretized using the diamond difference scheme
-as follows
+The diffusion correction formulation in Eq.\ \ref{eq:svdc} is discretized using the diamond
+difference scheme as follows
 %
@@
   calculation boundary conditions using Eq.\ \ref{eq:sn-bc}}
-  \shortintertext{3. Calculate $\phi^{m+\sfrac{1}{2}}$ with the $S_N$ transport sub-solver on
+  \shortintertext{3. Calculate $\phi^{m+\sfrac{1}{2}}$ with the $S_N$ transport subsolver on
   $V_1$}
-  \shortintertext{4. Generate \glspl{SVDC} with $\phi^{m+\sfrac{1}{2}}$ and $J^{m+\sfrac{1}{2}}$
-  using Eq.\ \ref{eq:svdc-num}}
-  \shortintertext{5. Calculate $\phi^{m+1}$ with the newly generated \glspl{SVDC} and the neutron
-  diffusion solver on $V_0$}
+  \shortintertext{4. Generate $D^s_g$ parameters with $\phi^{m+\sfrac{1}{2}}$ and
+    $J^{m+\sfrac{1}{2}}$ using Eq.\ \ref{eq:svdc-num}}
+  \shortintertext{5. Calculate $\phi^{m+1}$ with the newly generated $D^s_g$ parameters and the
+    neutron diffusion solver on $V_0$}
   \shortintertext{6. Check whether convergence is reached}
@@
 
-The $S_N$ sub-solver only covers $V_1$, so it does not calculate an independent $k$
+The $S_N$ subsolver only covers $V_1$, so it does not calculate an independent $k$
 estimate. Instead, the $k^m$ from the neutron diffusion calculation scales the fission neutron
-source term in the $S_N$ sub-solver.
+source term in the $S_N$ subsolver.
+
+\chapter{2-D MSRE Normalized Channel Fission Rate Distributions} \label{chap:2d-fiss-rate}
+
+This appendix contains normalized channel fission rate distributions for the 2-D full-core
+\gls{MSRE} model in Section \ref{sec:2d-nts-results}. I recommend viewing these figures digitally
+due to the small font.
+
+\begin{figure}[htb!]
+  \centering
+  \includegraphics[width=.4\columnwidth]{msre-full-legend}
+  \caption{Legend for the normalized channel fission rate distribution plots.}
+  \label{fig:full-legend}
+\end{figure}
+
+\begin{figure}[htb!]
+  \centering
+  \includegraphics[width=\columnwidth]{msre-full-0-power}
+  \caption{Normalized channel fission rate distribution of the 2-D \gls{MSRE} full-core model with
+  no rod inserted.}
+  \label{fig:0-rod}
+\end{figure}
+
+\begin{figure}[htb!]
+  \centering
+  \includegraphics[width=\columnwidth]{msre-full-1-power}
+  \caption{Normalized channel fission rate distribution of the 2-D \gls{MSRE} full-core model with
+  Rod 1 inserted.}
+  \label{fig:1-rod}
+\end{figure}
+
+\begin{figure}[htb!]
+  \centering
+  \includegraphics[width=\columnwidth]{msre-full-12-power}
+  \caption{Normalized channel fission rate distribution of the 2-D \gls{MSRE} full-core model with
+  Rods 1 \& 2 inserted.}
+  \label{fig:12-rod}
+\end{figure}
+
+\begin{figure}[htb!]
+  \centering
+  \includegraphics[width=\columnwidth]{msre-full-123-power}
+  \caption{Normalized channel fission rate distribution of the 2-D \gls{MSRE} full-core model with
+  Rods 1, 2 \& 3 inserted.}
+  \label{fig:123-rod}
+\end{figure}

### 2024-08-23 hybrid.tex

@@
 section, the derivation for the \gls{SAAF} formulation of the $S_N$ method and its
-implementation in Moltres follows closely the derivation developed by Wang et al.
+implementation in Moltres follows closely the derivation developed by Wang et al.\
 \cite{wang_diffusion_2014, wang_rattlesnake_2018}.
@@
 can be introduced by modifying the neutron diffusion equations.
-In this work, I explored two options for applying
-transport corrections to the neutron diffusion equations: applying
-diffusion corrections or adding a drift correction term.
-I performed preliminary investigations on 1-D test cases with both types of transport correction. I
-eventually elected to use the drift correction term for the remaining test cases and reactor model
+In this work, I explored two options for applying transport corrections to the neutron diffusion
+equations: a diffusion correction scheme and a drift correction scheme.
+I performed preliminary investigations on 1-D \gls{MSRE} models with both transport correction
+schemes. I eventually elected to use the drift correction scheme for the 2-D and 3-D reactor model
 demonstrations due to superior numerical properties. I will make clear the reasons for this
 selection through 1-D test case results in Chapter \ref{chap:msre}.
+
+Consequently, most of this chapter focuses on the $S_N$ and the drift correction-based hybrid
+method implementations on Moltres. On the other hand, I implemented the diffusion correction-based
+hybrid method using the Python programming language for quick
+prototyping. Implementation details of the $S_N$, neutron diffusion, and hybrid methods on Python
+are available in Appendix \ref{chap:implementation}. Both correction schemes share the iteration
+algorithm described in Section \ref{sec:hybrid-algorithm}.
 
@@
 to illustrate the buffer region and the behavior of $\vec{D}_g$. Case 3b is a 1-D slab problem
-consisting of control rod, air, fuel-graphite repeating lattice, and core vessel regions as shown
-in Figure \ref{fig:3b-geometry}. The fuel regions are neutron-multiplying media containing
+consisting of control rod, air, fuel-graphite repeating lattice, and reactor vessel regions as
+shown in Figure \ref{fig:3b-geometry}. The fuel regions are neutron-multiplying media containing
 $^\text{235}$U fissile material. I took material compositions of each region from the \gls{MSRE}

### 2024-09-11 acros.tex
saaf sn method
sn terms
add verification chapter for referencing, hybrid chapter edits
turbulence verification section

@@
 \newacronym{BDBE}{BDBE}{Beyond Design Basis Event}
+\newacronym{BFS}{BFS}{backward-facing step}
 \newacronym{BIDS}{BIDS}{Berkeley Institute for Data Science}
@@
 \newacronym{CI}{CI}{continuous integration}
+\newacronym{CMFD}{CMFD}{coarse-mesh finite difference}
 \newacronym{CNEN}{CNEN}{Comiss\~{a}o Nacional de Energia Nuclear}
@@
 \newacronym{DNP}{DNP}{delayed neutron precursor}
-\newacronym{DNS}{DNS}{Direct numerical simulation}
+\newacronym{DNS}{DNS}{direct numerical simulation}
 \newacronym{DOE}{DOE}{Department of Energy}
@@
 \newacronym{GIF}{GIF}{Generation IV International Forum}
+\newacronym{GMRES}{GMRES}{generalized minimal residual}
 \newacronym{GPAM}{GPAM}{Generic Performance Assessment Model}
@@
 \newacronym{HLW}{HLW}{high level waste}
+\newacronym{HOLO}{HOLO}{high-order/low-order}
 \newacronym{HPC}{HPC}{high-performance computing}
@@
 \newacronym{INL}{INL}{Idaho National Laboratory}
-\newacronym{INS}{INS}{incompressible Navier-Stokes}
+\newacronym{INS}{INS}{Incompressible Navier-Stokes}
+\newacronym{INSAD}{INSAD}{Incompressible Navier-Stokes Automatic Differentiation}
 \newacronym{IO}{I/O}{input/output}
@@
 \newacronym{NAGRA}{NAGRA}{National Cooperative for the Disposal of Radioactive Waste}
+\newacronym{NASA}{NASA}{National Aeronautics and Space Administration}
+\newacronym{NDA}{NDA}{nonlinear diffusion acceleration}
 \newacronym{NEAMS}{NEAMS}{Nuclear Energy Advanced Modeling and Simulation}
@@
 \newacronym{OCRWM}{OCRWM}{Office of Civilian Radioactive Waste Management}
-\newacronym{OOP}{OOP}{Object-oriented programming}
+\newacronym{OOP}{OOP}{object-oriented programming}
 \newacronym{ORION}{ORION}{ORION}
@@
 \newacronym{PH}{PRONGHORN}{PRONGHORN}
+\newacronym{PJFNK}{PJFNK}{preconditioned Jacobian-free Newton-Krylov}
 \newacronym{PoliMi}{PoliMi}{Politecnico di Milano}
@@
 \newacronym{RSM}{RSM}{Reynolds stress model}
+\newacronym{SA}{SA}{Spalart-Allmaras}
+\newacronym{SAAF}{SAAF}{self-adjoint angular flux}
 \newacronym{SAM}{SAM}{System Analysis Module}

### 2024-09-11 turbulence.tex
turbulence verification section

As detailed in Section \ref{sec:th}, Moltres compiles with the MOOSE Navier-Stokes
\cite{peterson_overview_2018} and Heat Transfer modules for incompressible flow
modeling capabilities by default. Moltres couples with these modules natively because they are all
built on the MOOSE framework.

% Past work with Moltres have modeled incompressible salt flow and temperature advection-diffusion
% with the \gls{INS} or \gls{INSAD} implementations of
% the Navier-Stokes equations. Coupling with compressible, porous media, and
% finite volume implementations within the Navier-Stokes module is possible through native
% \gls{MOOSE} multiphysics coupling capabilities but has not been demonstrated yet.

To address the lack of turbulence modeling in Moltres, I implemented a \gls{SA} turbulence model
\cite{spalart_one-equation_1994}, described in Section \ref{sec:lit-turb}, with \gls{SUPG}
stabilization on Moltres. On balance the \gls{SA} model is a complete (does not require prior
knowledge of the actual turbulence behavior) and computationally efficient one-equation turbulence
model for approximating wall-bounded turbulent flows. The \gls{SA} model implementation in Moltres
couples seamlessly with the continuous \gls{FEM} \gls{INSAD} \cite{lindsay_automatic_2021} model
from the Navier-Stokes module. Alongside the \gls{SA} model, Moltres also now has turbulent
diffusion physics kernels for temperature and the delayed neutron precursors.

The SA model in Moltres follows the \gls{SA} model with rotation correction (SA-R) implementation
\cite{aupoix_extensions_2003, dacles-mariani_numericalexperimental_1995} as described on the
\gls{NASA} Turbulence Modeling Resource website \cite{rumsey_turbulence_nodate}. The $f_{t2}$
term is togglable using the \texttt{use\_ft2\_term} input parameter (false by default). Moltres'
implementation solves for the modified dynamic viscosity $\tilde{\mu}$ as follows:

\begin{gather}
  \rho \frac{\partial\tilde{\mu}}{\partial t} + \rho \mathbf{u}\cdot\nabla\tilde{\mu} = \rho c_{b1}
  \left(1-f_{t2}\right)\tilde{S}\tilde{\mu} + \frac{1}{\sigma}\{\nabla\cdot\left[\left(\mu+
  \tilde{\mu}\right)\nabla\tilde{\mu}\right] + c_{b2}|\nabla\tilde{\mu}|^2\} - \left(c_{w1}f_w -
  \frac{c_{b1}}{\kappa^2}f_{t2}\right)\left(
  \frac{\tilde{\mu}}{d}\right)^2
  \shortintertext{where}
  \begin{align*}
    \mu_t &= \tilde{\mu}f_{v1} = \text{turbulent eddy viscosity}, &
    f_{v1} &= \frac{\chi^3}{\chi^3 + c_{v1}^3}, \\
    \chi &= \frac{\tilde{\mu}}{\mu}, &
    \tilde{S} &= \Omega + \frac{\tilde{\nu}}{\kappa^2 d^2} f_{v2}, \\
    f_{v2} &= 1 - \frac{\chi}{1+\chi f_{v1}}, &
    \Omega &= \sqrt{2W_{ij}W_{ij}} = \text{vorticity magnitude}, \\
    W_{ij} &= \frac{1}{2}\left(\frac{\partial u_i}{\partial x_j} - \frac{\partial u_j}{\partial x_i}
    \right), &
      f_w &= g\left(\frac{1 + c_{w3}^6}{g^6 + c_{w3}^6}\right)^{1/6}, \\
      g &= r + c_{w2}\left(r^6 - r\right), &
      r &= \text{min}\left(\frac{\tilde{\nu}}{\tilde{S}\kappa^2d^2}, 10\right), \\
      f_{t2} &= c_{t3} \exp{\left(-c_{t4}\chi^2\right)},
  \end{align*}
\shortintertext{and the constants are}
  \sigma = \frac{2}{3}, \ c_{b1} = 0.1355, \ c_{b2} = 0.622, \ \kappa = 0.41, \
  c_{w1} = \frac{c_{b1}}{\kappa^2} + \frac{1+c_{b2}}{\sigma}, \nonumber \\
  c_{w2} = 0.3, \ c_{w3} = 2, \
  c_{v1} = 7.1, \ c_{t3} = 1.2, \ c_{t4} = 0.5 \ \text{.} \nonumber
\end{gather}

The \gls{SA} model is verified simple verification and validation tests of the \gls{SA} model in Moltres using
reference problems for channel, pipe, and \gls{BFS} flow. The Moltres test dataset is available on
Zenodo \cite{park_dataset_2023}. The following subsections show plots comparing
results from Moltres to reference data from literature.

\subsection{Turbulent Channel Flow Verification Test}

The channel flow test is based on the \gls{DNS} of turbulent channel flow
with Re$_\tau\approx395$ by Moser et al.\ \cite{moser_direct_1999}. Corresponds to Re = 13750.

\begin{figure}[htb!]
  \centering
  \begin{subfigure}[b]{0.48\columnwidth}
    \centering
    \includegraphics[width=\columnwidth]{channel_vel}
    \caption{Normalized velocity distribution across the channel.}
    \label{fig:}
  \end{subfigure}
  \hfill
  \begin{subfigure}[b]{0.48\columnwidth}
    \centering
    \includegraphics[width=\columnwidth]{channel_nondim}
    \caption{Dimensionless velocity vs.\ dimensionless wall distance.}
    \label{fig:}
  \end{subfigure}
  \begin{subfigure}[b]{0.48\columnwidth}
    \centering
    \includegraphics[width=\columnwidth]{channel_stress}
    \caption{Normalized stress distribution across the channel.}
    \label{fig:}
  \end{subfigure}
  \caption{Comparison of turbulent channel flow results at Re$_\tau\approx395$ against reference
  \gls{DNS} results \cite{moser_direct_1999}.}
  \label{fig:}
\end{figure}

\subsection{Turbulent Pipe Flow Validation Test}

The pipe flow test is based on the turbulent pipe flow experiment with Re $\approx 40000$ by
Laufer \cite{laufer_structure_1954}.

\begin{figure}[htb]
  \centering
  \begin{subfigure}[b]{0.48\columnwidth}
    \centering
    \includegraphics[width=\columnwidth]{pipe_vel}
    \caption{Normalized velocity distribution across the pipe.}
    \label{fig:}
  \end{subfigure}
  \hfill
  \begin{subfigure}[b]{0.48\columnwidth}
    \centering
    \includegraphics[width=\columnwidth]{pipe_nondim}
    \caption{Dimensionless velocity vs.\ dimensionless wall distance}
    \label{fig:}
  \end{subfigure}
  \begin{subfigure}[b]{0.48\columnwidth}
    \centering
    \includegraphics[width=\columnwidth]{pipe_stress}
    \caption{Normalized turbulent shear stress distribution across the pipe.}
    \label{fig:}
  \end{subfigure}
  \caption{Comparison of turbulent pipe flow results at Re $\approx 40000$ against reference
  experimental data \cite{laufer_structure_1954}.}
\end{figure}

\subsection{Backward-Facing Step Flow Validation Test}

The \gls{BFS} flow test is based on the BFS flow experiment with Re $\approx36000$ by Driver \&
Seegmiller \cite{driver_features_1985}. The \gls{SA} model implementation in Moltres performs
largely similarly to the reference SA model results
provided on the \gls{NASA} Turbulence Modeling Resource website \cite{rumsey_turbulence_nodate}.

\begin{table}[htb]
  \centering
  \small
  \caption{\gls{BFS} flow reattachment length estimates normalized by step height $H$.}
  \begin{tabular}{l c c c}
    \toprule
    & {Experimental data \cite{driver_features_1985}} & {Reference \gls{SA} model
  \cite{rumsey_turbulence_nodate}} & {Moltres \gls{SA} model} \\
    \midrule
    Reattachment length [-] & {$6.26 \pm 0.10$} & 6.1 & 6.36 \\
    \bottomrule
  \end{tabular}
  \label{table:bfs}
\end{table}

\begin{figure}[htb]
  \centering
  \hfill
  \begin{subfigure}[b]{0.4\columnwidth}
    \centering
    \includegraphics[width=\columnwidth]{bfs_upstream_vel}
    \caption{Normalized velocity distribution at $x/H=-4$ upstream of step.}
    \label{fig:}
  \end{subfigure}
  \hfill
  \begin{subfigure}[b]{0.4\columnwidth}
    \centering
    \includegraphics[width=\columnwidth]{bfs_downstream_vel}
    \caption{Normalized velocity distribution at various $x/H$ locations downstream of step.}
    \label{fig:}
  \end{subfigure} \hfill \\
  \centering
  \hfill
  \begin{subfigure}[b]{0.4\columnwidth}
    \centering
    \includegraphics[width=\columnwidth]{bfs_cf}
    \caption{Skin friction coefficient along the bottom wall.}
    \label{fig:}
  \end{subfigure}
  \hfill
  \begin{subfigure}[b]{0.4\columnwidth}
    \centering
    \includegraphics[width=\columnwidth]{bfs_cp}
    \caption{Skin pressure coefficient along the bottom wall.}
    \label{fig:}
  \end{subfigure} \hfill \\
  \centering
  \begin{subfigure}[b]{0.4\columnwidth}
    \centering
    \includegraphics[width=\columnwidth]{bfs_stress}
    \caption{Normalized turbulent shear stress distribution at various $x/H$ locations downstream
    of step.}
    \label{fig:}
  \end{subfigure}
  \caption{Comparison of backward facing step flow results at Re $\approx 36000$ against reference
  experimental data and computational data from the CFL3D code.}
  \label{fig:}
\end{figure}

\begin{figure}[htb!]
  \centering
  \includegraphics[width=\columnwidth]{bfs}
  \caption{}
  \label{fig:}
\end{figure}

% !media bfs_upstream_vel.png
%        id=bfs-upstream-vel
%        caption=Normalized velocity distribution at $x/H$ equal -4 (upstream of step).
%        style=width:33%;padding:10px;float:left;
% 
% !media bfs_downstream_vel.png
%        id=bfs-downstream-vel
%        caption=Normalized velocity distribution at various $x/H$ locations (downstream of step).
%        style=width:33%;padding:10px;float:left;
% 
% !media bfs_cf.png
%        id=bfs-cf
%        caption=Skin friction coefficient along the bottom wall.
%        style=width:33%;padding:10px;float:left;
% 
% !media bfs_cp.png
%        id=bfs-cp
%        caption=Skin pressure coefficient along the bottom wall.
%        style=width:33%;padding:10px;margin-left:17%;float:left;
% 
% !media bfs_stress.png
%        id=bfs-stress
%        caption=Normalized turbulent shear stress distribution at various $x/H$ locations
%        (downstream of step).
%        style=width:33%;padding:10px;margin-right:17%;float:left;
% 
% !media bfs.png
%        id=bfs
%        caption=Close-up of the velocity magnitude and streamlines around the backward-facing step
%        style=width:80%;float:left;margin-left:10%;margin-right:10%;

### 2024-09-24 lit.tex
turbulence verification section
turbulence model verification section

@@
 
-\section{Turbulence Modeling in MSRs}
+\section{Turbulence Modeling in MSRs} \label{sec:lit-turb}
 
@@
 \gls{MSFR}, salt flow in the central core region is highly turbulent and
-reaches Reynolds numbers on the order of $10^5$.
+reaches Reynolds numbers on the order of $10^5$ based on core dimensions and salt flow rate
+\cite{rouch_preliminary_2014}.
 
@@
 turbulence rate of dissipation to obtain the eddy viscosity distribution. The
-most common and best-performing one-equation model is the Spalart-Allmaras
+most common one-equation model is the Spalart-Allmaras
 model, which directly provides an equation for the eddy viscosity with several
@@
 Investigations with
-one-equation models reveal the need for an extra equation to account for
+one-equation models revealed the need for an extra equation to account for
 turbulent length scales separately from turbulent velocity. Thus, two-equation

### 2024-09-24 turbulence.tex
turbulence model verification section

@@
 
-% Past work with Moltres have modeled incompressible salt flow and temperature advection-diffusion
-% with the \gls{INS} or \gls{INSAD} implementations of
-% the Navier-Stokes equations. Coupling with compressible, porous media, and
-% finite volume implementations within the Navier-Stokes module is possible through native
-% \gls{MOOSE} multiphysics coupling capabilities but has not been demonstrated yet.
-
 To address the lack of turbulence modeling in Moltres, I implemented a \gls{SA} turbulence model
 \cite{spalart_one-equation_1994}, described in Section \ref{sec:lit-turb}, with \gls{SUPG}
-stabilization on Moltres. On balance the \gls{SA} model is a complete (does not require prior
+stabilization on Moltres. The \gls{SA} model estimates the turbulent eddy viscosity defined by the
+eddy viscosity hypothesis applied to the \gls{RANS} equations \cite{rodi_turbulence_2017}.
+On balance the \gls{SA} model is a complete (does not require prior
 knowledge of the actual turbulence behavior) and computationally efficient one-equation turbulence
 model for approximating wall-bounded turbulent flows. The \gls{SA} model implementation in Moltres
-couples seamlessly with the continuous \gls{FEM} \gls{INSAD} \cite{lindsay_automatic_2021} model
+couples seamlessly with the continuous \gls{FEM} \gls{INSAD}
+\cite{peterson_overview_2018, lindsay_automatic_2021} model
 from the Navier-Stokes module. Alongside the \gls{SA} model, Moltres also now has turbulent
@@
 
-The SA model in Moltres follows the \gls{SA} model with rotation correction (SA-R) implementation
+The SA model in Moltres follows the \gls{SA} model with a rotation correction scheme
 \cite{aupoix_extensions_2003, dacles-mariani_numericalexperimental_1995} as described on the
-\gls{NASA} Turbulence Modeling Resource website \cite{rumsey_turbulence_nodate}. The $f_{t2}$
-term is togglable using the \texttt{use\_ft2\_term} input parameter (false by default). Moltres'
-implementation solves for the modified dynamic viscosity $\tilde{\mu}$ as follows:
+\gls{NASA} Turbulence Modeling Resource website \cite{rumsey_turbulence_nodate}. The rotation
+correction reduces eddy viscosity in regions of rotational but non-turbulent flow where the
+original \gls{SA} model overestimates eddy viscosity. The \gls{SA} model implementation in Moltres
+solves for the modified dynamic viscosity $\tilde{\mu}$ (as opposed to $\tilde{\nu}$) as follows:
 
@@
 
-The \gls{SA} model is verified simple verification and validation tests of the \gls{SA} model in Moltres using
-reference problems for channel, pipe, and \gls{BFS} flow. The Moltres test dataset is available on
-Zenodo \cite{park_dataset_2023}. The following subsections show plots comparing
-results from Moltres to reference data from literature.
+The $f_{t2}$ turbulence trip term is togglable using the \texttt{use\_ft2\_term} input parameter
+(false by default) if turbulence trip (initiation) is not necessary
+\cite{rumsey_turbulence_nodate}. The following subsections cover verification and validation tests
+of the \gls{SA} model in Moltres using reference problems for turbulent channel, pipe, and
+\gls{BFS} flow. Moltres input files
+for all three reference problems are available at
+\url{https://github.com/arfc/moltres/tree/devel/problems/2023-basic-turbulence-cases}.
+The Moltres test dataset is available on Zenodo at this reference listing \cite{park_dataset_2023}.
 
@@
 
-The channel flow test is based on the \gls{DNS} of turbulent channel flow
-with Re$_\tau\approx395$ by Moser et al.\ \cite{moser_direct_1999}. Corresponds to Re = 13750.
+Moser et al.\ \cite{moser_direct_1999} performed \gls{DNS} simulations of turbulent channel flow
+for friction Reynolds number, Re$_\tau\approx395$ (corresponds to Reynolds number, Re
+$\approx 13750$). Their results serve as the reference solution for this turbulent channel flow
+test.
+
+The Moltres model for this test is a 2-D 140 m$\times$0.5 m half-channel. The main flow direction
+is in the positive $x$ direction with the inlet and outlet on the left and right ends,
+respectively. The channel wall lies along the top boundary ($y=0.5$ m) while the bottom boundary
+($y=0$ m) serves as a symmetry axis for the half-channel geometry. Figure \ref{fig:channel-mesh}
+shows a close-up view of the refined mesh near the inlet and along the top wall boundary. The mesh
+for the rest of the channel geometry follows the same mesh resolution as the rightmost column of
+elements shown on the right side of Figure \ref{fig:channel-mesh}. The dimensionless wall distance
+parameter $y^+$ of the first mesh element along wall boundary for fully developed flow at the end
+of the channel is 0.974, meeting the
+$y^+ \lesssim 1$ requirement for properly wall-resolved flow. Table \ref{table:channel} lists
+relevant flow parameters for the turbulent channel flow test. The Moltres channel flow model
+ran as a time-dependent simulation with adaptive time-stepping starting from a timestep size of 0.1
+s up to a maximum timestep size of 10 s with a timestep size growth factor of 1.1. The simulation
+automatically terminated at $t=186$ s when Moltres detected that the flow profile reached
+steady-state, i.e., the flow and turbulent viscosity distributions remained unchanged between
+successive timesteps.
+
+\begin{figure}[htb!]
+  \centering
+  \includegraphics[width=\columnwidth]{channel_mesh}
+  \caption{Close-up view of the refined mesh near the inlet (left boundary) for the channel flow
+    test. The mesh for the rest of the channel follows the mesh resolution of the rightmost column
+  of elements.}
+  \label{fig:channel-mesh}
+\end{figure}
+
+\begin{table}[htb]
+  \centering
+  \small
+  \caption{Relevant turbulent channel flow problem parameters. The $\tilde{\mu}_\text{inlet}$ value
+  at the inlet is set to fives times the $\mu$ value as recommended for the \gls{SA} model
+  \cite{spalart_one-equation_1994}.}
+  \begin{tabular}{l S}
+    \toprule
+    Property & {Value} \\
+    \midrule
+    Density, $\rho$ [kg m$^{-3}$] & 1.0 \\
+    Inlet velocity, $v_x$ [m s$^{-1}$] & 1.0 \\
+    Dynamic viscosity, $\mu$ [kg m$^{-1}$ s$^{-1}$] & 7.272e-5 \\
+    Reynolds number, Re [-] & 1.375e4 \\
+    Modified viscosity along inlet, $\tilde{\mu}_\text{inlet}$ [kg m$^{-1}$ s$^{-1}$] & 3.636e-4 \\
+    Modified viscosity along wall, $\tilde{\mu}_\text{wall}$ [kg m$^{-1}$ s$^{-1}$] & 0.0 \\
+    \bottomrule
+  \end{tabular}
+  \label{table:channel}
+\end{table}
+
+Figure \ref{fig:channel-verification} shows plots of normalized and nondimensionalized velocities,
+wall distances, and stresses from the \gls{DNS} data \cite{moser_direct_1999}
+and the Moltres \gls{SA} model. The Moltres \gls{SA} model is largely consistent with the reference
+\gls{DNS} flow data and reproduces the expected trends in the velocity and stress distributions
+across the channel and as a function of the wall distance.
 
@@
     \caption{Normalized velocity distribution across the channel.}
-    \label{fig:}
+    \label{fig:channel-vel}
   \end{subfigure}
@@
     \caption{Dimensionless velocity vs.\ dimensionless wall distance.}
-    \label{fig:}
+    \label{fig:channel-nondim}
   \end{subfigure}
@@
     \caption{Normalized stress distribution across the channel.}
-    \label{fig:}
+    \label{fig:channel-stress}
   \end{subfigure}
   \caption{Comparison of turbulent channel flow results at Re$_\tau\approx395$ against reference
-  \gls{DNS} results \cite{moser_direct_1999}.}
-  \label{fig:}
+  \gls{DNS} results \cite{moser_direct_1999}. The Moltres \gls{SA} model agrees consistently with
+  the reference data.}
+  \label{fig:channel-verification}
 \end{figure}
@@
 
-The pipe flow test is based on the turbulent pipe flow experiment with Re $\approx 40000$ by
-Laufer \cite{laufer_structure_1954}.
+In 1954, Laufer performed turbulent pipe flow experiments for Re $\approx 40000$
+\cite{laufer_structure_1954}. Their results serve as the reference solution for this turbulent pipe
+flow test.
+
+\begin{figure}[htb!]
+  \centering
+  \begin{subfigure}[b]{0.42\columnwidth}
+    \centering
+    \includegraphics[width=\columnwidth]{pipe_mesh}
+    \caption{Close-up view of the refined mesh near the inlet for the pipe flow test.}
+    \label{fig:pipe-mesh-1}
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}[b]{0.56\columnwidth}
+    \centering
+    \includegraphics[width=\columnwidth]{pipe_mesh_zoom}
+    \caption{Close-up view of the refined mesh near the inlet and wall boundary interface.}
+    \label{fig:pipe-mesh-2}
+  \end{subfigure}
+  \caption{Close-up views of the refined mesh near the inlet (bottom boundary) and wall (right
+  boundary). The mesh for the rest of the channel follows the mesh resolution of the topmost row of
+  elements in the left subfigure.}
+  \label{fig:pipe-mesh}
+\end{figure}
+
+The Moltres model for this test is a 2-D axisymmetric (R-Z coordinates) half-pipe with a radius of
+0.5 m and a length of 150 m. The main flow direction
+is in the positive $z$ direction with the inlet and outlet on the bottom and top ends,
+respectively. The channel wall lies along the right boundary ($r=0.5$ m) while the left boundary
+($r=0$ m) serves as a symmetry axis for the half-pipe geometry. Figure \ref{fig:pipe-mesh} shows
+two close-up views of the refined mesh near the inlet and along the right wall boundary. The mesh
+for the rest of the pipe geometry follows the same mesh resolution as the topmost row of
+elements shown on the near the top of Figure \ref{fig:pipe-mesh-1}. The dimensionless wall distance
+parameter $y^+$ of the first mesh element along wall boundary for fully developed flow at the end
+of the pipe is 0.847, meeting the
+$y^+ \lesssim 1$ requirement for properly wall-resolved flow. Table \ref{table:channel} lists
+relevant flow parameters for the turbulent pipe flow test. The Moltres channel flow model ran as a
+time-dependent simulation with adaptive time-stepping starting from a timestep size of 0.01
+s up to a maximum timestep size of 10 s with a timestep size growth factor of 1.1. The simulation
+automatically terminated at $t=165$ s when Moltres detected that the flow profile reached
+steady-state, i.e., the flow and turbulent viscosity distributions remained unchanged between
+successive timesteps.
+
+\begin{table}[htb]
+  \centering
+  \small
+  \caption{Relevant turbulent pipe flow problem parameters. The $\tilde{\mu}_\text{inlet}$ value
+  at the inlet is set to fives times the $\mu$ value as recommended for the \gls{SA} model
+  \cite{spalart_one-equation_1994}.}
+  \begin{tabular}{l S}
+    \toprule
+    Property & {Value} \\
+    \midrule
+    Density, $\rho$ [kg m$^{-3}$] & 1.0 \\
+    Inlet velocity, $v_z$ [m s$^{-1}$] & 1.0 \\
+    Dynamic viscosity, $\mu$ [kg m$^{-1}$ s$^{-1}$] & 2.5e-5 \\
+    Reynolds number, Re [-] & 4.0e4 \\
+    Modified viscosity along inlet, $\tilde{\mu}_\text{inlet}$ [kg m$^{-1}$ s$^{-1}$] & 1.25e-4 \\
+    Modified viscosity along wall, $\tilde{\mu}_\text{wall}$ [kg m$^{-1}$ s$^{-1}$] & 0.0 \\
+    \bottomrule
+  \end{tabular}
+  \label{table:pipe}
+\end{table}
+
+Figure \ref{fig:pipe-verification} shows plots of normalized and nondimensionalized velocities,
+wall distances, and stresses from the experimental data \cite{laufer_structure_1954}
+and the Moltres \gls{SA} model. The Moltres \gls{SA} model is largely consistent with the reference
+experimental flow data and reproduces the expected trends in the velocity and stress distributions
+across the channel and as a function of the wall distance.
 
@@
     \caption{Normalized velocity distribution across the pipe.}
-    \label{fig:}
+    \label{fig:pipe-vel}
   \end{subfigure}
@@
     \caption{Dimensionless velocity vs.\ dimensionless wall distance}
-    \label{fig:}
+    \label{fig:pipe-nondim}
   \end{subfigure}
@@
     \caption{Normalized turbulent shear stress distribution across the pipe.}
-    \label{fig:}
+    \label{fig:pipe-stress}
   \end{subfigure}
@@
   experimental data \cite{laufer_structure_1954}.}
+  \label{fig:pipe-verification}
 \end{figure}
+
+\FloatBarrier
 
@@
 
-The \gls{BFS} flow test is based on the BFS flow experiment with Re $\approx36000$ by Driver \&
-Seegmiller \cite{driver_features_1985}. The \gls{SA} model implementation in Moltres performs
-largely similarly to the reference SA model results
-provided on the \gls{NASA} Turbulence Modeling Resource website \cite{rumsey_turbulence_nodate}.
+Driver \& Seegmiller performed the \gls{BFS} flow experiment for Re $\approx36000$ (based on the
+step height)
+\cite{driver_features_1985}. Their results serve as the reference solution for this \gls{BFS} test.
+Additionally, the \gls{NASA} Turbulence Modeling Resource website \cite{rumsey_turbulence_nodate}
+provides \gls{BFS} simulation results generated from the regular \gls{SA} model in the CFL3D
+Navier-Stokes CFD code developed at \gls{NASA} \cite{krist_cfl3d_1998}. The CFL3D \gls{SA} model
+does not contain the rotation correction scheme
+\cite{aupoix_extensions_2003, dacles-mariani_numericalexperimental_1995} present in the Moltres
+\gls{SA} model.
+
+\begin{figure}[htb!]
+  \centering
+  \includegraphics[width=0.9\columnwidth]{bfs_mesh}
+  \caption{Close-up view of the mesh for the \gls{BFS} flow test. The step is situated at $x=110$ m
+  with a height of $H=1$ m.}
+  \label{fig:bfs-mesh}
+\end{figure}
+
+The Moltres model for this test is a 2-D 160 m-long channel. The main flow direction is in the
+positive $x$ direction with the inlet and outlet on the left and right ends, respectively. The
+channel is 8 m tall before the 1 m-tall step at $x=110$ m. Figure \ref{fig:bfs-mesh} shows a
+close-up view of the mesh around the step. The dimensionless wall distance parameter $y^+$ of the
+first mesh elements along wall boundaries just prior to the step and at the end of the channel are
+0.735 and 0.555, respectively, meeting the $y^+ \lesssim 1$ requirement for properly wall-resolved
+flow. Table \ref{table:bfs} lists relevant flow parameters for the turbulent \gls{BFS} flow test.
+The Moltres \gls{BFS} flow model ran as two separate simulations for the upstream turbulent flow
+profile to fully develop from $x=0$ m to $x=104$ m and the \gls{BFS} flow profile from $x=104$ m to
+$x=160$ m. The fully-developed flow profile at $x=104$ m from the first simulation served as the
+inlet flow profile for the second simulation. Both simulations ran with adaptive time-stepping
+starting from a timestep size of 0.1 s up to a maximum timestep size of 10 s with a timestep size
+growth factor of 1.1. The upstream and \gls{BFS} simulations automatically terminated at $t=212$ s
+and $t=223$ s when Moltres detected that the flow profiles reached steady-state.
+
+Figure \ref{fig:bfs} shows the velocity magnitude and streamlines around the step at $x=110$ m. The
+streamlines illustrate the recirculation zones created by flow separation past the step.
+Figure \ref{fig:bfs-plots} shows the normalized velocity distributions at various distances from
+step, the skin friction and skin pressure coefficients along the bottom wall, and the normalized
+turbulent shear stress distributions at various distances downstream of the step. The \gls{SA}
+model with the rotation correction scheme in Moltres performs largely similarly to the reference
+\gls{SA} model results provided on the \gls{NASA} Turbulence Modeling Resource website
+\cite{rumsey_turbulence_nodate}. Compared with the reference experimental data, the Moltres
+\gls{SA} model predicts more accurate velocity (Figure \ref{fig:bfs-downstream}) and turbulent
+stress distributions (Figure \ref{fig:bfs-stress}) along $x/H=1$ downstream of the
+step than the CFL3D \gls{SA} model. This indicates that the Moltres
+\gls{SA} model reproduces the sizes and shapes of the largest and second-largest recirculation
+zones (see Figure \ref{fig:bfs}) more accurately than the CFL3D \gls{SA} model. Table
+\ref{table:bfs-reattach} further supports the prior observation since the flow reattachment length
+estimate from Moltres is more consistent with the experimental data than CFL3D. The Moltres
+\gls{SA} velocity distributions in Figure \ref{fig:bfs-downstream} are largely consistent with the
+experimental data. Lastly, the reader should
+note that this turbulent \gls{BFS} flow setup induces complex flow separation effects despite the
+apparent simplicity of the geometry. All \gls{RANS}-based model results from the \gls{NASA}
+turbulence modeling resource website \cite{rumsey_turbulence_nodate} fail to accurately reproduce
+the turbulent shear stress distribution (Figure \ref{fig:bfs-stress}) from the reference
+experimental data. Therefore, the discrepancies observed in the turbulent shear stresses from the
+\gls{SA} model are within expectations of a \gls{RANS}-based model.
+
+\begin{table}[htb]
+  \centering
+  \small
+  \caption{Relevant turbulent \gls{BFS} flow problem parameters. The $\tilde{\mu}_\text{inlet}$ value
+  at the inlet is set to fives times the $\mu$ value as recommended for the \gls{SA} model
+  \cite{spalart_one-equation_1994}.}
+  \begin{tabular}{l S}
+    \toprule
+    Property & {Value} \\
+    \midrule
+    Density, $\rho$ [kg m$^{-3}$] & 1.0 \\
+    Inlet velocity, $v_z$ [m s$^{-1}$] & 1.0 \\
+    Dynamic viscosity, $\mu$ [kg m$^{-1}$ s$^{-1}$] & 2.778e-5 \\
+    Reynolds number based on step height $H=1$ m, Re$_H$ [-] & 3.6e4 \\
+    Modified viscosity along inlet, $\tilde{\mu}_\text{inlet}$ [kg m$^{-1}$ s$^{-1}$] & 1.389e-4 \\
+    Modified viscosity along wall, $\tilde{\mu}_\text{wall}$ [kg m$^{-1}$ s$^{-1}$] & 0.0 \\
+    \bottomrule
+  \end{tabular}
+  \label{table:bfs}
+\end{table}
 
@@
     \toprule
-    & {Experimental data \cite{driver_features_1985}} & {Reference \gls{SA} model
+    & {Experimental data \cite{driver_features_1985}} & {CFL3D \gls{SA} model
   \cite{rumsey_turbulence_nodate}} & {Moltres \gls{SA} model} \\
@@
   \end{tabular}
-  \label{table:bfs}
+  \label{table:bfs-reattach}
 \end{table}
+
+\begin{figure}[htb!]
+  \centering
+  \includegraphics[width=\columnwidth]{bfs}
+  \caption{Velocity magnitude distribution and streamlines around the backward-facing step. The
+  streamlines illustrate the primary and secondary recirculation zones induced by flow past the
+  step.}
+  \label{fig:bfs}
+\end{figure}
 
@@
   \hfill
-  \begin{subfigure}[b]{0.4\columnwidth}
+  \begin{subfigure}[b]{0.38\columnwidth}
     \centering
@@
     \caption{Normalized velocity distribution at $x/H=-4$ upstream of step.}
-    \label{fig:}
+    \label{fig:bfs-upstream}
   \end{subfigure}
   \hfill
-  \begin{subfigure}[b]{0.4\columnwidth}
+  \begin{subfigure}[b]{0.38\columnwidth}
     \centering
     \includegraphics[width=\columnwidth]{bfs_downstream_vel}
-    \caption{Normalized velocity distribution at various $x/H$ locations downstream of step.}
-    \label{fig:}
+    \caption{Normalized velocity distributions downstream of step.}
+    \label{fig:bfs-downstream}
   \end{subfigure} \hfill \\
@@
     \caption{Skin friction coefficient along the bottom wall.}
-    \label{fig:}
+    \label{fig:bfs-cf}
   \end{subfigure}
   \hfill
-  \begin{subfigure}[b]{0.4\columnwidth}
+  \begin{subfigure}[b]{0.38\columnwidth}
     \centering
@@
     \caption{Skin pressure coefficient along the bottom wall.}
-    \label{fig:}
+    \label{fig:bfs-cp}
   \end{subfigure} \hfill \\
   \centering
-  \begin{subfigure}[b]{0.4\columnwidth}
+  \begin{subfigure}[b]{0.38\columnwidth}
     \centering
     \includegraphics[width=\columnwidth]{bfs_stress}
-    \caption{Normalized turbulent shear stress distribution at various $x/H$ locations downstream
+    \caption{Normalized turbulent shear stress distributions downstream
     of step.}
-    \label{fig:}
+    \label{fig:bfs-stress}
   \end{subfigure}
-  \caption{Comparison of backward facing step flow results at Re $\approx 36000$ against reference
-  experimental data and computational data from the CFL3D code.}
-  \label{fig:}
+  \caption{Comparison of backward facing step flow results against reference
+  experimental data and computational data from CFL3D. $x/H$ values are normalized horizontal
+  distances relative to the step.}
+  \label{fig:bfs-plots}
 \end{figure}
 
-\begin{figure}[htb!]
-  \centering
-  \includegraphics[width=\columnwidth]{bfs}
-  \caption{}
-  \label{fig:}
-\end{figure}
+\subsection{Summary}
 
-% !media bfs_upstream_vel.png
-%        id=bfs-upstream-vel
-%        caption=Normalized velocity distribution at $x/H$ equal -4 (upstream of step).
-%        style=width:33%;padding:10px;float:left;
-% 
-% !media bfs_downstream_vel.png
-%        id=bfs-downstream-vel
-%        caption=Normalized velocity distribution at various $x/H$ locations (downstream of step).
-%        style=width:33%;padding:10px;float:left;
-% 
-% !media bfs_cf.png
-%        id=bfs-cf
-%        caption=Skin friction coefficient along the bottom wall.
-%        style=width:33%;padding:10px;float:left;
-% 
-% !media bfs_cp.png
-%        id=bfs-cp
-%        caption=Skin pressure coefficient along the bottom wall.
-%        style=width:33%;padding:10px;margin-left:17%;float:left;
-% 
-% !media bfs_stress.png
-%        id=bfs-stress
-%        caption=Normalized turbulent shear stress distribution at various $x/H$ locations
-%        (downstream of step).
-%        style=width:33%;padding:10px;margin-right:17%;float:left;
-% 
-% !media bfs.png
-%        id=bfs
-%        caption=Close-up of the velocity magnitude and streamlines around the backward-facing step
-%        style=width:80%;float:left;margin-left:10%;margin-right:10%;
+\glspl{MSR} feature turbulent flow along various sections of the primary molten salt loop which in
+turn induce turbulent temperature and \glspl{DNP} transport. This section presented the
+implementation, verification, and validation of the \gls{SA} turbulence model
+\cite{spalart_one-equation_1994} in Moltres for turbulence
+modeling capabilities. The one-equation \gls{RANS}-based \gls{SA} model is computationally
+efficient relative to more complex \gls{RANS}-based and high-fidelity models while providing better
+accuracy than algebraic models when modeling flows with flow separation and significant streamline
+curvatures \cite{wilcox_turbulence_2006}.
+
+Verification and validation results for turbulent channel, pipe, and \gls{BFS} flow tests showed
+that the Moltres \gls{SA} model is largely consistent with reference simulation or experimental
+data in the literature. The Moltres \gls{SA} model includes a rotation correction scheme
+\cite{aupoix_extensions_2003, dacles-mariani_numericalexperimental_1995} and notably outperforms
+the reference \gls{SA} model in the turbulent \gls{BFS} flow test with more accurate velocity
+distributions and flow reattachment length estimate.

### 2024-10-01 msre-pump.tex
MSRE pump transient section

\subsection{Description of MSRE Pump Transient Tests}

The \gls{MSRE} was an experimental molten salt reactor constructed and operated at \gls{ORNL} in
the 1960s \cite{haubenreich_experience_1970}. It is a 8-MW$_{\text{th}}$, thermal-spectrum reactor
with a graphite moderator and a LiF-BeF$_2$-ZrF$_4$-UF$_4$ fuel-molten salt mixture. Under normal
power operation, the fuel salt flows out of the core and deposits heat in the heat exchanger before
being pumped back into the core. Due to
\glspl{DNP} being produced in the molten salt coolant, flow rates have significant impacts on the
\gls{DNP} distribution and reactivity. Delayed neutrons born in regions of lower neutronic
importance, e.g., at the periphery or outside the core, are more likely to be lost through neutron
leakage or parasitic absorption.

\gls{ORNL} researchers performed fuel pump start-up and coast-down transient experiments to
determine the ``transient effects of fuel flow-rate changes on reactivity''
\cite{prince_zero-power_1968}. These experiments were a part of \gls{MSRE} zero-power dynamics
tests performed in June 1965, the same month the \gls{MSRE} achieved initial criticality. Figures
\ref{fig:msre-startup} and \ref{fig:msre-coastdown} show the pump speeds and secondary salt coolant
flow rates measured during the transients. The changing flow rates caused advective changes to the
\gls{DNP} distribution and the \gls{DNF} in the reactor core.

During both experiments, the reactivity effects of the \gls{DNP} drift were measured by
allowing the flux servo controller to maintain criticality. The controller maintained core
criticality by adjusting the control rod insertion height in response to \gls{DNP} drift-induced
reactivity changes. The reactivity changes over time can be calculated by comparing the various
control rod positions (Figure \ref{fig:msre-pump-rod}) with the control rod integral worth curve
(Figure \ref{fig:msre-rod-worth}) obtained during control rod calibration experiments.

\begin{figure}[htb]
  \centering
  \begin{minipage}[t]{0.49\textwidth}
    \centering
    \includegraphics[width=\columnwidth]{msre-startup}
    \caption{Start-up pump speed and coolant flow rate \cite{prince_zero-power_1968}.}
    \label{fig:msre-startup}
  \end{minipage}
  \hfill
  \begin{minipage}[t]{0.49\textwidth}
    \centering
    \includegraphics[width=\columnwidth]{msre-coastdown}
    \caption{Coast-down pump speed and coolant flow rate \cite{prince_zero-power_1968}.}
    \label{fig:msre-coastdown}
  \end{minipage}
\end{figure}

\begin{figure}[htb]
  \centering
  \begin{minipage}[t]{0.49\textwidth}
    \centering
    \includegraphics[width=\columnwidth]{msre-transient}
    \caption{Control rod response to fuel pump start-up and coast-down
    \cite{prince_zero-power_1968}.}
    \label{fig:msre-pump-rod}
  \end{minipage}
  \hfill
  \begin{minipage}[t]{0.49\textwidth}
    \centering
    \includegraphics[width=\columnwidth]{msre-rod-worth}
    \caption{Integral control rod worth of Rod 1 \cite{prince_zero-power_1968}.}
    \label{fig:msre-rod-worth}
  \end{minipage}
\end{figure}

\subsection{Modeling Approach}

\subsection{Results}

\subsection{Summary}

\FloatBarrier

### 2024-10-04 msre-pump.tex
msre transient section additions

@@
+This section covers a \gls{VV} study based on the \gls{MSRE} zero-power pump
+start-up and coast-down pump transient tests. In collaboration with Aaron Reynolds (formerly from
+Oregon State University, we verified and validated the looped \gls{DNP} flow modeling capability of
+our respective reactor simulation software, Moltres and QuasiMolto \cite{reynolds_analysis_2023}.
+
 \subsection{Description of MSRE Pump Transient Tests}
@@
 leakage or parasitic absorption.
+
+\begin{figure}[htb]
+  \centering
+  \begin{minipage}[t]{0.49\textwidth}
+    \centering
+    \includegraphics[width=\columnwidth]{msre-startup}
+    \caption{Start-up pump speed and coolant flow rate \cite{prince_zero-power_1968}.}
+    \label{fig:msre-startup}
+  \end{minipage}
+  \hfill
+  \begin{minipage}[t]{0.49\textwidth}
+    \centering
+    \includegraphics[width=\columnwidth]{msre-coastdown}
+    \caption{Coast-down pump speed and coolant flow rate \cite{prince_zero-power_1968}.}
+    \label{fig:msre-coastdown}
+  \end{minipage}
+\end{figure}
 
@@
     \centering
-    \includegraphics[width=\columnwidth]{msre-startup}
-    \caption{Start-up pump speed and coolant flow rate \cite{prince_zero-power_1968}.}
-    \label{fig:msre-startup}
-  \end{minipage}
-  \hfill
-  \begin{minipage}[t]{0.49\textwidth}
-    \centering
-    \includegraphics[width=\columnwidth]{msre-coastdown}
-    \caption{Coast-down pump speed and coolant flow rate \cite{prince_zero-power_1968}.}
-    \label{fig:msre-coastdown}
-  \end{minipage}
-\end{figure}
-
-\begin{figure}[htb]
-  \centering
-  \begin{minipage}[t]{0.49\textwidth}
-    \centering
     \includegraphics[width=\columnwidth]{msre-transient}
@@
 
+\subsection{Description of QuasiMolto}
+
+QuasiMolto \cite{reynolds_analysis_2023} is an open-source multiphysics research code for
+simulating circulating fuel reactor kinetics. It uses the multilevel \gls{QD} methodology described
+in Section \ref{sec:lit-tc} to model multigroup neutron transport with \gls{DNP} flow and
+temperature advection-diffusion. Consisting of three levels of neutron calculations, the highest
+level solves the multigroup high-order transport equations for Eddington and boundary factors which
+populate previously unknown quantities in the mid-level multigroup low-order \gls{QD} equations.
+In turn, the mid-level equations solve for group-collapsed Eddington and boundary factors for the
+low-level effective low-order transport equations. The effective low-order transport equations
+couple with the \gls{DNP} and temperature governing equations for multiphysics reactivity feedback
+from temperature and \gls{DNP} flow. Readers may refer to the linked reference
+\cite{reynolds_analysis_2023} for more details on QuasiMolto.
+
+QuasiMolto applies the simple corner balance method
+\cite{adams_subcell_1997} for spatial discretization.
+For this study, the QuasiMolto \gls{MSRE} model employed $S_2$ neutron transport to facilitate
+fair comparisons with the neutron diffusion method in Moltres. For time-dependent \gls{DNP} flow
+modeling, QuasiMolto applies implicit Euler scheme for the reaction
+and diffusion terms, and explicit Superbee flux limiter scheme for the advection term.
+
 \subsection{Modeling Approach}
 
-\subsection{Results}
+For this study, we adopted a 2-D axisymmetric (R-Z) model of the \gls{MSRE} consisting of
+alternating vertical regions of molten salt and graphite moderator. This is the same \gls{MSRE}
+model Reynolds developed for his PhD thesis \cite{reynolds_multilevel_2020} and a related
+journal publication \cite{reynolds_analysis_2023}. Figure \ref{fig:pump-geom} depicts the 2-D
+axisymmetric \gls{MSRE} model. The fuel salt channels are 1.25-cm wide and spaced at 5-cm intervals
+in the graphite matrix. The fuel salt volume fraction of this model is 0.236, differing
+slightly to the \gls{MSRE} design specification of 0.225 \cite{robertson_msre_1965}. The geometry
+mesh is uniform with a width of $\Delta r=0.078125$ cm and height of $\Delta z=5$ cm.
+Vacuum boundary conditions for the neutron flux apply along the top, bottom, and right boundaries.
+
+The salt flows
+upwards at a spatially uniform velocity rate of 18.085 cm s$^{-1}$. \glspl{DNP} flow out of the
+core model and through a separate 285-cm long 1-D pipe model before being reintroduced through the
+bottom core channel inlets. The 1-D pipe length maintains the overall salt circulation time of
+25.2 s from the \gls{MSRE} design specification \cite{robertson_msre_1965}. Note that the salt
+velocity and 1-D pipe length in this study differs from the \gls{MSRE} model described in Reynolds'
+paper \cite{reynolds_analysis_2023}. We reduced the velocity to compensate for the absence of upper
+and lower core plena by extending salt residence time in the salt-graphite lattice region.
+
+\begin{figure}[htb]
+  \centering
+  \includegraphics[width=0.6\columnwidth]{msre-2d}
+  \caption{2-D axisymmetric model of the \gls{MSRE}.The gray and white regions represent fuel salt
+  and graphite, respectively. The leftmost fuel salt channel falls along the axis of symmetry and
+  is half-width. Not drawn to scale.}
+  \label{fig:pump-geom}
+\end{figure}
+
+Reynolds generated group constants for two neutron energy groups and six \gls{DNP} groups using the
+NEWT transport module within the SCALE code system \cite{rearden_scale_2018} and the ENDF/B-VII
+nuclear data library \cite{chadwick_endf/b-vii.1_2011}. The SCALE model geometry comprised a
+representative unit cell geometry
+consisting of a circular fuel region in the center of a square graphite region and reflective
+boundary conditions on all boundaries. The unit cell dimensions preserved material compositions
+and salt-to-graphite ratios from the \gls{MSRE} design specifications.
+
+Prior to the pump transient simulations, we performed code-to-code verification by comparing the
+$k_\text{eff}$ and neutron flux and \gls{DNP} distributions of the 2-D \gls{MSRE} model under
+static, no salt flow conditions. We sampled the neutron flux and \gls{DNP} distributions on
+mesh corners and edge centers along the centerline ($r=0$ cm) and midplane ($z=85$ cm) of the
+reactor geometry.
+
+In lieu of reactivity control using a control rod in the pump
+transient simulations, QuasiMolto and Moltres solved their respective
+two-group $k$-eigenvalue neutronics models at every time-step. The $1/k_\text{eff}$ scaling factor
+in the fission neutron source term keeps the simulation at a pseudo-critical state. The \gls{DNP}
+source term is also scaled by $1/k_\text{eff}$ at every time-step. We interpolated the \gls{MSRE}
+salt flow velocity data in Figures \ref{fig:msre-startup} and \ref{fig:msre-coastdown} to apply in
+our pump start-up and coast-down simulations. The pump start-up simulation starts from a static
+state with no salt flow while the pump coast-down simulation starts from a steady flow state with
+operating salt flow.
+
+\subsection{Results \& Discussion}
+
+We start with a comparison of the Moltres and QuasiMolto \gls{MSRE} numerical models under static
+and steady salt flow conditions. Table \ref{table:msre-pump-keff} lists the $k_\text{eff}$
+estimates under static and steady flow conditions, and the reactivity changes due to \gls{DNP}
+flow. Moltres and QuasiMolto show excellent agreement with 22 and 21 pcm absolute differences in
+$k_\text{eff}$ under static and steady flow conditions, respectively. Consequently, the
+$\Delta\rho$ estimates from \gls{DNP} flow fall within 1 pcm difference of each other.
+
+\begin{table}[htb]
+  \centering
+  \caption{Multiplication factors $k_\text{eff}$ under static and steady salt flow conditions, and
+  reactivity changes $\Delta\rho$ due to \gls{DNP} flow from the Moltres and QuasiMolto \gls{MSRE}
+  models.}
+  \begin{tabular}{l S S S}
+    \toprule
+    \multirow{2}{*}{Code} & \multicolumn{2}{c}{$k_\text{eff}$} & {$\Delta\rho$ due to} \\
+                          & {Static} & {Steady flow} & {\gls{DNP} flow} \\
+                          \cmidrule(r){1-1} \cmidrule(rl){2-3} \cmidrule(l){4-4}
+    Moltres & 1.05625 & 1.05311 & -282.4 \\
+    QuasiMolto & 1.05603 & 1.05290 & -281.7 \\
+    \bottomrule
+  \end{tabular}
+  \label{table:msre-pump-keff}
+\end{table}
+
+Moving on to comparing spatial distributions, we start with the centerline neutron flux
+distributions. Figure \ref{fig:centerline-flux-dist} shows the normalized group 1 and 2 neutron
+fluxes exhibiting nearly perfect overlap. The flux ratio distributions in Figure
+\ref{fig:centerline-flux-ratio} illustrate the strong agreement between Moltres and QuasiMolto at
+all sampled locations with the relative differences falling below 0.8\%. We attribute
+the oscillations in the flux ratios to differences in spatial discretization schemes. Moltres
+uses \gls{FEM} with 1st-order flux values computed at the mesh corners. QuasiMolto uses
+the simple corner balance method which subdivides each mesh element into four equal subcells and
+computes volume-averaged flux values in each subcell. As such, Moltres and QuasiMolto require
+interpolation for sampling mesh edge center and mesh corner values, respectively. The flux ratios
+sampled at the top and bottom boundaries show the greatest differences of up to 0.8\%. We attribute
+these differences at the boundaries to differences in neutronics methods. The neutron diffusion
+method vacuum boundary condition in Moltres depends on the scalar flux gradient (1-st
+order derivative) while the $S_2$ method vacuum boundary condition in QuasiMolto depends on the
+angular flux value. Combined with the different spatial discretization schemes, flux differences at
+the boundary are expected in numerical calculations. Nevertheless, the overall ratio magnitudes
+remain small, thereby supporting code-to-code consistency between Moltres and QuasiMolto.
+
+\begin{figure}[htb]
+  \centering
+  \begin{subfigure}[b]{0.48\columnwidth}
+    \centering
+    \includegraphics[width=\columnwidth]{centerline_flux}
+    \caption{Normalized centerline neutron fluxes.}
+    \label{fig:centerline-flux-dist}
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}[b]{0.48\columnwidth}
+    \centering
+    \includegraphics[width=\columnwidth]{centerline_flux_ratio}
+    \caption{Ratio of centerline neutron fluxes.}
+    \label{fig:centerline-flux-ratio}
+  \end{subfigure}
+  \caption{Centerline neutron flux distributions and ratios comparing QuasiMolto and Moltres
+  \gls{MSRE} models under static conditions.}
+  \label{fig:centerline-flux}
+\end{figure}
+
+\begin{figure}[htb]
+  \centering
+  \begin{subfigure}[b]{0.48\columnwidth}
+    \centering
+    \includegraphics[width=\columnwidth]{midplane_flux}
+    \caption{Normalized midplane neutron fluxes.}
+    \label{fig:centerline-flux-dist}
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}[b]{0.48\columnwidth}
+    \centering
+    \includegraphics[width=\columnwidth]{midplane_flux_ratio}
+    \caption{Ratio of midplane neutron fluxes.}
+    \label{fig:centerline-flux-ratio}
+  \end{subfigure}
+  \caption{Midplane neutron flux distributions and ratios comparing QuasiMolto and Moltres
+  \gls{MSRE} models under static conditions.}
+  \label{fig:centerline-flux}
+\end{figure}
+
+\begin{figure}[htb]
+  \centering
+  \begin{subfigure}[b]{0.48\columnwidth}
+    \centering
+    \includegraphics[width=\columnwidth]{centerline_pre}
+    \caption{Normalized centerline \gls{DNP} distributions.}
+    \label{fig:centerline-pre-dist}
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}[b]{0.48\columnwidth}
+    \centering
+    \includegraphics[width=\columnwidth]{centerline_pre_ratio}
+    \caption{Ratio of centerline \gls{DNP} distributions.}
+    \label{fig:centerline-pre-ratio}
+  \end{subfigure}
+  \caption{Centerline \gls{DNP} distribution from Moltres and ratios comparing QuasiMolto and
+  Moltres models under static \gls{MSRE} conditions. The ratio curves for every \gls{DNP} group
+  exhibit nearly perfect overlap.}
+  \label{fig:centerline-pre}
+\end{figure}
+
+\begin{figure}[htb]
+  \centering
+  \begin{subfigure}[b]{0.48\columnwidth}
+    \centering
+    \includegraphics[width=\columnwidth]{midplane_pre}
+    \caption{Normalized midplane \gls{DNP} distributions.}
+    \label{fig:midplane-pre-dist}
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}[b]{0.48\columnwidth}
+    \centering
+    \includegraphics[width=\columnwidth]{midplane_pre_ratio}
+    \caption{Ratio of midplane \gls{DNP} distributions.}
+    \label{fig:midplane-pre-ratio}
+  \end{subfigure}
+  \caption{Midplane \gls{DNP} distribution from Moltres and ratios comparing QuasiMolto and Moltres
+  models under static \gls{MSRE} conditions. \glspl{DNP} exist only in the fuel channel regions.
+  The ratio curves for every \gls{DNP} group exhibit nearly perfect overlap.}
+  \label{fig:midplane-pre}
+\end{figure}
+
+\begin{figure}[htb]
+  \centering
+  \includegraphics[width=0.9\columnwidth]{start-up-v2-reactivity}
+  \caption{Reactivity (relative to static no-flow conditions) during the \gls{MSRE} pump start-up
+  transient. The Moltres and QuasiMolto models agree closely with each other. The numerical models
+  report greater peak and final reactivity change than the \gls{ORNL} \gls{MSRE} experimental
+  data.}
+  \label{fig:start-up-reactivity}
+\end{figure}
+
+\begin{figure}[htb]
+  \centering
+  \includegraphics[width=0.9\columnwidth]{coast-down-v2-reactivity}
+  \caption{Reactivity (relative to static no-flow conditions) during the \gls{MSRE} pump coast-down
+  transient. The Moltres and QuasiMolto models agree closely with each other. The numerical models
+  report greater reactivity change than the \gls{ORNL} \gls{MSRE} experimental data throughout the
+  transient.}
+  \label{fig:coast-down-reactivity}
+\end{figure}
 

### 2024-10-06 benchmark.tex
verification chapter edits

@@
-\section{CNRS Benchmark}
+\section{CNRS Benchmark Verification Study} \label{sec:cnrs-benchmark}
 
@@
 cross-section libraries, or actual errors in the software. I adapted the
-contents of this chapter from my journal article titled
-``\textit{Verification of Moltres for Multiphysics Simulations of Fast-Spectrum
-Molten Salt Reactors}'' and published in the Annals of Nuclear Energy journal
+contents of this chapter from my journal article in the Annals of Nuclear Energy, titled
+\textit{Verification of Moltres for Multiphysics Simulations of Fast-Spectrum
+Molten Salt Reactors}
 \cite{park_verification_2022}.

### 2024-10-06 moltres.tex
turbulence verification section
verification chapter edits

@@
 
-This chapter provides an overview of Moltres as a multiphysics
+This section provides an overview of Moltres as a multiphysics
 simulation software for molten salt reactors. 
@@
 
-Moltres relies on MOOSE's \texttt{Heat} \texttt{Conduction} and
+Moltres relies on MOOSE's \texttt{Heat} \texttt{Transfer} and
 \texttt{Navier-Stokes} physics modules for its thermal-hydraulics modeling

### 2024-10-06 msre-pump.tex
verification chapter edits

@@
 start-up and coast-down pump transient tests. In collaboration with Aaron Reynolds (formerly from
-Oregon State University, we verified and validated the looped \gls{DNP} flow modeling capability of
+Oregon State University) we verified and validated the looped \gls{DNP} flow modeling capability of
 our respective reactor simulation software, Moltres and QuasiMolto \cite{reynolds_analysis_2023}.
+This study offers a more formal assessment of \gls{DNP} flow modeling in Moltres than previous
+work described in Section \ref{sec:moltres-previous}, by verifying Moltres against QuasiMolto in
+the context of a past experiment on the \gls{MSRE}. Moltres and QuasiMolto also benefited from this
+collaborative study through the discovery and correction of previously unnoticed bugs.
 
@@
 
-\subsection{Description of QuasiMolto}
+\subsection{Description of Reactor Software}
+
+Section \ref{sec:moltres-description} provides general details on the neutronics and \gls{DNP}
+models implemented with \gls{FEM} in Moltres. In this study, the neutron flux and \gls{DNP}
+variables were approximated using continuous 1st-order Lagrange and discontinuous 2nd-order
+monomial basis functions, respectively. The calculations for the neutron flux in the core,
+\gls{DNP} in the core, and \gls{DNP} outside the core were tightly coupled using the fixed point
+iteration capability available through the MOOSE MultiApp system \cite{gaston_physics-based_2015}.
 
@@
 
-QuasiMolto applies the simple corner balance method
-\cite{adams_subcell_1997} for spatial discretization.
+QuasiMolto applies the simple corner balance method \cite{adams_subcell_1997} for spatial
+discretization. The simple corner balance method is a subcell, finite-volume method that subdivides
+each mesh element into four equal subcells on each mesh corner.
 For this study, the QuasiMolto \gls{MSRE} model employed $S_2$ neutron transport to facilitate
@@
 
-For this study, we adopted a 2-D axisymmetric (R-Z) model of the \gls{MSRE} consisting of
+For this study, we adopted a 2-D axisymmetric (R-Z) geometry of the \gls{MSRE} consisting of
 alternating vertical regions of molten salt and graphite moderator. This is the same \gls{MSRE}
-model Reynolds developed for his PhD thesis \cite{reynolds_multilevel_2020} and a related
+geometry from Reynolds' PhD thesis \cite{reynolds_multilevel_2020} and
 journal publication \cite{reynolds_analysis_2023}. Figure \ref{fig:pump-geom} depicts the 2-D
-axisymmetric \gls{MSRE} model. The fuel salt channels are 1.25-cm wide and spaced at 5-cm intervals
+axisymmetric \gls{MSRE} model. The fuel salt channels are 1.25-cm-wide and spaced at 5 cm intervals
 in the graphite matrix. The fuel salt volume fraction of this model is 0.236, differing
@@
 upwards at a spatially uniform velocity rate of 18.085 cm s$^{-1}$. \glspl{DNP} flow out of the
-core model and through a separate 285-cm long 1-D pipe model before being reintroduced through the
+core model and through a separate 285-cm-long 1-D pipe model before being reintroduced through the
 bottom core channel inlets. The 1-D pipe length maintains the overall salt circulation time of
@@
 static, no salt flow conditions. We sampled the neutron flux and \gls{DNP} distributions on
-mesh corners and edge centers along the centerline ($r=0$ cm) and midplane ($z=85$ cm) of the
-reactor geometry.
+mesh corners and edge centers along the centerline ($r=0$ cm) and on mesh edge centers along
+the midplane ($z=85$ cm) of the reactor geometry.
 
 In lieu of reactivity control using a control rod in the pump
-transient simulations, QuasiMolto and Moltres solved their respective
+transient simulations, Moltres and QuasiMolto solved their respective
 two-group $k$-eigenvalue neutronics models at every time-step. The $1/k_\text{eff}$ scaling factor
@@
 state with no salt flow while the pump coast-down simulation starts from a steady flow state with
-operating salt flow.
+operating salt flow. Both simulations used a fixed time-step size of 0.1 s.
 
@@
 $k_\text{eff}$ under static and steady flow conditions, respectively. Consequently, the
-$\Delta\rho$ estimates from \gls{DNP} flow fall within 1 pcm difference of each other.
+$\Delta\rho$ estimates from \gls{DNP} flow fall within less than 1 pcm difference of each other.
 
@@
 distributions. Figure \ref{fig:centerline-flux-dist} shows the normalized group 1 and 2 neutron
-fluxes exhibiting nearly perfect overlap. The flux ratio distributions in Figure
-\ref{fig:centerline-flux-ratio} illustrate the strong agreement between Moltres and QuasiMolto at
-all sampled locations with the relative differences falling below 0.8\%. We attribute
+fluxes exhibiting nearly perfect overlap. The QuasiMolto-to-Moltres flux ratio distributions in
+Figure \ref{fig:centerline-flux-ratio} illustrate the strong agreement between Moltres and
+QuasiMolto at all sampled locations with the relative differences falling below 0.8\%. We attribute
 the oscillations in the flux ratios to differences in spatial discretization schemes. Moltres
@@
 the boundary are expected in numerical calculations. Nevertheless, the overall ratio magnitudes
-remain small, thereby supporting code-to-code consistency between Moltres and QuasiMolto.
+remain small, thereby showing that Moltres and QuasiMolto are consistent with each other.
 
-\begin{figure}[htb]
+\begin{figure}[t]
   \centering
@@
   \label{fig:centerline-flux}
-\end{figure}
-
-\begin{figure}[htb]
-  \centering
   \begin{subfigure}[b]{0.48\columnwidth}
@@
     \caption{Normalized midplane neutron fluxes.}
-    \label{fig:centerline-flux-dist}
+    \label{fig:midplane-flux-dist}
   \end{subfigure}
@@
     \caption{Ratio of midplane neutron fluxes.}
-    \label{fig:centerline-flux-ratio}
+    \label{fig:midplane-flux-ratio}
   \end{subfigure}
@@
   \gls{MSRE} models under static conditions.}
-  \label{fig:centerline-flux}
+  \label{fig:midplane-flux}
 \end{figure}
 
-\begin{figure}[htb]
+The midplane normalized neutron flux and flux ratio distributions in Figures
+\ref{fig:midplane-flux-dist} and \ref{fig:midplane-flux-ratio} also show consistency between
+Moltres and QuasiMolto through nearly perfect flux overlap and small flux ratios. Minor
+discretization errors arising from comparing continuous (Moltres) and discontinuous (QuasiMolto)
+fluxes contribute to discrete jumps in the flux ratios at the salt-graphite interfaces. Despite
+these deviations,
+the maximum midplane flux ratio of 1.001 remains much lower than the maximum centerline flux ratio
+of 1.008. Fine mesh resolution along the radial direction helps the vacuum boundary flux values
+converge. The centerline and midplane \gls{DNP} distributions and ratios in Figures
+\ref{fig:centerline-pre} and \ref{fig:midplane-pre} continue to show good agreement between Moltres
+and QuasiMolto. Figures \ref{fig:centerline-pre-dist} and \ref{fig:midplane-pre-dist} depict only
+the normalized centerline and midplane \gls{DNP} distributions from Moltres to reduce visual
+clutter. Combining all observations thus far, Moltres and QuasiMolto are consistent with each other
+under static and steady salt flow conditions.
+
+\begin{figure}[t]
   \centering
@@
   \caption{Centerline \gls{DNP} distribution from Moltres and ratios comparing QuasiMolto and
-  Moltres models under static \gls{MSRE} conditions. The ratio curves for every \gls{DNP} group
-  exhibit nearly perfect overlap.}
+  Moltres models under static \gls{MSRE} conditions. The ratio curves overlap perfectly.}
   \label{fig:centerline-pre}
-\end{figure}
-
-\begin{figure}[htb]
-  \centering
   \begin{subfigure}[b]{0.48\columnwidth}
@@
   models under static \gls{MSRE} conditions. \glspl{DNP} exist only in the fuel channel regions.
-  The ratio curves for every \gls{DNP} group exhibit nearly perfect overlap.}
+  The ratio curves overlap perfectly.}
   \label{fig:midplane-pre}
@@
 
-\begin{figure}[htb]
+\begin{figure}[t]
   \centering
-  \includegraphics[width=0.9\columnwidth]{start-up-v2-reactivity}
-  \caption{Reactivity (relative to static no-flow conditions) during the \gls{MSRE} pump start-up
-  transient. The Moltres and QuasiMolto models agree closely with each other. The numerical models
-  report greater peak and final reactivity change than the \gls{ORNL} \gls{MSRE} experimental
-  data.}
+  \includegraphics[width=0.8\columnwidth]{start-up-v2-reactivity}
+  \caption{Reactivity loss (relative to static no-flow conditions) during the \gls{MSRE} pump
+  start-up transient. The Moltres and QuasiMolto models agree closely with each other. The
+  numerical models report greater peak and final reactivity change than the \gls{ORNL} \gls{MSRE}
+  experimental data.}
   \label{fig:start-up-reactivity}
@@
 
-\begin{figure}[htb]
+Figure \ref{fig:start-up-reactivity} shows the change in reactivity loss relative to the static
+condition during the pump start-up transient. Moltres and QuasiMolto report nearly identical
+responses throughout the transient. Both codes and the \gls{ORNL} experimental data indicate a
+steep increase in reactivity loss caused by the loss of \glspl{DNP} flowing out of the core with no
+\glspl{DNP} flowing in. The reactivity loss peaks and plateaus starting around $t=12$ s for Moltres
+and QuasiMolto and $t=13$ s for the \gls{ORNL} data. From the \gls{ORNL} data, the reactivity loss
+drops after 2 s and stabilizes at around 225 pcm. From $t=30$ s onwards, the \gls{ORNL} reactivity
+readings appear to be noisy with no obvious trend. In contrast, Moltres and
+QuasiMolto report reactivity oscillations caused by the initial pocket of \gls{DNP} reentering the
+core every salt recirculation cycle of 25.2 s. The oscillations eventually dissipate as the
+\glspl{DNP} in the pocket progressively decay away over time. Reactivity loss from both codes
+level out at around 282 pcm which is consistent with the reactivity change due to \gls{DNP} flow
+from the steady salt flow case in Table \ref{table:msre-pump-keff}.
+
+Three key differences stand out between the numerical results and \gls{ORNL} data. 1) The
+\gls{ORNL} data logged a smaller reactivity loss of 225 pcm compared to 282 pcm from the
+numerical results. 2) The \gls{ORNL} data logged a shorter reactivity loss plateau on the first
+peak than the numerical results. 3) The \gls{ORNL} data did not log significant reactivity loss
+oscillations after the first peak. These differences in the reactivity loss profile stem mainly
+from the absence of upper and lower plena in the numerical model geometry (Figure
+\ref{fig:pump-geom}).
+
+Since the upper and lower plena comprise largely of salt, they retain some of the \glspl{DNP} which
+we neglected in our numerical model geometry. Robertson \cite{robertson_msre_1965} reports that
+salt residence times in the plena are approximately 4 s each. Therefore, Moltres and QuasiMolto
+reported greater reactivity loss under steady flow conditions. With a combined salt residence time
+of 17.1 s in the lower plenum, active core, and upper
+plenum regions out of a full recirculation time of 25.2 s \cite{robertson_msre_1965}, there should
+not be reactivity plateaus during the pump start-up transient. Lastly, Moltres and QuasiMolto
+applied uniform salt flow and neglected turbulent mixing of \gls{DNP} in the plena and out-of-core
+components. Without mixing, the numerical models preserved the initially static pocket of
+\glspl{DNP} as it advected around the salt loop and produced the reactivity oscillations observed
+in Figure \ref{fig:start-up-reactivity}. Conversely, the plateaus serve as
+useful reference features for verifying advection modeling accuracy for code-to-code verification
+in the context of \gls{MSR} simulations. This pump transient setup can be used to identify
+advection schemes that suffer from excessive numerical diffusion or dispersion effects. The fact
+that Moltres and QuasiMolto produced nearly identical results, despite significant differences in
+their numerical schemes, lends credibility to both sets of results for the purpose of code
+verification.
+
+\begin{figure}[b]
   \centering
-  \includegraphics[width=0.9\columnwidth]{coast-down-v2-reactivity}
-  \caption{Reactivity (relative to static no-flow conditions) during the \gls{MSRE} pump coast-down
-  transient. The Moltres and QuasiMolto models agree closely with each other. The numerical models
-  report greater reactivity change than the \gls{ORNL} \gls{MSRE} experimental data throughout the
-  transient.}
+  \includegraphics[width=0.8\columnwidth]{coast-down-v2-reactivity}
+  \caption{Reactivity loss (relative to static no-flow conditions) during the \gls{MSRE} pump
+  coast-down transient. The Moltres and QuasiMolto models agree closely with each other. The
+  numerical models report greater reactivity change than the \gls{ORNL} \gls{MSRE} experimental
+  data throughout the transient.}
   \label{fig:coast-down-reactivity}
@@
 
+Compared to the pump start-up transient, the reactivity loss over time for the pump coast-down
+transient shows a simpler, monotonic decline as shown in Figure \ref{fig:coast-down-reactivity}.
+At the start of the transient, reactivity loss remains stable for 2 s due to inertia in salt flow.
+Following this, the reactivity loss profile transitions into a steep exponential decline as
+\glspl{DNP} accumulate in the increasingly stagnant salt. Other than the mismatch in reactivity
+loss magnitude, Moltres and QuasiMolto reproduce the expected behavior observed in the \gls{ORNL}
+pump coast-down data.
+
 \subsection{Summary}
 
+In collaboration with the developer of QuasiMolto \cite{reynolds_analysis_2023}, we performed a
+\gls{VV} study based on the \gls{MSRE}
+zero-power pump start-up and coast-down transient. This study was designed to assess
+the looped \gls{DNP} flow modeling capabilities in Moltres and QuasiMolto reactor simulation
+software. We used a 2-D axisymmetric model of the \gls{MSRE} to numerically model the pump
+transients.
+
+Moltres and QuasiMolto were consistent with each other in $k$-eigenvalue neutronics simulations
+under static and steady salt flow conditions prior to pump transient simulations. The
+$k_\text{eff}$ estimates from both codes were within 22 and 21 pcm of each other under the static
+and steady flow conditions and the calculated reactivity loss due to \gls{DNP} flow agreed within
+less than 1 pcm. The centerline and midplane flux and \gls{DNP} distributions showed nearly perfect
+overlap with relative differences falling under 0.8\%.
+
+Moltres and QuasiMolto remained consistent with each other in the pump start-up and
+coast-down scenarios. Both codes reproduced most of the trends in reactivity loss from \gls{DNP}
+flow observed in the \gls{ORNL}
+\gls{MSRE} experimental data. Several differences between the numerical and experimental data arose
+primarily due to the absence of upper and lower plena in the numerical model and the omission of
+turbulent mixing. Nevertheless, the numerical results for the pump start-up transient
+simulation captured distinct reactivity oscillations corresponding to the \gls{MSRE} salt
+recirculation period. These oscillations served as valuable reference points for verifying \gls{DNP}
+advection modeling accuracy between Moltres and QuasiMolto. Researchers looking to verify
+their \gls{MSR} simulation software may use the results of this study as a benchmark.
+
 \FloatBarrier

### 2024-10-06 verification.tex
turbulence verification section
MSRE pump transient section
verification chapter edits

@@
+Section \ref{sec:msr-vv} identifies salt flow-induced \gls{DNP} drift and the strong coupling
+between neutronics and thermal-hydraulics as two critical multiphysics interactions that must be
+accurately modeled in \gls{MSR} simulation software. This chapter describes Moltres and
+subsequently covers several \gls{VV} studies for verifying and validating specific modeling
+capabilities in Moltres which relate to the aforementioned multiphysics interactions.
+
+Section \ref{sec:moltres-description} details the neutronics, thermal-hydraulics, and multiphysics
+coupling schemes in Moltres for \gls{MSR} multiphysics simulations. Following the model
+descriptions, the section also summarizes and assesses previous \gls{MSR} analyses
+performed with Moltres. Section \ref{sec:cnrs-benchmark} presents verification results from Moltres
+for the CNRS benchmark study. Section \ref{sec:msre-pump} presents key findings from a
+collaborative \gls{VV} study based on the \gls{MSRE} pump transient tests conducted in the 1960s.
+Section \ref{sec:turbulence} describes the Spalart-Allmaras turbulence model implementation in
+Moltres and presents the model \gls{VV} results against common wall-bounded turbulent flow case
+studies from literature.
+
 \input{moltres}
 \input{benchmark}
-\section{MSRE Zero-Power Pump Experiments}
-\section{Turbulence Model Verification}
+
+\section{MSRE Zero-Power Pump Experiments} \label{sec:msre-pump}
+
+\input{msre-pump}
+
+\section{Moltres Spalart-Allmaras Turbulence Model Verification} \label{sec:turbulence}
+
+\input{turbulence}
+
+\section{Summary}

### 2024-10-30 verification.tex
Fix minor errors

--- 2024-07-22 verification.tex
@@
+Section \ref{sec:msr-vv} identified salt flow-induced \gls{DNP} drift and the strong coupling
+between neutronics and thermal-hydraulics as two critical multiphysics interactions that must be
+accurately modeled in \gls{MSR} simulation software. This chapter describes Moltres and
+subsequently presents several \gls{VV} results and analyses for verifying and validating specific
+modeling capabilities in Moltres which relate to the aforementioned multiphysics interactions.
+Section \ref{sec:moltres-description} details the neutronics, thermal-hydraulics, and multiphysics
+coupling schemes in Moltres for \gls{MSR} multiphysics simulations. Following the model
+descriptions, the section also summarizes and assesses previous \gls{MSR} analyses
+performed with Moltres. Section \ref{sec:cnrs-benchmark} presents verification results from Moltres
+for the CNRS benchmark study. Section \ref{sec:msre-pump} presents key findings from a
+collaborative \gls{VV} study based on the \gls{MSRE} pump transient tests conducted in the 1960s.
+Section \ref{sec:turbulence} describes the \gls{SA} turbulence model implementation in
+Moltres and presents the model \gls{VV} results against common wall-bounded turbulent flow case
+studies from literature.
+
 \input{moltres}
 \input{benchmark}
-\section{MSRE Zero-Power Pump Experiments}
-\section{Turbulence Model Verification}
+
+\input{msre-pump}
+
+\input{turbulence}
+
+\FloatBarrier
+
+\section{Summary}
+
+Moltres is a highly scalable \gls{MOOSE}-based multiphysics application for advanced reactor
+simulations with a strong focus on \glspl{MSR} modeling. It features the multigroup neutron
+diffusion and thermal-hydraulics models, along with robust and flexible coupling capabilities for
+modeling multiphysics interactions in reactors.
+
+This chapter presented two \gls{VV} studies aimed at primarily assessing Moltres for modeling
+two key multiphysics interactions in \glspl{MSR}, namely looped \gls{DNP} flow modeling and the
+strong coupling between neutronics and thermal-hydraulics. The first study is the CNRS numerical
+benchmark for fast-spectrum \gls{MSR} modeling. The simulation results showed that Moltres is
+consistent with the four \gls{MSR} simulation software that participated in the CNRS benchmark
+study \cite{tiberga_results_2020} over a wide range of neutronics and thermal-hydraulics simulation
+metrics. Moltres underpredicts the temperature along the top boundary due to unphysical,
+discontinuous velocity boundary conditions defined at the top corners of the lid-driven cavity
+flow setup. Nevertheless, the temperature deviations are limited to the top boundary and Moltres
+is otherwise in excellent agreement with the benchmark participants.
+
+The second study is a collaborative work with the developer of QuasiMolto for assessing \gls{DNP}
+flow modeling in the \gls{MSRE} zero-power pump start-up and coast-down transient tests conducted
+in the 1960s \cite{prince_zero-power_1968}. Despite the widely different numerical schemes between
+Moltres and QuasiMolto, both codes were consistent with each other under all considered scenarios.
+$k_\text{eff}$ estimates from both codes were 22 pcm and 21 pcm apart under static and steady salt
+flow conditions. Consequently, the calculated reactivity loss due to \gls{DNP} flow agreed within
+less than 1 pcm. Spatial flux and \gls{DNP} distributions showed nearly perfect overlap with less
+than 0.8\% relative difference between Moltres and QuasiMolto. Moltres and QuasiMolto remained
+consistent with each other in the pump start-up and coast-down transients. We attributed
+discrepancies between the numerical results and \gls{ORNL} experimental data to the absence of
+upper and lower plena in the numerical model. The numerical study remains a useful benchmark for
+assessing \gls{DNP} advection modeling in \gls{MSR} simulation software due to clearly defined
+salt recirculation behavior reflected in the results.
+
+Lastly, this chapter presented the implementation and verification of the \gls{SA} model
+\cite{spalart_one-equation_1994} in Moltres for enabling wall-bounded turbulent molten salt flow
+modeling. The \gls{SA} model includes a rotation correction scheme
+\cite{aupoix_extensions_2003, dacles-mariani_numericalexperimental_1995} for improved accuracy when
+modeling turbulent flow with curved streamlines. The \gls{SA} model performed well under the
+turbulent channel flow verification test \cite{moser_direct_1999}, the turbulent pipe flow
+validation test \cite{laufer_structure_1954}, and the \gls{BFS} flow validation test
+\cite{driver_features_1985}.

### 2024-10-30 cnrs-benchmark/benchmark.tex

@@
 2m$\times$2m square cavity filled with LiF-BeF$_2$-UF$_4$ molten salt at an
-initial temperature of 900K \cite{tiberga_results_2020}.
+initial temperature of 900 K \cite{tiberga_results_2020}.
 Standard vacuum boundary conditions apply for neutron flux along all

### 2024-10-30 moltres.tex

@@
 This section provides an overview of Moltres as a multiphysics
-simulation software for molten salt reactors. 
+simulation software for \glspl{MSR}. 
 Section \ref{sec:moltres-features}
@@
 This section describes the various physics models available in Moltres to model
-coupled neutronics and thermal-hydraulics in \glspl{MSR}. Section \ref{sec:nts}
-discusses the neutronics model in Moltres, Section \ref{sec:th} discusses
-the thermal-hydraulics model, and Section \ref{sec:moltres-loop} discusses the
-external loop model for \gls{DNP} looping and heat removal via external heat
-exchangers.
+coupled neutronics and thermal-hydraulics in \glspl{MSR}.
 
@@
 
-\paragraph{Critical Assessment} \label{sec:msre-critique}
+%\paragraph{Critical Assessment} \label{sec:msre-critique}
 
@@
 
-\paragraph{Critical Assessment} \label{sec:msfr-critique}
+%\paragraph{Critical Assessment} \label{sec:msfr-critique}
 

### 2024-11-04 hybrid-nts.tex
Add 3-D results content

--- 2024-08-23 msre.tex
@@
 to quantify errors arising from neutron energy group discretization and the scattering cross
-section simplifications.
+section approximations.
 
@@
 
-\subsection{2-D Neutronics Model Setup}
+\subsection{2-D Neutronics Model Setup} \label{sec:2d-model-setup}
 
@@
   \item Set channel pitch, width, and length to 2, 0.2, and 1.2 inches, respectively.
-  \item Homogenized the INOR, graphite, and molten salt regions in the sample basket.
+  \item Homogenized the INOR-8, graphite, and molten salt regions in the sample basket.
   \item Modified the outer edge of the core graphite to be circular with no jagged edges.
@@
 
-\subsubsection{2-D quarter-core results}
+\subsubsection{2-D Quarter-Core}
 
@@
 
-\subsubsection{2-D full-core results}
+\subsubsection{2-D Full-Core}
 
@@
 
-TBD
-
-%\subsection{3-D Model Setup}
-
-%\subsection{3-D Numerical Results}
+\subsection{3-D Neutronics Model Setup} \label{sec:3d-model-setup}
+
+Figure \ref{fig:msre-picture} shows the vertical cross section of the actual \gls{MSRE} vessel.
+The actual design has rounded upper and lower plena to help direct salt flow through the system.
+This work uses a modified 3-D numerical \gls{MSRE} model as shown by the vertical cross section
+in Figure \ref{fig:msre-geom-vert}. The model excludes the toroidal flow distributor near the top
+of the vessel and reshapes the plena into right cylinders. It has the same horizontal cross section
+as the 2-D model (Figure \ref{fig:full-geom}) throughout most of the salt-graphite lattice. In
+addition to modifications in the 2-D model listed in Section \ref{sec:2d-model-setup}, the
+3-D model includes the following modifications:
+
+\begin{itemize}
+  \item Removed the toroidal flow distributor near the top of the vessel.
+  \item Reshaped the upper and lower plena into right cylinders while keeping the salt volume
+    constant.
+  \item Flattened the conical tops of the graphite lattice blocks.
+  \item Regularized the complicated graphite lattice structure at the bottom.
+  \item Set a fixed outer vessel thickness of 2.54 cm.
+  \item Homogenized the lower plenum by 90.8\% salt- 9.2\% INOR-8 (vessel) by volume.
+  \item Removed the reactor access nozzle and the various components within it.
+  \item Added a graphite block at the bottom of the control rod thimbles.
+\end{itemize}
+
+\begin{figure}[p]
+  \centering
+  \includegraphics[width=0.97\columnwidth]{msre-picture}
+  \caption{Vertical cross section of the actual \gls{MSRE} vessel. It has rounded upper and lower
+  plena, and a toroidal flow distributor that distributes the salt inflow evenly as it enters the
+  core through holes on the inside vessel surface of the distributor.}
+  \label{fig:msre-picture}
+\end{figure}
+
+\begin{figure}[p]
+  \centering
+  \includegraphics[width=\columnwidth]{msre-geom-vert}
+  \caption{Vertical cross section of the 3-D numerical \gls{MSRE} model.}
+  \label{fig:msre-geom-vert}
+\end{figure}
+
+Table \ref{table:reactor-dimensions} lists relevant dimensions of the 3-D \gls{MSRE} model.
+All material compositions remain
+the same as the 2-D models and the \gls{MSRE} benchmark specifications \cite{fratoni_molten_2020}.
+All 3-D neutronics simulations in this section used the molten salt composition with 65.25 kg
+$^{235}$U loaded (Table \ref{table:2d-salt-composition}), except the simulations for temperature
+reactivity coefficient calculations which had a higher $^{235}$U loading of 67.00 kg (Table
+\ref{table:67-salt-composition}.
+
+\begin{table}[htb]
+  \centering
+  \caption{List of reactor dimensions.}
+  \label{table:reactor-dimensions}
+\end{table}
+
+\begin{table}[htb]
+  \small
+  \centering
+  \setlength\tabcolsep{4pt}
+  \caption{\gls{MSRE} molten salt composition when the $^{235}$U loading was at 65.25 kg.}
+  \begin{tabular}{l S S S S S S S}
+    \toprule
+    Element & {Li} & {Be} & {Zr} & {Hf} & {$^{234}$U} & {$^{235}$U} & {$^{236}$U} \\
+    \midrule
+    Composition [kg] & 507.42 & 293.96 & 513.97 & 0.0029 & 0.68 & 67.00 & 0.28 \\
+    \bottomrule
+  \end{tabular}
+  \begin{tabular}{l S S S S S S}
+    \toprule
+    Element & {$^{238}$U} & {Fe} & {Cr} & {Ni} & {O} & {F} \\
+    \midrule
+    Composition [kg] & 142.02 & 0.75 & 0.13 & 0.14 & 2.27 & 3104.22 \\
+    \bottomrule
+  \end{tabular}
+  \label{table:67-salt-composition}
+\end{table}
+
+All neutronics simulations ran on OpenMC under continuous energy mode and on Moltres with the
+hybrid $S_N$-diffusion method. The hybrid simulations ran with a maximum stabilization factor of
+$c=250$, a void constant $\varsigma=0.5$, and a uncorrected diffusion coefficient cap of $D_g=2.5$.
+Refer to Sections \ref{sec:void-treatment} and \ref{sec:diffcoef-cap} for the relevant discussion
+on void treatment for near-void regions and the impact of large diffusion coefficient values on
+solver performance. As for the standard neutron diffusion method, it either converges very slowly
+or does not converge at all when $D_g$ values are uncapped due to excessive diffusive effects in
+the air-filled control rod thimbles. Group-wise $D_g$ values, generated using OpenMC, ranged from
+100 cm to 850 cm in the air regions as compared to the typical 0.5 cm to 2.0 cm range observed in
+non-air regions. Therefore, all neutron diffusion method simulation results in the following
+section were measured with the diffusion coefficient capped at $D_g=2.5$.
+
+Unlike the 2-D hybrid simulations which ran with $S_8$ angular discretization, the 3-D hybrid
+simulations ran with $S_6$ angular discretization due to significantly
+larger computational and memory requirements demanded by the 3-D simulations. Further discussion on
+computational performance may be found at the end of Section \ref{sec:3d-nts}. All simulations,
+except those for temperature reactivity coefficient calculations, ran on isothermal 3-D models at
+911 K.
+
+\subsection{3-D Numerical Results} \label{sec:3d-nts}
+
+\subsubsection{Initial Criticality Configuration}
+
+The \gls{MSRE} achieved initial criticality with Rod 1 inserted to a height of 46.6 inches relative
+to the fully inserted position. For reference, the rods are at a height of 51 inches when fully
+withdrawn. Table \ref{table:initial-crit}
+shows the $k_\text{eff}$ values from \gls{MSRE} experimental data, numerical benchmark data
+\cite{fratoni_molten_2020}, and this work. All numerical models overestimate the $k_\text{eff}$
+relative to the experimental value. In the \gls{MSRE} numerical benchmark report, the authors
+attribute this $k_\text{eff}$ discrepancy largely to discrepancies in the carbon cross section data
+in nuclear data libraries. They noted similar $k_\text{eff}$ overestimations by 1-2\% in benchmark
+results for the HTTR, a graphite-moderated, helium-cooled reactor. The OpenMC and Moltres
+simulations in this work used the same
+ENDF/B-VII.1 nuclear data library as the \gls{MSRE} numerical benchmark report.
+
+\begin{table}[htb]
+  \centering
+  \caption{$k_\text{eff}$ values from \gls{MSRE} experimental data, the \gls{MSRE} numerical
+  benchmark \cite{fratoni_molten_2020}, and the OpenMC and Moltres models in this work.}
+  \begin{tabular}{l S[table-format=1.5(2)]}
+    \toprule
+     & {$k_\text{eff}$} \\
+     \cmidrule(l){1-1}
+    \gls{MSRE} experimental data & 1.00000(420) \\
+    Serpent 2 (Numerical benchmark) & 1.02132(3) \\
+    OpenMC (This work) & 1.01308(20) \\
+    Hybrid (This work) & 1.01955 \\
+    Diffusion (This work) & 1.01885 \\
+    \bottomrule
+  \end{tabular}
+  \label{table:initial-crit}
+\end{table}
+
+Of the four $k_\text{eff}$ values, we expect to see the closest agreement between the Serpent 2
+model from the numerical benchmark and OpenMC from this work. However, the OpenMC \gls{MSRE} model
+differs from the Serpent 2 \gls{MSRE} models due to geometry modifications listed in Sections
+\ref{sec:2d-model-setup} and \ref{sec:3d-model-setup}. The benchmark report quantified the impact
+of several geometry modifications, namely the removal of the flow distributor, the homogenization
+of the sample basket, and the removal of the thermal shield. These modifications contribute to
+-$0.098\%$, -$0.037\%$, and -$0.885\%$ change in the $k_\text{eff}$, respectively, for a final
+$k_\text{eff}$ estimate of $1.01093(3)$. This brings the OpenMC $k_\text{eff}$ estimate of
+$1.01308(20)$ to closer agreement with the numerical benchmark value.
+
+The hybrid method $k_\text{eff}$ estimate is 0.00647 higher than the OpenMC estimate. For this
+initial criticality configuration, the neutron diffusion method $k_\text{eff}$ estimate is closer
+to OpenMC than the hybrid method. Overall, the OpenMC and Moltres models show good agreement with
+the Serpent 2 model from the numerical benchmark.
+Figures * show the neutron group fluxes on the interior of the 3-D models. The salt-graphite
+lattice structure is visible from the group flux distributions. Group 7 and 8 neutron fluxes are
+particularly small near the control rods due to strong thermal neutron capture by Gadolinium in the
+rods.
+
+\subsubsection{Control Rod Worth}
+
+After achieving initial criticality, \gls{ORNL} researchers ran control rod calibration experiments
+with the \gls{MSRE} at zero-power (typically at about 10 W power output).
+
+For control rod worth calculations, the OpenMC and Moltres simulations ran  with Rod 1 at fourteen
+different rod positions from full insertion (0 inch) to full withdrawal (51 inches). The rod
+positional limits of 0 inch and 51 inches represents the full range of deliberate rod motion
+allowable by the control rod drive motor \cite{robertson_msre_1965}.
+
+\begin{figure}[p]
+  \centering
+  \includegraphics[width=0.8\columnwidth]{rod-worth-curve}
+  \caption{Reactivity inserted by Rod 1 at various rod positions relative to the full insertion.}
+  \label{fig:rod-worth}
+\end{figure}
+
+\subsubsection{Temperature Reactivity Coefficient}
+
+\subsubsection{Computational Performance}
 
 \section{Summary}
-
-TBD
 

### 2024-11-04 hybrid.tex
Add 3-D results content

@@
 
-\subsection{Spatial Discretization and Void Treatment}
+\subsection{Spatial Discretization and Void Treatment} \ref{sec:void-treatment}
 
@@
     \begin{cases}
-      \frac{1}{\Sigma_{t,g}} \mbox{ for } h\Sigma_{t,g} \geq \varsigma \\
-      \frac{h}{\varsigma} \mbox{ for } h\Sigma_{t,g} < \varsigma
+      \frac{1}{c\Sigma_{t,g}} \mbox{ for } ch\Sigma_{t,g} \geq \varsigma \\
+      \frac{h}{\varsigma} \mbox{ for } ch\Sigma_{t,g} < \varsigma
     \end{cases}, & \\
     h &= \mbox{mesh element size,} & \\
+    c &= \mbox{maximum stabilization factor,} & \\
     \varsigma &= \mbox{void constant.} &
@@
 \end{gather}
+
+\subsection{Setting Caps on the Diffusion Coefficient Values} \ref{sec:diffcoef-cap}
+
+From experience, the $S_N$-diffusion outer iterations under both hybrid schemes or the standard
+$S_N$ method with nonlinear diffusion acceleration converged at a faster rate when the uncorrected
+diffusion coefficient values are capped at around $D_g=2.5$ cm. Excessively large $D_g$ values,
+which occur in near-void regions, cause significant computational performance degradation. The caps
+have no impact on any method because the diffusion term containing $D_g$ is canceled out by the
+transport correction terms as the solution converges. This behavior was also confirmed through
+email correspondence with Dr.\ Yaqi Wang regarding nonlinear diffusion acceleration of the
+\gls{SAAF} $S_N$ method \cite{wang_diffusion_2014}.
 

### 2024-11-04 hybrid-nts.tex
More 3-D modeling content

@@
   \includegraphics[width=\columnwidth]{msre-geom-vert}
-  \caption{Vertical cross section of the 3-D numerical \gls{MSRE} model.}
+  \caption{Vertical cross section of the 3-D numerical \gls{MSRE} model offset by 5.08 cm to show
+  the control rod thimble and homogenized sample basket. The control rod is in its fully inserted
+  position. The homogenized lower plenum consists of 90.8\% salt and 9.2\% INOR-8 alloy by volume.}
   \label{fig:msre-geom-vert}
@@
 reactivity coefficient calculations which had a higher $^{235}$U loading of 67.00 kg (Table
-\ref{table:67-salt-composition}.
+\ref{table:67-salt-composition}).
 
@@
   \caption{List of reactor dimensions.}
+  \begin{tabular}{l l}
+    Height [cm] & 215 \\
+    TODO: Populate table & \\
   \label{table:reactor-dimensions}
@@
      & {$k_\text{eff}$} \\
-     \cmidrule(l){1-1}
+     \cmidrule(l){2-2}
     \gls{MSRE} experimental data & 1.00000(420) \\
@@
 
-After achieving initial criticality, \gls{ORNL} researchers ran control rod calibration experiments
-with the \gls{MSRE} at zero-power (typically at about 10 W power output).
+After achieving initial criticality, \gls{MSRE} researchers ran control rod calibration experiments
+with the \gls{MSRE} at zero-power (typically at about 10 W power output)
+\cite{prince_zero-power_1968}. They measured differential rod worths by measuring the time taken
+for the reactor power to increase by 20 W following a prescribed withdrawal of Rod 1 and applying
+the inhour equation to obtain the reactivity inserted. They repeated this process at different rod
+heights after successive additions of highly enriched uranium capsules. The final integral rod
+worth curve was normalized to the critical $^{235}$U loading of 65.25 kg
+\cite{prince_zero-power_1968}.
 
@@
 
+Figure \ref{fig:rod-worth} shows the integral rod worth curves from the \gls{MSRE} experimental
+data and the OpenMC and Moltres numerical models. By taking the $k_\text{eff}$ when Rod 1 is fully
+inserted as reference, all rod worth curves start at 0 pcm. The change in integral rod worth,
+represented by the gradient in the plot, is steepest as the tip of the rod passes through the
+center of the core. The hybrid method performs well in reproducing the rod worth curve from OpenMC;
+all hybrid method rod worth values fall within one or two sigma range of OpenMC.
+
+Due to the experimental procedure of computing integral rod worth from cumulative differential rod
+worth measurements, the uncertainty range of the \gls{MSRE} data scales with rod position.
+Both OpenMC and the hybrid method report consistently larger integral rod worths than the
+\gls{MSRE} data across all rod positions. Nevertheless, they are in good agreement with the
+\gls{MSRE} data as compared to the neutron diffusion method, which significantly overestimates the
+integral rod worth.
+
+Table \ref{table:rod-worth} shows the total rod worth of Rod 1.
+
 \begin{figure}[p]
@@
 
+\begin{table}[htb]
+  \centering
+  \caption{Total rod worth of Rod 1 when fully inserted.}
+  \begin{tabular}{l S[table-format=4(2)] S}
+    \toprule
+     & {$\rho_\text{worth}$ [pcm]} & {Percentage discrepancy [\%]}\\
+     \cmidrule(l){2-2}
+    \gls{MSRE} experimental data & 2250(67) & {-}\\
+    OpenMC (This work) & 2364(44) & 5.1 \\
+    Hybrid (This work) & 2345 & 4.2 \\
+    Diffusion (This work) & 2725 & 21.1 \\
+    \bottomrule
+  \end{tabular}
+  \label{table:rod-worth}
+\end{table}
+
 \subsubsection{Temperature Reactivity Coefficient}

### 2024-11-05 hybrid-nts.tex
Further chapter 5 content

@@
 
-\subsection{1-D Neutronics Numerical Results}
+\subsection{1-D Neutronics Numerical Results \& Discussion}
 
@@
 
-\subsection{2-D Neutronics Numerical Results} \label{sec:2d-nts-results}
+\subsection{2-D Neutronics Numerical Results \& Discussion} \label{sec:2d-nts-results}
 
@@
     TODO: Populate table & \\
+  \end{tabular}
   \label{table:reactor-dimensions}
@@
 
-\subsection{3-D Numerical Results} \label{sec:3d-nts}
+\subsection{3-D Numerical Results \& Discussion} \label{sec:3d-nts}
 
@@
 to the fully inserted position. For reference, the rods are at a height of 51 inches when fully
-withdrawn. Table \ref{table:initial-crit}
-shows the $k_\text{eff}$ values from \gls{MSRE} experimental data, numerical benchmark data
-\cite{fratoni_molten_2020}, and this work. All numerical models overestimate the $k_\text{eff}$
-relative to the experimental value. In the \gls{MSRE} numerical benchmark report, the authors
+withdrawn. Table \ref{table:initial-crit} shows the $k_\text{eff}$ values from \gls{MSRE}
+experimental data, numerical benchmark data \cite{fratoni_molten_2020}, and this work. All
+numerical models overestimate the $k_\text{eff}$ relative to the experimental value by 1-2\%
+(1000-2000 pcm). In the \gls{MSRE} numerical benchmark report, the authors
 attribute this $k_\text{eff}$ discrepancy largely to discrepancies in the carbon cross section data
@@
 lattice structure is visible from the group flux distributions. Group 7 and 8 neutron fluxes are
-particularly small near the control rods due to strong thermal neutron capture by Gadolinium in the
-rods.
+particularly small near the control rods due to strong thermal neutron capture cross section of
+gadolinium.
 
@@
 
-Due to the experimental procedure of computing integral rod worth from cumulative differential rod
-worth measurements, the uncertainty range of the \gls{MSRE} data scales with rod position.
+Due to the experimental procedure of computing integral rod worth through cumulative differential
+rod worth measurements, the uncertainty range of the \gls{MSRE} data scales with rod position.
 Both OpenMC and the hybrid method report consistently larger integral rod worths than the
@@
 integral rod worth.
-
-Table \ref{table:rod-worth} shows the total rod worth of Rod 1.
-
-\begin{figure}[p]
+Table \ref{table:rod-worth} shows the total rod worth of Rod 1. Similar to observations from Figure
+\ref{fig:rod-worth}, OpenMC and the hybrid method report total rod worths marginally larger than
+the \gls{MSRE} data by 5.1\% and 4.2\%, respectively, while the diffusion method exceeds the
+experimental value by 21.1\%.
+
+\begin{figure}[t]
   \centering
@@
 
-\begin{table}[htb]
+\begin{table}[t]
   \centering
@@
      & {$\rho_\text{worth}$ [pcm]} & {Percentage discrepancy [\%]}\\
-     \cmidrule(l){2-2}
+     \cmidrule(rl){2-2} \cmidrule(l){3-3}
     \gls{MSRE} experimental data & 2250(67) & {-}\\
@@
 
+The most likely factor for OpenMC and the hybrid method overpredicting the control rod worth is the
+modeling of control rods as solid annular cylinders as opposed to 36 separate control elements
+strung along a flexible stainless steel cable. Figure \ref{fig:rod-bead} depicts a poison element
+and its physical dimensions. Other than the Gd$_2$O$_3$-Al$_2$O$_3$ poison material, each control
+element has a top and bottom closure made of INOR-8 alloy. According to the physical dimensions,
+the closures make up 15.1\% of the total element height. The 3-D numerical \gls{MSRE} model in this
+work omits the outer and inner INOR-8 tubes and the flexible stainless cable, and replaces the top
+and bottom closure with poison material. Fully resolving the closures in Moltres would require
+excessively fine axial mesh resolution and advanced adaptive meshing features not currently
+available. A straightforward corrective measure for future \gls{MSRE} simulations would be to
+retain the solid annular cylinder structure and adjust the rod dimensions or the rod material
+composition to match the expected worth from \gls{MSRE} data or OpenMC simulations with the beaded
+geometry.
+
+\begin{figure}[t]
+  \begin{subfigure}[b]{0.33\columnwidth}
+    \centering
+    \includegraphics[width=\columnwidth]{rod-bead-2}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.65\columnwidth}
+    \centering
+    \includegraphics[width=\columnwidth]{rod-bead}
+  \end{subfigure}
+  \caption{Cutaway view and dimensions of a control rod poison element.}
+  \label{fig:rod-bead}
+\end{figure}
+
+Overall, the results show that the hybrid $S_N$-diffusion method accurately reproduces the control
+rod worth at various levels of insertion. As observed in the 1-D and 2-D models, the 3-D model
+$k_\text{eff}$ values exhibit some discrepancies relative to OpenMC, but the discrepancies do not
+affect rod worth estimates because they are systematic errors. This \gls{MSRE} hybrid model may be
+improved in future work by characterizing and applying neutron albedo at the vacuum boundaries.
+
 \subsubsection{Temperature Reactivity Coefficient}
 
+This work compares isothermal temperature reactivity coefficients calculated from $k_\text{eff}$
+values measured at 872 K and 972 K with 67.86 $^{235}$U loading. This approach is identical to that
+of the \gls{MSRE} experimental measurements \cite{prince_zero-power_1968} and the \gls{MSRE}
+numerical benchmark calculations \cite{fratoni_molten_2020}. The temperature reactivity coefficient
+is calculated by dividing the difference in reactivities by the temperature change (100 K). All
+simulations included the effects of thermal expansion on the material density but left all reactor
+physical dimensions unchanged.
+
+Table \ref{table:temp-coef} compares the isothermal reactivity coefficients from the experimental
+and numerical sources. The OpenMC and Moltres models in this work show closer agreement with the
+\gls{MSRE} data than the Serpent 2 model from the \gls{MSRE} numerical benchmark. All numerical
+results fall within one standard deviation of the \gls{MSRE} data.
+
+\begin{table}[t]
+  \small
+  \centering
+  \caption{Isothermal temperature reactivity coefficients with 67.86 kg $^{235}$U loading from
+  \gls{MSRE} data, the \gls{MSRE} numerical benchmark \cite{fratoni_molten_2020}, and the OpenMC
+  and Moltres models in this work.}
+  \begin{tabular}{l S[separate-uncertainty=true,table-format=2.2(2)] S}
+    \toprule
+     & {Temperature coefficient [pcm K$^{-1}$]} & {Percentage discrepancy [\%]}\\
+     \cmidrule(rl){2-2} \cmidrule(l){3-3}
+    \gls{MSRE} experimental data & 13.41(150) & {-}\\
+    Serpent 2 (Numerical benchmark) & 12.51(23) & -6.7 \\
+    OpenMC (This work) & 13.85(43) & 3.3 \\
+    Hybrid (This work) & {TBD} &  \\
+    Diffusion (This work) & {TBD} &  \\
+    \bottomrule
+  \end{tabular}
+  \label{table:temp-coef}
+\end{table}
+
 \subsubsection{Computational Performance}
+
+% Strong scaling tests & discussion
+
+% Difficulties with 3-D modeling
+% SN void treatment constants, Transfer times
 

### 2024-11-07 hybrid-nts.tex
Chapter 5 summary

@@
 testbed for \gls{HPC} scaling tests, and preparation for the time-dependent multiphysics
-simulations of \gls{MSRE} reactivity insertion transients.
+simulations in Chapter 6.
 
@@
 Figures \ref{fig:1/4-geom} and \ref{fig:full-geom} show the 2-D quarter-core and full-core
-\gls{MSRE} models investigated. I adapted the models from the horizontal cross section of the base
-\gls{MSRE} benchmark model developed by \cite{fratoni_molten_2020} with the following differences
-in the 2-D \gls{MSRE} model for this work:
+\gls{MSRE} models investigated. I adapted the models from the horizontal cross section of the
+\gls{MSRE} numerical benchmark model \cite{fratoni_molten_2020} with the
+following differences in the 2-D \gls{MSRE} model for this work:
 
@@
   \begin{tabular}{l l}
+    \toprule
     Height [cm] & 215 \\
+    \midrule
     TODO: Populate table & \\
+    \bottomrule
   \end{tabular}
@@
 center of the core. The hybrid method performs well in reproducing the rod worth curve from OpenMC;
-all hybrid method rod worth values fall within one or two sigma range of OpenMC.
+all hybrid method rod worth values fall within one or two standard deviations of OpenMC.
 
@@
     OpenMC (This work) & 13.85(43) & 3.3 \\
-    Hybrid (This work) & {TBD} &  \\
-    Diffusion (This work) & {TBD} &  \\
+    Hybrid (This work) & 13.98 & 4.3 \\
+    Diffusion (This work) & 14.01 & 4.5 \\
     \bottomrule
@@
 
+Beyond demonstrating the hybrid $S_N$-diffusion method, its computational performance is of
+interest as a candidate for 3-D, time-dependent reactor simulations. All simulations ran on the
+Polaris supercomputer system at \gls{ANL}. Each compute node consists of a 2.8-GHz AMD EPYC Milan
+7543P 32-core CPU, four NVIDIA A100 GPUs, and 512 GB of RAM. The GPUs were not utilized in this
+work. Moltres uses the \gls{MPI} communication protocol for parallelizing simulations. For this
+work, simulations ran with 1-to-1 ratio of \gls{MPI} ranks and processor cores (i.e., 32 \gls{MPI}
+ranks for each compute node).
+
+Every 3-D hybrid method $k$-eigenvalue
+simulation of the \gls{MSRE} consisted of 747,688,760 \glspl{DOF} and 1,690,777,728 \glspl{DOF} for
+the neutron diffusion and $S_N$ components, respectively. Each \gls{DOF} corresponds to a scalar
+or angular group flux variable value defined at a mesh node. Each simulation ran with the distributed
+mesh feature, which partitions and distributes the mesh across however many \gls{MPI} ranks used
+in the simulation. All variable and mesh-linked data are also distributed with the mesh. This
+greatly reduces memory requirements compared to storing the full mesh for each \gls{MPI} rank.
+
+Each 3-D \gls{MSRE} simulation required at least forty nodes for sufficient memory to run. This
+sets a strict minimum requirement for the 3-D hybrid simulations. On forty nodes, the simulations
+required approximately 2.2 h to complete for a total of 88 node-hours or 2816 core-hours. The
+distributed mesh capability comes with the cost of increased data transfer times between the $S_N$
+and diffusion solvers due to increased inter-core data transfers. Transfers accounted for
+approximately 0.675 h or 30 \% of the total simulation time. Future research into optimizing mesh
+distribution and transfer-related data caching may help to minimize data transfers and speed-up
+hybrid method simulations. The solution time for the initial neutron diffusion calculation with zero
+drift correction also takes up approximately 30 \% of the total simulation time. After subtracting
+the minor computational cost of evaluating the drift term even if it is zero, the hybrid method
+requires about four times the solution time of the standard neutron diffusion method.
+
+For the hybrid method to be viable for large time-dependent simulations, it must scale reasonably
+well on computing clusters. For this work, I investigated strong scaling, which measures how the
+solution time decreases when more cores are used while keeping the problem size fixed. This strong
+scaling study used 3-D quarter-core models run on 10, 20, 30, and 40 compute nodes.
+Figure * shows the results of this scaling study.
+The hybrid method implementation scales well as shown by the nearly linear scaling between the
+solution time and the number of processors. Therefore, the hybrid method appears promising for the
+modeling of larger reactor analyses with little scaling performance penalty.
+
 % Strong scaling tests & discussion
-
 % Difficulties with 3-D modeling
@@
 
-%The hybrid method does not alter the $k_\text{eff}$ and flux estimates relative to the neutron
-%diffusion method in cases where the neutron diffusion method provides
+This chapter covers work demonstrating the hybrid $S_N$-diffusion method implemented in Moltres
+through 1-D, 2-D, and 3-D neutronics simulations of \gls{MSRE} models. The hybrid method combines
+the $S_N$
+and diffusion methods by applying transport corrections generated using the high-fidelity $S_N$
+method near control rods while treating the rest of the problem with the lower-fidelity diffusion
+method for tractable 3-D time-dependent reactor analyses. This is achieved by identifying and
+setting up a subregion containing the control rod and its vicinity for $S_N$ transport
+calculations. The $S_N$ and diffusion solvers are coupled through boundary sources at the subregion
+boundaries. The hybrid method applies an adaptive algorithm for discarding inaccurate transport
+correction values near the subregion boundaries and preserving the smoothness of the neutron flux
+gradient. In this chapter, the 1-D simulations provided verification of the hybrid method for
+estimating control rod worth. The simulations also compared two different implementations of the
+hybrid method which differed by the use of drift or diffusion correction schemes.
+The 2-D and 3-D simulations extended the hybrid method to multi-dimensional problems. Additionally,
+the 3-D simulations validated the numerical \gls{MSRE} model against \gls{MSRE} experimental data
+and provided useful insight on the computational performance of the hybrid method for
+large reactor analyses.
+
+Six 1-D test cases based on the \gls{MSRE} with increasing geometric complexity were run using the
+standard $S_N$ method, the hybrid method, and standard neutron diffusion method and verified against
+reference OpenMC Monte Carlo neutron transport simulations under continuous neutron energy (OpenMC-CE) and
+multigroup (OpenMC-MG) modes. The $S_N$ implementation with nonlinear diffusion acceleration
+showed good agreement with OpenMC-MG which indicated that discrepancies between those codes and
+OpenMC-CE largely arose from the eight neutron energy group structure for multigroup simulations.
+Both hybrid method implementations, with drift or diffusion correction schemes, showed similar
+performance in the 1-D cases. While the $k_\text{eff}$ estimates from the hybrid method deviated
+from the $S_N$ method and OpenMC-MG by 100 to 400 pcm, the hybrid method accurately reproduces
+control rod worths. Rod worth estimates from OpenMC-MG, $S_N$, and hybrid methods differed from
+OpenMC-CE by 2 to 3 \%. The hybrid method improves on the neutron diffusion method whose rod worth
+error estimate is 8 \%. The hybrid method also produced better agreement in neutron flux
+distributions with the neutron transport methods compared to the neutron diffusion method.
+
+Analysis on the drift and diffusion correction schemes showed that the diffusion correction
+parameter led to numerous discontinuities tending to infinity values near points in the problem
+domain where flux peaks and troughs occur. These discontinuities and negative diffusion coefficient
+values are unphysical and would lead to numerical stability issues. Therefore, the hybrid method
+implementation with drift correction scheme was chosen for the remaining 1-D analyses and the
+2-D and 3-D analyses.
+
+Analysis on the impact of $S_N$ correction subregion size found that it has little impact on the
+control rod worth estimates as long as the subregion size is kept consistent (i.e., same size) when
+performing control rod worth measurements. Rod worths varied by less than 0.35 \% across the
+different subregion sizes investigated. A separate analysis showed that the $k_\text{eff}$
+estimates from the hybrid method exhibit superlinear convergence rate when tightening the $S_N$
+convergence tolerance value. The convergence tolerance value also impacts the number of outer
+iterations between $S_N$ and diffusion solvers. After just two outer iterations, the error in
+$k_\text{eff}$ fell to below 10$^{-7}$.
+
+The hybrid method continued to show improved control rod estimates over the neutron diffusion
+method in 2-D quarter-core and full-core \gls{MSRE} simulations with up to three rods inserted in the
+full-core model. The hybrid method reported rod worth error magnitudes of less than 40 pcm relative
+to OpenMC-CE, which are significant improvements over the neutron diffusion method error estimates
+ranging from 569 pcm to 1484 pcm. The hybrid method also showed significant improvements in the
+absolute mean and maximum errors in fuel channel power distributions relative to the reference
+OpenMC-CE distribution.
+
+For 3-D simulations, the \gls{MSRE} experimental data and the \gls{MSRE} numerical benchmark
+report \cite{fratoni_molten_2020} both served as references for model validation. The numerical
+benchmark is included in the International Reactor Physics Experiment Evaluation Project (IRPhEP)
+handbook. The
+$k_\text{eff}$ estimates of the \gls{MSRE} at initial criticality from OpenMC, the hybrid method,
+and the neutron diffusion method in this work showed good agreement with the \gls{MSRE} numerical
+benchmark. All numerical $k_\text{eff}$ estimates, including that from the benchmark report
+\cite{fratoni_molten_2020}, exceeded the \gls{MSRE} experimental value by 1-2 \% due to possible
+biases and uncertainties in the neutronic properties of graphite. Temperature reactivity
+coefficient values from OpenMC, the hybrid method, and the neutron diffusion method also show good
+agreement with the \gls{MSRE} data within experimental uncertainty with percentage discrepancies of
+about 3 \%, but differ from the numerical benchmark which has a percentage discrepancy of -6.7 \%.
+
+Control rod worth calculations were also performed for Rod 1 of the \gls{MSRE} under various levels
+of insertion from its fully inserted to its fully withdrawn state. OpenMC and the hybrid method
+showed excellent agreement at every stage within one standard deviation. Due to geometric
+approximations of the control rods in the numerical models, they overestimated the total
+rod worth relative to the \gls{MSRE} experimental data by approximately 4-5 \%. The hybrid method
+compares favorably against the neutron diffusion method which overestimated the total worth by
+21.1 \%.
+
+The 3-D simulations performed well under the strong scaling test which studies how the solution
+time scales with the number of processors used to run the simulation. The 3-D simulations showed
+close to linear scaling within the scope of the strong scaling test. Therefore, the hybrid method
+exhibits good computational scaling performance which is essential for large 3-D reactor analyses.
+Excessively large memory requirements were mitigated through the distributed mesh capability in
+Moltres. Some performance optimizations may be possible to reduce data transfer times
+between the $S_N$ and diffusion solvers which currently take up about 30 \% of the total solution
+time as a result of the inter-processor communications.
+
+Overall, the hybrid $S_N$-diffusion method provides a significant improvement over the neutron
+diffusion method at only four times the computational expense. Control rod estimates
+show good agreement with reference Monte Carlo solutions and \gls{MSRE} experimental data.
+Further investigations into characterizing and minimizing neutron leakage rates at vacuum
+boundaries may help with reducing $k_\text{eff}$ discrepancies with reference Monte Carlo
+solutions.

### 2024-11-07 msre.tex
Rename files

@@
 
-\subsection{2-D Neutronics Model Setup}
+\subsection{2-D Neutronics Model Setup} \label{sec:2d-model-setup}
 
 Figures \ref{fig:1/4-geom} and \ref{fig:full-geom} show the 2-D quarter-core and full-core
-\gls{MSRE} models investigated. I adapted the models from the horizontal cross section of the base
-\gls{MSRE} benchmark model developed by \cite{fratoni_molten_2020} with the following differences
-in the 2-D \gls{MSRE} model for this work:
+\gls{MSRE} models investigated. I adapted the models from the horizontal cross section of the
+\gls{MSRE} numerical benchmark model \cite{fratoni_molten_2020} with the
+following differences in the 2-D \gls{MSRE} model for this work:
 
@@
 
-\subsection{2-D Neutronics Numerical Results} \label{sec:2d-nts-results}
-
-\subsubsection{2-D quarter-core results}
+\subsection{2-D Neutronics Numerical Results \& Discussion} \label{sec:2d-nts-results}
+
+\subsubsection{2-D Quarter-Core}
 
@@
 
-TBD
-
-%\subsection{3-D Model Setup}
-
-%\subsection{3-D Numerical Results}
+\subsection{3-D Neutronics Model Setup} \label{sec:3d-model-setup}
+
+Figure \ref{fig:msre-picture} shows the vertical cross section of the actual \gls{MSRE} vessel.
+The actual design has rounded upper and lower plena to help direct salt flow through the system.
+This work uses a modified 3-D numerical \gls{MSRE} model as shown by the vertical cross section
+in Figure \ref{fig:msre-geom-vert}. The model excludes the toroidal flow distributor near the top
+of the vessel and reshapes the plena into right cylinders. It has the same horizontal cross section
+as the 2-D model (Figure \ref{fig:full-geom}) throughout most of the salt-graphite lattice. In
+addition to modifications in the 2-D model listed in Section \ref{sec:2d-model-setup}, the
+3-D model includes the following modifications:
+
+\begin{itemize}
+  \item Removed the toroidal flow distributor near the top of the vessel.
+  \item Reshaped the upper and lower plena into right cylinders while keeping the salt volume
+    constant.
+  \item Flattened the conical tops of the graphite lattice blocks.
+  \item Regularized the complicated graphite lattice structure at the bottom.
+  \item Set a fixed outer vessel thickness of 2.54 cm.
+  \item Homogenized the lower plenum by 90.8\% salt- 9.2\% INOR-8 (vessel) by volume.
+  \item Removed the reactor access nozzle and the various components within it.
+  \item Added a graphite block at the bottom of the control rod thimbles.
+\end{itemize}
+
+\begin{figure}[p]
+  \centering
+  \includegraphics[width=0.97\columnwidth]{msre-picture}
+  \caption{Vertical cross section of the actual \gls{MSRE} vessel. It has rounded upper and lower
+  plena, and a toroidal flow distributor that distributes the salt inflow evenly as it enters the
+  core through holes on the inside vessel surface of the distributor.}
+  \label{fig:msre-picture}
+\end{figure}
+
+\begin{figure}[p]
+  \centering
+  \includegraphics[width=\columnwidth]{msre-geom-vert}
+  \caption{Vertical cross section of the 3-D numerical \gls{MSRE} model offset by 5.08 cm to show
+  the control rod thimble and homogenized sample basket. The control rod is in its fully inserted
+  position. The homogenized lower plenum consists of 90.8\% salt and 9.2\% INOR-8 alloy by volume.}
+  \label{fig:msre-geom-vert}
+\end{figure}
+
+Table \ref{table:reactor-dimensions} lists relevant dimensions of the 3-D \gls{MSRE} model.
+All material compositions remain
+the same as the 2-D models and the \gls{MSRE} benchmark specifications \cite{fratoni_molten_2020}.
+All 3-D neutronics simulations in this section used the molten salt composition with 65.25 kg
+$^{235}$U loaded (Table \ref{table:2d-salt-composition}), except the simulations for temperature
+reactivity coefficient calculations which had a higher $^{235}$U loading of 67.00 kg (Table
+\ref{table:67-salt-composition}).
+
+\begin{table}[htb]
+  \centering
+  \caption{List of reactor dimensions.}
+  \begin{tabular}{l l}
+    \toprule
+    Height [cm] & 215 \\
+    \midrule
+    TODO: Populate table & \\
+    \bottomrule
+  \end{tabular}
+  \label{table:reactor-dimensions}
+\end{table}
+
+\begin{table}[htb]
+  \small
+  \centering
+  \setlength\tabcolsep{4pt}
+  \caption{\gls{MSRE} molten salt composition when the $^{235}$U loading was at 65.25 kg.}
+  \begin{tabular}{l S S S S S S S}
+    \toprule
+    Element & {Li} & {Be} & {Zr} & {Hf} & {$^{234}$U} & {$^{235}$U} & {$^{236}$U} \\
+    \midrule
+    Composition [kg] & 507.42 & 293.96 & 513.97 & 0.0029 & 0.68 & 67.00 & 0.28 \\
+    \bottomrule
+  \end{tabular}
+  \begin{tabular}{l S S S S S S}
+    \toprule
+    Element & {$^{238}$U} & {Fe} & {Cr} & {Ni} & {O} & {F} \\
+    \midrule
+    Composition [kg] & 142.02 & 0.75 & 0.13 & 0.14 & 2.27 & 3104.22 \\
+    \bottomrule
+  \end{tabular}
+  \label{table:67-salt-composition}
+\end{table}
+
+All neutronics simulations ran on OpenMC under continuous energy mode and on Moltres with the
+hybrid $S_N$-diffusion method. The hybrid simulations ran with a maximum stabilization factor of
+$c=250$, a void constant $\varsigma=0.5$, and a uncorrected diffusion coefficient cap of $D_g=2.5$.
+Refer to Sections \ref{sec:void-treatment} and \ref{sec:diffcoef-cap} for the relevant discussion
+on void treatment for near-void regions and the impact of large diffusion coefficient values on
+solver performance. As for the standard neutron diffusion method, it either converges very slowly
+or does not converge at all when $D_g$ values are uncapped due to excessive diffusive effects in
+the air-filled control rod thimbles. Group-wise $D_g$ values, generated using OpenMC, ranged from
+100 cm to 850 cm in the air regions as compared to the typical 0.5 cm to 2.0 cm range observed in
+non-air regions. Therefore, all neutron diffusion method simulation results in the following
+section were measured with the diffusion coefficient capped at $D_g=2.5$.
+
+Unlike the 2-D hybrid simulations which ran with $S_8$ angular discretization, the 3-D hybrid
+simulations ran with $S_6$ angular discretization due to significantly
+larger computational and memory requirements demanded by the 3-D simulations. Further discussion on
+computational performance may be found at the end of Section \ref{sec:3d-nts}. All simulations,
+except those for temperature reactivity coefficient calculations, ran on isothermal 3-D models at
+911 K.
+
+\subsection{3-D Numerical Results \& Discussion} \label{sec:3d-nts}
+
+\subsubsection{Initial Criticality Configuration}
+
+The \gls{MSRE} achieved initial criticality with Rod 1 inserted to a height of 46.6 inches relative
+to the fully inserted position. For reference, the rods are at a height of 51 inches when fully
+withdrawn. Table \ref{table:initial-crit} shows the $k_\text{eff}$ values from \gls{MSRE}
+experimental data, numerical benchmark data \cite{fratoni_molten_2020}, and this work. All
+numerical models overestimate the $k_\text{eff}$ relative to the experimental value by 1-2\%
+(1000-2000 pcm). In the \gls{MSRE} numerical benchmark report, the authors
+attribute this $k_\text{eff}$ discrepancy largely to discrepancies in the carbon cross section data
+in nuclear data libraries. They noted similar $k_\text{eff}$ overestimations by 1-2\% in benchmark
+results for the HTTR, a graphite-moderated, helium-cooled reactor. The OpenMC and Moltres
+simulations in this work used the same
+ENDF/B-VII.1 nuclear data library as the \gls{MSRE} numerical benchmark report.
+
+\begin{table}[htb]
+  \centering
+  \caption{$k_\text{eff}$ values from \gls{MSRE} experimental data, the \gls{MSRE} numerical
+  benchmark \cite{fratoni_molten_2020}, and the OpenMC and Moltres models in this work.}
+  \begin{tabular}{l S[table-format=1.5(2)]}
+    \toprule
+     & {$k_\text{eff}$} \\
+     \cmidrule(l){2-2}
+    \gls{MSRE} experimental data & 1.00000(420) \\
+    Serpent 2 (Numerical benchmark) & 1.02132(3) \\
+    OpenMC (This work) & 1.01308(20) \\
+    Hybrid (This work) & 1.01955 \\
+    Diffusion (This work) & 1.01885 \\
+    \bottomrule
+  \end{tabular}
+  \label{table:initial-crit}
+\end{table}
+
+Of the four $k_\text{eff}$ values, we expect to see the closest agreement between the Serpent 2
+model from the numerical benchmark and OpenMC from this work. However, the OpenMC \gls{MSRE} model
+differs from the Serpent 2 \gls{MSRE} models due to geometry modifications listed in Sections
+\ref{sec:2d-model-setup} and \ref{sec:3d-model-setup}. The benchmark report quantified the impact
+of several geometry modifications, namely the removal of the flow distributor, the homogenization
+of the sample basket, and the removal of the thermal shield. These modifications contribute to
+-$0.098\%$, -$0.037\%$, and -$0.885\%$ change in the $k_\text{eff}$, respectively, for a final
+$k_\text{eff}$ estimate of $1.01093(3)$. This brings the OpenMC $k_\text{eff}$ estimate of
+$1.01308(20)$ to closer agreement with the numerical benchmark value.
+
+The hybrid method $k_\text{eff}$ estimate is 0.00647 higher than the OpenMC estimate. For this
+initial criticality configuration, the neutron diffusion method $k_\text{eff}$ estimate is closer
+to OpenMC than the hybrid method. Overall, the OpenMC and Moltres models show good agreement with
+the Serpent 2 model from the numerical benchmark.
+Figures * show the neutron group fluxes on the interior of the 3-D models. The salt-graphite
+lattice structure is visible from the group flux distributions. Group 7 and 8 neutron fluxes are
+particularly small near the control rods due to strong thermal neutron capture cross section of
+gadolinium.
+
+\subsubsection{Control Rod Worth}
+
+After achieving initial criticality, \gls{MSRE} researchers ran control rod calibration experiments
+with the \gls{MSRE} at zero-power (typically at about 10 W power output)
+\cite{prince_zero-power_1968}. They measured differential rod worths by measuring the time taken
+for the reactor power to increase by 20 W following a prescribed withdrawal of Rod 1 and applying
+the inhour equation to obtain the reactivity inserted. They repeated this process at different rod
+heights after successive additions of highly enriched uranium capsules. The final integral rod
+worth curve was normalized to the critical $^{235}$U loading of 65.25 kg
+\cite{prince_zero-power_1968}.
+
+For control rod worth calculations, the OpenMC and Moltres simulations ran with Rod 1 at fourteen
+different rod positions from full insertion (0 inch) to full withdrawal (51 inches). The rod
+positional limits of 0 inch and 51 inches represents the full range of deliberate rod motion
+allowable by the control rod drive motor \cite{robertson_msre_1965}.
+
+Figure \ref{fig:rod-worth} shows the integral rod worth curves from the \gls{MSRE} experimental
+data and the OpenMC and Moltres numerical models. By taking the $k_\text{eff}$ when Rod 1 is fully
+inserted as reference, all rod worth curves start at 0 pcm. The change in integral rod worth,
+represented by the gradient in the plot, is steepest as the tip of the rod passes through the
+center of the core. The hybrid method performs well in reproducing the rod worth curve from OpenMC;
+all hybrid method rod worth values fall within one or two standard deviations of OpenMC.
+
+Due to the experimental procedure of computing integral rod worth through cumulative differential
+rod worth measurements, the uncertainty range of the \gls{MSRE} data scales with rod position.
+Both OpenMC and the hybrid method report consistently larger integral rod worths than the
+\gls{MSRE} data across all rod positions. Nevertheless, they are in good agreement with the
+\gls{MSRE} data as compared to the neutron diffusion method, which significantly overestimates the
+integral rod worth.
+Table \ref{table:rod-worth} shows the total rod worth of Rod 1. Similar to observations from Figure
+\ref{fig:rod-worth}, OpenMC and the hybrid method report total rod worths marginally larger than
+the \gls{MSRE} data by 5.1\% and 4.2\%, respectively, while the diffusion method exceeds the
+experimental value by 21.1\%.
+
+\begin{figure}[t]
+  \centering
+  \includegraphics[width=0.8\columnwidth]{rod-worth-curve}
+  \caption{Reactivity inserted by Rod 1 at various rod positions relative to the full insertion.}
+  \label{fig:rod-worth}
+\end{figure}
+
+\begin{table}[t]
+  \centering
+  \caption{Total rod worth of Rod 1 when fully inserted.}
+  \begin{tabular}{l S[table-format=4(2)] S}
+    \toprule
+     & {$\rho_\text{worth}$ [pcm]} & {Percentage discrepancy [\%]}\\
+     \cmidrule(rl){2-2} \cmidrule(l){3-3}
+    \gls{MSRE} experimental data & 2250(67) & {-}\\
+    OpenMC (This work) & 2364(44) & 5.1 \\
+    Hybrid (This work) & 2345 & 4.2 \\
+    Diffusion (This work) & 2725 & 21.1 \\
+    \bottomrule
+  \end{tabular}
+  \label{table:rod-worth}
+\end{table}
+
+The most likely factor for OpenMC and the hybrid method overpredicting the control rod worth is the
+modeling of control rods as solid annular cylinders as opposed to 36 separate control elements
+strung along a flexible stainless steel cable. Figure \ref{fig:rod-bead} depicts a poison element
+and its physical dimensions. Other than the Gd$_2$O$_3$-Al$_2$O$_3$ poison material, each control
+element has a top and bottom closure made of INOR-8 alloy. According to the physical dimensions,
+the closures make up 15.1\% of the total element height. The 3-D numerical \gls{MSRE} model in this
+work omits the outer and inner INOR-8 tubes and the flexible stainless cable, and replaces the top
+and bottom closure with poison material. Fully resolving the closures in Moltres would require
+excessively fine axial mesh resolution and advanced adaptive meshing features not currently
+available. A straightforward corrective measure for future \gls{MSRE} simulations would be to
+retain the solid annular cylinder structure and adjust the rod dimensions or the rod material
+composition to match the expected worth from \gls{MSRE} data or OpenMC simulations with the beaded
+geometry.
+
+\begin{figure}[t]
+  \begin{subfigure}[b]{0.33\columnwidth}
+    \centering
+    \includegraphics[width=\columnwidth]{rod-bead-2}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.65\columnwidth}
+    \centering
+    \includegraphics[width=\columnwidth]{rod-bead}
+  \end{subfigure}
+  \caption{Cutaway view and dimensions of a control rod poison element. Retrieved from
+  \cite{tolson_msre_1967} and \cite{robertson_msre_1965}.}
+  \label{fig:rod-bead}
+\end{figure}
+
+Overall, the results show that the hybrid $S_N$-diffusion method accurately reproduces the control
+rod worth at various levels of insertion. As observed in the 1-D and 2-D models, the 3-D model
+$k_\text{eff}$ values exhibit some discrepancies relative to OpenMC, but the discrepancies do not
+affect rod worth estimates because they are systematic errors. This \gls{MSRE} hybrid model may be
+improved in future work by characterizing and applying neutron albedo at the vacuum boundaries.
+
+\subsubsection{Temperature Reactivity Coefficient}
+
+This work compares isothermal temperature reactivity coefficients calculated from $k_\text{eff}$
+values measured at 872 K and 972 K with 67.86 $^{235}$U loading. This approach is identical to that
+of the \gls{MSRE} experimental measurements \cite{prince_zero-power_1968} and the \gls{MSRE}
+numerical benchmark calculations \cite{fratoni_molten_2020}. The temperature reactivity coefficient
+is calculated by dividing the difference in reactivities by the temperature change (100 K). All
+simulations included the effects of thermal expansion on the material density but left all reactor
+physical dimensions unchanged.
+
+Table \ref{table:temp-coef} compares the isothermal reactivity coefficients from the experimental
+and numerical sources. The OpenMC and Moltres models in this work show closer agreement with the
+\gls{MSRE} data than the Serpent 2 model from the \gls{MSRE} numerical benchmark. All numerical
+results fall within one standard deviation of the \gls{MSRE} data.
+
+\begin{table}[t]
+  \small
+  \centering
+  \caption{Isothermal temperature reactivity coefficients with 67.86 kg $^{235}$U loading from
+  \gls{MSRE} data, the \gls{MSRE} numerical benchmark \cite{fratoni_molten_2020}, and the OpenMC
+  and Moltres models in this work.}
+  \begin{tabular}{l S[separate-uncertainty=true,table-format=2.2(2)] S}
+    \toprule
+     & {Temperature coefficient [pcm K$^{-1}$]} & {Percentage discrepancy [\%]}\\
+     \cmidrule(rl){2-2} \cmidrule(l){3-3}
+    \gls{MSRE} experimental data & 13.41(150) & {-}\\
+    Serpent 2 (Numerical benchmark) & 12.51(23) & -6.7 \\
+    OpenMC (This work) & 13.85(43) & 3.3 \\
+    Hybrid (This work) & 13.98 & 4.3 \\
+    Diffusion (This work) & 14.01 & 4.5 \\
+    \bottomrule
+  \end{tabular}
+  \label{table:temp-coef}
+\end{table}
+
+\subsubsection{Computational Performance}
+
+Beyond demonstrating the hybrid $S_N$-diffusion method, its computational performance is of
+interest as a candidate for 3-D, time-dependent reactor simulations. All simulations ran on the
+Polaris supercomputer system at \gls{ANL}. Each compute node consists of a 2.8-GHz AMD EPYC Milan
+7543P 32-core CPU, four NVIDIA A100 GPUs, and 512 GB of RAM. The GPUs were not utilized in this
+work. Moltres uses the \gls{MPI} communication protocol for parallelizing simulations. For this
+work, simulations ran with 1-to-1 ratio of \gls{MPI} ranks and processor cores (i.e., 32 \gls{MPI}
+ranks for each compute node).
+
+Every 3-D hybrid method $k$-eigenvalue
+simulation of the \gls{MSRE} consisted of 747,688,760 \glspl{DOF} and 1,690,777,728 \glspl{DOF} for
+the neutron diffusion and $S_N$ components, respectively. Each \gls{DOF} corresponds to a scalar
+or angular group flux variable value defined at a mesh node. Each simulation ran with the distributed
+mesh feature, which partitions and distributes the mesh across however many \gls{MPI} ranks used
+in the simulation. All variable and mesh-linked data are also distributed with the mesh. This
+greatly reduces memory requirements compared to storing the full mesh for each \gls{MPI} rank.
+
+Each 3-D \gls{MSRE} simulation required at least forty nodes for sufficient memory to run. This
+sets a strict minimum requirement for the 3-D hybrid simulations. On forty nodes, the simulations
+required approximately 2.2 h to complete for a total of 88 node-hours or 2816 core-hours. The
+distributed mesh capability comes with the cost of increased data transfer times between the $S_N$
+and diffusion solvers due to increased inter-core data transfers. Transfers accounted for
+approximately 0.675 h or 30 \% of the total simulation time. Future research into optimizing mesh
+distribution and transfer-related data caching may help to minimize data transfers and speed-up
+hybrid method simulations. The solution time for the initial neutron diffusion calculation with zero
+drift correction also takes up approximately 30 \% of the total simulation time. After subtracting
+the minor computational cost of evaluating the drift term even if it is zero, the hybrid method
+requires about four times the solution time of the standard neutron diffusion method.
+
+For the hybrid method to be viable for large time-dependent simulations, it must scale reasonably
+well on computing clusters. For this work, I investigated strong scaling, which measures how the
+solution time decreases when more cores are used while keeping the problem size fixed. This strong
+scaling study used 3-D quarter-core models run on 10, 20, 30, and 40 compute nodes.
+Figure * shows the results of this scaling study.
+The hybrid method implementation scales well as shown by the nearly linear scaling between the
+solution time and the number of processors. Therefore, the hybrid method appears promising for the
+modeling of larger reactor analyses with little scaling performance penalty.
+
+% Strong scaling tests & discussion
+% Difficulties with 3-D modeling
+% SN void treatment constants, Transfer times
 
@@
 
-TBD
-
-%The hybrid method does not alter the $k_\text{eff}$ and flux estimates relative to the neutron
-%diffusion method in cases where the neutron diffusion method provides
+This chapter covers work demonstrating the hybrid $S_N$-diffusion method implemented in Moltres
+through 1-D, 2-D, and 3-D neutronics simulations of \gls{MSRE} models. The hybrid method combines
+the $S_N$
+and diffusion methods by applying transport corrections generated using the high-fidelity $S_N$
+method near control rods while treating the rest of the problem with the lower-fidelity diffusion
+method for tractable 3-D time-dependent reactor analyses. This is achieved by identifying and
+setting up a subregion containing the control rod and its vicinity for $S_N$ transport
+calculations. The $S_N$ and diffusion solvers are coupled through boundary sources at the subregion
+boundaries. The hybrid method applies an adaptive algorithm for discarding inaccurate transport
+correction values near the subregion boundaries and preserving the smoothness of the neutron flux
+gradient. In this chapter, the 1-D simulations provided verification of the hybrid method for
+estimating control rod worth. The simulations also compared two different implementations of the
+hybrid method which differed by the use of drift or diffusion correction schemes.
+The 2-D and 3-D simulations extended the hybrid method to multi-dimensional problems. Additionally,
+the 3-D simulations validated the numerical \gls{MSRE} model against \gls{MSRE} experimental data
+and provided useful insight on the computational performance of the hybrid method for
+large reactor analyses.
+
+Six 1-D test cases based on the \gls{MSRE} with increasing geometric complexity were run using the
+standard $S_N$ method, the hybrid method, and standard neutron diffusion method and verified against
+reference OpenMC Monte Carlo neutron transport simulations under continuous neutron energy (OpenMC-CE) and
+multigroup (OpenMC-MG) modes. The $S_N$ implementation with nonlinear diffusion acceleration
+showed good agreement with OpenMC-MG which indicated that discrepancies between those codes and
+OpenMC-CE largely arose from the eight neutron energy group structure for multigroup simulations.
+Both hybrid method implementations, with drift or diffusion correction schemes, showed similar
+performance in the 1-D cases. While the $k_\text{eff}$ estimates from the hybrid method deviated
+from the $S_N$ method and OpenMC-MG by 100 to 400 pcm, the hybrid method accurately reproduces
+control rod worths. Rod worth estimates from OpenMC-MG, $S_N$, and hybrid methods differed from
+OpenMC-CE by 2 to 3 \%. The hybrid method improves on the neutron diffusion method whose rod worth
+error estimate is 8 \%. The hybrid method also produced better agreement in neutron flux
+distributions with the neutron transport methods compared to the neutron diffusion method.
+
+Analysis on the drift and diffusion correction schemes showed that the diffusion correction
+parameter led to numerous discontinuities tending to infinity values near points in the problem
+domain where flux peaks and troughs occur. These discontinuities and negative diffusion coefficient
+values are unphysical and would lead to numerical stability issues. Therefore, the hybrid method
+implementation with drift correction scheme was chosen for the remaining 1-D analyses and the
+2-D and 3-D analyses.
+
+Analysis on the impact of $S_N$ correction subregion size found that it has little impact on the
+control rod worth estimates as long as the subregion size is kept consistent (i.e., same size) when
+performing control rod worth measurements. Rod worths varied by less than 0.35 \% across the
+different subregion sizes investigated. A separate analysis showed that the $k_\text{eff}$
+estimates from the hybrid method exhibit superlinear convergence rate when tightening the $S_N$
+convergence tolerance value. The convergence tolerance value also impacts the number of outer
+iterations between $S_N$ and diffusion solvers. After just two outer iterations, the error in
+$k_\text{eff}$ fell to below 10$^{-7}$.
+
+The hybrid method continued to show improved control rod estimates over the neutron diffusion
+method in 2-D quarter-core and full-core \gls{MSRE} simulations with up to three rods inserted in the
+full-core model. The hybrid method reported rod worth error magnitudes of less than 40 pcm relative
+to OpenMC-CE, which are significant improvements over the neutron diffusion method error estimates
+ranging from 569 pcm to 1484 pcm. The hybrid method also showed significant improvements in the
+absolute mean and maximum errors in fuel channel power distributions relative to the reference
+OpenMC-CE distribution.
+
+For 3-D simulations, the \gls{MSRE} experimental data and the \gls{MSRE} numerical benchmark
+report \cite{fratoni_molten_2020} both served as references for model validation. The numerical
+benchmark is included in the International Reactor Physics Experiment Evaluation Project (IRPhEP)
+handbook. The
+$k_\text{eff}$ estimates of the \gls{MSRE} at initial criticality from OpenMC, the hybrid method,
+and the neutron diffusion method in this work showed good agreement with the \gls{MSRE} numerical
+benchmark. All numerical $k_\text{eff}$ estimates, including that from the benchmark report
+\cite{fratoni_molten_2020}, exceeded the \gls{MSRE} experimental value by 1-2 \% due to possible
+biases and uncertainties in the neutronic properties of graphite. Temperature reactivity
+coefficient values from OpenMC, the hybrid method, and the neutron diffusion method also show good
+agreement with the \gls{MSRE} data within experimental uncertainty with percentage discrepancies of
+about 3 \%, but differ from the numerical benchmark which has a percentage discrepancy of -6.7 \%.
+
+Control rod worth calculations were also performed for Rod 1 of the \gls{MSRE} under various levels
+of insertion from its fully inserted to its fully withdrawn state. OpenMC and the hybrid method
+showed excellent agreement at every stage within one standard deviation. Due to geometric
+approximations of the control rods in the numerical models, they overestimated the total
+rod worth relative to the \gls{MSRE} experimental data by approximately 4-5 \%. The hybrid method
+compares favorably against the neutron diffusion method which overestimated the total worth by
+21.1 \%.
+
+The 3-D simulations performed well under the strong scaling test which studies how the solution
+time scales with the number of processors used to run the simulation. The 3-D simulations showed
+close to linear scaling within the scope of the strong scaling test. Therefore, the hybrid method
+exhibits good computational scaling performance which is essential for large 3-D reactor analyses.
+Excessively large memory requirements were mitigated through the distributed mesh capability in
+Moltres. Some performance optimizations may be possible to reduce data transfer times
+between the $S_N$ and diffusion solvers which currently take up about 30 \% of the total solution
+time as a result of the inter-processor communications.
+
+Overall, the hybrid $S_N$-diffusion method provides a significant improvement over the neutron
+diffusion method at only four times the computational expense. Control rod estimates
+show good agreement with reference Monte Carlo solutions and \gls{MSRE} experimental data.
+Further investigations into characterizing and minimizing neutron leakage rates at vacuum
+boundaries may help with reducing $k_\text{eff}$ discrepancies with reference Monte Carlo
+solutions.

### 2024-11-07 transient.tex
Rename files

\section{Rod Drop Experiment}

\section{Reactivity Insertion Experiment}

### 2024-11-07 transient.tex
Chapter 6 intro

@@
-\section{Rod Drop Experiment}
+Chapter \ref{chap:msre} verified the hybrid $S_N$-diffusion method implemented in Moltres through
+1-D, 2-D, and 3-D
+$k$-eigenvalue simulations of the \gls{MSRE} against the OpenMC Monte Carlo neutron transport code
+for molten salt reactor control rod modeling. The 3-D simulations validated the 3-D \gls{MSRE}
+model in this work against \gls{MSRE} experimental data and the \gls[MSRE} numerical benchmark
+study \cite{fratoni_molten_2020} in the context of the initial criticality and rod worth
+measurement experiments. The 3-D simulations also showed that the hybrid method runs with
+reasonable computational costs compared to the standard neutron diffusion method and exhibited
+efficient strong scaling across compute nodes.
 
-\section{Reactivity Insertion Experiment}
+This chapter builds on the foundations of Chapter \ref{chap:msre} to verify and validate the hybrid
+method for time-dependent 3-D simulations of reactivity-initiated transients. This chapter includes
+two demonstrations using the hybrid method of the \gls{MSRE} rod drop and reactivity insertion
+experiments. Beyond time-dependent modeling, the simulations will also verify and validate
+compatibility between the existing neutronics/thermal-hydraulics coupling capabilities in Moltres
+and the newly-implemented hybrid method.
+
+Section \ref{sec:rod-drop}
+
+\section{Rod Drop Experiment} \label{sec:rod-drop}
+
+\subsection{Model Setup}
+
+\subsection{Rod Drop Numerical Results \& Discussion}
+
+\section{Reactivity Insertion Experiment} \label{sec:reactivity-insertion}
+
+\subsection{Model Setup}
+
+\subsection{Reactivity Insertion Numerical Results \& Discussion}
+
+\section{Summary} \label{sec:transient-summary}

### 2024-11-07 transient.tex
Rod drop content

@@
 
-Section \ref{sec:rod-drop}
+Section \ref{sec:rod-drop} covers the model setup, numerical results, and discussion of the rod
+drop experiment. Section \ref{sec:reactivity-insertion} covers the model setup, numerical results,
+and discussion of the reactivity insertion experiment. Lastly, Section \ref{sec:transient-summary}
+summarizes the goals and outcomes in this chapter.
 
 \section{Rod Drop Experiment} \label{sec:rod-drop}
+
+\gls{MSRE} researchers conducted a series of rod drop experiments at three different $^{235}$U
+loadings \cite{prince_zero-power_1968}. These experimented were conducted from an initially critical
+and low power state with an initial neutron count rate of approximately 30,000 from neutron detectors
+placed near the reactor. A rapid-action camera photographed the integral count shown on the
+detectors about once every second starting from a few seconds before the rod drop to about thirty
+seconds after the rod drop. This work aims to replicate the integral count curve following a rod
+drop of Rod 1 with the $^{235}$U loading corresponding to "30 capsule additions" of highly-enriched
+$^{235}$U capsules.
 

### 2024-11-08 hybrid.tex
Chapter 6 content

@@
 
-\subsection{Spatial Discretization and Void Treatment} \ref{sec:void-treatment}
+\subsection{Spatial Discretization and Void Treatment} \label{sec:void-treatment}
 
@@
 
-\subsection{Setting Caps on the Diffusion Coefficient Values} \ref{sec:diffcoef-cap}
+\subsection{Setting Caps on the Diffusion Coefficient Values} \label{sec:diffcoef-cap}
 

### 2024-11-08 msre.tex
Chapter 6 content

@@
 testbed for \gls{HPC} scaling tests, and preparation for the time-dependent multiphysics
-simulations in Chapter 6.
+simulations in Chapter \ref{chap:transient}.
 
@@
 Beyond demonstrating the hybrid $S_N$-diffusion method, its computational performance is of
-interest as a candidate for 3-D, time-dependent reactor simulations. All simulations ran on the
+interest as a candidate for time-dependent 3-D reactor simulations. All simulations ran on the
 Polaris supercomputer system at \gls{ANL}. Each compute node consists of a 2.8-GHz AMD EPYC Milan
@@
 This chapter covers work demonstrating the hybrid $S_N$-diffusion method implemented in Moltres
-through 1-D, 2-D, and 3-D neutronics simulations of \gls{MSRE} models. The hybrid method combines
-the $S_N$
+through 1-D, 2-D, and 3-D $k$-eigenvalue simulations of \gls{MSRE} models. The hybrid method
+combines stengths of the $S_N$
 and diffusion methods by applying transport corrections generated using the high-fidelity $S_N$
 method near control rods while treating the rest of the problem with the lower-fidelity diffusion
-method for tractable 3-D time-dependent reactor analyses. This is achieved by identifying and
+method for tractable time-dependent 3-D reactor analyses. This is achieved by identifying and
 setting up a subregion containing the control rod and its vicinity for $S_N$ transport

### 2024-11-08 transient.tex
Chapter 6 content

@@
 for molten salt reactor control rod modeling. The 3-D simulations validated the 3-D \gls{MSRE}
-model in this work against \gls{MSRE} experimental data and the \gls[MSRE} numerical benchmark
+model in this work against \gls{MSRE} experimental data and the \gls{MSRE} numerical benchmark
 study \cite{fratoni_molten_2020} in the context of the initial criticality and rod worth
@@
 This chapter builds on the foundations of Chapter \ref{chap:msre} to verify and validate the hybrid
-method for time-dependent 3-D simulations of reactivity-initiated transients. This chapter includes
-two demonstrations using the hybrid method of the \gls{MSRE} rod drop and reactivity insertion
-experiments. Beyond time-dependent modeling, the simulations will also verify and validate
-compatibility between the existing neutronics/thermal-hydraulics coupling capabilities in Moltres
-and the newly-implemented hybrid method.
+method for neutronics modeling in time-dependent 3-D simulations of reactivity-initiated
+transients. This chapter presents two sets of simulation results involving the hybrid method and
+other capabilities in Moltres of the \gls{MSRE} rod drop and reactivity insertion
+experiments. Beyond time-dependent modeling with the hybrid method, the simulations will also
+demonstrate compatibility between the existing neutronics/thermal-hydraulics coupling
+capabilities in Moltres and the newly-implemented hybrid method.
 
@@
 loadings \cite{prince_zero-power_1968}. These experimented were conducted from an initially critical
-and low power state with an initial neutron count rate of approximately 30,000 from neutron detectors
-placed near the reactor. A rapid-action camera photographed the integral count shown on the
-detectors about once every second starting from a few seconds before the rod drop to about thirty
-seconds after the rod drop. This work aims to replicate the integral count curve following a rod
+and low power state with an initial neutron count rate of approximately 30,000 counts per second
+from neutron detectors placed near the reactor. A rapid-action camera photographed the integral
+count shown on the detectors about once every second starting from a few seconds before the rod
+drop to about thirty seconds after the rod drop.
+
+This work aims to replicate the integral count curve following a rod
 drop of Rod 1 with the $^{235}$U loading corresponding to "30 capsule additions" of highly-enriched
-$^{235}$U capsules.
+$^{235}$U capsules. Each capsule contains 85 g of $^{235}$U, of which 94.94 \% enters the core due
+to residual fuel salt remaining in the drain tanks after mixing \cite{fratoni_molten_2020}.
+
+\subsection{Model Setup} \label{sec:rod-drop-setup}
+
+The key experimental parameters to simulate the rod drop experiment are: the exact $^{235}$U
+loading, the initial and final rod heights, the total reactivity withdrawn, and the speed of the
+rod drop.
+
+The \gls{MSRE} achieved initial criticality after eight capsule additions with Rod 1 inserted to a
+height of 46.6 inches relative to the fully inserted height \cite{prince_zero-power_1968}.
+From the \gls{MSRE} experimental rod worth
+data, the rod worth evaluates to approximately 72 pcm relative to the fully withdrawn state; the
+$k_\text{eff}$ of the \gls{MSRE} at initial criticality with Rod 1 fully withdrawn would be
+1.00072. According to the same \gls{MSRE} report, the
+authors back-calculated the critical $^{235}$U loading with Rod 1 fully withdrawn to be 65.25 kg.
+They also provided a $^{235}$U concentration reactivity coefficient of 0.223 $(\delta k/k)/\delta m
+/m)$. The excess $^{235}$U mass at initial criticality compensating for Rod 1 inserted to 46.6
+inches can estimated through the following steps:
+%
+\begin{align}
+  \frac{\Delta k / k}{\Delta m / m} &= 0.223 \nonumber \\
+  \frac{m}{k}\frac{dk}{dm} &= 0.223 \nonumber \\
+  \int^{1.00072}_1 \frac{1}{k}\ dk &= 0.223 \int^{65.25+\Delta m}_{65.25} \frac{1}{m}\ dm \nonumber \\
+  \ln 1.00072 &= 0.223 \ln\left(\frac{65.25+\Delta m}{65.25}\right) \nonumber \\
+  \Delta m &= 0.2109 \mbox{ kg}.
+\end{align}
+%
+This change in mass is equivalent to approximately 2.61 highly-enriched $^{235}$U capsules.
+Therefore, the $^{235}$U loading for the rod drop experiment is:
+%
+\begin{gather}
+  m_\text{rod drop} = 65.25 \text{kg} + (30 - 8 + 2.61) * 0.085 \text{kg} * 0.9494 = 67.24 \text{kg}
+\end{gather}
+%
+The $k_\text{eff}$ of the \gls{MSRE} with 67.24 kg loading and all rods fully withdrawn is:
+%
+\begin{align}
+  \int^{k_{67.24 \text{kg}}}_1 \frac{1}{k}\ dk &= 0.223 \int^{67.24}_{65.25} \frac{1}{m}\ dm \nonumber \\
+    k_{67.24 \text{kg}} &= 1.00671
+\end{align}
+
+\begin{figure}[t]
+  \centering
+  \includegraphics[width=0.6\columnwidth]{msre-rod-worth}
+  \caption{Integral control rod worth of Rod 1 \cite{prince_zero-power_1968} at critical and final
+  $^{235}$U loading. The final loading rod worth curve is scaled down by 1.087 relative to the
+  critical loading rod worth curve.}
+  \label{fig:msre-rod-worth-2}
+\end{figure}
+
+When loading more $^{235}$U into the core, the shape of the integral rod worth does not change,
+but the magnitude scales down by a factor of 1.087 between the critical and final loading
+as shown in Figure \ref{fig:msre-rod-worth-2}. After applying this scaling factor to the monotone cubic
+spline interpolated \gls{MSRE} rod worth curve, the initial height of Rod 1 to keep the \gls{MSRE}
+with 67.24 kg loading critical is approximately 30.75 inches. During the rod drop experiment, the rod
+drops from this height to its fully inserted state at 0 inches. This drop induces a total
+reactivity withdrawal of 1500 pcm.
+
+The rod drop simulation uses the hybrid $S_N$-diffusion method in Moltres on the verified 3-D
+\gls{MSRE} model from Chapter
+\ref{chap:msre} at the critical $^{235}$ loading. Refer to Section \ref{sec:3d-model-setup} for
+details on the 3-D model. The initial rod height for the \gls{MSRE} model to induce a similar total
+reactivity withdrawal of 1500 pcm from Rod 1 drop is 29.4 inches. The acceleration of the
+rod is approximately 15 ft s$^{-2}$ from prior testing \cite{prince_zero-power_1968}. At this
+acceleration, the rod reaches its fully inserted height at $t=0.4046$ s. Figure \ref{fig:rod-height}
+shows the evolution of rod height at each timestep within the first second of the rod drop simulation.
+
+\begin{figure}[htb!]
+  \centering
+  \includegraphics[width=0.7\columnwidth]{rod-height}
+  \caption{Evolution of rod height evaluated at each timestep within the first second of the rod
+  drop simulation. The rod reaches its fully inserted height at $t=0.4046$ s.}
+  \label{fig:rod-height}
+\end{figure}
+
+A significant change for the rod drop simulation compared to the $k$-eigenvalue 3-D simulation,
+other than the time-dependence, is the inclusion of \glspl{DNP} for the delayed neutron response.
+I used the \gls{MOOSE} MultiApp capability available in Moltres to set up a nested iterative
+coupling structure for the hybrid
+$S_N$-diffusion method with \glspl{DNP} such that the neutron diffusion and \gls{DNP} solvers are
+coupled through inner iterations and the $S_N$ solver is coupled them through outer iterations.
+Figure \ref{fig:rod-drop-coupling} illustrates this iterative coupling structure and the data
+transfers between the solvers.
+
+\begin{figure}[htb!]
+  \centering
+%  \includegraphics[width=0.8\columnwidth]{rod-height}
+  \caption{The nested iteration structure coupling the $S_N$, neutron diffusion, and \gls{DNP}
+  solvers for the rod drop simulation using the hybrid $S_N$-diffusion method.}
+  \label{fig:rod-drop-coupling}
+\end{figure}
+
+The output of a $k$-eigenvalue simulation with the rod at the prescribed initial height provides
+the initial group flux and \gls{DNP} distributions for the rod drop simulation. The time-depedent
+rod drop simulation uses the \gls{MOOSE} Restart capability to load the variable distributions from
+the prior $k$-eigenvalue simulation. The total fission neutron production rate at every timestep
+is normalized to match the 30,000 counts per second at $t=0$ s from the \gls{MSRE} rod drop
+experimental measurements.
+
+\subsection{Rod Drop Numerical Results \& Discussion}
+
+Figure \ref{fig:integral-count} compares the integral neutron count during the rod drop experiment
+from \gls{MSRE} experimental data and the hybrid method numerical results. The hybrid method
+reproduces the expected shape of the integral
+neutron count characterized by the drop in neutron count rate leading to a gradual flattening of
+the integral count curve. The discrepancy in integral neutron count at around $t=16$ s is x \%.
+Experimental uncertainties surrounding the rod drop experiment include the actual $^{235}$U
+loading, initial rod height, rod acceleration, and the initial neutron count rate. These uncertainties
+may have affected the reactivity withdrawal estimates in Section \ref{sec:rod-drop-setup}.
+
+\begin{figure}[t]
+  \centering
+  \includegraphics[width=0.8\columnwidth]{integral-count}
+  \caption{Integral neutron count during the rod drop experiment from \gls{MSRE} experimental data
+  and hybrid method numerical results. The hybrid method agrees closely with the \gls{MSRE} data
+  with a discrepancy of x \% at around $t=16$ s.}
+  \label{fig:integral-count}
+\end{figure}
+
+Figures * show snapshots of the neutron group flux distributions at various times during the rod
+drop simulation. The neutron flux magnitudes fall as the reactor enters a sub-critical state. The
+Rod 1 induces a significant drop in neutron fluxes near the center of the core.
+
+\section{Reactivity Insertion Experiment} \label{sec:reactivity-insertion}
+
+After a long operational run on $^{235}$U fuel, \gls{ORNL} extended the \gls{MSRE} program to
+demonstrate reactor operation with $^{233}$U fuel. The \gls{MSRE} then became the first nuclear
+reactor to run on $^{233}$U bred from $^{232}$Th. During a series of tests for experimental
+dynamic analyses, \gls{MSRE} researchers conducted reactivity insertion tests with the reactor
+operating at 1, 5, and 8 MW \cite{steffy_experimental_1970}.
+In contrast to the rod drop experiments, these reactivity insertion
+experiments were conducted with the \gls{MSRE} in power operation.
+
+This work aims to replicate the power response following a reactivity insertion of 24.8 pcm of the
+\gls{MSRE} initially at 1 MW. Figure \ref{msre-reactivity-insertion} shows the theoretical and
+experimental power response following the reactivity insertion. Three separate publications
+\cite{zanetti_innovative_2015, singh_study_2019, mochizuki_validation_2022} noted that the plot is
+mislabeled with the wrong reactivity insertion value of 0.0139 \% $\delta k/k$, which had been the
+reactivity insertion value for the 8 MW case. The correct reactivity insertion value is 24.8 pcm as
+mentioned prior.
+
+\begin{figure}[t]
+  \centering
+  \includegraphics[width=0.8\columnwidth]{msre-reactivity-insertion}
+  \caption{Power response following a step reactivity insertion of 24.8 pcm with the reactor
+    initially at 1 MW \cite{steffy_experimental_1970}. Note that the authors of the \gls{MSRE}
+    report mislabeled the reactivity insertion value on the plot
+  \cite{zanetti_innovative_2015, singh_study_2019, mochizuki_validation_2022}.}
+  \label{fig:msre-reactivity-insertion}
+\end{figure}
 
@@
 
-\subsection{Rod Drop Numerical Results \& Discussion}
+For the reactivity insertion simulation, we must first identify the $^{233}$U salt composition used
+in the experiment. The \gls{MSRE} $^{233}$ power operations occurred during Runs 16 to 18 according
+to \gls{MSRE} reports \cite{thoma_chemical_1971, burke_analysis_2019}. Table \ref{table:u233-salt}
+lists the proportion of elements in the \gls{MSRE} fuel salt during Runs 16 to 18. Noted that the
+components do not sum up to 100 \% due to experimental uncertainty in chemical analysis and the
+presence of other elements that were not measured. Taking the nominal uranium value to be true, I scaled
+up the compositions of non-uranium elements to make up the remaining $(100\%-0.8155\%)$ of the salt. I
+kept the nominal weight percentage of uranium constant because it is the most accurate value in the table
+and has the biggest impact on the reactivity.
+%
+\begin{table}[t]
+  \small
+  \centering
+  \setlength\tabcolsep{4pt}
+  \caption{\gls{MSRE} fuel salt mean composition during the power operations with $^{233}$U \cite{thoma_chemical_1971}.
+  All values were measured via chemical analysis except the nominal U weight percentage, which was
+  calculated from inventory tracking of cumulative U additions to the salt.}
+  \begin{tabular}{l S[table-format=2.2(2)] S[table-format=1.2(2)] S[table-format=2.2(2)] S[table-format=2.2(3)] S[table-format=1.3(2)] S}
+    \toprule
+    Element & {Li} & {Be} & {Zr} & {F} & {U} & {U (Nominal)} \\
+    \midrule
+    Weight percentage [\%]  & 11.33(39) & 6.71(30) & 10.99(29) & 68.71(126) & 0.809(11) & 0.8155 \\
+    \bottomrule
+  \end{tabular}
+  \begin{tabular}{l S[table-format=2(1)] S[table-format=3(2)] S[table-format=2(2)] S}
+    \toprule
+    Element & {Cr} & {Fe} & {Ni} & {O} \\
+    \midrule
+    Parts per million [ppm] & 72(9) & 135(24) & 46(12) & 61 \\
+    \bottomrule
+  \end{tabular}
+  \label{table:u233-salt}
+\end{table}
 
-\section{Reactivity Insertion Experiment} \label{sec:reactivity-insertion}
+For the isotopic content of uranium, I used a composition restruction strategy developed by Burke
+\cite{burke_analysis_2019}. Their analysis determined that the $^{233}$U salt consists of residual
+salt from $^{235}$U operations, a depleted uranium feed, and a stock $^{233}$U feed. The residue
+and depleted uranium feed masses, taken from Burke's work, are 1.935 kg and 0.890 kg, respectively.
+The stock feed uranium
+mass is estimated to be 35.234 kg reported on January 12 1969 \cite{thoma_chemical_1971}. The
+experiment was performed on January 16 1969 \cite{steffy_experimental_1970}. Table * shows
+the breakdown and total isotopic content of uranium in the salt. The density of $^{233}$U salt is
+2.264 g cm$^{-3}$ at 911 K \cite{thoma_chemical_1971}.
+%
+\begin{table}[t]
+  \small
+  \centering
+  \setlength\tabcolsep{4pt}
+  \caption{Uranium isotopic composition reconstruction table \cite{burke_analysis_2019}.}
+  \begin{tabular}{l S S S S S S S}
+    \toprule
+    \multirow{2}{*}{Uranium Source} & {\multirow{2}{*}{Mass [kg]}} & \multicolumn{6}{c}{Weight percentage [\%]} \\
+                                    & & {$^{232}$U} & {$^{233}$U} & {$^{234}$U} & {$^{235}$U} & {$^{236}$U} & {$^{238}$U} \\
+                                    \cmidrule(r){1-1} \cmidrule(rl){2-2} \cmidrule(l){3-8}
+    Residue & 1.935 & & & 0.343 & 32.705 & 0.483 & 66.469 \\
+    Depleted & 0.890 & & & & 0.3 & & 99.7 \\
+    Stock & 35.234 & 0.02 & 91.49 & 7.6 & 0.7 & 0.05 & 0.14 \\
+    \bottomrule
+  \end{tabular}
+  \label{table:u233-composition}
+\end{table}
 
-\subsection{Model Setup}
+Due to constraints on computational resources, the reactivity insertion simulation ran a 3-D
+quarter-core \gls{MSRE} model. Refer to Figure \ref{fig:1/4-geom} for the horizontal cross section
+view of the 3-D quarter-core model. Figure * shows the side view from a vertical side of the
+quarter-core and the direction of salt flow in each major region. The regions are the salt-graphite
+lattice, the upper plenum, the out-of-core components, the downcomer, and the lower plenum. The
+model assumes a flat salt velocity profile in each region. The actual total volumetric flow rate of
+the \gls{MSRE} is 1200 gpm (gallons per minute) \cite{robertson_msre_1965}. Dividing this value by
+four for the quarter-core model and converting to SI units gives a volumetric flow rate of
+0.01893 m$^3$ s$^{-1}$.
+
+Table \ref{table:msre-flow} lists the dimensions, salt flow velocities, and salt residence
+times in each region to maintain the calculated volumetric flow rate. The salt residence times are
+consistent with \gls{MSRE} specifications \cite{robertson_msre_1965}. The out-of-core region is
+modeled as a 1-D pipe due to its lower neutronic importance other than \gls{DNP} decay and heat
+removal through the heat exchanger. The out-of-core region salt velocity is set to the same magnitude
+as the downcomer for ease of implementation with the height adjusted accordingly to maintain the
+out-of-core salt residence time. Note that the velocity in the downcomer is negative since salt
+flows downwards unlike in the other regions.
+%
+\begin{table}[t]
+  \small
+  \centering
+  \caption{Dimensions, salt flow velocities, and salt residence times in each region of the \gls{MSRE}.}
+  \begin{tabular}{l S S S S}
+    \toprule
+    Component & {Salt cross-sectional area [m$^2$]} & {Height [m]} & {Velocity [m s$^{-1}$]} & {Time [s]} \\
+    \cmidrule(r){1-1} \cmidrule(l){2-5}
+    Upper plenum & 0.42411 & 0.2539 & 0.04463 & 5.69 \\
+    Lattice      & 0.08646 & 1.6637 & 0.21891 & 7.60 \\
+    Lower plenum & 0.42614 & 0.1875 & 0.04935 & 3.80 \\
+    Out-of-core  & {-}     & 5.3081 & 0.65532 & 8.1 \\
+    Downcomer    & 0.02888 & 1.6637 & -0.65532 & 2.539 \\
+    \bottomrule
+  \end{tabular}
+  \label{table:msre-flow}
+\end{table}
+
+% Unlike with $^{235}$U salt, there is no widely recognized benchmark value
+% for the density of $^{233}$U salt. During \gls{MSRE} operation, researchers were still developing
+% reliable techniques for measuring the density of molten salt mixtures \cite{thoma_chemical_1971}.
+% \gls{MSRE} $^{233}$U fuel salt with has a lower density than $^{235}$U salt largely due to its
+% lower $^{238}$U content.
+
+The MultiApp nested iterative coupling structure for this reactivity insertion simulation is similar
+to the implementation for the rod drop simulation. Figure \ref{fig:insertion-coupling}
+provides an updated illustration of the nexted iteration coupling with temperature. The \gls{DNP}
+and temperature distributions are solved together with additive Schwarz method preconditioning in
+Moltres since they are both advective problems. The 1-D out-of-core model accounts for \gls{DNP}
+loss and heat removal outside the core. Delayed neutrons produced in this region are considered
+lost and do not contribute to the neutron flux in the core.
+
+\begin{figure}[htb!]
+  \centering
+%  \includegraphics[width=0.8\columnwidth]{rod-height}
+  \caption{The nested iteration structure coupling the $S_N$, neutron diffusion,
+  \gls{DNP}-temperature, and out-of-core
+  solvers for the rod drop simulation using the hybrid $S_N$-diffusion method.}
+  \label{fig:insertion-coupling}
+\end{figure}
+
+The reactivity insertion simulation requires a prior ``steady-state'' simulation to obtain the group
+flux, \gls{DNP}, and temperature distributions under 1 MW power operation and steady salt flow. The
+``steady-state'' simulation is a time-dependent simulation with timestep sizes of 1 s to advance
+the model towards 1 MW steady state operation. Following this, the reactivity insertion simulation
+uses the \gls{MOOSE} Restart capability to load in variable distributions for the initial conditions.
+
+\subsection{Preliminary Rod Worth Calculations}
+
+The control rod worth of the \gls{MSRE} quarter-core model with $^{233}$U fuel must be determined
+to estimate the initial rod height and rod travel necessary for the reactivity insertion
+simulation. We start this process with comparisons of Rod 1 worths between a full-core OpenMC model
+and experimental data. Table \ref{table:u233-rod-worth} compares the total rod worth of Rod 1 at
+922 K between the OpenMC model from this work
+and \gls{MSRE} data. The rod worth estimate from OpenMC agrees closely with the predicted (i.e.,
+calculated) \gls{MSRE} value. The measured \gls{MSRE} rod worth value is about 7 \% smaller than
+the predicted value. \gls{MSRE} researchers noted poor precision in their rod worth measurements
+due to ``wear and aging'' in the control rod drive assembly \cite{engel_zero-power_1972}. This may
+be responsible for some uncertainty. However, the actual rod worth is likely closer to the measured
+value than the predicted value, and the \gls{MSRE} model of this work is likely to contain as yet
+unknown biases. Nevertheless, the rod worth estimate is close enough that it should not affect the
+\gls{MSRE} power response following a fixed reactivity insertion.
+%
+\begin{table}[t]
+  \centering
+  \caption{Total rod worth of Rod 1 at 922 K from OpenMC full-core \gls{MSRE} model and \gls{MSRE}
+  experimental data.}
+  \begin{tabular}{l S[table-format=4(3)] S}
+    \toprule
+    & {Rod worth [pcm]} & {Percentage discrepancy [\%]} \\
+    \cmidrule(r){1-1} \cmidrule(l){2-3}
+    MSRE (measured) & 2580(129) & {-} \\
+    MSRE (predicted) & 2750 & 6.6 \\
+    OpenMC (this work) & 2771 & 7.4\\
+    \bottomrule
+  \end{tabular}
+  \label{table:u233-rod-worth}
+\end{table}
+
+The 3-D quarter-core \gls{MSRE} model for the reactivity insertion simulation contains only one
+control rod, but it is equivalent to four rods due to rotational symmetry of the quarter-core model.
+The total rod worth of the 3-D quarter-core \gls{MSRE} model using the hybrid method on Moltres is
+9702 pcm, just under four times the worth of Rod 1 from Table \ref{table:u233-rod-worth} due to
+rod shadowing effects that reduce the worth of adjacent control rods. Figure \ref{fig:u233-rod-worth-curve} shows the rod
+worth values from Moltres for the 3-D quarter-core model and the monotone cubic spline interpolation
+curve. Despite having fewer data points than the $^{235}$U rod worth analysis (Section
+\ref{sec:3d-nts}), this interpolation curve is valid for the reactivity insertion experiment since
+the rod heights will strictly remain within the middle
+range where the change in rod worth is approximately linear to changes in rod height.
+
+The required
+rod travel in the quarter-core model for a 24.8 pcm reactivity insertion in this linear region
+is 0.105 inches (0.267 cm). The estimated rod travel required to induce the same reactivity
+insertion in the actual \gls{MSRE} reactor is 0.396 inches (1.00 cm) obtained by scaling with the
+total measured \gls{MSRE} rod worth in Table \ref{table:u233-rod-worth}. Using
+an estimated rod speed of 1.27 cm s$^{-1}$ \cite{robertson_msre_1965}, the reactivity insertion
+occurs over a 0.791-s ramp (reactivity insertions at this timescale are considered step reactivity
+insertions). The rod is assumed to accelerate and decelerate instantaneously given
+the low rod speed.
+The reactivity insertion simulation ran with timestep sizes of 0.198 s during rod movement and
+0.25 s after the rod reaches its final position. These timestep sizes were chosen based on
+considerations balancing accuracy and available computational resources. The main purpose of this
+work is the demonstration of the hybrid method for time-dependent control rod modeling; a detailed
+time-stepping truncation error analysis is out of the scope of this work.
+
+\begin{figure}[t]
+  \centering
+  \includegraphics[width=0.8\columnwidth]{u233-rod-worth-curve}
+  \caption{Rod worth data points and the monotone cubic spline interpolation curve from $k$-eigenvalue
+  hybrid method simulations on the 3-D quarter-core \gls{MSRE} model.}
+  \label{fig:u233-rod-worth-curve}
+\end{figure}
 
@@
 
+Figure \ref{fig:insertion-power} shows the \gls{MSRE} data and simulation results for change in
+power output following the 24.8 pcm reactivity
+insertion. From the experimental data, the power output rises rapidly at the start of the
+reactivity insertion before oscillating back down below the initial power output. Power output
+peaks again later at around $t=360$ s. Moltres reproduces the same trend in power output, but 
+over/under-predicts the first power peak by y \%. It also over/under-predicts the subsequent power
+oscillations.
+Significant contributors to the discrepancies are the uncertainties in the physical properties of
+the fuel salt, temperature reactivity coefficient, and the dynamic response of the heat exchanger.
+
+Figure \ref{fig:insertion-temperature} shows the maximum fuel salt and graphite temperatures, and
+the mean salt inlet and outlet
+temperatures following the reactivity insertion. The increase in maximum temperatures is more
+gradual than the increase in power output as the salt and graphite take time to heat up. The salt
+outlet temperature follows closely behind as the heated salt flows out of the core. The salt inlet
+temperature lags behind the other temperature measurements due to the time taken for the heated
+salt to travel through the out-of-core region. The temperature values appear to settle at higher
+temperatures to compensate for the inserted reactivity.
+
+\begin{figure}[t]
+  \centering
+%  \includegraphics[width=0.8\columnwidth]{u233-rod-worth-curve}
+  \caption{Change in power output following a step reactivity insertion of 24.8 pcm with the
+  \gls{MSRE} initially at 1 MW.}
+  \label{fig:insertion-power}
+\end{figure}
+
+\begin{figure}[t]
+  \centering
+%  \includegraphics[width=0.8\columnwidth]{u233-rod-worth-curve}
+  \caption{Change in maximum fuel salt and graphite temperatures, and the mean salt inlet and outlet
+  temperatures following a step reactivity insertion of 24.8 pcm with the \gls{MSRE} initially at 1
+  MW.}
+  \label{fig:insertion-temperature}
+\end{figure}
+
 \section{Summary} \label{sec:transient-summary}
+
+The hybrid $S_N$-diffusion method combines the strengths of the $S_N$ and neutron
+diffusion methods accurately model control rods while remaining tractable on modest computational
+resources. This chapter covers work demonstrating the hybrid $S_N$-diffusion method through two
+time-dependent reactivity-initiated simulations.The hybrid $S_N$-diffusion solver implemented in
+Moltres is coupled to preexisting \gls{DNP} and temperature solver capabilities to model
+transient behavior in reactors.
+
+The first time-dependent simulation is based on a zero-power \gls{MSRE} rod drop experiment. The
+rod drop is estimated to induce a 1500 pcm reactivity withdrawal in the \gls{MSRE}. Preliminary
+calculations determined the required initial rod height in the numerical model to replicate the
+same reactivity withdrawal effect and eliminate rod worth discrepancies. Following an initial
+$k$-eigenvalue simulation to obtain initial flux and \gls{DNP} distributions, the rod drop
+simulation ran with total fission neutron production rate measurements at each timestep. The
+integral neutron production, normalized to match the initial experimental neutron count rate,
+showed good agreement with the experimental integral neutron count curve following the rod drop.
+The numerical result has a maximum discrepancy of x \% at $t=16$ s. Experimental uncertainties in
+the actual $^{235}$U content, initial rod height, rod acceleration, and the initial neutron count
+rate may be responsible for some of the discrepancy. Snapshots of the 3-D neutron group flux
+distributions at various timesteps show expected localized flux decreases around the control rod
+as it moves through the core.
+
+The second time-dependent simulation is based on a reactivity insertion experiment on the
+$^{233}$U-fueled \gls{MSRE}. A control rod withdrawal induces a 24.8 pcm step reactivity
+insertion on the \gls{MSRE} initially running at 1 MW with steady salt flow. Moltres reproduces
+the general shape of the first power peak and the subsequent power oscillations. The power peak
+magnitude differs by y \% compared to \gls{MSRE} data. The fuel
+salt, graphite, inlet, and outlet temperature measurements show delayed responses to the increase
+in power output due to the time taken for the salt and graphite to heat up. Over the simulation
+time, the numerical \gls{MSRE} model settles into higher temperature distributions than the initial
+state to compensate for the reactivity insertion.
+
+Overall, this work has successfully demonstrated the hybrid method for coupled time-dependent
+reactor analyses of reactivity-initiated transients involving control rods.

### 2024-11-13 abstract.tex
Full draft + edits

\glspl{MSR} are advanced reactors noted for strong passive safety features.
They present unique challenges in multiphysics reactor modeling \& simulation arising from
strong temperature reactivity feedback, delayed neutron precursor flow, and turbulent heat transport
in fuel salt regions. Simulating the complex multiphysics interactions in
\glspl{MSR} requires robust, flexible, and highly scalable multiphysics software. Many \gls{MSR}
designs also still retain control rods which can complicate time-dependent reactivity-initiated
transient simulations on reactor software relying on neutron diffusion theory.
This work builds on existing capabilities in Moltres, a MOOSE-based \gls{MSR} simulation software,
to tackle these challenges and support efforts towards \gls{MSR} deployment.

This work verified and validated existing multiphysics coupling capabilities in Moltres in two
comparative studies. In both studies, Moltres showed good agreement with other \gls{MSR} simulation tools
involving coupled neutronics and thermal-hydraulics problems. This work also introduces a
turbulence model in Moltres to support future \gls{MSR} analyses involving turbulent delayed
neutron precursor and temperature transport. Lastly, this work introduces a novel hybrid
$S_N$-diffusion method for accurate control rod modeling in time-dependent \gls{MSR} simulations.
The hybrid method combines the strengths of both approaches by generating transport corrections
using the $S_N$ method near control rods. The $S_N$ and neutron diffusion solvers are coupled
through an adaptive boundary coupling algorithm. This algorithm allows the solver to adapt to the
transport correction parameters and preserve smooth neutron flux gradients across the interface.
In 1-D, 2-D, and 3-D $k$-eigenvalue simulations, the hybrid method produced accurate control
rod worth estimates, relative to reference neutron transport solutions and experimental data,
at approximately four times the cost of the neutron diffusion method.
Demonstrations of the hybrid method for time-dependent rod drop and reactivity insertion
simulations also perfomed well in reproducing expected trends observed in experimental data.
Analysis of its computational performance indicated the possibility of further optimizations
beyond its current implementation. With the hybrid method's spatial resolution and
efficient computational performance, Moltres could enable accurate and cost-effective simulations
of asymmetric transient in \glspl{MSR}.

### 2024-11-13 conclusion.tex
Full draft + edits

Five decades since the \gls{MSRE} \cite{haubenreich_experience_1970}, \glspl{MSR} are once again
drawing significant interest as an advanced reactor concept with enhanced safety, sustainability,
efficiency, and cost aspects over the
current generation of predominantly \glspl{LWR}. This observation is supported by the
nuclear start-ups developing \gls{MSR} designs and public funding from various governments around
the world in the race to reduce carbon emissions. \gls{MSR} simulation tools are essential in
supporting this wave of \gls{MSR} research and development and reaching the eventual \gls{MSR}
commercialization to meet our growing energy needs.
\glspl{MSR} feature physics interactions, such as \gls{DNP} flow, turbulent heat transport in fuel
salt regions, and strong temperature reactivity feedback, that may render some conventional reactor
software less suitable for \gls{MSR} analysis.

The overarching goal of this work was to enhance Moltres, a \gls{MOOSE}-based application, as a
reliable, intermediate-fidelity simulation tool for multiphysics \gls{MSR} analysis that scales
well from workstations to larger computing clusters. I successfully met this goal through two main
objectives:
%
\begin{enumerate}
  \item Verification and validation of multiphysics capabilities in Moltres
  \item Development of a novel hybrid $S_N$-diffusion method for improved control rod modeling in
    Moltres
\end{enumerate}

I conducted two \gls{VV} studies on Moltres to meet the first objective.
The first study follows the CNRS benchmark
study proposed and investigated by several European research groups \cite{tiberga_results_2020}.
The second study is a collaborative effort with Dr Aaron Reynolds (formerly of Oregon State
University) based on the \gls{MSRE} zero-power pump experiments \cite{prince_zero-power_1968}.
I also implemented and verified a Spalart-Allmaras turbulence model in Moltres to enable future
turbulent salt flow simulations. Chapter \ref{chap:verification} covers the \gls{VV} studies and
turbulence model implementation. Chapters \ref{chap:hybrid}, \ref{chap:msre}, and
\ref{chap:transient} cover the development, verification, and demonstration of the hybrid
$S_N$-diffusion method to meet the second objective.

Chapter \ref{chap:verification} provided a detailed description of Moltres and its existing
capabilities prior to this work. Moltres is a multiphysics simulation software for \glspl{MSR}. It
has also been adopted, by itself or coupled to other \gls{MOOSE} applications, for high-temperature
gas-cooled reactor and lead-cooled fast reactor analyses
\cite{fairhurst-agosta_multi-physics_2020, ji_numerical_2024}. Moltres features a modular code
architecture that enables the integration of its own multigroup neutron diffusion and \gls{DNP}
models with the Squirrel \gls{MOOSE}-based application \cite{noauthor_arfcsquirrel_nodate} and the
\gls{MOOSE} Navier-Stokes module \cite{peterson_overview_2018} for thermal-hydraulics modeling.
The \gls{MOOSE} framework provides user- and developer-friendly interfaces that enable fast
prototyping and implementation of experimental modeling features.
The underlying libMesh \cite{kirk_libmesh_2006} and PETSc \cite{satish_petsc_2019} libraries
provide state-of-the-art meshing and \gls{FEM} solver capabilities that are robust, efficient, and
highly scalable.

Chapter \ref{chap:verification} continued with the CNRS benchmark verification results from
Moltres. The CNRS benchmark \cite{tiberga_results_2020} is a numerical benchmark study
designed to assess multiphysics coupling in \gls{MSR} simulation tools for fast-spectrum
\gls{MSR} modeling. The benchmark starts with three steady-state single-physics problems, followed by four
steady-state multiphysics problems incrementally introducing various multiphysics coupling
interactions, and lastly a series of transient multiphysics problems for dynamic analysis. Moltres
showed good agreement with the four participating \gls{MSR} simulation tools published in the CNRS
benchmark publication. Most of the test metric measurements from Moltres fell within one standard
deviation of the benchmark average. The only discrepancies arose in temperature readings along the
top boundary of the 2-D model due to discontinuous and unphysical velocity boundary conditions
prescribed by the leaky lid-driven cavity flow profile. All spatial distributions of fission rate
density, delayed neutron source, temperature, and velocity were nearly indistinguishable compared
to the benchmark participants.

Chapter \ref{chap:verification} also covered the collaboration study with the developer of
QuasiMolto \cite{reynolds_analysis_2023} for a \gls{VV} study based on the \gls{MSRE} zero-power
pump start-up and coast-down experiments \cite{prince_zero-power_1968}. This study was designed to
assess the looped \gls{DNP} flow modeling capabilities in Moltres and QuasiMolto. QuasiMolto is a
circulating fuel reactor simulation software that employs a multi-level quasi-diffusion method
\cite{tamang_multilevel_2014} for 2-D axisymmetric multiphysics simulations of \glspl{MSR}.
We developed a 2-D axisymmetric model of the \gls{MSRE} for this study. Despite the widely
different numerical schemes employed by Moltres and QuasiMolto, both codes were highly consistent
with each other in $k$-eigenvalue neutronics simulations under static and steady salt flow
conditions prior to pump transient simulations. The $k_\text{eff}$ estimates from both codes fell
within 22 and 21 pcm of each other under static and steady flow conditions. Consequently, reactivity
loss estimates due to \gls{DNP} decay outside the core agreed within 1 pcm. Spatial flux and
\gls{DNP} distributions showed nearly perfect overlap with relative differences under 0.8 \%. Both
codes remained consistent with each other in the pump start-up and coast-down transient simulations
when comparing the reactivity loss over time induced by changes in \gls{DNP} flow. The numerical
results deviated from \gls{MSRE} experimental data due to the absence of upper and lower plena in
the numerical models and the omission of turbulent mixing. These affected the peak reactivity loss
measurements and the flattening of reactivity oscillations in the pump start-up experiment.
Nevertheless, the distinct reactivity oscillations serve as valuable reference points for
verifying \gls{DNP} advection modeling accuracy between Moltres and QuasiMolto, and for other
researchers looking to verify their \gls{MSR} simulation tools.

Chapter \ref{chap:verification} concluded with the implementation and verification of the
Spalart-Allmaras turbulence model in Moltres for enabling wall-bounded turbulent molten salt flow
modeling. The implementation includes a rotation correction scheme
\cite{aupoix_extensions_2003, dacles-mariani_numericalexperimental_1995} for improved accuracy
when modeling turbulent flow with curved streamlines. The model performed well under the turbulent
channel flow verification test \cite{moser_direct_1999}, the turbulent pipe flow validation test
\cite{laufer_structure_1954}, and the backward-facing step flow validation test
\cite{driver_features_1985}.

Chapter \ref{chap:hybrid} introduced the hybrid $S_N$-diffusion method for \gls{MSR} control rod
modeling in time-dependent reactor analyses. The hybrid $S_N$-diffusion method improves on the
standard neutron diffusion method by iteratively applying transport corrections generated from
solving the $S_N$ neutron transport method in subdomains containing highly neutron-absorbing
control rods.
The chapter starts with the derivation for the newly-implemented $S_N$ neutron transport method in
Moltres. It uses the \gls{SAAF} formulation of the $S_N$ equations, which are highly efficient and
scalable with HYPRE multigrid preconditioners \cite{hypre_hypre_2022} from PETSc. I formulated
two possible transport correction schemes for the hybrid method: a diffusion correction and a
drift correction scheme. These schemes fit into modified forms of the neutron diffusion equations
to correct flux errors in near control rods where diffusion theory is not valid. I also developed
an adaptive boundary coupling algorithm to couple the $S_N$ and neutron diffusion solvers. The
algorithm automatically truncates transport correction parameters near the boundaries of the
$S_N$ problem subdomain to discard inaccurate correction parameters and preserve smooth neutron
flux gradients across the interface. The chapter concludes with the numerical implemetation details
of the $S_N$ method with level-symmetric quadrature sets and modifications to the existing neutron
diffusion solver for the hybrid method. The $S_N$ and diffusion solvers are coupled through fixed
point iterations using the \gls{MOOSE} \texttt{MultiApp} and \texttt{Transfers} systems.

Chapter \ref{chap:msre} demonstrated the hybrid $S_N$-diffusion method in $k$-eigenvalue simulations
of 1-D, 2-D, and 3-D \gls{MSRE} models. The 1-D simulations involve six test cases with increasing
geometric complexity modeled after the air-filled control rod thimbles, salt-graphite lattice
structure, and vessel regions in the \gls{MSRE}. Simulations with the OpenMC Monte Carlo neutron
transport code and the $S_8$ method in Moltres provided reference solutions for the hybrid and
neutron diffusion methods to be assessed against. Some discrepancies arose from the eight neutron
energy group structure as evidenced by differences in $k_\text{eff}$ and flux distributions between
OpenMC simulations on continuous energy (OpenMC-CE) and multigroup (OpenMC-MG) modes. Otherwise, the
multigroup $S_8$ method showed good agreement with OpenMC-MG. While $k_\text{eff}$ estimates from
the hybrid methods with the diffusion and drift correction schemes deviated from the $S_8$ method
and OpenMC-MG by 100-400 pcm, the hybrid methods
accurately reproduced control rod worth estimates within 0.5 \% of them (2-3 \% relative to
OpenMC-CE). The neutron diffusion method fared significantly worse at 5.5 \% relative to $S_8$ and
OpenMC-MG, and 8 \% relative to OpenMC-CE. An analysis of the drift and diffusion correction
schemes showed the drift correction scheme to be superior due to singularities and discontinuities
in the diffusion correction parameter. Therefore, the drift correction scheme was chosen for
the remainder of this work. Another analysis showed little impact on control rod worth estimates
from varying the $S_N$-resolved subdomain size as long as the size is kept constant between
$k$-eigenvalue simulations used to calculate the rod worth. The hybrid $S_N$-diffusion method
exhibited a superlinear convergence rate with the number of fixed point iterations, leading to the
$k_\text{eff}$ error estimate falling below $10^{-7}$ after two iterations.

Chapter \ref{chap:msre} continued with 2-D quarter-core and full-core \gls{MSRE} simulations. The
\gls{MSRE} models were nearly identical to the horizontal cross section of the actual \gls{MSRE}
reactor. The hybrid method maintained its accurate control rod modeling capability in the 2-D
simulations involving the incrementally insertion of three control rods in the \gls{MSRE} model.
The hybrid method reported rod worth error magnitudes of less than 40 pcm relative to OpenMC-CE,
which were significant improvements over the neutron diffusion method error estimates ranging from
569 pcm to 1484 pcm. The hybrid method also showed significant improvements in the absolute mean
and maximum errors in fuel channel power distributions over the neutron diffusion method.

The 3-D simulations in Chapter \ref{chap:msre} doubled as validation for the 3-D \gls{MSRE} model
against reference \gls{MSRE} experimental data and the \gls{MSRE} numerical benchmark study
\cite{fratoni_molten_2020} in
the International Reactor Physics Experiment Evaluation Project (IRPhEP) handbook. OpenMC and
hybrid method estimates of $k_\text{eff}$ of the \gls{MSRE} when the experiment first achieved criticality
$k_\text{eff}$ estimates of the \gls{MSRE} at its initial critcality configuration from OpenMC, the
hybrid method, and the neutron diffusion method in this work showed good agreement with the Serpent
model from the \gls{MSRE} numerical benchmark. All numerical estimates exceeded the experimental
value by 1-2 \% due to possible biases and uncertainties in the nuclear data library for graphite
\cite{fratoni_molten_2020}. Temperature reactivity coefficient values from this work also showed
good agreement with \gls{MSRE} data, within experimental uncertainty, with percentage discrepancies
of about 3 \%. In the subsequent control rod worth study, OpenMC and the hybrid method showed nearly
perfect overlap in the integral rod worth curve throughout the entire length of rod travel. Due to
geometric approximations of the control rods in the numerical models, they overestimated the total
rod worth relative to \gls{MSRE} data by approximately 4-5 \%. The hybrid method significantly
outperforms the neutron diffusion method which overestimated the total worth by 21.1 \%.
When comparing solution times, the hybrid method took approximately four times as long as the
neutron diffusion method.
The 3-D hybrid method simulations exhibited nearly linear scaling in strong scaling tests, which
is promising for future projects involving larger reactor models. Some performance optimizations
may be possible to reduce data transfer times between the $S_N$ and diffusion solvers which took
up about 30 \% of the total solution time due to inter-processor communications.

Lastly, Chapter \ref{chap:transient} demonstrated the hybrid method in time-dependent simulations
based on a \gls{MSRE} zero-power rod drop experiment \cite{prince_zero-power_1968} and a \gls{MSRE}
reactivity insertion experiment during power operation \cite{steffy_experimental_1970}. These
simulations required coupling the hybrid method solver to \gls{DNP} and temperature solvers in
Moltres through nested iteration coupling structures. I set up the rod drop simulation through a
$k$-eigenvalue simulation with the control rod at its initial height before using the simulation
output as the initial condition for the time-dependent rod drop simulation. Moltres reproduced the
expected shape of the integral neutron count curve from \gls{MSRE} experimental data with a maximum
discrepancy of x \% at $t=16$ after the rod drop. Snapshots of the neutron flux spatial
distribution at various timesteps show expected localized flux depressions around the control rod
as it moves through the core.
The reactivity insertion simulation started with the \gls{MSRE} initially at 1 MW power output and
steady salt flow. Following a 24.8 pcm step reactivity insertion, Moltres reproduced the general
shape of the first power peak and the subsequent power oscillations. The power peak magnitude
is y \% higher than the peak in the experimental data. Temperatures in the \gls{MSRE} model rise
and settle at higher values to compensate for the reactivity insertion.

\section{Limitations and Future Work}

The second \gls{VV} study verified the looped \gls{DNP} flow modeling capability in Moltres under
steady-state and transient scenarios. In the original pump experiments, the control rod was driven
by a ``flux servo controller'' that adjusts the rod position in response to flux changes to maintain
criticality. Our study in this work approximated this action through $k$-eigenvalue solvers
coupled to time-dependent \gls{DNP} flow solvers. Potential future work would be to implement a
control system such as a PID (Proportional-Integral-Derivative) controller in combination with
the newly-implemented hybrid method to create a more representative model of the pump experiments.
This would enhance model validation and open up additional research directions which require a
control system.

This work implemented and verified the Spalart-Allmaras turbulence model in Moltres. The turbulence
model \gls{VV} tests indicated significant mesh refinement requirements near the wall. Mesh
refinement scales with the flow Reynolds number. With \gls{MSR} designs such as the \gls{MSFR}
reaching Reynolds numbers on the order of $10^6$, future work in this area should focus on
implementing wall functions. Wall functions eliminate the need for fine mesh near the wall by
approximating the log-law velocity profile near the wall. Verification and demonstration of the
turbulence model, beyond general \gls{VV} tests, in \gls{MSR} turbulent salt flow problems is also
crucial for the continued development of Moltres as a \gls{MSR} simulation tool.

Chapters \ref{chap:msre} and \ref{chap:transient} demonstrated the hybrid $S_N$-diffusion method's
accuracy in control rod worth calculations. However, some discrepancies still persist in the
$k_\text{eff}$ estimates of individual reactor states. These discrepancies affect the average
temperature of the reactor in multiphysics simulations to compensate for the mismatch in
reactivity. These discrepancies could be reduced through
neutron leakage correction at the vacuum boundaries. Remedies for this issue exist in a number of
forms such as group-wise or matrix albedo boundary conditions.

The large 3-D simulations in this work uncovered significant memory usage by the $S_N$ solver, even
with the distributed mesh feature to spread mesh and variable storage across the compute nodes.
Each full-core simulation required at least 40 nodes with 512 GB memory each to run.
This issue may be a stumbling block for using the hybrid method on smaller computing clusters that
have less memory per processor. A custom preconditioner and solver routine in Moltres could help to
reduce memory usage by referencing the same stored Jacobian C++ variable for neutron angular flux
variables in the same neutron energy group and on the same \gls{FEM} quadrature point. Their
Jacobian formulations are identical aside from the level-symmetric ordinate and weight. Another
area for optimization in the hybrid method is the data transfer times between the $S_N$ and
diffusion solvers. Due to mesh distribution, each 3-D simulation spent 30 \% of its solution time
on data transfers between processors. This could be minimized through optimizations in the mesh
distribution and the data transfer caching systems.

Finally, future work could demonstrate the hybrid method to its fullest potential through the
simulation of asymmetric reactivity-initiated transients. Asymmetry refers to significant changes
in the neutron flux shape during a transient scenario. The core benefit of the hybrid method is
the spatial resolution it provides to time-dependent simulations involving control rod movement.
This objective could be met through modeling most other \gls{MSR} designs whose control and shim
rods are not centrally located.

### 2024-11-13 hybrid.tex
Full draft + edits

@@
 hybrid method using the Python programming language for quick
-prototyping. Implementation details of the $S_N$, neutron diffusion, and hybrid methods on Python
+prototyping. Implementation details of the $S_N$, neutron diffusion, and hybrid methods in Python
 are available in Appendix \ref{chap:implementation}. Both correction schemes share the iteration
@@
       $x=10$ cm.}
-    \label{fig:3b-drift}
+    \label{fig:3b-drift-1}
 \end{figure}
@@
 to encompass the control rod region and its vicinity from $x=0$ to $10$ cm based on the geometry in
-Figure \ref{fig:3b-geometry}. Figure \ref{fig:3b-drift} shows the $\vec{D}_1$ and $\vec{D}_8$
+Figure \ref{fig:3b-geometry}. Figure \ref{fig:3b-drift-1} shows the $\vec{D}_1$ and $\vec{D}_8$
 distributions from $x=0$ to $15$ from the reference $S_8$ and hybrid simulations. The hybrid

### 2024-11-13 introduction.tex
Full draft + edits

\section{Motivation}

A liquid-fueled \gls{MSR} is an advanced nuclear reactor with fissile material
dissolved in a molten salt mixture. This fuel salt mixture doubles as the primary reactor coolant
for transferring heat from the reactor core to the primary heat exchangers. Figure \ref{fig:msr}
shows a schematic diagram of a thermal-spectrum, channel-type \gls{MSR}. In channel-type
\glspl{MSR}, the fuel salt flows through vertical channels in the reactor core alongside neutron
moderators (e.g., graphite, heavy water, or molten sodium hydroxide). 
%
\begin{figure}[htb!]
	\centering
	\includegraphics[width=.7\columnwidth]{msr}
	\caption{Schematic diagram of the \gls{MSR} concept. Retrieved from
	\cite{u.s._doe_nuclear_energy_research_advisory_committee_technology_2002}.}
	\label{fig:msr}
\end{figure}

The \gls{MSR} is one of six advanced reactor concepts selected for improved safety, sustainability,
efficiency, and cost over the current generation of predominantly \glspl{LWR} at the \gls{GIF}
\cite{u.s._doe_nuclear_energy_research_advisory_committee_technology_2002}.
Due to the high thermal expansion coefficient, \glspl{MSR} possess an inherently robust
safety feature in the strong negative fuel temperature coefficient of
reactivity \cite{elsheikh_safety_2013}. This reactivity coefficient limits the
maximum temperature that the reactor core would experience in an accident
scenario such as an unprotected reactivity insertion because the subsequent
rise in core temperatures induces a significant drop in reactivity,
quickly neutralizing the initial reactivity insertion. \glspl{MSR} also
operate at a large thermal margin to boiling and can rely on natural
circulation in the event of a pump failure. For an additional layer of safety, many \gls{MSR}
designs incorporate a drain plug made of actively-cooled frozen salt, which
melts when the core temperatures exceed safety thresholds. Hot molten salt
in the core would then flow into a drain tank designed to hold the fuel salt in
a subcritical configuration to disrupt any further chain fission reactions.

Some \glspl{MSR}, like the \gls{MSBR} or the \gls{MSFR}, can
incorporate the thorium fuel cycle for improved sustainability arising from the
use of abundant natural thorium resources and reduced transuranic waste
\cite{heuer_towards_2014}. The latter consequence also reduces costs
associated with long-term nuclear waste storage. Online fuel reprocessing raises the capacity
factor by reducing reactor downtime during reactor operation \cite{dolan_1_2017}.
Molten salt coolant loops can operate at near atmospheric pressures which eliminates the need for a
thick pressure vessel and drives down construction costs. We can make further economic arguments
supporting \glspl{MSR} in the context of the
carbon-constrained future envisioned in the \gls{IEA}'s \gls{NZE} roadmap
\cite{iea_net_2021}. The roadmap for minimizing carbon emissions requires solar photovoltaic- and
wind-dominated energy markets, which can results in highly variable and non-dispatchable
electricity generation. The resulting volatility in electricity prices
encourages the construction of heat storage and peak power
production plants. At the same time, demand for carbon-neutral
fuel will rise as electrification is economically unfeasible
for some industries, such as the aviation and marine sectors, which depend on
energy-dense fuels for propulsion.
The most cost-efficient options for heat storage, peak power production, and carbon-neutral fuel
production all require high-temperature heat \cite{forsberg_market_2020}.
This requirement favors \glspl{MSR}, which can deliver heat at higher average
temperatures than \glspl{LWR} and \glspl{SFR}.

Significant hurdles still stand in the way of commercial \gls{MSR} deployment \cite{dolan_27_2017}.
Major technological challenges include the need for further research on \gls{MSR} safety analysis,
irradiation and corrosion behavior of \gls{MSR} structural components, fission product tracking
and processing, \gls{MSR} nuclear safeguards measures, and the development of \gls{MSR} hydraulic
and heat exchanger components.

\subsection{Past \& Present \gls{MSR} Research \& Development}

\gls{ORNL} researchers first conceived the \gls{MSR} concept in pursuit of a high-temperature
liquid fuel reactor for the US Aircraft Nuclear Propulsion program in
the 1950s \cite{rosenthal_molten-salt_1970}. They
built the first ever operational \gls{MSR}, the 2.5 MW$_{\text{th}}$
\gls{ARE} reactor at \gls{ORNL}. The successful demonstration of the \gls{ARE} spurred further
research into adapting \glspl{MSR} for civilian power generation \cite{rosenthal_molten-salt_1970}.
Continued \gls{MSR} research efforts culminated in designing constructing, and successfully
operating of the 8-MW$_{\text{th}}$, thermal-spectrum \gls{MSRE} with
graphite channels and a LiF-BeF$_2$-ZrF$_4$-UF$_4$ fuel salt mixture
\cite{haubenreich_experience_1970}. In addition to other operational achievements, the
\gls{MSRE} became the first reactor to run on $^{233}$U fuel bred from $^{232}$Th. Building on
their experience with the \gls{MSRE}, \gls{ORNL} proposed a new program for constructing and
operating a demonstration reactor based on the \gls{MSBR} concept, which they had also developed
\cite{macpherson_molten_1985}. The \gls{MSBR} is a thermal-spectrum \gls{MSR} with fertile
$^{232}$Th isotopes mixed directly into the \gls{FLiBe} molten salt for $^{233}$U breeding. The
\gls{MSBR} was to rely on continuous online reprocessing to add fertile
material and remove fission product neutron poisons.
However, the \gls{MSBR} project was canceled prior to the demonstration stage in
favor of the \gls{LMFBR}, which had benefited from a head start in development and more substantial
political backing \cite{macpherson_molten_1985}.

Following a relative lull lasting until the late 1990s, renewed research efforts and interest from
the \gls{GIF}
provided new impetus for \gls{MSR} research and development. As of the end of 2022, numerous
\gls{MSR} designs exist at various stages of development. Leading \gls{MSR} designs, in terms of
development, licensing, or demonstration, include the \gls{MSFR} \cite{merle_optimized_2007},
\gls{MCFR} \cite{terrapower_terrapower_2021}, TMSR-LF1 \cite{zhang_review_2018}, and \gls{IMSR}
\cite{leblanc_18_2017}. The \gls{MSFR} is a fast-spectrum breeder reactor developed through
collaboration among European institutes with funding support from the
European Union. Figure \ref{fig:msfr} shows a schematic diagram of the \gls{MSFR}. As opposed to
the multi-channel design of the \gls{MSRE} and \gls{MSBR}, the \gls{MSFR} is a pool-type design and
the reactor core consists of a large molten salt pool without graphite moderators to avoid
frequent graphite replacements and positive graphite temperature reactivity feedback. The
\gls{MCFR} is a similar pool-type reactor under active development at TerraPower. TerraPower and
Southern Company embarked on a joint project to design, construct, and operate a prototype
\gls{MCRE} design with funding support from the \gls{DOE}'s \gls{ARDP}. \gls{CAS} launched the
\gls{TMSR} program in 2011 to develop and construct both solid-fueled and liquid-fueled \gls{TMSR}
designs \cite{zou_research_2019}. They finished construction of TMSR-LF1, a 2-MW$_{\text{th}}$
liquid-fueled prototype, in August 2021 and received approval for reactor commissioning in August
2022. Lastly, Canada-based Terrestrial Energy is also developing its \gls{IMSR}, a small modular
\gls{MSR} based on the \gls{MSRE}. It passed a joint technical review by the
Canadian and US nuclear regulators in July 2022.

Developing \gls{MSR} simulation software plays a vital role in supporting \gls{MSR}
development. Accurate reactor modeling capabilities are vital because they
accelerate reactor design and optimization by enabling quicker iteration through numerous design
changes. \gls{MSR} simulation software are also essential tools in reactor safety analysis and
licensing efforts since reactor developers must demonstrate and verify that the \gls{MSR} systems
perform as designed and remain safe under various accident scenarios. 
%
\begin{figure}[htb!]
	\centering
	\includegraphics[width=.7\columnwidth]{msfr}
	\caption{Schematic diagram of the \gls{MSFR}. Retrieved from 
	\cite{allibert_7_2016}.}
	\label{fig:msfr}
\end{figure}

\subsection{Challenges in Multiphysics \gls{MSR} Modeling \& Simulation}

While modeling \glspl{MSR} is not necessarily more difficult than modeling
solid-fueled reactors, we must adapt our software tools to accurately model the
unique phenomena found in these circulating-fuel reactors. The differences in
the challenges of simulating \glspl{MSR} compared to solid-fueled reactors stem
mainly from the liquid fuel form of the fuel salt \cite{diamond_phenomena_2018,
huff_identifying_2019}.

Liquids generally exhibit greater thermal expansion per unit increase in temperature than solids.
The decreased fuel salt density due to thermal expansion
increases the likelihood of neutrons escaping the fuel region
and being absorbed by non-fissile material elsewhere in the reactor.
Consequently, combined with the temperature-dependent Doppler broadening of
resonance capture cross sections, \glspl{MSR} possess stronger negative fuel
temperature reactivity feedback than their solid-fueled counterparts
\cite{elsheikh_safety_2013}. These
phenomena ultimately result in strong interactions between the neutron fluxes
and the fuel salt temperatures, given that neutron fluxes affect fuel salt temperatures
through fission heat generation and fuel salt temperatures affect neutron
fluxes through the mechanisms described previously.

With the fuel salt also providing cooling in the core through advective heat
transfer, velocity flow
profiles in the fuel salts strongly impact the temperature distribution via
advection-dominated heat transfer \cite{diamond_phenomena_2018}. The flow-driven
temperature distribution contrasts
with the relatively static temperature profiles in fuel pins and
other solid fuel types, which are physically separate from the coolant.

\Glspl{DNP} flow freely within the primary coolant loop instead of
being held in place as in solid-fueled reactors. Thus, the delayed neutron
source distribution varies significantly depending on the flow profile and
velocity of the salt. In addition, the reactor loses some delayed neutrons from out-of-core
\gls{DNP} decay. These delayed neutrons are considered lost as they are emitted
in subcritical regions and are unlikely to contribute to further fission
reactions in the active core. The reduced delayed neutron fraction in the core
contributes to a greater prompt power spike following a reactivity insertion
event than solid-fueled reactors, absent any temperature reactivity
feedback.

Molten-salt flow along various parts of the coolant loop may fall within the turbulent flow
regime, characterized by chaotic eddies, vortices, and other flow instabilities.
Turbulent flow effects further complicate multiphysics interactions of flow with the temperature
and \gls{DNP} distribution. Turbulent flow effects contribute significantly to advection-dominated
heat and particle transfer in molten salt systems, thereby causing enhanced mixing. Therefore,
multiphysics software for \gls{MSR} analysis require adequate flow modeling capabilities with
support for \gls{DNP} drift and some form of turbulence modeling.

\gls{MSR} simulation tools require transient modeling capabilities to simulate transient
accident scenarios. Furthermore, several transient scenarios involve control rods, either as active
safety mechanisms or as accident initiators through unintended control rod ejection. Therefore,
\gls{MSR} simulation tools also require accurate control rod modeling capabilities to run
numerical analyses of accident scenarios involving control rods.

\subsection{Moltres for Multiphysics \gls{MSR} Analysis}

Moltres \cite{lindsay_moltres_2017} is an open-source multiphysics reactor simulation software
explicitly developed considering \gls{MSR} characteristics in mind. Moltres is
built on the \gls{MOOSE} \cite{giudicelli_30_2024} open-source finite-element framework,
which facilitates multiphysics coupling between different
\gls{MOOSE}-based and \gls{MOOSE}-wrapped applications. The \gls{MOOSE} framework also provides
Moltres with advanced meshing and numerical solver capabilities through interfacing with libMesh
\cite{kirk_libmesh_2006} and PETSc \cite{satish_petsc_2019} open-source libraries. Therefore,
Moltres supports up to 3-D unstructured meshes, scales well on high-performance computing systems,
and provides a flexible multiphysics coupling system, which can be tailored for different types of
simulations.

Moltres models coupled neutronics and thermal-hydraulics in reactors. While
generally applicable to most reactor concepts, much of
Moltres' development focuses on meeting the needs of \gls{MSR} multiphysics simulations.
Together with \gls{MOOSE}'s \texttt{Heat}
\texttt{Conduction} and \texttt{Navier-Stokes} \cite{peterson_overview_2018}
modules, Moltres solves the multigroup neutron diffusion
equations for an arbitrary number of energy and precursor groups and
thermal-hydraulics equations simultaneously on the same mesh (or separately solved and coupled
through fixed-point iterations if desired).

Lindsay et al. \cite{lindsay_introduction_2018}
demonstrated Moltres' multiphysics \gls{MSR} modeling capabilities with 1-D salt
flow modeling in axisymmetric and 3-D models of the \gls{MSRE}. The neutron flux and
temperature distributions agreed qualitatively with legacy
\gls{MSRE} data albeit with some minor quantitative discrepancies due to
simplifications and assumptions in the reactor geometry. I demonstrated Moltres' capabilities for
1) looping of \gls{DNP} drift back into the reactor core, 2) coupling the \gls{DNP}
drift to numerically calculated salt flow profiles within the reactor core,
and 3) a decay heat model to simulate decay heat from fission products, with a 2-D axisymmetric
design of the \gls{MSFR} in my Master's thesis \cite{park_advancement_2020}.

While the Moltres
model of the \gls{MSFR} showed good agreement with other studies in most steady-state and transient
simulation cases, the Moltres model showed significant discrepancies during pump-initiated
transient scenarios without a proper turbulence model. Instead, the model applied a
uniform eddy viscosity assumption, which proved inadequate under non-steady flow. In order to
advance Moltres as a multiphysics simulation software for \gls{MSR} analysis, Moltres requires a
turbulence modeling capability to capture adequate turbulent flow phenomena and their interactions
with other physics present in \glspl{MSR}. Of the various classes of turbulence models, a
\gls{RANS}-based model is most suitable for coupling with Moltres when considering computational
efficiency. The \gls{RANS}-based equations are time-averaged fluid flow equations
derived by separating the flow variable into its time-averaged and fluctuating components.
\gls{RANS}-based models are also well-validated, representing the most widely used turbulence
models for engineering applications.

In addition to new features, Moltres would benefit from further \gls{VV} of its existing
multiphysics capabilities to identify software bugs and ensure reliable results. Accurate modeling
and simulation of salt flow-induced \gls{DNP} drift and the strong coupling between neutronics and
thermal-hydraulics are crucial for \gls{MSR} simulation tools.
% Therefore, the proposed work will include relevant \gls{VV} studies for Moltres.

\subsection{Control Rod Modeling in \glspl{MSR}}

While some \gls{MSR} concepts completely eliminated control rods from their signs, many still
retain them. Control rods consist of strong
neutron absorbers like boron, cadmium, or gadolinium to absorb and thus reduce the number of
neutrons available to contribute to further fission chain reactions. All reactors start operation
with excess reactivity to ensure enough fissile material is available to maintain power operation
and last through to the next scheduled refueling. Control rods and burnable poisons cancel out the
excess reactivity. Control rods are also helpful
for ramping the power up or down during start-up, shut-down, or load-following operations. Most
importantly, control rods provide a safety mechanism for quickly shutting down a reactor under
dangerous conditions. Rapid control rod insertion is the primary mechanism of a reactor scram to
reliably introduce a large negative reactivity insertion in response to an unintended reactivity
insertion or other events that may threaten the reactor's safe operation.

Amongst other tasks in reactor modeling, reactor physics analysts aim to accurately reproduce
control rod worth, which is measured as the relative reduction
in the neutron multiplication factor $k_{eff}$ due to the insertion of the control rod in the
reactor. They are also interested in the accompanying change in flux shape near the
control rod. Neutrons entering the control rods are more likely to be absorbed than
adjacent regions in the reactor core. Therefore, control rods induce highly anisotropic neutron
angular fluxes and sharp gradients in the neutron flux in their vicinity. Control rods also cause
shifts in the neutron energy spectrum because their absorption cross sections are much higher in
the lower neutron energy range; more energetic neutrons generally have a higher probability of
escaping the control rod region. As a consequence, modeling control rods accurately requires
high-fidelity computational methods to capture the angular dependence in the neutron
flux near the highly absorbing medium. Existing reactor analysis workflows often rely on Monte
Carlo methods for high-fidelity neutron transport calculations.

However, traditional neutron transport methods, including Monte Carlo methods, are computationally expensive
and thus are mainly used for time-independent neutronic analyses. For solid-fueled reactors,
static control rod analyses can be justified due to the relatively static temperature distribution
shape in the fuel and disparate timescales of heat transfer from the fuel to
the coolant (seconds) and neutron flux changes (microseconds to milliseconds). In
\glspl{MSR}, this distinction blurs due to molten salt serving the dual role of fuel and
coolant and the strong fuel temperature reactivity feedback. For time-dependent multiphysics
simulations coupling neutronics to \gls{TH} and other physics present in nuclear reactors, most
reactor analyses rely on neutron diffusion theory for modeling neutronics. The issue with applying
neutron diffusion theory near control rods arises because neutron diffusion theory approximates
neutron transport theory by integrating over the angular variables. As a result, diffusion-based
methods perform poorly in reactor systems with control rods.

% To tackle the issue of modeling control rods in \glspl{MSR} with the multigroup neutron diffusion
% solver in Moltres, I will develop and implement a novel hybrid method for incorporating neutron
% transport-derived corrections to the neutron diffusion method while largely preserving the
% computational efficiency of the neutron diffusion method over neutron transport methods. This will
% allow for computationally feasible transient analysis of \glspl{MSR} that incorporate thermal and
% neutronic feedback on variable timescales.

\section{Research Objectives}

The overarching goal of this work is to improve on Moltres as a reliable, intermediate-fidelity
simulation tool for multiphysics \gls{MSR} analysis that can be run on workstations as well as on
larger computing clusters. Thus, new and existing capabilities in Moltres must
be accurate within accepted bounds, rigorously verified, computationally efficient, and highly
scalable. These conditions form the underlying development principles of this work.

The scope of this dissertation can be divided into two main objectives.
%
%\begin{enumerate}[itemindent=20pt, listparindent=1.5em, label=\textbf{\arabic*}]
%  \item \textbf{Verification and validation of multiphysics capabilities in
%    Moltres}

    The first objective is to verify and validate multiphysics capabilities in Moltres for
    modeling and simulating
    the strong interactions between the neutronics and thermal-hydraulics in \glspl{MSR}. This
    will be accomplished through two separate studies: verifying Moltres against the CNRS
    benchmark \cite{tiberga_results_2020} for fast-spectrum \gls{MSR} analysis, and modeling the
    \gls{MSRE} zero-power pump start-up and coast-down experiments. I will perform the second study
    in collaboration with the developer of QuasiMolto \cite{reynolds_analysis_2023} as a \gls{VV}
    study.
%    with the additional aim of creating a reproducible numerical benchmark of the listed
%    \gls{MSRE} experiments.
%
%  \item \textbf{Implementation and verification of a \gls{RANS}-based turbulence model in Moltres}
    This objective includes implementing and verifying a Spalart-Allmaras turbulence model in
    Moltres to support future turbulent flow
    simulations for \gls{MSR} analyses. This will address the absence of turbulence modeling
    capability in Moltres and the consequent restriction on Moltres for modeling \gls{MSR} systems
    with complex turbulent flow.

%  \item \textbf{Development of a novel hybrid $S_N$-diffusion method for improved control rod modeling in Moltres}

    The second objective is to
    develop a novel hybrid $S_N$-diffusion method for accurate control rod modeling in
    time-dependent \gls{MSR} analyses. The hybrid method uses the discrete ordinates ($S_N$)
    neutron transport method to generate transport corrections in regions near control rods for
    drift correction terms in modified neutron diffusion equations. The $S_N$ method is applied on
    a reduced problem domain to ensure the hybrid method remains tractable on small to moderate
    computing clusters.
%\end{enumerate}

\section{Outline}

Chapter 2 presents a literature review of existing multiphysics \gls{MSR} simulation software,
\gls{VV} studies on \gls{MSR} modeling and simulation, turbulence modeling in \gls{MSR} systems,
and transport-correction techniques for neutron diffusion methods.
Chapter 3 provides an in-depth description of Moltres and its existing capabilities in the context
of previously published work. This chapter then presents Moltres \gls{VV} results for the CNRS
benchmark and the numerical \gls{MSRE} zero-power pump experiment studies. Following these studies,
the chapter presents the implementation and verification of the Spalart-Allmaras turbulence model
in Moltres.
Chapter 4 presents the theory and numerical implementation of the hybrid
$S_N$-diffusion method in Moltres. These include the $S_N$ method implementation, the transport
correction formulations, the iteration algorithm, the $S_N$-diffusion coupling implementation, and
general implementation details relating to the underlying numerical solver in Moltres.
Chapter 5 presents the verification and demonstration of the hybrid $S_N$-diffusion method for
$k_\text{eff}$ and rod worth calculations through
$k$-eigenvalue simulations of 1-D, 2-D, and 3-D \gls{MSRE} models. The hybrid method is verified
against reference results from OpenMC Monte Carlo neutron transport code and \gls{MSRE}
experimental data.
Chapter 6 presents the demonstration of the hybrid $S_N$-diffusion method in time-dependent
reactivity-initiated simulations modeling the \gls{MSRE} rod drop and reactivity insertion
experiments.
Chapter 7 concludes this dissertation by summarizing the results presented, identifying limitations
in this work, and providing potential research directions to address those limitations or extend
this work.

### 2024-11-13 main.tex
saaf sn method
add verification chapter for referencing, hybrid chapter edits
Fix typos and add numerical implementation section
final edits for first draft of chapter 4; add image files for chapter 2,3,4
MSRE chapter edits
msre chapter edits
chapter 5 draft
turbulence verification section
Chapter 5 summary
Chapter 6 content
Full draft + edits

@@
 %% Package and Class "uiucthesis2021" for use with LaTeX2e.
-\documentclass[edeposit,fullpage,11pt]{uiucthesis2021}
+\documentclass[11pt]{uiucthesis2021}
 \doublespacing
@@
 \usepackage{enumitem}
+\setlist[itemize]{noitemsep}
 \usepackage{diagbox}
@@
 \usepackage{datatool}
-\usepackage[numbers]{natbib}
+% \usepackage[numbers]{natbib}
 \usepackage{notoccite}
+
+\usepackage[utf8]{inputenc}
+\usepackage[english]{babel}
+\usepackage[bibstyle=ieee,backend=bibtex8,sorting=none,eprint=false,isbn=false,doi=false]{biblatex}
+\usepackage{textalpha}
+\addbibresource{./bibliography.bib}
+\AtEveryBibitem{%
+  \clearfield{note}%
+}
 
@@
 
-\title{A Hybrid Neutronics Method for Time-Dependent Control Rod Modeling in Molten Salt Reactors}
+%\title{A Hybrid Neutronics Method for Time-Dependent Control Rod Modeling in Molten Salt Reactors}
+\title{Advancements in Moltres for Time-Dependent Multiphysics Molten Salt Reactor Modeling}
 \author{Sun Myung Park}
 \department{Nuclear, Plasma \& Radiological Engineering}
-\prelim
 \degreeyear{2024}
-\committee{Dr Madicken Munk, Chair \\
+\committee{Associate Professor Kathryn D. Huff, Chair \& Co-Director of Research \\
+           Assistant Professor Madicken Munk, Co-Director of Research, Oregon State University \\
            Professor Rizwan Uddin \\
-           Associate Professor Tomasz Kozlowski \\
+           Professor Tomasz Kozlowski \\
            Professor Paul Fischer}
+
+\newcounter{counterforappendices}
 
@@
 
-%\input{abstract}
+\input{abstract}
 
@@
 
-%\begin{acknowledgments}
-%
-%%\input{acks}
-%
-%\end{acknowledgments}
-%\addcontentsline{toc}{chapter}{Acknowledgments}
+\begin{acknowledgments}
+
+%\input{acks}
+
+\end{acknowledgments}
+\addcontentsline{toc}{chapter}{Acknowledgments}
 
@@
 \label{chap:intro}
-%\input{introduction}
-%\glsresetall
+\input{introduction}
+\glsresetall
 
@@
 \glsresetall
-%
-%\chapter{Description of Moltres and Assessment of its Current Capabilities}
-%\label{chap:moltres}
-%\input{moltres}
-%\glsresetall
-%
-%\chapter{Verification of Moltres for Multiphysics Molten Salt
-%Reactor Modeling}
-%\label{chap:benchmark}
-%\input{benchmark}
-%\glsresetall
-%
-%\chapter{Hybrid S$_\text{N}$-Diffusion Method for Improved Control Rod Modeling
-%in Neutron Diffusion Methods}
-%\label{chap:hybrid}
-%\input{hybrid}
-%\glsresetall
-%
-%\chapter{Proposed Work}
-%\label{chap:proposedwork}
-%\input{proposed-work}
-%\glsresetall
+
+\chapter{Verification of Multiphysics Features in Moltres}
+\label{chap:verification}
+\input{verification}
+\glsresetall
+
+\chapter{Hybrid $S_N$-Diffusion Method}
+\label{chap:hybrid}
+\input{hybrid}
+\glsresetall
+
+\chapter{Demonstration of the Hybrid Method with the MSRE}
+\label{chap:msre}
+\input{msre}
+\glsresetall
+
+\chapter{Demonstration of Time-Dependent Reactivity-Initiated Transients}
+\label{chap:transient}
+\input{transient}
+\glsresetall
+
+\chapter{Conclusion}
+\label{chap:conclusion}
+\input{conclusion}
+\glsresetall
 
@@
 
-\bibliographystyle{IEEEtran}
-\bibliography{bibliography}
+{
+\setstretch{1}
+% \bibliographystyle{IEEEtran}
+\printbibliography[heading=bibintoc,title={References}]
+}
+
+\clearpage
+\setcounter{counterforappendices}{\value{page}}
+\mainmatter
+\setcounter{page}{\value{counterforappendices}}
+\appendix
+\input{appendix}
 

### 2024-11-13 msre.tex
Full draft + edits

@@
 highly neutron-absorbing regions. I developed and implemented the diffusion and drift correction
-schemes for the hybrid method on Python and Moltres, respectively. This chapter demonstrates
+schemes for the hybrid method in Python and using Moltres, respectively. This chapter demonstrates
 the hybrid method through 1-D, 2-D, and 3-D $k$-eigenvalue simulations of the \gls{MSRE}.
@@
 %
-\begin{itemize}
-  \item $\Sigma_{t,g}$: Macroscopic total cross section in group $g$,
-  \item $\Sigma_{r,g}$: Macroscopic removal cross section in group $g$,
-  \item $\Sigma_s^{g'\rightarrow g}$: Macroscopic group-to-group scattering cross section matrix,
-  \item $\Sigma_{s,l}^{g'\rightarrow g}$: $l$-th Legendre moment of the macroscopic
-    group-to-group scattering cross section matrix,
-  \item $\Sigma_{sp,l}^{g'\rightarrow g}$: $l$-th Legendre moment of the macroscopic
-    group-to-group scattering production cross section matrix,
-  \item $D_g$: $P_1$-based diffusion coefficient in group $g$,
-  \item $\nu\Sigma_{f,g}$: Product of the average number of neutrons produced per fission and the
-    macroscopic fission cross section in group $g$,
-  \item $\chi_g$: Neutron fission spectrum in group $g$.
-\end{itemize}
+\begin{align*}
+  \Sigma_{t,g}:& \mbox{ Macroscopic total cross section in group $g$,}
+  \Sigma_{r,g}:& \mbox{ Macroscopic removal cross section in group $g$,}
+  \Sigma_s^{g'\rightarrow g}:& \mbox{ Macroscopic group-to-group scattering cross section matrix,}
+  \Sigma_{s,l}^{g'\rightarrow g}:& \mbox{ $l$-th Legendre moment of the macroscopic
+  group-to-group scattering cross section matrix,}
+  \Sigma_{sp,l}^{g'\rightarrow g}:& \mbox{ $l$-th Legendre moment of the macroscopic
+  group-to-group scattering production cross section matrix,}
+    D_g:& \mbox{ $P_1$-based diffusion coefficient in group $g$,}
+  \nu\Sigma_{f,g}:& \mbox{ Product of the average number of neutrons produced per fission and the
+  macroscopic fission cross section in group $g$,}
+    \chi_g:& \mbox{ Neutron fission spectrum in group $g$.}
+\end{align*}
 %
@@
 Moltres follows the methodology for solving the \gls{SAAF} formulation of the $S_N$ equations as
-described in Section \ref{sec:saaf} while the implementation on Python follows the methodology
+described in Section \ref{sec:saaf} while the implementation in Python follows the methodology
 for solving the original first-order formulation as described in Section \ref{sec:python-sn}.
@@
 The implementations on Moltres discretize the problem domain in space using the finite element
-method while the implementations on Python use finite difference or diamond difference schemes. The
+method while the implementations in Python use finite difference or diamond difference schemes. The
 most crucial difference lies in the transport correction formulations for the hybrid method. The
@@
 \section{1-D Neutronics Models \& Results} \label{sec:1d-results}
+
+This section presents the models, results, and discussion for the 1-D test cases.
 
@@
 
+This section presents the models, results, and discussion of the 2-D cases.
+
 \subsection{2-D Neutronics Model Setup} \label{sec:2d-model-setup}
@@
 \begin{itemize}
-  \item Ignored thermal expansion of all reactor components.
+  \item Neglected thermal expansion of all reactor components.
   \item Modeled the control rods as perfectly annular cylinders instead of beaded control elements
-    strung on flexible steel cables. Ignored the inconel shells encasing the control elements and
+    strung on flexible steel cables. Neglected the inconel shells encasing the control elements and
     the steel helical cables linking the control elements together.
@@
 \subsection{2-D Neutronics Numerical Results \& Discussion} \label{sec:2d-nts-results}
+
+This subsection presents the results and discussion of the 2-D quarter-core and full-core models.
 
@@
 
+This section presents the models, results, and discussion of the 3-D \gls{MSRE} simulations.
+
 \subsection{3-D Neutronics Model Setup} \label{sec:3d-model-setup}
@@
   \caption{List of reactor dimensions.}
-  \begin{tabular}{l l}
-    \toprule
-    Height [cm] & 215 \\
+  \begin{tabular}{l S}
+    \toprule
+    Reactor Specifications & {Value} \\
     \midrule
-    TODO: Populate table & \\
+    Height [m] & 2.15 \\
+    Radius [m] & 0.762 \\
+    Vessel thickness [m] & 0.0254 \\
+    Fuel channel length [m] & 0.03048 \\
+    Fuel channel width [m] & 0.01016 \\
+    Graphite lattice pitch [m] & 0.05080 \\
     \bottomrule
@@
      & {$k_\text{eff}$} \\
-     \cmidrule(l){2-2}
+     \midrule
     \gls{MSRE} experimental data & 1.00000(420) \\
@@
 greatly reduces memory requirements compared to storing the full mesh for each \gls{MPI} rank.
-
 Each 3-D \gls{MSRE} simulation required at least forty nodes for sufficient memory to run. This
-sets a strict minimum requirement for the 3-D hybrid simulations. On forty nodes, the simulations
+sets a strict minimum requirement for the 3-D hybrid simulations. A significant fraction of
+memory use in Moltres comes from storing the residual and Jacobian quantities at each \gls{FEM}
+quadrature point. Given that terms in the $S_N$ equations share the same Jacobian formulation
+across all ordinate directions, memory use (and compute time) could be optimized by reducing
+duplicate Jacobian evaluations.
+
+On forty nodes, the simulations
 required approximately 2.2 h to complete for a total of 88 node-hours or 2816 core-hours. The
@@
 approximately 0.675 h or 30 \% of the total simulation time. Future research into optimizing mesh
-distribution and transfer-related data caching may help to minimize data transfers and speed-up
+distribution and transfer-related data caching may help to minimize data transfers and speed up
 hybrid method simulations. The solution time for the initial neutron diffusion calculation with zero
@@
 benchmark. All numerical $k_\text{eff}$ estimates, including that from the benchmark report
-\cite{fratoni_molten_2020}, exceeded the \gls{MSRE} experimental value by 1-2 \% due to possible
-biases and uncertainties in the neutronic properties of graphite. Temperature reactivity
+\cite{fratoni_molten_2020}, exceeded the \gls{MSRE} experimental value by 1-2 \% possibly due to
+biases and uncertainties in the nuclear data library for graphite. Temperature reactivity
 coefficient values from OpenMC, the hybrid method, and the neutron diffusion method also show good

### 2024-11-13 transient.tex
Full draft + edits

@@
 
-This work aims to replicate the integral count curve following a rod
-drop of Rod 1 with the $^{235}$U loading corresponding to "30 capsule additions" of highly-enriched
-$^{235}$U capsules. Each capsule contains 85 g of $^{235}$U, of which 94.94 \% enters the core due
-to residual fuel salt remaining in the drain tanks after mixing \cite{fratoni_molten_2020}.
+This work aims to replicate the integral count curve following a rod drop of Rod 1 with the
+$^{235}$U loading corresponding to ``30 capsule additions'' of highly-enriched $^{235}$U capsules
+into the ``pump bowl''. Each capsule contains 85 g of $^{235}$U, of which 94.94 \% enters the core
+due to residual fuel salt remaining in the drain tanks after mixing \cite{fratoni_molten_2020}.
 
@@
   \int^{1.00072}_1 \frac{1}{k}\ dk &= 0.223 \int^{65.25+\Delta m}_{65.25} \frac{1}{m}\ dm \nonumber \\
-  \ln 1.00072 &= 0.223 \ln\left(\frac{65.25+\Delta m}{65.25}\right) \nonumber \\
+  \ln \left(1.00072\right) &= 0.223 \ln\left(\frac{65.25+\Delta m}{65.25}\right) \nonumber \\
   \Delta m &= 0.2109 \mbox{ kg}.
@@
 This work aims to replicate the power response following a reactivity insertion of 24.8 pcm of the
-\gls{MSRE} initially at 1 MW. Figure \ref{msre-reactivity-insertion} shows the theoretical and
+\gls{MSRE} initially at 1 MW. Figure \ref{fig:msre-reactivity-insertion} shows the theoretical and
 experimental power response following the reactivity insertion. Three separate publications
@@
 For the reactivity insertion simulation, we must first identify the $^{233}$U salt composition used
-in the experiment. The \gls{MSRE} $^{233}$ power operations occurred during Runs 16 to 18 according
+in the experiment. Power operations with the $^{233}$U-fueled \gls{MSRE} core
+occurred during Runs 16 to 18 according
 to \gls{MSRE} reports \cite{thoma_chemical_1971, burke_analysis_2019}. Table \ref{table:u233-salt}
-lists the proportion of elements in the \gls{MSRE} fuel salt during Runs 16 to 18. Noted that the
+lists the proportion of elements in the \gls{MSRE} fuel salt during Runs 16 to 18. Note that the
 components do not sum up to 100 \% due to experimental uncertainty in chemical analysis and the
@@
 salt, graphite, inlet, and outlet temperature measurements show delayed responses to the increase
-in power output due to the time taken for the salt and graphite to heat up. Over the simulation
-time, the numerical \gls{MSRE} model settles into higher temperature distributions than the initial
+in power output due to the time taken for the salt and graphite to heat up. The temperature values
+eventually settle higher than the initial
 state to compensate for the reactivity insertion.

### 2024-11-13 turbulence.tex
Full draft + edits

@@
+\section{Moltres Spalart-Allmaras Turbulence Model Verification} \label{sec:turbulence}
+
 As detailed in Section \ref{sec:th}, Moltres compiles with the MOOSE Navier-Stokes
@@
 
-To address the lack of turbulence modeling in Moltres, I implemented a \gls{SA} turbulence model
+To address the lack of turbulence modeling in Moltres, I implemented a Spalart-Allmaras turbulence model
 \cite{spalart_one-equation_1994}, described in Section \ref{sec:lit-turb}, with \gls{SUPG}
-stabilization on Moltres. The \gls{SA} model estimates the turbulent eddy viscosity defined by the
+stabilization on Moltres. The Spalart-Allmaras model estimates the turbulent eddy viscosity defined by the
 eddy viscosity hypothesis applied to the \gls{RANS} equations \cite{rodi_turbulence_2017}.
-On balance the \gls{SA} model is a complete (does not require prior
+On balance the Spalart-Allmaras model is a complete (does not require prior
 knowledge of the actual turbulence behavior) and computationally efficient one-equation turbulence
-model for approximating wall-bounded turbulent flows. The \gls{SA} model implementation in Moltres
+model for approximating wall-bounded turbulent flows. The Spalart-Allmaras model implementation in Moltres
 couples seamlessly with the continuous \gls{FEM} \gls{INSAD}
 \cite{peterson_overview_2018, lindsay_automatic_2021} model
-from the Navier-Stokes module. Alongside the \gls{SA} model, Moltres also now has turbulent
+from the Navier-Stokes module. Alongside the Spalart-Allmaras model, Moltres also now has turbulent
 diffusion physics kernels for temperature and the delayed neutron precursors.
 
-The SA model in Moltres follows the \gls{SA} model with a rotation correction scheme
+The SA model in Moltres follows the Spalart-Allmaras model with a rotation correction scheme
 \cite{aupoix_extensions_2003, dacles-mariani_numericalexperimental_1995} as described on the
@@
 correction reduces eddy viscosity in regions of rotational but non-turbulent flow where the
-original \gls{SA} model overestimates eddy viscosity. The \gls{SA} model implementation in Moltres
+original Spalart-Allmaras model overestimates eddy viscosity. The Spalart-Allmaras model implementation in Moltres
 solves for the modified dynamic viscosity $\tilde{\mu}$ (as opposed to $\tilde{\nu}$) as follows:
@@
   \begin{align*}
-    \mu_t &= \tilde{\mu}f_{v1} = \text{turbulent eddy viscosity}, &
+    \mu_t &= \tilde{\mu}f_{v1} = \text{turbulent eddy viscosity}, \\
     f_{v1} &= \frac{\chi^3}{\chi^3 + c_{v1}^3}, \\
-    \chi &= \frac{\tilde{\mu}}{\mu}, &
+    \chi &= \frac{\tilde{\mu}}{\mu}, \\
     \tilde{S} &= \Omega + \frac{\tilde{\nu}}{\kappa^2 d^2} f_{v2}, \\
-    f_{v2} &= 1 - \frac{\chi}{1+\chi f_{v1}}, &
+    f_{v2} &= 1 - \frac{\chi}{1+\chi f_{v1}}, \\
     \Omega &= \sqrt{2W_{ij}W_{ij}} = \text{vorticity magnitude}, \\
     W_{ij} &= \frac{1}{2}\left(\frac{\partial u_i}{\partial x_j} - \frac{\partial u_j}{\partial x_i}
-    \right), &
+    \right), \\
       f_w &= g\left(\frac{1 + c_{w3}^6}{g^6 + c_{w3}^6}\right)^{1/6}, \\
-      g &= r + c_{w2}\left(r^6 - r\right), &
+      g &= r + c_{w2}\left(r^6 - r\right), \\
       r &= \text{min}\left(\frac{\tilde{\nu}}{\tilde{S}\kappa^2d^2}, 10\right), \\
@@
 \cite{rumsey_turbulence_nodate}. The following subsections cover verification and validation tests
-of the \gls{SA} model in Moltres using reference problems for turbulent channel, pipe, and
+of the Spalart-Allmaras model in Moltres using reference problems for turbulent channel, pipe, and
 \gls{BFS} flow. Moltres input files
@@
 
-The Moltres model for this test is a 2-D 140 m$\times$0.5 m half-channel. The main flow direction
+The Moltres model for this test is a 2-D 140 m$\times$0.5 m half-channel
+(Figure \ref{fig:channel-geom}). The main flow direction
 is in the positive $x$ direction with the inlet and outlet on the left and right ends,
@@
 
-\begin{figure}[htb!]
+\begin{figure}[p]
+  \centering
+  \includegraphics[width=0.9\columnwidth]{channel-geom}
+  \caption{Channel geometry for the turbulent channel flow verification test. The red box indicates
+  the region shown by the close-up view in Figure \ref{fig:channel-mesh}.}
+  \label{fig:channel-geom}
+%\end{figure}
+%%
+%\begin{figure}[htb!]
   \centering
@@
   \caption{Relevant turbulent channel flow problem parameters. The $\tilde{\mu}_\text{inlet}$ value
-  at the inlet is set to fives times the $\mu$ value as recommended for the \gls{SA} model
+  at the inlet is set to fives times the $\mu$ value as recommended for the Spalart-Allmaras model
   \cite{spalart_one-equation_1994}.}
@@
 wall distances, and stresses from the \gls{DNS} data \cite{moser_direct_1999}
-and the Moltres \gls{SA} model. The Moltres \gls{SA} model is largely consistent with the reference
+and the Moltres Spalart-Allmaras model. The Moltres Spalart-Allmaras model is largely consistent with the reference
 \gls{DNS} flow data and reproduces the expected trends in the velocity and stress distributions
@@
   \caption{Comparison of turbulent channel flow results at Re$_\tau\approx395$ against reference
-  \gls{DNS} results \cite{moser_direct_1999}. The Moltres \gls{SA} model agrees consistently with
+  \gls{DNS} results \cite{moser_direct_1999}. The Moltres Spalart-Allmaras model agrees consistently with
   the reference data.}
@@
 
-\begin{figure}[htb!]
+\begin{figure}[p]
   \centering
-  \begin{subfigure}[b]{0.42\columnwidth}
+  \includegraphics[width=0.8\columnwidth]{pipe-geom}
+  \caption{Pipe geometry for the turbulent pipe flow verification test. The red box indicates
+  the region shown by the close-up views in Figure \ref{fig:pipe-mesh}.}
+  \label{fig:pipe-geom}
+%\end{figure}
+%%
+%\begin{figure}[htb!]
+  \centering
+  \begin{subfigure}[b]{0.40\columnwidth}
     \centering
@@
   \hfill
-  \begin{subfigure}[b]{0.56\columnwidth}
+  \begin{subfigure}[b]{0.53\columnwidth}
     \centering
@@
 The Moltres model for this test is a 2-D axisymmetric (R-Z coordinates) half-pipe with a radius of
-0.5 m and a length of 150 m. The main flow direction
+0.5 m and a length of 150 m (Figure \ref{fig:pipe-geom}). The main flow direction
 is in the positive $z$ direction with the inlet and outlet on the bottom and top ends,
@@
   \caption{Relevant turbulent pipe flow problem parameters. The $\tilde{\mu}_\text{inlet}$ value
-  at the inlet is set to fives times the $\mu$ value as recommended for the \gls{SA} model
+  at the inlet is set to fives times the $\mu$ value as recommended for the Spalart-Allmaras model
   \cite{spalart_one-equation_1994}.}
@@
 wall distances, and stresses from the experimental data \cite{laufer_structure_1954}
-and the Moltres \gls{SA} model. The Moltres \gls{SA} model is largely consistent with the reference
+and the Moltres Spalart-Allmaras model. The Moltres Spalart-Allmaras model is largely consistent with the reference
 experimental flow data and reproduces the expected trends in the velocity and stress distributions
@@
 Additionally, the \gls{NASA} Turbulence Modeling Resource website \cite{rumsey_turbulence_nodate}
-provides \gls{BFS} simulation results generated from the regular \gls{SA} model in the CFL3D
-Navier-Stokes CFD code developed at \gls{NASA} \cite{krist_cfl3d_1998}. The CFL3D \gls{SA} model
+provides \gls{BFS} simulation results generated from the regular Spalart-Allmaras model in the CFL3D
+Navier-Stokes CFD code developed at \gls{NASA} \cite{krist_cfl3d_1998}. The CFL3D Spalart-Allmaras model
 does not contain the rotation correction scheme
 \cite{aupoix_extensions_2003, dacles-mariani_numericalexperimental_1995} present in the Moltres
-\gls{SA} model.
+Spalart-Allmaras model.
 
-\begin{figure}[htb!]
+\begin{figure}[p]
+  \centering
+  \includegraphics[width=0.9\columnwidth]{backstep-geom}
+  \caption{Backward step geometry for the turbulent \gls{BFS} flow verification test. The red box indicates
+  the region shown by the close-up view in Figure \ref{fig:bfs-mesh}.}
+  \label{fig:backstep-geom}
+%\begin{figure}[htb!]
   \centering
@@
 
-The Moltres model for this test is a 2-D 160 m-long channel. The main flow direction is in the
+The Moltres model for this test is a 2-D 160 m-long channel (Figure \ref{fig:backstep-geom}.
+The main flow direction is in the
 positive $x$ direction with the inlet and outlet on the left and right ends, respectively. The
@@
 
-Figure \ref{fig:bfs} shows the velocity magnitude and streamlines around the step at $x=110$ m. The
-streamlines illustrate the recirculation zones created by flow separation past the step.
-Figure \ref{fig:bfs-plots} shows the normalized velocity distributions at various distances from
-step, the skin friction and skin pressure coefficients along the bottom wall, and the normalized
-turbulent shear stress distributions at various distances downstream of the step. The \gls{SA}
-model with the rotation correction scheme in Moltres performs largely similarly to the reference
-\gls{SA} model results provided on the \gls{NASA} Turbulence Modeling Resource website
-\cite{rumsey_turbulence_nodate}. Compared with the reference experimental data, the Moltres
-\gls{SA} model predicts more accurate velocity (Figure \ref{fig:bfs-downstream}) and turbulent
-stress distributions (Figure \ref{fig:bfs-stress}) along $x/H=1$ downstream of the
-step than the CFL3D \gls{SA} model. This indicates that the Moltres
-\gls{SA} model reproduces the sizes and shapes of the largest and second-largest recirculation
-zones (see Figure \ref{fig:bfs}) more accurately than the CFL3D \gls{SA} model. Table
-\ref{table:bfs-reattach} further supports the prior observation since the flow reattachment length
-estimate from Moltres is more consistent with the experimental data than CFL3D. The Moltres
-\gls{SA} velocity distributions in Figure \ref{fig:bfs-downstream} are largely consistent with the
-experimental data. Lastly, the reader should
-note that this turbulent \gls{BFS} flow setup induces complex flow separation effects despite the
-apparent simplicity of the geometry. All \gls{RANS}-based model results from the \gls{NASA}
-turbulence modeling resource website \cite{rumsey_turbulence_nodate} fail to accurately reproduce
-the turbulent shear stress distribution (Figure \ref{fig:bfs-stress}) from the reference
-experimental data. Therefore, the discrepancies observed in the turbulent shear stresses from the
-\gls{SA} model are within expectations of a \gls{RANS}-based model.
-
 \begin{table}[htb]
@@
   \caption{Relevant turbulent \gls{BFS} flow problem parameters. The $\tilde{\mu}_\text{inlet}$ value
-  at the inlet is set to fives times the $\mu$ value as recommended for the \gls{SA} model
+  at the inlet is set to fives times the $\mu$ value as recommended for the Spalart-Allmaras model
   \cite{spalart_one-equation_1994}.}
@@
 
+Figure \ref{fig:bfs} shows the velocity magnitude and streamlines around the step at $x=110$ m. The
+streamlines illustrate the recirculation zones created by flow separation past the step.
+Figure \ref{fig:bfs-plots} shows the normalized velocity distributions at various distances from
+step, the skin friction and skin pressure coefficients along the bottom wall, and the normalized
+turbulent shear stress distributions at various distances downstream of the step. The Spalart-Allmaras
+model with the rotation correction scheme in Moltres performs largely similarly to the reference
+Spalart-Allmaras model results provided on the \gls{NASA} Turbulence Modeling Resource website
+\cite{rumsey_turbulence_nodate}. Compared with the reference experimental data, the Moltres
+Spalart-Allmaras model predicts more accurate velocity (Figure \ref{fig:bfs-downstream}) and turbulent
+stress distributions (Figure \ref{fig:bfs-stress}) along $x/H=1$ downstream of the
+step than the CFL3D Spalart-Allmaras model. This indicates that the Moltres
+Spalart-Allmaras model reproduces the sizes and shapes of the largest and second-largest recirculation
+zones (see Figure \ref{fig:bfs}) more accurately than the CFL3D Spalart-Allmaras model. Table
+\ref{table:bfs-reattach} further supports the prior observation since the flow reattachment length
+estimate from Moltres is more consistent with the experimental data than CFL3D. The Moltres
+Spalart-Allmaras velocity distributions in Figure \ref{fig:bfs-downstream} are largely consistent with the
+experimental data. Lastly, the reader should
+note that this turbulent \gls{BFS} flow setup induces complex flow separation effects despite the
+apparent simplicity of the geometry. All \gls{RANS}-based model results from the \gls{NASA}
+turbulence modeling resource website \cite{rumsey_turbulence_nodate} fail to accurately reproduce
+the turbulent shear stress distribution (Figure \ref{fig:bfs-stress}) from the reference
+experimental data. Therefore, the discrepancies observed in the turbulent shear stresses from the
+Spalart-Allmaras model are within expectations of a \gls{RANS}-based model.
+
 \begin{table}[htb]
@@
   \caption{\gls{BFS} flow reattachment length estimates normalized by step height $H$.}
-  \begin{tabular}{l c c c}
+  \begin{tabular}{l S[table-format=1.2(2)]}
     \toprule
-    & {Experimental data \cite{driver_features_1985}} & {CFL3D \gls{SA} model
-  \cite{rumsey_turbulence_nodate}} & {Moltres \gls{SA} model} \\
+    Source & {Reattachment length [-]} \\
     \midrule
-    Reattachment length [-] & {$6.26 \pm 0.10$} & 6.1 & 6.36 \\
+    Experimental data \cite{driver_features_1985} & 6.26(10) \\
+    CFL3D Spalart-Allmaras model \cite{rumsey_turbulence_nodate} & 6.1 \\
+    Moltres Spalart-Allmaras model & 6.36 \\
     \bottomrule
@@
 turn induce turbulent temperature and \glspl{DNP} transport. This section presented the
-implementation, verification, and validation of the \gls{SA} turbulence model
+implementation, verification, and validation of the Spalart-Allmaras turbulence model
 \cite{spalart_one-equation_1994} in Moltres for turbulence
-modeling capabilities. The one-equation \gls{RANS}-based \gls{SA} model is computationally
+modeling capabilities. The one-equation \gls{RANS}-based Spalart-Allmaras model is computationally
 efficient relative to more complex \gls{RANS}-based and high-fidelity models while providing better
@@
 Verification and validation results for turbulent channel, pipe, and \gls{BFS} flow tests showed
-that the Moltres \gls{SA} model is largely consistent with reference simulation or experimental
-data in the literature. The Moltres \gls{SA} model includes a rotation correction scheme
+that the Moltres Spalart-Allmaras model is largely consistent with reference simulation or experimental
+data in the literature. The Moltres Spalart-Allmaras model includes a rotation correction scheme
 \cite{aupoix_extensions_2003, dacles-mariani_numericalexperimental_1995} and notably outperforms
-the reference \gls{SA} model in the turbulent \gls{BFS} flow test with more accurate velocity
+the CFL3D Spalart-Allmaras model in the turbulent \gls{BFS} flow test with more accurate velocity
 distributions and flow reattachment length estimate.

### 2024-11-13 verification.tex
Full draft + edits

@@
 collaborative \gls{VV} study based on the \gls{MSRE} pump transient tests conducted in the 1960s.
-Section \ref{sec:turbulence} describes the \gls{SA} turbulence model implementation in
+Section \ref{sec:turbulence} describes the Spalart-Allmaras turbulence model implementation in
 Moltres and presents the model \gls{VV} results against common wall-bounded turbulent flow case
@@
 
-Lastly, this chapter presented the implementation and verification of the \gls{SA} model
+Lastly, this chapter presented the implementation and verification of the Spalart-Allmaras model
 \cite{spalart_one-equation_1994} in Moltres for enabling wall-bounded turbulent molten salt flow
-modeling. The \gls{SA} model includes a rotation correction scheme
+modeling. The Spalart-Allmaras model includes a rotation correction scheme
 \cite{aupoix_extensions_2003, dacles-mariani_numericalexperimental_1995} for improved accuracy when
-modeling turbulent flow with curved streamlines. The \gls{SA} model performed well under the
+modeling turbulent flow with curved streamlines. The Spalart-Allmaras model performed well under the
 turbulent channel flow verification test \cite{moser_direct_1999}, the turbulent pipe flow

### 2024-11-19 msre.tex
add ray effect subsection

@@
 
-\begin{figure}[htb!]
+\begin{figure}[p]
   \centering
@@
 
+\subsubsection{$S_N$ Discretization \& Ray Effects}
+
+Angular resolution in 2-D (and later 3-D) flux simulations is important for accurately capturing
+the angular dependence in the neutron angular flux. This subsection compares the $k_\text{eff}$ and
+neutron scalar flux distribution solutions from the hybrid method with the $S_6$, $S_8$, $S_{10}$,
+and $S_{12}$ methods for the 2-D full-core with no rods inserted. Table
+\ref{table:2d-sn-convergence} shows the $k_\text{eff}$ estimates from those simulation results.
+The $k_\text{eff}$ varies by less than 1 pcm, thereby indicating that the angular discretization
+order used by the hybrid method has little impact on the integral reactor parameters.
+
+Figures \ref{fig:ray-effect-12} and \ref{fig:ray-effect-34} show the first four neutron group flux
+distributions from the hybrid method in the central reactor region consisting of three
+air-filled thimbles and a homogenized sample basket (bottom right). The geometry corresponds
+directly to Figure \ref{fig:full-geom-closeup} without the three control rods. Group 1 fluxes
+in the thimble form a distinctive ray effect artifacts arising from the regular geometrical shapes and
+insufficient angular resolution. The artifacts become less distinct with increasing angular
+discretization levels. They also diminish with increasing neutron energy groups due to shorter
+neutron mean free paths leading to more diffusive behavior.
+
+By Group 3 and 4, the artifacts are indiscernible to the eye across all angular discretization
+orders. Given that $k_\text{eff}$ is heavily weighted towards the later (slower) energy groups due
+to larger fission cross sections, it is clear why Table \ref{table:2d-sn-convergence} shows little
+differences in $k_\text{eff}$ across the angular discretization orders. Therefore, $S_8$ is
+adequate for control rod worth calculations in the \gls{MSRE}. 
+Further investigations into the angular discretization level may be warranted when applying the
+hybrid method to reactors with more air-filled regions such as high-temperature gas-cooled
+reactors.
+
+\begin{table}[t]
+  \centering
+  \caption{Hybrid method $k_\text{eff}$ estimates for the 2-D full-core model with zero rods
+  inserted and $S_N$ schemes from $S_6$ to $S_{12}$.}
+  \begin{tabular}{l S}
+    \toprule
+    Angular discretization & {$k_\text{eff}$} \\
+    \midrule
+    $S_6$ & 1.108414 \\
+    $S_8$ & 1.108423 \\
+    $S_{10}$ & 1.108422 \\
+    $S_{12}$ & 1.108420 \\
+    \bottomrule
+  \end{tabular}
+  \label{table:2d-sn-convergence}
+\end{table}
+
+\begin{figure}[p]
+  \small
+  \centering
+  \includegraphics[width=0.7\columnwidth]{ray-effect-12}
+  \caption{Group 1 and 2 neutron flux distributions in the hybrid $S_N$-diffusion method correction
+  region with $S_6$, $S_8$, $S_{10}$, \& $S_{12}$. Ray effects diminish with increasing
+  angular fidelity, but still persist with $S_{12}$ in group 1 and 2.}
+  \label{fig:ray-effect-12}
+\end{figure}
+
+\begin{figure}[p]
+  \small
+  \centering
+  \includegraphics[width=0.7\columnwidth]{ray-effect-34}
+  \caption{Group 3 and 4 neutron flux distributions in the hybrid $S_N$-diffusion method correction
+  region with $S_6$, $S_8$, $S_{10}$, \& $S_{12}$. Ray effects are barely visible in group
+  3 and 4 due to shorter neutron mean free paths.}
+  \label{fig:ray-effect-34}
+\end{figure}
+
+\FloatBarrier
+
 \section{3-D Neutronics Models \& Results} \label{sec:3d-results}

### 2024-11-23 msre.tex
Scaling test and rod cusping

@@
   \includegraphics[width=\columnwidth]{msre-quarter-no-rod-power}
-  \caption{Normalized channel fission rate distribution of the 2-D \gls{MSRE} quarter-core model.
+  \caption{Normalized channel fission rate distribution of the 2-D \gls{MSRE} quarter-core model
+    with the rod withdrawn.
   This figure depicts the 1/8th-core distribution due to symmetry about the diagonal.}
@@
   \includegraphics[width=\columnwidth]{msre-quarter-rod-power}
-  \caption{Normalized channel fission rate distribution of the 2-D \gls{MSRE} quarter-core model.
+  \caption{Normalized channel fission rate distribution of the 2-D \gls{MSRE} quarter-core model
+    with the rod inserted.
   This figure depicts the 1/8th-core distribution due to symmetry about the diagonal.}
@@
 The total fission rate includes fission rate contributions from peripheral fuel salt regions.
+Division by total fission rate normalizes the fission rates across the different neutronics
+methods.
 Figures \ref{fig:1/4-no-rod} and \ref{fig:1/4-rod} show the normalized fission rate distributions
@@
   \item Set a fixed outer vessel thickness of 2.54 cm.
-  \item Homogenized the lower plenum by 90.8\% salt- 9.2\% INOR-8 (vessel) by volume.
+  \item Homogenized the lower plenum by 90.8\% salt-9.2\% INOR-8 (vessel) by volume.
   \item Removed the reactor access nozzle and the various components within it.
@@
 hybrid $S_N$-diffusion method. The hybrid simulations ran with a maximum stabilization factor of
-$c=250$, a void constant $\varsigma=0.5$, and a uncorrected diffusion coefficient cap of $D_g=2.5$.
+$c=250$, a void constant of $\varsigma=0.5$, and a uncorrected diffusion coefficient cap of $D_g=2.5$.
 Refer to Sections \ref{sec:void-treatment} and \ref{sec:diffcoef-cap} for the relevant discussion
@@
 work. Moltres uses the \gls{MPI} communication protocol for parallelizing simulations. For this
-work, simulations ran with 1-to-1 ratio of \gls{MPI} ranks and processor cores (i.e., 32 \gls{MPI}
-ranks for each compute node).
+work, simulations ran with 1-to-1 ratio of \gls{MPI} ranks to processor cores (i.e., 32 \gls{MPI}
+ranks per compute node).
 
 Every 3-D hybrid method $k$-eigenvalue
-simulation of the \gls{MSRE} consisted of 747,688,760 \glspl{DOF} and 1,690,777,728 \glspl{DOF} for
-the neutron diffusion and $S_N$ components, respectively. Each \gls{DOF} corresponds to a scalar
+simulation of the \gls{MSRE} consisted of 750M \glspl{DOF} and 1.69B \glspl{DOF} for
+% 747,688,760 \glspl{DOF} and 1,690,777,728 \glspl{DOF} for
+the neutron diffusion and $S_N$ subsolvers, respectively. Each \gls{DOF} corresponds to a scalar
 or angular group flux variable value defined at a mesh node. Each simulation ran with the distributed
@@
 
-On forty nodes, the simulations
+On forty nodes, each full-core rod worth simulation
 required approximately 2.2 h to complete for a total of 88 node-hours or 2816 core-hours. The
 distributed mesh capability comes with the cost of increased data transfer times between the $S_N$
-and diffusion solvers due to increased inter-core data transfers. Transfers accounted for
+and diffusion solvers due to increased inter-processor data transfers. Transfers accounted for
 approximately 0.675 h or 30 \% of the total simulation time. Future research into optimizing mesh
@@
 hybrid method simulations. The solution time for the initial neutron diffusion calculation with zero
-drift correction also takes up approximately 30 \% of the total simulation time. After subtracting
-the minor computational cost of evaluating the drift term even if it is zero, the hybrid method
+drift correction also takes up approximately 30 \% of the total simulation time. After taking into
+consideration the computational cost of evaluating the drift term, the hybrid method
 requires about four times the solution time of the standard neutron diffusion method.
@@
 solution time decreases when more cores are used while keeping the problem size fixed. This strong
-scaling study used 3-D quarter-core models run on 10, 20, 30, and 40 compute nodes.
-Figure * shows the results of this scaling study.
-The hybrid method implementation scales well as shown by the nearly linear scaling between the
-solution time and the number of processors. Therefore, the hybrid method appears promising for the
-modeling of larger reactor analyses with little scaling performance penalty.
+scaling study used 3-D quarter-core models run on 10, 20, 40, and 80 compute nodes.
+Figure \ref{fig:scaling} shows the results of this scaling study. The total wall time scales at a
+sublinear rate of approximately $q=0.7$. The total wall time is broken down into two components:
+the solver wall time (i.e., initialization, numerical solver, etc.) and the transfer time (i.e,
+data transfer between the $S_N$ and diffusion solvers). The solver wall time scales at a sublinear
+rate of approximately $q=0.8$, while the transfer wall time scales at approximately $q=0.6$ from
+ten to forty nodes before registering a wall time increase when going from forty nodes to eighty nodes.
+It is likely that the increase in inter-processor data transfers due to increased data distribution
+outweighed the increase in computational power when running the problem with more compute nodes.
+Further investigation is necessary to confirm this hypothesis.
+
+Nevertheless, the total wall time exhibits decent strong scaling performance. For Moltres
+simulations in the next chapter, I optimized the data transfer setup which reduced the
+proportion of transfer wall time in the total wall time to approximately 18 \%. Therefore, the total
+wall times would scale closer to the $q=0.8$ rate observed with solver wall times.
+This is a promising indication that the hybrid method would scale well for the modeling of larger
+reactor analyses.
+
+\begin{figure}[t]
+  \centering
+  \begin{subfigure}[b]{0.49\columnwidth}
+    \centering
+    \includegraphics[width=\columnwidth]{scaling-total}
+    \caption{Total wall time}
+  \end{subfigure} \\
+  \begin{subfigure}[b]{0.49\columnwidth}
+    \centering
+    \includegraphics[width=\columnwidth]{scaling-solver}
+    \caption{Solver wall time}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.49\columnwidth}
+    \centering
+    \includegraphics[width=\columnwidth]{scaling-transfer}
+    \caption{Transfer wall time}
+  \end{subfigure}
+  \caption{The total, solver, and transfer wall time of hybrid method simulations of the 3-D
+  quarter-core model on 10, 20, 40, and 80 compute nodes (32 processors per node) of the Polaris
+  supercomputer. The solver
+  time scales almost linearly ($q=0.8$) with the number of processors, while the transfer time
+  scales poorly with an increase in wall time at 80 nodes. All axes are in log scale.}
+  \label{fig:scaling}
+\end{figure}
 
@@
 exhibits good computational scaling performance which is essential for large 3-D reactor analyses.
-Excessively large memory requirements were mitigated through the distributed mesh capability in
-Moltres. Some performance optimizations may be possible to reduce data transfer times
-between the $S_N$ and diffusion solvers which currently take up about 30 \% of the total solution
+The simulation setups mitigated excessively large memory requirements through the distributed mesh
+capability in Moltres. Some performance optimizations may be possible to reduce data transfer times
+between the $S_N$ and diffusion solvers which currently take up about 18 \% of the total solution
 time as a result of the inter-processor communications.

### 2024-11-23 transient.tex
Scaling test and rod cusping

@@
 experimental measurements.
+
+\subsection{Rod Cusping Effect Correction}
+
+In deterministic codes with spatial discretization, the material interfaces of a partially-inserted
+control rod may not always align with mesh elements. This results in a heterogenous
+mesh element that is composed of a control rod material at the top and a non-rod material at the
+bottom (for control rods that are inserted from above). The rod cusping effect occurs when
+homogenizing mixed mesh elements using the volume weighting method to combine dissimilar material
+neutron group constants. Rod cusping error typically decreases with increasing axial mesh
+resolution because the size of the mixed mesh elements become smaller relative to the entire
+problem domain.
+
+Several rod cusping correction methods exist in literature that are variations of the flux-volume
+weighting approach to generate more accurate homogenized group constants in the mixed mesh
+elements \cite{yamamoto_cell_2004, graham_subplane_2018, schunert_control_2019}. However, due to
+a failed flux-volume weighting implementation and time constraints, this work uses an empirical
+approach for rod cusping effect correction. The approach involves precomputing the $k_\text{eff}$
+with the control rod at various levels of insertion within a mesh element. These simulations use
+the simple volume weighting approach to compute a homogenized group constant value as follows:
+%
+\begin{gather}
+  \Sigma_\text{vw} = \frac{V_\text{r} \Sigma_\text{r} + V_\text{nr} \Sigma_\text{nr}}{V_\text{r} + V_\text{nr}} \label{eq:vw}
+  \shortintertext{where}
+  \begin{align*}
+    \Sigma_\text{vw} &= \mbox{ homogenized group constant in mixed mesh element,} \\
+    V_\text{r} &= \mbox{ rod volume fraction in mixed mesh element,} \\
+    \Sigma_\text{r} &= \mbox{ rod group constant value,} \\
+    V_\text{nr} &= \mbox{ non-rod volume fraction in mixed mesh element,} \\
+    \Sigma_\text{nr} &= \mbox{ non-rod group constant value.}
+  \end{align*}
+\end{gather}
+%
+Note that vertical material and mesh interfaces in the 3-D \gls{MSRE} model always align perfectly
+because the control rod moves vertically. Therefore, the mixed mesh elements occur only around the
+bottom tip of the control rod. Additionally, the volume fraction is equivalent to the height
+fraction, and all mixed mesh elements at the same height share the same volume fraction. Figure
+\ref{fig:rod-cusping-1} shows how the fraction of reactivity change changes with the control rod
+volume fraction when the control rod is at around mid-reactor and full insertion heights. The
+figure depicts the nonlinear relation between the reactivity change and volume fraction that the
+``cusping effect'' is named after. Within this small range of rod motion, the relation should be
+approximately linear.
+
+\begin{figure}[t]
+    \centering
+    \begin{subfigure}[t]{.49\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{rod-cusping-1}
+        \caption{Uncorrected volume weighting}
+        \label{fig:rod-cusping-1}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{.49\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{rod-cusping-2}
+        \caption{Corrected volume weighting}
+        \label{fig:rod-cusping-2}
+    \end{subfigure}
+    \caption{Fraction of reactivity change against control rod volume fraction in mixed mesh
+    elements with the rod inserted at mid-reactor height (27.4 to 29.4 inches) and full insertion
+    height (0.0 to 1.9 inches).}
+    \label{fig:rod-cusping}
+\end{figure}
+
+Recognizing that the rod cusping effect approximately follows a $y=x^p$ curve profile, I applied
+curve-fitting to obtain the reactivity changes using corrected volume fractions shown in Figure
+\ref{fig:rod-cusping-2}. The corrected volume fraction $V_r'$ is calculated as follows
+%
+\begin{align}
+  V_r' &= cV_r^p
+  \shortintertext{where}
+  p &= 2.5148, \nonumber \\
+  c &= 0.43807. \nonumber
+  \end{align}
+%
+The corrected volume fraction produces the desired approximately linear relation between reactivity
+change and control rod volume fraction. While this empirical approach worked reasonably well for
+this study, a more robust technique will be necessary for future control rod simulations beyond
+this work.
 

### 2024-11-26 transient.tex
Flowchart & u233 temp coef

@@
 
-\begin{figure}[htb!]
+\begin{figure}[t]
+  \tikzstyle{every node}=[font=\small]
   \centering
-%  \includegraphics[width=0.8\columnwidth]{rod-height}
+  \begin{tikzpicture}
+    \node (1) [solver] {\textbf{$\bm{S_N}$ transport subsolver}};
+    \node (2) [solver, right of=1, xshift=4cm] {\textbf{Neutron\\ diffusion\\ subsolver}};
+    \node (3) [solver, right of=2, xshift=4cm] {\textbf{DNP \& temperature subsolver}};
+    \draw [arrow] (1.10) -- node[anchor=south, text width=4cm, align=center] {$\vec{D}_g, \gamma_g$} (2.170);
+    \draw [arrow] (2.190) -- node[anchor=north, text width=4cm, align=center] {$\phi_g, C_i, T$} (1.350);
+    \draw [arrow] (2.10) -- node[anchor=south, text width=4cm, align=center] {$Q_f, Q$} (3.170);
+    \draw [arrow] (3.190) -- node[anchor=north, text width=4cm, align=center] {$C_i, T$} (2.350);
+    \node (5) [bound, fit=(2) (3), inner sep=8pt, label={[align=left]above:\textbf{Inner iteration}}] {};
+    \node (6) [bound, fit=(1) (5), inner sep=15pt, label={[align=left]above:\textbf{Outer iteration}}] {};
+  \end{tikzpicture}
   \caption{The nested iteration structure coupling the $S_N$, neutron diffusion, and \gls{DNP}
@@
 mass is estimated to be 35.234 kg reported on January 12 1969 \cite{thoma_chemical_1971}. The
-experiment was performed on January 16 1969 \cite{steffy_experimental_1970}. Table * shows
+experiment was performed on January 16 1969 \cite{steffy_experimental_1970}. Table 
+\ref{table:u233-composition} shows
 the breakdown and total isotopic content of uranium in the salt. The density of $^{233}$U salt is
@@
 
-\begin{figure}[htb!]
+\begin{figure}[t]
+  \tikzstyle{every node}=[font=\small]
   \centering
-%  \includegraphics[width=0.8\columnwidth]{rod-height}
+  \begin{tikzpicture}
+    \node (1) [solver] {\textbf{$\bm{S_N}$ transport subsolver}};
+    \node (2) [solver, right of=1, xshift=4cm] {\textbf{Neutron\\ diffusion\\ subsolver}};
+    \node (3) [solver, right of=2, xshift=4cm] {\textbf{In-core DNP \& temperature subsolver}};
+    \node (4) [solver, below of=3, yshift=-3cm] {\textbf{Ex-core DNP \& temperature subsolver}};
+    \draw [arrow] (1.10) -- node[anchor=south, text width=4cm, align=center] {$\vec{D}_g, \gamma_g$} (2.170);
+    \draw [arrow] (2.190) -- node[anchor=north, text width=4cm, align=center] {$\phi_g, C_i, T$} (1.350);
+    \draw [arrow] (2.10) -- node[anchor=south, text width=4cm, align=center] {$Q_f, Q$} (3.170);
+    \draw [arrow] (3.190) -- node[anchor=north, text width=4cm, align=center] {$C_i, T$} (2.350);
+    \draw [arrow] (3.280) -- node[anchor=west, text width=1cm, align=center] {$C^\text{out}_i$\\ $T^\text{out}$} (4.80);
+    \draw [arrow] (4.100) -- node[anchor=east, text width=1cm, align=center] {$C^\text{in}_i$\\ $T^\text{in}$} (3.260);
+    \node (5) [bound, fit=(3) (4), label={[align=center,rotate=90]left:\textbf{Salt loop}\\}] {};
+    \node (6) [bound, fit=(2) (5), label={[align=center,rotate=90]left:\textbf{Inner iteration}\\}] {};
+    \node (7) [bound, fit=(1) (6), label={[align=center,rotate=90]left:\textbf{Outer iteration}\\}] {};
+  \end{tikzpicture}
   \caption{The nested iteration structure coupling the $S_N$, neutron diffusion,
-  \gls{DNP}-temperature, and out-of-core
-  solvers for the rod drop simulation using the hybrid $S_N$-diffusion method.}
+  in-core \gls{DNP}-temperature, and ex-core
+  solvers for the reactivity insertion simulation using the hybrid $S_N$-diffusion method.}
   \label{fig:insertion-coupling}
@@
 
+\subsection{Temperature Reactivity Coefficient}
+
+This work compares isothermal temperature reactivity coefficients calculated from $k_\text{eff}$
+values measured at 872 K and 972 K with the $^{233}$-fueled \gls{MSRE} at operating fuel
+concentration. This approach is identical to that
+of the \gls{MSRE} experimental measurements \cite{engel_zero-power_1972} and the temperature
+coefficient calculation procedure used for the $^{235}$U-fueled \gls{MSRE} in Section
+\ref{sec:temp-coef}.
+
+Table \ref{table:temp-coef-u233} compares the isothermal reactivity coefficients from the experimental
+and numerical sources. No uncertainty data is available from the relevant \gls{ORNL} report
+\cite{engel_zero-power_1972}. However, the authors noted significantly different readings ranging
+from 12.4 to 15.3 pcm K$^{-1}$ due to gas bubbles introduced by the salt pump. The
+presence of gas bubbles led to smaller reactivity coefficient measurements. The numerical OpenMC
+and hybrid models overestimate the temperature reactivity coefficient magnitude by 12 \% and 11 \%,
+respectively, relative to the experimental data.
+
+\begin{table}[t]
+  \small
+  \centering
+  \caption{Isothermal temperature reactivity coefficients with $^{233}$U-fueled \gls{MSRE} from
+    \gls{MSRE} experimental data, \gls{MSRE} predicted value, and the OpenMC
+  and Moltres models in this work.}
+  \begin{tabular}{l S[separate-uncertainty=true,table-format=2.2(2)] S}
+    \toprule
+     & {Temperature coefficient [pcm K$^{-1}$]} & {Percentage discrepancy [\%]}\\
+     \cmidrule(rl){2-2} \cmidrule(l){3-3}
+    \gls{MSRE} experimental data & -15.3 & {-}\\
+    \gls{MSRE} predicted value & -15.8 & -3.4 \\
+    OpenMC (This work) & -17.14 & -12 \\
+    Hybrid (This work) & -16.96 & -11 \\
+    \bottomrule
+  \end{tabular}
+  \label{table:temp-coef-u233}
+\end{table}
+
 \subsection{Reactivity Insertion Numerical Results \& Discussion}

### 2024-12-04 msre.tex
Flowchart & u233 temp coef
Fix equation and typo

@@
 \begin{align}
-  \text{Normalized fission rate of channel }i &= \frac{\text{Fission rate in channel }i}{
-  \text{Total fission rate}}\cdot\frac{\text{Total number of channels}}{
-    \text{Total fission rate in channels}} \nonumber \\
-                                              &= \frac{\sum^G_{g=1}\int_{V_i}\Sigma_{f,g}
-  \phi_g(\vec{r})}{\sum^G_{g=1}\int_V\Sigma_{f,g}\phi_g(\vec{r})}\cdot\frac{N_i}{
-  \sum^{N_i}_{i=1}\sum^G_{g=1}\int_{V_i}\Sigma_{f,g}\phi_g(\vec{r})}.
+%  \text{Normalized fission rate of channel }i &= \frac{\text{Fission rate in channel }i}{
+%  \text{Total fission rate}}\cdot\frac{\text{Total number of channels}}{
+%    \text{Total fission rate in channels}} \nonumber \\
+%                                              &= \frac{\sum^G_{g=1}\int_{V_i}\Sigma_{f,g}
+%  \phi_g(\vec{r})}{\sum^G_{g=1}\int_V\Sigma_{f,g}\phi_g(\vec{r})}\cdot\frac{N_i}{
+%  \sum^{N_i}_{i=1}\sum^G_{g=1}\int_{V_i}\Sigma_{f,g}\phi_g(\vec{r})}.
+  \text{Normalized fission rate of channel }i &= \frac{(\text{Total number of channels})\times
+    (\text{Fission rate in channel }i)}{\text{Total fission rate in channels}} \nonumber \\
+                                              &= \frac{N_i\sum^G_{g=1}\int_{V_i}\Sigma_{f,g}
+  \phi_g(\vec{r}) dV}{
+  \sum^{N_i}_{i=1}\sum^G_{g=1}\int_{V_i}\Sigma_{f,g}\phi_g(\vec{r}) dV}.
 \end{align}
@@
 Figures \ref{fig:1/4-no-rod} and \ref{fig:1/4-rod} show the normalized fission rate distributions
-of the quarter-core model for the ``No Rod'' and ``Rod'' cases, respectively. I presented 1/8th
-slices of the core in the figures because the distributions are symmetrical about the $y=x$ line
+of the quarter-core model for the ``No Rod'' and ``Rod'' cases, respectively. The figures present
+1/8th slices of the core because the distributions are symmetrical about the $y=x$ line
 outside of statistical variances for OpenMC-CE. I combined the midpoints between the centroids of
@@
 
-\subsubsection{Temperature Reactivity Coefficient}
+\subsubsection{Temperature Reactivity Coefficient} \label{sec:temp-coef}
 

### 2024-12-04 presentation/1d.tex
Presentation commit

--- 2024-11-04 hybrid-nts.tex
@@
-In Chapter \ref{chap:hybrid}, I laid out the mathematical background and numerical implementation
-of the hybrid $S_N$-diffusion method for improving neutronic accuracy in reactor systems containing
-highly neutron-absorbing regions. I developed and implemented the diffusion and drift correction
-schemes for the hybrid method on Python and Moltres, respectively. This chapter demonstrates
-the hybrid method through 1-D, 2-D, and 3-D $k$-eigenvalue simulations of the \gls{MSRE}.
-I employed the OpenMC Monte Carlo neutral particle transport software \cite{romano_openmc:_2015} to
-generate group constants for the hybrid method and reference solutions for method verification.
-The 1-D models serve to verify the newly-implemented $S_N$ and hybrid methods
-against OpenMC as well as to analyze the impact of key input parameters on the hybrid method. The
-2-D models test the hybrid method performance through increased spatial dimensionality and reduced
-geometric asymmetry relative to the 1-D models. Lastly, the 3-D models serve as neutronics model
-validation against the \gls{MSRE} benchmark model by Fratoni et al.\ \cite{fratoni_molten_2020}, a
-testbed for \gls{HPC} scaling tests, and preparation for the time-dependent multiphysics
-simulations of \gls{MSRE} reactivity insertion transients.
+\begin{frame}
+  \frametitle{Hybrid $S_N$-Diffusion Method: 1-D Neutronics Eigenvalue Simulations}
+  \textbf{1-D Neutronics Model Geometries}
+  \begin{figure}[htb!]
+    \centering
+    \includegraphics[width=0.9\columnwidth]{case-geometry}
+    \caption{Geometries of the 1-D test cases. The material labeled ``mixture'' represents a
+      homogeneous mixture of fuel and graphite at a ratio of 22.5\%-77.5\% by volume. All geometries
+      have reflective boundary conditions on the boundary at $x=0$ cm. The right-side boundaries are
+      reflective for Cases 1a and 1b, and vacuum for Cases 2a, 2b, 3a, and 3b.}
+    \label{fig:case-geom}
+  \end{figure}
+\end{frame}
 
-Section \ref{sec:msre-gc} details the OpenMC model setup for group constant generation. Section
-\ref{sec:nts-methods} describes and labels the different neutronics methods employed for the
-$k$-eigenvalue simulations. Section \ref{sec:test-metrics} details basic test metrics relevant to
-neutronics verification and validation. Sections \ref{sec:1d-results}, \ref{sec:2d-results}, and
-\ref{sec:3d-results} discuss the model setup, results, and discussions for the 1-D, 2-D, and 3-D
-\gls{MSRE} models, respectively.
+\begin{frame}
+  \frametitle{Hybrid $S_N$-Diffusion Method: 1-D Neutronics Eigenvalue Simulations}
+  \begin{columns}
+    \column{5.5cm}
+    \textbf{1-D Neutronics Model Setup}
+    \vspace{.2cm}
 
-\section{Group Constant Data Generation on OpenMC} \label{sec:msre-gc}
-
-Group constants relevant to the neutron diffusion, $S_N$ neutron transport, or both
-methods are:
-%
     \begin{itemize}
-  \item $\Sigma_{t,g}$: Macroscopic total cross section in group $g$,
-  \item $\Sigma_{r,g}$: Macroscopic removal cross section in group $g$,
-  \item $\Sigma_s^{g'\rightarrow g}$: Macroscopic group-to-group scattering cross section matrix,
-  \item $\Sigma_{s,l}^{g'\rightarrow g}$: $l$-th Legendre moment of the macroscopic
-    group-to-group scattering cross section matrix,
-  \item $\Sigma_{sp,l}^{g'\rightarrow g}$: $l$-th Legendre moment of the macroscopic
-    group-to-group scattering production cross section matrix,
-  \item $D_g$: $P_1$-based diffusion coefficient in group $g$,
-  \item $\nu\Sigma_{f,g}$: Product of the average number of neutrons produced per fission and the
-    macroscopic fission cross section in group $g$,
-  \item $\chi_g$: Neutron fission spectrum in group $g$.
+      \item Material compositions derived from the MSRE design
+      \item Reduced gadolinium content in control rod to reduce rod worth from 50000 pcm to 20000 pcm
+      \item Eight neutron energy groups
+      \item Group constants generated using OpenMC with up to 2nd-order Legendre expansions of scattering
+        matrices
+      \item Uniform temperature at 900 K
     \end{itemize}
-%
-I generated the group constants using OpenMC's multigroup cross section generation capability
-\cite{boyd_multigroup_2019}. A group constant postprocessing script for Moltres
-(\texttt{moltres\_xs.py}) parses OpenMC simulation output files and rewrites the group constant
-data into a JSON format file suitable for Moltres simulations. While OpenMC provides most of the
-group constants directly, the postprocessing script calculates $\Sigma_{r,g}$ as
-%
-\begin{align}
-  \Sigma_{r,g} =& \sum^G_{g'\neq g}\Sigma_s^{g\rightarrow g'}+\Sigma_{a,g}-\left(\Sigma_{sp}^{g
-    \rightarrow g} - \Sigma_s^{g\rightarrow g}\right)
-  \shortintertext{where}
-      \Sigma_{a,g} =& \mbox{ macroscopic absorption cross section in group $g$,} \nonumber \\
-      \Sigma_{sp}^{g\rightarrow g} =& \mbox{ macroscopic scattering production cross section from
-      group $g$ to $g$.} \nonumber
-\end{align}
-%
-$\Sigma_{r,g}$ primarily represents the loss of neutrons from group $g$ through outscattering and
-absorption. $\Sigma_{sp}^{g\rightarrow g}$ incorporates neutron multiplication effects from neutron
-knockout reactions into the scattering cross section. Neutronics codes commonly identify neutron
-knockout reactions as scattering reactions, but $\Sigma_{r,g}$ is a convenient parameter in which
-to incorporate neutron knockout effects in the neutron diffusion method. I ran all test cases
-for $S_N$ and hybrid method calculations with scattering cross sections approximated by up to 2nd
-Legendre moments ($L=2$).
-
-OpenMC uses the $P_1$ flux-limited formulation \cite{pomraning_flux-limited_1984} by default for
-calculating $D_g$ as follows
-%
-\begin{align}
-  D_g =& \frac{1}{3\Sigma_{tr,g}},
-  \shortintertext{where}
-  \Sigma_{tr,g} =& \frac{\langle\Sigma_{t,g}\phi_g\rangle-\langle\Sigma_{s1,g}\phi_g\rangle}
-  {\langle\phi_g\rangle}, \nonumber \\
-  \langle\Sigma_{t,g}\phi_g\rangle =& \int_{r\in V}dr \int_{4\pi}d\Omega\int^{E_{g-1}}_{E_g}dE\
-  \Sigma_{t,g}(r,E)\Psi(r,E,\Omega), \nonumber \\
-  \langle\Sigma_{s1,g}\phi_g\rangle =& \int_{r\in V}dr \int_{4\pi}d\Omega\int^{E_{g-1}}_{E_g}dE
-  \int_{4\pi}d\Omega'\int^{\infty}_0dE'\int^1_{-1}d\mu\ \mu\Sigma_s(r,E'\rightarrow E,\Omega'\cdot
-  \Omega)\phi(r,E',\Omega'), \nonumber \\
-  \langle \phi \rangle =& \int_{r\in V}dr\int_{4\pi}d\Omega\int^{E_{g-1}}_{E_g}dE\ \Psi(r,E,\Omega)
-  .\nonumber
-\end{align}
-
-I ran all \gls{MSRE} neutronics simulations with the eight neutron energy group structure proposed
-by Jaradat for \gls{MSRE} analysis \cite{jaradat_development_2021-1}.
-Table \ref{table:energy-group} shows the upper neutron energy bounds of the eight-group structure.
-I selected this energy group structure as a compromise between accuracy and computational expense
-among the various group structures investigated by Jaradat ranging from four to sixteen groups.
-
-\begin{table}[htb]
+    \column{5.5cm}
+    \begin{table}[h]
       \centering
@@
     \end{table}
+  \end{columns}
+\end{frame}
 
-I set the material compositions and densities of every component in the \gls{MSRE} models using
-reference data from \gls{MSRE} benchmark specifications collated by Fratoni et al.
-\cite{fratoni_molten_2020}. All OpenMC models utilize the ENDF/B-VII.1 nuclear data library
-\cite{chadwick_endf/b-vii.1_2011} and include thermal scattering cross section data for beryllium
-and carbon target nuclei in the fuel salt and graphite moderator. All models have a uniform
-temperature of 900 K unless otherwise stated. Detailed information pertaining specifically to the
-1-D, 2-D, and 3-D OpenMC models may be found in the respective subsections below.
+\begin{frame}
+  \frametitle{Hybrid $S_N$-Diffusion Method: 1-D Neutronics Eigenvalue Simulations}
+  \textbf{1-D Neutronics Model Numerical Solvers}
+  \vspace{.2cm}
 
-\section{Neutronics Methods} \label{sec:nts-methods}
+  All 1-D cases ran on each of the following numerical solvers:
+  \begin{enumerate}
+    \item OpenMC in continuous energy mode (OpenMC-CE)
+    \item OpenMC in multigroup mode (OpenMC-MG)
+    \item Standard neutron diffusion method (Moltres \& Python)
+    \item $S_8$ method (Moltres \& Python)
+    \item Hybrid $S_8$-diffusion method (Moltres \& Python)
+  \end{enumerate}
+  \vspace{.2cm}
 
-I ran all cases on OpenMC to generate the required group constants and reference $k_\text{eff}$
-values to assess the accuracy of the deterministic methods ($S_N$, neutron diffusion, hybrid 
-methods) against OpenMC in continuous-energy (OpenMC-CE) and multigroup (OpenMC-MG) modes. Table
-\ref{table:var} details how OpenMC and the deterministic methods represent the position, direction
-of travel, neutron energy, and angle-dependence in $\Sigma_s$. The hybrid $S_N$-diffusion method
-features representations from either $S_N$ or neutron diffusion depending on the subregion under
-consideration. Comparing OpenMC-MG results with OpenMC-CE results allows us
-to quantify errors arising from neutron energy group discretization and the scattering cross
-section approximations.
-
-\begin{table}[htb!]
-  \centering
-  \footnotesize
-  \caption{Variable representations in OpenMC under continuous-energy (OpenMC-CE) and multigroup
-  (OpenMC-MG) modes, and in the $S_N$ neutron transport and neutron diffusion.}
-  \begin{tabular}{l l l l l}
-    \toprule
-    Variable & OpenMC-CE & OpenMC-MG & $S_N$ Transport & Diffusion \\
-    \midrule
-    Position, $\bm{r}$ & Continuous & Continuous & Discrete & Discrete \\
-    Direction of travel, $\bm{\hat{\Omega}}$ & Continuous & Continuous & Discrete & N/A \\
-    Energy, $E$ & Continuous & Discrete & Discrete & Discrete \\
-    Angle-dependence in $\Sigma_s$ & Continuous & $L=2$ & $L=2$ & N/A \\
-    \bottomrule
-  \end{tabular}
-  \label{table:var}
-\end{table}
-
-The rest of this section pertains to the deterministic neutronics methods used in this work for
-modeling the \gls{MSRE}.
-For the 1-D models, I performed neutronics calculations with the $S_N$, neutron diffusion, and
-hybrid method implementations on Moltres and Python. Table \ref{table:nts-methods} highlights the
-differences between the two implementations. The $S_N$ neutron transport method implementation on
-Moltres follows the methodology for solving the \gls{SAAF} formulation of the $S_N$ equations as
-described in Section \ref{sec:saaf} while the implementation on Python follows the methodology
-for solving the original first-order formulation as described in Section \ref{sec:python-sn}.
-
-The implementations on Moltres discretize the problem domain in space using the finite element
-method while the implementations on Python use finite difference or diamond difference schemes. The
-most crucial difference lies in the transport correction formulations for the hybrid method. The
-Moltres implementation uses the drift correction scheme described in Section
-\ref{sec:drift-correction} while the Python implementation uses the diffusion correction
-scheme described in Section \ref{sec:diffusion-correction}.
-
-In this chapter, I refer to the methods and results from the Moltres and Python implementations
-with the Moltres- or Python- prefix, i.e., Moltres-$S_N$, Moltres-diffusion, Moltres-hybrid,
-Python-$S_N$, Python-diffusion, and Python-hybrid.
-
-For 2- and 3-D models, I abandoned the diffusion correction formulation in favor of the drift
-correction formulation due to superior numerical performance in the latter. Readers may find the
-analysis supporting this decision in the analysis of 1-D results in Section \ref{sec:1d-results}.
-
-\begin{table}[htb!]
-  \centering
-  \footnotesize
-  \caption{Implementation differences of neutronics methods on Moltres and Python.}
-  \begin{tabular}{p{0.14\linewidth} p{0.25\linewidth} p{0.15\linewidth} p{0.15\linewidth} p{0.17\linewidth}}
-    \toprule
-    Implementations & Spatial discretization & $S_N$ formulation & $S_N$ acceleration &
-    Hybrid transport\newline correction \\
-    \midrule
-    Moltres & Finite element method & SAAF & Nonlinear diffusion acceleration & Drift correction \\
-    Python & Finite difference (Diffusion)\newline Diamond difference ($S_N$) & First-order & None & Diffusion correction \\
-    \bottomrule
-  \end{tabular}
-  \label{table:nts-methods}
-\end{table}
-
-\section{Basic Test Metrics} \label{sec:test-metrics}
-
-In this section, I define several metrics for comparing simulation results among the various
-neutronics methods.
-
-For steady-state eigenvalue calculations, all solvers directly provide the effective neutron
-multiplication factor ($k_\text{eff}$), the rate of neutron population increase between successive
-neutron generations. Reactivity is defined as
-%
+  \textbf{Reactivity \& Reactivity Difference}
   \begin{gather}
-  \rho \equiv \frac{k_\text{eff}-1}{k_\text{eff}}.
+    \mbox{Reactivity } \rho \equiv \frac{k_\text{eff}-1}{k_\text{eff}}.
   \end{gather}
-%
-Differences in reactivity, for either comparisons between neutronics methods or control rod worth
-calculations, can be calculated as
-%
   \begin{gather}
-  \Delta\rho_{1,2} = \rho_1 - \rho_2 =
+    \Delta\rho = \rho_1 - \rho_2 =
     \frac{k_{\text{eff},1}-k_{\text{eff},2}}{k_{\text{eff},1}k_{\text{eff},2}}.
   \end{gather}
+\end{frame}
 
-Spatially dependent variables (e.g., neutron group flux, neutron drift) are sampled at fixed 0.1 cm
-intervals along the specified lines (e.g., horizontal, diagonal, etc.) regardless of the actual
-mesh resolution.
-
-\section{1-D Neutronics Models \& Results} \label{sec:1d-results}
-
-\subsection{1-D Neutronics Model Setup \& Geometries}
-
-I designed six 1-D test cases with increasing complexity to verify the $S_N$ and hybrid
-$S_N$-diffusion implementations and to test the performance of the hybrid
-method in response to various geometrical features. The last two cases resemble the
-reference \gls{MSRE} design \cite{robertson_msre_1965}, which has centrally located control rods
-and air-filled rod guide tubes. Figure \ref{fig:case-geom} shows the geometries of Cases 1a to
-3b. All geometries have reflective boundary conditions at $x=0$ cm, thereby
-forming half-core or repeating unit cell models. Cases 1a and 1b are repeating unit
-cell models with reflecting boundaries on the right-side boundaries. Cases 2a, 2b, 3a, and 3b
-are 1-D half-core models with vacuum boundaries on the right-side boundaries.
-
-\begin{figure}[htb!]
-  \centering
-  \includegraphics[width=0.8\columnwidth]{case-geometry}
-  \caption{Geometries of the 1-D test cases. The material labeled ``mixture'' represents a
-    homogeneous mixture of fuel and graphite at a ratio of 22.5\%-77.5\% by volume. All geometries
-    have reflective boundary conditions on the boundary at $x=0$ cm. The right-side boundaries are
-    reflective for Cases 1a and 1b, and vacuum for Cases 2a, 2b, 3a, and 3b.}
-  \label{fig:case-geom}
-\end{figure}
-
-Cases 1a and 1b are simple test cases containing neutron multiplying regions. Cases 1a and 1b serve
-to verify basic neutronics phenomena modeling (e.g., fission, scattering, absorption) of the newly
-implemented $S_N$ and neutron diffusion methods in Python and $S_N$ method in Moltres. I did not
-apply the hybrid method on these cases. Cases 2a and 2b represent 1-D models of the \gls{MSRE} with
-air-filled control rod guide tube regions, a homogenized fuel-graphite lattice regions, and outer
-vessel regions. Case 2b additionally contains a 0.5cm-thick control rod region to contrast with
-Case 2a for control rod worth calculations. Cases 3a and 3b retain the heterogeneous fuel-graphite
-lattice geometry present in the \gls{MSRE}.
-
-In my preliminary investigations, the original control rod material composition
-(70 wt\% Gd$_2$O$_3$-30 wt\% Al$_2$O$_3$) led to excessively large control rod worth estimates of
-approximately 55000 pcm in the 1-D models, effectively halving the $k_\text{eff}$. The worth
-estimates are ten times larger
-than the 5500 pcm expected from full rod insertion of all three rods in the actual \gls{MSRE}
-\cite{fratoni_molten_2020}. Therefore, I greatly reduced the proportion of Gd$_2$O$_3$ in the
-control rod to 0.35 wt\% for a fairer assessment of rod worth in the 1-D models. This change
-brought the new rod worth estimates down to below 20000 pcm.
-
-\subsection{1-D Neutronics Model Mesh Convergence Tests}
-
-I ran mesh convergence studies based on Case 3b for the neutron diffusion and $S_N$ neutron
-transport methods. Given that the hybrid $S_N$-diffusion method comprises of a combination of those
-methods, I found hybrid simulations were fully spatially resolved on meshes that are
-appropriately resolved for both neutron diffusion and $S_N$ simulations.
-
-\begin{figure}[htb!]
-  \centering
-  \begin{subfigure}[b]{0.48\columnwidth}
-    \centering
-    \includegraphics[width=\columnwidth]{diffusion-mesh-convergence-k}
-    \caption{Neutron diffusion method}
-    \label{fig:diff-mesh-k}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}[b]{0.48\columnwidth}
-    \centering
-    \includegraphics[width=\columnwidth]{sn-mesh-convergence-k}
-    \caption{$S_8$ neutron transport method}
-    \label{fig:sn-mesh-k}
-  \end{subfigure}
-  \caption{Convergence of multiplication factor ($k_\text{eff}$) estimates for Case 3b across five
-    levels of mesh refinement. Higher refinement levels correspond to smaller mesh element sizes.}
-  \label{fig:mesh-k}
-\end{figure}
-
-Figures \ref{fig:diff-mesh-k} and \ref{fig:sn-mesh-k} show the convergence of $k_\text{eff}$
-estimates observed for Case 3b simulations with neutron diffusion and $S_8$ neutron transport
-methods. The $k_\text{eff}$ values plateau at 0.9713203 and 0.9813790 for the neutron diffusion and
-$S_8$ methods, respectively. By mesh refinement level 2, both methods fall within $5\times 10^{-7}$
-range of the $k_\text{eff}$ value at level 4.
-
-Figures \ref{fig:diff-mesh-flux} and \ref{fig:sn-mesh-flux} show the absolute differences in the
-group flux distributions at mesh refinement levels 1 to 3 relative to the reference group flux
-distribution at level 4 (maximum refinement). I omitted data from level 0 to improve the fidelity
-of the plots as flux differences drop significantly by one order of magnitude with each level of
-refinement.
-With these results, I concluded that the mesh resolution at level 2 is a sufficient guideline for
-the mesh resolutions used in the subsequent 1-D analyses. The absolute differences in both
-$k_\text{eff}$ and group flux distributions at level 2 remain below $5\times10^{-7}$ relative to
-level 4.
-
-\begin{figure}[htb!]
-  \centering
-  \includegraphics[width=\columnwidth]{diffusion-mesh-convergence-flux}
-  \caption{Absolute difference in group fluxes ($\phi_g$) from the neutron diffusion method for
-  mesh refinement levels 1 to 4 relative to the reference group flux distributions at mesh
-  refinement level 5.}
-  \label{fig:diff-mesh-flux}
-\end{figure}
-
-\begin{figure}[htb!]
-  \centering
-  \includegraphics[width=\columnwidth]{sn-mesh-convergence-flux}
-  \caption{Absolute difference in group fluxes ($\phi_g$) from the $S_8$ neutron transport method
-  for mesh refinement levels 1 to 4 relative to the reference group flux distributions at mesh
-  refinement level 5.}
-  \label{fig:sn-mesh-flux}
-\end{figure}
-
-\FloatBarrier
-
-\subsection{1-D Neutronics Numerical Results}
-
-\subsubsection{Reactivity}
-
-Figure \ref{fig:1d-rho} shows the difference in reactivities of OpenMC-MG, $S_8$,
-neutron diffusion, and hybrid methods relative to OpenMC-CE. The standard deviations of
-reactivity values from OpenMC-CE are approximately 40 pcm for Cases 1a and 1b and 60 pcm for the
-rest as depicted by the blue highlighted area in Figure \ref{fig:1d-rho}. For Cases 1a and 1b, the
-OpenMC-MG, $S_8$, and neutron diffusion methods show excellent agreement with OpenMC-CE as the
-reactivity values fall within the one or two standard deviation range. The Case 1 results show
-the Moltres and Python implementations are consistent with each other given that they are spatially
-well-resolved.
-
-\begin{figure}[htb!]
+\begin{frame}
+  \frametitle{Hybrid $S_N$-Diffusion Method: 1-D Neutronics Eigenvalue Simulations}
+  \textbf{1-D Neutronics Model Reactivity Results}
+  \begin{figure}[h]
     \centering
@@
   \end{figure}
+\end{frame}
 
-For all cases, the OpenMC-MG and the $S_8$ methods show consistent agreement with one another. 
-While they deviate from OpenMC-CE by approximately 350 pcm for Cases 2a and 3a, they fall within
-two standard deviations of OpenMC-CE for Cases 2b and 3b. When increasing the maximum Legendre
-polynomial order to approximate the angular dependence in $\Sigma_s^{g'\rightarrow g}$ from
-2nd-order to 3rd-order, the reactivity changed by only 1 pcm. Therefore, I attribute the reactivity
-discrepancies to the eight neutron energy group structure which remains the
-only significant difference between OpenMC-CE and the multigroup methods.
-
-The neutron diffusion and hybrid methods agree closely with one another for Cases 2a and 3a which
-exclude the control rod region. This shows that the hybrid method provides similar $k_\text{eff}$
-estimates to the neutron diffusion method in 1-D simulations without highly neutron-absorbing
-regions. Compared to the OpenMC-MG and $S_8$ reactivity values, the neutron diffusion and hybrid
-method reactivity values agree closer with the reference OpenMC-CE value. However, this is likely
-due to favorable error cancellation; Figures \ref{fig:3a-flux} and \ref{fig:3a-flux-diff} show
-OpenMC-MG and the $S_8$ methods reproduce
-the OpenMC-CE flux distribution more accurately than the neutron diffusion and hybrid methods.
-
-In Cases 2b and 3b, the neutron diffusion method largely fails to accurately capture the
-effect of introducing the control rod region as evidenced by the -1500 pcm and -1150 pcm
-discrepancies relative to OpenMC-CE. The hybrid method fares better with -400 pcm and -300 pcm
-discrepancies.
-
-\begin{figure}[htb!]
+\begin{frame}
+  \frametitle{Hybrid $S_N$-Diffusion Method: 1-D Neutronics Eigenvalue Simulations}
+  \textbf{1-D Neutronics Model Control Rod Worth Results}
+  \begin{figure}[h]
     \centering
@@
   \end{figure}
+\end{frame}
 
-\subsubsection{Control rod worth}
-
-Moving the discussion to control rod worths, figure \ref{fig:1d-worth} shows the percentage
-difference in rod worths for Cases 2 and 3 of OpenMC-MG, $S_8$, neutron diffusion, and hybrid
-methods relative to OpenMC-CE. I calculated the rod worth estimates from each method by taking the
-difference in reactivity between the ``rod out'' (a) and ``rod in'' (b) configurations of Cases 2
-and 3. All neutronics methods overestimate the rod worth relative to OpenMC-CE, resulting in
-positive percentage difference values observed in the figure.
-The neutron diffusion methods clearly show up as outliers with rod worth estimates in excess
-of 8\%. Most of the remaining methods cluster around 2.5\% and 3\% percentage difference for Cases
-2 and 3 with the exception of the Python-hybrid method for Case 3.
-
-Given that the hybrid methods rely on transport corrections derived from the $S_N$ method, the
-$S_8$ rod worth estimates serve as the
-reference point for hybrid method verification. Additionally, errors arising from the multigroup
-approximation affect both hybrid and $S_N$ simulations to similar degrees. While the Python-hybrid
-method produces the lowest percentage difference relative to OpenMC-CE for Case 3, the
-Python-hybrid method deviates from the $S_8$ ``multigroup'' reference rod worths more than the
-Moltres-hybrid method. Nevertheless, both hybrid method implementations provide significant
-improvements in rod worth estimates over the neutron diffusion method.
-
-\subsubsection{Neutron flux distribution}
-
-This discussion on the neutron flux distributions focuses on Cases 3a and 3b which most closely
-resemble the actual \gls{MSRE} reactor geometry. Cases 3a and 3b resemble the \gls{MSRE} reactor
-with a control rod withdrawn and inserted, respectively. Figures \ref{fig:3a-flux} and
-\ref{fig:3b-flux} show the neutron group flux distributions for Cases 3a and 3b from OpenMC-CE and
-OpenMC-MG. Differences between the two methods arise from the multigroup approximation, notably in
-group 2 for Case 3a and groups 2, 6, 7, 8 for Case 3b. Future work aimed at finding a better
-neutron energy group structure should at minimum focus on improving fidelity in the listed energy
-groups.
-
-For a fair evaluation of the deterministic multigroup methods without distortions from the
-multigroup approximation, I chose to compare their flux distributions with the OpenMC-MG flux
-distribution. Figures \ref{fig:3a-flux-diff} and \ref{fig:3b-flux-diff} show the absolute
-difference in neutron group flux distributions of the Moltres-$S_8$, Moltres-diffusion,
-Moltres-hybrid, and Python-hybrid methods relative to OpenMC-MG. The figures omit Python-$S_8$ and
-Python-diffusion as their distributions are nearly identical to their Moltres-based counterparts.
-The neutron diffusion and hybrid methods fare worse than the $S_8$ method at capturing the
-oscillatory pattern in groups 1, 2, 5, 7, and 8 arising from the fuel-graphite lattice geometry
-along $x=5$ cm to 10 cm. In both Cases 3a and 3b, the neutron diffusion method (yellow plot lines)
-exhibits significant flux deviations in groups 1, 2, 7, and 8 near $x=0$ cm.
-
-Both hybrid methods
-provide significant improvements in neutron flux distributions compared to the neutron diffusion
-method as shown by the generally smaller flux difference values particularly near $x=0$ cm. The
-figures also show Moltres-hybrid (green) provides more accurate flux estimates than
-Python-hybrid (red) in all instances. The differences indicate deficiencies in the diffusion
-correction scheme in Python-hybrid which lead to poorer performance compared to the drift
-correction scheme in Moltres-hybrid.
-
+\begin{frame}
+%  \frametitle{Hybrid $S_N$-Diffusion Method: 1-D Neutronics Eigenvalue Simulations}
+  \begin{columns}
+    \column{5.5cm}
     \begin{figure}[htb!]
@@
     \end{figure}
-
+    \column{5.5cm}
     \begin{figure}[htb!]
@@
     \end{figure}
+  \end{columns}
+\end{frame}
 
+\begin{frame}
+%  \frametitle{Hybrid $S_N$-Diffusion Method: 1-D Neutronics Eigenvalue Simulations}
+  \begin{columns}
+    \column{5.5cm}
+    \textbf{Case 3a Neutron Flux Distributions}
+    \begin{itemize}
+      \item The neutron diffusion and hybrid methods fare worse than the $S_8$ method at capturing
+        the oscillatory flux pattern.
+      \item The hybrid method performs better than the neutron diffusion method near $x=0$ cm where
+        the correction region is situated.
+    \end{itemize}
+    \column{5.5cm}
     \begin{figure}[htb!]
       \centering
-  \includegraphics[width=\columnwidth]{case-3a-flux-diff}
+      \includegraphics[width=.975\columnwidth]{case-3a-flux-diff}
       \caption{Absolute difference in neutron group flux distributions for Case 3a from Moltres-$S_8$,
@@
     \end{figure}
+  \end{columns}
+\end{frame}
 
+\begin{frame}
+%  \frametitle{Hybrid $S_N$-Diffusion Method: 1-D Neutronics Eigenvalue Simulations}
+  \begin{columns}
+    \column{5.5cm}
+    \textbf{Case 3b Neutron Flux Distributions}
+    \begin{itemize}
+      \item The neutron diffusion method performance worsens with the control rod region introduced
+        near $x=0$.
+    \end{itemize}
+    \column{5.5cm}
     \begin{figure}[htb!]
       \centering
-  \includegraphics[width=\columnwidth]{case-3b-flux-diff}
+      \includegraphics[width=.975\columnwidth]{case-3b-flux-diff}
       \caption{Absolute difference in neutron group flux distributions for Case 3b from Moltres-$S_8$,
@@
     \end{figure}
+  \end{columns}
+\end{frame}
 
-\FloatBarrier
-
-\subsubsection{Drift \& diffusion correction distribution}
-
-Figures \ref{fig:3b-drift} and \ref{fig:3b-svdc} show the distributions of the drift and diffusion
-correction parameters, respectively, in each neutron energy group for Case 3b. As a reminder to the
-reader, the hybrid $S_N$-diffusion methods generate these correction parameters in subregions
-containing highly neutron absorbing materials. I implemented the drift correction scheme for
-Moltres-hybrid and the diffusion correction scheme for Python-hybrid. The hybrid methods
-selectively apply the $S_N$ method to improve neutronic modeling within the subregions. I computed
-the reference distributions (yellow data points) in both plots from reference flux solutions of the
-standard $S_8$ method.
-
-As discussed in Section \ref{sec:buffer-region}, both sets of correction parameters generated by
-the two correction schemes match their respective reference values within the correction subregion
-from $x=0$ cm to 10 cm except near the subregion boundary at $x=10$ cm. Therefore, both schemes
-provide improved flux estimates in the subregion compared to the neutron diffusion method.
-The prior reactivity and flux distribution results indicate Moltres-hybrid is more accurate
-than Python-hybrid relative to OpenMC-MG as the reference.
-
-In Figure \ref{fig:3b-drift}, the drift
-correction distributions (Moltres-hybrid) for all eight energy groups are well-defined throughout
-the subregion and are continuous except at material interfaces. Consequently, the drift correction
-scheme provides accurate flux corrections for about 75\% of the subregion based on the buffer
-region cutoff criteria which occurs around $x=7.5$ cm where most of the drift distributions are
-zero (refer to Section \ref{sec:buffer-region}). By contrast, the diffusion correction
-distributions in Figure \ref{fig:3b-svdc} contains numerous discontinuities tending towards
-positive or negative infinity in all eight energy groups. The discontinuities arise when
-computing corrected diffusion coefficients (Eq. \ref{eq:svdc}) near flux peaks and troughs
-(stationary points) where the flux gradients approach zero. The location of stationary points from
-the diffusion and $S_N$ calculations in the hybrid method do not necessarily coincide, resulting in
-discontinuous flux gradients in the final flux solution. Discontinuities are generally unavoidable
-in most problems because the fluxes typically reach local minima in control rod regions.
-Furthermore, unlike negative drift coefficients, negative diffusion coefficients are unphysical.
-Taking everything into consideration, the diffusion correction scheme is less accurate and more
-numerically unstable than the drift correction scheme.
-
-The subsequent discussions on the correction region size and the number of outer
-iterations focus on the drift correction scheme of the hybrid method.
-
-\begin{figure}[htb!]
-  \centering
-  \includegraphics[width=\columnwidth]{case-3b-drift}
-  \caption{Multigroup drift correction ($\vec{D}_g$) $x$-component distributions from the
-  Moltres-hybrid and Moltres-$S_8$ solvers.}
-  \label{fig:3b-drift}
-\end{figure}
-
-\begin{figure}[htb!]
-  \centering
-  \includegraphics[width=\columnwidth]{case-3b-svdc}
-  \caption{Multigroup diffusion correction ($D^s_g$) $x$-component distributions from the
-    Python-hybrid and Python-$S_8$ solvers.}
-  \label{fig:3b-svdc}
-\end{figure}
-
-\FloatBarrier
-
-\subsubsection{Impact of correction subregion sizes}
-
-The hybrid $S_N$-diffusion method relies on $S_N$ calculations in the correction subregion to
-generate flux corrections. Minimizing the size of the correction subregion and the $S_N$ subproblem
-is essential for the hybrid method to be computationally competitive for time-dependent full-core
-simulations. I investigated the effect of the correction subregion size on the $k_\text{eff}$ and
-control rod worth estimates with Cases 3a and 3b by varying the subregion sizes from 10 cm to 40 cm
-at 5 cm-intervals.
-
-Figures \ref{fig:v1-size-a-k} and \ref{fig:v1-size-b-k} show the $k_\text{eff}$ estimates from the
-hybrid method for Cases 3a and 3b, respectively. In both cases, the $k_\text{eff}$ values initially
-decrease as the correction subregion sizes increase before reversing in trend when the subregion
-size reaches 35 cm and beyond. The $k_\text{eff}$ values vary by up to 164 pcm for Case 3a and 109
-pcm for Case 3b. An important observation here is that the $k_\text{eff}$ values do not
-monotonically converge towards the $k_\text{eff}$ estimate from the $S_8$ method. The data implies
-other significant sources of discrepancies exist beyond those being corrected in the
-correction subregion by the hybrid method.
-
-% In general,
-% discrepancies arise between the neutron diffusion method and $S_8$ from the neutron diffusion 
-% method's failure to accurately model the neutron flux in the fuel-graphite lattice region. The
-% figures imply that those discrepancies result in a positive change in $k_\text{eff}$ since the
-% $k_\text{eff}$ decreases as larger fractions of the problem domain benefit from flux corrections.
-% The remaining discrepancy in $k_\text{eff}$ between the hybrid method with a correction subregion
-% size of 40 cm and the $S_8$ method arises from
-
-\begin{figure}[htb!]
+\begin{frame}
+  \frametitle{Hybrid $S_N$-Diffusion Method: 1-D Neutronics Eigenvalue Simulations}
+  \textbf{Impact of Correction Subregion Sizes on $k$}
+  \begin{itemize}
+    \item Minimizing the correction region size is essential for the hybrid method to be
+      computationally efficient for time-dependent simulations.
+    \item $k$ varies by up to 164 pcm for Case 3a and 109 pcm for Case 3b.
+    \item The hybrid method $k$ value does not converge monotonically towards the $S_8$ method $k$ value,
+      implying other sources of discrepancies.
+  \end{itemize}
+  \begin{figure}[h]
     \centering
@@
   \end{figure}
+\end{frame}
 
-Figure \ref{fig:v1-size-rho} shows the percentage difference in rod worth relative to OpenMC-CE. 
-Due to the identical trends observed in the $k_\text{eff}$ estimates of both cases, the control rod
-worth estimates do not change significantly when the correction subregion
-sizes change. This indicates limiting the correction subregion size to save
-computational cost on the expensive $S_N$ calculations has a negligible impact on rod worth
-estimates. The rod worth estimates for all investigated correction subregion sizes remain within
-0.2\% of the $S_8$ method.
-
+\begin{frame}
+  \frametitle{Hybrid $S_N$-Diffusion Method: 1-D Neutronics Eigenvalue Simulations}
+  \textbf{Impact of Correction Subregion Sizes on Rod Worth}
+  \begin{itemize}
+    \item Rod worth estimates vary non-monotonically with increasing correction subregion size.
+    \item Rod worth estimates remain within 0.2 \% of the $S_8$ method rod worth.
+    \item The hybrid method produces accurate rod worth estimates as long as the correction region
+      size is kept consistent.
+  \end{itemize}
   \begin{figure}[htb!]
     \centering
-  \includegraphics[width=0.8\columnwidth]{correction-size-rho}
+    \includegraphics[width=0.6\columnwidth]{correction-size-rho}
     \caption{Percentage difference in rod worth from the hybrid method relative to OpenMC-CE for
@@
   \end{figure}
+\end{frame}
 
-The $k_\text{eff}$ and rod worth results show the hybrid method is able to accurately
-reproduce rod worth estimates as long as the correction subregion size remains consistent. The
-correction subregion size must also be sufficiently large to capture the highly neutron-absorptive
-rod's influence on the neutron flux around the rod at $x=0$ cm. The drift correction parameter
-distributions in Figure \ref{fig:3b-drift} indicate that the influence of the rod on the drift, and
-consequently the flux, extend to approximately $x=5$ cm before the drift distribution settles on a
-regular repeating pattern influenced by the fuel-graphite lattice structure. I applied a consistent
-correction subregion size of 10 cm for Cases 2 and 3 since it sufficiently meets the requirements
-discussed while also minimizing the computational cost of the $S_8$ subsolver. Other reactor
-designs consisting of different material components may require different minimum correction
-subregion sizes.
+\begin{frame}
+  \frametitle{Hybrid $S_N$-Diffusion Method: 1-D Neutronics Eigenvalue Simulations}
+  \textbf{Relaxing the $S_N$ Convergence Tolerance}
+  \begin{itemize}
+    \item The hybrid $S_N$-diffusion method iteratively couples the neutron diffusion solver with
+      the $S_N$ subsolver.
+    \item Diffusion-based acceleration schemes for neutron transport methods typically do not
+      require the neutron transport calculations to fully converge.
+    \item Transport correction parameters tend to converge faster than scalar flux.
+    \item Relaxing the $S_N$ subsolver convergence tolerance would provide computational savings to
+      the hybrid method.
+  \end{itemize}
+\end{frame}
 
-\subsubsection{Relaxing the $S_N$ subsolver convergence threshold}
-
-As described in Section \ref{sec:hybrid-algorithm}, the hybrid $S_N$-diffusion method works by
-iteratively coupling a neutron diffusion solver and a $S_N$ subsolver with the latter generating
-drift correction parameters for the former. Excluding the initial neutron diffusion calculation,
-each outer (fixed point) iteration comprises of an inner $S_N$ calculation on the correction
-subregion and an inner neutron diffusion calculation on the entire domain. The \gls{FEM} algorithm
-in Moltres determines an individual solver reaches convergence when the combined
-residual contributions fall below a specified convergence tolerance threshold. The fixed point
-iteration scheme under the \texttt{MultiApps} system in Moltres for coupling multiple individual
-solvers (referred to as apps) requires users to specify a main app for the final convergence check
-after each outer iteration to terminate the iterations. The scheme also requires users to set a
-fixed point convergence tolerance value independent of the tolerance values of each individual
-solver. Naturally, the fixed point tolerance value must be equal to or greater than the tolerance
-value of the main individual solver for the possibility of overall convergence. The hybrid method
-implementation in Moltres sets the neutron diffusion solver as the main app and the $S_N$ subsolver
-as the sub app.
-
-The hybrid method shares some similarities with diffusion-based \gls{HOLO} acceleration schemes
-(refer to Section \ref{sec:drift-correction}) for the neutron transport methods which do not
-require the neutron transport side of the calculations to fully converge
-\cite{reynolds_analysis_2023, wang_diffusion_2014}.
-The typical convergence threshold value required for converged neutron flux calculations in Moltres
-is $10^{-8}$; I set the convergence tolerance value to $10^{-8}$ for the neutron diffusion solver
-and the fixed point scheme. I varied the convergence tolerance value for the $S_N$ subsolver to 
-find the optimal value for maintaining sufficient solution accuracy and minimizing computational
-costs; tightening the $S_N$ convergence tolerance value generally results in more outer iterations
-performed before the solver meets the fixed point convergence tolerance.
-Table \ref{table:sn-tol} lists the number of outer iterations required by the hybrid method for
-a given set of $S_8$ subsolver convergence tolerance values in Case 3b.
-
-\begin{table}[htb]
+\begin{frame}
+  \frametitle{Hybrid $S_N$-Diffusion Method: 1-D Neutronics Eigenvalue Simulations}
+  \begin{columns}
+    \column{7cm}
+    \textbf{Relaxing the $S_N$ Convergence Tolerance}
+    \begin{itemize}
+      \item The number of outer iterations generally increases with the $S_8$ convergence tolerance.
+      \item The hybrid method exhibits superlinear ($q=1.333$) convergence in $k$ with respect to
+        the $S_8$ convergence tolerance value.
+    \end{itemize}
+    \begin{table}[h]
       \centering
@@
       convergence tolerance values imposed on the $S_8$ subsolver.}
+      \small
+      \setlength\tabcolsep{2pt}
       \begin{tabular}{l S S S S S S}
@@
     \end{table}
-
-Figure \ref{fig:sn-tol} shows the $k_\text{eff}$ error estimates relative to the reference value
-when $\epsilon_\text{tol}=10^{-8}$. The hybrid method exhibits superlinear convergence ($q=1.333$)
-with respect to the $S_N$ subsolver convergence tolerance value. Based on the results, setting
-$\epsilon_\text{tol}$ to $10^{-5}$ is the best balance between accuracy and computational
-cost because the $k_\text{eff}$ error is less than 0.01 pcm and the number of outer iterations
-would increase from 2 to 3 if the tolerance value is further tightened.
-
-\begin{figure}[htb!]
+    \column{4cm}
+    \begin{figure}[h]
       \centering
-  \includegraphics[width=0.6\columnwidth]{sn-tol}
+      \includegraphics[width=\columnwidth]{sn-tol}
       \caption{$k_\text{eff}$ error estimates of Case 3b for a range of convergence tolerance values
       imposed on the $S_8$ subsolver relative to the reference $k_\text{eff}$ value when
-  $\epsilon_\text{tol}=10^{-8}$. The $q=1.333$ line represents the approximate rate of
-  convergence.}
+      $\epsilon_\text{tol}=10^{-8}$.}
       \label{fig:sn-tol}
     \end{figure}
-
-\section{2-D Neutronics Models \& Results} \label{sec:2d-results}
-
-\subsection{2-D Neutronics Model Setup} \label{sec:2d-model-setup}
-
-Figures \ref{fig:1/4-geom} and \ref{fig:full-geom} show the 2-D quarter-core and full-core
-\gls{MSRE} models investigated. I adapted the models from the horizontal cross section of the base
-\gls{MSRE} benchmark model developed by \cite{fratoni_molten_2020} with the following differences
-in the 2-D \gls{MSRE} model for this work:
-
-\begin{itemize}
-  \item Ignored thermal expansion of all reactor components.
-  \item Modeled the control rods as perfectly annular cylinders instead of beaded control elements
-    strung on flexible steel cables. Ignored the inconel shells encasing the control elements and
-    the steel helical cables linking the control elements together.
-  \item Set channel pitch, width, and length to 2, 0.2, and 1.2 inches, respectively.
-  \item Homogenized the INOR-8, graphite, and molten salt regions in the sample basket.
-  \item Modified the outer edge of the core graphite to be circular with no jagged edges.
-  \item Replaced partial channels with full-size channels near the edge of the core graphite.
-  \item Ignored the thermal insulation and shield structures outside the core.
-\end{itemize}
-
-The differences listed above, along with other changes in the 3-D model, result in negligible
-changes to the $k_\text{eff}$ (refer to Section 5.6) relative to the \gls{MSRE} benchmark model
-\cite{fratoni_molten_2020}.
-
-\begin{figure}[htb!]
-  \centering
-  \includegraphics[width=\columnwidth]{quarter-core-geom}
-  \caption{2-D \gls{MSRE} quarter-core model based on the horizontal cross section of the actual
-  \gls{MSRE} geometry.}
-  \label{fig:1/4-geom}
-\end{figure}
-
-\begin{figure}[htb!]
-  \centering
-  \includegraphics[width=\columnwidth]{full-core-geom}
-  \caption{2-D \gls{MSRE} full-core model based on the horizontal cross section of the actual
-  \gls{MSRE} geometry.}
-  \label{fig:full-geom}
-\end{figure}
-
-\begin{table}[htb]
-  \small
-  \centering
-  \setlength\tabcolsep{4pt}
-  \caption{\gls{MSRE} molten salt composition when the $^{235}$U loading was at 65.25 kg.}
-  \begin{tabular}{l S S S S S S S}
-    \toprule
-    Element & {Li} & {Be} & {Zr} & {Hf} & {$^{234}$U} & {$^{235}$U} & {$^{236}$U} \\
-    \midrule
-    Composition [kg] & 507.27 & 293.96 & 513.97 & 0.0029 & 0.67 & 65.25 & 0.27 \\
-    \bottomrule
-  \end{tabular}
-  \begin{tabular}{l S S S S S S}
-    \toprule
-    Element & {$^{238}$U} & {Fe} & {Cr} & {Ni} & {O} & {F} \\
-    \midrule
-    Composition [kg] & 141.91 & 0.75 & 0.13 & 0.14 & 2.27 & 3103.22 \\
-    \bottomrule
-  \end{tabular}
-  \label{table:2d-salt-composition}
-\end{table}
-
-INOR-8, also referred to as Hastelloy N, is a nickel-based alloy used for structural components in
-the \gls{MSRE} \cite{robertson_msre_1965}.
-Table \ref{table:2d-salt-composition} lists the molten salt composition for the 2-D models. The
-$^{235}$U loading of 65.25 kg is the amount at which the \gls{MSRE} first achieved criticality. The
-control rods consist of the original \gls{MSRE} 70 wt\% Gd$_2$O$_3$-30 wt\% Al$_2$O$_3$ mixture.
-The models impose neutron vacuum boundary conditions on the outermost boundary of the INOR-8
-vessel structure. The quarter-core model imposes neutron reflecting boundary conditions along the
-left and bottom straight edges (Figure \ref{fig:1/4-geom}). Note that the quarter-core and
-full-core models are not identical because the full-core model contains a sample basket instead of
-a control rod in the lower-right segment of Figure \ref{fig:full-geom-closeup}.
-
-\begin{figure}[htb!]
-  \centering
-  \includegraphics[width=0.7\columnwidth]{full-core-closeup}
-  \caption{Detailed view of the control rod thimbles and sample basket in 2-D \gls{MSRE} full-core
-    model.}
-  \label{fig:full-geom-closeup}
-\end{figure}
-
-I ran neutronics simulations on the quarter- and full-core models on OpenMC under both continuous
-energy (OpenMC-CE) and multigroup (OpenMC-MG) modes, and on Moltres with the hybrid $S_N$-diffusion
-and neutron diffusion methods. The hybrid method ran with the $S_8$ angular discretization scheme
-and up to 2nd-order scattering approximations for the $S_N$ subsolver. I simulated the
-quarter-core model with and without the control rod in the air-filled rod thimble; air replaces the
-control rod material when the rod is removed. For the full-core model, I considered four rod
-configurations: no rods present, rod 1 present, rods 1 \& 2 present, and rods 1, 2 \& present.
-Other than providing a close-up view of the core center, Figure \ref{fig:full-geom-closeup} also
-presents the correction subregion on which the $S_8$ subsolver of the hybrid method generates
-drift correction parameters for the coupled diffusion solver.
-
-\FloatBarrier
-
-\subsection{2-D Neutronics Numerical Results} \label{sec:2d-nts-results}
-
-\subsubsection{2-D Quarter-Core}
-
-Table \ref{table:quarter-core} details the $k_\text{eff}$ and control rod worth estimates and error
-values relative to OpenMC-CE. For the ``No Rod'' case, OpenMC-CE estimated the $k_\text{eff}$ to be
-1.11209(43). The OpenMC-MG, hybrid, and neutron diffusion methods performed
-similarly to one another with reactivity errors ranging from 618 pcm to 773 pcm relative to
-OpenMC-CE. Based on $k_\text{eff}$ alone, the hybrid method performed marginally worse than the
-neutron diffusion method. However, for the ``Rod'' case, then neutron diffusion method saw a
-change in sign of the error to -816 pcm. The errors for OpenMC-MG and the hybrid method
-remained positive and reduced to 446 pcm and 760 pcm, respectively. Consequently, when taking the
-difference in reactivities to estimate the rod worth, the neutron diffusion method performed
-significantly worse than OpenMC-MG and the hybrid method. OpenMC-CE estimated the rod worth to be
-8370(53) pcm. The neutron diffusion method overestimated the rod worth by 1484 pcm. The hybrid
-method improved significantly on the neutron diffusion method with a rod worth estimate that is
-only 13 pcm
-higher than OpenMC-CE. OpenMC-MG reported a larger error of 172 pcm likely due to favorable error
-cancellation with the hybrid method and statistical uncertainties with the Monte Carlo methods.
-The OpenMC-CE and OpenMC-MG rod worth estimates agree within two standard deviations of each other.
-
-\begin{table}[htb]
-  \small
-  \centering
-  \caption{$k_\text{eff}$ and control rod worth estimates for the 2-D quarter-core \gls{MSRE}
-    model. Error values are relative to OpenMC-CE.}
-  \begin{tabular}{l S[table-format=1.5(2)] S S[table-format=1.5(2)] S S[table-format=4(2)] S}
-    \toprule
-    \multirow{2}{*}{Method} & \multicolumn{2}{c}{No Rod} & \multicolumn{2}{c}{Rod} & \multicolumn{2}{c}{Rod worth} \\
-                            & {$k_\text{eff}$} & {Error [pcm]} & {$k_\text{eff}$} & {Error [pcm]} & {$\Delta\rho_\text{worth}$ [pcm]} & {Error [pcm]} \\
-                            \cmidrule(r){1-1} \cmidrule(rl){2-3} \cmidrule(rl){4-5} \cmidrule(l){6-7}
-	  OpenMC-CE & 1.11209(43) & {-} & 1.01740(42) & {-} & 8370(53) & {-} \\
-	  OpenMC-MG & 1.11979(42) & 618 & 1.02204(41) & 446 & 8541(51) & 172 \\
-      Diffusion & 1.12059 & 682 & 1.00903 & -816 & 9867 & 1484 \\
-      Hybrid & 1.12174 & 773 & 1.02532 & 760 & 8383 & 13 \\
-    \bottomrule
-  \end{tabular}
-  \label{table:quarter-core}
-\end{table}
-
-\begin{figure}[htb!]
-  \centering
-  \includegraphics[width=\columnwidth]{msre-quarter-no-rod-power}
-  \caption{Normalized channel fission rate distribution of the 2-D \gls{MSRE} quarter-core model.
-  This figure depicts the 1/8th-core distribution due to symmetry about the diagonal.}
-  \label{fig:1/4-no-rod}
-\end{figure}
-
-\begin{figure}[htb!]
-  \centering
-  \includegraphics[width=\columnwidth]{msre-quarter-rod-power}
-  \caption{Normalized channel fission rate distribution of the 2-D \gls{MSRE} quarter-core model.
-  This figure depicts the 1/8th-core distribution due to symmetry about the diagonal.}
-  \label{fig:1/4-rod}
-\end{figure}
-
-\FloatBarrier
-
-I also calculated the normalized channel fission rate distribution by integrating the fission rate
-within each fuel salt channel and normalizing the values such that the integrated fission rate in
-each channel averages to 1 as follows
-%
-\begin{align}
-  \text{Normalized fission rate of channel }i &= \frac{\text{Fission rate in channel }i}{
-  \text{Total fission rate}}\cdot\frac{\text{Total number of channels}}{
-    \text{Total fission rate in channels}} \nonumber \\
-                                              &= \frac{\sum^G_{g=1}\int_{V_i}\Sigma_{f,g}
-  \phi_g(\vec{r})}{\sum^G_{g=1}\int_V\Sigma_{f,g}\phi_g(\vec{r})}\cdot\frac{N_i}{
-  \sum^{N_i}_{i=1}\sum^G_{g=1}\int_{V_i}\Sigma_{f,g}\phi_g(\vec{r})}.
-\end{align}
-%
-The total fission rate includes fission rate contributions from peripheral fuel salt regions.
-Figures \ref{fig:1/4-no-rod} and \ref{fig:1/4-rod} show the normalized fission rate distributions
-of the quarter-core model for the ``No Rod'' and ``Rod'' cases, respectively. I presented 1/8th
-slices of the core in the figures because the distributions are symmetrical about the $y=x$ line
-outside of statistical variances for OpenMC-CE. I combined the midpoints between the centroids of
-neighboring fuel salt channels to form the diamond lattices shown in the figures.
-Each diamond element in the figures corresponds a horizontally or vertically oriented channel
-depicted in Figure \ref{fig:1/4-geom}. The lattice also provides a guide for subdividing the
-annular fuel regions around the control rod thimbles. The fission rate
-distributions omit four channels along the outer edge due to the proximity to peripheral annular
-fuel regions and the consequent difficulty in accurately tallying fission rates.
-
-In the ``No Rod'' case, the peak channel fission rate occured several channels away from the
-physical center of the reactor core. The presence of the empty (air-filled) rod thimble depressed
-fission rates near the center. The neutron diffusion and hybrid methods report similar percentage
-errors relative to OpenMC-CE except near the center. While the hybrid method benefits from drift
-corrections, the neutron diffusion method's reliance on scalar diffusion coefficients tends to
-artificially flatten neutron fluxes in near-void regions. Therefore, the largest percentage errors
-from the neutron diffusion method occurred in the second and third diamond elements
-diagonally up from the bottom which represent the annular fuel region surrounding the rod thimble.
-In the ``Rod'' case, the introduction of the control rod shifted the peak further away from the
-center as expected of a highly neutron-absorbing material. The rod further worsened fission rate
-errors for the neutron diffusion method with large errors of 1\% or more at the core center and
-periphery. On the other hand, the fission rate errors for the hybrid method largely remained below
-1\%.
-
-\begin{table}[htb]
-  \small
-  \centering
-  \caption{Absolute mean and maximum percentage errors in the normalized channel fission rates of
-  the 2-D \gls{MSRE} quarter-core models relative to OpenMC. The mean relative standard deviation of
-  OpenMC normalized channel fission rates is 0.20\%.}
-  \begin{tabular}{l S S S S}
-    \toprule
-    \multirow{2}{*}{Method} & \multicolumn{2}{c}{No Rod} & \multicolumn{2}{c}{Rod} \\
-                            & {Mean [\%]} & {Maximum [\%]} & {Mean [\%]} & {Maximum [\%]} \\
-                            \cmidrule(r){1-1} \cmidrule(rl){2-3} \cmidrule(l){4-5}
-    Diffusion & 0.40 & 2.63 & 2.01 & 17.44 \\
-    Hybrid & 0.40 & 1.32 & 0.43 & 3.08 \\
-    \bottomrule
-  \end{tabular}
-  \label{table:quarter-core-power}
-\end{table}
-
-Table \ref{table:quarter-core-power} shows the absolute mean and maximum percentage errors of the
-normalized channel fission rates from the neutron diffusion and hybrid methods relative to
-OpenMC-CE. Overall, the hybrid method reported smaller percentage errors than the neutron diffusion
-method. The absolute maximum percentage error from the neutron diffusion method jumped to 17.44\%
-as compared to the 3.08\% for the hybrid method in the ``Rod'' case. Along with the figures, these
-results demonstrate the hybrid method provides improved fission rate distribution accuracy in
-the existence of near-void and highly neutron-absorbing regions.
-
-\FloatBarrier
-
-\subsubsection{2-D Full-Core}
-
-I performed the same $k_\text{eff}$ and fission rate distribution analyses with the 2-D full-core
-model under four configurations starting from no rod inserted to all three rods inserted. The
-$k_\text{eff}$ and error estimates in Table \ref{table:full-core-k} show the same trends observed
-with the 2-D quarter-core model. OpenMC-MG and the hybrid method exhibit moderate but consistent
-reactivity errors of approximately the 550 pcm and 800 pcm, respectively. The neutron diffusion
-method reported errors ranging from 692 pcm when no rod is inserted to -470 pcm when all three rods
-are inserted. Consequently, in Table \ref{table:full-core-worth}, OpenMC-MG and the hybrid method
-report accurate control rod worths within 1-3 standard deviations of OpenMC-CE while the neutron
-diffusion method increasingly overestimates the rod worth as more rods are inserted.
-
-\begin{table}[htb]
-  \small
-  \centering
-  \caption{$k_\text{eff}$ estimates for the 2-D full-core \gls{MSRE} model with the indicated rods
-    inserted. Error values are relative to OpenMC-CE.}
-  \setlength\tabcolsep{2pt}
-  \begin{tabular}{l S[table-format=1.5(2)] S S[table-format=1.5(2)] S S[table-format=1.5(2)] S S[table-format=1.5(2)] S}
-    \toprule
-    \multirow{3}{*}{Method} & \multicolumn{2}{c}{No Rod} & \multicolumn{2}{c}{Rod 1} & \multicolumn{2}{c}{Rod 1 \& 2} & \multicolumn{2}{c}{Rod 1, 2 \& 3} \\
-                            & {\multirow{2}{*}{$k_\text{eff}$}} & {Error} & {\multirow{2}{*}{$k_\text{eff}$}} & {Error} & {\multirow{2}{*}{$k_\text{eff}$}} & {Error} & {\multirow{2}{*}{$k_\text{eff}$}} & {Error} \\
-                            & & {[pcm]} & & {[pcm]} & & {[pcm]} & & {[pcm]} \\
-                            \cmidrule(r){1-1} \cmidrule(rl){2-3} \cmidrule(rl){4-5} \cmidrule(rl){6-7} \cmidrule(l){8-9}
-    OpenMC-CE & 1.09859(20) & {-} & 1.06980(21) & {-} & 1.04690(18) & {-} & 1.02688(18) & {-} \\
-    OpenMC-MG & 1.10637(19) & 640 & 1.07633(20) & 567 & 1.05235(20) & 495 & 1.03262(17) & 541 \\
-    Diffusion & 1.10701 & 692 & 1.07120 & 122 & 1.04415 & -252 & 1.02195 & -470 \\
-    Hybrid & 1.10843 & 808 & 1.07906 & 802 & 1.05553 & 781 & 1.03582 & 841 \\
-    \bottomrule
-  \end{tabular}
-  \label{table:full-core-k}
-\end{table}
-
-\begin{table}[htb]
-  \small
-  \centering
-  \caption{Control rod worth estimates for the 2-D full-core \gls{MSRE} with the
-  indicated rods inserted. Error values are relative to OpenMC-CE.}
-  \setlength\tabcolsep{5pt}
-  \begin{tabular}{l S[table-format=4(2)] S S[table-format=4(2)] S S[table-format=4(2)] S}
-    \toprule
-    \multirow{2}{*}{Method} & \multicolumn{2}{c}{Rod 1} & \multicolumn{2}{c}{Rod 1 \& 2} & \multicolumn{2}{c}{Rod 1, 2 \& 3} \\
-                            & {$\Delta\rho_\text{worth}$ [pcm]} & {Error [pcm]} & {$\Delta\rho_\text{worth}$ [pcm]} & {Error [pcm]} & {$\Delta\rho_\text{worth}$ [pcm]} & {Error [pcm]} \\
-                            \cmidrule(r){1-1} \cmidrule(rl){2-3} \cmidrule(rl){4-5} \cmidrule(l){6-7}
-    OpenMC-CE & 2450(25) & {-} & 4494(23) & {-} & 6357(24) & {-} \\
-    OpenMC-MG & 2523(23) & 73 & 4640(24) & 146 & 6455(22) & 98 \\
-    Diffusion & 3019 & 569 & 5439 & 945 & 7519 & 1162 \\
-    Hybrid & 2455 & 5 & 4521 & 27 & 6323 & -34 \\
-    \bottomrule
-  \end{tabular}
-  \label{table:full-core-worth}
-\end{table}
-
-Plots of the normalized channel fission rate distributions for all four rod configurations are
-available in Appendix \ref{chap:2d-fiss-rate}. They are best viewed in digital pdf format with
-magnification due to the numerous fuel salt channels and small fonts.
-
-Table \ref{table:full-core-power} shows the absolute mean and maximum percentage errors in the
-normalized channel fission rates from the neutron diffusion and hybrid methods relative to
-OpenMC-CE. Both mean and maximum percentage errors from the neutron diffusion method steadily
-increase with the number of rods inserted. The mean percentage error from the hybrid method remains
-constant at 0.43\% while the maximum percentage error increases marginally from 1.45\% to 2.52\%
-when all three rods are inserted. These results demonstrate the hybrid method remains
-effective at improving control rod worth and fission rate distribution estimates in less symmetric
-geometries.
-
-\begin{table}[htb]
-  \footnotesize
-  \centering
-  \caption{Absolute mean and maximum percentage errors in the normalized channel fission rates of
-  the 2-D \gls{MSRE} full-core models relative to OpenMC. The mean relative standard deviation of
-  OpenMC normalized channel fission rates is 0.27\%.}
-  \setlength\tabcolsep{2.5pt}
-  \begin{tabular}{l S S S S S S S S}
-    \toprule
-    \multirow{2}{*}{Method} & \multicolumn{2}{c}{No Rod} & \multicolumn{2}{c}{Rod 1} & \multicolumn{2}{c}{Rod 1 \& 2} & \multicolumn{2}{c}{Rod 1, 2 \& 3} \\
-                            & {Mean [\%]} & {Maximum [\%]} & {Mean [\%]} & {Maximum [\%]} & {Mean [\%]} & {Maximum [\%]} & {Mean [\%]} & {Maximum [\%]} \\
-                            \cmidrule(r){1-1} \cmidrule(rl){2-3} \cmidrule(rl){4-5} \cmidrule(rl){6-7} \cmidrule(l){8-9}
-    Diffusion & 0.45 & 2.95 & 0.94 & 12.61 & 1.35 & 15.34 & 1.67 & 17.09 \\
-    Hybrid & 0.43 & 1.45 & 0.43 & 1.82 & 0.43 & 2.26 & 0.43 & 2.52 \\
-    \bottomrule
-  \end{tabular}
-  \label{table:full-core-power}
-\end{table}
-
-\FloatBarrier
-
-\section{3-D Neutronics Models \& Results} \label{sec:3d-results}
-
-\subsection{3-D Neutronics Model Setup} \label{sec:3d-model-setup}
-
-Figure \ref{fig:msre-picture} shows the vertical cross section of the actual \gls{MSRE} vessel.
-The actual design has rounded upper and lower plena to help direct salt flow through the system.
-This work uses a modified 3-D numerical \gls{MSRE} model as shown by the vertical cross section
-in Figure \ref{fig:msre-geom-vert}. The model excludes the toroidal flow distributor near the top
-of the vessel and reshapes the plena into right cylinders. It has the same horizontal cross section
-as the 2-D model (Figure \ref{fig:full-geom}) throughout most of the salt-graphite lattice. In
-addition to modifications in the 2-D model listed in Section \ref{sec:2d-model-setup}, the
-3-D model includes the following modifications:
-
-\begin{itemize}
-  \item Removed the toroidal flow distributor near the top of the vessel.
-  \item Reshaped the upper and lower plena into right cylinders while keeping the salt volume
-    constant.
-  \item Flattened the conical tops of the graphite lattice blocks.
-  \item Regularized the complicated graphite lattice structure at the bottom.
-  \item Set a fixed outer vessel thickness of 2.54 cm.
-  \item Homogenized the lower plenum by 90.8\% salt- 9.2\% INOR-8 (vessel) by volume.
-  \item Removed the reactor access nozzle and the various components within it.
-  \item Added a graphite block at the bottom of the control rod thimbles.
-\end{itemize}
-
-\begin{figure}[p]
-  \centering
-  \includegraphics[width=0.97\columnwidth]{msre-picture}
-  \caption{Vertical cross section of the actual \gls{MSRE} vessel. It has rounded upper and lower
-  plena, and a toroidal flow distributor that distributes the salt inflow evenly as it enters the
-  core through holes on the inside vessel surface of the distributor.}
-  \label{fig:msre-picture}
-\end{figure}
-
-\begin{figure}[p]
-  \centering
-  \includegraphics[width=\columnwidth]{msre-geom-vert}
-  \caption{Vertical cross section of the 3-D numerical \gls{MSRE} model offset by 5.08 cm to show
-  the control rod thimble and homogenized sample basket. The control rod is in its fully inserted
-  position. The homogenized lower plenum consists of 90.8\% salt and 9.2\% INOR-8 alloy by volume.}
-  \label{fig:msre-geom-vert}
-\end{figure}
-
-Table \ref{table:reactor-dimensions} lists relevant dimensions of the 3-D \gls{MSRE} model.
-All material compositions remain
-the same as the 2-D models and the \gls{MSRE} benchmark specifications \cite{fratoni_molten_2020}.
-All 3-D neutronics simulations in this section used the molten salt composition with 65.25 kg
-$^{235}$U loaded (Table \ref{table:2d-salt-composition}), except the simulations for temperature
-reactivity coefficient calculations which had a higher $^{235}$U loading of 67.00 kg (Table
-\ref{table:67-salt-composition}).
-
-\begin{table}[htb]
-  \centering
-  \caption{List of reactor dimensions.}
-  \begin{tabular}{l l}
-    Height [cm] & 215 \\
-    TODO: Populate table & \\
-  \label{table:reactor-dimensions}
-\end{table}
-
-\begin{table}[htb]
-  \small
-  \centering
-  \setlength\tabcolsep{4pt}
-  \caption{\gls{MSRE} molten salt composition when the $^{235}$U loading was at 65.25 kg.}
-  \begin{tabular}{l S S S S S S S}
-    \toprule
-    Element & {Li} & {Be} & {Zr} & {Hf} & {$^{234}$U} & {$^{235}$U} & {$^{236}$U} \\
-    \midrule
-    Composition [kg] & 507.42 & 293.96 & 513.97 & 0.0029 & 0.68 & 67.00 & 0.28 \\
-    \bottomrule
-  \end{tabular}
-  \begin{tabular}{l S S S S S S}
-    \toprule
-    Element & {$^{238}$U} & {Fe} & {Cr} & {Ni} & {O} & {F} \\
-    \midrule
-    Composition [kg] & 142.02 & 0.75 & 0.13 & 0.14 & 2.27 & 3104.22 \\
-    \bottomrule
-  \end{tabular}
-  \label{table:67-salt-composition}
-\end{table}
-
-All neutronics simulations ran on OpenMC under continuous energy mode and on Moltres with the
-hybrid $S_N$-diffusion method. The hybrid simulations ran with a maximum stabilization factor of
-$c=250$, a void constant $\varsigma=0.5$, and a uncorrected diffusion coefficient cap of $D_g=2.5$.
-Refer to Sections \ref{sec:void-treatment} and \ref{sec:diffcoef-cap} for the relevant discussion
-on void treatment for near-void regions and the impact of large diffusion coefficient values on
-solver performance. As for the standard neutron diffusion method, it either converges very slowly
-or does not converge at all when $D_g$ values are uncapped due to excessive diffusive effects in
-the air-filled control rod thimbles. Group-wise $D_g$ values, generated using OpenMC, ranged from
-100 cm to 850 cm in the air regions as compared to the typical 0.5 cm to 2.0 cm range observed in
-non-air regions. Therefore, all neutron diffusion method simulation results in the following
-section were measured with the diffusion coefficient capped at $D_g=2.5$.
-
-Unlike the 2-D hybrid simulations which ran with $S_8$ angular discretization, the 3-D hybrid
-simulations ran with $S_6$ angular discretization due to significantly
-larger computational and memory requirements demanded by the 3-D simulations. Further discussion on
-computational performance may be found at the end of Section \ref{sec:3d-nts}. All simulations,
-except those for temperature reactivity coefficient calculations, ran on isothermal 3-D models at
-911 K.
-
-\subsection{3-D Numerical Results} \label{sec:3d-nts}
-
-\subsubsection{Initial Criticality Configuration}
-
-The \gls{MSRE} achieved initial criticality with Rod 1 inserted to a height of 46.6 inches relative
-to the fully inserted position. For reference, the rods are at a height of 51 inches when fully
-withdrawn. Table \ref{table:initial-crit}
-shows the $k_\text{eff}$ values from \gls{MSRE} experimental data, numerical benchmark data
-\cite{fratoni_molten_2020}, and this work. All numerical models overestimate the $k_\text{eff}$
-relative to the experimental value. In the \gls{MSRE} numerical benchmark report, the authors
-attribute this $k_\text{eff}$ discrepancy largely to discrepancies in the carbon cross section data
-in nuclear data libraries. They noted similar $k_\text{eff}$ overestimations by 1-2\% in benchmark
-results for the HTTR, a graphite-moderated, helium-cooled reactor. The OpenMC and Moltres
-simulations in this work used the same
-ENDF/B-VII.1 nuclear data library as the \gls{MSRE} numerical benchmark report.
-
-\begin{table}[htb]
-  \centering
-  \caption{$k_\text{eff}$ values from \gls{MSRE} experimental data, the \gls{MSRE} numerical
-  benchmark \cite{fratoni_molten_2020}, and the OpenMC and Moltres models in this work.}
-  \begin{tabular}{l S[table-format=1.5(2)]}
-    \toprule
-     & {$k_\text{eff}$} \\
-     \cmidrule(l){2-2}
-    \gls{MSRE} experimental data & 1.00000(420) \\
-    Serpent 2 (Numerical benchmark) & 1.02132(3) \\
-    OpenMC (This work) & 1.01308(20) \\
-    Hybrid (This work) & 1.01955 \\
-    Diffusion (This work) & 1.01885 \\
-    \bottomrule
-  \end{tabular}
-  \label{table:initial-crit}
-\end{table}
-
-Of the four $k_\text{eff}$ values, we expect to see the closest agreement between the Serpent 2
-model from the numerical benchmark and OpenMC from this work. However, the OpenMC \gls{MSRE} model
-differs from the Serpent 2 \gls{MSRE} models due to geometry modifications listed in Sections
-\ref{sec:2d-model-setup} and \ref{sec:3d-model-setup}. The benchmark report quantified the impact
-of several geometry modifications, namely the removal of the flow distributor, the homogenization
-of the sample basket, and the removal of the thermal shield. These modifications contribute to
--$0.098\%$, -$0.037\%$, and -$0.885\%$ change in the $k_\text{eff}$, respectively, for a final
-$k_\text{eff}$ estimate of $1.01093(3)$. This brings the OpenMC $k_\text{eff}$ estimate of
-$1.01308(20)$ to closer agreement with the numerical benchmark value.
-
-The hybrid method $k_\text{eff}$ estimate is 0.00647 higher than the OpenMC estimate. For this
-initial criticality configuration, the neutron diffusion method $k_\text{eff}$ estimate is closer
-to OpenMC than the hybrid method. Overall, the OpenMC and Moltres models show good agreement with
-the Serpent 2 model from the numerical benchmark.
-Figures * show the neutron group fluxes on the interior of the 3-D models. The salt-graphite
-lattice structure is visible from the group flux distributions. Group 7 and 8 neutron fluxes are
-particularly small near the control rods due to strong thermal neutron capture by Gadolinium in the
-rods.
-
-\subsubsection{Control Rod Worth}
-
-After achieving initial criticality, \gls{MSRE} researchers ran control rod calibration experiments
-with the \gls{MSRE} at zero-power (typically at about 10 W power output)
-\cite{prince_zero-power_1968}. They measured differential rod worths by measuring the time taken
-for the reactor power to increase by 20 W following a prescribed withdrawal of Rod 1 and applying
-the inhour equation to obtain the reactivity inserted. They repeated this process at different rod
-heights after successive additions of highly enriched uranium capsules. The final integral rod
-worth curve was normalized to the critical $^{235}$U loading of 65.25 kg
-\cite{prince_zero-power_1968}.
-
-For control rod worth calculations, the OpenMC and Moltres simulations ran with Rod 1 at fourteen
-different rod positions from full insertion (0 inch) to full withdrawal (51 inches). The rod
-positional limits of 0 inch and 51 inches represents the full range of deliberate rod motion
-allowable by the control rod drive motor \cite{robertson_msre_1965}.
-
-Figure \ref{fig:rod-worth} shows the integral rod worth curves from the \gls{MSRE} experimental
-data and the OpenMC and Moltres numerical models. By taking the $k_\text{eff}$ when Rod 1 is fully
-inserted as reference, all rod worth curves start at 0 pcm. The change in integral rod worth,
-represented by the gradient in the plot, is steepest as the tip of the rod passes through the
-center of the core. The hybrid method performs well in reproducing the rod worth curve from OpenMC;
-all hybrid method rod worth values fall within one or two sigma range of OpenMC.
-
-Due to the experimental procedure of computing integral rod worth from cumulative differential rod
-worth measurements, the uncertainty range of the \gls{MSRE} data scales with rod position.
-Both OpenMC and the hybrid method report consistently larger integral rod worths than the
-\gls{MSRE} data across all rod positions. Nevertheless, they are in good agreement with the
-\gls{MSRE} data as compared to the neutron diffusion method, which significantly overestimates the
-integral rod worth.
-
-Table \ref{table:rod-worth} shows the total rod worth of Rod 1.
-
-\begin{figure}[p]
-  \centering
-  \includegraphics[width=0.8\columnwidth]{rod-worth-curve}
-  \caption{Reactivity inserted by Rod 1 at various rod positions relative to the full insertion.}
-  \label{fig:rod-worth}
-\end{figure}
-
-\begin{table}[htb]
-  \centering
-  \caption{Total rod worth of Rod 1 when fully inserted.}
-  \begin{tabular}{l S[table-format=4(2)] S}
-    \toprule
-     & {$\rho_\text{worth}$ [pcm]} & {Percentage discrepancy [\%]}\\
-     \cmidrule(l){2-2}
-    \gls{MSRE} experimental data & 2250(67) & {-}\\
-    OpenMC (This work) & 2364(44) & 5.1 \\
-    Hybrid (This work) & 2345 & 4.2 \\
-    Diffusion (This work) & 2725 & 21.1 \\
-    \bottomrule
-  \end{tabular}
-  \label{table:rod-worth}
-\end{table}
-
-\subsubsection{Temperature Reactivity Coefficient}
-
-\subsubsection{Computational Performance}
-
-\section{Summary}
-
-%The hybrid method does not alter the $k_\text{eff}$ and flux estimates relative to the neutron
-%diffusion method in cases where the neutron diffusion method provides
+  \end{columns}
+\end{frame}

### 2024-12-04 presentation/2d.tex
Presentation commit

--- 2024-11-04 hybrid-nts.tex
@@
-In Chapter \ref{chap:hybrid}, I laid out the mathematical background and numerical implementation
-of the hybrid $S_N$-diffusion method for improving neutronic accuracy in reactor systems containing
-highly neutron-absorbing regions. I developed and implemented the diffusion and drift correction
-schemes for the hybrid method on Python and Moltres, respectively. This chapter demonstrates
-the hybrid method through 1-D, 2-D, and 3-D $k$-eigenvalue simulations of the \gls{MSRE}.
-I employed the OpenMC Monte Carlo neutral particle transport software \cite{romano_openmc:_2015} to
-generate group constants for the hybrid method and reference solutions for method verification.
-The 1-D models serve to verify the newly-implemented $S_N$ and hybrid methods
-against OpenMC as well as to analyze the impact of key input parameters on the hybrid method. The
-2-D models test the hybrid method performance through increased spatial dimensionality and reduced
-geometric asymmetry relative to the 1-D models. Lastly, the 3-D models serve as neutronics model
-validation against the \gls{MSRE} benchmark model by Fratoni et al.\ \cite{fratoni_molten_2020}, a
-testbed for \gls{HPC} scaling tests, and preparation for the time-dependent multiphysics
-simulations of \gls{MSRE} reactivity insertion transients.
-
-Section \ref{sec:msre-gc} details the OpenMC model setup for group constant generation. Section
-\ref{sec:nts-methods} describes and labels the different neutronics methods employed for the
-$k$-eigenvalue simulations. Section \ref{sec:test-metrics} details basic test metrics relevant to
-neutronics verification and validation. Sections \ref{sec:1d-results}, \ref{sec:2d-results}, and
-\ref{sec:3d-results} discuss the model setup, results, and discussions for the 1-D, 2-D, and 3-D
-\gls{MSRE} models, respectively.
-
-\section{Group Constant Data Generation on OpenMC} \label{sec:msre-gc}
-
-Group constants relevant to the neutron diffusion, $S_N$ neutron transport, or both
-methods are:
-%
-\begin{itemize}
-  \item $\Sigma_{t,g}$: Macroscopic total cross section in group $g$,
-  \item $\Sigma_{r,g}$: Macroscopic removal cross section in group $g$,
-  \item $\Sigma_s^{g'\rightarrow g}$: Macroscopic group-to-group scattering cross section matrix,
-  \item $\Sigma_{s,l}^{g'\rightarrow g}$: $l$-th Legendre moment of the macroscopic
-    group-to-group scattering cross section matrix,
-  \item $\Sigma_{sp,l}^{g'\rightarrow g}$: $l$-th Legendre moment of the macroscopic
-    group-to-group scattering production cross section matrix,
-  \item $D_g$: $P_1$-based diffusion coefficient in group $g$,
-  \item $\nu\Sigma_{f,g}$: Product of the average number of neutrons produced per fission and the
-    macroscopic fission cross section in group $g$,
-  \item $\chi_g$: Neutron fission spectrum in group $g$.
-\end{itemize}
-%
-I generated the group constants using OpenMC's multigroup cross section generation capability
-\cite{boyd_multigroup_2019}. A group constant postprocessing script for Moltres
-(\texttt{moltres\_xs.py}) parses OpenMC simulation output files and rewrites the group constant
-data into a JSON format file suitable for Moltres simulations. While OpenMC provides most of the
-group constants directly, the postprocessing script calculates $\Sigma_{r,g}$ as
-%
-\begin{align}
-  \Sigma_{r,g} =& \sum^G_{g'\neq g}\Sigma_s^{g\rightarrow g'}+\Sigma_{a,g}-\left(\Sigma_{sp}^{g
-    \rightarrow g} - \Sigma_s^{g\rightarrow g}\right)
-  \shortintertext{where}
-      \Sigma_{a,g} =& \mbox{ macroscopic absorption cross section in group $g$,} \nonumber \\
-      \Sigma_{sp}^{g\rightarrow g} =& \mbox{ macroscopic scattering production cross section from
-      group $g$ to $g$.} \nonumber
-\end{align}
-%
-$\Sigma_{r,g}$ primarily represents the loss of neutrons from group $g$ through outscattering and
-absorption. $\Sigma_{sp}^{g\rightarrow g}$ incorporates neutron multiplication effects from neutron
-knockout reactions into the scattering cross section. Neutronics codes commonly identify neutron
-knockout reactions as scattering reactions, but $\Sigma_{r,g}$ is a convenient parameter in which
-to incorporate neutron knockout effects in the neutron diffusion method. I ran all test cases
-for $S_N$ and hybrid method calculations with scattering cross sections approximated by up to 2nd
-Legendre moments ($L=2$).
-
-OpenMC uses the $P_1$ flux-limited formulation \cite{pomraning_flux-limited_1984} by default for
-calculating $D_g$ as follows
-%
-\begin{align}
-  D_g =& \frac{1}{3\Sigma_{tr,g}},
-  \shortintertext{where}
-  \Sigma_{tr,g} =& \frac{\langle\Sigma_{t,g}\phi_g\rangle-\langle\Sigma_{s1,g}\phi_g\rangle}
-  {\langle\phi_g\rangle}, \nonumber \\
-  \langle\Sigma_{t,g}\phi_g\rangle =& \int_{r\in V}dr \int_{4\pi}d\Omega\int^{E_{g-1}}_{E_g}dE\
-  \Sigma_{t,g}(r,E)\Psi(r,E,\Omega), \nonumber \\
-  \langle\Sigma_{s1,g}\phi_g\rangle =& \int_{r\in V}dr \int_{4\pi}d\Omega\int^{E_{g-1}}_{E_g}dE
-  \int_{4\pi}d\Omega'\int^{\infty}_0dE'\int^1_{-1}d\mu\ \mu\Sigma_s(r,E'\rightarrow E,\Omega'\cdot
-  \Omega)\phi(r,E',\Omega'), \nonumber \\
-  \langle \phi \rangle =& \int_{r\in V}dr\int_{4\pi}d\Omega\int^{E_{g-1}}_{E_g}dE\ \Psi(r,E,\Omega)
-  .\nonumber
-\end{align}
-
-I ran all \gls{MSRE} neutronics simulations with the eight neutron energy group structure proposed
-by Jaradat for \gls{MSRE} analysis \cite{jaradat_development_2021-1}.
-Table \ref{table:energy-group} shows the upper neutron energy bounds of the eight-group structure.
-I selected this energy group structure as a compromise between accuracy and computational expense
-among the various group structures investigated by Jaradat ranging from four to sixteen groups.
-
-\begin{table}[htb]
-  \centering
-  \caption{Neutron energy group structure in this work. Originally devised by Jaradat
-  \cite{jaradat_development_2021-1}.}
-  \begin{tabular}{r S}
-    \toprule
-    Group & {Upper energy bound [eV]} \\
-    \midrule
-    1 & 2.000$\times 10^7$ \\
-    2 & 1.353$\times 10^6$ \\
-    3 & 6.734$\times 10^4$ \\
-    4 & 9.118$\times 10^3$ \\
-    5 & 1.487$\times 10^2$ \\
-    6 & 4.000$\times 10^0$ \\
-    7 & 6.250$\times 10^{-1}$ \\
-    8 & 8.000$\times 10^{-2}$ \\
-    \bottomrule
-  \end{tabular}
-  \label{table:energy-group}
-\end{table}
-
-I set the material compositions and densities of every component in the \gls{MSRE} models using
-reference data from \gls{MSRE} benchmark specifications collated by Fratoni et al.
-\cite{fratoni_molten_2020}. All OpenMC models utilize the ENDF/B-VII.1 nuclear data library
-\cite{chadwick_endf/b-vii.1_2011} and include thermal scattering cross section data for beryllium
-and carbon target nuclei in the fuel salt and graphite moderator. All models have a uniform
-temperature of 900 K unless otherwise stated. Detailed information pertaining specifically to the
-1-D, 2-D, and 3-D OpenMC models may be found in the respective subsections below.
-
-\section{Neutronics Methods} \label{sec:nts-methods}
-
-I ran all cases on OpenMC to generate the required group constants and reference $k_\text{eff}$
-values to assess the accuracy of the deterministic methods ($S_N$, neutron diffusion, hybrid 
-methods) against OpenMC in continuous-energy (OpenMC-CE) and multigroup (OpenMC-MG) modes. Table
-\ref{table:var} details how OpenMC and the deterministic methods represent the position, direction
-of travel, neutron energy, and angle-dependence in $\Sigma_s$. The hybrid $S_N$-diffusion method
-features representations from either $S_N$ or neutron diffusion depending on the subregion under
-consideration. Comparing OpenMC-MG results with OpenMC-CE results allows us
-to quantify errors arising from neutron energy group discretization and the scattering cross
-section approximations.
-
-\begin{table}[htb!]
-  \centering
-  \footnotesize
-  \caption{Variable representations in OpenMC under continuous-energy (OpenMC-CE) and multigroup
-  (OpenMC-MG) modes, and in the $S_N$ neutron transport and neutron diffusion.}
-  \begin{tabular}{l l l l l}
-    \toprule
-    Variable & OpenMC-CE & OpenMC-MG & $S_N$ Transport & Diffusion \\
-    \midrule
-    Position, $\bm{r}$ & Continuous & Continuous & Discrete & Discrete \\
-    Direction of travel, $\bm{\hat{\Omega}}$ & Continuous & Continuous & Discrete & N/A \\
-    Energy, $E$ & Continuous & Discrete & Discrete & Discrete \\
-    Angle-dependence in $\Sigma_s$ & Continuous & $L=2$ & $L=2$ & N/A \\
-    \bottomrule
-  \end{tabular}
-  \label{table:var}
-\end{table}
-
-The rest of this section pertains to the deterministic neutronics methods used in this work for
-modeling the \gls{MSRE}.
-For the 1-D models, I performed neutronics calculations with the $S_N$, neutron diffusion, and
-hybrid method implementations on Moltres and Python. Table \ref{table:nts-methods} highlights the
-differences between the two implementations. The $S_N$ neutron transport method implementation on
-Moltres follows the methodology for solving the \gls{SAAF} formulation of the $S_N$ equations as
-described in Section \ref{sec:saaf} while the implementation on Python follows the methodology
-for solving the original first-order formulation as described in Section \ref{sec:python-sn}.
-
-The implementations on Moltres discretize the problem domain in space using the finite element
-method while the implementations on Python use finite difference or diamond difference schemes. The
-most crucial difference lies in the transport correction formulations for the hybrid method. The
-Moltres implementation uses the drift correction scheme described in Section
-\ref{sec:drift-correction} while the Python implementation uses the diffusion correction
-scheme described in Section \ref{sec:diffusion-correction}.
-
-In this chapter, I refer to the methods and results from the Moltres and Python implementations
-with the Moltres- or Python- prefix, i.e., Moltres-$S_N$, Moltres-diffusion, Moltres-hybrid,
-Python-$S_N$, Python-diffusion, and Python-hybrid.
-
-For 2- and 3-D models, I abandoned the diffusion correction formulation in favor of the drift
-correction formulation due to superior numerical performance in the latter. Readers may find the
-analysis supporting this decision in the analysis of 1-D results in Section \ref{sec:1d-results}.
-
-\begin{table}[htb!]
-  \centering
-  \footnotesize
-  \caption{Implementation differences of neutronics methods on Moltres and Python.}
-  \begin{tabular}{p{0.14\linewidth} p{0.25\linewidth} p{0.15\linewidth} p{0.15\linewidth} p{0.17\linewidth}}
-    \toprule
-    Implementations & Spatial discretization & $S_N$ formulation & $S_N$ acceleration &
-    Hybrid transport\newline correction \\
-    \midrule
-    Moltres & Finite element method & SAAF & Nonlinear diffusion acceleration & Drift correction \\
-    Python & Finite difference (Diffusion)\newline Diamond difference ($S_N$) & First-order & None & Diffusion correction \\
-    \bottomrule
-  \end{tabular}
-  \label{table:nts-methods}
-\end{table}
-
-\section{Basic Test Metrics} \label{sec:test-metrics}
-
-In this section, I define several metrics for comparing simulation results among the various
-neutronics methods.
-
-For steady-state eigenvalue calculations, all solvers directly provide the effective neutron
-multiplication factor ($k_\text{eff}$), the rate of neutron population increase between successive
-neutron generations. Reactivity is defined as
-%
-\begin{gather}
-  \rho \equiv \frac{k_\text{eff}-1}{k_\text{eff}}.
-\end{gather}
-%
-Differences in reactivity, for either comparisons between neutronics methods or control rod worth
-calculations, can be calculated as
-%
-\begin{gather}
-  \Delta\rho_{1,2} = \rho_1 - \rho_2 =
-  \frac{k_{\text{eff},1}-k_{\text{eff},2}}{k_{\text{eff},1}k_{\text{eff},2}}.
-\end{gather}
-
-Spatially dependent variables (e.g., neutron group flux, neutron drift) are sampled at fixed 0.1 cm
-intervals along the specified lines (e.g., horizontal, diagonal, etc.) regardless of the actual
-mesh resolution.
-
-\section{1-D Neutronics Models \& Results} \label{sec:1d-results}
-
-\subsection{1-D Neutronics Model Setup \& Geometries}
-
-I designed six 1-D test cases with increasing complexity to verify the $S_N$ and hybrid
-$S_N$-diffusion implementations and to test the performance of the hybrid
-method in response to various geometrical features. The last two cases resemble the
-reference \gls{MSRE} design \cite{robertson_msre_1965}, which has centrally located control rods
-and air-filled rod guide tubes. Figure \ref{fig:case-geom} shows the geometries of Cases 1a to
-3b. All geometries have reflective boundary conditions at $x=0$ cm, thereby
-forming half-core or repeating unit cell models. Cases 1a and 1b are repeating unit
-cell models with reflecting boundaries on the right-side boundaries. Cases 2a, 2b, 3a, and 3b
-are 1-D half-core models with vacuum boundaries on the right-side boundaries.
-
-\begin{figure}[htb!]
-  \centering
-  \includegraphics[width=0.8\columnwidth]{case-geometry}
-  \caption{Geometries of the 1-D test cases. The material labeled ``mixture'' represents a
-    homogeneous mixture of fuel and graphite at a ratio of 22.5\%-77.5\% by volume. All geometries
-    have reflective boundary conditions on the boundary at $x=0$ cm. The right-side boundaries are
-    reflective for Cases 1a and 1b, and vacuum for Cases 2a, 2b, 3a, and 3b.}
-  \label{fig:case-geom}
-\end{figure}
-
-Cases 1a and 1b are simple test cases containing neutron multiplying regions. Cases 1a and 1b serve
-to verify basic neutronics phenomena modeling (e.g., fission, scattering, absorption) of the newly
-implemented $S_N$ and neutron diffusion methods in Python and $S_N$ method in Moltres. I did not
-apply the hybrid method on these cases. Cases 2a and 2b represent 1-D models of the \gls{MSRE} with
-air-filled control rod guide tube regions, a homogenized fuel-graphite lattice regions, and outer
-vessel regions. Case 2b additionally contains a 0.5cm-thick control rod region to contrast with
-Case 2a for control rod worth calculations. Cases 3a and 3b retain the heterogeneous fuel-graphite
-lattice geometry present in the \gls{MSRE}.
-
-In my preliminary investigations, the original control rod material composition
-(70 wt\% Gd$_2$O$_3$-30 wt\% Al$_2$O$_3$) led to excessively large control rod worth estimates of
-approximately 55000 pcm in the 1-D models, effectively halving the $k_\text{eff}$. The worth
-estimates are ten times larger
-than the 5500 pcm expected from full rod insertion of all three rods in the actual \gls{MSRE}
-\cite{fratoni_molten_2020}. Therefore, I greatly reduced the proportion of Gd$_2$O$_3$ in the
-control rod to 0.35 wt\% for a fairer assessment of rod worth in the 1-D models. This change
-brought the new rod worth estimates down to below 20000 pcm.
-
-\subsection{1-D Neutronics Model Mesh Convergence Tests}
-
-I ran mesh convergence studies based on Case 3b for the neutron diffusion and $S_N$ neutron
-transport methods. Given that the hybrid $S_N$-diffusion method comprises of a combination of those
-methods, I found hybrid simulations were fully spatially resolved on meshes that are
-appropriately resolved for both neutron diffusion and $S_N$ simulations.
-
-\begin{figure}[htb!]
-  \centering
-  \begin{subfigure}[b]{0.48\columnwidth}
-    \centering
-    \includegraphics[width=\columnwidth]{diffusion-mesh-convergence-k}
-    \caption{Neutron diffusion method}
-    \label{fig:diff-mesh-k}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}[b]{0.48\columnwidth}
-    \centering
-    \includegraphics[width=\columnwidth]{sn-mesh-convergence-k}
-    \caption{$S_8$ neutron transport method}
-    \label{fig:sn-mesh-k}
-  \end{subfigure}
-  \caption{Convergence of multiplication factor ($k_\text{eff}$) estimates for Case 3b across five
-    levels of mesh refinement. Higher refinement levels correspond to smaller mesh element sizes.}
-  \label{fig:mesh-k}
-\end{figure}
-
-Figures \ref{fig:diff-mesh-k} and \ref{fig:sn-mesh-k} show the convergence of $k_\text{eff}$
-estimates observed for Case 3b simulations with neutron diffusion and $S_8$ neutron transport
-methods. The $k_\text{eff}$ values plateau at 0.9713203 and 0.9813790 for the neutron diffusion and
-$S_8$ methods, respectively. By mesh refinement level 2, both methods fall within $5\times 10^{-7}$
-range of the $k_\text{eff}$ value at level 4.
-
-Figures \ref{fig:diff-mesh-flux} and \ref{fig:sn-mesh-flux} show the absolute differences in the
-group flux distributions at mesh refinement levels 1 to 3 relative to the reference group flux
-distribution at level 4 (maximum refinement). I omitted data from level 0 to improve the fidelity
-of the plots as flux differences drop significantly by one order of magnitude with each level of
-refinement.
-With these results, I concluded that the mesh resolution at level 2 is a sufficient guideline for
-the mesh resolutions used in the subsequent 1-D analyses. The absolute differences in both
-$k_\text{eff}$ and group flux distributions at level 2 remain below $5\times10^{-7}$ relative to
-level 4.
-
-\begin{figure}[htb!]
-  \centering
-  \includegraphics[width=\columnwidth]{diffusion-mesh-convergence-flux}
-  \caption{Absolute difference in group fluxes ($\phi_g$) from the neutron diffusion method for
-  mesh refinement levels 1 to 4 relative to the reference group flux distributions at mesh
-  refinement level 5.}
-  \label{fig:diff-mesh-flux}
-\end{figure}
-
-\begin{figure}[htb!]
-  \centering
-  \includegraphics[width=\columnwidth]{sn-mesh-convergence-flux}
-  \caption{Absolute difference in group fluxes ($\phi_g$) from the $S_8$ neutron transport method
-  for mesh refinement levels 1 to 4 relative to the reference group flux distributions at mesh
-  refinement level 5.}
-  \label{fig:sn-mesh-flux}
-\end{figure}
-
-\FloatBarrier
-
-\subsection{1-D Neutronics Numerical Results}
-
-\subsubsection{Reactivity}
-
-Figure \ref{fig:1d-rho} shows the difference in reactivities of OpenMC-MG, $S_8$,
-neutron diffusion, and hybrid methods relative to OpenMC-CE. The standard deviations of
-reactivity values from OpenMC-CE are approximately 40 pcm for Cases 1a and 1b and 60 pcm for the
-rest as depicted by the blue highlighted area in Figure \ref{fig:1d-rho}. For Cases 1a and 1b, the
-OpenMC-MG, $S_8$, and neutron diffusion methods show excellent agreement with OpenMC-CE as the
-reactivity values fall within the one or two standard deviation range. The Case 1 results show
-the Moltres and Python implementations are consistent with each other given that they are spatially
-well-resolved.
-
-\begin{figure}[htb!]
-  \centering
-  \includegraphics[width=\columnwidth]{rho}
-  \caption{Difference in reactivity $\rho$ of all neutronics methods investigated relative
-  to OpenMC-CE.}
-  \label{fig:1d-rho}
-\end{figure}
-
-For all cases, the OpenMC-MG and the $S_8$ methods show consistent agreement with one another. 
-While they deviate from OpenMC-CE by approximately 350 pcm for Cases 2a and 3a, they fall within
-two standard deviations of OpenMC-CE for Cases 2b and 3b. When increasing the maximum Legendre
-polynomial order to approximate the angular dependence in $\Sigma_s^{g'\rightarrow g}$ from
-2nd-order to 3rd-order, the reactivity changed by only 1 pcm. Therefore, I attribute the reactivity
-discrepancies to the eight neutron energy group structure which remains the
-only significant difference between OpenMC-CE and the multigroup methods.
-
-The neutron diffusion and hybrid methods agree closely with one another for Cases 2a and 3a which
-exclude the control rod region. This shows that the hybrid method provides similar $k_\text{eff}$
-estimates to the neutron diffusion method in 1-D simulations without highly neutron-absorbing
-regions. Compared to the OpenMC-MG and $S_8$ reactivity values, the neutron diffusion and hybrid
-method reactivity values agree closer with the reference OpenMC-CE value. However, this is likely
-due to favorable error cancellation; Figures \ref{fig:3a-flux} and \ref{fig:3a-flux-diff} show
-OpenMC-MG and the $S_8$ methods reproduce
-the OpenMC-CE flux distribution more accurately than the neutron diffusion and hybrid methods.
-
-In Cases 2b and 3b, the neutron diffusion method largely fails to accurately capture the
-effect of introducing the control rod region as evidenced by the -1500 pcm and -1150 pcm
-discrepancies relative to OpenMC-CE. The hybrid method fares better with -400 pcm and -300 pcm
-discrepancies.
-
-\begin{figure}[htb!]
-  \centering
-  \includegraphics[width=\columnwidth]{worth}
-  \caption{Percentage difference in rod worth for Cases 2 and 3 of all neutronics methods
-  investigated relative to OpenMC-CE.}
-  \label{fig:1d-worth}
-\end{figure}
-
-\subsubsection{Control rod worth}
-
-Moving the discussion to control rod worths, figure \ref{fig:1d-worth} shows the percentage
-difference in rod worths for Cases 2 and 3 of OpenMC-MG, $S_8$, neutron diffusion, and hybrid
-methods relative to OpenMC-CE. I calculated the rod worth estimates from each method by taking the
-difference in reactivity between the ``rod out'' (a) and ``rod in'' (b) configurations of Cases 2
-and 3. All neutronics methods overestimate the rod worth relative to OpenMC-CE, resulting in
-positive percentage difference values observed in the figure.
-The neutron diffusion methods clearly show up as outliers with rod worth estimates in excess
-of 8\%. Most of the remaining methods cluster around 2.5\% and 3\% percentage difference for Cases
-2 and 3 with the exception of the Python-hybrid method for Case 3.
-
-Given that the hybrid methods rely on transport corrections derived from the $S_N$ method, the
-$S_8$ rod worth estimates serve as the
-reference point for hybrid method verification. Additionally, errors arising from the multigroup
-approximation affect both hybrid and $S_N$ simulations to similar degrees. While the Python-hybrid
-method produces the lowest percentage difference relative to OpenMC-CE for Case 3, the
-Python-hybrid method deviates from the $S_8$ ``multigroup'' reference rod worths more than the
-Moltres-hybrid method. Nevertheless, both hybrid method implementations provide significant
-improvements in rod worth estimates over the neutron diffusion method.
-
-\subsubsection{Neutron flux distribution}
-
-This discussion on the neutron flux distributions focuses on Cases 3a and 3b which most closely
-resemble the actual \gls{MSRE} reactor geometry. Cases 3a and 3b resemble the \gls{MSRE} reactor
-with a control rod withdrawn and inserted, respectively. Figures \ref{fig:3a-flux} and
-\ref{fig:3b-flux} show the neutron group flux distributions for Cases 3a and 3b from OpenMC-CE and
-OpenMC-MG. Differences between the two methods arise from the multigroup approximation, notably in
-group 2 for Case 3a and groups 2, 6, 7, 8 for Case 3b. Future work aimed at finding a better
-neutron energy group structure should at minimum focus on improving fidelity in the listed energy
-groups.
-
-For a fair evaluation of the deterministic multigroup methods without distortions from the
-multigroup approximation, I chose to compare their flux distributions with the OpenMC-MG flux
-distribution. Figures \ref{fig:3a-flux-diff} and \ref{fig:3b-flux-diff} show the absolute
-difference in neutron group flux distributions of the Moltres-$S_8$, Moltres-diffusion,
-Moltres-hybrid, and Python-hybrid methods relative to OpenMC-MG. The figures omit Python-$S_8$ and
-Python-diffusion as their distributions are nearly identical to their Moltres-based counterparts.
-The neutron diffusion and hybrid methods fare worse than the $S_8$ method at capturing the
-oscillatory pattern in groups 1, 2, 5, 7, and 8 arising from the fuel-graphite lattice geometry
-along $x=5$ cm to 10 cm. In both Cases 3a and 3b, the neutron diffusion method (yellow plot lines)
-exhibits significant flux deviations in groups 1, 2, 7, and 8 near $x=0$ cm.
-
-Both hybrid methods
-provide significant improvements in neutron flux distributions compared to the neutron diffusion
-method as shown by the generally smaller flux difference values particularly near $x=0$ cm. The
-figures also show Moltres-hybrid (green) provides more accurate flux estimates than
-Python-hybrid (red) in all instances. The differences indicate deficiencies in the diffusion
-correction scheme in Python-hybrid which lead to poorer performance compared to the drift
-correction scheme in Moltres-hybrid.
-
-\begin{figure}[htb!]
-  \centering
-  \includegraphics[width=\columnwidth]{case-3a-flux}
-  \caption{Case 3a neutron group flux distributions from OpenMC-CE and OpenMC-MG.}
-  \label{fig:3a-flux}
-\end{figure}
-
-\begin{figure}[htb!]
-  \centering
-  \includegraphics[width=\columnwidth]{case-3b-flux}
-  \caption{Case 3b neutron group flux distributions from OpenMC-CE and OpenMC-MG.}
-  \label{fig:3b-flux}
-\end{figure}
-
-\begin{figure}[htb!]
-  \centering
-  \includegraphics[width=\columnwidth]{case-3a-flux-diff}
-  \caption{Absolute difference in neutron group flux distributions for Case 3a from Moltres-$S_8$,
-  Moltres-diffusion, Moltres-hybrid, and Python-hybrid relative to OpenMC-MG.}
-  \label{fig:3a-flux-diff}
-\end{figure}
-
-\begin{figure}[htb!]
-  \centering
-  \includegraphics[width=\columnwidth]{case-3b-flux-diff}
-  \caption{Absolute difference in neutron group flux distributions for Case 3b from Moltres-$S_8$,
-  Moltres-diffusion, Moltres-hybrid, and Python-hybrid relative to OpenMC-MG.}
-  \label{fig:3b-flux-diff}
-\end{figure}
-
-\FloatBarrier
-
-\subsubsection{Drift \& diffusion correction distribution}
-
-Figures \ref{fig:3b-drift} and \ref{fig:3b-svdc} show the distributions of the drift and diffusion
-correction parameters, respectively, in each neutron energy group for Case 3b. As a reminder to the
-reader, the hybrid $S_N$-diffusion methods generate these correction parameters in subregions
-containing highly neutron absorbing materials. I implemented the drift correction scheme for
-Moltres-hybrid and the diffusion correction scheme for Python-hybrid. The hybrid methods
-selectively apply the $S_N$ method to improve neutronic modeling within the subregions. I computed
-the reference distributions (yellow data points) in both plots from reference flux solutions of the
-standard $S_8$ method.
-
-As discussed in Section \ref{sec:buffer-region}, both sets of correction parameters generated by
-the two correction schemes match their respective reference values within the correction subregion
-from $x=0$ cm to 10 cm except near the subregion boundary at $x=10$ cm. Therefore, both schemes
-provide improved flux estimates in the subregion compared to the neutron diffusion method.
-The prior reactivity and flux distribution results indicate Moltres-hybrid is more accurate
-than Python-hybrid relative to OpenMC-MG as the reference.
-
-In Figure \ref{fig:3b-drift}, the drift
-correction distributions (Moltres-hybrid) for all eight energy groups are well-defined throughout
-the subregion and are continuous except at material interfaces. Consequently, the drift correction
-scheme provides accurate flux corrections for about 75\% of the subregion based on the buffer
-region cutoff criteria which occurs around $x=7.5$ cm where most of the drift distributions are
-zero (refer to Section \ref{sec:buffer-region}). By contrast, the diffusion correction
-distributions in Figure \ref{fig:3b-svdc} contains numerous discontinuities tending towards
-positive or negative infinity in all eight energy groups. The discontinuities arise when
-computing corrected diffusion coefficients (Eq. \ref{eq:svdc}) near flux peaks and troughs
-(stationary points) where the flux gradients approach zero. The location of stationary points from
-the diffusion and $S_N$ calculations in the hybrid method do not necessarily coincide, resulting in
-discontinuous flux gradients in the final flux solution. Discontinuities are generally unavoidable
-in most problems because the fluxes typically reach local minima in control rod regions.
-Furthermore, unlike negative drift coefficients, negative diffusion coefficients are unphysical.
-Taking everything into consideration, the diffusion correction scheme is less accurate and more
-numerically unstable than the drift correction scheme.
-
-The subsequent discussions on the correction region size and the number of outer
-iterations focus on the drift correction scheme of the hybrid method.
-
-\begin{figure}[htb!]
-  \centering
-  \includegraphics[width=\columnwidth]{case-3b-drift}
-  \caption{Multigroup drift correction ($\vec{D}_g$) $x$-component distributions from the
-  Moltres-hybrid and Moltres-$S_8$ solvers.}
-  \label{fig:3b-drift}
-\end{figure}
-
-\begin{figure}[htb!]
-  \centering
-  \includegraphics[width=\columnwidth]{case-3b-svdc}
-  \caption{Multigroup diffusion correction ($D^s_g$) $x$-component distributions from the
-    Python-hybrid and Python-$S_8$ solvers.}
-  \label{fig:3b-svdc}
-\end{figure}
-
-\FloatBarrier
-
-\subsubsection{Impact of correction subregion sizes}
-
-The hybrid $S_N$-diffusion method relies on $S_N$ calculations in the correction subregion to
-generate flux corrections. Minimizing the size of the correction subregion and the $S_N$ subproblem
-is essential for the hybrid method to be computationally competitive for time-dependent full-core
-simulations. I investigated the effect of the correction subregion size on the $k_\text{eff}$ and
-control rod worth estimates with Cases 3a and 3b by varying the subregion sizes from 10 cm to 40 cm
-at 5 cm-intervals.
-
-Figures \ref{fig:v1-size-a-k} and \ref{fig:v1-size-b-k} show the $k_\text{eff}$ estimates from the
-hybrid method for Cases 3a and 3b, respectively. In both cases, the $k_\text{eff}$ values initially
-decrease as the correction subregion sizes increase before reversing in trend when the subregion
-size reaches 35 cm and beyond. The $k_\text{eff}$ values vary by up to 164 pcm for Case 3a and 109
-pcm for Case 3b. An important observation here is that the $k_\text{eff}$ values do not
-monotonically converge towards the $k_\text{eff}$ estimate from the $S_8$ method. The data implies
-other significant sources of discrepancies exist beyond those being corrected in the
-correction subregion by the hybrid method.
-
-% In general,
-% discrepancies arise between the neutron diffusion method and $S_8$ from the neutron diffusion 
-% method's failure to accurately model the neutron flux in the fuel-graphite lattice region. The
-% figures imply that those discrepancies result in a positive change in $k_\text{eff}$ since the
-% $k_\text{eff}$ decreases as larger fractions of the problem domain benefit from flux corrections.
-% The remaining discrepancy in $k_\text{eff}$ between the hybrid method with a correction subregion
-% size of 40 cm and the $S_8$ method arises from
-
-\begin{figure}[htb!]
-  \centering
-  \begin{subfigure}[b]{0.49\columnwidth}
-    \centering
-    \includegraphics[width=\columnwidth]{correction-size-a-k}
-    \caption{Case 3a}
-    \label{fig:v1-size-a-k}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}[b]{0.49\columnwidth}
-    \centering
-    \includegraphics[width=\columnwidth]{correction-size-b-k}
-    \caption{Case 3b}
-    \label{fig:v1-size-b-k}
-  \end{subfigure}
-  \caption{$k_\text{eff}$ estimates from the hybrid method for Cases 3a and 3b with different
-  correction subregion sizes. The horizontal lines indicate $k_\text{eff}$ estimates from the
-  OpenMC-CE, OpenMC-MG, and $S_8$ methods.}
-  \label{fig:v1-size-k}
-\end{figure}
-
-Figure \ref{fig:v1-size-rho} shows the percentage difference in rod worth relative to OpenMC-CE. 
-Due to the identical trends observed in the $k_\text{eff}$ estimates of both cases, the control rod
-worth estimates do not change significantly when the correction subregion
-sizes change. This indicates limiting the correction subregion size to save
-computational cost on the expensive $S_N$ calculations has a negligible impact on rod worth
-estimates. The rod worth estimates for all investigated correction subregion sizes remain within
-0.2\% of the $S_8$ method.
-
-\begin{figure}[htb!]
-  \centering
-  \includegraphics[width=0.8\columnwidth]{correction-size-rho}
-  \caption{Percentage difference in rod worth from the hybrid method relative to OpenMC-CE for
-    Cases 3a and 3b with different correction subregion sizes. The horizontal lines indicate
-    equivalent rod worth differences from the OpenMC-MG and $S_8$ methods.}
-  \label{fig:v1-size-rho}
-\end{figure}
-
-The $k_\text{eff}$ and rod worth results show the hybrid method is able to accurately
-reproduce rod worth estimates as long as the correction subregion size remains consistent. The
-correction subregion size must also be sufficiently large to capture the highly neutron-absorptive
-rod's influence on the neutron flux around the rod at $x=0$ cm. The drift correction parameter
-distributions in Figure \ref{fig:3b-drift} indicate that the influence of the rod on the drift, and
-consequently the flux, extend to approximately $x=5$ cm before the drift distribution settles on a
-regular repeating pattern influenced by the fuel-graphite lattice structure. I applied a consistent
-correction subregion size of 10 cm for Cases 2 and 3 since it sufficiently meets the requirements
-discussed while also minimizing the computational cost of the $S_8$ subsolver. Other reactor
-designs consisting of different material components may require different minimum correction
-subregion sizes.
-
-\subsubsection{Relaxing the $S_N$ subsolver convergence threshold}
-
-As described in Section \ref{sec:hybrid-algorithm}, the hybrid $S_N$-diffusion method works by
-iteratively coupling a neutron diffusion solver and a $S_N$ subsolver with the latter generating
-drift correction parameters for the former. Excluding the initial neutron diffusion calculation,
-each outer (fixed point) iteration comprises of an inner $S_N$ calculation on the correction
-subregion and an inner neutron diffusion calculation on the entire domain. The \gls{FEM} algorithm
-in Moltres determines an individual solver reaches convergence when the combined
-residual contributions fall below a specified convergence tolerance threshold. The fixed point
-iteration scheme under the \texttt{MultiApps} system in Moltres for coupling multiple individual
-solvers (referred to as apps) requires users to specify a main app for the final convergence check
-after each outer iteration to terminate the iterations. The scheme also requires users to set a
-fixed point convergence tolerance value independent of the tolerance values of each individual
-solver. Naturally, the fixed point tolerance value must be equal to or greater than the tolerance
-value of the main individual solver for the possibility of overall convergence. The hybrid method
-implementation in Moltres sets the neutron diffusion solver as the main app and the $S_N$ subsolver
-as the sub app.
-
-The hybrid method shares some similarities with diffusion-based \gls{HOLO} acceleration schemes
-(refer to Section \ref{sec:drift-correction}) for the neutron transport methods which do not
-require the neutron transport side of the calculations to fully converge
-\cite{reynolds_analysis_2023, wang_diffusion_2014}.
-The typical convergence threshold value required for converged neutron flux calculations in Moltres
-is $10^{-8}$; I set the convergence tolerance value to $10^{-8}$ for the neutron diffusion solver
-and the fixed point scheme. I varied the convergence tolerance value for the $S_N$ subsolver to 
-find the optimal value for maintaining sufficient solution accuracy and minimizing computational
-costs; tightening the $S_N$ convergence tolerance value generally results in more outer iterations
-performed before the solver meets the fixed point convergence tolerance.
-Table \ref{table:sn-tol} lists the number of outer iterations required by the hybrid method for
-a given set of $S_8$ subsolver convergence tolerance values in Case 3b.
-
-\begin{table}[htb]
-  \centering
-  \caption{Number of outer iterations in hybrid method calculations of Case 3b for a given set of
-  convergence tolerance values imposed on the $S_8$ subsolver.}
-  \begin{tabular}{l S S S S S S}
-    \toprule
-    $S_8$ subsolver tolerance, $\epsilon_\text{tol}$ & {$10^{-8}$} & {$10^{-7}$} & {$10^{-6}$} & {$10^{-5}$} & {$10^{-4}$} & {$10^{-3}$} \\
-    \midrule
-    Number of outer iterations & 3 & 3 & 3 & 2 & 2 & 1 \\
-    \bottomrule
-  \end{tabular}
-  \label{table:sn-tol}
-\end{table}
-
-Figure \ref{fig:sn-tol} shows the $k_\text{eff}$ error estimates relative to the reference value
-when $\epsilon_\text{tol}=10^{-8}$. The hybrid method exhibits superlinear convergence ($q=1.333$)
-with respect to the $S_N$ subsolver convergence tolerance value. Based on the results, setting
-$\epsilon_\text{tol}$ to $10^{-5}$ is the best balance between accuracy and computational
-cost because the $k_\text{eff}$ error is less than 0.01 pcm and the number of outer iterations
-would increase from 2 to 3 if the tolerance value is further tightened.
-
-\begin{figure}[htb!]
-  \centering
-  \includegraphics[width=0.6\columnwidth]{sn-tol}
-  \caption{$k_\text{eff}$ error estimates of Case 3b for a range of convergence tolerance values
-  imposed on the $S_8$ subsolver relative to the reference $k_\text{eff}$ value when
-  $\epsilon_\text{tol}=10^{-8}$. The $q=1.333$ line represents the approximate rate of
-  convergence.}
-  \label{fig:sn-tol}
-\end{figure}
-
-\section{2-D Neutronics Models \& Results} \label{sec:2d-results}
-
-\subsection{2-D Neutronics Model Setup} \label{sec:2d-model-setup}
-
-Figures \ref{fig:1/4-geom} and \ref{fig:full-geom} show the 2-D quarter-core and full-core
-\gls{MSRE} models investigated. I adapted the models from the horizontal cross section of the base
-\gls{MSRE} benchmark model developed by \cite{fratoni_molten_2020} with the following differences
-in the 2-D \gls{MSRE} model for this work:
-
-\begin{itemize}
-  \item Ignored thermal expansion of all reactor components.
-  \item Modeled the control rods as perfectly annular cylinders instead of beaded control elements
-    strung on flexible steel cables. Ignored the inconel shells encasing the control elements and
-    the steel helical cables linking the control elements together.
-  \item Set channel pitch, width, and length to 2, 0.2, and 1.2 inches, respectively.
-  \item Homogenized the INOR-8, graphite, and molten salt regions in the sample basket.
-  \item Modified the outer edge of the core graphite to be circular with no jagged edges.
-  \item Replaced partial channels with full-size channels near the edge of the core graphite.
-  \item Ignored the thermal insulation and shield structures outside the core.
-\end{itemize}
-
-The differences listed above, along with other changes in the 3-D model, result in negligible
-changes to the $k_\text{eff}$ (refer to Section 5.6) relative to the \gls{MSRE} benchmark model
-\cite{fratoni_molten_2020}.
-
+\begin{frame}
+  \frametitle{Hybrid $S_N$-Diffusion Method: 2-D Neutronics Eigenvalue Simulations}
+  \textbf{2-D Neutronics Quarter-Core \& Full-Core MSRE Models}
+  \begin{columns}
+    \column{5.5cm}
   \begin{figure}[htb!]
@@
   \end{figure}
-
+  \column{5.5cm}
   \begin{figure}[htb!]
@@
   \end{figure}
+\end{columns}
+\end{frame}
 
-\begin{table}[htb]
-  \small
-  \centering
-  \setlength\tabcolsep{4pt}
-  \caption{\gls{MSRE} molten salt composition when the $^{235}$U loading was at 65.25 kg.}
-  \begin{tabular}{l S S S S S S S}
-    \toprule
-    Element & {Li} & {Be} & {Zr} & {Hf} & {$^{234}$U} & {$^{235}$U} & {$^{236}$U} \\
-    \midrule
-    Composition [kg] & 507.27 & 293.96 & 513.97 & 0.0029 & 0.67 & 65.25 & 0.27 \\
-    \bottomrule
-  \end{tabular}
-  \begin{tabular}{l S S S S S S}
-    \toprule
-    Element & {$^{238}$U} & {Fe} & {Cr} & {Ni} & {O} & {F} \\
-    \midrule
-    Composition [kg] & 141.91 & 0.75 & 0.13 & 0.14 & 2.27 & 3103.22 \\
-    \bottomrule
-  \end{tabular}
-  \label{table:2d-salt-composition}
-\end{table}
+\begin{frame}
+  \frametitle{Hybrid $S_N$-Diffusion Method: 2-D Neutronics Eigenvalue Simulations}
+  \begin{columns}
+    \column{6.5cm}
+    \textbf{2-D Neutronics MSRE Model Setup}
+    \vspace{.2cm}
 
-INOR-8, also referred to as Hastelloy N, is a nickel-based alloy used for structural components in
-the \gls{MSRE} \cite{robertson_msre_1965}.
-Table \ref{table:2d-salt-composition} lists the molten salt composition for the 2-D models. The
-$^{235}$U loading of 65.25 kg is the amount at which the \gls{MSRE} first achieved criticality. The
-control rods consist of the original \gls{MSRE} 70 wt\% Gd$_2$O$_3$-30 wt\% Al$_2$O$_3$ mixture.
-The models impose neutron vacuum boundary conditions on the outermost boundary of the INOR-8
-vessel structure. The quarter-core model imposes neutron reflecting boundary conditions along the
-left and bottom straight edges (Figure \ref{fig:1/4-geom}). Note that the quarter-core and
-full-core models are not identical because the full-core model contains a sample basket instead of
-a control rod in the lower-right segment of Figure \ref{fig:full-geom-closeup}.
-
+    Geometry approximations:
+    \begin{itemize}
+      \item Modeled the control rods as perfectly annular cylinders
+      \item Homogenized the nickel alloy, graphite, and molten salt regions in the sample basket
+      \item Neglected the thermal insulation and shield structures outside the core
+      \item Replaced partial fuel channels and jagged graphite edges on the core periphery
+    \end{itemize}
+    All material compositions correspond to the original MSRE specifications at initial criticality.
+    \column{4.5cm}
     \begin{figure}[htb!]
       \centering
-  \includegraphics[width=0.7\columnwidth]{full-core-closeup}
+      \includegraphics[width=\columnwidth]{full-core-closeup}
       \caption{Detailed view of the control rod thimbles and sample basket in 2-D \gls{MSRE} full-core
@@
     \end{figure}
+  \end{columns}
+\end{frame}
 
-I ran neutronics simulations on the quarter- and full-core models on OpenMC under both continuous
-energy (OpenMC-CE) and multigroup (OpenMC-MG) modes, and on Moltres with the hybrid $S_N$-diffusion
-and neutron diffusion methods. The hybrid method ran with the $S_8$ angular discretization scheme
-and up to 2nd-order scattering approximations for the $S_N$ subsolver. I simulated the
-quarter-core model with and without the control rod in the air-filled rod thimble; air replaces the
-control rod material when the rod is removed. For the full-core model, I considered four rod
-configurations: no rods present, rod 1 present, rods 1 \& 2 present, and rods 1, 2 \& present.
-Other than providing a close-up view of the core center, Figure \ref{fig:full-geom-closeup} also
-presents the correction subregion on which the $S_8$ subsolver of the hybrid method generates
-drift correction parameters for the coupled diffusion solver.
-
-\FloatBarrier
-
-\subsection{2-D Neutronics Numerical Results} \label{sec:2d-nts-results}
-
-\subsubsection{2-D Quarter-Core}
-
-Table \ref{table:quarter-core} details the $k_\text{eff}$ and control rod worth estimates and error
-values relative to OpenMC-CE. For the ``No Rod'' case, OpenMC-CE estimated the $k_\text{eff}$ to be
-1.11209(43). The OpenMC-MG, hybrid, and neutron diffusion methods performed
-similarly to one another with reactivity errors ranging from 618 pcm to 773 pcm relative to
-OpenMC-CE. Based on $k_\text{eff}$ alone, the hybrid method performed marginally worse than the
-neutron diffusion method. However, for the ``Rod'' case, then neutron diffusion method saw a
-change in sign of the error to -816 pcm. The errors for OpenMC-MG and the hybrid method
-remained positive and reduced to 446 pcm and 760 pcm, respectively. Consequently, when taking the
-difference in reactivities to estimate the rod worth, the neutron diffusion method performed
-significantly worse than OpenMC-MG and the hybrid method. OpenMC-CE estimated the rod worth to be
-8370(53) pcm. The neutron diffusion method overestimated the rod worth by 1484 pcm. The hybrid
-method improved significantly on the neutron diffusion method with a rod worth estimate that is
-only 13 pcm
-higher than OpenMC-CE. OpenMC-MG reported a larger error of 172 pcm likely due to favorable error
-cancellation with the hybrid method and statistical uncertainties with the Monte Carlo methods.
-The OpenMC-CE and OpenMC-MG rod worth estimates agree within two standard deviations of each other.
-
+\begin{frame}
+  \frametitle{Hybrid $S_N$-Diffusion Method: 2-D Neutronics Eigenvalue Simulations}
+  \textbf{2-D Quarter-Core $k$ \& Rod Worth Results}
   \begin{table}[htb]
@@
       model. Error values are relative to OpenMC-CE.}
+    \setlength\tabcolsep{2pt}
     \begin{tabular}{l S[table-format=1.5(2)] S S[table-format=1.5(2)] S S[table-format=4(2)] S}
@@
   \end{table}
+  \vspace{.2cm}
 
-\begin{figure}[htb!]
+  The hybrid method produces accurate rod worth estimates while the neutron diffusion method
+      significantly overestimates rod worth.
+\end{frame}
+
+\begin{frame}
+%  \frametitle{Hybrid $S_N$-Diffusion Method: 2-D Neutronics Eigenvalue Simulations}
+%  \textbf{Normalized Fuel Channel Power Distribution}
+  \begin{columns}
+    \column{6cm}
+    \centerline{\textbf{Control Rod Withdrawn}}
+  \begin{figure}[p]
     \centering
     \includegraphics[width=\columnwidth]{msre-quarter-no-rod-power}
-  \caption{Normalized channel fission rate distribution of the 2-D \gls{MSRE} quarter-core model.
-  This figure depicts the 1/8th-core distribution due to symmetry about the diagonal.}
+    \caption{Normalized channel fission rate distribution of the 2-D \gls{MSRE} quarter-core model
+    with the rod withdrawn.}
     \label{fig:1/4-no-rod}
   \end{figure}
-
-\begin{figure}[htb!]
+    \column{6cm}
+    \centerline{\textbf{Control Rod Inserted}}
+  \begin{figure}[p]
     \centering
     \includegraphics[width=\columnwidth]{msre-quarter-rod-power}
-  \caption{Normalized channel fission rate distribution of the 2-D \gls{MSRE} quarter-core model.
-  This figure depicts the 1/8th-core distribution due to symmetry about the diagonal.}
+    \caption{Normalized channel fission rate distribution of the 2-D \gls{MSRE} quarter-core model
+    with the rod inserted.}
     \label{fig:1/4-rod}
   \end{figure}
+\end{columns}
+\end{frame}
 
-\FloatBarrier
-
-I also calculated the normalized channel fission rate distribution by integrating the fission rate
-within each fuel salt channel and normalizing the values such that the integrated fission rate in
-each channel averages to 1 as follows
-%
-\begin{align}
-  \text{Normalized fission rate of channel }i &= \frac{\text{Fission rate in channel }i}{
-  \text{Total fission rate}}\cdot\frac{\text{Total number of channels}}{
-    \text{Total fission rate in channels}} \nonumber \\
-                                              &= \frac{\sum^G_{g=1}\int_{V_i}\Sigma_{f,g}
-  \phi_g(\vec{r})}{\sum^G_{g=1}\int_V\Sigma_{f,g}\phi_g(\vec{r})}\cdot\frac{N_i}{
-  \sum^{N_i}_{i=1}\sum^G_{g=1}\int_{V_i}\Sigma_{f,g}\phi_g(\vec{r})}.
-\end{align}
-%
-The total fission rate includes fission rate contributions from peripheral fuel salt regions.
-Figures \ref{fig:1/4-no-rod} and \ref{fig:1/4-rod} show the normalized fission rate distributions
-of the quarter-core model for the ``No Rod'' and ``Rod'' cases, respectively. I presented 1/8th
-slices of the core in the figures because the distributions are symmetrical about the $y=x$ line
-outside of statistical variances for OpenMC-CE. I combined the midpoints between the centroids of
-neighboring fuel salt channels to form the diamond lattices shown in the figures.
-Each diamond element in the figures corresponds a horizontally or vertically oriented channel
-depicted in Figure \ref{fig:1/4-geom}. The lattice also provides a guide for subdividing the
-annular fuel regions around the control rod thimbles. The fission rate
-distributions omit four channels along the outer edge due to the proximity to peripheral annular
-fuel regions and the consequent difficulty in accurately tallying fission rates.
-
-In the ``No Rod'' case, the peak channel fission rate occured several channels away from the
-physical center of the reactor core. The presence of the empty (air-filled) rod thimble depressed
-fission rates near the center. The neutron diffusion and hybrid methods report similar percentage
-errors relative to OpenMC-CE except near the center. While the hybrid method benefits from drift
-corrections, the neutron diffusion method's reliance on scalar diffusion coefficients tends to
-artificially flatten neutron fluxes in near-void regions. Therefore, the largest percentage errors
-from the neutron diffusion method occurred in the second and third diamond elements
-diagonally up from the bottom which represent the annular fuel region surrounding the rod thimble.
-In the ``Rod'' case, the introduction of the control rod shifted the peak further away from the
-center as expected of a highly neutron-absorbing material. The rod further worsened fission rate
-errors for the neutron diffusion method with large errors of 1\% or more at the core center and
-periphery. On the other hand, the fission rate errors for the hybrid method largely remained below
-1\%.
-
+\begin{frame}
+  \frametitle{Hybrid $S_N$-Diffusion Method: 2-D Neutronics Eigenvalue Simulations}
+  \textbf{2-D Quarter-Core Normalized Fuel Channel Power Distribution}
   \begin{table}[htb]
@@
   \end{table}
+  \vspace{.2cm}
 
-Table \ref{table:quarter-core-power} shows the absolute mean and maximum percentage errors of the
-normalized channel fission rates from the neutron diffusion and hybrid methods relative to
-OpenMC-CE. Overall, the hybrid method reported smaller percentage errors than the neutron diffusion
-method. The absolute maximum percentage error from the neutron diffusion method jumped to 17.44\%
-as compared to the 3.08\% for the hybrid method in the ``Rod'' case. Along with the figures, these
-results demonstrate the hybrid method provides improved fission rate distribution accuracy in
-the existence of near-void and highly neutron-absorbing regions.
+  \begin{itemize}
+    \item The hybrid method improves channel fission rate estimates, especially for the rodded case.
+    \item Significant improvement in maximum percentage error of the channel fission rate.
+  \end{itemize}
+\end{frame}
 
-\FloatBarrier
-
-\subsubsection{2-D Full-Core}
-
-I performed the same $k_\text{eff}$ and fission rate distribution analyses with the 2-D full-core
-model under four configurations starting from no rod inserted to all three rods inserted. The
-$k_\text{eff}$ and error estimates in Table \ref{table:full-core-k} show the same trends observed
-with the 2-D quarter-core model. OpenMC-MG and the hybrid method exhibit moderate but consistent
-reactivity errors of approximately the 550 pcm and 800 pcm, respectively. The neutron diffusion
-method reported errors ranging from 692 pcm when no rod is inserted to -470 pcm when all three rods
-are inserted. Consequently, in Table \ref{table:full-core-worth}, OpenMC-MG and the hybrid method
-report accurate control rod worths within 1-3 standard deviations of OpenMC-CE while the neutron
-diffusion method increasingly overestimates the rod worth as more rods are inserted.
-
-\begin{table}[htb]
-  \small
-  \centering
-  \caption{$k_\text{eff}$ estimates for the 2-D full-core \gls{MSRE} model with the indicated rods
-    inserted. Error values are relative to OpenMC-CE.}
-  \setlength\tabcolsep{2pt}
-  \begin{tabular}{l S[table-format=1.5(2)] S S[table-format=1.5(2)] S S[table-format=1.5(2)] S S[table-format=1.5(2)] S}
-    \toprule
-    \multirow{3}{*}{Method} & \multicolumn{2}{c}{No Rod} & \multicolumn{2}{c}{Rod 1} & \multicolumn{2}{c}{Rod 1 \& 2} & \multicolumn{2}{c}{Rod 1, 2 \& 3} \\
-                            & {\multirow{2}{*}{$k_\text{eff}$}} & {Error} & {\multirow{2}{*}{$k_\text{eff}$}} & {Error} & {\multirow{2}{*}{$k_\text{eff}$}} & {Error} & {\multirow{2}{*}{$k_\text{eff}$}} & {Error} \\
-                            & & {[pcm]} & & {[pcm]} & & {[pcm]} & & {[pcm]} \\
-                            \cmidrule(r){1-1} \cmidrule(rl){2-3} \cmidrule(rl){4-5} \cmidrule(rl){6-7} \cmidrule(l){8-9}
-    OpenMC-CE & 1.09859(20) & {-} & 1.06980(21) & {-} & 1.04690(18) & {-} & 1.02688(18) & {-} \\
-    OpenMC-MG & 1.10637(19) & 640 & 1.07633(20) & 567 & 1.05235(20) & 495 & 1.03262(17) & 541 \\
-    Diffusion & 1.10701 & 692 & 1.07120 & 122 & 1.04415 & -252 & 1.02195 & -470 \\
-    Hybrid & 1.10843 & 808 & 1.07906 & 802 & 1.05553 & 781 & 1.03582 & 841 \\
-    \bottomrule
-  \end{tabular}
-  \label{table:full-core-k}
-\end{table}
-
+\begin{frame}
+  \frametitle{Hybrid $S_N$-Diffusion Method: 2-D Neutronics Eigenvalue Simulations}
+  \textbf{2-D Full-Core Rod Worth Results}
   \begin{table}[htb]
@@
     indicated rods inserted. Error values are relative to OpenMC-CE.}
-  \setlength\tabcolsep{5pt}
+    \setlength\tabcolsep{2pt}
     \begin{tabular}{l S[table-format=4(2)] S S[table-format=4(2)] S S[table-format=4(2)] S}
@@
   \end{table}
+  \vspace{.2cm}
 
-Plots of the normalized channel fission rate distributions for all four rod configurations are
-available in Appendix \ref{chap:2d-fiss-rate}. They are best viewed in digital pdf format with
-magnification due to the numerous fuel salt channels and small fonts.
+  The hybrid method remains effective at improving control rod worth estimates in the full-core
+  model.
+\end{frame}
 
-Table \ref{table:full-core-power} shows the absolute mean and maximum percentage errors in the
-normalized channel fission rates from the neutron diffusion and hybrid methods relative to
-OpenMC-CE. Both mean and maximum percentage errors from the neutron diffusion method steadily
-increase with the number of rods inserted. The mean percentage error from the hybrid method remains
-constant at 0.43\% while the maximum percentage error increases marginally from 1.45\% to 2.52\%
-when all three rods are inserted. These results demonstrate the hybrid method remains
-effective at improving control rod worth and fission rate distribution estimates in less symmetric
-geometries.
-
+\begin{frame}
+  \frametitle{Hybrid $S_N$-Diffusion Method: 2-D Neutronics Eigenvalue Simulations}
+  \textbf{2-D Full-Core Normalized Fuel Channel Power Distribution}
   \begin{table}[htb]
@@
     OpenMC normalized channel fission rates is 0.27\%.}
-  \setlength\tabcolsep{2.5pt}
+    \setlength\tabcolsep{1pt}
     \begin{tabular}{l S S S S S S S S}
@@
   \end{table}
-
-\FloatBarrier
-
-\section{3-D Neutronics Models \& Results} \label{sec:3d-results}
-
-\subsection{3-D Neutronics Model Setup} \label{sec:3d-model-setup}
-
-Figure \ref{fig:msre-picture} shows the vertical cross section of the actual \gls{MSRE} vessel.
-The actual design has rounded upper and lower plena to help direct salt flow through the system.
-This work uses a modified 3-D numerical \gls{MSRE} model as shown by the vertical cross section
-in Figure \ref{fig:msre-geom-vert}. The model excludes the toroidal flow distributor near the top
-of the vessel and reshapes the plena into right cylinders. It has the same horizontal cross section
-as the 2-D model (Figure \ref{fig:full-geom}) throughout most of the salt-graphite lattice. In
-addition to modifications in the 2-D model listed in Section \ref{sec:2d-model-setup}, the
-3-D model includes the following modifications:
+  \vspace{.2cm}
 
   \begin{itemize}
-  \item Removed the toroidal flow distributor near the top of the vessel.
-  \item Reshaped the upper and lower plena into right cylinders while keeping the salt volume
-    constant.
-  \item Flattened the conical tops of the graphite lattice blocks.
-  \item Regularized the complicated graphite lattice structure at the bottom.
-  \item Set a fixed outer vessel thickness of 2.54 cm.
-  \item Homogenized the lower plenum by 90.8\% salt- 9.2\% INOR-8 (vessel) by volume.
-  \item Removed the reactor access nozzle and the various components within it.
-  \item Added a graphite block at the bottom of the control rod thimbles.
+    \item Mean percentage error for the hybrid method remains consistent at 0.43 \%.
+    \item Significant improvement in maximum percentage error of the channel fission rate.
   \end{itemize}
-
-\begin{figure}[p]
-  \centering
-  \includegraphics[width=0.97\columnwidth]{msre-picture}
-  \caption{Vertical cross section of the actual \gls{MSRE} vessel. It has rounded upper and lower
-  plena, and a toroidal flow distributor that distributes the salt inflow evenly as it enters the
-  core through holes on the inside vessel surface of the distributor.}
-  \label{fig:msre-picture}
-\end{figure}
-
-\begin{figure}[p]
-  \centering
-  \includegraphics[width=\columnwidth]{msre-geom-vert}
-  \caption{Vertical cross section of the 3-D numerical \gls{MSRE} model offset by 5.08 cm to show
-  the control rod thimble and homogenized sample basket. The control rod is in its fully inserted
-  position. The homogenized lower plenum consists of 90.8\% salt and 9.2\% INOR-8 alloy by volume.}
-  \label{fig:msre-geom-vert}
-\end{figure}
-
-Table \ref{table:reactor-dimensions} lists relevant dimensions of the 3-D \gls{MSRE} model.
-All material compositions remain
-the same as the 2-D models and the \gls{MSRE} benchmark specifications \cite{fratoni_molten_2020}.
-All 3-D neutronics simulations in this section used the molten salt composition with 65.25 kg
-$^{235}$U loaded (Table \ref{table:2d-salt-composition}), except the simulations for temperature
-reactivity coefficient calculations which had a higher $^{235}$U loading of 67.00 kg (Table
-\ref{table:67-salt-composition}).
-
-\begin{table}[htb]
-  \centering
-  \caption{List of reactor dimensions.}
-  \begin{tabular}{l l}
-    Height [cm] & 215 \\
-    TODO: Populate table & \\
-  \label{table:reactor-dimensions}
-\end{table}
-
-\begin{table}[htb]
-  \small
-  \centering
-  \setlength\tabcolsep{4pt}
-  \caption{\gls{MSRE} molten salt composition when the $^{235}$U loading was at 65.25 kg.}
-  \begin{tabular}{l S S S S S S S}
-    \toprule
-    Element & {Li} & {Be} & {Zr} & {Hf} & {$^{234}$U} & {$^{235}$U} & {$^{236}$U} \\
-    \midrule
-    Composition [kg] & 507.42 & 293.96 & 513.97 & 0.0029 & 0.68 & 67.00 & 0.28 \\
-    \bottomrule
-  \end{tabular}
-  \begin{tabular}{l S S S S S S}
-    \toprule
-    Element & {$^{238}$U} & {Fe} & {Cr} & {Ni} & {O} & {F} \\
-    \midrule
-    Composition [kg] & 142.02 & 0.75 & 0.13 & 0.14 & 2.27 & 3104.22 \\
-    \bottomrule
-  \end{tabular}
-  \label{table:67-salt-composition}
-\end{table}
-
-All neutronics simulations ran on OpenMC under continuous energy mode and on Moltres with the
-hybrid $S_N$-diffusion method. The hybrid simulations ran with a maximum stabilization factor of
-$c=250$, a void constant $\varsigma=0.5$, and a uncorrected diffusion coefficient cap of $D_g=2.5$.
-Refer to Sections \ref{sec:void-treatment} and \ref{sec:diffcoef-cap} for the relevant discussion
-on void treatment for near-void regions and the impact of large diffusion coefficient values on
-solver performance. As for the standard neutron diffusion method, it either converges very slowly
-or does not converge at all when $D_g$ values are uncapped due to excessive diffusive effects in
-the air-filled control rod thimbles. Group-wise $D_g$ values, generated using OpenMC, ranged from
-100 cm to 850 cm in the air regions as compared to the typical 0.5 cm to 2.0 cm range observed in
-non-air regions. Therefore, all neutron diffusion method simulation results in the following
-section were measured with the diffusion coefficient capped at $D_g=2.5$.
-
-Unlike the 2-D hybrid simulations which ran with $S_8$ angular discretization, the 3-D hybrid
-simulations ran with $S_6$ angular discretization due to significantly
-larger computational and memory requirements demanded by the 3-D simulations. Further discussion on
-computational performance may be found at the end of Section \ref{sec:3d-nts}. All simulations,
-except those for temperature reactivity coefficient calculations, ran on isothermal 3-D models at
-911 K.
-
-\subsection{3-D Numerical Results} \label{sec:3d-nts}
-
-\subsubsection{Initial Criticality Configuration}
-
-The \gls{MSRE} achieved initial criticality with Rod 1 inserted to a height of 46.6 inches relative
-to the fully inserted position. For reference, the rods are at a height of 51 inches when fully
-withdrawn. Table \ref{table:initial-crit}
-shows the $k_\text{eff}$ values from \gls{MSRE} experimental data, numerical benchmark data
-\cite{fratoni_molten_2020}, and this work. All numerical models overestimate the $k_\text{eff}$
-relative to the experimental value. In the \gls{MSRE} numerical benchmark report, the authors
-attribute this $k_\text{eff}$ discrepancy largely to discrepancies in the carbon cross section data
-in nuclear data libraries. They noted similar $k_\text{eff}$ overestimations by 1-2\% in benchmark
-results for the HTTR, a graphite-moderated, helium-cooled reactor. The OpenMC and Moltres
-simulations in this work used the same
-ENDF/B-VII.1 nuclear data library as the \gls{MSRE} numerical benchmark report.
-
-\begin{table}[htb]
-  \centering
-  \caption{$k_\text{eff}$ values from \gls{MSRE} experimental data, the \gls{MSRE} numerical
-  benchmark \cite{fratoni_molten_2020}, and the OpenMC and Moltres models in this work.}
-  \begin{tabular}{l S[table-format=1.5(2)]}
-    \toprule
-     & {$k_\text{eff}$} \\
-     \cmidrule(l){2-2}
-    \gls{MSRE} experimental data & 1.00000(420) \\
-    Serpent 2 (Numerical benchmark) & 1.02132(3) \\
-    OpenMC (This work) & 1.01308(20) \\
-    Hybrid (This work) & 1.01955 \\
-    Diffusion (This work) & 1.01885 \\
-    \bottomrule
-  \end{tabular}
-  \label{table:initial-crit}
-\end{table}
-
-Of the four $k_\text{eff}$ values, we expect to see the closest agreement between the Serpent 2
-model from the numerical benchmark and OpenMC from this work. However, the OpenMC \gls{MSRE} model
-differs from the Serpent 2 \gls{MSRE} models due to geometry modifications listed in Sections
-\ref{sec:2d-model-setup} and \ref{sec:3d-model-setup}. The benchmark report quantified the impact
-of several geometry modifications, namely the removal of the flow distributor, the homogenization
-of the sample basket, and the removal of the thermal shield. These modifications contribute to
--$0.098\%$, -$0.037\%$, and -$0.885\%$ change in the $k_\text{eff}$, respectively, for a final
-$k_\text{eff}$ estimate of $1.01093(3)$. This brings the OpenMC $k_\text{eff}$ estimate of
-$1.01308(20)$ to closer agreement with the numerical benchmark value.
-
-The hybrid method $k_\text{eff}$ estimate is 0.00647 higher than the OpenMC estimate. For this
-initial criticality configuration, the neutron diffusion method $k_\text{eff}$ estimate is closer
-to OpenMC than the hybrid method. Overall, the OpenMC and Moltres models show good agreement with
-the Serpent 2 model from the numerical benchmark.
-Figures * show the neutron group fluxes on the interior of the 3-D models. The salt-graphite
-lattice structure is visible from the group flux distributions. Group 7 and 8 neutron fluxes are
-particularly small near the control rods due to strong thermal neutron capture by Gadolinium in the
-rods.
-
-\subsubsection{Control Rod Worth}
-
-After achieving initial criticality, \gls{MSRE} researchers ran control rod calibration experiments
-with the \gls{MSRE} at zero-power (typically at about 10 W power output)
-\cite{prince_zero-power_1968}. They measured differential rod worths by measuring the time taken
-for the reactor power to increase by 20 W following a prescribed withdrawal of Rod 1 and applying
-the inhour equation to obtain the reactivity inserted. They repeated this process at different rod
-heights after successive additions of highly enriched uranium capsules. The final integral rod
-worth curve was normalized to the critical $^{235}$U loading of 65.25 kg
-\cite{prince_zero-power_1968}.
-
-For control rod worth calculations, the OpenMC and Moltres simulations ran with Rod 1 at fourteen
-different rod positions from full insertion (0 inch) to full withdrawal (51 inches). The rod
-positional limits of 0 inch and 51 inches represents the full range of deliberate rod motion
-allowable by the control rod drive motor \cite{robertson_msre_1965}.
-
-Figure \ref{fig:rod-worth} shows the integral rod worth curves from the \gls{MSRE} experimental
-data and the OpenMC and Moltres numerical models. By taking the $k_\text{eff}$ when Rod 1 is fully
-inserted as reference, all rod worth curves start at 0 pcm. The change in integral rod worth,
-represented by the gradient in the plot, is steepest as the tip of the rod passes through the
-center of the core. The hybrid method performs well in reproducing the rod worth curve from OpenMC;
-all hybrid method rod worth values fall within one or two sigma range of OpenMC.
-
-Due to the experimental procedure of computing integral rod worth from cumulative differential rod
-worth measurements, the uncertainty range of the \gls{MSRE} data scales with rod position.
-Both OpenMC and the hybrid method report consistently larger integral rod worths than the
-\gls{MSRE} data across all rod positions. Nevertheless, they are in good agreement with the
-\gls{MSRE} data as compared to the neutron diffusion method, which significantly overestimates the
-integral rod worth.
-
-Table \ref{table:rod-worth} shows the total rod worth of Rod 1.
-
-\begin{figure}[p]
-  \centering
-  \includegraphics[width=0.8\columnwidth]{rod-worth-curve}
-  \caption{Reactivity inserted by Rod 1 at various rod positions relative to the full insertion.}
-  \label{fig:rod-worth}
-\end{figure}
-
-\begin{table}[htb]
-  \centering
-  \caption{Total rod worth of Rod 1 when fully inserted.}
-  \begin{tabular}{l S[table-format=4(2)] S}
-    \toprule
-     & {$\rho_\text{worth}$ [pcm]} & {Percentage discrepancy [\%]}\\
-     \cmidrule(l){2-2}
-    \gls{MSRE} experimental data & 2250(67) & {-}\\
-    OpenMC (This work) & 2364(44) & 5.1 \\
-    Hybrid (This work) & 2345 & 4.2 \\
-    Diffusion (This work) & 2725 & 21.1 \\
-    \bottomrule
-  \end{tabular}
-  \label{table:rod-worth}
-\end{table}
-
-\subsubsection{Temperature Reactivity Coefficient}
-
-\subsubsection{Computational Performance}
-
-\section{Summary}
-
-%The hybrid method does not alter the $k_\text{eff}$ and flux estimates relative to the neutron
-%diffusion method in cases where the neutron diffusion method provides
+\end{frame}

### 2024-12-04 presentation/3d-transient.tex
Presentation commit

\begin{frame}
  \frametitle{Time-Dependent Multiphysics Simulations with the Hybrid Method}
  I investigated two transient scenarios based on actual MSRE experiments:
  \begin{block}{\textbf{MSRE Rod Drop Experiment}}
    \begin{itemize}
      \item Neutronic response of an initially critical, zero-power MSRE to a
        rod drop of Rod 1 \cite{prince_zero-power_1968}
      \item Corresponds to a reactivity withdrawal of -1500 pcm
      \item Requires delayed neutron precursor (DNP) modeling
    \end{itemize}
  \end{block}
  \begin{block}{\textbf{MSRE Reactivity Insertion Experiment}}
    \begin{itemize}
      \item Coupled power response of the MSRE initially at 1 MW to a rod
        withdrawal \cite{engel_zero-power_1972}
      \item Reactivity insertion of 24.8 pcm
      \item MSRE fueled with $^{233}$U
      \item Requires DNP and temperature advection-diffusion modeling
    \end{itemize}
  \end{block}
\end{frame}

\begin{frame}
  \frametitle{MSRE Rod Drop Experiment}
  \textbf{Rod Height}
  \begin{figure}[htb!]
    \centering
    \includegraphics[width=0.7\columnwidth]{rod-height}
    \caption{Evolution of rod height evaluated at each timestep within the first second of the rod
    drop simulation. The rod reaches its fully inserted height at $t=0.4046$ s.}
    \label{fig:rod-height}
  \end{figure}
\end{frame}

\begin{frame}
  \frametitle{MSRE Rod Drop Experiment}
  \textbf{Nested Coupling Iteration Structure for the Rod Drop Simulation}
  \begin{figure}[t]
    \tikzstyle{every node}=[font=\small]
    \centering
    \includegraphics[width=\columnwidth]{images/nest-1}
    \caption{The nested iteration structure coupling the $S_N$, neutron diffusion, and \gls{DNP}
    solvers for the rod drop simulation using the hybrid $S_N$-diffusion method.}
    \label{fig:rod-drop-coupling}
  \end{figure}
\end{frame}

\begin{frame}
  \frametitle{MSRE Rod Drop Experiment}
  \textbf{Integral Neutron Count Following Rod Drop}
  \begin{figure}[t]
    \centering
    \includegraphics[width=0.55\columnwidth]{integral-count}
    \caption{Integral neutron count during the rod drop experiment from \gls{MSRE} experimental data
    and hybrid method numerical results.}
    \label{fig:integral-count}
  \end{figure}
  \begin{itemize}
    \item Moltres underpredicted the integral neutron count rate by approximately 20 \%.
    \item Potential sources of error
    \begin{itemize}
      \item Uncertainty in the initial neutron count rate
      \item Miscalculation of -1500 pcm reactivity withdrawal
    \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{MSRE Reactivity Insertion Experiment}
  \textbf{Nested Coupling Iteration Structure for the Reactivity Insertion Simulation}
  \begin{figure}[t]
    \centering
    \includegraphics[width=\columnwidth]{images/nest-2}
    \caption{The nested iteration structure coupling the $S_N$, neutron diffusion,
    in-core \gls{DNP}-temperature, and ex-core
    solvers for the reactivity insertion simulation using the hybrid $S_N$-diffusion method.}
    \label{fig:insertion-coupling}
  \end{figure}
\end{frame}

### 2024-12-04 presentation/3d.tex
Presentation commit

--- 2024-11-19 msre.tex
@@
-In Chapter \ref{chap:hybrid}, I laid out the mathematical background and numerical implementation
-of the hybrid $S_N$-diffusion method for improving neutronic accuracy in reactor systems containing
-highly neutron-absorbing regions. I developed and implemented the diffusion and drift correction
-schemes for the hybrid method in Python and using Moltres, respectively. This chapter demonstrates
-the hybrid method through 1-D, 2-D, and 3-D $k$-eigenvalue simulations of the \gls{MSRE}.
-I employed the OpenMC Monte Carlo neutral particle transport software \cite{romano_openmc:_2015} to
-generate group constants for the hybrid method and reference solutions for method verification.
-The 1-D models serve to verify the newly-implemented $S_N$ and hybrid methods
-against OpenMC as well as to analyze the impact of key input parameters on the hybrid method. The
-2-D models test the hybrid method performance through increased spatial dimensionality and reduced
-geometric asymmetry relative to the 1-D models. Lastly, the 3-D models serve as neutronics model
-validation against the \gls{MSRE} benchmark model by Fratoni et al.\ \cite{fratoni_molten_2020}, a
-testbed for \gls{HPC} scaling tests, and preparation for the time-dependent multiphysics
-simulations in Chapter \ref{chap:transient}.
-
-Section \ref{sec:msre-gc} details the OpenMC model setup for group constant generation. Section
-\ref{sec:nts-methods} describes and labels the different neutronics methods employed for the
-$k$-eigenvalue simulations. Section \ref{sec:test-metrics} details basic test metrics relevant to
-neutronics verification and validation. Sections \ref{sec:1d-results}, \ref{sec:2d-results}, and
-\ref{sec:3d-results} discuss the model setup, results, and discussions for the 1-D, 2-D, and 3-D
-\gls{MSRE} models, respectively.
-
-\section{Group Constant Data Generation on OpenMC} \label{sec:msre-gc}
-
-Group constants relevant to the neutron diffusion, $S_N$ neutron transport, or both
-methods are:
-%
-\begin{align*}
-  \Sigma_{t,g}:& \mbox{ Macroscopic total cross section in group $g$,}
-  \Sigma_{r,g}:& \mbox{ Macroscopic removal cross section in group $g$,}
-  \Sigma_s^{g'\rightarrow g}:& \mbox{ Macroscopic group-to-group scattering cross section matrix,}
-  \Sigma_{s,l}^{g'\rightarrow g}:& \mbox{ $l$-th Legendre moment of the macroscopic
-  group-to-group scattering cross section matrix,}
-  \Sigma_{sp,l}^{g'\rightarrow g}:& \mbox{ $l$-th Legendre moment of the macroscopic
-  group-to-group scattering production cross section matrix,}
-    D_g:& \mbox{ $P_1$-based diffusion coefficient in group $g$,}
-  \nu\Sigma_{f,g}:& \mbox{ Product of the average number of neutrons produced per fission and the
-  macroscopic fission cross section in group $g$,}
-    \chi_g:& \mbox{ Neutron fission spectrum in group $g$.}
-\end{align*}
-%
-I generated the group constants using OpenMC's multigroup cross section generation capability
-\cite{boyd_multigroup_2019}. A group constant postprocessing script for Moltres
-(\texttt{moltres\_xs.py}) parses OpenMC simulation output files and rewrites the group constant
-data into a JSON format file suitable for Moltres simulations. While OpenMC provides most of the
-group constants directly, the postprocessing script calculates $\Sigma_{r,g}$ as
-%
-\begin{align}
-  \Sigma_{r,g} =& \sum^G_{g'\neq g}\Sigma_s^{g\rightarrow g'}+\Sigma_{a,g}-\left(\Sigma_{sp}^{g
-    \rightarrow g} - \Sigma_s^{g\rightarrow g}\right)
-  \shortintertext{where}
-      \Sigma_{a,g} =& \mbox{ macroscopic absorption cross section in group $g$,} \nonumber \\
-      \Sigma_{sp}^{g\rightarrow g} =& \mbox{ macroscopic scattering production cross section from
-      group $g$ to $g$.} \nonumber
-\end{align}
-%
-$\Sigma_{r,g}$ primarily represents the loss of neutrons from group $g$ through outscattering and
-absorption. $\Sigma_{sp}^{g\rightarrow g}$ incorporates neutron multiplication effects from neutron
-knockout reactions into the scattering cross section. Neutronics codes commonly identify neutron
-knockout reactions as scattering reactions, but $\Sigma_{r,g}$ is a convenient parameter in which
-to incorporate neutron knockout effects in the neutron diffusion method. I ran all test cases
-for $S_N$ and hybrid method calculations with scattering cross sections approximated by up to 2nd
-Legendre moments ($L=2$).
-
-OpenMC uses the $P_1$ flux-limited formulation \cite{pomraning_flux-limited_1984} by default for
-calculating $D_g$ as follows
-%
-\begin{align}
-  D_g =& \frac{1}{3\Sigma_{tr,g}},
-  \shortintertext{where}
-  \Sigma_{tr,g} =& \frac{\langle\Sigma_{t,g}\phi_g\rangle-\langle\Sigma_{s1,g}\phi_g\rangle}
-  {\langle\phi_g\rangle}, \nonumber \\
-  \langle\Sigma_{t,g}\phi_g\rangle =& \int_{r\in V}dr \int_{4\pi}d\Omega\int^{E_{g-1}}_{E_g}dE\
-  \Sigma_{t,g}(r,E)\Psi(r,E,\Omega), \nonumber \\
-  \langle\Sigma_{s1,g}\phi_g\rangle =& \int_{r\in V}dr \int_{4\pi}d\Omega\int^{E_{g-1}}_{E_g}dE
-  \int_{4\pi}d\Omega'\int^{\infty}_0dE'\int^1_{-1}d\mu\ \mu\Sigma_s(r,E'\rightarrow E,\Omega'\cdot
-  \Omega)\phi(r,E',\Omega'), \nonumber \\
-  \langle \phi \rangle =& \int_{r\in V}dr\int_{4\pi}d\Omega\int^{E_{g-1}}_{E_g}dE\ \Psi(r,E,\Omega)
-  .\nonumber
-\end{align}
-
-I ran all \gls{MSRE} neutronics simulations with the eight neutron energy group structure proposed
-by Jaradat for \gls{MSRE} analysis \cite{jaradat_development_2021-1}.
-Table \ref{table:energy-group} shows the upper neutron energy bounds of the eight-group structure.
-I selected this energy group structure as a compromise between accuracy and computational expense
-among the various group structures investigated by Jaradat ranging from four to sixteen groups.
-
-\begin{table}[htb]
-  \centering
-  \caption{Neutron energy group structure in this work. Originally devised by Jaradat
-  \cite{jaradat_development_2021-1}.}
-  \begin{tabular}{r S}
-    \toprule
-    Group & {Upper energy bound [eV]} \\
-    \midrule
-    1 & 2.000$\times 10^7$ \\
-    2 & 1.353$\times 10^6$ \\
-    3 & 6.734$\times 10^4$ \\
-    4 & 9.118$\times 10^3$ \\
-    5 & 1.487$\times 10^2$ \\
-    6 & 4.000$\times 10^0$ \\
-    7 & 6.250$\times 10^{-1}$ \\
-    8 & 8.000$\times 10^{-2}$ \\
-    \bottomrule
-  \end{tabular}
-  \label{table:energy-group}
-\end{table}
-
-I set the material compositions and densities of every component in the \gls{MSRE} models using
-reference data from \gls{MSRE} benchmark specifications collated by Fratoni et al.
-\cite{fratoni_molten_2020}. All OpenMC models utilize the ENDF/B-VII.1 nuclear data library
-\cite{chadwick_endf/b-vii.1_2011} and include thermal scattering cross section data for beryllium
-and carbon target nuclei in the fuel salt and graphite moderator. All models have a uniform
-temperature of 900 K unless otherwise stated. Detailed information pertaining specifically to the
-1-D, 2-D, and 3-D OpenMC models may be found in the respective subsections below.
-
-\section{Neutronics Methods} \label{sec:nts-methods}
-
-I ran all cases on OpenMC to generate the required group constants and reference $k_\text{eff}$
-values to assess the accuracy of the deterministic methods ($S_N$, neutron diffusion, hybrid 
-methods) against OpenMC in continuous-energy (OpenMC-CE) and multigroup (OpenMC-MG) modes. Table
-\ref{table:var} details how OpenMC and the deterministic methods represent the position, direction
-of travel, neutron energy, and angle-dependence in $\Sigma_s$. The hybrid $S_N$-diffusion method
-features representations from either $S_N$ or neutron diffusion depending on the subregion under
-consideration. Comparing OpenMC-MG results with OpenMC-CE results allows us
-to quantify errors arising from neutron energy group discretization and the scattering cross
-section approximations.
-
-\begin{table}[htb!]
-  \centering
-  \footnotesize
-  \caption{Variable representations in OpenMC under continuous-energy (OpenMC-CE) and multigroup
-  (OpenMC-MG) modes, and in the $S_N$ neutron transport and neutron diffusion.}
-  \begin{tabular}{l l l l l}
-    \toprule
-    Variable & OpenMC-CE & OpenMC-MG & $S_N$ Transport & Diffusion \\
-    \midrule
-    Position, $\bm{r}$ & Continuous & Continuous & Discrete & Discrete \\
-    Direction of travel, $\bm{\hat{\Omega}}$ & Continuous & Continuous & Discrete & N/A \\
-    Energy, $E$ & Continuous & Discrete & Discrete & Discrete \\
-    Angle-dependence in $\Sigma_s$ & Continuous & $L=2$ & $L=2$ & N/A \\
-    \bottomrule
-  \end{tabular}
-  \label{table:var}
-\end{table}
-
-The rest of this section pertains to the deterministic neutronics methods used in this work for
-modeling the \gls{MSRE}.
-For the 1-D models, I performed neutronics calculations with the $S_N$, neutron diffusion, and
-hybrid method implementations on Moltres and Python. Table \ref{table:nts-methods} highlights the
-differences between the two implementations. The $S_N$ neutron transport method implementation on
-Moltres follows the methodology for solving the \gls{SAAF} formulation of the $S_N$ equations as
-described in Section \ref{sec:saaf} while the implementation in Python follows the methodology
-for solving the original first-order formulation as described in Section \ref{sec:python-sn}.
-
-The implementations on Moltres discretize the problem domain in space using the finite element
-method while the implementations in Python use finite difference or diamond difference schemes. The
-most crucial difference lies in the transport correction formulations for the hybrid method. The
-Moltres implementation uses the drift correction scheme described in Section
-\ref{sec:drift-correction} while the Python implementation uses the diffusion correction
-scheme described in Section \ref{sec:diffusion-correction}.
-
-In this chapter, I refer to the methods and results from the Moltres and Python implementations
-with the Moltres- or Python- prefix, i.e., Moltres-$S_N$, Moltres-diffusion, Moltres-hybrid,
-Python-$S_N$, Python-diffusion, and Python-hybrid.
-
-For 2- and 3-D models, I abandoned the diffusion correction formulation in favor of the drift
-correction formulation due to superior numerical performance in the latter. Readers may find the
-analysis supporting this decision in the analysis of 1-D results in Section \ref{sec:1d-results}.
-
-\begin{table}[htb!]
-  \centering
-  \footnotesize
-  \caption{Implementation differences of neutronics methods on Moltres and Python.}
-  \begin{tabular}{p{0.14\linewidth} p{0.25\linewidth} p{0.15\linewidth} p{0.15\linewidth} p{0.17\linewidth}}
-    \toprule
-    Implementations & Spatial discretization & $S_N$ formulation & $S_N$ acceleration &
-    Hybrid transport\newline correction \\
-    \midrule
-    Moltres & Finite element method & SAAF & Nonlinear diffusion acceleration & Drift correction \\
-    Python & Finite difference (Diffusion)\newline Diamond difference ($S_N$) & First-order & None & Diffusion correction \\
-    \bottomrule
-  \end{tabular}
-  \label{table:nts-methods}
-\end{table}
-
-\section{Basic Test Metrics} \label{sec:test-metrics}
-
-In this section, I define several metrics for comparing simulation results among the various
-neutronics methods.
-
-For steady-state eigenvalue calculations, all solvers directly provide the effective neutron
-multiplication factor ($k_\text{eff}$), the rate of neutron population increase between successive
-neutron generations. Reactivity is defined as
-%
-\begin{gather}
-  \rho \equiv \frac{k_\text{eff}-1}{k_\text{eff}}.
-\end{gather}
-%
-Differences in reactivity, for either comparisons between neutronics methods or control rod worth
-calculations, can be calculated as
-%
-\begin{gather}
-  \Delta\rho_{1,2} = \rho_1 - \rho_2 =
-  \frac{k_{\text{eff},1}-k_{\text{eff},2}}{k_{\text{eff},1}k_{\text{eff},2}}.
-\end{gather}
-
-Spatially dependent variables (e.g., neutron group flux, neutron drift) are sampled at fixed 0.1 cm
-intervals along the specified lines (e.g., horizontal, diagonal, etc.) regardless of the actual
-mesh resolution.
-
-\section{1-D Neutronics Models \& Results} \label{sec:1d-results}
-
-This section presents the models, results, and discussion for the 1-D test cases.
-
-\subsection{1-D Neutronics Model Setup \& Geometries}
-
-I designed six 1-D test cases with increasing complexity to verify the $S_N$ and hybrid
-$S_N$-diffusion implementations and to test the performance of the hybrid
-method in response to various geometrical features. The last two cases resemble the
-reference \gls{MSRE} design \cite{robertson_msre_1965}, which has centrally located control rods
-and air-filled rod guide tubes. Figure \ref{fig:case-geom} shows the geometries of Cases 1a to
-3b. All geometries have reflective boundary conditions at $x=0$ cm, thereby
-forming half-core or repeating unit cell models. Cases 1a and 1b are repeating unit
-cell models with reflecting boundaries on the right-side boundaries. Cases 2a, 2b, 3a, and 3b
-are 1-D half-core models with vacuum boundaries on the right-side boundaries.
-
-\begin{figure}[htb!]
-  \centering
-  \includegraphics[width=0.8\columnwidth]{case-geometry}
-  \caption{Geometries of the 1-D test cases. The material labeled ``mixture'' represents a
-    homogeneous mixture of fuel and graphite at a ratio of 22.5\%-77.5\% by volume. All geometries
-    have reflective boundary conditions on the boundary at $x=0$ cm. The right-side boundaries are
-    reflective for Cases 1a and 1b, and vacuum for Cases 2a, 2b, 3a, and 3b.}
-  \label{fig:case-geom}
-\end{figure}
-
-Cases 1a and 1b are simple test cases containing neutron multiplying regions. Cases 1a and 1b serve
-to verify basic neutronics phenomena modeling (e.g., fission, scattering, absorption) of the newly
-implemented $S_N$ and neutron diffusion methods in Python and $S_N$ method in Moltres. I did not
-apply the hybrid method on these cases. Cases 2a and 2b represent 1-D models of the \gls{MSRE} with
-air-filled control rod guide tube regions, a homogenized fuel-graphite lattice regions, and outer
-vessel regions. Case 2b additionally contains a 0.5cm-thick control rod region to contrast with
-Case 2a for control rod worth calculations. Cases 3a and 3b retain the heterogeneous fuel-graphite
-lattice geometry present in the \gls{MSRE}.
-
-In my preliminary investigations, the original control rod material composition
-(70 wt\% Gd$_2$O$_3$-30 wt\% Al$_2$O$_3$) led to excessively large control rod worth estimates of
-approximately 55000 pcm in the 1-D models, effectively halving the $k_\text{eff}$. The worth
-estimates are ten times larger
-than the 5500 pcm expected from full rod insertion of all three rods in the actual \gls{MSRE}
-\cite{fratoni_molten_2020}. Therefore, I greatly reduced the proportion of Gd$_2$O$_3$ in the
-control rod to 0.35 wt\% for a fairer assessment of rod worth in the 1-D models. This change
-brought the new rod worth estimates down to below 20000 pcm.
-
-\subsection{1-D Neutronics Model Mesh Convergence Tests}
-
-I ran mesh convergence studies based on Case 3b for the neutron diffusion and $S_N$ neutron
-transport methods. Given that the hybrid $S_N$-diffusion method comprises of a combination of those
-methods, I found hybrid simulations were fully spatially resolved on meshes that are
-appropriately resolved for both neutron diffusion and $S_N$ simulations.
-
-\begin{figure}[htb!]
-  \centering
-  \begin{subfigure}[b]{0.48\columnwidth}
-    \centering
-    \includegraphics[width=\columnwidth]{diffusion-mesh-convergence-k}
-    \caption{Neutron diffusion method}
-    \label{fig:diff-mesh-k}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}[b]{0.48\columnwidth}
-    \centering
-    \includegraphics[width=\columnwidth]{sn-mesh-convergence-k}
-    \caption{$S_8$ neutron transport method}
-    \label{fig:sn-mesh-k}
-  \end{subfigure}
-  \caption{Convergence of multiplication factor ($k_\text{eff}$) estimates for Case 3b across five
-    levels of mesh refinement. Higher refinement levels correspond to smaller mesh element sizes.}
-  \label{fig:mesh-k}
-\end{figure}
-
-Figures \ref{fig:diff-mesh-k} and \ref{fig:sn-mesh-k} show the convergence of $k_\text{eff}$
-estimates observed for Case 3b simulations with neutron diffusion and $S_8$ neutron transport
-methods. The $k_\text{eff}$ values plateau at 0.9713203 and 0.9813790 for the neutron diffusion and
-$S_8$ methods, respectively. By mesh refinement level 2, both methods fall within $5\times 10^{-7}$
-range of the $k_\text{eff}$ value at level 4.
-
-Figures \ref{fig:diff-mesh-flux} and \ref{fig:sn-mesh-flux} show the absolute differences in the
-group flux distributions at mesh refinement levels 1 to 3 relative to the reference group flux
-distribution at level 4 (maximum refinement). I omitted data from level 0 to improve the fidelity
-of the plots as flux differences drop significantly by one order of magnitude with each level of
-refinement.
-With these results, I concluded that the mesh resolution at level 2 is a sufficient guideline for
-the mesh resolutions used in the subsequent 1-D analyses. The absolute differences in both
-$k_\text{eff}$ and group flux distributions at level 2 remain below $5\times10^{-7}$ relative to
-level 4.
-
-\begin{figure}[htb!]
-  \centering
-  \includegraphics[width=\columnwidth]{diffusion-mesh-convergence-flux}
-  \caption{Absolute difference in group fluxes ($\phi_g$) from the neutron diffusion method for
-  mesh refinement levels 1 to 4 relative to the reference group flux distributions at mesh
-  refinement level 5.}
-  \label{fig:diff-mesh-flux}
-\end{figure}
-
-\begin{figure}[htb!]
-  \centering
-  \includegraphics[width=\columnwidth]{sn-mesh-convergence-flux}
-  \caption{Absolute difference in group fluxes ($\phi_g$) from the $S_8$ neutron transport method
-  for mesh refinement levels 1 to 4 relative to the reference group flux distributions at mesh
-  refinement level 5.}
-  \label{fig:sn-mesh-flux}
-\end{figure}
-
-\FloatBarrier
-
-\subsection{1-D Neutronics Numerical Results \& Discussion}
-
-\subsubsection{Reactivity}
-
-Figure \ref{fig:1d-rho} shows the difference in reactivities of OpenMC-MG, $S_8$,
-neutron diffusion, and hybrid methods relative to OpenMC-CE. The standard deviations of
-reactivity values from OpenMC-CE are approximately 40 pcm for Cases 1a and 1b and 60 pcm for the
-rest as depicted by the blue highlighted area in Figure \ref{fig:1d-rho}. For Cases 1a and 1b, the
-OpenMC-MG, $S_8$, and neutron diffusion methods show excellent agreement with OpenMC-CE as the
-reactivity values fall within the one or two standard deviation range. The Case 1 results show
-the Moltres and Python implementations are consistent with each other given that they are spatially
-well-resolved.
-
-\begin{figure}[htb!]
-  \centering
-  \includegraphics[width=\columnwidth]{rho}
-  \caption{Difference in reactivity $\rho$ of all neutronics methods investigated relative
-  to OpenMC-CE.}
-  \label{fig:1d-rho}
-\end{figure}
-
-For all cases, the OpenMC-MG and the $S_8$ methods show consistent agreement with one another. 
-While they deviate from OpenMC-CE by approximately 350 pcm for Cases 2a and 3a, they fall within
-two standard deviations of OpenMC-CE for Cases 2b and 3b. When increasing the maximum Legendre
-polynomial order to approximate the angular dependence in $\Sigma_s^{g'\rightarrow g}$ from
-2nd-order to 3rd-order, the reactivity changed by only 1 pcm. Therefore, I attribute the reactivity
-discrepancies to the eight neutron energy group structure which remains the
-only significant difference between OpenMC-CE and the multigroup methods.
-
-The neutron diffusion and hybrid methods agree closely with one another for Cases 2a and 3a which
-exclude the control rod region. This shows that the hybrid method provides similar $k_\text{eff}$
-estimates to the neutron diffusion method in 1-D simulations without highly neutron-absorbing
-regions. Compared to the OpenMC-MG and $S_8$ reactivity values, the neutron diffusion and hybrid
-method reactivity values agree closer with the reference OpenMC-CE value. However, this is likely
-due to favorable error cancellation; Figures \ref{fig:3a-flux} and \ref{fig:3a-flux-diff} show
-OpenMC-MG and the $S_8$ methods reproduce
-the OpenMC-CE flux distribution more accurately than the neutron diffusion and hybrid methods.
-
-In Cases 2b and 3b, the neutron diffusion method largely fails to accurately capture the
-effect of introducing the control rod region as evidenced by the -1500 pcm and -1150 pcm
-discrepancies relative to OpenMC-CE. The hybrid method fares better with -400 pcm and -300 pcm
-discrepancies.
-
-\begin{figure}[htb!]
-  \centering
-  \includegraphics[width=\columnwidth]{worth}
-  \caption{Percentage difference in rod worth for Cases 2 and 3 of all neutronics methods
-  investigated relative to OpenMC-CE.}
-  \label{fig:1d-worth}
-\end{figure}
-
-\subsubsection{Control rod worth}
-
-Moving the discussion to control rod worths, figure \ref{fig:1d-worth} shows the percentage
-difference in rod worths for Cases 2 and 3 of OpenMC-MG, $S_8$, neutron diffusion, and hybrid
-methods relative to OpenMC-CE. I calculated the rod worth estimates from each method by taking the
-difference in reactivity between the ``rod out'' (a) and ``rod in'' (b) configurations of Cases 2
-and 3. All neutronics methods overestimate the rod worth relative to OpenMC-CE, resulting in
-positive percentage difference values observed in the figure.
-The neutron diffusion methods clearly show up as outliers with rod worth estimates in excess
-of 8\%. Most of the remaining methods cluster around 2.5\% and 3\% percentage difference for Cases
-2 and 3 with the exception of the Python-hybrid method for Case 3.
-
-Given that the hybrid methods rely on transport corrections derived from the $S_N$ method, the
-$S_8$ rod worth estimates serve as the
-reference point for hybrid method verification. Additionally, errors arising from the multigroup
-approximation affect both hybrid and $S_N$ simulations to similar degrees. While the Python-hybrid
-method produces the lowest percentage difference relative to OpenMC-CE for Case 3, the
-Python-hybrid method deviates from the $S_8$ ``multigroup'' reference rod worths more than the
-Moltres-hybrid method. Nevertheless, both hybrid method implementations provide significant
-improvements in rod worth estimates over the neutron diffusion method.
-
-\subsubsection{Neutron flux distribution}
-
-This discussion on the neutron flux distributions focuses on Cases 3a and 3b which most closely
-resemble the actual \gls{MSRE} reactor geometry. Cases 3a and 3b resemble the \gls{MSRE} reactor
-with a control rod withdrawn and inserted, respectively. Figures \ref{fig:3a-flux} and
-\ref{fig:3b-flux} show the neutron group flux distributions for Cases 3a and 3b from OpenMC-CE and
-OpenMC-MG. Differences between the two methods arise from the multigroup approximation, notably in
-group 2 for Case 3a and groups 2, 6, 7, 8 for Case 3b. Future work aimed at finding a better
-neutron energy group structure should at minimum focus on improving fidelity in the listed energy
-groups.
-
-For a fair evaluation of the deterministic multigroup methods without distortions from the
-multigroup approximation, I chose to compare their flux distributions with the OpenMC-MG flux
-distribution. Figures \ref{fig:3a-flux-diff} and \ref{fig:3b-flux-diff} show the absolute
-difference in neutron group flux distributions of the Moltres-$S_8$, Moltres-diffusion,
-Moltres-hybrid, and Python-hybrid methods relative to OpenMC-MG. The figures omit Python-$S_8$ and
-Python-diffusion as their distributions are nearly identical to their Moltres-based counterparts.
-The neutron diffusion and hybrid methods fare worse than the $S_8$ method at capturing the
-oscillatory pattern in groups 1, 2, 5, 7, and 8 arising from the fuel-graphite lattice geometry
-along $x=5$ cm to 10 cm. In both Cases 3a and 3b, the neutron diffusion method (yellow plot lines)
-exhibits significant flux deviations in groups 1, 2, 7, and 8 near $x=0$ cm.
-
-Both hybrid methods
-provide significant improvements in neutron flux distributions compared to the neutron diffusion
-method as shown by the generally smaller flux difference values particularly near $x=0$ cm. The
-figures also show Moltres-hybrid (green) provides more accurate flux estimates than
-Python-hybrid (red) in all instances. The differences indicate deficiencies in the diffusion
-correction scheme in Python-hybrid which lead to poorer performance compared to the drift
-correction scheme in Moltres-hybrid.
-
-\begin{figure}[htb!]
-  \centering
-  \includegraphics[width=\columnwidth]{case-3a-flux}
-  \caption{Case 3a neutron group flux distributions from OpenMC-CE and OpenMC-MG.}
-  \label{fig:3a-flux}
-\end{figure}
-
-\begin{figure}[htb!]
-  \centering
-  \includegraphics[width=\columnwidth]{case-3b-flux}
-  \caption{Case 3b neutron group flux distributions from OpenMC-CE and OpenMC-MG.}
-  \label{fig:3b-flux}
-\end{figure}
-
-\begin{figure}[htb!]
-  \centering
-  \includegraphics[width=\columnwidth]{case-3a-flux-diff}
-  \caption{Absolute difference in neutron group flux distributions for Case 3a from Moltres-$S_8$,
-  Moltres-diffusion, Moltres-hybrid, and Python-hybrid relative to OpenMC-MG.}
-  \label{fig:3a-flux-diff}
-\end{figure}
-
-\begin{figure}[htb!]
-  \centering
-  \includegraphics[width=\columnwidth]{case-3b-flux-diff}
-  \caption{Absolute difference in neutron group flux distributions for Case 3b from Moltres-$S_8$,
-  Moltres-diffusion, Moltres-hybrid, and Python-hybrid relative to OpenMC-MG.}
-  \label{fig:3b-flux-diff}
-\end{figure}
-
-\FloatBarrier
-
-\subsubsection{Drift \& diffusion correction distribution}
-
-Figures \ref{fig:3b-drift} and \ref{fig:3b-svdc} show the distributions of the drift and diffusion
-correction parameters, respectively, in each neutron energy group for Case 3b. As a reminder to the
-reader, the hybrid $S_N$-diffusion methods generate these correction parameters in subregions
-containing highly neutron absorbing materials. I implemented the drift correction scheme for
-Moltres-hybrid and the diffusion correction scheme for Python-hybrid. The hybrid methods
-selectively apply the $S_N$ method to improve neutronic modeling within the subregions. I computed
-the reference distributions (yellow data points) in both plots from reference flux solutions of the
-standard $S_8$ method.
-
-As discussed in Section \ref{sec:buffer-region}, both sets of correction parameters generated by
-the two correction schemes match their respective reference values within the correction subregion
-from $x=0$ cm to 10 cm except near the subregion boundary at $x=10$ cm. Therefore, both schemes
-provide improved flux estimates in the subregion compared to the neutron diffusion method.
-The prior reactivity and flux distribution results indicate Moltres-hybrid is more accurate
-than Python-hybrid relative to OpenMC-MG as the reference.
-
-In Figure \ref{fig:3b-drift}, the drift
-correction distributions (Moltres-hybrid) for all eight energy groups are well-defined throughout
-the subregion and are continuous except at material interfaces. Consequently, the drift correction
-scheme provides accurate flux corrections for about 75\% of the subregion based on the buffer
-region cutoff criteria which occurs around $x=7.5$ cm where most of the drift distributions are
-zero (refer to Section \ref{sec:buffer-region}). By contrast, the diffusion correction
-distributions in Figure \ref{fig:3b-svdc} contains numerous discontinuities tending towards
-positive or negative infinity in all eight energy groups. The discontinuities arise when
-computing corrected diffusion coefficients (Eq. \ref{eq:svdc}) near flux peaks and troughs
-(stationary points) where the flux gradients approach zero. The location of stationary points from
-the diffusion and $S_N$ calculations in the hybrid method do not necessarily coincide, resulting in
-discontinuous flux gradients in the final flux solution. Discontinuities are generally unavoidable
-in most problems because the fluxes typically reach local minima in control rod regions.
-Furthermore, unlike negative drift coefficients, negative diffusion coefficients are unphysical.
-Taking everything into consideration, the diffusion correction scheme is less accurate and more
-numerically unstable than the drift correction scheme.
-
-The subsequent discussions on the correction region size and the number of outer
-iterations focus on the drift correction scheme of the hybrid method.
-
-\begin{figure}[htb!]
-  \centering
-  \includegraphics[width=\columnwidth]{case-3b-drift}
-  \caption{Multigroup drift correction ($\vec{D}_g$) $x$-component distributions from the
-  Moltres-hybrid and Moltres-$S_8$ solvers.}
-  \label{fig:3b-drift}
-\end{figure}
-
-\begin{figure}[htb!]
-  \centering
-  \includegraphics[width=\columnwidth]{case-3b-svdc}
-  \caption{Multigroup diffusion correction ($D^s_g$) $x$-component distributions from the
-    Python-hybrid and Python-$S_8$ solvers.}
-  \label{fig:3b-svdc}
-\end{figure}
-
-\FloatBarrier
-
-\subsubsection{Impact of correction subregion sizes}
-
-The hybrid $S_N$-diffusion method relies on $S_N$ calculations in the correction subregion to
-generate flux corrections. Minimizing the size of the correction subregion and the $S_N$ subproblem
-is essential for the hybrid method to be computationally competitive for time-dependent full-core
-simulations. I investigated the effect of the correction subregion size on the $k_\text{eff}$ and
-control rod worth estimates with Cases 3a and 3b by varying the subregion sizes from 10 cm to 40 cm
-at 5 cm-intervals.
-
-Figures \ref{fig:v1-size-a-k} and \ref{fig:v1-size-b-k} show the $k_\text{eff}$ estimates from the
-hybrid method for Cases 3a and 3b, respectively. In both cases, the $k_\text{eff}$ values initially
-decrease as the correction subregion sizes increase before reversing in trend when the subregion
-size reaches 35 cm and beyond. The $k_\text{eff}$ values vary by up to 164 pcm for Case 3a and 109
-pcm for Case 3b. An important observation here is that the $k_\text{eff}$ values do not
-monotonically converge towards the $k_\text{eff}$ estimate from the $S_8$ method. The data implies
-other significant sources of discrepancies exist beyond those being corrected in the
-correction subregion by the hybrid method.
-
-% In general,
-% discrepancies arise between the neutron diffusion method and $S_8$ from the neutron diffusion 
-% method's failure to accurately model the neutron flux in the fuel-graphite lattice region. The
-% figures imply that those discrepancies result in a positive change in $k_\text{eff}$ since the
-% $k_\text{eff}$ decreases as larger fractions of the problem domain benefit from flux corrections.
-% The remaining discrepancy in $k_\text{eff}$ between the hybrid method with a correction subregion
-% size of 40 cm and the $S_8$ method arises from
-
-\begin{figure}[htb!]
-  \centering
-  \begin{subfigure}[b]{0.49\columnwidth}
-    \centering
-    \includegraphics[width=\columnwidth]{correction-size-a-k}
-    \caption{Case 3a}
-    \label{fig:v1-size-a-k}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}[b]{0.49\columnwidth}
-    \centering
-    \includegraphics[width=\columnwidth]{correction-size-b-k}
-    \caption{Case 3b}
-    \label{fig:v1-size-b-k}
-  \end{subfigure}
-  \caption{$k_\text{eff}$ estimates from the hybrid method for Cases 3a and 3b with different
-  correction subregion sizes. The horizontal lines indicate $k_\text{eff}$ estimates from the
-  OpenMC-CE, OpenMC-MG, and $S_8$ methods.}
-  \label{fig:v1-size-k}
-\end{figure}
-
-Figure \ref{fig:v1-size-rho} shows the percentage difference in rod worth relative to OpenMC-CE. 
-Due to the identical trends observed in the $k_\text{eff}$ estimates of both cases, the control rod
-worth estimates do not change significantly when the correction subregion
-sizes change. This indicates limiting the correction subregion size to save
-computational cost on the expensive $S_N$ calculations has a negligible impact on rod worth
-estimates. The rod worth estimates for all investigated correction subregion sizes remain within
-0.2\% of the $S_8$ method.
-
-\begin{figure}[htb!]
-  \centering
-  \includegraphics[width=0.8\columnwidth]{correction-size-rho}
-  \caption{Percentage difference in rod worth from the hybrid method relative to OpenMC-CE for
-    Cases 3a and 3b with different correction subregion sizes. The horizontal lines indicate
-    equivalent rod worth differences from the OpenMC-MG and $S_8$ methods.}
-  \label{fig:v1-size-rho}
-\end{figure}
-
-The $k_\text{eff}$ and rod worth results show the hybrid method is able to accurately
-reproduce rod worth estimates as long as the correction subregion size remains consistent. The
-correction subregion size must also be sufficiently large to capture the highly neutron-absorptive
-rod's influence on the neutron flux around the rod at $x=0$ cm. The drift correction parameter
-distributions in Figure \ref{fig:3b-drift} indicate that the influence of the rod on the drift, and
-consequently the flux, extend to approximately $x=5$ cm before the drift distribution settles on a
-regular repeating pattern influenced by the fuel-graphite lattice structure. I applied a consistent
-correction subregion size of 10 cm for Cases 2 and 3 since it sufficiently meets the requirements
-discussed while also minimizing the computational cost of the $S_8$ subsolver. Other reactor
-designs consisting of different material components may require different minimum correction
-subregion sizes.
-
-\subsubsection{Relaxing the $S_N$ subsolver convergence threshold}
-
-As described in Section \ref{sec:hybrid-algorithm}, the hybrid $S_N$-diffusion method works by
-iteratively coupling a neutron diffusion solver and a $S_N$ subsolver with the latter generating
-drift correction parameters for the former. Excluding the initial neutron diffusion calculation,
-each outer (fixed point) iteration comprises of an inner $S_N$ calculation on the correction
-subregion and an inner neutron diffusion calculation on the entire domain. The \gls{FEM} algorithm
-in Moltres determines an individual solver reaches convergence when the combined
-residual contributions fall below a specified convergence tolerance threshold. The fixed point
-iteration scheme under the \texttt{MultiApps} system in Moltres for coupling multiple individual
-solvers (referred to as apps) requires users to specify a main app for the final convergence check
-after each outer iteration to terminate the iterations. The scheme also requires users to set a
-fixed point convergence tolerance value independent of the tolerance values of each individual
-solver. Naturally, the fixed point tolerance value must be equal to or greater than the tolerance
-value of the main individual solver for the possibility of overall convergence. The hybrid method
-implementation in Moltres sets the neutron diffusion solver as the main app and the $S_N$ subsolver
-as the sub app.
-
-The hybrid method shares some similarities with diffusion-based \gls{HOLO} acceleration schemes
-(refer to Section \ref{sec:drift-correction}) for the neutron transport methods which do not
-require the neutron transport side of the calculations to fully converge
-\cite{reynolds_analysis_2023, wang_diffusion_2014}.
-The typical convergence threshold value required for converged neutron flux calculations in Moltres
-is $10^{-8}$; I set the convergence tolerance value to $10^{-8}$ for the neutron diffusion solver
-and the fixed point scheme. I varied the convergence tolerance value for the $S_N$ subsolver to 
-find the optimal value for maintaining sufficient solution accuracy and minimizing computational
-costs; tightening the $S_N$ convergence tolerance value generally results in more outer iterations
-performed before the solver meets the fixed point convergence tolerance.
-Table \ref{table:sn-tol} lists the number of outer iterations required by the hybrid method for
-a given set of $S_8$ subsolver convergence tolerance values in Case 3b.
-
-\begin{table}[htb]
-  \centering
-  \caption{Number of outer iterations in hybrid method calculations of Case 3b for a given set of
-  convergence tolerance values imposed on the $S_8$ subsolver.}
-  \begin{tabular}{l S S S S S S}
-    \toprule
-    $S_8$ subsolver tolerance, $\epsilon_\text{tol}$ & {$10^{-8}$} & {$10^{-7}$} & {$10^{-6}$} & {$10^{-5}$} & {$10^{-4}$} & {$10^{-3}$} \\
-    \midrule
-    Number of outer iterations & 3 & 3 & 3 & 2 & 2 & 1 \\
-    \bottomrule
-  \end{tabular}
-  \label{table:sn-tol}
-\end{table}
-
-Figure \ref{fig:sn-tol} shows the $k_\text{eff}$ error estimates relative to the reference value
-when $\epsilon_\text{tol}=10^{-8}$. The hybrid method exhibits superlinear convergence ($q=1.333$)
-with respect to the $S_N$ subsolver convergence tolerance value. Based on the results, setting
-$\epsilon_\text{tol}$ to $10^{-5}$ is the best balance between accuracy and computational
-cost because the $k_\text{eff}$ error is less than 0.01 pcm and the number of outer iterations
-would increase from 2 to 3 if the tolerance value is further tightened.
-
-\begin{figure}[htb!]
-  \centering
-  \includegraphics[width=0.6\columnwidth]{sn-tol}
-  \caption{$k_\text{eff}$ error estimates of Case 3b for a range of convergence tolerance values
-  imposed on the $S_8$ subsolver relative to the reference $k_\text{eff}$ value when
-  $\epsilon_\text{tol}=10^{-8}$. The $q=1.333$ line represents the approximate rate of
-  convergence.}
-  \label{fig:sn-tol}
-\end{figure}
-
-\section{2-D Neutronics Models \& Results} \label{sec:2d-results}
-
-This section presents the models, results, and discussion of the 2-D cases.
-
-\subsection{2-D Neutronics Model Setup} \label{sec:2d-model-setup}
-
-Figures \ref{fig:1/4-geom} and \ref{fig:full-geom} show the 2-D quarter-core and full-core
-\gls{MSRE} models investigated. I adapted the models from the horizontal cross section of the
-\gls{MSRE} numerical benchmark model \cite{fratoni_molten_2020} with the
-following differences in the 2-D \gls{MSRE} model for this work:
-
-\begin{itemize}
-  \item Neglected thermal expansion of all reactor components.
-  \item Modeled the control rods as perfectly annular cylinders instead of beaded control elements
-    strung on flexible steel cables. Neglected the inconel shells encasing the control elements and
-    the steel helical cables linking the control elements together.
-  \item Set channel pitch, width, and length to 2, 0.2, and 1.2 inches, respectively.
-  \item Homogenized the INOR-8, graphite, and molten salt regions in the sample basket.
-  \item Modified the outer edge of the core graphite to be circular with no jagged edges.
-  \item Replaced partial channels with full-size channels near the edge of the core graphite.
-  \item Ignored the thermal insulation and shield structures outside the core.
-\end{itemize}
-
-The differences listed above, along with other changes in the 3-D model, result in negligible
-changes to the $k_\text{eff}$ (refer to Section 5.6) relative to the \gls{MSRE} benchmark model
-\cite{fratoni_molten_2020}.
-
-\begin{figure}[htb!]
-  \centering
-  \includegraphics[width=\columnwidth]{quarter-core-geom}
-  \caption{2-D \gls{MSRE} quarter-core model based on the horizontal cross section of the actual
-  \gls{MSRE} geometry.}
-  \label{fig:1/4-geom}
-\end{figure}
-
-\begin{figure}[htb!]
-  \centering
-  \includegraphics[width=\columnwidth]{full-core-geom}
-  \caption{2-D \gls{MSRE} full-core model based on the horizontal cross section of the actual
-  \gls{MSRE} geometry.}
-  \label{fig:full-geom}
-\end{figure}
-
-\begin{table}[htb]
-  \small
-  \centering
-  \setlength\tabcolsep{4pt}
-  \caption{\gls{MSRE} molten salt composition when the $^{235}$U loading was at 65.25 kg.}
-  \begin{tabular}{l S S S S S S S}
-    \toprule
-    Element & {Li} & {Be} & {Zr} & {Hf} & {$^{234}$U} & {$^{235}$U} & {$^{236}$U} \\
-    \midrule
-    Composition [kg] & 507.27 & 293.96 & 513.97 & 0.0029 & 0.67 & 65.25 & 0.27 \\
-    \bottomrule
-  \end{tabular}
-  \begin{tabular}{l S S S S S S}
-    \toprule
-    Element & {$^{238}$U} & {Fe} & {Cr} & {Ni} & {O} & {F} \\
-    \midrule
-    Composition [kg] & 141.91 & 0.75 & 0.13 & 0.14 & 2.27 & 3103.22 \\
-    \bottomrule
-  \end{tabular}
-  \label{table:2d-salt-composition}
-\end{table}
-
-INOR-8, also referred to as Hastelloy N, is a nickel-based alloy used for structural components in
-the \gls{MSRE} \cite{robertson_msre_1965}.
-Table \ref{table:2d-salt-composition} lists the molten salt composition for the 2-D models. The
-$^{235}$U loading of 65.25 kg is the amount at which the \gls{MSRE} first achieved criticality. The
-control rods consist of the original \gls{MSRE} 70 wt\% Gd$_2$O$_3$-30 wt\% Al$_2$O$_3$ mixture.
-The models impose neutron vacuum boundary conditions on the outermost boundary of the INOR-8
-vessel structure. The quarter-core model imposes neutron reflecting boundary conditions along the
-left and bottom straight edges (Figure \ref{fig:1/4-geom}). Note that the quarter-core and
-full-core models are not identical because the full-core model contains a sample basket instead of
-a control rod in the lower-right segment of Figure \ref{fig:full-geom-closeup}.
-
-\begin{figure}[htb!]
-  \centering
-  \includegraphics[width=0.7\columnwidth]{full-core-closeup}
-  \caption{Detailed view of the control rod thimbles and sample basket in 2-D \gls{MSRE} full-core
-    model.}
-  \label{fig:full-geom-closeup}
-\end{figure}
-
-I ran neutronics simulations on the quarter- and full-core models on OpenMC under both continuous
-energy (OpenMC-CE) and multigroup (OpenMC-MG) modes, and on Moltres with the hybrid $S_N$-diffusion
-and neutron diffusion methods. The hybrid method ran with the $S_8$ angular discretization scheme
-and up to 2nd-order scattering approximations for the $S_N$ subsolver. I simulated the
-quarter-core model with and without the control rod in the air-filled rod thimble; air replaces the
-control rod material when the rod is removed. For the full-core model, I considered four rod
-configurations: no rods present, rod 1 present, rods 1 \& 2 present, and rods 1, 2 \& present.
-Other than providing a close-up view of the core center, Figure \ref{fig:full-geom-closeup} also
-presents the correction subregion on which the $S_8$ subsolver of the hybrid method generates
-drift correction parameters for the coupled diffusion solver.
-
-\FloatBarrier
-
-\subsection{2-D Neutronics Numerical Results \& Discussion} \label{sec:2d-nts-results}
-
-This subsection presents the results and discussion of the 2-D quarter-core and full-core models.
-
-\subsubsection{2-D Quarter-Core}
-
-Table \ref{table:quarter-core} details the $k_\text{eff}$ and control rod worth estimates and error
-values relative to OpenMC-CE. For the ``No Rod'' case, OpenMC-CE estimated the $k_\text{eff}$ to be
-1.11209(43). The OpenMC-MG, hybrid, and neutron diffusion methods performed
-similarly to one another with reactivity errors ranging from 618 pcm to 773 pcm relative to
-OpenMC-CE. Based on $k_\text{eff}$ alone, the hybrid method performed marginally worse than the
-neutron diffusion method. However, for the ``Rod'' case, then neutron diffusion method saw a
-change in sign of the error to -816 pcm. The errors for OpenMC-MG and the hybrid method
-remained positive and reduced to 446 pcm and 760 pcm, respectively. Consequently, when taking the
-difference in reactivities to estimate the rod worth, the neutron diffusion method performed
-significantly worse than OpenMC-MG and the hybrid method. OpenMC-CE estimated the rod worth to be
-8370(53) pcm. The neutron diffusion method overestimated the rod worth by 1484 pcm. The hybrid
-method improved significantly on the neutron diffusion method with a rod worth estimate that is
-only 13 pcm
-higher than OpenMC-CE. OpenMC-MG reported a larger error of 172 pcm likely due to favorable error
-cancellation with the hybrid method and statistical uncertainties with the Monte Carlo methods.
-The OpenMC-CE and OpenMC-MG rod worth estimates agree within two standard deviations of each other.
-
-\begin{table}[htb]
-  \small
-  \centering
-  \caption{$k_\text{eff}$ and control rod worth estimates for the 2-D quarter-core \gls{MSRE}
-    model. Error values are relative to OpenMC-CE.}
-  \begin{tabular}{l S[table-format=1.5(2)] S S[table-format=1.5(2)] S S[table-format=4(2)] S}
-    \toprule
-    \multirow{2}{*}{Method} & \multicolumn{2}{c}{No Rod} & \multicolumn{2}{c}{Rod} & \multicolumn{2}{c}{Rod worth} \\
-                            & {$k_\text{eff}$} & {Error [pcm]} & {$k_\text{eff}$} & {Error [pcm]} & {$\Delta\rho_\text{worth}$ [pcm]} & {Error [pcm]} \\
-                            \cmidrule(r){1-1} \cmidrule(rl){2-3} \cmidrule(rl){4-5} \cmidrule(l){6-7}
-	  OpenMC-CE & 1.11209(43) & {-} & 1.01740(42) & {-} & 8370(53) & {-} \\
-	  OpenMC-MG & 1.11979(42) & 618 & 1.02204(41) & 446 & 8541(51) & 172 \\
-      Diffusion & 1.12059 & 682 & 1.00903 & -816 & 9867 & 1484 \\
-      Hybrid & 1.12174 & 773 & 1.02532 & 760 & 8383 & 13 \\
-    \bottomrule
-  \end{tabular}
-  \label{table:quarter-core}
-\end{table}
-
+\begin{frame}
+  \frametitle{Hybrid $S_N$-Diffusion Method: 3-D Neutronics Eigenvalue Simulations}
+  \begin{columns}
+    \column{5.5cm}
+  \textbf{3-D Full-Core MSRE Model}
   \begin{figure}[p]
     \centering
-  \includegraphics[width=\columnwidth]{msre-quarter-no-rod-power}
-  \caption{Normalized channel fission rate distribution of the 2-D \gls{MSRE} quarter-core model.
-  This figure depicts the 1/8th-core distribution due to symmetry about the diagonal.}
-  \label{fig:1/4-no-rod}
+    \includegraphics[width=0.85\columnwidth]{msre-picture}
+    \caption{Vertical cross section of the actual \gls{MSRE} vessel.}
+    \label{fig:msre-picture}
   \end{figure}
-
+    \column{6cm}
   \begin{figure}[p]
     \centering
-  \includegraphics[width=\columnwidth]{msre-quarter-rod-power}
-  \caption{Normalized channel fission rate distribution of the 2-D \gls{MSRE} quarter-core model.
-  This figure depicts the 1/8th-core distribution due to symmetry about the diagonal.}
-  \label{fig:1/4-rod}
-\end{figure}
-
-\FloatBarrier
-
-I also calculated the normalized channel fission rate distribution by integrating the fission rate
-within each fuel salt channel and normalizing the values such that the integrated fission rate in
-each channel averages to 1 as follows
-%
-\begin{align}
-  \text{Normalized fission rate of channel }i &= \frac{\text{Fission rate in channel }i}{
-  \text{Total fission rate}}\cdot\frac{\text{Total number of channels}}{
-    \text{Total fission rate in channels}} \nonumber \\
-                                              &= \frac{\sum^G_{g=1}\int_{V_i}\Sigma_{f,g}
-  \phi_g(\vec{r})}{\sum^G_{g=1}\int_V\Sigma_{f,g}\phi_g(\vec{r})}\cdot\frac{N_i}{
-  \sum^{N_i}_{i=1}\sum^G_{g=1}\int_{V_i}\Sigma_{f,g}\phi_g(\vec{r})}.
-\end{align}
-%
-The total fission rate includes fission rate contributions from peripheral fuel salt regions.
-Figures \ref{fig:1/4-no-rod} and \ref{fig:1/4-rod} show the normalized fission rate distributions
-of the quarter-core model for the ``No Rod'' and ``Rod'' cases, respectively. I presented 1/8th
-slices of the core in the figures because the distributions are symmetrical about the $y=x$ line
-outside of statistical variances for OpenMC-CE. I combined the midpoints between the centroids of
-neighboring fuel salt channels to form the diamond lattices shown in the figures.
-Each diamond element in the figures corresponds a horizontally or vertically oriented channel
-depicted in Figure \ref{fig:1/4-geom}. The lattice also provides a guide for subdividing the
-annular fuel regions around the control rod thimbles. The fission rate
-distributions omit four channels along the outer edge due to the proximity to peripheral annular
-fuel regions and the consequent difficulty in accurately tallying fission rates.
-
-In the ``No Rod'' case, the peak channel fission rate occured several channels away from the
-physical center of the reactor core. The presence of the empty (air-filled) rod thimble depressed
-fission rates near the center. The neutron diffusion and hybrid methods report similar percentage
-errors relative to OpenMC-CE except near the center. While the hybrid method benefits from drift
-corrections, the neutron diffusion method's reliance on scalar diffusion coefficients tends to
-artificially flatten neutron fluxes in near-void regions. Therefore, the largest percentage errors
-from the neutron diffusion method occurred in the second and third diamond elements
-diagonally up from the bottom which represent the annular fuel region surrounding the rod thimble.
-In the ``Rod'' case, the introduction of the control rod shifted the peak further away from the
-center as expected of a highly neutron-absorbing material. The rod further worsened fission rate
-errors for the neutron diffusion method with large errors of 1\% or more at the core center and
-periphery. On the other hand, the fission rate errors for the hybrid method largely remained below
-1\%.
-
-\begin{table}[htb]
-  \small
-  \centering
-  \caption{Absolute mean and maximum percentage errors in the normalized channel fission rates of
-  the 2-D \gls{MSRE} quarter-core models relative to OpenMC. The mean relative standard deviation of
-  OpenMC normalized channel fission rates is 0.20\%.}
-  \begin{tabular}{l S S S S}
-    \toprule
-    \multirow{2}{*}{Method} & \multicolumn{2}{c}{No Rod} & \multicolumn{2}{c}{Rod} \\
-                            & {Mean [\%]} & {Maximum [\%]} & {Mean [\%]} & {Maximum [\%]} \\
-                            \cmidrule(r){1-1} \cmidrule(rl){2-3} \cmidrule(l){4-5}
-    Diffusion & 0.40 & 2.63 & 2.01 & 17.44 \\
-    Hybrid & 0.40 & 1.32 & 0.43 & 3.08 \\
-    \bottomrule
-  \end{tabular}
-  \label{table:quarter-core-power}
-\end{table}
-
-Table \ref{table:quarter-core-power} shows the absolute mean and maximum percentage errors of the
-normalized channel fission rates from the neutron diffusion and hybrid methods relative to
-OpenMC-CE. Overall, the hybrid method reported smaller percentage errors than the neutron diffusion
-method. The absolute maximum percentage error from the neutron diffusion method jumped to 17.44\%
-as compared to the 3.08\% for the hybrid method in the ``Rod'' case. Along with the figures, these
-results demonstrate the hybrid method provides improved fission rate distribution accuracy in
-the existence of near-void and highly neutron-absorbing regions.
-
-\FloatBarrier
-
-\subsubsection{2-D Full-Core}
-
-I performed the same $k_\text{eff}$ and fission rate distribution analyses with the 2-D full-core
-model under four configurations starting from no rod inserted to all three rods inserted. The
-$k_\text{eff}$ and error estimates in Table \ref{table:full-core-k} show the same trends observed
-with the 2-D quarter-core model. OpenMC-MG and the hybrid method exhibit moderate but consistent
-reactivity errors of approximately the 550 pcm and 800 pcm, respectively. The neutron diffusion
-method reported errors ranging from 692 pcm when no rod is inserted to -470 pcm when all three rods
-are inserted. Consequently, in Table \ref{table:full-core-worth}, OpenMC-MG and the hybrid method
-report accurate control rod worths within 1-3 standard deviations of OpenMC-CE while the neutron
-diffusion method increasingly overestimates the rod worth as more rods are inserted.
-
-\begin{table}[htb]
-  \small
-  \centering
-  \caption{$k_\text{eff}$ estimates for the 2-D full-core \gls{MSRE} model with the indicated rods
-    inserted. Error values are relative to OpenMC-CE.}
-  \setlength\tabcolsep{2pt}
-  \begin{tabular}{l S[table-format=1.5(2)] S S[table-format=1.5(2)] S S[table-format=1.5(2)] S S[table-format=1.5(2)] S}
-    \toprule
-    \multirow{3}{*}{Method} & \multicolumn{2}{c}{No Rod} & \multicolumn{2}{c}{Rod 1} & \multicolumn{2}{c}{Rod 1 \& 2} & \multicolumn{2}{c}{Rod 1, 2 \& 3} \\
-                            & {\multirow{2}{*}{$k_\text{eff}$}} & {Error} & {\multirow{2}{*}{$k_\text{eff}$}} & {Error} & {\multirow{2}{*}{$k_\text{eff}$}} & {Error} & {\multirow{2}{*}{$k_\text{eff}$}} & {Error} \\
-                            & & {[pcm]} & & {[pcm]} & & {[pcm]} & & {[pcm]} \\
-                            \cmidrule(r){1-1} \cmidrule(rl){2-3} \cmidrule(rl){4-5} \cmidrule(rl){6-7} \cmidrule(l){8-9}
-    OpenMC-CE & 1.09859(20) & {-} & 1.06980(21) & {-} & 1.04690(18) & {-} & 1.02688(18) & {-} \\
-    OpenMC-MG & 1.10637(19) & 640 & 1.07633(20) & 567 & 1.05235(20) & 495 & 1.03262(17) & 541 \\
-    Diffusion & 1.10701 & 692 & 1.07120 & 122 & 1.04415 & -252 & 1.02195 & -470 \\
-    Hybrid & 1.10843 & 808 & 1.07906 & 802 & 1.05553 & 781 & 1.03582 & 841 \\
-    \bottomrule
-  \end{tabular}
-  \label{table:full-core-k}
-\end{table}
-
-\begin{table}[htb]
-  \small
-  \centering
-  \caption{Control rod worth estimates for the 2-D full-core \gls{MSRE} with the
-  indicated rods inserted. Error values are relative to OpenMC-CE.}
-  \setlength\tabcolsep{5pt}
-  \begin{tabular}{l S[table-format=4(2)] S S[table-format=4(2)] S S[table-format=4(2)] S}
-    \toprule
-    \multirow{2}{*}{Method} & \multicolumn{2}{c}{Rod 1} & \multicolumn{2}{c}{Rod 1 \& 2} & \multicolumn{2}{c}{Rod 1, 2 \& 3} \\
-                            & {$\Delta\rho_\text{worth}$ [pcm]} & {Error [pcm]} & {$\Delta\rho_\text{worth}$ [pcm]} & {Error [pcm]} & {$\Delta\rho_\text{worth}$ [pcm]} & {Error [pcm]} \\
-                            \cmidrule(r){1-1} \cmidrule(rl){2-3} \cmidrule(rl){4-5} \cmidrule(l){6-7}
-    OpenMC-CE & 2450(25) & {-} & 4494(23) & {-} & 6357(24) & {-} \\
-    OpenMC-MG & 2523(23) & 73 & 4640(24) & 146 & 6455(22) & 98 \\
-    Diffusion & 3019 & 569 & 5439 & 945 & 7519 & 1162 \\
-    Hybrid & 2455 & 5 & 4521 & 27 & 6323 & -34 \\
-    \bottomrule
-  \end{tabular}
-  \label{table:full-core-worth}
-\end{table}
-
-Plots of the normalized channel fission rate distributions for all four rod configurations are
-available in Appendix \ref{chap:2d-fiss-rate}. They are best viewed in digital pdf format with
-magnification due to the numerous fuel salt channels and small fonts.
-
-Table \ref{table:full-core-power} shows the absolute mean and maximum percentage errors in the
-normalized channel fission rates from the neutron diffusion and hybrid methods relative to
-OpenMC-CE. Both mean and maximum percentage errors from the neutron diffusion method steadily
-increase with the number of rods inserted. The mean percentage error from the hybrid method remains
-constant at 0.43\% while the maximum percentage error increases marginally from 1.45\% to 2.52\%
-when all three rods are inserted. These results demonstrate the hybrid method remains
-effective at improving control rod worth and fission rate distribution estimates in less symmetric
-geometries.
-
-\begin{table}[htb]
-  \footnotesize
-  \centering
-  \caption{Absolute mean and maximum percentage errors in the normalized channel fission rates of
-  the 2-D \gls{MSRE} full-core models relative to OpenMC. The mean relative standard deviation of
-  OpenMC normalized channel fission rates is 0.27\%.}
-  \setlength\tabcolsep{2.5pt}
-  \begin{tabular}{l S S S S S S S S}
-    \toprule
-    \multirow{2}{*}{Method} & \multicolumn{2}{c}{No Rod} & \multicolumn{2}{c}{Rod 1} & \multicolumn{2}{c}{Rod 1 \& 2} & \multicolumn{2}{c}{Rod 1, 2 \& 3} \\
-                            & {Mean [\%]} & {Maximum [\%]} & {Mean [\%]} & {Maximum [\%]} & {Mean [\%]} & {Maximum [\%]} & {Mean [\%]} & {Maximum [\%]} \\
-                            \cmidrule(r){1-1} \cmidrule(rl){2-3} \cmidrule(rl){4-5} \cmidrule(rl){6-7} \cmidrule(l){8-9}
-    Diffusion & 0.45 & 2.95 & 0.94 & 12.61 & 1.35 & 15.34 & 1.67 & 17.09 \\
-    Hybrid & 0.43 & 1.45 & 0.43 & 1.82 & 0.43 & 2.26 & 0.43 & 2.52 \\
-    \bottomrule
-  \end{tabular}
-  \label{table:full-core-power}
-\end{table}
-
-\FloatBarrier
-
-\subsubsection{$S_N$ Discretization \& Ray Effects}
-
-Angular resolution in 2-D (and later 3-D) flux simulations is important for accurately capturing
-the angular dependence in the neutron angular flux. This subsection compares the $k_\text{eff}$ and
-neutron scalar flux distribution solutions from the hybrid method with the $S_6$, $S_8$, $S_{10}$,
-and $S_{12}$ methods for the 2-D full-core with no rods inserted. Table
-\ref{table:2d-sn-convergence} shows the $k_\text{eff}$ estimates from those simulation results.
-The $k_\text{eff}$ varies by less than 1 pcm, thereby indicating that the angular discretization
-order used by the hybrid method has little impact on the integral reactor parameters.
-
-Figures \ref{fig:ray-effect-12} and \ref{fig:ray-effect-34} show the first four neutron group flux
-distributions from the hybrid method in the central reactor region consisting of three
-air-filled thimbles and a homogenized sample basket (bottom right). The geometry corresponds
-directly to Figure \ref{fig:full-geom-closeup} without the three control rods. Group 1 fluxes
-in the thimble form a distinctive ray effect artifacts arising from the regular geometrical shapes and
-insufficient angular resolution. The artifacts become less distinct with increasing angular
-discretization levels. They also diminish with increasing neutron energy groups due to shorter
-neutron mean free paths leading to more diffusive behavior.
-
-By Group 3 and 4, the artifacts are indiscernible to the eye across all angular discretization
-orders. Given that $k_\text{eff}$ is heavily weighted towards the later (slower) energy groups due
-to larger fission cross sections, it is clear why Table \ref{table:2d-sn-convergence} shows little
-differences in $k_\text{eff}$ across the angular discretization orders. Therefore, $S_8$ is
-adequate for control rod worth calculations in the \gls{MSRE}. 
-Further investigations into the angular discretization level may be warranted when applying the
-hybrid method to reactors with more air-filled regions such as high-temperature gas-cooled
-reactors.
-
-\begin{table}[t]
-  \centering
-  \caption{Hybrid method $k_\text{eff}$ estimates for the 2-D full-core model with zero rods
-  inserted and $S_N$ schemes from $S_6$ to $S_{12}$.}
-  \begin{tabular}{l S}
-    \toprule
-    Angular discretization & {$k_\text{eff}$} \\
-    \midrule
-    $S_6$ & 1.108414 \\
-    $S_8$ & 1.108423 \\
-    $S_{10}$ & 1.108422 \\
-    $S_{12}$ & 1.108420 \\
-    \bottomrule
-  \end{tabular}
-  \label{table:2d-sn-convergence}
-\end{table}
-
-\begin{figure}[p]
-  \small
-  \centering
-  \includegraphics[width=0.7\columnwidth]{ray-effect-12}
-  \caption{Group 1 and 2 neutron flux distributions in the hybrid $S_N$-diffusion method correction
-  region with $S_6$, $S_8$, $S_{10}$, \& $S_{12}$. Ray effects diminish with increasing
-  angular fidelity, but still persist with $S_{12}$ in group 1 and 2.}
-  \label{fig:ray-effect-12}
-\end{figure}
-
-\begin{figure}[p]
-  \small
-  \centering
-  \includegraphics[width=0.7\columnwidth]{ray-effect-34}
-  \caption{Group 3 and 4 neutron flux distributions in the hybrid $S_N$-diffusion method correction
-  region with $S_6$, $S_8$, $S_{10}$, \& $S_{12}$. Ray effects are barely visible in group
-  3 and 4 due to shorter neutron mean free paths.}
-  \label{fig:ray-effect-34}
-\end{figure}
-
-\FloatBarrier
-
-\section{3-D Neutronics Models \& Results} \label{sec:3d-results}
-
-This section presents the models, results, and discussion of the 3-D \gls{MSRE} simulations.
-
-\subsection{3-D Neutronics Model Setup} \label{sec:3d-model-setup}
-
-Figure \ref{fig:msre-picture} shows the vertical cross section of the actual \gls{MSRE} vessel.
-The actual design has rounded upper and lower plena to help direct salt flow through the system.
-This work uses a modified 3-D numerical \gls{MSRE} model as shown by the vertical cross section
-in Figure \ref{fig:msre-geom-vert}. The model excludes the toroidal flow distributor near the top
-of the vessel and reshapes the plena into right cylinders. It has the same horizontal cross section
-as the 2-D model (Figure \ref{fig:full-geom}) throughout most of the salt-graphite lattice. In
-addition to modifications in the 2-D model listed in Section \ref{sec:2d-model-setup}, the
-3-D model includes the following modifications:
-
-\begin{itemize}
-  \item Removed the toroidal flow distributor near the top of the vessel.
-  \item Reshaped the upper and lower plena into right cylinders while keeping the salt volume
-    constant.
-  \item Flattened the conical tops of the graphite lattice blocks.
-  \item Regularized the complicated graphite lattice structure at the bottom.
-  \item Set a fixed outer vessel thickness of 2.54 cm.
-  \item Homogenized the lower plenum by 90.8\% salt- 9.2\% INOR-8 (vessel) by volume.
-  \item Removed the reactor access nozzle and the various components within it.
-  \item Added a graphite block at the bottom of the control rod thimbles.
-\end{itemize}
-
-\begin{figure}[p]
-  \centering
-  \includegraphics[width=0.97\columnwidth]{msre-picture}
-  \caption{Vertical cross section of the actual \gls{MSRE} vessel. It has rounded upper and lower
-  plena, and a toroidal flow distributor that distributes the salt inflow evenly as it enters the
-  core through holes on the inside vessel surface of the distributor.}
-  \label{fig:msre-picture}
-\end{figure}
-
-\begin{figure}[p]
-  \centering
-  \includegraphics[width=\columnwidth]{msre-geom-vert}
+    \includegraphics[width=.95\columnwidth]{msre-geom-vert}
     \caption{Vertical cross section of the 3-D numerical \gls{MSRE} model offset by 5.08 cm to show
-  the control rod thimble and homogenized sample basket. The control rod is in its fully inserted
-  position. The homogenized lower plenum consists of 90.8\% salt and 9.2\% INOR-8 alloy by volume.}
+    the control rod thimble and homogenized sample basket.}
     \label{fig:msre-geom-vert}
   \end{figure}
+\end{columns}
+\end{frame}
 
-Table \ref{table:reactor-dimensions} lists relevant dimensions of the 3-D \gls{MSRE} model.
-All material compositions remain
-the same as the 2-D models and the \gls{MSRE} benchmark specifications \cite{fratoni_molten_2020}.
-All 3-D neutronics simulations in this section used the molten salt composition with 65.25 kg
-$^{235}$U loaded (Table \ref{table:2d-salt-composition}), except the simulations for temperature
-reactivity coefficient calculations which had a higher $^{235}$U loading of 67.00 kg (Table
-\ref{table:67-salt-composition}).
+\begin{frame}
+  \frametitle{Hybrid $S_N$-Diffusion Method: 3-D Neutronics Eigenvalue Simulations}
+  \textbf{3-D Full-Core MSRE Model Details}
+  \begin{itemize}
+    \item Hybrid $S_6$-diffusion method
+    \item $^{235}$U concentration at initial criticality
+    \item Uniform temperature at 911 K
+    \item Stabilization factor of $c=250$
+    \item Void constant of $\varsigma=0.5$
+    \item Uncorrected diffusion coefficient values capped at $D_g = 2.5$
+  \end{itemize}
+  \vspace{.2cm}
 
-\begin{table}[htb]
-  \centering
-  \caption{List of reactor dimensions.}
-  \begin{tabular}{l S}
-    \toprule
-    Reactor Specifications & {Value} \\
-    \midrule
-    Height [m] & 2.15 \\
-    Radius [m] & 0.762 \\
-    Vessel thickness [m] & 0.0254 \\
-    Fuel channel length [m] & 0.03048 \\
-    Fuel channel width [m] & 0.01016 \\
-    Graphite lattice pitch [m] & 0.05080 \\
-    \bottomrule
-  \end{tabular}
-  \label{table:reactor-dimensions}
-\end{table}
+  These parameters are required for numerical stability in the air-filled control rod
+  thimble region.
+  \vspace{.2cm}
 
-\begin{table}[htb]
-  \small
-  \centering
-  \setlength\tabcolsep{4pt}
-  \caption{\gls{MSRE} molten salt composition when the $^{235}$U loading was at 65.25 kg.}
-  \begin{tabular}{l S S S S S S S}
-    \toprule
-    Element & {Li} & {Be} & {Zr} & {Hf} & {$^{234}$U} & {$^{235}$U} & {$^{236}$U} \\
-    \midrule
-    Composition [kg] & 507.42 & 293.96 & 513.97 & 0.0029 & 0.68 & 67.00 & 0.28 \\
-    \bottomrule
-  \end{tabular}
-  \begin{tabular}{l S S S S S S}
-    \toprule
-    Element & {$^{238}$U} & {Fe} & {Cr} & {Ni} & {O} & {F} \\
-    \midrule
-    Composition [kg] & 142.02 & 0.75 & 0.13 & 0.14 & 2.27 & 3104.22 \\
-    \bottomrule
-  \end{tabular}
-  \label{table:67-salt-composition}
-\end{table}
+  All hybrid method results are compared with MSRE experimental data, the MSRE numerical benchmark
+  report data (Serpent 2 model) \cite{fratoni_molten_2020}, and the OpenMC model in this work.
+\end{frame}
 
-All neutronics simulations ran on OpenMC under continuous energy mode and on Moltres with the
-hybrid $S_N$-diffusion method. The hybrid simulations ran with a maximum stabilization factor of
-$c=250$, a void constant $\varsigma=0.5$, and a uncorrected diffusion coefficient cap of $D_g=2.5$.
-Refer to Sections \ref{sec:void-treatment} and \ref{sec:diffcoef-cap} for the relevant discussion
-on void treatment for near-void regions and the impact of large diffusion coefficient values on
-solver performance. As for the standard neutron diffusion method, it either converges very slowly
-or does not converge at all when $D_g$ values are uncapped due to excessive diffusive effects in
-the air-filled control rod thimbles. Group-wise $D_g$ values, generated using OpenMC, ranged from
-100 cm to 850 cm in the air regions as compared to the typical 0.5 cm to 2.0 cm range observed in
-non-air regions. Therefore, all neutron diffusion method simulation results in the following
-section were measured with the diffusion coefficient capped at $D_g=2.5$.
-
-Unlike the 2-D hybrid simulations which ran with $S_8$ angular discretization, the 3-D hybrid
-simulations ran with $S_6$ angular discretization due to significantly
-larger computational and memory requirements demanded by the 3-D simulations. Further discussion on
-computational performance may be found at the end of Section \ref{sec:3d-nts}. All simulations,
-except those for temperature reactivity coefficient calculations, ran on isothermal 3-D models at
-911 K.
-
-\subsection{3-D Numerical Results \& Discussion} \label{sec:3d-nts}
-
-\subsubsection{Initial Criticality Configuration}
-
-The \gls{MSRE} achieved initial criticality with Rod 1 inserted to a height of 46.6 inches relative
-to the fully inserted position. For reference, the rods are at a height of 51 inches when fully
-withdrawn. Table \ref{table:initial-crit} shows the $k_\text{eff}$ values from \gls{MSRE}
-experimental data, numerical benchmark data \cite{fratoni_molten_2020}, and this work. All
-numerical models overestimate the $k_\text{eff}$ relative to the experimental value by 1-2\%
-(1000-2000 pcm). In the \gls{MSRE} numerical benchmark report, the authors
-attribute this $k_\text{eff}$ discrepancy largely to discrepancies in the carbon cross section data
-in nuclear data libraries. They noted similar $k_\text{eff}$ overestimations by 1-2\% in benchmark
-results for the HTTR, a graphite-moderated, helium-cooled reactor. The OpenMC and Moltres
-simulations in this work used the same
-ENDF/B-VII.1 nuclear data library as the \gls{MSRE} numerical benchmark report.
-
+\begin{frame}
+  \frametitle{Hybrid $S_N$-Diffusion Method: 3-D Neutronics Eigenvalue Simulations}
+  \textbf{MSRE at Initial Criticality}
   \begin{table}[htb]
@@
 
-Of the four $k_\text{eff}$ values, we expect to see the closest agreement between the Serpent 2
-model from the numerical benchmark and OpenMC from this work. However, the OpenMC \gls{MSRE} model
-differs from the Serpent 2 \gls{MSRE} models due to geometry modifications listed in Sections
-\ref{sec:2d-model-setup} and \ref{sec:3d-model-setup}. The benchmark report quantified the impact
-of several geometry modifications, namely the removal of the flow distributor, the homogenization
-of the sample basket, and the removal of the thermal shield. These modifications contribute to
--$0.098\%$, -$0.037\%$, and -$0.885\%$ change in the $k_\text{eff}$, respectively, for a final
-$k_\text{eff}$ estimate of $1.01093(3)$. This brings the OpenMC $k_\text{eff}$ estimate of
-$1.01308(20)$ to closer agreement with the numerical benchmark value.
+  \begin{itemize}
+    \item The Serpent 2 benchmark model reports a $k_\text{eff}$ of 1.01093(3) after considering some
+      of the geometrical modifications used in the OpenMC model.
+    \item The hybrid and neutron diffusion models agree with the OpenMC model within 700 pcm.
+  \end{itemize}
+\end{frame}
 
-The hybrid method $k_\text{eff}$ estimate is 0.00647 higher than the OpenMC estimate. For this
-initial criticality configuration, the neutron diffusion method $k_\text{eff}$ estimate is closer
-to OpenMC than the hybrid method. Overall, the OpenMC and Moltres models show good agreement with
-the Serpent 2 model from the numerical benchmark.
-Figures * show the neutron group fluxes on the interior of the 3-D models. The salt-graphite
-lattice structure is visible from the group flux distributions. Group 7 and 8 neutron fluxes are
-particularly small near the control rods due to strong thermal neutron capture cross section of
-gadolinium.
-
-\subsubsection{Control Rod Worth}
-
-After achieving initial criticality, \gls{MSRE} researchers ran control rod calibration experiments
-with the \gls{MSRE} at zero-power (typically at about 10 W power output)
-\cite{prince_zero-power_1968}. They measured differential rod worths by measuring the time taken
-for the reactor power to increase by 20 W following a prescribed withdrawal of Rod 1 and applying
-the inhour equation to obtain the reactivity inserted. They repeated this process at different rod
-heights after successive additions of highly enriched uranium capsules. The final integral rod
-worth curve was normalized to the critical $^{235}$U loading of 65.25 kg
-\cite{prince_zero-power_1968}.
-
-For control rod worth calculations, the OpenMC and Moltres simulations ran with Rod 1 at fourteen
-different rod positions from full insertion (0 inch) to full withdrawal (51 inches). The rod
-positional limits of 0 inch and 51 inches represents the full range of deliberate rod motion
-allowable by the control rod drive motor \cite{robertson_msre_1965}.
-
-Figure \ref{fig:rod-worth} shows the integral rod worth curves from the \gls{MSRE} experimental
-data and the OpenMC and Moltres numerical models. By taking the $k_\text{eff}$ when Rod 1 is fully
-inserted as reference, all rod worth curves start at 0 pcm. The change in integral rod worth,
-represented by the gradient in the plot, is steepest as the tip of the rod passes through the
-center of the core. The hybrid method performs well in reproducing the rod worth curve from OpenMC;
-all hybrid method rod worth values fall within one or two standard deviations of OpenMC.
-
-Due to the experimental procedure of computing integral rod worth through cumulative differential
-rod worth measurements, the uncertainty range of the \gls{MSRE} data scales with rod position.
-Both OpenMC and the hybrid method report consistently larger integral rod worths than the
-\gls{MSRE} data across all rod positions. Nevertheless, they are in good agreement with the
-\gls{MSRE} data as compared to the neutron diffusion method, which significantly overestimates the
-integral rod worth.
-Table \ref{table:rod-worth} shows the total rod worth of Rod 1. Similar to observations from Figure
-\ref{fig:rod-worth}, OpenMC and the hybrid method report total rod worths marginally larger than
-the \gls{MSRE} data by 5.1\% and 4.2\%, respectively, while the diffusion method exceeds the
-experimental value by 21.1\%.
-
+\begin{frame}
+  \frametitle{Hybrid $S_N$-Diffusion Method: 3-D Neutronics Eigenvalue Simulations}
+  \textbf{MSRE Rod Worth Measurements}
   \begin{figure}[t]
@@
   \end{figure}
+\end{frame}
 
+\begin{frame}
+  \frametitle{Hybrid $S_N$-Diffusion Method: 3-D Neutronics Eigenvalue Simulations}
+  \textbf{MSRE Rod Worth Measurements}
   \begin{table}[t]
@@
   \end{table}
+  \vspace{.2cm}
 
-The most likely factor for OpenMC and the hybrid method overpredicting the control rod worth is the
-modeling of control rods as solid annular cylinders as opposed to 36 separate control elements
-strung along a flexible stainless steel cable. Figure \ref{fig:rod-bead} depicts a poison element
-and its physical dimensions. Other than the Gd$_2$O$_3$-Al$_2$O$_3$ poison material, each control
-element has a top and bottom closure made of INOR-8 alloy. According to the physical dimensions,
-the closures make up 15.1\% of the total element height. The 3-D numerical \gls{MSRE} model in this
-work omits the outer and inner INOR-8 tubes and the flexible stainless cable, and replaces the top
-and bottom closure with poison material. Fully resolving the closures in Moltres would require
-excessively fine axial mesh resolution and advanced adaptive meshing features not currently
-available. A straightforward corrective measure for future \gls{MSRE} simulations would be to
-retain the solid annular cylinder structure and adjust the rod dimensions or the rod material
-composition to match the expected worth from \gls{MSRE} data or OpenMC simulations with the beaded
-geometry.
-
+  Slight overprediction by OpenMC and the hybrid method due to approximating the control rod
+  as an annular cylinder instead of beaded rod elements.
   \begin{figure}[t]
-  \begin{subfigure}[b]{0.33\columnwidth}
+    \begin{subfigure}[b]{0.22\columnwidth}
       \centering
@@
     \end{subfigure}
-  \begin{subfigure}[b]{0.65\columnwidth}
+    \begin{subfigure}[b]{0.44\columnwidth}
       \centering
@@
   \end{figure}
+\end{frame}
 
-Overall, the results show that the hybrid $S_N$-diffusion method accurately reproduces the control
-rod worth at various levels of insertion. As observed in the 1-D and 2-D models, the 3-D model
-$k_\text{eff}$ values exhibit some discrepancies relative to OpenMC, but the discrepancies do not
-affect rod worth estimates because they are systematic errors. This \gls{MSRE} hybrid model may be
-improved in future work by characterizing and applying neutron albedo at the vacuum boundaries.
+\begin{frame}
+  \frametitle{Hybrid $S_N$-Diffusion Method: 3-D Neutronics Eigenvalue Simulations}
+  \textbf{Computational Performance}
+  \begin{itemize}
+    \item All simulations ran on the Polaris supercomputer system at Argonne National Laboratory.
+    \item Each 3-D full-core simulation consisted of 750M and 1.69B degrees of freedom for the
+      neutron diffusion and $S_N$ subsolvers, respectively.
+    \item 3-D full-core simulations required at least 40 compute nodes with 512 GB RAM each due to
+      strict memory requirements (significant improvements in memory use is possible) and took 2.2
+      h to complete.
+    \item The hybrid method took about four times longer than the neutron diffusion method.
+    \item 30 \% of compute time spent on data transfers between the neutron diffusion and $S_N$
+      subsolvers (Reduced to 18 \% in subsequent simulations after optimizations)
+  \end{itemize}
+\end{frame}
 
-\subsubsection{Temperature Reactivity Coefficient}
+\begin{frame}
+  \frametitle{Hybrid $S_N$-Diffusion Method: 3-D Neutronics Eigenvalue Simulations}
+  \textbf{Strong Scaling Test}
+  \vspace{.2cm}
 
-This work compares isothermal temperature reactivity coefficients calculated from $k_\text{eff}$
-values measured at 872 K and 972 K with 67.86 $^{235}$U loading. This approach is identical to that
-of the \gls{MSRE} experimental measurements \cite{prince_zero-power_1968} and the \gls{MSRE}
-numerical benchmark calculations \cite{fratoni_molten_2020}. The temperature reactivity coefficient
-is calculated by dividing the difference in reactivities by the temperature change (100 K). All
-simulations included the effects of thermal expansion on the material density but left all reactor
-physical dimensions unchanged.
-
-Table \ref{table:temp-coef} compares the isothermal reactivity coefficients from the experimental
-and numerical sources. The OpenMC and Moltres models in this work show closer agreement with the
-\gls{MSRE} data than the Serpent 2 model from the \gls{MSRE} numerical benchmark. All numerical
-results fall within one standard deviation of the \gls{MSRE} data.
-
-\begin{table}[t]
-  \small
+  Performed a strong scaling test on the 3-D quarter-core MSRE model on 10, 20, 40, and 80 compute
+  nodes. The $S_N$ and diffusion subsolvers scale well throughout the test.
+  The $S_N$-diffusion data transfer processes scale poorly beyond 40 nodes.
+  \begin{figure}[t]
     \centering
-  \caption{Isothermal temperature reactivity coefficients with 67.86 kg $^{235}$U loading from
-  \gls{MSRE} data, the \gls{MSRE} numerical benchmark \cite{fratoni_molten_2020}, and the OpenMC
-  and Moltres models in this work.}
-  \begin{tabular}{l S[separate-uncertainty=true,table-format=2.2(2)] S}
-    \toprule
-     & {Temperature coefficient [pcm K$^{-1}$]} & {Percentage discrepancy [\%]}\\
-     \cmidrule(rl){2-2} \cmidrule(l){3-3}
-    \gls{MSRE} experimental data & 13.41(150) & {-}\\
-    Serpent 2 (Numerical benchmark) & 12.51(23) & -6.7 \\
-    OpenMC (This work) & 13.85(43) & 3.3 \\
-    Hybrid (This work) & 13.98 & 4.3 \\
-    Diffusion (This work) & 14.01 & 4.5 \\
-    \bottomrule
-  \end{tabular}
-  \label{table:temp-coef}
-\end{table}
-
-\subsubsection{Computational Performance}
-
-Beyond demonstrating the hybrid $S_N$-diffusion method, its computational performance is of
-interest as a candidate for time-dependent 3-D reactor simulations. All simulations ran on the
-Polaris supercomputer system at \gls{ANL}. Each compute node consists of a 2.8-GHz AMD EPYC Milan
-7543P 32-core CPU, four NVIDIA A100 GPUs, and 512 GB of RAM. The GPUs were not utilized in this
-work. Moltres uses the \gls{MPI} communication protocol for parallelizing simulations. For this
-work, simulations ran with 1-to-1 ratio of \gls{MPI} ranks and processor cores (i.e., 32 \gls{MPI}
-ranks for each compute node).
-
-Every 3-D hybrid method $k$-eigenvalue
-simulation of the \gls{MSRE} consisted of 747,688,760 \glspl{DOF} and 1,690,777,728 \glspl{DOF} for
-the neutron diffusion and $S_N$ components, respectively. Each \gls{DOF} corresponds to a scalar
-or angular group flux variable value defined at a mesh node. Each simulation ran with the distributed
-mesh feature, which partitions and distributes the mesh across however many \gls{MPI} ranks used
-in the simulation. All variable and mesh-linked data are also distributed with the mesh. This
-greatly reduces memory requirements compared to storing the full mesh for each \gls{MPI} rank.
-Each 3-D \gls{MSRE} simulation required at least forty nodes for sufficient memory to run. This
-sets a strict minimum requirement for the 3-D hybrid simulations. A significant fraction of
-memory use in Moltres comes from storing the residual and Jacobian quantities at each \gls{FEM}
-quadrature point. Given that terms in the $S_N$ equations share the same Jacobian formulation
-across all ordinate directions, memory use (and compute time) could be optimized by reducing
-duplicate Jacobian evaluations.
-
-On forty nodes, the simulations
-required approximately 2.2 h to complete for a total of 88 node-hours or 2816 core-hours. The
-distributed mesh capability comes with the cost of increased data transfer times between the $S_N$
-and diffusion solvers due to increased inter-core data transfers. Transfers accounted for
-approximately 0.675 h or 30 \% of the total simulation time. Future research into optimizing mesh
-distribution and transfer-related data caching may help to minimize data transfers and speed up
-hybrid method simulations. The solution time for the initial neutron diffusion calculation with zero
-drift correction also takes up approximately 30 \% of the total simulation time. After subtracting
-the minor computational cost of evaluating the drift term even if it is zero, the hybrid method
-requires about four times the solution time of the standard neutron diffusion method.
-
-For the hybrid method to be viable for large time-dependent simulations, it must scale reasonably
-well on computing clusters. For this work, I investigated strong scaling, which measures how the
-solution time decreases when more cores are used while keeping the problem size fixed. This strong
-scaling study used 3-D quarter-core models run on 10, 20, 30, and 40 compute nodes.
-Figure * shows the results of this scaling study.
-The hybrid method implementation scales well as shown by the nearly linear scaling between the
-solution time and the number of processors. Therefore, the hybrid method appears promising for the
-modeling of larger reactor analyses with little scaling performance penalty.
-
-% Strong scaling tests & discussion
-% Difficulties with 3-D modeling
-% SN void treatment constants, Transfer times
-
-\section{Summary}
-
-This chapter covers work demonstrating the hybrid $S_N$-diffusion method implemented in Moltres
-through 1-D, 2-D, and 3-D $k$-eigenvalue simulations of \gls{MSRE} models. The hybrid method
-combines stengths of the $S_N$
-and diffusion methods by applying transport corrections generated using the high-fidelity $S_N$
-method near control rods while treating the rest of the problem with the lower-fidelity diffusion
-method for tractable time-dependent 3-D reactor analyses. This is achieved by identifying and
-setting up a subregion containing the control rod and its vicinity for $S_N$ transport
-calculations. The $S_N$ and diffusion solvers are coupled through boundary sources at the subregion
-boundaries. The hybrid method applies an adaptive algorithm for discarding inaccurate transport
-correction values near the subregion boundaries and preserving the smoothness of the neutron flux
-gradient. In this chapter, the 1-D simulations provided verification of the hybrid method for
-estimating control rod worth. The simulations also compared two different implementations of the
-hybrid method which differed by the use of drift or diffusion correction schemes.
-The 2-D and 3-D simulations extended the hybrid method to multi-dimensional problems. Additionally,
-the 3-D simulations validated the numerical \gls{MSRE} model against \gls{MSRE} experimental data
-and provided useful insight on the computational performance of the hybrid method for
-large reactor analyses.
-
-Six 1-D test cases based on the \gls{MSRE} with increasing geometric complexity were run using the
-standard $S_N$ method, the hybrid method, and standard neutron diffusion method and verified against
-reference OpenMC Monte Carlo neutron transport simulations under continuous neutron energy (OpenMC-CE) and
-multigroup (OpenMC-MG) modes. The $S_N$ implementation with nonlinear diffusion acceleration
-showed good agreement with OpenMC-MG which indicated that discrepancies between those codes and
-OpenMC-CE largely arose from the eight neutron energy group structure for multigroup simulations.
-Both hybrid method implementations, with drift or diffusion correction schemes, showed similar
-performance in the 1-D cases. While the $k_\text{eff}$ estimates from the hybrid method deviated
-from the $S_N$ method and OpenMC-MG by 100 to 400 pcm, the hybrid method accurately reproduces
-control rod worths. Rod worth estimates from OpenMC-MG, $S_N$, and hybrid methods differed from
-OpenMC-CE by 2 to 3 \%. The hybrid method improves on the neutron diffusion method whose rod worth
-error estimate is 8 \%. The hybrid method also produced better agreement in neutron flux
-distributions with the neutron transport methods compared to the neutron diffusion method.
-
-Analysis on the drift and diffusion correction schemes showed that the diffusion correction
-parameter led to numerous discontinuities tending to infinity values near points in the problem
-domain where flux peaks and troughs occur. These discontinuities and negative diffusion coefficient
-values are unphysical and would lead to numerical stability issues. Therefore, the hybrid method
-implementation with drift correction scheme was chosen for the remaining 1-D analyses and the
-2-D and 3-D analyses.
-
-Analysis on the impact of $S_N$ correction subregion size found that it has little impact on the
-control rod worth estimates as long as the subregion size is kept consistent (i.e., same size) when
-performing control rod worth measurements. Rod worths varied by less than 0.35 \% across the
-different subregion sizes investigated. A separate analysis showed that the $k_\text{eff}$
-estimates from the hybrid method exhibit superlinear convergence rate when tightening the $S_N$
-convergence tolerance value. The convergence tolerance value also impacts the number of outer
-iterations between $S_N$ and diffusion solvers. After just two outer iterations, the error in
-$k_\text{eff}$ fell to below 10$^{-7}$.
-
-The hybrid method continued to show improved control rod estimates over the neutron diffusion
-method in 2-D quarter-core and full-core \gls{MSRE} simulations with up to three rods inserted in the
-full-core model. The hybrid method reported rod worth error magnitudes of less than 40 pcm relative
-to OpenMC-CE, which are significant improvements over the neutron diffusion method error estimates
-ranging from 569 pcm to 1484 pcm. The hybrid method also showed significant improvements in the
-absolute mean and maximum errors in fuel channel power distributions relative to the reference
-OpenMC-CE distribution.
-
-For 3-D simulations, the \gls{MSRE} experimental data and the \gls{MSRE} numerical benchmark
-report \cite{fratoni_molten_2020} both served as references for model validation. The numerical
-benchmark is included in the International Reactor Physics Experiment Evaluation Project (IRPhEP)
-handbook. The
-$k_\text{eff}$ estimates of the \gls{MSRE} at initial criticality from OpenMC, the hybrid method,
-and the neutron diffusion method in this work showed good agreement with the \gls{MSRE} numerical
-benchmark. All numerical $k_\text{eff}$ estimates, including that from the benchmark report
-\cite{fratoni_molten_2020}, exceeded the \gls{MSRE} experimental value by 1-2 \% possibly due to
-biases and uncertainties in the nuclear data library for graphite. Temperature reactivity
-coefficient values from OpenMC, the hybrid method, and the neutron diffusion method also show good
-agreement with the \gls{MSRE} data within experimental uncertainty with percentage discrepancies of
-about 3 \%, but differ from the numerical benchmark which has a percentage discrepancy of -6.7 \%.
-
-Control rod worth calculations were also performed for Rod 1 of the \gls{MSRE} under various levels
-of insertion from its fully inserted to its fully withdrawn state. OpenMC and the hybrid method
-showed excellent agreement at every stage within one standard deviation. Due to geometric
-approximations of the control rods in the numerical models, they overestimated the total
-rod worth relative to the \gls{MSRE} experimental data by approximately 4-5 \%. The hybrid method
-compares favorably against the neutron diffusion method which overestimated the total worth by
-21.1 \%.
-
-The 3-D simulations performed well under the strong scaling test which studies how the solution
-time scales with the number of processors used to run the simulation. The 3-D simulations showed
-close to linear scaling within the scope of the strong scaling test. Therefore, the hybrid method
-exhibits good computational scaling performance which is essential for large 3-D reactor analyses.
-Excessively large memory requirements were mitigated through the distributed mesh capability in
-Moltres. Some performance optimizations may be possible to reduce data transfer times
-between the $S_N$ and diffusion solvers which currently take up about 30 \% of the total solution
-time as a result of the inter-processor communications.
-
-Overall, the hybrid $S_N$-diffusion method provides a significant improvement over the neutron
-diffusion method at only four times the computational expense. Control rod estimates
-show good agreement with reference Monte Carlo solutions and \gls{MSRE} experimental data.
-Further investigations into characterizing and minimizing neutron leakage rates at vacuum
-boundaries may help with reducing $k_\text{eff}$ discrepancies with reference Monte Carlo
-solutions.
+    \begin{subfigure}[b]{0.32\columnwidth}
+      \centering
+      \includegraphics[width=\columnwidth]{scaling-total}
+      \caption{Total wall time}
+    \end{subfigure}
+    \begin{subfigure}[b]{0.32\columnwidth}
+      \centering
+      \includegraphics[width=\columnwidth]{scaling-solver}
+      \caption{Solver wall time}
+    \end{subfigure}
+    \begin{subfigure}[b]{0.32\columnwidth}
+      \centering
+      \includegraphics[width=\columnwidth]{scaling-transfer}
+      \caption{Transfer wall time}
+    \end{subfigure}
+    \caption{The total, solver, and transfer wall time of hybrid method simulations of the 3-D
+    quarter-core model on 10, 20, 40, and 80 compute nodes (32 processors per node) of the Polaris
+    supercomputer. All axes are in log scale.}
+    \label{fig:scaling}
+  \end{figure}
+\end{frame}

### 2024-12-04 presentation/cnrs.tex
Presentation commit

\begin{frame}
  \frametitle{V\&V Study 1: Verification of Moltres with the CNRS Benchmark}

  Published in \textit{S.M. Park, M. Munk, "Verification of Moltres for Multiphysics Simulations of
    Fast-Spectrum Molten Salt Reactors," Annals of Nuclear Energy, vol. 173, Aug 2022.}
  \begin{columns}
    \column[t]{6.5cm}
    \textbf{Description of the CNRS Benchmark \cite{tiberga_results_2020}}
    \vspace{.2cm}

      \begin{itemize}
        \item A numerical benchmark for multiphysics software dedicated to modeling ``pool-type''
          fast-spectrum MSRs
        \item Problem Setup
          \begin{itemize}
            \item LiF-BeF$_2$-UF$_4$ molten salt
            \item Six neutron energy groups
            \item Eight precursor groups
            \item Volumetric conjugate heat sink term
            \item Incompressible flow
          \end{itemize}
        \item Consists of three phases
          \begin{itemize}
            \item Phase 0: Single-physics verification
            \item Phase 1: Steady-state coupling
            \item Phase 2: Time-dependent coupling
          \end{itemize}
      \end{itemize}
    \column[t]{3.5cm}
    \begin{figure}
      \centering
      \includegraphics[width=\columnwidth]{../images/cnrs-geometry}
      \caption{CNRS benchmark problem domain \cite{tiberga_results_2020}}
    \end{figure}
  \end{columns}
\end{frame}

\begin{frame}
  \frametitle{V\&V Study 1: Verification of Moltres with the CNRS Benchmark}
  \begin{block}{\textbf{Study Outcome}}
    Moltres is consistent with the other multiphysics software in the benchmark problems.
  \end{block}
  \begin{columns}
    \hfill
    \column[t]{4cm}
    \vspace{.3cm}
    \begin{figure}
      \centering
      \includegraphics[width=\columnwidth]{../images/full-coupled}
      \caption{Step 1.4 - Temperature distribution and flow streamlines for the fully
      coupled system \cite{park_verification_2022}.}
    \end{figure}
    \column[t]{4cm}
    \begin{figure}
      \centering
      \includegraphics[width=\columnwidth]{../images/2-1-gain-plot}
      \caption{Step 2.1 - Bode gain plot of the frequency response of the fully
      coupled system \cite{park_verification_2022}.}
    \end{figure}
    \column[t]{4cm}
    \begin{figure}
      \centering
      \includegraphics[width=\columnwidth]{../images/2-1-phase-plot}
      \caption{Step 2.1 - Bode phase plot of the frequency response of the fully
      coupled system \cite{park_verification_2022}.}
    \end{figure}
    \hfill
  \end{columns}
\end{frame}

### 2024-12-04 presentation/conclusion.tex
Presentation commit

%\begin{frame}
%  \frametitle{Conclusion}
%  \begin{block}{\textbf{Objectives}}
%    \begin{enumerate}
%      \item \textbf{Improve Multiphysics Capabilities in Moltres}
%      \begin{itemize}
%        \item \textbf{Verify and Validate Existing Multiphysics Coupling Capabilities in Moltres}
%        \item \textbf{Implement a RANS-based Turbulence Model in Moltres}
%      \end{itemize}
%      \item \textbf{Develop a Hybrid Neutronics Method for Control Rod Modeling in Moltres}
%    \end{enumerate}
%  \end{block}
%  \begin{block}{\textbf{Research Impact}}
%    \begin{itemize}
%      \item \textbf{Improve Moltres as a simulation tool for MSR multiphysics modeling}
%      \item \textbf{Allow for a wider range of MSR simulation scenarios on Moltres involving
%        turbulent flow and control rod movement}
%      \item \textbf{Analysis of neutron transport effects in MSRs that are neglected by neutron
%        diffusion methods}
%%      \item \textbf{Publication of numerical data for MSRE control rod insertion/withdrawal
%%        simulations}
%    \end{itemize}
%  \end{block}
%\end{frame}

\begin{frame}
  \frametitle{Conclusion}
  \begin{block}{\textbf{Moltres Multiphysics V\&V Studies}}
    \textbf{Relevance}
    \begin{itemize}
      \item Existing multiphysics coupling capabilities in Moltres must be verified and validated
    \end{itemize}
    \textbf{Results Presented}
    \begin{itemize}
      \item CNRS numerical benchmark study with Moltres
      \item MSRE pump transient V\&V study with Moltres
    \end{itemize}
    \textbf{Major takeaways}
    \begin{itemize}
      \item Moltres is highly consistent with other MSR simulation tools in accurately modeling:
      \begin{itemize}
        \item Salt flow-induced temperature and DNP drift
        \item Temperature ractivity feedback due to Doppler broadening and thermal expansion
        \item Buoyancy-driven flow due to temperature gradients
        \item Time-varying and looped precursor flow
      \end{itemize}
    \end{itemize}
  \end{block}
\end{frame}

\begin{frame}
  \frametitle{Conclusion}
  \begin{block}{\textbf{Spalart-Allmaras Turbulence Model Implementation in Moltres}}
    \textbf{Relevance}
    \begin{itemize}
      \item Moltres requires turbulence modeling capability to model turbulent salt
        flow in MSRs
    \end{itemize}
    \textbf{Results Presented}
    \begin{itemize}
      \item Implementation of a Spalart-Allmaras turbulence model in Moltres
      \item V\&V studies of turbulent channel, turbulent pipe, and backward-facing step flow
    \end{itemize}
    \textbf{Major Takeaways}
    \begin{itemize}
      \item The Spalart-Allmaras turbulence model in Moltres is accurate for turbulent channel and
        pipe flows
      \item It is also consistent with other implementations when simulating the
        backward-facing step flow problem.
    \end{itemize}
  \end{block}
\end{frame}

\begin{frame}
  \frametitle{Conclusion}
  \begin{block}{\textbf{Hybrid $S_N$-Diffusion Method}}
    \textbf{Relevance}
    \begin{itemize}
      \item No MSR studies exist for time-dependent MSR simulations involving reactivity-initiated
        transients using control rods
    \end{itemize}
    \textbf{Results Presented}
    \begin{itemize}
      \item Development and implementation of a hybrid $S_N$-diffusion method for
        accurate control rod modeling in Moltres
      \item Verification, validation, and demonstration of the hybrid $S_N$-diffusion
        method through 1-D, 2-D, \& 3-D analyses of the MSRE
    \end{itemize}
    \textbf{Major Takeaways}
    \begin{itemize}
      \item The hybrid method is an iterative method that applies a $S_N$ subsolver around control
        rod regions for improved rod worth estimates
      \item The hybrid method applies an adaptive domain decomposition technique to
        couple the $S_N$ subsolver and the neutron diffusion solver
      \item The hybrid method accurately reproduces control rod worths at only four times the
        computational cost of the neutron diffusion method
      \item The hybrid method exhibits decent scalability on high-performance computing clusters
    \end{itemize}
  \end{block}
\end{frame}

### 2024-12-04 presentation/future.tex
Presentation commit

\begin{frame}
  \begin{block}{\textbf{Future Work}}
    \begin{itemize}
      \item Extend the Spalart-Allmaras turbulence model implementation
      \begin{itemize}
        \item Investigate more effective preconditioning schemes beyond the
          current additive Schwarz method
        \item Implement wall functions to reduce fine mesh requirements near the wall for high
          Reynolds flows (Re$\approx 10^6$)
      \end{itemize}
      \item Improve the computational performance of the hybrid $S_N$-diffusion method
      \begin{itemize}
        \item General code optimizations
        \item Reduce compute and memory use through shared FEM Jacobian evaluations for all angular
          fluxes at the same position
        \item Custom mesh distribution and data transfer caching routines to minimize data transfer time
      \end{itemize}
      \item Further 3-D full-core reactor analyses with the hybrid method
      \begin{itemize}
        \item Other MSRs designs with control rods include the Chinese TMSR-LF1,
          Terrapower's MCRE, and the Seaborg Technology's CMSR
        \item These reactor designs allow for asymmetric reactivity-initiated transients due to
          non-centrally located control rods.
      \end{itemize}
    \end{itemize}
  \end{block}
\end{frame}

### 2024-12-04 presentation/msr.tex
Presentation commit

\begin{frame}
  \frametitle{What is a Molten Salt Reactor (MSR)?}
	\begin{itemize}
	  \item Advanced nuclear reactor fueled by fissile
		material dissolved in a molten salt mixture
	  \item In most MSR designs, the fuel salt mixture doubles as the primary coolant for the
        reactor
	  \item Both thermal- and fast-spectrum configurations are viable
	\end{itemize}
	\begin{figure}
	  \centering
	  \includegraphics[width=.5\textwidth]{./images/msr}
      \caption{Schematic diagram of a general channel-type MSR concept
      \cite{u.s._doe_nuclear_energy_research_advisory_committee_technology_2002}.}
	  \label{fig:msr}
	\end{figure}
\end{frame}

%\begin{frame}
%  \frametitle{What is a Molten Salt Reactor (MSR)?}
%  \textbf{Advantages of MSRs over other reactor types}
%  \begin{itemize}
%    \item Robust passive safety
%      \begin{itemize}
%        \item strong temperature feedback
%        \item large thermal margin to boiling
%      \end{itemize}
%    \item Freeze plug drains fuel salt into containment tanks under emergency situations
%    \item Online refueling and reprocessing can:
%      \begin{itemize}
%        \item reduce reactor downtime
%        \item reduce required fissile inventory in reactor at any given time
%        \item reduce transuranic waste produced
%        \item allow for $^{233}$U breeding from $^{232}$Th
%      \end{itemize}
%    \item Viable as fast-spectrum breeder/burner reactor
%    \item Can provide high-temperature heat for industrial heat applications
%  \end{itemize}
%\end{frame}

\begin{frame}
  \frametitle{Molten Salt Reactor Modeling \& Simulation}
  \textbf{While modeling MSRs is not necessarily more difficult than modeling solid-fueled
  reactors, we must adapt our software tools to accurately model phenomena unique to MSRs.}\\

  \textbf{Phenomena which pose challenges for MSR modeling}
  \begin{itemize}
	\item Strong multiphysics interactions involving neutron flux, temperature, and flow in the
      reactor core
	  \begin{itemize}
		\item Strong temperature feedback due to thermal expansion of liquid fuel salt
		\item Movement of \gls{DNP} along primary coolant loop
        \item Advection-dominated heat transfer
	  \end{itemize}
    \item Loss of delayed neutrons to out-of-core decay
    \item Complex turbulent flow effects in MSRs
  \end{itemize}
\end{frame}

### 2024-12-04 presentation/msre-transients.tex
Presentation commit

\begin{frame}
  \frametitle{V\&V Study 2: MSRE Pump Start-up \& Coast-Down Transients}
  \textbf{Description of MSRE Pump Experiments}
  \begin{itemize}
    \item Pump start-up and coast-down transient flow-rate tests on the MSRE under zero-power
  conditions
    \item Starting from zero-power critical states with/without forced flow, the fuel salt pump
  was coasted down/started up
    \item Reactor was kept critical by a flux-servo controller moving the rod in response to flux
      changes
  \end{itemize}
  \begin{columns}
    \column[t]{5.5cm}
    \begin{figure}
      \centering
      \includegraphics[width=.8\columnwidth]{images/msre-startup}
      \caption{MSRE start-up pump speed and flow rate experimental data
      \cite{prince_zero-power_1968}.}
    \end{figure}
    \column[t]{5.5cm}
    \begin{figure}
      \centering
      \includegraphics[width=.95\columnwidth]{images/msre-coastdown}
      \caption{MSRE coast-down pump speed and flow rate experimental data
      \cite{prince_zero-power_1968}.}
    \end{figure}
  \end{columns}
\end{frame}

%\begin{frame}
%  \frametitle{V\&V Study 2: MSRE Pump Start-up \& Coast-Down Transients}
%  \begin{block}{\textbf{MSR Phenomena Involved}}
%    \begin{itemize}
%      \item DNP drift under time-varying flow
%      \item Loss of delayed neutrons due to out-of-core DNP decay
%    \end{itemize}
%  \end{block}
%  \begin{block}{\textbf{Aim of the Study}}
%    Reproduce the reactivity curve of the control rod response with a 2-D axisymmetric model of the
%    MSRE in Moltres.
%  \end{block}
%  \begin{block}{\textbf{Study Objectives}}
%    \begin{itemize}
%      \item Develop a V\&V study based on the MSRE pump start-up \& coast-down
%        transients that is easily reproducible
%      \item Verify Moltres against QuasiMolto in collaboration with Aaron Reynolds
%        (formerly at Oregon State University)
%      \item Validate Moltres against MSRE experimental data for modeling \gls{DNP} drift
%    \end{itemize}
%  \end{block}
%\end{frame}

\begin{frame}
  \frametitle{V\&V Study 2: MSRE Pump Start-up \& Coast-Down Transients}
  \begin{columns}
    \column{6.5cm}
    \begin{block}{\textbf{MSR Phenomena Involved}}
      \begin{itemize}
        \item DNP drift under time-varying flow
        \item Loss of delayed neutrons due to out-of-core DNP decay
      \end{itemize}
    \end{block}
    \begin{block}{\textbf{Aim of the Study}}
      \begin{itemize}
         \item Develop a V\&V study based on the MSRE pump start-up \& coast-down
           transients that is easily reproducible
         \item Verify Moltres against QuasiMolto in collaboration with Aaron Reynolds
           (formerly at Oregon State University)
         \item Validate Moltres against MSRE experimental data for modeling \gls{DNP} drift
       \end{itemize}
    \end{block}
    \begin{block}{\textbf{Modeling Approach}}
      \begin{itemize}
        \item Two neutron energy groups and six DNP groups
        \item Group constants generated using a representative unit cell model in SCALE
        \item Ignored upper and lower plena in the geometry
        \item Looped salt flow with a salt circulation time of 25.2 s
        \item $k$-eigenvalue calculation at every time-step in place of reactivity control
        \item Timestep size of 0.1 s
      \end{itemize}
    \end{block}
    \column{4.5cm}
    \begin{figure}[htb]
      \centering
      \includegraphics[width=0.8\columnwidth]{images/msre-2d}
      \caption{2-D axisymmetric model of the \gls{MSRE}.The gray and white regions represent fuel salt
      and graphite, respectively. Fuel salt channels are 1.25-cm wide and spaced at 5-cm intervals.}
      \label{fig:pump-geom}
    \end{figure}
  \end{columns}
\end{frame}

\begin{frame}
  \frametitle{V\&V Study 2: MSRE Pump Start-up \& Coast-Down Transients}
  \textbf{Static \& Steady-Flow Conditions}
  \begin{table}[htb]
    \centering
    \caption{Multiplication factors $k_\text{eff}$ under static and steady salt flow conditions, and
    reactivity changes $\Delta\rho$ due to \gls{DNP} flow from the Moltres and QuasiMolto \gls{MSRE}
    models.}
    \begin{tabular}{l S S S}
      \toprule
      \multirow{2}{*}{Code} & \multicolumn{2}{c}{$k_\text{eff}$} & {$\Delta\rho$ due to} \\
                            & {Static} & {Steady flow} & {\gls{DNP} flow} \\
                            \cmidrule(r){1-1} \cmidrule(rl){2-3} \cmidrule(l){4-4}
      Moltres & 1.05625 & 1.05311 & -282.4 \\
      QuasiMolto & 1.05603 & 1.05290 & -281.7 \\
      \bottomrule
    \end{tabular}
    \label{table:msre-pump-keff}
  \end{table}

  \begin{itemize}
    \item Excellent agreement between Moltres and QuasiMolto under static and steady-flow conditions prior
  to the pump transients.
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{V\&V Study 2: MSRE Pump Start-up \& Coast-Down Transients}
  \textbf{Neutron Flux Distributions Under Static Conditions}
  \begin{figure}[h]
    \centering
    \begin{subfigure}[b]{0.48\columnwidth}
      \centering
      \includegraphics[width=\columnwidth]{centerline_flux}
      \caption{Normalized centerline neutron fluxes.}
      \label{fig:centerline-flux-dist}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.48\columnwidth}
      \centering
      \includegraphics[width=\columnwidth]{midplane_flux}
      \caption{Normalized midplane neutron fluxes.}
      \label{fig:midplane-flux-dist}
    \end{subfigure}
    \caption{Neutron flux distributions comparing QuasiMolto and Moltres
    \gls{MSRE} models under static conditions. Maximum relative difference $\approx 0.8 \%$.}
    \label{fig:centerline-flux}
  \end{figure}
\end{frame}

\begin{frame}
  \frametitle{V\&V Study 2: MSRE Pump Start-up \& Coast-Down Transients}
  \textbf{DNP Distributions Under Static Conditions}
  \begin{figure}[t]
    \centering
    \begin{subfigure}[b]{0.48\columnwidth}
      \centering
      \includegraphics[width=\columnwidth]{centerline_pre}
      \caption{Normalized centerline \gls{DNP} distributions.}
      \label{fig:centerline-pre-dist}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.48\columnwidth}
      \centering
      \includegraphics[width=\columnwidth]{midplane_pre}
      \caption{Normalized midplane \gls{DNP} distributions.}
      \label{fig:midplane-pre-dist}
    \end{subfigure}
    \caption{\gls{DNP} distribution from Moltres and ratios comparing QuasiMolto and
    Moltres models under static \gls{MSRE} conditions. Maximum relative difference $\approx 0.1 \%$.}
    \label{fig:centerline-pre}
  \end{figure}
\end{frame}

\begin{frame}
  \frametitle{V\&V Study 2: MSRE Pump Start-up \& Coast-Down Transients}
  \textbf{Pump Start-Up Transient Results}
  \begin{columns}
  \column{5.5cm}
  \begin{figure}[t]
    \centering
    \includegraphics[width=\columnwidth]{start-up-v2-reactivity}
    \caption{Reactivity loss (relative to static no-flow conditions) during the \gls{MSRE} pump
    start-up transient.}
    \label{fig:start-up-reactivity}
  \end{figure}
  \column{5.5cm}
  \textbf{Observations}
  \begin{itemize}
    \item Steep initial reactivity loss due to DNP outflow
    \item Reactivity loss plateaus temporarily before dropping due to DNP reentry
    \item Circulation of the excess DNP cause reactivity oscillations
    \item Oscillations eventually dissipate due to decay of excess DNP
    \item Excellent agreement between Moltres and QuasiMolto
  \end{itemize}
  \end{columns}
\end{frame}

\begin{frame}
  \frametitle{V\&V Study 2: MSRE Pump Start-up \& Coast-Down Transients}
  \textbf{Pump Start-Up Transient Results}
  \begin{columns}
  \column{5.5cm}
  \begin{figure}[t]
    \centering
    \includegraphics[width=\columnwidth]{start-up-v2-reactivity}
    \caption{Reactivity loss (relative to static no-flow conditions) during the \gls{MSRE} pump
    start-up transient.}
    \label{fig:start-up-reactivity}
  \end{figure}
  \column{5.5cm}
  \textbf{Discrepancies against MSRE Data}
  \begin{itemize}
    \item MSRE reported a smaller final reactivity loss
    \item MSRE did not report a reactivity plateau (due to DNP present in the upper and lower plena)
    \item MSRE did not log significant reactivity oscillations after the first peak
  \end{itemize}
  \end{columns}
\end{frame}

\begin{frame}
  \frametitle{V\&V Study 2: MSRE Pump Start-up \& Coast-Down Transients}
  \textbf{Pump Coast-Down Transient Results}
  \begin{columns}
  \column{5.5cm}
  \begin{figure}[t]
    \centering
    \includegraphics[width=\columnwidth]{coast-down-v2-reactivity}
    \caption{Reactivity loss (relative to static no-flow conditions) during the \gls{MSRE} pump
    coast-down transient.}
    \label{fig:start-up-reactivity}
  \end{figure}
  \column{5.5cm}
  \textbf{Observations}
  \begin{itemize}
    \item Reactivity loss remains briefly stable at the start due to flow inertia
    \item Monotonic decline in reactivity loss over time
  \end{itemize}
  \end{columns}
\end{frame}

\begin{frame}
  \frametitle{V\&V Study 2: MSRE Pump Start-up \& Coast-Down Transients}
  \textbf{Study Outcome}
  \begin{itemize}
    \item Moltres and QuasiMolto are highly consistent with each other in their looped DNP flow
      modeling capabilities
    \item Moltres and QuasiMolto reproduced key reactivity behavior in response to varying salt
      flow velocities
    \item Some discrepancies observed between the numerical and experimental results due to the
      omission of upper and lower plena
    \item This study can be used for verifying advection capabilities in other MSR simulation tools
    \item The distinct oscillations serve as valuable reference points to facilitate code-to-code
      comparisons
  \end{itemize}
\end{frame}

### 2024-12-04 presentation/objectives.tex
Presentation commit

\begin{frame}
  \frametitle{Objectives}
  \begin{block}{\textbf{Overarching Goal of this Work}}
    \textbf{Improve Moltres as a MSR simulation tool by enabling a wider range of MSR simulation
    scenarios.}
  \end{block}
  \begin{block}{\textbf{Objectives}}
    \begin{enumerate}
      \item \textbf{Improve Multiphysics Capabilities in Moltres}
      \begin{itemize}
        \item \textbf{Verify and Validate Existing Multiphysics Coupling Capabilities in Moltres}
        \item \textbf{Implement a RANS-based Turbulence Model in Moltres}
      \end{itemize}
      \item \textbf{Develop a Hybrid Neutronics Method for Control Rod Modeling in Moltres}
    \end{enumerate}
  \end{block}
\end{frame}

### 2024-12-04 presentation/review.tex
Presentation commit

\begin{frame}
  \frametitle{Hybrid $S_N$-Diffusion Method: Literature Review}
  \textbf{Transport-Correction Techniques For Neutron Diffusion-Based Solvers}
  \vspace{.3cm}

  Techniques for augmenting the neutron diffusion method (or equivalent low-order equations) with
  corrections derived from neutron transport:
  \begin{itemize}
    \item Ronen method \cite{ronen_accurate_2004, tomatis_ronen_2021, gross_comprehensive_2023}
    \item Multilevel quasi-diffusion method \cite{goldin_quasi-diffusion_1964, anistratov_computational_2012, tamang_multilevel_2014}
    \item Multischeme transport method \cite{wang_hybrid_2017}
    \item Hybrid transport-diffusion method \cite{anistratov_computational_2012, stehle_hybrid_2014}
  \end{itemize}
  These techniques generally require:
  \begin{itemize}
    \item High-order neutron transport solver
    \item Corrective terms for the neutron diffusion equations
    \item Boundary coupling scheme (for spatial domain decomposition)
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Hybrid $S_N$-Diffusion Method: Literature Review}
  \textbf{The Ronen Method \cite{ronen_accurate_2004, tomatis_ronen_2021, gross_comprehensive_2023}}
  \vspace{.2cm}

  Starting with an initial neutron diffusion flux solution, a transport
  expression is used to iteratively improve the flux solution by updating the
  diffusion coefficients:
  %
  \begin{align}
    D(\vec{r},E) =& -\frac{J_{tr}(\vec{r},E)}{\nabla \phi(\vec{r},E)}
    \label{eq:ronen}
    \shortintertext{where}
    J_{tr} =& \mbox{ transport-derived neutron current.} \nonumber
  \end{align}
  %
  In 1-D slab geometry, the integral expression for calculating the current is:
  %
  \begin{align}
    J(x,E) = \frac{1}{2}\int^a_0 dx'\ &E_2[\tau(x',x,E)sgn(x-x')q_0(x',E) \nonumber \\
    &+\frac{3}{2}\int^a_0dx' \ E_3[\tau(x',x,E)]q_1(x',E)
  \end{align}
  \textbf{Limitation: Demonstrated for 1-D geometries only.}
\end{frame}

%\begin{frame}
%  \frametitle{Hybrid $S_N$-Diffusion Method: Literature Review}
%  \textbf{Transport-Derived Diffusion Closures \cite{pounders_diffusion_2009}}
%  \vspace{.2cm}
%  
%  Pounders \& Rahnema developed two separate methods which also
%  generate space-dependent diffusion coefficients for transport corrections.
%  \vspace{.1cm}
%  \begin{columns}
%    \column[t]{.5\textwidth}
%    \textbf{1) Averaged Eddington Factor Diffusion Theory}
%    %
%    \begin{align}
%      \small
%      E_g(z) =& \frac{\int^1_{-1} \mu^2\psi(z,\mu)d\mu}{\int^1_{-1} \psi(z,\mu)d\mu}
%    \end{align}
%    %
%    \begin{align}
%      \small
%      D^{AEF}_g(z) =& E^i_g\left[\hat{\Sigma}_{t,g}-\sum^G_{g'=1}\hat{\Sigma}^{g'\rightarrow g}_{s1}
%      \frac{\hat{J}_{g'}}{\hat{J}_g}\right]^{-1}
%    \end{align}
%    \hfill
%    \column[t]{.5\textwidth}
%    \textbf{2) High-Order Empirical Diffusion Coefficients}
%    \begin{align}
%      D^i_g =& -\frac{\left(z_{i+1}-z_i\right) \bar{J}_g}{\left[\phi_g(z_{i+1})-\phi_g(z_i)\right]}
%      \label{eq:emp}
%    \end{align}
%  \end{columns}
%  \textbf{Limitation: Both methods require a priori knowledge of the neutron flux and current
%  solution.}
%\end{frame}

\begin{frame}
  \frametitle{Hybrid $S_N$-Diffusion Method: Literature Review}
  \textbf{Multilevel Quasi-Diffusion Method}
  \vspace{.2cm}

  Gol'din originally developed the quasi-diffusion (QD) method as a transport method acceleration
  scheme \cite{goldin_quasi-diffusion_1964}.
  \begin{gather}
    \mbox{\small Zeroth angular moment: } 
    \nabla\cdot J(\vec{r}) + \Sigma_t(\vec{r})\phi(\vec{r}) = \Sigma_{s,0}(\vec{r})\phi(\vec{r}) +
    Q_0(\vec{r}) \label{eq:qd-low-0th} \\
    \mbox{\small First angular moment: }
    \nabla\cdot\left(\vec{\vec{E}}(\vec{r})\phi(\vec{r})\right) + \Sigma_t(\vec{r})J(\vec{r}) =
    \Sigma_{s,1}(\vec{r})J(\vec{r}) + Q_1(\vec{r}) \label{eq:qd-low-1st}
    \shortintertext{\small where}
    \vec{\vec{E}} = \frac{\int_{4\pi}d\hat{\Omega}\ \hat{\Omega}\hat{\Omega}\psi(\vec{r})}{
    \int_{4\pi}d\hat{\Omega}\ \psi(\vec{r})} = \mbox{\small Eddington tensor}.
  \end{gather}
  %
  Gol'din later extended it to a three-level method by deriving effective gray (one-group)
  low-order QD equations \cite{anistratov_solution_1986}.
\end{frame}

\begin{frame}
  \frametitle{Hybrid $S_N$-Diffusion Method: Literature Review}
  \textbf{Multilevel Quasi-Diffusion Method}
  \vspace{.2cm}

  Tamang \& Anistratov applied the QD method to 1-D time-dependent multiphysics neutron transport
  and heat transfer problems by iteratively coupling the gray low-order equations to the heat
  transfer and precursor equations \cite{tamang_multilevel_2014}.
  \vspace{.2cm}

  Reynolds \& Palmer applied the same approach for steady-state simulations of a 2-D axisymmetric,
  two-group MSRE model \cite{reynolds_analysis_2023}.
  %
  \begin{figure}[h]
    \centering
    \includegraphics[width=.6\columnwidth]{images/reynolds-flux}
    \caption{Fast and thermal fluxes of the 2-D axisymmetric MSRE model by Reynolds \& Palmer
    \cite{reynolds_analysis_2023}.}
    \label{fig:reynolds-flux}
  \end{figure}
\end{frame}

\begin{frame}
  \frametitle{Hybrid $S_N$-Diffusion Method: Literature Review}
  \textbf{Multischeme Transport Method}
  \vspace{.2cm}

  Implemented in the Griffin reactor physics application. This method involves making predetermined
  domain decompositions to be treated with the $S_N$, $P_N$, or diffusion methods. Adjoining
  subdomains are coupled through the Lagrange multiplier interface condition:
  \begin{gather}
    -\int_S \left[\left(\llbracket\Psi\rrbracket,\Lambda^*\right)_{\Gamma} +
    \left(\llbracket\Psi^*\rrbracket,\Lambda\right)_{\Gamma}\right] d\hat{\Omega}
    \shortintertext{where}
    \begin{align}
      \llbracket\Psi\rrbracket &= \Psi^+ - \Psi^-, 
      \Psi^\pm = \lim_{s\rightarrow 0^+} \Psi(\vec{x}+s\hat{n}), \nonumber \\
      \Lambda &= \mbox{ Lagrange multiplier},
      \Gamma = \mbox{ interface boundary}, \nonumber \\
      \hat{n} &= \mbox{ normal vector at interface boundary}, \nonumber
    \end{align}
  \end{gather}
  or the upwinding method:
  \begin{gather}
    \int_S \left(|\hat{\Omega}\cdot\hat{n}|\llbracket\Psi^*\rrbracket,\Psi^-\right)_{\Gamma^+}
    d\hat{\Omega} -
    \int_S\left(|\hat{\Omega}\cdot\hat{n}|\llbracket\Psi^*\rrbracket,\Psi^+\right)_{\Gamma^-}
    d\hat{\Omega}
  \end{gather}
\end{frame}

\begin{frame}
  \frametitle{Hybrid $S_N$-Diffusion Method: Literature Review}
  \textbf{Hybrid Transport-Diffusion Method}
  \vspace{.2cm}

  Anistratov \& Stehle developed an adaptive domain decomposition scheme using Eddington tensor
  estimates as a metric for deducing whether a region is diffusive or
  transport-like \cite{anistratov_computational_2012, stehle_hybrid_2014}.
  \vspace{.2cm}

  It uses two-level neutronics method based on the Second-Moment method. The low-order
  second-moment equations at the $(s+1)$th iteration are:
  \begin{gather}
    \nabla\cdot J^{(s+1)}+\Sigma_a\phi^{(s+1)} = q, \label{eq:2nd-moment-1} \\
    \frac{1}{3}\nabla\cdot\phi^{(s+1)}+\Sigma_t J^{(s+1)} = \nabla\cdot F^{(s+1/2)}
    \label{eq:2nd-moment-2}
    \shortintertext{where}
    F^{(s+1/2)}_{\alpha\beta} = \int_{4\pi}
    \left(\frac{1}{3}\delta_{\alpha\beta}-\Omega_\alpha\Omega_\beta\right)
    \Psi^{(s+1/2)} d\Omega \label{eq:2nd-moment-3}
  \end{gather}
  %
  The transport subdomains require interface conditions in the form of boundary sources from
  neighboring diffusion subdomains.
\end{frame}

\begin{frame}
  \frametitle{Hybrid $S_N$-Diffusion Method: Literature Review}
  \begin{block}{\textbf{Summary}}
    \begin{itemize}
      \item Corrective terms for the neutron diffusion equations in various forms have been
        effectively applied across various methods.
      \item Domain decomposition is necessary for limiting computational cost in large 3-D reactor
        simulations.
      \item Interface conditions are required when applying domain decomposition to iteratively
        couple the transport and diffusion solvers.
%      \item Demonstrations of the Ronen method are limited to 1-D geometries due to the difficulty of
%        deriving transport operators for complex 2-D and 3-D geometries
    \end{itemize}
  \end{block}
\end{frame}

### 2024-12-04 presentation/theory.tex
Presentation commit

\begin{frame}
  \frametitle{Hybrid $S_N$-Diffusion Method: Theory}
  \textbf{Hybrid $\bm{S_N}$-Diffusion Method}
  \begin{block}{\textbf{Aim}}
    To enable accurate control rod modeling in time-dependent molten salt reactor simulations
  \end{block}
  \begin{block}{\textbf{Method Overview}}
    \begin{itemize}
      \item Applies the discrete ordinates ($S_N$) method to a small subregion around a
        control rod to generate corrections for the neutron diffusion equation
      \item Limits computationally expensive $S_N$ calculations to small subdomains
      \item Retains the computational efficiency of the neutron diffusion method
    \end{itemize}
  \end{block}
\end{frame}

\begin{frame}
  \frametitle{Hybrid $S_N$-Diffusion Method: Theory}
  \textbf{Multigroup Discrete Ordinates $\bm{S_N}$ Neutron Transport Equations}
  \vspace{.2cm}

  The multigroup $S_N$ equations defined on the 3-D spatial domain 
  $\mathcal{D}$ and 2-D unit sphere angular domain $\mathcal{S}$ is:
  \begin{multline}
    \hat{\Omega}\cdot\nabla\Psi_g(\vec{r},\hat{\Omega},t) + \Sigma_{t,g}
    \Psi_g(\vec{r},\hat{\Omega},t) =
    \sum^G_{g'=1}\int_\mathcal{S} \Sigma_s^{g'\rightarrow g}(\hat{\Omega}'\rightarrow\hat{\Omega})
    \Psi_{g'}(\vec{r},\hat{\Omega}',t)d\hat{\Omega}' \\
    + \frac{1}{4\pi}\frac{\chi_{p,g}(1-\beta)}{k}\sum^G_{g'=1} \nu\Sigma_{f,g'} \phi_{g'}(\vec{r},t)
    + \frac{1}{4\pi}\sum^I_{i=1}\chi_{d,g}
    \lambda_i C_i(\vec{r},t)
    \label{eq:mg-nte}
  \end{multline}
  with boundary conditions:
  %
  \begin{gather}
    \Psi_g(\vec{r},\hat{\Omega}) = \Psi^\text{inc}_g(\vec{r},\hat{\Omega}) +
    \alpha^s_g\Psi_g(\vec{r},\hat{\Omega}_r)
    \mbox{ on } \vec{r} \in \partial\mathcal{D} \mbox{ and } \hat{\Omega}\cdot\hat{n}_b < 0,
    \shortintertext{where}
    \begin{align*}
%      \chi_{p,g} &= \mbox{prompt fission neutron spectrum in group $g$,} \\
%      \beta &= \sum^I_{i=1} \beta_i = \mbox{total delayed neutron fraction,} \\
%      \chi_{d,g} &= \mbox{delayed fission neutron spectrum in group $g$,} \\
%      \lambda_i &= \mbox{decay constant of precursor group $i$,} \\
%      C_i &= \mbox{delayed neutron precursor concentration for group $i$,} \\
      \Psi^\text{inc}_g &= \mbox{incident surface source in group $g$,} \\
      \alpha^s_g &= \mbox{specular reflectivity on }\partial \mathcal{D} \mbox{ for group }g, \\
      \hat{\Omega}_r &= \hat{\Omega}-2(\hat{\Omega}\cdot \hat{n}_b)\hat{n}_b = \mbox{reflection angle}, \\
      \hat{n}_b &= \mbox{outward unit normal vector on the boundary.}
    \end{align*}
  \end{gather}
\end{frame}

\begin{frame}
  \frametitle{Hybrid $S_N$-Diffusion Method: Theory}
  \textbf{Self-Adjoint Angular Flux (SAAF) Formulation of the Multigroup $\bm{S_N}$ Equations}
  \vspace{.2cm}

  Second-order linear neutron transport equation derived by obtaining the analytical solution of
  angular flux and substituting it back into the $S_N$ equations.

  \begin{block}{\textbf{Implementation Details}}
    \begin{itemize}
      \item Expressed in weak form for the continuous finite element method (CFEM)
      \item Compatible with the efficient and scalable Hypre-BoomerAMG preconditioning
      \item Modified formulation to handle $1/\Sigma_{t,g}$ factor in void regions (similar to
        Streamline-Upwind/Petrov Galerkin (SUPG) stabilization scheme) \cite{wang_diffusion_2014}
      \item Level-symmetric quadrature set for angular discretization (up to $S_{18}$)
      \item Nonlinear diffusion acceleration scheme \cite{wang_diffusion_2014}
    \end{itemize}
  \end{block}
\end{frame}

\begin{frame}
  \frametitle{Hybrid $S_N$-Diffusion Method: Theory}
  \textbf{Weak Formulation of the Multigroup SAAF $S_N$ Equations}
  \begin{gather}
    \shortintertext{Streaming term:}
    \sum^G_{g=1}\sum^{N_d}_{d=1}w_d\left(\hat{\Omega}_d\cdot\nabla\Psi^*_{g,d},\tau_g\hat{\Omega}
    \cdot\nabla\Psi_{g,d}-(1-\tau_g\Sigma_{t,g})\Psi_{g,d}\right)_\mathcal{D} \\
    \shortintertext{Collision term: }
    \sum^G_{g=1}\sum^{N_d}_{d=1}w_d\left(\Psi^*_{g,d},\Sigma_{t,g}\Psi_{g,d}\right)_\mathcal{D} \\
    \shortintertext{Scattering term:}
    \sum^G_{g=1}\sum^{N_d}_{d=1}w_d\left(\Psi^*_{g,d}+\tau_g\hat{\Omega}_d\cdot\nabla\Psi^*_{g,d},
    \sum^G_{g'=1}\sum^L_{l=0}\Sigma^{g'\rightarrow g}_{s,l}\sum^l_{m=-l}
    \frac{2l+1}{w}Y_{l,m}(\hat{\Omega}_d)\phi_{g',l,m}\right)_\mathcal{D} \\
    \shortintertext{Fission source term: }
    \sum^G_{g=1}\sum^{N_d}_{d=1}w_d\left(\Psi^*_{g,d}+\tau_g\hat{\Omega}_d\cdot\nabla\Psi^*_{g,d},
    \frac{1}{w}\frac{\chi_{p,g}(1-\beta)}{k}\sum^G_{g'=1}\nu\Sigma_{f,g'}\phi_{g'}\right)_\mathcal{D}
  \end{gather}
\end{frame}

\begin{frame}
  \frametitle{Hybrid $S_N$-Diffusion Method: Theory}
  \textbf{Weak Formulation of the Multigroup SAAF $S_N$ Equations}
  \begin{gather}
    \shortintertext{Delayed neutron source term:}
    \sum^G_{g=1}\sum^{N_d}_{d=1}w_d\left(\Psi^*_{g,d}+\tau_g\hat{\Omega}_d\cdot\nabla\Psi^*_{g,d},
    \frac{1}{w}\sum ^I_{i=1}\chi_{d,g}\lambda_i C_i\right)_\mathcal{D}
    \shortintertext{Boundary source term:}
    \begin{cases}
      \sum^G_{g=1}\sum^{N_d}_{d=1}w_d\left(\Psi^*_{g,d},
      \hat{\Omega}_d\cdot\hat{n}_b\Psi_{g,d}\right)_{\partial\mathcal{D}},
      & \hat{\Omega}\cdot\hat{n}_b>0,\vec{r}\in\partial\mathcal{D} \\
      \sum^G_{g=1}\sum^{N_d}_{d=1}w_d\left(\Psi^*_{g,d},
      \hat{\Omega}_d\cdot\hat{n}_b\Psi^\text{inc}_{g,d}\right)_{\partial\mathcal{D}},
      & \hat{\Omega}\cdot\hat{n}_b<0,\vec{r}\in\partial\mathcal{D}
    \end{cases} \label{eq:boundary-source} \\
    \shortintertext{Reflecting boundary term:}
    \begin{cases}
      \sum^G_{g=1}\sum^{N_d}_{d=1}w_d\left(\Psi^*_{g,d},
      \hat{\Omega}_d\cdot\hat{n}_b\Psi_{g,d}\right)_{\partial\mathcal{D}},
      & \hat{\Omega}\cdot\hat{n}_b>0,\vec{r}\in\partial\mathcal{D}_s \\
      \sum^G_{g=1}\sum^{N_d}_{d=1}w_d\left(\Psi^*_{g,d},
      \hat{\Omega}_d\cdot\hat{n}_b\Psi_{g,d_r}\right)_{\partial\mathcal{D}},
      & \hat{\Omega}\cdot\hat{n}_b<0,\vec{r}\in\partial\mathcal{D}_s
    \end{cases} \label{eq:reflecting-bc}
  \end{gather}
\end{frame}

\begin{frame}
  \frametitle{Hybrid $S_N$-Diffusion Method: Theory}
  \textbf{Weak Formulation of the Multigroup SAAF $S_N$ Equations}
  \begin{gather}
    \shortintertext{Void stabilization parameter \cite{wang_diffusion_2014}:}
    \tau_g =
    \begin{cases}
      \frac{1}{c\Sigma_{t,g}} \mbox{ for } ch\Sigma_{t,g} \geq \varsigma \\
      \frac{h}{\varsigma} \mbox{ for } ch\Sigma_{t,g} < \varsigma
    \end{cases}, \\
    \shortintertext{where}
    \begin{align*}
      h &= \mbox{mesh element size,} & \\
      c &= \mbox{maximum stabilization factor,} & \\
      \varsigma &= \mbox{void constant.} &
    \end{align*}
  \end{gather}
  $c=1$ and $\varsigma=0.5$ by default
\end{frame}

\begin{frame}
  \frametitle{Hybrid $S_N$-Diffusion Method: Theory}
  \textbf{Multigroup Neutron Diffusion Equations}
  \begin{gather}
    - \nabla \cdot D_g \nabla \phi_g + \Sigma^r_g \phi_g =
    \sum^G_{g' \neq g} \Sigma^s_{g' \rightarrow g} \phi_{g'}
    + \frac{\chi_{p,g} \left( 1-\beta \right)}{k} \sum^G_{g'=1} \nu \Sigma^f_{g'}
    \phi_{g'} + \chi^d_g \sum^I_i \lambda_i C_i \label{eq:neutron} %\\
  \end{gather}
  %
  Traditionally, $D_g$ is determined through region-wide neutron interaction tallies in
  high-fidelity neutron transport simulations as follows:
  \begin{align}
    D_g =& \frac{1}{3\Sigma_{t,g}} \quad \mbox{(isotropic)} \\
    D_g =& \frac{1}{3\Sigma_{tr,g}} = \frac{1}{3\left(\Sigma_{t,g}-
    \Sigma_{s1,g}\right)}
    \quad \mbox{(linearly anisotropic)} \label{eq:p1-diffcoef}
    \shortintertext{where}
    \Sigma_{tr} =& \mbox{ macroscopic transport cross section} \nonumber
  \end{align}
\end{frame}

\begin{frame}
  \frametitle{Hybrid $S_N$-Diffusion Method: Theory}
  \textbf{Diffusion Correction Scheme (Implemented in Python)}
  \vspace{.2cm}

  In this work, I investigated two transport correction schemes. The first scheme is the diffusion
  correction scheme inspired by Fick's first law of diffusion:
  \begin{gather}
    D^s_g(x) = -J^{tr}_g(x)\bigg/\frac{d\phi^{tr}_g(x)}{dx} \label{eq:svdc}
  \end{gather}
  %
  where $D^s_g$ is the diffusion correction parameter, and the $tr$ superscript denotes the
  transport-derived neutron current and scalar flux solutions from the $S_N$ method.
  \vspace{.2cm}

  $D^s_g$ provides pointwise corrections to closely match the diffusion flux solution to the $S_N$
  flux solution.
  By replacing $D_g$ with $D^s_g$, we are effectively adding the following transport correction term
  %
  \begin{gather}
    -\frac{\partial}{\partial x}(D^s_g-D_g)\frac{\partial\phi_g}{\partial x}
  \end{gather}
  to the neutron diffusion equations.
\end{frame}

\begin{frame}
  \frametitle{Hybrid $S_N$-Diffusion Method: Theory}
  \textbf{Drift Correction Scheme (Implemented on Moltres)}
  \vspace{.2cm}

  The second scheme is the drift correction scheme arising from adding a drift term to the neutron
  diffusion equations:
  \begin{gather}
  \scalebox{0.95}{$
    \vec{D}_g = \frac{\sum^{N_d}_{d=1}w_d\left(\tau_g\hat{\Omega}_d\hat{\Omega}_d\cdot\nabla\Psi_{g,d}
    + \left(\tau_g\Sigma_{t,g}-1\right)\hat{\Omega}_d\Psi_{g,d}
    - \tau_g\sum^G_{g'=1}\Sigma^{g'\rightarrow g}_{s,1}\hat{\Omega}_d\Psi_{g',d}
- D_g\nabla\Psi_{g,d}\right)}{\sum^{N_d}_{d=1}w_d\Psi_{g,d}},$} \label{eq:drift} \\
  \scalebox{0.95}{$
    \gamma_g =
    \frac{\sum_{\hat{\Omega}_d\cdot\hat{n}_b > 0}w_d |\hat{\Omega}_d\cdot\hat{n}_b |
  \Psi_{g,d}}{\sum^{N_d}_{d=1}w_d\Psi_{g,d}}.$} \label{eq:bound-coef}
  \end{gather}
  \vspace{.2cm}

  The drift term also provides pointwise corrections to the neutron diffusion equations. This
  formulation is derived from the SAAF-$S_N$ equations by integrating over the angular domain.
\end{frame}

\begin{frame}
  \frametitle{Hybrid $S_N$-Diffusion Method: Theory}
  \begin{columns}
    \column{.35\textwidth}
    \textbf{Hybrid $S_N$-Diffusion Method Algorithm}
    \vspace{.5cm}

  {\small
  $V_0$: Full problem domain
  \vspace{.1cm}

  $V_1$: Problem subdomain containing control rod region
  \vspace{.1cm}

  $D^{s,(i)}_g$: diffusion correction parameter in the $i$-th iteration
  \vspace{.1cm}

  $\vec{D}^{(i)}_g$: drift correction parameter in the $i$-th iteration
\vspace{4cm}}
  \column{.65\textwidth}
  \begin{figure}
    \centering
    \includegraphics[width=.45\textwidth]{images/algorithm}
    \begin{minipage}[b]{.49\textwidth}
      \caption{Algorithm flowchart for the hybrid $S_N$-diffusion method.}
    \end{minipage}
  \end{figure}
\end{columns}
\end{frame}

\begin{frame}
  \frametitle{Hybrid $S_N$-Diffusion Method: Theory}
  \textbf{$S_N$ Subsolver Boundary Conditions}
  \vspace{.2cm}

  For the hybrid $S_N$-diffusion method to converge, it requires appropriate boundary conditions for
  the $S_N$ subproblem.
  We may obtain
  the best estimate by applying the $P_1$ approximation for evaluating the neutron angular flux along
  the discrete ordinates $\hat{\Omega}_d$ of the $S_N$ angular quadrature set as follows:
  \begin{align}
    \Psi_{g,d} &\approx \frac{1}{4\pi}\left(\phi^\text{diff}_g+3\hat{\Omega}_d\cdot
    \vec{J}^\text{diff}_g\right) \nonumber \\
    &=\frac{1}{4\pi}\left(\phi^\text{diff}_g-3\hat{\Omega}_d\cdot D_g\nabla\phi^\text{diff}_g\right)
  \end{align}
  Therefore, the boundary source term for the $S_N$ subsolver is:
  \begin{gather}
    \Psi^\text{inc}_{g,d} = \frac{1}{w}
    \left(\phi^\text{diff}_g-3\hat{\Omega}_d\cdot D_g\nabla\phi^\text{diff}_g\right)
  \end{gather}
\end{frame}

\begin{frame}
  \frametitle{Hybrid $S_N$-Diffusion Method: Theory}
  \textbf{Correction Region ($\bm{V_1}$) and Buffer Zone}
  \begin{itemize}
    \item Recall that the full problem domain and the correction region are defined as $V_0$
      and $V_1$, where $V_1 \subseteq V_0$
    \item The approximate $S_N$ boundary conditions will generally yield some deviations in the
      flux distribution since neutron fluxes in realistic reactor systems are at least slightly
      anisotropic throughout most of the system
    \item However, the influence of boundary conditions on the ratio of $J$ and $\frac{d\phi}{dx}$
      does not extend far from the boundary $\partial V_1$ in optically thick media; transport
      correction parameters are accurate everywhere except near $\partial V_1$
    \item There exists a buffer zone near $\partial V_1$ where transport correction parameters
      are inaccurate
  \end{itemize}
%  \pause
%  $\Rightarrow$ Approach for preliminary results: Define $V_1$ such that it is large enough to
%  provide sufficient transport
%  corrections and accommodate inaccurate SVDCs near $\partial V_1$. Discard inaccurate SVDCs
%  in favor of the default $P_1$-based diffusion coefficients.
\end{frame}

\begin{frame}
  \frametitle{Hybrid $S_N$-Diffusion Method: Theory}
  \textbf{Correction Region ($\bm{V_1}$) and Buffer Zone}
  \begin{figure}[htb!]
    \centering
    \includegraphics[width=.5\columnwidth]{case-3b-geometry}
    \caption{1-D geometry for Case 3b.}
    \label{fig:3b-geometry}
  \end{figure}
  \begin{figure}[htb!]
      \centering
      \begin{subfigure}[t]{.49\textwidth}
          \centering
          \includegraphics[width=.9\textwidth]{case-3b-group-1-drift}
      \end{subfigure}
      \hfill
      \begin{subfigure}[t]{.49\textwidth}
          \centering
          \includegraphics[width=.9\textwidth]{case-3b-group-8-drift}
      \end{subfigure}
      \caption{The reference and hybrid drift ($\vec{D}_g$) distributions for group 1 and 8 calculated
        from $S_8$ and $S_8$-diffusion simulations. The correction subregion $V_1$ spans $x=0$ cm to
        $x=10$ cm.}
      \label{fig:3b-drift-1}
  \end{figure}
\end{frame}

\begin{frame}
  \frametitle{Hybrid $S_N$-Diffusion Method: Theory}
  \textbf{Correction Region ($\bm{V_1}$) and Buffer Zone}
  \vspace{.2cm}

  A natural/intuitive criterion for the location of the buffer zone cutoff boundary
  would be wherever the components of the drift correction variable $\vec{D}_g$ is zero, i.e.,
  wherever the components change signs.
  \begin{enumerate}
    \item At points where the $\vec{D}_g$ components are zero, the flux is approximately isotropic
      along the axes corresponding to the components.
    \item This choice preserves the smoothness of the neutron flux gradient.
  \end{enumerate}
  \begin{figure}[htb!]
      \centering
      \begin{subfigure}[t]{.49\textwidth}
          \centering
          \includegraphics[width=.9\textwidth]{case-3b-group-1-drift}
      \end{subfigure}
      \hfill
      \begin{subfigure}[t]{.49\textwidth}
          \centering
          \includegraphics[width=.9\textwidth]{case-3b-group-8-drift}
      \end{subfigure}
      \caption{The reference and hybrid drift ($\vec{D}_g$) distributions for group 1 and 8 calculated
        from $S_8$ and $S_8$-diffusion simulations. The correction subregion $V_1$ spans $x=0$ cm to
        $x=10$ cm.}
      \label{fig:3b-drift-1}
  \end{figure}
\end{frame}

\begin{frame}
  \frametitle{Hybrid $S_N$-Diffusion Method: Theory}
  \textbf{Numerical Implementation}
  \vspace{.2cm}

  The SAAF-$S_N$ and hybrid $S_N$-diffusion method with the drift correction scheme were
  implemented on Moltres.
  \begin{itemize}
    \item Preconditioned Jacobian-free Newton-Krylov (PJFNK) solver \cite{knoll_jacobian-free_2004}
    \item Hypre-BoomerAMG (Algebraic multigrid) preconditioning \cite{hypre_hypre_2022}
    \item MultiApp and Transfers systems from MOOSE for iterative coupling and data transfers
    \item Supporting material and utility C++ classes for loading group constant data and performing
      angular quadrature calculations
  \end{itemize}
  \vspace{.2cm}

  A 1-D hybrid $S_N$-diffusion method with the diffusion correction
  scheme was implemented in Python. (This scheme was abandoned after 1-D analyses due to division by
  zero errors wherever the flux gradients approach zero.)
  \begin{itemize}
    \item 1-D $S_N$ method with diamond differencing scheme
    \item 1-D neutron diffusion method with finite differencing scheme
  \end{itemize}
\end{frame}

### 2024-12-04 presentation/turb-moltres.tex
Presentation commit

\begin{frame}
  \frametitle{Spalart-Allmaras Turbulence Model in Moltres}
  \begin{block}{\textbf{Why the Spalart-Allmaras Model for Moltres?}}
    A computationally efficient one-equation model that provides reasonable estimates for turbulent
    effects in MSRs and is tractable for large problems on small computing clusters.
  \end{block}
  \begin{block}{\textbf{Subobjectives}}
    \textbf{Implementation of a Spalart-Allmaras turbulence model in Moltres}
    \begin{itemize}
      \item Implement a Spalart-Allmaras model in Moltres
      \item Verify and validate the model against reference turbulent flow solutions
    \end{itemize}
  \end{block}
\end{frame}

\begin{frame}
  \frametitle{Spalart-Allmaras Turbulence Model in Moltres}
  \textbf{Spalart-Allmaras Model with Rotation Correction Scheme
  \cite{spalart_one-equation_1994, aupoix_extensions_2003, dacles-mariani_numericalexperimental_1995}}\\
  \\
  Governing equation for the modified dynamic viscosity, $\tilde{\mu}$:
  \begin{multline}
    \rho \frac{\partial\tilde{\mu}}{\partial t} + \rho \mathbf{u}\cdot\nabla\tilde{\mu} = \rho c_{b1}
    \left(1-f_{t2}\right)\tilde{S}\tilde{\mu} + \frac{1}{\sigma}\{\nabla\cdot\left[\left(\mu+
    \tilde{\mu}\right)\nabla\tilde{\mu}\right] + c_{b2}|\nabla\tilde{\mu}|^2\} \\
    - \left(c_{w1}f_w - \frac{c_{b1}}{\kappa^2}f_{t2}\right)\left(\frac{\tilde{\mu}}{d}\right)^2
  \end{multline}
  \begin{align*}
    \shortintertext{where}
      \mu_t &= \tilde{\mu}f_{v1} = \text{turbulent eddy viscosity}, \\
      f_{v1} &= \frac{\chi^3}{\chi^3 + c_{v1}^3}, \\
      \chi &= \frac{\tilde{\mu}}{\mu}.
  \end{align*}
%      \tilde{S} &= \Omega + \frac{\tilde{\nu}}{\kappa^2 d^2} f_{v2}, \\
%      f_{v2} &= 1 - \frac{\chi}{1+\chi f_{v1}}, \\
%      \Omega &= \sqrt{2W_{ij}W_{ij}} = \text{vorticity magnitude}, \\
%      W_{ij} &= \frac{1}{2}\left(\frac{\partial u_i}{\partial x_j} - \frac{\partial u_j}{\partial x_i}
%      \right), \\
%        f_w &= g\left(\frac{1 + c_{w3}^6}{g^6 + c_{w3}^6}\right)^{1/6}, \\
%        g &= r + c_{w2}\left(r^6 - r\right), \\
%        r &= \text{min}\left(\frac{\tilde{\nu}}{\tilde{S}\kappa^2d^2}, 10\right), \\
%        f_{t2} &= c_{t3} \exp{\left(-c_{t4}\chi^2\right)},
%    \end{align*}
%  \shortintertext{and the constants are}
%    \sigma = \frac{2}{3}, \ c_{b1} = 0.1355, \ c_{b2} = 0.622, \ \kappa = 0.41, \
%    c_{w1} = \frac{c_{b1}}{\kappa^2} + \frac{1+c_{b2}}{\sigma}, \nonumber \\
%    c_{w2} = 0.3, \ c_{w3} = 2, \
%    c_{v1} = 7.1, \ c_{t3} = 1.2, \ c_{t4} = 0.5 \ \text{.} \nonumber
\end{frame}

\begin{frame}
  \frametitle{Spalart-Allmaras Turbulence Model in Moltres}
  \begin{block}{\textbf{Model Implementation Details}}
  \begin{itemize}
    \item Built on top of an existing incompressible Navier-Stokes model in MOOSE
    \item Uses \gls{SUPG} stabilization scheme for numerical stability
    \item $f_{t2}$ turbulence trip term is togglable (Negligible difference in fully turbulent flows)
    \item Rotation correction scheme \cite{aupoix_extensions_2003, dacles-mariani_numericalexperimental_1995}
      reduces eddy viscosity in regions of rotational but non-turbulent flow
  \end{itemize}
  \end{block}
  \pause
  \begin{block}{\textbf{Verification \& Validation Tests for the Turbulence Model}}
    \begin{enumerate}
      \item Turbulent Channel Flow Verification Test at Re $\approx 13750$
      \begin{itemize}
        \item Reference results: \gls{DNS} results by Moser et al.\ \cite{moser_direct_1999}
      \end{itemize}
      \item Turbulent Pipe Flow Validation Test at Re $\approx 40000$
      \begin{itemize}
        \item Reference results: Experimental results by Laufer \cite{laufer_structure_1954}
      \end{itemize}
      \item Backward-Facing Step Flow Validation Test
      \begin{itemize}
        \item Reference results: Experimental results by Driver \& Seegmiller \cite{driver_features_1985}
        \item Reference results: Spalart-Allmaras model results from CFL3D CFD code \cite{krist_cfl3d_1998}
      \end{itemize}
    \end{enumerate}
  \end{block}
\end{frame}

\begin{frame}
  \frametitle{Spalart-Allmaras Turbulence Model in Moltres}
  \textbf{Turbulent Channel Flow Verification Test Results}
  \begin{figure}[htb!]
    \centering
    \begin{subfigure}[b]{0.32\columnwidth}
      \centering
      \includegraphics[width=\columnwidth]{channel_vel}
      \caption{Normalized velocity distribution across the channel.}
      \label{fig:channel-vel}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.32\columnwidth}
      \centering
      \includegraphics[width=\columnwidth]{channel_nondim}
      \caption{Dimensionless velocity vs.\ dimensionless wall distance.}
      \label{fig:channel-nondim}
    \end{subfigure}
    \begin{subfigure}[b]{0.32\columnwidth}
      \centering
      \includegraphics[width=\columnwidth]{channel_stress}
      \caption{Normalized stress distribution across the channel.}
      \label{fig:channel-stress}
    \end{subfigure}
    \caption{Comparison of turbulent channel flow results at Re$_\tau\approx395$ against reference
    \gls{DNS} results \cite{moser_direct_1999}. The Moltres Spalart-Allmaras model agrees consistently with
    the reference data.}
    \label{fig:channel-verification}
  \end{figure}
\end{frame}

\begin{frame}
  \frametitle{Spalart-Allmaras Turbulence Model in Moltres}
  \textbf{Turbulent Pipe Flow Validation Test Results}
  \begin{figure}[htb]
    \centering
    \begin{subfigure}[b]{0.32\columnwidth}
      \centering
      \includegraphics[width=\columnwidth]{pipe_vel}
      \caption{Normalized velocity distribution across the pipe.}
      \label{fig:pipe-vel}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.32\columnwidth}
      \centering
      \includegraphics[width=\columnwidth]{pipe_nondim}
      \caption{Dimensionless velocity vs.\ dimensionless wall distance}
      \label{fig:pipe-nondim}
    \end{subfigure}
    \begin{subfigure}[b]{0.32\columnwidth}
      \centering
      \includegraphics[width=\columnwidth]{pipe_stress}
      \caption{Normalized turbulent shear stress across the pipe.}
      \label{fig:pipe-stress}
    \end{subfigure}
    \caption{Comparison of turbulent pipe flow results at Re $\approx 40000$ against reference
    experimental data \cite{laufer_structure_1954}.}
    \label{fig:pipe-verification}
  \end{figure}
\end{frame}

\begin{frame}
  \frametitle{Spalart-Allmaras Turbulence Model in Moltres}
  \textbf{Backward-Facing Step Flow Validation Test Results}
  \begin{columns}
    \column{5.5cm}
    \begin{figure}[h]
      \centering
      \includegraphics[width=\columnwidth]{backstep-geom-pres}
      \caption{Backward step geometry for the turbulent BFS flow verification test. The red box indicates
      the region shown by the close-up view in Figure \ref{fig:bfs-mesh}.}
      \label{fig:backstep-geom}
    \end{figure}
    \column{5.5cm}
    \begin{figure}[h]
      \centering
      \includegraphics[width=\columnwidth]{bfs_mesh}
      \caption{Close-up view of the mesh for the BFS flow test. The step is situated at $x=110$ m
      with a height of $H=1$ m.}
      \label{fig:bfs-mesh}
    \end{figure}
  \end{columns}
\end{frame}

\begin{frame}
  \frametitle{Spalart-Allmaras Turbulence Model in Moltres}
  \textbf{Backward-Facing Step Flow Validation Test Results}
  \begin{columns}
    \column{5.5cm}
    \begin{figure}[htb!]
      \centering
      \includegraphics[width=\columnwidth]{bfs}
      \caption{Velocity magnitude distribution and streamlines around the backward-facing step. The
      streamlines illustrate the primary and secondary recirculation zones.}
      \label{fig:bfs}
    \end{figure}
    \column{5.5cm}
    \begin{table}[htb]
      \centering
      \scriptsize
      \caption{BFS flow reattachment length estimates normalized by step height $H$.}
      \begin{tabular}{l S[table-format=1.2(2)]}
        \toprule
        Source & {Reattachment length [-]} \\
        \midrule
        Experimental data & 6.26(10) \\
        CFL3D & 6.1 \\
        Moltres & 6.36 \\
        \bottomrule
      \end{tabular}
      \label{table:bfs-reattach}
    \end{table}
  \end{columns}
\end{frame}

\begin{frame}
  \frametitle{Spalart-Allmaras Turbulence Model in Moltres}
  \textbf{Backward-Facing Step Flow Validation Test Results}
  \begin{figure}[htb]
    \centering
    \hfill
    \begin{subfigure}[b]{0.24\columnwidth}
      \centering
      \includegraphics[width=\columnwidth]{bfs_downstream_vel}
      \caption{Normalized velocity distributions downstream of step.}
      \label{fig:bfs-downstream}
    \end{subfigure} %\hfill \\
%    \centering
    \hfill
    \begin{subfigure}[b]{0.24\columnwidth}
      \centering
      \includegraphics[width=\columnwidth]{bfs_cf}
      \caption{Skin friction coefficient along the bottom wall.}
      \label{fig:bfs-cf}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.24\columnwidth}
      \centering
      \includegraphics[width=\columnwidth]{bfs_cp}
      \caption{Skin pressure coefficient along the bottom wall.}
      \label{fig:bfs-cp}
    \end{subfigure}% \hfill \\
%    \centering
    \begin{subfigure}[b]{0.24\columnwidth}
      \centering
      \includegraphics[width=\columnwidth]{bfs_stress}
      \caption{Normalized turbulent shear stress downstream
      of step.}
      \label{fig:bfs-stress}
    \end{subfigure}
    \caption{Comparison of backward facing step flow results against reference
    experimental data and computational data from CFL3D. $x/H$ values are normalized horizontal
    distances relative to the step.}
    \label{fig:bfs-plots}
  \end{figure}
\end{frame}

\begin{frame}
  \frametitle{Spalart-Allmaras Turbulence Model in Moltres}
  \textbf{Spalart-Allmaras Model V\&V Test Summary}
  \begin{itemize}
    \item Good agreement between Moltres and reference solutions in the turbulent channel and pipe
      flow tests
    \item Good agreement between Moltres and CFL3D in the backward-facing step flow test
    \item Some discrepancy between Moltres and experimental data in the backward-facing step flow
      test, but Moltres performed better than CFL3D at predicting flow reattachment length and
      velocity/stress distributions along $x/H=1$.
  \end{itemize}
\end{frame}

### 2024-12-05 presentation/3d-transient.tex
slide edits

@@
 \begin{frame}
-  \frametitle{Time-Dependent Multiphysics Simulations with the Hybrid Method}
-  I investigated two transient scenarios based on actual MSRE experiments:
+  \frametitle{Time-Dependent Simulations with the Hybrid Method}
+%  I investigated two transient scenarios based on actual MSRE experiments:
+  Time-dependent reactivity-initiated simulation based on an MSRE rod drop experiment.
   \begin{block}{\textbf{MSRE Rod Drop Experiment}}
@@
       \item Requires delayed neutron precursor (DNP) modeling
+      \item Induces a prompt response, followed by a delayed response, in the neutron count rate
     \end{itemize}
   \end{block}
-  \begin{block}{\textbf{MSRE Reactivity Insertion Experiment}}
-    \begin{itemize}
-      \item Coupled power response of the MSRE initially at 1 MW to a rod
-        withdrawal \cite{engel_zero-power_1972}
-      \item Reactivity insertion of 24.8 pcm
-      \item MSRE fueled with $^{233}$U
-      \item Requires DNP and temperature advection-diffusion modeling
-    \end{itemize}
-  \end{block}
+%  \begin{block}{\textbf{MSRE Reactivity Insertion Experiment}}
+%    \begin{itemize}
+%      \item Coupled power response of the MSRE initially at 1 MW to a rod
+%        withdrawal \cite{engel_zero-power_1972}
+%      \item Reactivity insertion of 24.8 pcm
+%      \item MSRE fueled with $^{233}$U
+%      \item Requires DNP and temperature advection-diffusion modeling
+%    \end{itemize}
+%  \end{block}
 \end{frame}
@@
   \frametitle{MSRE Rod Drop Experiment}
-  \textbf{Integral Neutron Count Following Rod Drop}
+  \textbf{Neutron Count Rate Following Rod Drop}
+  \begin{columns}
+    \column{5.5cm}
     \begin{figure}[t]
       \centering
-    \includegraphics[width=0.55\columnwidth]{integral-count}
+      \includegraphics[width=\columnwidth]{count-rate}
+      \caption{Neutron count rate during the rod drop experiment from Moltres rod drop simulation.}
+      \label{fig:count-rate}
+    \end{figure}
+    \column{5.5cm}
+    \begin{itemize}
+      \item Steep initial decline in neutron count rate as the rod drops
+      \item Decline in neutron count rate slows at $t=0.5$ s due to presence of DNPs
+      \item Convergence issues prevented the simulation from converging from $t = 0.8$ s
+      \item I raised the convergence tolerance to help the simulation continue
+    \end{itemize}
+  \end{columns}
+%  \begin{itemize}
+%    \item Moltres underpredicted the integral neutron count rate by approximately 20 \%.
+%    \item Potential sources of error
+%    \begin{itemize}
+%      \item Uncertainty in the initial neutron count rate
+%      \item Miscalculation of -1500 pcm reactivity withdrawal
+%    \end{itemize}
+%  \end{itemize}
+\end{frame}
+
+\begin{frame}
+  \frametitle{MSRE Rod Drop Experiment}
+  \textbf{Integral Neutron Count Following Rod Drop}
+  \begin{columns}
+    \column{5.5cm}
+    \begin{figure}[t]
+      \centering
+      \includegraphics[width=\columnwidth]{integral-count}
       \caption{Integral neutron count during the rod drop experiment from \gls{MSRE} experimental data
@@
     \end{figure}
+    \column{5.5cm}
     \begin{itemize}
-    \item Moltres underpredicted the integral neutron count rate by approximately 20 \%.
-    \item Potential sources of error
-    \begin{itemize}
-      \item Uncertainty in the initial neutron count rate
-      \item Miscalculation of -1500 pcm reactivity withdrawal
+      \item Moltres reproduces the expected trend in the integral neutron count rate
+      \item Slight underprediction relative to MSRE rod drop experimental data
+      \item Underprediction may be due to experimental uncertainty in $^{235}$U concentration,
+        initial \& final rod height, initial neutron count rate.
     \end{itemize}
-  \end{itemize}
+  \end{columns}
+%  \begin{itemize}
+%    \item Moltres underpredicted the integral neutron count rate by approximately 20 \%.
+%    \item Potential sources of error
+%    \begin{itemize}
+%      \item Uncertainty in the initial neutron count rate
+%      \item Miscalculation of -1500 pcm reactivity withdrawal
+%    \end{itemize}
+%  \end{itemize}
 \end{frame}
 
+%\begin{frame}
+%  \frametitle{MSRE Reactivity Insertion Experiment}
+%  \textbf{Nested Coupling Iteration Structure for the Reactivity Insertion Simulation}
+%  \begin{figure}[t]
+%    \centering
+%    \includegraphics[width=\columnwidth]{images/nest-2}
+%    \caption{The nested iteration structure coupling the $S_N$, neutron diffusion,
+%    in-core \gls{DNP}-temperature, and ex-core
+%    solvers for the reactivity insertion simulation using the hybrid $S_N$-diffusion method.}
+%    \label{fig:insertion-coupling}
+%  \end{figure}
+%\end{frame}
+
 \begin{frame}
-  \frametitle{MSRE Reactivity Insertion Experiment}
-  \textbf{Nested Coupling Iteration Structure for the Reactivity Insertion Simulation}
-  \begin{figure}[t]
-    \centering
-    \includegraphics[width=\columnwidth]{images/nest-2}
-    \caption{The nested iteration structure coupling the $S_N$, neutron diffusion,
-    in-core \gls{DNP}-temperature, and ex-core
-    solvers for the reactivity insertion simulation using the hybrid $S_N$-diffusion method.}
-    \label{fig:insertion-coupling}
-  \end{figure}
+  \frametitle{3-D Modeling with the Hybrid Method}
+  \begin{block}{\textbf{Difficulties Faced During 3-D Modeling}}
+  \begin{itemize}
+    \item Poor convergence rate relative to 1-D \& 2-D modeling
+    \begin{itemize}
+      \item Likely due to increased streaming effects in 3-D near-void (air) regions in the reactor
+    \end{itemize}
+    \item Significant memory requirements
+    \item Lagged control rod positions in fixed point iterations affecting convergence in time-dependent
+      simulations $\Rightarrow$ simulation requires smaller timestep sizes
+  \end{itemize}
+\end{block}
 \end{frame}

### 2024-12-05 presentation/control-rod.tex
slide edits

\begin{frame}
  \frametitle{Hybrid $S_N$-Diffusion Method: Motivation}
  \begin{columns}
    \hfill
    \column[t]{5cm}
    \textbf{Control Rods in MSRs}
    \vspace{.2cm}
    \begin{figure}
      \centering
      \includegraphics[width=\columnwidth]{images/msre-cutout}
      \caption{\footnotesize MSRE reactor vessel \cite{robertson_msre_1965}}
    \end{figure}
    \hfill
    \column[t]{6cm}
    \begin{figure}
      \centering
      \includegraphics[width=.8\columnwidth]{images/tmsr}
      \caption{\footnotesize Cross-sectional views of the TMSR-LF1
      \cite{liu_sensitivityuncertainty_2020}}
    \end{figure}
    \begin{figure}
      \centering
      \includegraphics[width=.5\columnwidth]{images/tap-msr-rods}
      \caption{\footnotesize Cross-sectional view of the TAP MSR \cite{lee_neutronics_2020}}
    \end{figure}
    \hfill
  \end{columns}
\end{frame}

\begin{frame}
  \frametitle{Hybrid $S_N$-Diffusion Method: Motivation}
  \textbf{Control rods provide a means of controlling the fission rate in nuclear reactors.} 
  \begin{itemize}
    \item Facilitate reactor start-up, shut-down, or load-following operations
    \item Consist of highly neutron-absorbing materials such as boron or gadolinium
  \end{itemize}
  \textbf{MSRs contain comparatively fewer control rods than most other reactor types due to:}
  \begin{itemize}
    \item Uniform liquid fuel burnup
    \item Strong passive safety of the liquid fuel form
    \item Low excess reactivity of fuel inventory
    \item Availability of other control mechanisms
  \end{itemize}
  \pause
  \textbf{Nevertheless, it is important to characterize control rod effects in all relevant
  transient scenarios.}
\end{frame}

\begin{frame}
  \frametitle{Hybrid $S_N$-Diffusion Method: Motivation}
  \textbf{Control rods induce highly anisotropic neutron fluxes and steep flux gradients in their
  vicinity.}
  \begin{block}{\textbf{The Control Rod Modeling Dilemma}}
    \begin{itemize}
      \item Neutron diffusion, $P_1$, and $SP_N$ methods perform poorly near control rod regions
        due to the highly anisotropic neutron fluxes and steep flux gradients
      \item High-fidelity neutron transport methods remain too computationally expensive for
        routine time-dependent multiphysics simulations
    \end{itemize}
  \end{block}
\end{frame}

\begin{frame}
  \frametitle{Hybrid $S_N$-Diffusion Method: Motivation}
  \textbf{Control Rod Modeling in MSR Multiphysics Studies}
  \vspace{.2cm}

  Common simplifications applied to control rod modeling include:
  \begin{itemize}
    \item Homogenized, coarse-mesh geometries containing static control rods with albedo neutron
      flux boundary conditions \cite{kophazi_development_2009} or transport-corrected cross
      sections \cite{cui_development_2021, jaradat_development_2021, yang_development_2022}
      \begin{itemize}
        \item Neutronics mesh differs from thermal-hydraulics mesh
        \item 1-D thermal-hydraulics model with uniform or prescribed heat deposition
        \item Static calibration factors may not be suitable for time-dependent simulations
      \end{itemize}
    \item Scaling the neutron source term to simulate moving control rods
      \cite{delpech_benchmark_2003, krepel_dyn3d-msr_2007, jaradat_development_2021,
      yang_development_2022}
      \begin{itemize}
        \item Does not capture local or asymmetric flux changes
      \end{itemize}
  \end{itemize}

  \begin{block}{\textbf{Technical Gap}}
    \textbf{There are no existing MSR simulation tools or studies involving moving control rods
    which are explicitly modeled.}
  \end{block}
\end{frame}

### 2024-12-05 presentation/future.tex
slide edits

@@
   \begin{block}{\textbf{Future Work}}
-    \begin{itemize}
-      \item Extend the Spalart-Allmaras turbulence model implementation
+    \textbf{Extend the Spalart-Allmaras turbulence model implementation}
       \begin{itemize}
@@
       \end{itemize}
-      \item Improve the computational performance of the hybrid $S_N$-diffusion method
+    \textbf{Improve the computational performance of the hybrid $S_N$-diffusion method}
       \begin{itemize}
@@
         \item Custom mesh distribution and data transfer caching routines to minimize data transfer time
+        \item Preconditioning method or other optimizations to improve convergence of modified neutron diffusion
+          method
       \end{itemize}
-      \item Further 3-D full-core reactor analyses with the hybrid method
+    \textbf{Further 3-D full-core reactor analyses with the hybrid method}
       \begin{itemize}
@@
       \end{itemize}
-    \end{itemize}
   \end{block}

### 2024-12-05 presentation/msre-transients.tex
slide edits

@@
   \begin{itemize}
-    \item Pump start-up and coast-down transient flow-rate tests on the MSRE under zero-power
-  conditions
     \item Starting from zero-power critical states with/without forced flow, the fuel salt pump
@@
   \end{itemize}
+  \begin{block}{\textbf{MSR Phenomena Involved}}
+    \begin{itemize}
+      \item DNP drift under time-varying flow
+      \item Loss of delayed neutrons due to out-of-core DNP decay
+    \end{itemize}
+  \end{block}
   \begin{columns}
-    \column[t]{5.5cm}
+    \column{5.5cm}
     \begin{figure}
@@
     \end{figure}
-    \column[t]{5.5cm}
+    \column{5.5cm}
     \begin{figure}
@@
     \column{6.5cm}
-    \begin{block}{\textbf{MSR Phenomena Involved}}
-      \begin{itemize}
-        \item DNP drift under time-varying flow
-        \item Loss of delayed neutrons due to out-of-core DNP decay
-      \end{itemize}
-    \end{block}
     \begin{block}{\textbf{Aim of the Study}}
       \begin{itemize}
-         \item Develop a V\&V study based on the MSRE pump start-up \& coast-down
-           transients that is easily reproducible
+         \item Develop a V\&V study based on the MSRE pump transients
          \item Verify Moltres against QuasiMolto in collaboration with Aaron Reynolds
@@
         \item Two neutron energy groups and six DNP groups
-        \item Group constants generated using a representative unit cell model in SCALE
         \item Ignored upper and lower plena in the geometry
         \item Looped salt flow with a salt circulation time of 25.2 s
-        \item $k$-eigenvalue calculation at every time-step in place of reactivity control
-        \item Timestep size of 0.1 s
+        \item $k$-eigenvalue calculation at every time-step ($dt=0.1$ s) in place of reactivity control
       \end{itemize}
@@
   \frametitle{V\&V Study 2: MSRE Pump Start-up \& Coast-Down Transients}
-  \textbf{Static \& Steady-Flow Conditions}
+  \textbf{$k_\text{eff}$ Under Static \& Steady-Flow Conditions}
   \begin{table}[htb]
@@
   \end{table}
-
   \begin{itemize}
@@
   \end{figure}
+  \begin{itemize}
+    \item Perfect overlap in centerline and midplane neutron fluxes between Moltres and QuasiMolto.
+  \end{itemize}
 \end{frame}
@@
   \end{figure}
+  \begin{itemize}
+    \item Perfect overlap in centerline and midplane DNP distributions between Moltres and QuasiMolto.
+  \end{itemize}
 \end{frame}
@@
   \frametitle{V\&V Study 2: MSRE Pump Start-up \& Coast-Down Transients}
-  \textbf{Study Outcome}
+  \begin{block}{\textbf{Study Outcome}}
     \begin{itemize}
@@
     \end{itemize}
+  \end{block}
 \end{frame}

### 2024-12-05 presentation/review.tex
slide edits

@@
   low-order QD equations \cite{anistratov_solution_1986}.
+  \begin{figure}[t]
+    \tikzstyle{every node}=[font=\small]
+    \centering
+    \begin{tikzpicture}
+      \node (1) [solver2] {Multigroup high-order transport};
+      \node (2) [solver2, right of=1, xshift=3cm] {Multigroup low-order QD};
+      \node (3) [solver2, right of=2, xshift=3cm] {Effective low-order transport};
+      \draw [arrow] (1.10) -- node[anchor=south, text width=2cm, align=center] {Group\\ Eddington
+        and boundary factors} (2.170);
+      \draw [arrow] (2.190) -- node[anchor=north, text width=2cm, align=center] {Effective flux
+        and precursors} (1.350);
+      \draw [arrow] (2.10) -- node[anchor=south, text width=2cm, align=center] {Collapsed Eddington
+        and boundary factors} (3.170);
+      \draw [arrow] (3.190) -- node[anchor=north, text width=2cm, align=center] {Effective flux
+        and precursors} (2.350);
+    \end{tikzpicture}
+    \caption{Algorithm flowchart for the three-level quasi-diffusion method.}
+    \label{fig:algorithm}
+  \end{figure}
 \end{frame}
@@
 
-  It uses two-level neutronics method based on the Second-Moment method. The low-order
-  second-moment equations at the $(s+1)$th iteration are:
+  It uses two-level neutronics method based on the Second-Moment method \cite{lewis_comparison_1976}.
+  The low-order second-moment equations on the $(s+1)$th iteration are:
   \begin{gather}

### 2024-12-05 presentation/theory.tex
slide edits

@@
   \textbf{Hybrid $\bm{S_N}$-Diffusion Method}
-  \begin{block}{\textbf{Aim}}
-    To enable accurate control rod modeling in time-dependent molten salt reactor simulations
-  \end{block}
+%  \begin{block}{\textbf{Aim}}
+%    To enable accurate control rod modeling in time-dependent molten salt reactor simulations
+%  \end{block}
   \begin{block}{\textbf{Method Overview}}
@@
       \item Retains the computational efficiency of the neutron diffusion method
+      \item Applies an adaptive algorithm to smooth flux gradients near the $S_N$-diffusion interface
     \end{itemize}
   \end{block}
+  \begin{figure}
+    \centering
+    \includegraphics[width=.7\textwidth]{hybrid-illustration}
+    \caption{Illustration of the problem domains of the $S_N$ and neutron diffusion methods in an
+    example 1-D problem.}
+  \end{figure}
 \end{frame}
@@
     \begin{itemize}
-      \item Expressed in weak form for the continuous finite element method (CFEM)
+      \item Implemented with finite element method (FEM)
       \item Compatible with the efficient and scalable Hypre-BoomerAMG preconditioning
-      \item Modified formulation to handle $1/\Sigma_{t,g}$ factor in void regions (similar to
+      \item Uses a modified formulation to handle $1/\Sigma_{t,g}$ factor in void regions (similar to
         Streamline-Upwind/Petrov Galerkin (SUPG) stabilization scheme) \cite{wang_diffusion_2014}
@@
   \end{gather}
-  $c=1$ and $\varsigma=0.5$ by default
+  $c=1$ and $\varsigma=0.5$ by default.
+  \vspace{.2cm}
+
+  The SAAF-$S_N$ equations require this stabilization scheme
+  in near-void regions where $\Sigma_{t,g}$ is very small.
 \end{frame}
@@
   In this work, I investigated two transport correction schemes. The first scheme is the diffusion
-  correction scheme inspired by Fick's first law of diffusion:
+  correction scheme similar to Fick's first law of diffusion:
   \begin{gather}
@@
 
-  $D^s_g$ provides pointwise corrections to closely match the diffusion flux solution to the $S_N$
-  flux solution.
+%  $D^s_g$ provides pointwise corrections to closely match the diffusion flux solution to the $S_N$
+%  flux solution.
   By replacing $D_g$ with $D^s_g$, we are effectively adding the following transport correction term
@@
   \begin{gather}
-    -\frac{\partial}{\partial x}(D^s_g-D_g)\frac{\partial\phi_g}{\partial x}
+    -\nabla\cdot (D^s_g-D_g)\nabla\phi_g
   \end{gather}
   to the neutron diffusion equations.
+  \vspace{.2cm}
+
+  $\Rightarrow$ Abandoned after 1-D investigations due to division-by-zero issues near flux maximas
+  and minimas.
 \end{frame}
@@
 
-  The second scheme is the drift correction scheme arising from adding a drift term to the neutron
-  diffusion equations:
+  The second scheme is the drift correction scheme arising from adding a drift term
+  ($\vec{D}_g\cdot\nabla \phi_g$) to the neutron diffusion equations:
   \begin{gather}
@@
   The drift term also provides pointwise corrections to the neutron diffusion equations. This
-  formulation is derived from the SAAF-$S_N$ equations by integrating over the angular domain.
+  formulation is derived from the SAAF-$S_N$ equations by integrating over the angular domain and
+  eliminating terms shared by the neutron diffusion equations.
+\end{frame}
+
+\begin{frame}
+  \frametitle{Hybrid $S_N$-Diffusion Method: Theory}
+  \textbf{$S_N$ Subsolver Boundary Conditions}
+  \vspace{.2cm}
+
+  For the hybrid $S_N$-diffusion method to converge, it requires appropriate boundary conditions for
+  the $S_N$ subproblem.
+  We may obtain
+  the best estimate by applying the $P_1$ approximation for evaluating the neutron angular flux along
+  the discrete ordinates $\hat{\Omega}_d$ of the $S_N$ angular quadrature set as follows:
+  \begin{align}
+    \Psi_{g,d} &\approx \frac{1}{4\pi}\left(\phi^\text{diff}_g+3\hat{\Omega}_d\cdot
+    \vec{J}^\text{diff}_g\right) \nonumber \\
+    &=\frac{1}{4\pi}\left(\phi^\text{diff}_g-3\hat{\Omega}_d\cdot D_g\nabla\phi^\text{diff}_g\right)
+  \end{align}
+  Therefore, the boundary source term for the $S_N$ subsolver is:
+  \begin{gather}
+    \Psi^\text{inc}_{g,d} = \frac{1}{w}
+    \left(\phi^\text{diff}_g-3\hat{\Omega}_d\cdot D_g\nabla\phi^\text{diff}_g\right)
+  \end{gather}
+  where $w$ is the sum of weights of the level-symmetric quadrature set.
 \end{frame}
@@
     \centering
-    \includegraphics[width=.45\textwidth]{images/algorithm}
+    \includegraphics[width=.46\textwidth]{images/algorithm}
     \begin{minipage}[b]{.49\textwidth}
@@
 \end{columns}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Hybrid $S_N$-Diffusion Method: Theory}
-  \textbf{$S_N$ Subsolver Boundary Conditions}
-  \vspace{.2cm}
-
-  For the hybrid $S_N$-diffusion method to converge, it requires appropriate boundary conditions for
-  the $S_N$ subproblem.
-  We may obtain
-  the best estimate by applying the $P_1$ approximation for evaluating the neutron angular flux along
-  the discrete ordinates $\hat{\Omega}_d$ of the $S_N$ angular quadrature set as follows:
-  \begin{align}
-    \Psi_{g,d} &\approx \frac{1}{4\pi}\left(\phi^\text{diff}_g+3\hat{\Omega}_d\cdot
-    \vec{J}^\text{diff}_g\right) \nonumber \\
-    &=\frac{1}{4\pi}\left(\phi^\text{diff}_g-3\hat{\Omega}_d\cdot D_g\nabla\phi^\text{diff}_g\right)
-  \end{align}
-  Therefore, the boundary source term for the $S_N$ subsolver is:
-  \begin{gather}
-    \Psi^\text{inc}_{g,d} = \frac{1}{w}
-    \left(\phi^\text{diff}_g-3\hat{\Omega}_d\cdot D_g\nabla\phi^\text{diff}_g\right)
-  \end{gather}
 \end{frame}

### 2024-12-05 presentation/turb-moltres.tex
slide edits

@@
   \begin{block}{\textbf{Why the Spalart-Allmaras Model for Moltres?}}
-    A computationally efficient one-equation model that provides reasonable estimates for turbulent
-    effects in MSRs and is tractable for large problems on small computing clusters.
+    A computationally efficient one-equation model that provides reasonable estimates for wall-bounded
+    turbulent flows and is tractable for large problems on small computing clusters.
   \end{block}
@@
     \caption{Comparison of turbulent channel flow results at Re$_\tau\approx395$ against reference
-    \gls{DNS} results \cite{moser_direct_1999}. The Moltres Spalart-Allmaras model agrees consistently with
-    the reference data.}
+    \gls{DNS} results \cite{moser_direct_1999}.}
     \label{fig:channel-verification}
   \end{figure}
+  \begin{itemize}
+    \item The Moltres Spalart-Allmaras model results are consistent with reference DNS channel flow data.
+  \end{itemize}
 \end{frame}
@@
     \label{fig:pipe-verification}
+  \begin{itemize}
+    \item The Moltres Spalart-Allmaras model results are consistent with reference experimental pipe flow data.
+  \end{itemize}
   \end{figure}
@@
     \end{table}
+    \begin{itemize}
+      \item Moltres is consistent with experimental data (within uncertainty range) while CFL3D
+        falls outside the uncertainty range.
+    \end{itemize}
   \end{columns}
@@
   \textbf{Backward-Facing Step Flow Validation Test Results}
+  \begin{columns}
+    \column{11cm}
   \begin{figure}[htb]
@@
     \hfill
-    \begin{subfigure}[b]{0.24\columnwidth}
+    \begin{subfigure}[t]{0.24\columnwidth}
       \centering
@@
 %    \centering
-    \begin{subfigure}[b]{0.24\columnwidth}
+    \begin{subfigure}[t]{0.24\columnwidth}
       \centering
@@
   \end{figure}
+\end{columns}
 \end{frame}
@@
   \frametitle{Spalart-Allmaras Turbulence Model in Moltres}
-  \textbf{Spalart-Allmaras Model V\&V Test Summary}
+  \begin{block}{\textbf{Spalart-Allmaras Model V\&V Test Summary}}
     \begin{itemize}

### 2024-12-09 abstract.tex
Draft for committee before defense

@@
 rod worth estimates, relative to reference neutron transport solutions and experimental data,
-at approximately four times the cost of the neutron diffusion method.
-Demonstrations of the hybrid method for time-dependent rod drop and reactivity insertion
-simulations also perfomed well in reproducing expected trends observed in experimental data.
+at approximately four times the computational cost of the neutron diffusion method.
+%Demonstrations of the hybrid method for time-dependent rod drop and reactivity insertion
+%simulations also performed well in reproducing expected trends observed in experimental data.
+A demonstration of the hybrid method for a time-dependent rod drop
+simulation also performed well in reproducing expected trends observed in experimental data.
 Analysis of its computational performance indicated the possibility of further optimizations
@@
 efficient computational performance, Moltres could enable accurate and cost-effective simulations
-of asymmetric transient in \glspl{MSR}.
+of asymmetric transients in \glspl{MSR}.

### 2024-12-09 conclusion.tex
Draft for committee before defense

@@
 may be possible to reduce data transfer times between the $S_N$ and diffusion solvers which took
-up about 30 \% of the total solution time due to inter-processor communications.
+up about 18 \% of the total solution time due to inter-processor communications.
 
-Lastly, Chapter \ref{chap:transient} demonstrated the hybrid method in time-dependent simulations
-based on a \gls{MSRE} zero-power rod drop experiment \cite{prince_zero-power_1968} and a \gls{MSRE}
-reactivity insertion experiment during power operation \cite{steffy_experimental_1970}. These
-simulations required coupling the hybrid method solver to \gls{DNP} and temperature solvers in
-Moltres through nested iteration coupling structures. I set up the rod drop simulation through a
+Lastly, Chapter \ref{chap:transient} demonstrated the hybrid method in a time-dependent simulation
+based on a \gls{MSRE} zero-power rod drop experiment \cite{prince_zero-power_1968}.
+The simulation required coupling the hybrid method solver to the pre-existing \gls{DNP} solver
+in Moltres through a nested iteration coupling setup.
+%and a \gls{MSRE}
+%reactivity insertion experiment during power operation \cite{steffy_experimental_1970}. These
+%simulation required coupling the hybrid method solver to \gls{DNP} and temperature solvers in
+%Moltres through nested iteration coupling structures.
+I set up the rod drop simulation through a
 $k$-eigenvalue simulation with the control rod at its initial height before using the simulation
 output as the initial condition for the time-dependent rod drop simulation. Moltres reproduced the
-expected shape of the integral neutron count curve from \gls{MSRE} experimental data with a maximum
-discrepancy of x \% at $t=16$ after the rod drop. Snapshots of the neutron flux spatial
-distribution at various timesteps show expected localized flux depressions around the control rod
-as it moves through the core.
-The reactivity insertion simulation started with the \gls{MSRE} initially at 1 MW power output and
-steady salt flow. Following a 24.8 pcm step reactivity insertion, Moltres reproduced the general
-shape of the first power peak and the subsequent power oscillations. The power peak magnitude
-is y \% higher than the peak in the experimental data. Temperatures in the \gls{MSRE} model rise
-and settle at higher values to compensate for the reactivity insertion.
+expected prompt and delayed response in the integral neutron count data observed in \gls{MSRE}
+experimental data. Convergence issues midway through the simulation had a minor impact on the
+solution precision. Overall, this work successfully demonstrated a time-dependent
+reactivity-initiated simulation using the hybrid method.
+%Snapshots of the neutron flux spatial
+%distribution at various timesteps show expected localized flux depressions around the control rod
+%as it moves through the core.
+%The reactivity insertion simulation started with the \gls{MSRE} initially at 1 MW power output and
+%steady salt flow. Following a 24.8 pcm step reactivity insertion, Moltres reproduced the general
+%shape of the first power peak and the subsequent power oscillations. The power peak magnitude
+%is y \% higher than the peak in the experimental data. Temperatures in the \gls{MSRE} model rise
+%and settle at higher values to compensate for the reactivity insertion.
 
@@
 
-The large 3-D simulations in this work uncovered significant memory usage by the $S_N$ solver, even
+The 3-D simulations in this work uncovered significant memory usage by the $S_N$ solver, even
 with the distributed mesh feature to spread mesh and variable storage across the compute nodes.
@@
 Jacobian formulations are identical aside from the level-symmetric ordinate and weight. Another
-area for optimization in the hybrid method is the data transfer times between the $S_N$ and
-diffusion solvers. Due to mesh distribution, each 3-D simulation spent 30 \% of its solution time
+area for optimization in the hybrid method is reducing data transfer times between the $S_N$ and
+diffusion solvers. Due to mesh distribution, each 3-D simulation spent 18 \% of its solution time
 on data transfers between processors. This could be minimized through optimizations in the mesh

### 2024-12-09 hybrid.tex
Draft for committee before defense

@@
 
-\section{\Gls{SAAF} $S_N$ Method} \label{sec:saaf}
+\section{Self-Adjoint Angular Flux (SAAF) $S_N$ Method} \label{sec:saaf}
 
@@
 section, the derivation for the \gls{SAAF} formulation of the $S_N$ method and its
-implementation in Moltres follows closely the derivation developed by Wang et al.\
+implementation in Moltres follows the derivation developed by Wang et al.\
 \cite{wang_diffusion_2014, wang_rattlesnake_2018}.
@@
 
-\subsection{\gls{SAAF} Formulation}
+\subsection{Self-Adjoint Angular Flux (SAAF) Formulation}
 
@@
   \shortintertext{where}
-  \hat{\Omega}_{d_r} = \hat{\Omega}_d - 2(\hat{\Omega}_d\cdot\hat{n}_b)\hat{n}_b. \nonumber
+  \hat{\Omega}_{d_r} = \hat{\Omega}_d - 2(\hat{\Omega}_d\cdot\hat{n}_b)\hat{n}_b = \mbox{ direction
+  of incident neutron travel}. \nonumber
 \end{gather}
@@
 
-\subsection{Diffusion Correction} \label{sec:diffusion-correction}
+\subsection{Diffusion Correction Scheme} \label{sec:diffusion-correction}
 
@@
 pointwise corrections to closely match the diffusion flux solution to the $S_N$ flux solution.
-By replacing $D_g$ with $D^s_g$, we are effectively adding the following transport correction term
-%
-\begin{gather}
+By replacing $D_g$ with $D^s_g$, we are effectively adding a transport correction term to the
+neutron diffusion equations.
+%
+\begin{multline}
+  \frac{1}{v_g} \frac{\partial \phi_g}{\partial t} - \frac{\partial}{\partial x} \cdot D_g
+  \frac{\partial \phi_g}{\partial x} 
   -\frac{\partial}{\partial x}(D^s_g-D_g)\frac{\partial\phi_g}{\partial x}
-\end{gather}
-to the neutron diffusion equations. Alternatively, we may define
-$\partial D\equiv\left(D^s_g-D_g)\right)$. This alternative form shows that diffusion
+  + \Sigma^r_g \phi_g = \\
+  \sum^G_{g' \neq g} \Sigma^s_{g' \rightarrow g} \phi_{g'}
+  + \chi^p_g \sum^G_{g'=1} \left( 1-\beta \right) \nu \Sigma^f_{g'}
+  \phi_{g'} + \chi^d_g \sum^I_i \lambda_i C_i \label{eq:neutron-diffusion-correction} %\\
+\end{multline}
+%\begin{gather}
+%  -\frac{\partial}{\partial x}(D^s_g-D_g)\frac{\partial\phi_g}{\partial x}
+%\end{gather}
+%to the neutron diffusion equations.
+Alternatively, we may define
+$\partial D\equiv\left(D^s_g-D_g\right)$. This alternative form shows that diffusion
 correction applies a multiplicative closure that scales with the flux gradient.
@@
 \cite{pounders_diffusion_2009} formulations for space-dependent diffusion coefficients in Eq.\
-\ref{eq:ronen} and Eq.\ \ref{eq:emp}, respectively. In comparison with
-the Ronen method, our approaches differ in how $\phi_{tr}$ and $J_{tr}$ are obtained. Starting with
+\ref{eq:ronen} and Eq.\ \ref{eq:emp}, respectively. This work differs from the Ronen method in how
+$\phi_{tr}$ and $J_{tr}$ are obtained. Starting with
 a standard neutron diffusion calculation, the Ronen method applies an analytically derived
@@
 
-For the empirical method developed by Pounders \& Rahnema, they assumed a priori knowledge of the
-reference flux solution and used it to generate piecewise-constant, space-dependent diffusion
-coefficients.
-Pounders \& Rahnema \cite{pounders_diffusion_2009} demonstrated the effectiveness of applying
+Pounders \& Rahnema's \cite{pounders_diffusion_2009} investigations focused on closures for the neutron diffusion method and
+assumed a priori knowledge of the reference flux solution.
+They demonstrated the effectiveness of applying
 pointwise corrections derived from analytical or Monte Carlo reference flux solutions. Compared
@@
 for piecewise-constant diffusion coefficients introduces some truncation error if the flux is
-non-linear within the mesh element. This design choice may be due to an intention to retain the
-diffusion coefficient as a constant in the $\frac{d}{dx}D\frac{d\phi}{dx}$ term of the neutron
+non-linear within the mesh element. This design choice retains the diffusion coefficient as a
+piece-wise constants in the $\frac{d}{dx}D\frac{d\phi}{dx}$ term of the neutron
 diffusion equation (Eq.\ \ref{eq:mg-diff}).
 
-Unlike their approach, my diffusion correction scheme (Eq. \ref{eq:svdc} allows for continuously
+Conversely, the diffusion correction scheme of this work (Eq. \ref{eq:svdc}) allows for continuously
 varying diffusion coefficients to reduce truncation error. In practice, the discretization order of
@@
 flux solutions \cite{gross_comprehensive_2023, pounders_diffusion_2009}.
-Howver, a significant challenge for diffusion corrections
+However, a significant challenge for diffusion corrections
 involves their resolution near neutron flux peaks. As observed in Eq.\ \ref{eq:svdc}, the neutron
@@
 resulting in diffusion coefficient values tending to positive or negative infinity when the
-flux gradient is close to zero. Pounders \& Rahnema avoided this issue by using
-larger mesh sizes to calculate their empirical diffusion coefficients. However, their
-remedy contradicts mesh convergence requirements and would worsen flux accuracy in regions with
+flux gradient is close to zero. Larger mesh sizes may help avoid this issue at the cost of
+worsening flux accuracy in regions with
 steep, non-linear fluxes, such as near control rods. For the Ronen method, Gross et al.\
@@
 
-\subsection{Drift Correction Term} \label{sec:drift-correction}
-
-Drift correction terms feature prominently in existing literature as transport correction
-terms for \gls{NDA} schemes or nodal diffusion methods
+\subsection{Drift Correction Scheme} \label{sec:drift-correction}
+
+Drift terms may be added to neutron diffusion or equivalent low-order equations to correct them
+such that the modified equations can reproduce the transport solution upon reaching iterative
+convergence. Drift correction terms feature prominently in existing literature as transport
+correction terms for \gls{NDA} schemes or nodal diffusion methods
 \cite{smith_nodal_1983, smith_assembly_1986, adams_fast_2002, wang_diffusion_2014}. \gls{NDA}
@@
 accelerating a neutron transport
-method (high-order) with a modified neutron diffusion method (low-order). Drift terms
-are added to the low-order neutron diffusion equations to correct them such that the modified
-equations can reproduce the transport solution upon reaching iterative convergence. In nodal
+method (high-order) with a modified neutron diffusion method (low-order). In nodal
 diffusion methods, the drift terms are used to rectify the incoming and outgoing neutron
@@
 
-The general form of drift correction terms to be added to the neutron diffusion equations is a
-first-order derivative term $\nabla\cdot \vec{D}_g\phi_g$ (note the vector notation). As such,
+The general form of a drift correction term for the neutron diffusion equations is a
+first-order derivative $\nabla\cdot \vec{D}_g\phi_g$ (note the vector notation). As such,
 drift terms apply
 multiplicative closures that scale with the flux. This contrasts with diffusion
-corrections which scale with the flux gradient instead. Therefore, drift terms do not suffer from
+corrections which scale with the flux gradient. Therefore, drift terms do not suffer from
 the division-by-zero errors encountered with diffusion corrections.
@@
 and $\partial\mathcal{D}_v$ represents sections of $\partial\mathcal{D}$ which are vacuum boundaries. We
-assumed that $\bm{Q}$ is an isotropic source because we do not deal with anisotropic sources in
+assume that $\bm{Q}$ is an isotropic source because we do not deal with anisotropic sources in
 this work outside of scattering.
@@
   + \left(\bm{\Phi}^*,\frac{\bm{\Phi}}{2}\right)_{\mathcal{D}_v}
-  + \left(\nabla\bm{\Phi}^*,\bm{\tau}\vec{\bm{r}}_1-\vec{\bm{J}}-\mathcal{D}\bm{\Phi}\right)_\mathcal{D}
+  + \overbrace{\left(\nabla\bm{\Phi}^*,\bm{\tau}\vec{\bm{r}}_1-\vec{\bm{J}}-\mathcal{D}\bm{\Phi}\right)_\mathcal{D}}^\text{(1)}
   + \\
-  \left(\bm{\Phi}^*,\bm{J}^\text{out}-\frac{\bm{\Phi}}{2}\right)_{\partial\mathcal{D}_v}
+  \underbrace{\left(\bm{\Phi}^*,\bm{J}^\text{out}-\frac{\bm{\Phi}}{2}\right)_{\partial\mathcal{D}_v}}_\text{(2)}
   = \left(\bm{\Phi}^*,\mathbb{S}_0\bm{\Phi}\right)_\mathcal{D}
@@
 We can observe that Eq.\ \ref{eq:modified-diff} is a modified neutron diffusion equation with
-transport corrections provided by the fifth and sixth terms. The fifth term is a drift term with
+transport corrections provided by the terms labeled (1) and (2). Term (1) is a drift term with
 the drift vector defined as
@@
 %
-and the sixth term is a boundary correction term with the boundary coefficient vector defined as
-%
-\begin{gather}
-  \bm{\gamma} \equiv \frac{\bm{J}^\text{out}}{\bm{\Phi}}-\frac{1}{2}\bm{I}.
-\end{gather}
-%
+and Term (2) is a boundary correction term with the boundary coefficient vector defined as
+%
+\begin{gather}
+  \bm{\gamma} \equiv \frac{\bm{J}^\text{out}}{\bm{\Phi}}-\frac{1}{2}\bm{I}_3
+\end{gather}
+where $\bm{I}_3$ is the 3-D identity matrix.
 With the $S_N$ angular discretization scheme, the drift and boundary correction vector components
@@
   \item With the $S_N$ subsolver, evaluate transport correction terms in $V_1$ using Eq.
-    \ref{eq:svdc} or \ref{eq:drift}.
+    \ref{eq:svdc} (diffusion correction) or \ref{eq:drift} (drift correction).
   \item Pass the transport correction terms to the neutron diffusion solver to apply corrections
@@
 With this relation, we can formulate the boundary source term for the $S_N$ subsolver by setting
-$\Psi^\text{inc}_{g,d}$ in Eq.\ \ref{eq:boundary-source} to
+the incident boundary flux source,
+$\Psi^\text{inc}_{g,d}$, in Eq.\ \ref{eq:boundary-source} to
 %
@@
 The $S_N$ subsolver boundary conditions will generally cause deviations
-in the flux and transport correction values since the $P_1$ approximation is accurate for up to
-linear anisotropies in the flux. However, as I will show in Chapter \ref{chap:msre},
+in the flux and transport correction values since the $P_1$ approximation captures
+linear anisotropic flux at best. However, as I will show in Chapter \ref{chap:msre},
 the influence of boundary conditions on the accuracy of
@@
 ($D^s_g$ or $\vec{D}_g$) generated with the $S_N$ subsolver within $V_1$ are accurate everywhere up
-to a few centimeters away from $\partial V_1$.
+to a few centimeters inside of $V_1$.
 
@@
 opposite to $\partial V_1$. A natural/intuitive criterion for the location of the cutoff boundary
-would be wherever the components of the drift correction variable $\vec{D}_g$ is zero, i.e.,
-wherever the components change signs. The reasons for this criterion are twofold. Firstly, at
+would be where the components of the drift correction variable $\vec{D}_g$ become zero (i.e.,
+where the components change signs). The reasons for this criterion are twofold. Firstly, at
 points where the $\vec{D}_g$ components are zero, the flux is approximately isotropic along the
@@
 
-\section{Numerical Implementation}
+\section{Numerical Implementation} \label{sec:numerical-implementation}
 
@@
 iterative coupling between the $S_N$ and neutron diffusion solvers for the hybrid $S_N$-diffusion method.
-The \texttt{MultiApp} and \texttt{Transfers} systems from \gls{MOOSE} provide flexible interfaces
-for setting up iterative solvers and transferring data between the solvers. A typical hybrid
-method simulation on Moltres starts with the neutron diffusion solver for the first estimate of the
-neutron flux distribution. Moltres transfers the neutron flux distribution to the $S_N$ solver to
-compute the boundary source distribution. The $S_N$ solver then generates its own flux estimate and
-computes drift and boundary correction variables for the neutron diffusion solver. Moltres
-transfers the correction variables to the neutron diffusion solver to compute its next flux
-estimate. This process iterates until the neutron flux distribution converges to within tolerance
-values provided in the input file.
+The \texttt{MultiApps} and \texttt{Transfers} systems from \gls{MOOSE} provide flexible interfaces
+for setting up iterative solvers and transferring data between the solvers.
+Excluding the initial neutron diffusion calculation,
+each outer (fixed point) iteration comprises of an inner $S_N$ calculation on the correction
+subregion and an inner neutron diffusion calculation on the entire domain. The \gls{FEM} algorithm
+in Moltres determines an individual solver reaches convergence when the combined
+residual contributions fall below a specified convergence tolerance threshold. The fixed point
+iteration scheme in the \texttt{MultiApps} system for coupling multiple individual
+solvers (referred to as apps) requires users to specify a main app for the final convergence check
+after each outer iteration to terminate the iterations. The scheme also requires users to set a
+fixed point convergence tolerance value independent of the tolerance values of each individual
+solver. Naturally, the fixed point tolerance value must be equal to or greater than the tolerance
+value of the main individual solver for the possibility of overall convergence. The hybrid method
+implementation in Moltres sets the neutron diffusion solver as the main app and the $S_N$ subsolver
+as the sub app.
+
+%A typical hybrid
+%method simulation on Moltres starts with the neutron diffusion solver for the first estimate of the
+%neutron flux distribution. Moltres transfers the neutron flux distribution to the $S_N$ solver to
+%compute the boundary source distribution. The $S_N$ solver then generates its own flux estimate and
+%computes drift and boundary correction variables for the neutron diffusion solver. Moltres
+%transfers the correction variables to the neutron diffusion solver to compute its next flux
+%estimate. This process iterates until the neutron flux distribution converges to within tolerance
+%values provided in the input file.
 
@@
 I presented two different formulations of transport correction that I investigated in this work for
-implementation in the hybrid method: diffusion correction and drift correction. Diffusion
+implementation in the hybrid method: diffusion correction and drift correction schemes. Diffusion
 correction involves replacing the diffusion coefficient in the neutron diffusion equations with
@@
 values near the $S_N$ subsolver boundary are inaccurate due to the approximate boundary conditions
-provided by the diffusion subsolver. Thus, I developed and presented a numerical fix for discarding
-inaccurate transport correction values and preserving the smoothness of the neutron flux and flux
-gradient solution.
+provided by the diffusion subsolver. Thus, I developed an adaptive boundary algorithm that
+establishes a buffer region in which transport correction parameters are discarded to improve
+$S_N$-diffusion coupling and preserve smoothness of the neutron flux gradients.
 
@@
 enable full-core, time-dependent reactor simulations on \gls{HPC} clusters.
-In the next chapter, I verify the hybrid method against Monte Carlo and $S_N$ neutron transport
+In the next chapter, I verify the hybrid method with both diffusion and drift correction schemes
+against Monte Carlo and $S_N$ neutron transport
 methods through various neutronics test cases and discuss some significant observations made in my

### 2024-12-09 introduction.tex
Draft for committee before defense

@@
 
-    The second objective is to
-    develop a novel hybrid $S_N$-diffusion method for accurate control rod modeling in
+    The second objective is to develop, implement, and demonstrate a novel hybrid $S_N$-diffusion
+    method for accurate control rod modeling in
     time-dependent \gls{MSR} analyses. The hybrid method uses the discrete ordinates ($S_N$)
@@
 experimental data.
-Chapter 6 presents the demonstration of the hybrid $S_N$-diffusion method in time-dependent
-reactivity-initiated simulations modeling the \gls{MSRE} rod drop and reactivity insertion
-experiments.
+%Chapter 6 presents the demonstration of the hybrid $S_N$-diffusion method in time-dependent
+%reactivity-initiated simulations modeling the \gls{MSRE} rod drop and reactivity insertion
+%experiments.
+Chapter 6 presents a demonstration of the hybrid $S_N$-diffusion method in a time-dependent
+reactivity-initiated simulation modeling a \gls{MSRE} rod drop experiment.
 Chapter 7 concludes this dissertation by summarizing the results presented, identifying limitations

### 2024-12-09 lit.tex
msre transient section additions
Draft for committee before defense

@@
 taken to capture various multiphysics interactions accurately. For instance, the strong coupling
-between neutron flux and temperature in \glspl{MSR} in \glspl{MSR} necessitates robust coupling
+between neutron flux and temperature in \glspl{MSR} necessitates robust coupling
 techniques. Na\"ive multiphysics coupling algorithms may result in longer computation times
@@
 \subsection{Transport-Correction Techniques With Neutron Diffusion-Based Solvers}
+\label{sec:lit-tc}
 
@@
 on the size of the absorber region and coefficients quantifying the escape probability of neutrons
-entering the absorber region. The $\lambda$ and the associated coefficients were derived
+entering the absorber region. $\lambda$ is derived
 analytically from the \gls{BTE} with some simplifying assumptions, such as having uniform or
 cosine-shaped incident neutron currents, isotropic scattering within the absorber, and mathematical
-approximations in ignoring higher-order terms in intermediate steps. Blackness theory emerged in
-the mid-20th century when computational resources were limited. Later, as computational
+approximations in ignoring higher-order terms in intermediate steps. Absorber blackness methods
+emerged in the mid-20th century when computational resources were limited. Later, as computational
 resources became more readily available and powerful enough for more complicated transport
@@
 for time-dependent \gls{MSR} analyses. Control rods induce highly anisotropic fluxes within
-themselves and their vicinities.
-Unfortunately, higher-fidelity neutron transport methods are typically too computationally
-expensive for multiphysics \gls{MSR} simulations, while faster methods like neutron
-diffusion perform poorly in regions with highly anisotropic fluxes. However, transport-correction
+themselves and their vicinities that the neutron diffusion approximation cannot model accurately.
+Inaccurate rod worth estimates subsequently affect transient reactor power and temperature
+distribution calculations. Higher-fidelity neutron transport methods remain too computationally
+expensive for time-dependent multiphysics \gls{MSR} simulations. However, transport-correction
 techniques can be applied to augment neutron diffusion calculations in these regions. While few

### 2024-12-09 main.tex
add ray effect subsection
tikz preamble for previous commit
Draft for committee before defense

@@
 \usepackage{subcaption}
+\usepackage[rightcaption]{sidecap}
+\sidecaptionvpos{figure}{c}
 \usepackage{hhline}
@@
 \usepackage{stmaryrd}
+\usepackage{textcomp}
 
@@
 \usepackage{tkz-euclide}
-\usetikzlibrary{positioning, arrows, decorations, shapes}
+\usetikzlibrary{positioning, arrows, decorations, shapes, fit, backgrounds}
 \usetikzlibrary{shapes.geometric,arrows}
@@
 \tikzstyle{arrow} = [thick,->,>=stealth]
+\tikzstyle{solver} = [rectangle, rounded corners, thick, minimum width=3cm, minimum height=1.5cm, text centered, draw=black, fill=illiniblue, text width=7em]
+\tikzstyle{bound} = [rectangle, rounded corners, thick, draw=black]
 
@@
 \degreeyear{2024}
-\committee{Associate Professor Kathryn D. Huff, Chair \& Co-Director of Research \\
-           Assistant Professor Madicken Munk, Co-Director of Research, Oregon State University \\
+\committee{Associate Professor Kathryn D. Huff, Co-Chair \\
+           Assistant Professor Madicken Munk, Co-Chair \\
            Professor Rizwan Uddin \\

### 2024-12-09 msre.tex
Draft for committee before defense

@@
-In Chapter \ref{chap:hybrid}, I laid out the mathematical background and numerical implementation
+Chapter \ref{chap:hybrid} laid out the mathematical background and numerical implementation
 of the hybrid $S_N$-diffusion method for improving neutronic accuracy in reactor systems containing
-highly neutron-absorbing regions. I developed and implemented the diffusion and drift correction
-schemes for the hybrid method in Python and using Moltres, respectively. This chapter demonstrates
+highly neutron-absorbing regions. Chapter \ref{chap:hybrid} specifically discussed two transport
+correction schemes for the hybrid method: the diffusion and drift correction
+schemes. This chapter demonstrates
 the hybrid method through 1-D, 2-D, and 3-D $k$-eigenvalue simulations of the \gls{MSRE}.
-I employed the OpenMC Monte Carlo neutral particle transport software \cite{romano_openmc:_2015} to
-generate group constants for the hybrid method and reference solutions for method verification.
+This chapter employs the OpenMC Monte Carlo neutral particle transport software \cite{romano_openmc:_2015} to
+generate group constants for the hybrid method and to generate reference solutions for method verification.
 The 1-D models serve to verify the newly-implemented $S_N$ and hybrid methods
-against OpenMC as well as to analyze the impact of key input parameters on the hybrid method. The
-2-D models test the hybrid method performance through increased spatial dimensionality and reduced
-geometric asymmetry relative to the 1-D models. Lastly, the 3-D models serve as neutronics model
-validation against the \gls{MSRE} benchmark model by Fratoni et al.\ \cite{fratoni_molten_2020}, a
-testbed for \gls{HPC} scaling tests, and preparation for the time-dependent multiphysics
-simulations in Chapter \ref{chap:transient}.
+and to analyze the impact of key input parameters on the hybrid method. The
+2-D models test the hybrid method performance through increased spatial dimensionality and
+geometric asymmetry relative to the 1-D models. Parametric studies using 2-D models on the $S_N$ angular
+discretization reveal its impact on the neutron flux distributions and $k_\text{eff}$.
+Lastly, the 3-D models serve as neutronics model
+validation against the \gls{MSRE} benchmark model by Fratoni et al.\ \cite{fratoni_molten_2020}.
+This chapter also investigates the computational performance of 3-D simulations on a \gls{HPC}
+cluster.
 
@@
 neutronics verification and validation. Sections \ref{sec:1d-results}, \ref{sec:2d-results}, and
-\ref{sec:3d-results} discuss the model setup, results, and discussions for the 1-D, 2-D, and 3-D
+\ref{sec:3d-results} discuss the model setups, results, and discussions for the 1-D, 2-D, and 3-D
 \gls{MSRE} models, respectively.
@@
 \begin{align*}
-  \Sigma_{t,g}:& \mbox{ Macroscopic total cross section in group $g$,}
-  \Sigma_{r,g}:& \mbox{ Macroscopic removal cross section in group $g$,}
-  \Sigma_s^{g'\rightarrow g}:& \mbox{ Macroscopic group-to-group scattering cross section matrix,}
-  \Sigma_{s,l}^{g'\rightarrow g}:& \mbox{ $l$-th Legendre moment of the macroscopic
-  group-to-group scattering cross section matrix,}
-  \Sigma_{sp,l}^{g'\rightarrow g}:& \mbox{ $l$-th Legendre moment of the macroscopic
-  group-to-group scattering production cross section matrix,}
-    D_g:& \mbox{ $P_1$-based diffusion coefficient in group $g$,}
-  \nu\Sigma_{f,g}:& \mbox{ Product of the average number of neutrons produced per fission and the
-  macroscopic fission cross section in group $g$,}
-    \chi_g:& \mbox{ Neutron fission spectrum in group $g$.}
+  \Sigma_{t,g}=& \mbox{ Macroscopic total cross section in group $g$,}\\
+  \Sigma_{r,g}=& \mbox{ Macroscopic removal cross section in group $g$,}\\
+  \Sigma_s^{g'\rightarrow g}=& \mbox{ Macroscopic group-to-group scattering} \\
+  \mbox{cross section matrix,}\\
+  \Sigma_{s,l}^{g'\rightarrow g}=& \mbox{ $l$-th Legendre moment of the macroscopic
+  group-to-group scattering cross section matrix,}\\
+  \Sigma_{sp,l}^{g'\rightarrow g}=& \mbox{ $l$-th Legendre moment of the macroscopic
+  group-to-group scattering production} \\
+  \mbox{cross section matrix,}\\
+  D_g=& \mbox{ $P_1$-based diffusion coefficient in group $g$,}\\
+  \nu\Sigma_{f,g}=& \mbox{ Product of the average number of neutrons produced per fission and} \\
+  \mbox{the macroscopic fission cross section in group $g$,}\\
+  \chi_g=& \mbox{ Neutron fission spectrum in group $g$.}\\
 \end{align*}
 %
-I generated the group constants using OpenMC's multigroup cross section generation capability
+OpenMC generates these group constants using its multigroup cross section generation capability
 \cite{boyd_multigroup_2019}. A group constant postprocessing script for Moltres
@@
 knockout reactions as scattering reactions, but $\Sigma_{r,g}$ is a convenient parameter in which
-to incorporate neutron knockout effects in the neutron diffusion method. I ran all test cases
-for $S_N$ and hybrid method calculations with scattering cross sections approximated by up to 2nd
-Legendre moments ($L=2$).
+to incorporate neutron knockout effects in the neutron diffusion method.
+All simulations in this thesis used 2nd-order-accurate scattering cross sections approximated using
+Legendre polynomials.
 
@@
 
-I ran all \gls{MSRE} neutronics simulations with the eight neutron energy group structure proposed
+All \gls{MSRE} neutronics simulations ran with the eight neutron energy group structure proposed
 by Jaradat for \gls{MSRE} analysis \cite{jaradat_development_2021-1}.
 Table \ref{table:energy-group} shows the upper neutron energy bounds of the eight-group structure.
-I selected this energy group structure as a compromise between accuracy and computational expense
-among the various group structures investigated by Jaradat ranging from four to sixteen groups.
+This structure balances accuracy and computational expense most appropriately for this work when
+compared to other various group structures investigated by Jaradat ranging from four to sixteen
+groups.
 
@@
 
-I set the material compositions and densities of every component in the \gls{MSRE} models using
-reference data from \gls{MSRE} benchmark specifications collated by Fratoni et al.
-\cite{fratoni_molten_2020}. All OpenMC models utilize the ENDF/B-VII.1 nuclear data library
-\cite{chadwick_endf/b-vii.1_2011} and include thermal scattering cross section data for beryllium
-and carbon target nuclei in the fuel salt and graphite moderator. All models have a uniform
+Material compositions and densities of all components in the \gls{MSRE} models are from reference
+\gls{MSRE} specifications \cite{robertson_msre_1965, fratoni_molten_2020}.
+All OpenMC simulations in this work used the ENDF/B-VII.1 nuclear data library
+\cite{chadwick_endf/b-vii.1_2011}. All models have a uniform
 temperature of 900 K unless otherwise stated. Detailed information pertaining specifically to the
@@
 consideration. Comparing OpenMC-MG results with OpenMC-CE results allows us
-to quantify errors arising from neutron energy group discretization and the scattering cross
-section approximations.
+to quantify errors arising from neutron energy group discretization and the 2nd-order scattering
+cross section approximations.
 
@@
     Energy, $E$ & Continuous & Discrete & Discrete & Discrete \\
-    Angle-dependence in $\Sigma_s$ & Continuous & $L=2$ & $L=2$ & N/A \\
+    Angle resolution in $\Sigma_s$ & Continuous & $L=2$ & $L=2$ & N/A \\
     \bottomrule
@@
 For the 1-D models, I performed neutronics calculations with the $S_N$, neutron diffusion, and
-hybrid method implementations on Moltres and Python. Table \ref{table:nts-methods} highlights the
-differences between the two implementations. The $S_N$ neutron transport method implementation on
+hybrid method implementations in Moltres and Python. Table \ref{table:nts-methods} highlights the
+differences between the two implementations. The $S_N$ neutron transport method implementation in
 Moltres follows the methodology for solving the \gls{SAAF} formulation of the $S_N$ equations as
@@
 
-The implementations on Moltres discretize the problem domain in space using the finite element
-method while the implementations in Python use finite difference or diamond difference schemes. The
+Moltres discretizes the problem domain in space using the finite element
+method while the Python implementation uses finite difference or diamond difference schemes. The
 most crucial difference lies in the transport correction formulations for the hybrid method. The
@@
 
-In this chapter, I refer to the methods and results from the Moltres and Python implementations
+This chapter refers to the methods and results from the Moltres and Python implementations
 with the Moltres- or Python- prefix, i.e., Moltres-$S_N$, Moltres-diffusion, Moltres-hybrid,
@@
 
-For 2- and 3-D models, I abandoned the diffusion correction formulation in favor of the drift
+For 2- and 3-D models, the diffusion correction formulation is abandoned in favor of the drift
 correction formulation due to superior numerical performance in the latter. Readers may find the
@@
 
-In this section, I define several metrics for comparing simulation results among the various
+This section defines several metrics for comparing simulation results among the various
 neutronics methods.
@@
 
-I designed six 1-D test cases with increasing complexity to verify the $S_N$ and hybrid
+This section covers six 1-D test cases with increasing complexity to verify the $S_N$ and hybrid
 $S_N$-diffusion implementations and to test the performance of the hybrid
@@
 
-In my preliminary investigations, the original control rod material composition
+In preliminary investigations, the original control rod material composition
 (70 wt\% Gd$_2$O$_3$-30 wt\% Al$_2$O$_3$) led to excessively large control rod worth estimates of
@@
 than the 5500 pcm expected from full rod insertion of all three rods in the actual \gls{MSRE}
-\cite{fratoni_molten_2020}. Therefore, I greatly reduced the proportion of Gd$_2$O$_3$ in the
-control rod to 0.35 wt\% for a fairer assessment of rod worth in the 1-D models. This change
+\cite{fratoni_molten_2020}. Therefore, the proportion of Gd$_2$O$_3$ was reduced to 0.35 wt\% for a
+fairer assessment of rod worth in the 1-D models. This change
 brought the new rod worth estimates down to below 20000 pcm.
@@
 
-I ran mesh convergence studies based on Case 3b for the neutron diffusion and $S_N$ neutron
-transport methods. Given that the hybrid $S_N$-diffusion method comprises of a combination of those
-methods, I found hybrid simulations were fully spatially resolved on meshes that are
-appropriately resolved for both neutron diffusion and $S_N$ simulations.
-
-\begin{figure}[htb!]
+This mesh convergence tests were run on Case 3b for the neutron diffusion and $S_N$ neutron
+transport methods. Table \ref{table:mesh-size} shows the mesh element sizes at mesh refinement
+level 0. Mesh element sizes approximately halve with each successive refinement level up to level 4.
+
+\begin{table}[t]
+  \centering
+  \caption{Mesh element sizes at mesh refinement level 0.}
+  \begin{tabular}{l S}
+    \toprule
+    Region & {Mesh size [cm]} \\
+    \midrule
+    Control rod (interior) & 0.04 \\
+    Control rod (near interface) & 0.025 \\
+    Air & 0.125 \\
+    Fuel & 0.1875 \\
+    Graphite & 0.19375 \\
+    Vessel & 0.2 \\
+    \bottomrule
+  \end{tabular}
+  \label{table:mesh-size}
+\end{table}
+
+\begin{figure}[t]
   \centering
@@
   \end{subfigure}
-  \caption{Convergence of multiplication factor ($k_\text{eff}$) estimates for Case 3b across five
-    levels of mesh refinement. Higher refinement levels correspond to smaller mesh element sizes.}
+  \caption{Convergence of multiplication factor ($k_\text{eff}$) estimates for Case 3b across four
+    levels of mesh refinement relative to the finest mesh resolution.}
   \label{fig:mesh-k}
@@
 estimates observed for Case 3b simulations with neutron diffusion and $S_8$ neutron transport
-methods. The $k_\text{eff}$ values plateau at 0.9713203 and 0.9813790 for the neutron diffusion and
-$S_8$ methods, respectively. By mesh refinement level 2, both methods fall within $5\times 10^{-7}$
-range of the $k_\text{eff}$ value at level 4.
+methods. The $k_\text{eff}$ values plateau at 0.97132 and 0.98138 for the neutron diffusion and
+$S_8$ methods, respectively. After two rounds of mesh refinements, both methods fall within
+$5\times 10^{-7}$ range of the reference $k_\text{eff}$ values, below the 0.1 pcm ($10^{-5}$)
+threshold for determining solution convergence.
 
@@
 group flux distributions at mesh refinement levels 1 to 3 relative to the reference group flux
-distribution at level 4 (maximum refinement). I omitted data from level 0 to improve the fidelity
+distribution at level 4 (maximum refinement). The plots omit data from level 0 to improve the fidelity
 of the plots as flux differences drop significantly by one order of magnitude with each level of
 refinement.
-With these results, I concluded that the mesh resolution at level 2 is a sufficient guideline for
-the mesh resolutions used in the subsequent 1-D analyses. The absolute differences in both
-$k_\text{eff}$ and group flux distributions at level 2 remain below $5\times10^{-7}$ relative to
-level 4.
+These results show that the mesh resolution at level 2 is a sufficient guideline for
+the mesh resolutions used in the subsequent 1-D analyses. Given that the hybrid $S_N$-diffusion
+method comprises of a combination of those
+methods, the hybrid simulations were fully spatially resolved on meshes that are
+appropriately resolved for both neutron diffusion and $S_N$ simulations.
 
@@
   \centering
-  \includegraphics[width=\columnwidth]{diffusion-mesh-convergence-flux}
+  \includegraphics[width=.95\columnwidth]{diffusion-mesh-convergence-flux}
   \caption{Absolute difference in group fluxes ($\phi_g$) from the neutron diffusion method for
-  mesh refinement levels 1 to 4 relative to the reference group flux distributions at mesh
-  refinement level 5.}
+  mesh refinement levels 1 to 3 relative to the reference group flux distributions at mesh
+  refinement level 4. Flux differences fall by an order of magnitude with each successive mesh
+  refinement.}
   \label{fig:diff-mesh-flux}
@@
   \centering
-  \includegraphics[width=\columnwidth]{sn-mesh-convergence-flux}
+  \includegraphics[width=.95\columnwidth]{sn-mesh-convergence-flux}
   \caption{Absolute difference in group fluxes ($\phi_g$) from the $S_8$ neutron transport method
-  for mesh refinement levels 1 to 4 relative to the reference group flux distributions at mesh
-  refinement level 5.}
+  for mesh refinement levels 1 to 3 relative to the reference group flux distributions at mesh
+  refinement level 4. Flux differences fall by an order of magnitude with each successive mesh
+  refinement.}
   \label{fig:sn-mesh-flux}
@@
 method reactivity values agree closer with the reference OpenMC-CE value. However, this is likely
-due to favorable error cancellation; Figures \ref{fig:3a-flux} and \ref{fig:3a-flux-diff} show
-OpenMC-MG and the $S_8$ methods reproduce
+due to favorable error cancellation in these lower-fidelity methods. Figures \ref{fig:3a-flux} and
+\ref{fig:3a-flux-diff} show OpenMC-MG and the $S_8$ methods reproduce
 the OpenMC-CE flux distribution more accurately than the neutron diffusion and hybrid methods.
@@
 of 8\%. Most of the remaining methods cluster around 2.5\% and 3\% percentage difference for Cases
-2 and 3 with the exception of the Python-hybrid method for Case 3.
+2 and 3 with the exception of the Python-hybrid method for Case 3 reporting a 1.5\% difference.
+The diffusion correction scheme in Python deviates from the drift correction
+scheme in Moltres due to numerical issues with division by zero around flux maximas and minimas
+in the fuel-graphite lattice region.
 
@@
 provide significant improvements in neutron flux distributions compared to the neutron diffusion
-method as shown by the generally smaller flux difference values particularly near $x=0$ cm. The
+method as shown by the generally smaller per-group flux deviations particularly near $x=0$ cm. The
 figures also show Moltres-hybrid (green) provides more accurate flux estimates than
@@
 selectively apply the $S_N$ method to improve neutronic modeling within the subregions. I computed
-the reference distributions (yellow data points) in both plots from reference flux solutions of the
+the reference distributions (orange data points) in both plots from reference flux solutions of the
 standard $S_8$ method.
@@
 The prior reactivity and flux distribution results indicate Moltres-hybrid is more accurate
-than Python-hybrid relative to OpenMC-MG as the reference.
+than Python-hybrid relative to OpenMC-MG as the reference. Therefore, subsequent hybrid method
+simulations in this work are solely with the hybrid method with the drift correction scheme in
+Moltres.
 
@@
 pcm for Case 3b. An important observation here is that the $k_\text{eff}$ values do not
-monotonically converge towards the $k_\text{eff}$ estimate from the $S_8$ method. The data implies
-other significant sources of discrepancies exist beyond those being corrected in the
-correction subregion by the hybrid method.
+monotonically converge towards the $k_\text{eff}$ estimate from the $S_8$ method. This behavior
+implies that neutron leakage error remains a significant source of discrepancy as long as the
+correction region does not encompass the outer boundary.
 
@@
 worth estimates do not change significantly when the correction subregion
-sizes change. This indicates limiting the correction subregion size to save
+sizes change. The rod worth estimates for all investigated correction subregion sizes remain within
+0.2\% of the $S_8$ method. This indicates limiting the correction subregion size to save
 computational cost on the expensive $S_N$ calculations has a negligible impact on rod worth
-estimates. The rod worth estimates for all investigated correction subregion sizes remain within
-0.2\% of the $S_8$ method.
+estimates.
 
@@
 The $k_\text{eff}$ and rod worth results show the hybrid method is able to accurately
-reproduce rod worth estimates as long as the correction subregion size remains consistent. The
+reproduce rod worth estimates as long as the correction subregion size remains consistent between
+$k$-eigenvalue simulations used for rod worth calculations. The
 correction subregion size must also be sufficiently large to capture the highly neutron-absorptive
@@
 
-As described in Section \ref{sec:hybrid-algorithm}, the hybrid $S_N$-diffusion method works by
-iteratively coupling a neutron diffusion solver and a $S_N$ subsolver with the latter generating
-drift correction parameters for the former. Excluding the initial neutron diffusion calculation,
-each outer (fixed point) iteration comprises of an inner $S_N$ calculation on the correction
-subregion and an inner neutron diffusion calculation on the entire domain. The \gls{FEM} algorithm
-in Moltres determines an individual solver reaches convergence when the combined
-residual contributions fall below a specified convergence tolerance threshold. The fixed point
-iteration scheme under the \texttt{MultiApps} system in Moltres for coupling multiple individual
-solvers (referred to as apps) requires users to specify a main app for the final convergence check
-after each outer iteration to terminate the iterations. The scheme also requires users to set a
-fixed point convergence tolerance value independent of the tolerance values of each individual
-solver. Naturally, the fixed point tolerance value must be equal to or greater than the tolerance
-value of the main individual solver for the possibility of overall convergence. The hybrid method
-implementation in Moltres sets the neutron diffusion solver as the main app and the $S_N$ subsolver
-as the sub app.
-
+As described in Sections \ref{sec:hybrid-algorithm} \& \ref{sec:numerical-implementation}, the
+hybrid $S_N$-diffusion method works by iteratively coupling a neutron diffusion solver and a $S_N$
+subsolver with the latter generating drift correction parameters for the former.
 The hybrid method shares some similarities with diffusion-based \gls{HOLO} acceleration schemes
@@
   \item Set channel pitch, width, and length to 2, 0.2, and 1.2 inches, respectively.
-  \item Homogenized the INOR-8, graphite, and molten salt regions in the sample basket.
+  \item Homogenized the INOR-8 (inconel alloy reactor vessel), graphite, and molten salt regions in the sample basket.
   \item Modified the outer edge of the core graphite to be circular with no jagged edges.
@@
 
-I ran neutronics simulations on the quarter- and full-core models on OpenMC under both continuous
-energy (OpenMC-CE) and multigroup (OpenMC-MG) modes, and on Moltres with the hybrid $S_N$-diffusion
+I ran neutronics simulations on the quarter- and full-core models using OpenMC under both continuous
+energy (OpenMC-CE) and multigroup (OpenMC-MG) modes, and using Moltres with the hybrid $S_N$-diffusion
 and neutron diffusion methods. The hybrid method ran with the $S_8$ angular discretization scheme
@@
 8370(53) pcm. The neutron diffusion method overestimated the rod worth by 1484 pcm. The hybrid
-method improved significantly on the neutron diffusion method with a rod worth estimate that is
-only 13 pcm
-higher than OpenMC-CE. OpenMC-MG reported a larger error of 172 pcm likely due to favorable error
-cancellation with the hybrid method and statistical uncertainties with the Monte Carlo methods.
+method improved significantly over the neutron diffusion method with a rod worth estimate that is
+only 13 pcm higher than OpenMC-CE. OpenMC-MG reported a larger error of 172 pcm. This is likely due
+to statistical uncertainty in OpenMC-MG and favorable error cancellation between different sources
+of error in the hybrid method (e.g., neutron leakage, heterogeneous core geometry).
+.
 The OpenMC-CE and OpenMC-MG rod worth estimates agree within two standard deviations of each other.
@@
 center as expected of a highly neutron-absorbing material. The rod further worsened fission rate
-errors for the neutron diffusion method with large errors of 1\% or more at the core center and
+errors for the neutron diffusion method with errors of 1\% or more at the core center and
 periphery. On the other hand, the fission rate errors for the hybrid method largely remained below
@@
 method. The absolute maximum percentage error from the neutron diffusion method jumped to 17.44\%
-as compared to the 3.08\% for the hybrid method in the ``Rod'' case. Along with the figures, these
+as compared to 3.08\% for the hybrid method in the ``Rod'' case. Along with the figures, these
 results demonstrate the hybrid method provides improved fission rate distribution accuracy in
@@
 when all three rods are inserted. These results demonstrate the hybrid method remains
-effective at improving control rod worth and fission rate distribution estimates in less symmetric
-geometries.
+effective at improving control rod worth and fission rate distribution estimates in geometries with
+asymmetric rod configurations.
 
@@
 The actual design has rounded upper and lower plena to help direct salt flow through the system.
-This work uses a modified 3-D numerical \gls{MSRE} model as shown by the vertical cross section
-in Figure \ref{fig:msre-geom-vert}. The model excludes the toroidal flow distributor near the top
+This work uses a modified 3-D numerical \gls{MSRE} model as shown by Figures
+\ref{fig:msre-geom-vert} and \ref{fig:wedge-view}. The model excludes the toroidal flow distributor
+near the top
 of the vessel and reshapes the plena into right cylinders. It has the same horizontal cross section
@@
   \includegraphics[width=0.97\columnwidth]{msre-picture}
-  \caption{Vertical cross section of the actual \gls{MSRE} vessel. It has rounded upper and lower
+  \caption{Vertical cross section of the actual \gls{MSRE} vessel \cite{robertson_msre_1965}.
+  It has rounded upper and lower
   plena, and a toroidal flow distributor that distributes the salt inflow evenly as it enters the
@@
   \label{fig:msre-geom-vert}
+\end{figure}
+
+\begin{figure}[p]
+  \centering
+  \includegraphics[width=\columnwidth]{wedge-view-high-res}
+  \caption{3-D section view of the 3-D numerical \gls{MSRE} model offset by 5.08 cm in the $x$ and
+  $y$ axes to show the three control rod thimbles. Rod 1 is inserted by 4.4 inches while Rods 2 \&
+  3 are fully withdrawn.}
+  \label{fig:wedge-view}
 \end{figure}
@@
   \centering
-  \caption{List of reactor dimensions.}
+  \caption{List of key reactor dimensions.}
   \begin{tabular}{l S}
@@
     \midrule
-    Height [m] & 2.15 \\
-    Radius [m] & 0.762 \\
-    Vessel thickness [m] & 0.0254 \\
-    Fuel channel length [m] & 0.03048 \\
-    Fuel channel width [m] & 0.01016 \\
-    Graphite lattice pitch [m] & 0.05080 \\
+    Total height [cm] & 215.5967 \\
+    Lower plenum height [cm] & 18.7530 \\
+    Upper plenum height [cm] & 25.3937 \\
+    Total radius [cm] & 76.200 \\
+    Graphite core radius [cm] & 70.168 \\
+    Core can inner radius [cm] & 70.485 \\
+    Core can outer radius [cm] & 71.120 \\
+    Vessel thickness [cm] & 2.540 \\
+    Fuel channel length [cm] & 3.048 \\
+    Fuel channel width [cm] & 1.016 \\
+    Graphite lattice pitch [cm] & 5.080 \\
+    Control rod inner radius [cm] & 1.0668 \\
+    Control rod outer radius [cm] & 1.3716 \\
+    Control rod thimble inner radius [cm] & 2.3749 \\
+    Control rod thimble outer radius [cm] & 2.540 \\
+    Homogenized sample basket radius [cm] & 2.8575 \\
+    Radius of fuel annulus around thimble [cm] & 3.2016 \\
     \bottomrule
@@
 
+\begin{table}[htb]
+  \small
+  \centering
+  \setlength\tabcolsep{4pt}
+  \caption{Axial discretization of the 3-D \gls{MSRE} model mesh.}
+  \begin{tabular}{l S S S S}
+    \toprule
+    Layer & {Layer height [cm]} & {Layer start height [cm]} & {Layer end height [cm]} & {Layer mesh resolution} \\
+    \midrule
+    1 & 2.5400 & 0.0000 & 2.5400 & 1 \\
+    2 & 18.7530 & 2.5400 & 21.2930 & 4 \\
+    3 & 5.0800 & 21.2930 & 26.3730 & 1 \\
+    4 & 14.6050 & 26.3730 & 40.9780 & 3 \\
+    5 & 139.7000 & 40.9780 & 180.6780 & 28 \\
+    6 & 6.9850 & 180.6780 & 187.6630 & 2 \\
+    7 & 25.3937 & 187.6630 & 213.0567 & 5 \\
+    8 & 2.5400 & 213.0567 & 215.5967 & 1 \\
+    \bottomrule
+  \end{tabular}
+  \label{table:axial-mesh}
+\end{table}
+
 \subsection{3-D Numerical Results \& Discussion} \label{sec:3d-nts}
@@
 the Serpent 2 model from the numerical benchmark.
-Figures * show the neutron group fluxes on the interior of the 3-D models. The salt-graphite
-lattice structure is visible from the group flux distributions. Group 7 and 8 neutron fluxes are
-particularly small near the control rods due to strong thermal neutron capture cross section of
-gadolinium.
+
+Figures * show a 3-D section view of the \gls{MSRE} model geometry and the group 1 \& 8 neutron
+flux distributions. The salt-graphite lattice structure is visible from the group flux
+distributions. The group 1 flux distribution largely retains a spherical shape. The group 8 neutron
+flux distribution is particularly suppressed near the control rods due to strong thermal neutron
+capture cross section of gadolinium.
+
+\begin{figure}[p]
+  \centering
+  \includegraphics[width=\columnwidth]{3d-view-high-res}
+  \caption{3-D section view of the \gls{MSRE} model geometry with Rod 1 inserted by 4.4 inches at
+  initial criticality.}
+  \label{fig:3d-view}
+\end{figure}
+
+\begin{figure}[p]
+  \centering
+  \includegraphics[width=\columnwidth]{ic-g1}
+  \caption{}
+  \label{fig:ic-g1}
+\end{figure}
+
+\begin{figure}[p]
+  \centering
+  \includegraphics[width=\columnwidth]{ic-g8}
+  \caption{}
+  \label{fig:ic-g8}
+\end{figure}
 
@@
 
-Table \ref{table:temp-coef} compares the isothermal reactivity coefficients from the experimental
-and numerical sources. The OpenMC and Moltres models in this work show closer agreement with the
-\gls{MSRE} data than the Serpent 2 model from the \gls{MSRE} numerical benchmark. All numerical
-results fall within one standard deviation of the \gls{MSRE} data.
-
-\begin{table}[t]
+\begin{table}[h]
   \small
@@
 \end{table}
+
+Table \ref{table:temp-coef} compares the isothermal reactivity coefficients from the experimental
+and numerical sources. The OpenMC and Moltres models in this work show closer agreement with the
+\gls{MSRE} data than the Serpent 2 model from the \gls{MSRE} numerical benchmark. All numerical
+results fall within one standard deviation of the \gls{MSRE} data.
 
@@
 
+\begin{figure}[t]
+  \centering
+  \begin{subfigure}[b]{0.49\columnwidth}
+    \centering
+    \includegraphics[width=\columnwidth]{scaling-total}
+    \caption{Total wall time}
+  \end{subfigure} \\
+  \begin{subfigure}[b]{0.49\columnwidth}
+    \centering
+    \includegraphics[width=\columnwidth]{scaling-solver}
+    \caption{Solver wall time}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.49\columnwidth}
+    \centering
+    \includegraphics[width=\columnwidth]{scaling-transfer}
+    \caption{Transfer wall time}
+  \end{subfigure}
+  \caption{The total, solver, and transfer wall time of hybrid method simulations of the 3-D
+  quarter-core model on 10, 20, 40, and 80 compute nodes (32 processors per node) of the Polaris
+  supercomputer. The solver
+  time scales almost linearly ($q=0.8$) with the number of processors, while the transfer time
+  scales poorly with an increase in wall time at 80 nodes. All axes are in log scale.}
+  \label{fig:scaling}
+\end{figure}
+
+3-D simulations generally required more nonlinear iterations than 2-D simulations for the neutron
+diffusion subsolver to
+converge. This is likely due to increased streaming effects in 3-D air regions and the
+consequent poor performance of the Hypre-BoomerAMG preconditioning method. The performance of this
+preconditioner worsens with strong advective effects from the drift correction term. Thus,
+alternative preconditioning methods suited for advection-dominated problems may help accelerate
+the hybrid method.
+
 For the hybrid method to be viable for large time-dependent simulations, it must scale reasonably
@@
 
-\begin{figure}[t]
-  \centering
-  \begin{subfigure}[b]{0.49\columnwidth}
-    \centering
-    \includegraphics[width=\columnwidth]{scaling-total}
-    \caption{Total wall time}
-  \end{subfigure} \\
-  \begin{subfigure}[b]{0.49\columnwidth}
-    \centering
-    \includegraphics[width=\columnwidth]{scaling-solver}
-    \caption{Solver wall time}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.49\columnwidth}
-    \centering
-    \includegraphics[width=\columnwidth]{scaling-transfer}
-    \caption{Transfer wall time}
-  \end{subfigure}
-  \caption{The total, solver, and transfer wall time of hybrid method simulations of the 3-D
-  quarter-core model on 10, 20, 40, and 80 compute nodes (32 processors per node) of the Polaris
-  supercomputer. The solver
-  time scales almost linearly ($q=0.8$) with the number of processors, while the transfer time
-  scales poorly with an increase in wall time at 80 nodes. All axes are in log scale.}
-  \label{fig:scaling}
-\end{figure}
+\FloatBarrier
 

### 2024-12-09 transient.tex
Draft for committee before defense

--- 2024-11-23 transient.tex
@@
 Section \ref{sec:rod-drop} covers the model setup, numerical results, and discussion of the rod
-drop experiment. Section \ref{sec:reactivity-insertion} covers the model setup, numerical results,
-and discussion of the reactivity insertion experiment. Lastly, Section \ref{sec:transient-summary}
-summarizes the goals and outcomes in this chapter.
+drop experiment.
+%Section \ref{sec:reactivity-insertion} covers the model setup, numerical results,
+%and discussion of the reactivity insertion experiment.
+Section \ref{sec:transient-summary}
+summarizes the findings from the rod drop experiment study.
 
@@
 $^{235}$U loading corresponding to ``30 capsule additions'' of highly-enriched $^{235}$U capsules
-into the ``pump bowl''. Each capsule contains 85 g of $^{235}$U, of which 94.94 \% enters the core
+into the \gls{MSRE} through the ``pump bowl''. Each capsule contains 85 g of $^{235}$U, of which
+94.94 \% enters the core
 due to residual fuel salt remaining in the drain tanks after mixing \cite{fratoni_molten_2020}.
@@
 
-\begin{figure}[htb!]
+\begin{figure}[t]
+  \tikzstyle{every node}=[font=\small]
   \centering
-%  \includegraphics[width=0.8\columnwidth]{rod-height}
+  \begin{tikzpicture}
+    \node (1) [solver] {\textbf{$\bm{S_N}$ transport subsolver}};
+    \node (2) [solver, right of=1, xshift=4cm] {\textbf{Neutron\\ diffusion\\ subsolver}};
+    \node (3) [solver, right of=2, xshift=4cm] {\textbf{DNP subsolver}};
+    \draw [arrow] (1.10) -- node[anchor=south, text width=4cm, align=center] {$\vec{D}_g, \gamma_g$} (2.170);
+    \draw [arrow] (2.190) -- node[anchor=north, text width=4cm, align=center] {$\phi_g, C_i$} (1.350);
+    \draw [arrow] (2.10) -- node[anchor=south, text width=4cm, align=center] {$Q_f, Q$} (3.170);
+    \draw [arrow] (3.190) -- node[anchor=north, text width=4cm, align=center] {$C_i$} (2.350);
+    \node (5) [bound, fit=(2) (3), inner sep=8pt, label={[align=left]above:\textbf{Inner iteration}}] {};
+    \node (6) [bound, fit=(1) (5), inner sep=15pt, label={[align=left]above:\textbf{Outer iteration}}] {};
+  \end{tikzpicture}
   \caption{The nested iteration structure coupling the $S_N$, neutron diffusion, and \gls{DNP}
@@
 
+Figure \ref{fig:count-rate} shows the neutron count rate following the rod drop from Moltres. It
+shows the steep initial decline in neutron count rate as the rod drops further into the core. The
+decline slows at around $t=0.5$ s due to the delayed response of \glspl{DNP}. During this
+time-dependent simulation, convergence issues prevented it from converging beyond $t=0.825$ s. The
+simulation managed to continue after raising the convergence tolerance value of the neutron
+diffusion subsolver and the fixed point iterations from $\epsilon_\text{tol}=10^{-8}$ to $10^{-7}$. However,
+this action led to under-converged scalar flux and neutron count results between $t=1.325$ s and
+$3.825$ s.
+
 Figure \ref{fig:integral-count} compares the integral neutron count during the rod drop experiment
@@
 neutron count characterized by the drop in neutron count rate leading to a gradual flattening of
-the integral count curve. The discrepancy in integral neutron count at around $t=16$ s is x \%.
+the integral count curve. Moltres underpredicts the integral neutron count throughout the
+simulation compared to the \gls{MSRE} experimental data.
 Experimental uncertainties surrounding the rod drop experiment include the actual $^{235}$U
-loading, initial rod height, rod acceleration, and the initial neutron count rate. These uncertainties
-may have affected the reactivity withdrawal estimates in Section \ref{sec:rod-drop-setup}.
+concentration, initial \& final rod heights, rod acceleration, and the initial neutron count rate.
+These uncertainties may have affected the reactivity withdrawal estimates in Section
+\ref{sec:rod-drop-setup}.
 
-\begin{figure}[t]
+Considering that the \gls{DNP} subsolver convergence remained unaffected throughout the simulation,
+it would have produced \gls{DNP} distributions that are fairly accurate. Only minor discrepancies
+may be present in the \gls{DNP} distributions through \gls{DNP} production from under-converged
+scalar fluxes. Figure \ref{fig:count-rate-interp} shows the neutron count rate if we replaced the
+under-converged data points between $t=1.325$ s and $3.825$ s with linearly interpolated count
+rates from $t=0.825$ s to $4.325$ s. Figure \ref{fig:integral-count-interp} shows the corresponding
+integral neutron count curve generated using the interpolated data. In this case, the integral
+count curve shape is more consistent with the experimental data, albeit seemingly in response to a
+larger reactivity withdrawal.
+
+\begin{figure}[p]
   \centering
+  \includegraphics[width=0.8\columnwidth]{count-rate}
+  \caption{Neutron count rate during the rod drop experiment from Moltres. The convergence tolerance
+  of the neutron diffusion subsolver and the fixed point iterations was raised from $\epsilon=10^{-8}$
+  to $10^{-7}$ from $t=1.32$ s onwards.}
+  \label{fig:count-rate}
   \includegraphics[width=0.8\columnwidth]{integral-count}
   \caption{Integral neutron count during the rod drop experiment from \gls{MSRE} experimental data
-  and hybrid method numerical results. The hybrid method agrees closely with the \gls{MSRE} data
-  with a discrepancy of x \% at around $t=16$ s.}
+  and Moltres.}
   \label{fig:integral-count}
@@
 
-Figures * show snapshots of the neutron group flux distributions at various times during the rod
-drop simulation. The neutron flux magnitudes fall as the reactor enters a sub-critical state. The
-Rod 1 induces a significant drop in neutron fluxes near the center of the core.
-
-\section{Reactivity Insertion Experiment} \label{sec:reactivity-insertion}
-
-After a long operational run on $^{235}$U fuel, \gls{ORNL} extended the \gls{MSRE} program to
-demonstrate reactor operation with $^{233}$U fuel. The \gls{MSRE} then became the first nuclear
-reactor to run on $^{233}$U bred from $^{232}$Th. During a series of tests for experimental
-dynamic analyses, \gls{MSRE} researchers conducted reactivity insertion tests with the reactor
-operating at 1, 5, and 8 MW \cite{steffy_experimental_1970}.
-In contrast to the rod drop experiments, these reactivity insertion
-experiments were conducted with the \gls{MSRE} in power operation.
-
-This work aims to replicate the power response following a reactivity insertion of 24.8 pcm of the
-\gls{MSRE} initially at 1 MW. Figure \ref{fig:msre-reactivity-insertion} shows the theoretical and
-experimental power response following the reactivity insertion. Three separate publications
-\cite{zanetti_innovative_2015, singh_study_2019, mochizuki_validation_2022} noted that the plot is
-mislabeled with the wrong reactivity insertion value of 0.0139 \% $\delta k/k$, which had been the
-reactivity insertion value for the 8 MW case. The correct reactivity insertion value is 24.8 pcm as
-mentioned prior.
-
-\begin{figure}[t]
+\begin{figure}[p]
   \centering
-  \includegraphics[width=0.8\columnwidth]{msre-reactivity-insertion}
-  \caption{Power response following a step reactivity insertion of 24.8 pcm with the reactor
-    initially at 1 MW \cite{steffy_experimental_1970}. Note that the authors of the \gls{MSRE}
-    report mislabeled the reactivity insertion value on the plot
-  \cite{zanetti_innovative_2015, singh_study_2019, mochizuki_validation_2022}.}
-  \label{fig:msre-reactivity-insertion}
+  \includegraphics[width=0.8\columnwidth]{count-rate-interp}
+  \caption{Neutron count rate during the rod drop experiment from Moltres with linearly interpolated
+    data between $t=1.325$ s and $3.825$ s.}
+  \label{fig:count-rate-interp}
+  \includegraphics[width=0.8\columnwidth]{integral-count-interp}
+  \caption{Integral neutron count during the rod drop experiment from \gls{MSRE} experimental data
+  and Moltres with linearly interpolated count rate data between $t=1.325$ s and $3.825$ s.}
+  \label{fig:integral-count-interp}
 \end{figure}
 
-\subsection{Model Setup}
+Lagging solution states during the first fixed point iteration in each time-step also slowed
+solution convergence, particularly when the control rod was in motion. This limited the simulation
+from taking larger time-steps. Further work could include more sophiscated coupling schemes to
+help the solver converge faster.
 
-For the reactivity insertion simulation, we must first identify the $^{233}$U salt composition used
-in the experiment. Power operations with the $^{233}$U-fueled \gls{MSRE} core
-occurred during Runs 16 to 18 according
-to \gls{MSRE} reports \cite{thoma_chemical_1971, burke_analysis_2019}. Table \ref{table:u233-salt}
-lists the proportion of elements in the \gls{MSRE} fuel salt during Runs 16 to 18. Note that the
-components do not sum up to 100 \% due to experimental uncertainty in chemical analysis and the
-presence of other elements that were not measured. Taking the nominal uranium value to be true, I scaled
-up the compositions of non-uranium elements to make up the remaining $(100\%-0.8155\%)$ of the salt. I
-kept the nominal weight percentage of uranium constant because it is the most accurate value in the table
-and has the biggest impact on the reactivity.
+%Figures * show snapshots of the neutron group flux distributions at various times during the rod
+%drop simulation. The neutron flux magnitudes fall as the reactor enters a sub-critical state. The
+%Rod 1 induces a significant drop in neutron fluxes near the center of the core.
+
+%\section{Reactivity Insertion Experiment} \label{sec:reactivity-insertion}
 %
-\begin{table}[t]
-  \small
-  \centering
-  \setlength\tabcolsep{4pt}
-  \caption{\gls{MSRE} fuel salt mean composition during the power operations with $^{233}$U \cite{thoma_chemical_1971}.
-  All values were measured via chemical analysis except the nominal U weight percentage, which was
-  calculated from inventory tracking of cumulative U additions to the salt.}
-  \begin{tabular}{l S[table-format=2.2(2)] S[table-format=1.2(2)] S[table-format=2.2(2)] S[table-format=2.2(3)] S[table-format=1.3(2)] S}
-    \toprule
-    Element & {Li} & {Be} & {Zr} & {F} & {U} & {U (Nominal)} \\
-    \midrule
-    Weight percentage [\%]  & 11.33(39) & 6.71(30) & 10.99(29) & 68.71(126) & 0.809(11) & 0.8155 \\
-    \bottomrule
-  \end{tabular}
-  \begin{tabular}{l S[table-format=2(1)] S[table-format=3(2)] S[table-format=2(2)] S}
-    \toprule
-    Element & {Cr} & {Fe} & {Ni} & {O} \\
-    \midrule
-    Parts per million [ppm] & 72(9) & 135(24) & 46(12) & 61 \\
-    \bottomrule
-  \end{tabular}
-  \label{table:u233-salt}
-\end{table}
+%After a long operational run on $^{235}$U fuel, \gls{ORNL} extended the \gls{MSRE} program to
+%demonstrate reactor operation with $^{233}$U fuel. The \gls{MSRE} then became the first nuclear
+%reactor to run on $^{233}$U bred from $^{232}$Th. During a series of tests for experimental
+%dynamic analyses, \gls{MSRE} researchers conducted reactivity insertion tests with the reactor
+%operating at 1, 5, and 8 MW \cite{steffy_experimental_1970}.
+%In contrast to the rod drop experiments, these reactivity insertion
+%experiments were conducted with the \gls{MSRE} in power operation.
+%
+%This work aims to replicate the power response following a reactivity insertion of 24.8 pcm of the
+%\gls{MSRE} initially at 1 MW. Figure \ref{fig:msre-reactivity-insertion} shows the theoretical and
+%experimental power response following the reactivity insertion. Three separate publications
+%\cite{zanetti_innovative_2015, singh_study_2019, mochizuki_validation_2022} noted that the plot is
+%mislabeled with the wrong reactivity insertion value of 0.0139 \% $\delta k/k$, which had been the
+%reactivity insertion value for the 8 MW case. The correct reactivity insertion value is 24.8 pcm as
+%mentioned prior.
+%
+%\begin{figure}[t]
+%  \centering
+%  \includegraphics[width=0.8\columnwidth]{msre-reactivity-insertion}
+%  \caption{Power response following a step reactivity insertion of 24.8 pcm with the reactor
+%    initially at 1 MW \cite{steffy_experimental_1970}. Note that the authors of the \gls{MSRE}
+%    report mislabeled the reactivity insertion value on the plot
+%  \cite{zanetti_innovative_2015, singh_study_2019, mochizuki_validation_2022}.}
+%  \label{fig:msre-reactivity-insertion}
+%\end{figure}
+%
+%\subsection{Model Setup}
+%
+%For the reactivity insertion simulation, we must first identify the $^{233}$U salt composition used
+%in the experiment. Power operations with the $^{233}$U-fueled \gls{MSRE} core
+%occurred during Runs 16 to 18 according
+%to \gls{MSRE} reports \cite{thoma_chemical_1971, burke_analysis_2019}. Table \ref{table:u233-salt}
+%lists the proportion of elements in the \gls{MSRE} fuel salt during Runs 16 to 18. Note that the
+%components do not sum up to 100 \% due to experimental uncertainty in chemical analysis and the
+%presence of other elements that were not measured. Taking the nominal uranium value to be true, I scaled
+%up the compositions of non-uranium elements to make up the remaining $(100\%-0.8155\%)$ of the salt. I
+%kept the nominal weight percentage of uranium constant because it is the most accurate value in the table
+%and has the biggest impact on the reactivity.
+%%
+%\begin{table}[t]
+%  \small
+%  \centering
+%  \setlength\tabcolsep{4pt}
+%  \caption{\gls{MSRE} fuel salt mean composition during the power operations with $^{233}$U \cite{thoma_chemical_1971}.
+%  All values were measured via chemical analysis except the nominal U weight percentage, which was
+%  calculated from inventory tracking of cumulative U additions to the salt.}
+%  \begin{tabular}{l S[table-format=2.2(2)] S[table-format=1.2(2)] S[table-format=2.2(2)] S[table-format=2.2(3)] S[table-format=1.3(2)] S}
+%    \toprule
+%    Element & {Li} & {Be} & {Zr} & {F} & {U} & {U (Nominal)} \\
+%    \midrule
+%    Weight percentage [\%]  & 11.33(39) & 6.71(30) & 10.99(29) & 68.71(126) & 0.809(11) & 0.8155 \\
+%    \bottomrule
+%  \end{tabular}
+%  \begin{tabular}{l S[table-format=2(1)] S[table-format=3(2)] S[table-format=2(2)] S}
+%    \toprule
+%    Element & {Cr} & {Fe} & {Ni} & {O} \\
+%    \midrule
+%    Parts per million [ppm] & 72(9) & 135(24) & 46(12) & 61 \\
+%    \bottomrule
+%  \end{tabular}
+%  \label{table:u233-salt}
+%\end{table}
+%
+%For the isotopic content of uranium, I used a composition restruction strategy developed by Burke
+%\cite{burke_analysis_2019}. Their analysis determined that the $^{233}$U salt consists of residual
+%salt from $^{235}$U operations, a depleted uranium feed, and a stock $^{233}$U feed. The residue
+%and depleted uranium feed masses, taken from Burke's work, are 1.935 kg and 0.890 kg, respectively.
+%The stock feed uranium
+%mass is estimated to be 35.234 kg reported on January 12 1969 \cite{thoma_chemical_1971}. The
+%experiment was performed on January 16 1969 \cite{steffy_experimental_1970}. Table 
+%\ref{table:u233-composition} shows
+%the breakdown and total isotopic content of uranium in the salt. The density of $^{233}$U salt is
+%2.264 g cm$^{-3}$ at 911 K \cite{thoma_chemical_1971}.
+%%
+%\begin{table}[t]
+%  \small
+%  \centering
+%  \setlength\tabcolsep{4pt}
+%  \caption{Uranium isotopic composition reconstruction table \cite{burke_analysis_2019}.}
+%  \begin{tabular}{l S S S S S S S}
+%    \toprule
+%    \multirow{2}{*}{Uranium Source} & {\multirow{2}{*}{Mass [kg]}} & \multicolumn{6}{c}{Weight percentage [\%]} \\
+%                                    & & {$^{232}$U} & {$^{233}$U} & {$^{234}$U} & {$^{235}$U} & {$^{236}$U} & {$^{238}$U} \\
+%                                    \cmidrule(r){1-1} \cmidrule(rl){2-2} \cmidrule(l){3-8}
+%    Residue & 1.935 & & & 0.343 & 32.705 & 0.483 & 66.469 \\
+%    Depleted & 0.890 & & & & 0.3 & & 99.7 \\
+%    Stock & 35.234 & 0.02 & 91.49 & 7.6 & 0.7 & 0.05 & 0.14 \\
+%    \bottomrule
+%  \end{tabular}
+%  \label{table:u233-composition}
+%\end{table}
+%
+%Due to constraints on computational resources, the reactivity insertion simulation ran a 3-D
+%quarter-core \gls{MSRE} model. Refer to Figure \ref{fig:1/4-geom} for the horizontal cross section
+%view of the 3-D quarter-core model. Figure * shows the side view from a vertical side of the
+%quarter-core and the direction of salt flow in each major region. The regions are the salt-graphite
+%lattice, the upper plenum, the out-of-core components, the downcomer, and the lower plenum. The
+%model assumes a flat salt velocity profile in each region. The actual total volumetric flow rate of
+%the \gls{MSRE} is 1200 gpm (gallons per minute) \cite{robertson_msre_1965}. Dividing this value by
+%four for the quarter-core model and converting to SI units gives a volumetric flow rate of
+%0.01893 m$^3$ s$^{-1}$.
+%
+%Table \ref{table:msre-flow} lists the dimensions, salt flow velocities, and salt residence
+%times in each region to maintain the calculated volumetric flow rate. The salt residence times are
+%consistent with \gls{MSRE} specifications \cite{robertson_msre_1965}. The out-of-core region is
+%modeled as a 1-D pipe due to its lower neutronic importance other than \gls{DNP} decay and heat
+%removal through the heat exchanger. The out-of-core region salt velocity is set to the same magnitude
+%as the downcomer for ease of implementation with the height adjusted accordingly to maintain the
+%out-of-core salt residence time. Note that the velocity in the downcomer is negative since salt
+%flows downwards unlike in the other regions.
+%%
+%\begin{table}[t]
+%  \small
+%  \centering
+%  \caption{Dimensions, salt flow velocities, and salt residence times in each region of the \gls{MSRE}.}
+%  \begin{tabular}{l S S S S}
+%    \toprule
+%    Component & {Salt cross-sectional area [m$^2$]} & {Height [m]} & {Velocity [m s$^{-1}$]} & {Time [s]} \\
+%    \cmidrule(r){1-1} \cmidrule(l){2-5}
+%    Upper plenum & 0.42411 & 0.2539 & 0.04463 & 5.69 \\
+%    Lattice      & 0.08646 & 1.6637 & 0.21891 & 7.60 \\
+%    Lower plenum & 0.42614 & 0.1875 & 0.04935 & 3.80 \\
+%    Out-of-core  & {-}     & 5.3081 & 0.65532 & 8.1 \\
+%    Downcomer    & 0.02888 & 1.6637 & -0.65532 & 2.539 \\
+%    \bottomrule
+%  \end{tabular}
+%  \label{table:msre-flow}
+%\end{table}
+%
+%% Unlike with $^{235}$U salt, there is no widely recognized benchmark value
+%% for the density of $^{233}$U salt. During \gls{MSRE} operation, researchers were still developing
+%% reliable techniques for measuring the density of molten salt mixtures \cite{thoma_chemical_1971}.
+%% \gls{MSRE} $^{233}$U fuel salt with has a lower density than $^{235}$U salt largely due to its
+%% lower $^{238}$U content. 
+%
+%The MultiApp nested iterative coupling structure for this reactivity insertion simulation is similar
+%to the implementation for the rod drop simulation. Figure \ref{fig:insertion-coupling}
+%provides an updated illustration of the nexted iteration coupling with temperature. The \gls{DNP}
+%and temperature distributions are solved together with additive Schwarz method preconditioning in
+%Moltres since they are both advective problems. The 1-D out-of-core model accounts for \gls{DNP}
+%loss and heat removal outside the core. Delayed neutrons produced in this region are considered
+%lost and do not contribute to the neutron flux in the core.
+%
+%\begin{figure}[t]
+%  \tikzstyle{every node}=[font=\small]
+%  \centering
+%  \begin{tikzpicture}
+%    \node (1) [solver] {\textbf{$\bm{S_N}$ transport subsolver}};
+%    \node (2) [solver, right of=1, xshift=4cm] {\textbf{Neutron\\ diffusion\\ subsolver}};
+%    \node (3) [solver, right of=2, xshift=4cm] {\textbf{In-core DNP \& temperature subsolver}};
+%    \node (4) [solver, below of=3, yshift=-3cm] {\textbf{Ex-core DNP \& temperature subsolver}};
+%    \draw [arrow] (1.10) -- node[anchor=south, text width=4cm, align=center] {$\vec{D}_g, \gamma_g$} (2.170);
+%    \draw [arrow] (2.190) -- node[anchor=north, text width=4cm, align=center] {$\phi_g, C_i, T$} (1.350);
+%    \draw [arrow] (2.10) -- node[anchor=south, text width=4cm, align=center] {$Q_f, Q$} (3.170);
+%    \draw [arrow] (3.190) -- node[anchor=north, text width=4cm, align=center] {$C_i, T$} (2.350);
+%    \draw [arrow] (3.280) -- node[anchor=west, text width=1cm, align=center] {$C^\text{out}_i$\\ $T^\text{out}$} (4.80);
+%    \draw [arrow] (4.100) -- node[anchor=east, text width=1cm, align=center] {$C^\text{in}_i$\\ $T^\text{in}$} (3.260);
+%    \node (5) [bound, fit=(3) (4), label={[align=center,rotate=90]left:\textbf{Salt loop}\\}] {};
+%    \node (6) [bound, fit=(2) (5), label={[align=center,rotate=90]left:\textbf{Inner iteration}\\}] {};
+%    \node (7) [bound, fit=(1) (6), label={[align=center,rotate=90]left:\textbf{Outer iteration}\\}] {};
+%  \end{tikzpicture}
+%  \caption{The nested iteration structure coupling the $S_N$, neutron diffusion,
+%  in-core \gls{DNP}-temperature, and ex-core
+%  solvers for the reactivity insertion simulation using the hybrid $S_N$-diffusion method.}
+%  \label{fig:insertion-coupling}
+%\end{figure}
+%
+%The reactivity insertion simulation requires a prior ``steady-state'' simulation to obtain the group
+%flux, \gls{DNP}, and temperature distributions under 1 MW power operation and steady salt flow. The
+%``steady-state'' simulation is a time-dependent simulation with timestep sizes of 1 s to advance
+%the model towards 1 MW steady state operation. Following this, the reactivity insertion simulation
+%uses the \gls{MOOSE} Restart capability to load in variable distributions for the initial conditions.
+%
+%\subsection{Preliminary Rod Worth Calculations}
+%
+%The control rod worth of the \gls{MSRE} quarter-core model with $^{233}$U fuel must be determined
+%to estimate the initial rod height and rod travel necessary for the reactivity insertion
+%simulation. We start this process with comparisons of Rod 1 worths between a full-core OpenMC model
+%and experimental data. Table \ref{table:u233-rod-worth} compares the total rod worth of Rod 1 at
+%922 K between the OpenMC model from this work
+%and \gls{MSRE} data. The rod worth estimate from OpenMC agrees closely with the predicted (i.e.,
+%calculated) \gls{MSRE} value. The measured \gls{MSRE} rod worth value is about 7 \% smaller than
+%the predicted value. \gls{MSRE} researchers noted poor precision in their rod worth measurements
+%due to ``wear and aging'' in the control rod drive assembly \cite{engel_zero-power_1972}. This may
+%be responsible for some uncertainty. However, the actual rod worth is likely closer to the measured
+%value than the predicted value, and the \gls{MSRE} model of this work is likely to contain as yet
+%unknown biases. Nevertheless, the rod worth estimate is close enough that it should not affect the
+%\gls{MSRE} power response following a fixed reactivity insertion.
+%%
+%\begin{table}[t]
+%  \centering
+%  \caption{Total rod worth of Rod 1 at 922 K from OpenMC full-core \gls{MSRE} model and \gls{MSRE}
+%  experimental data.}
+%  \begin{tabular}{l S[table-format=4(3)] S}
+%    \toprule
+%    & {Rod worth [pcm]} & {Percentage discrepancy [\%]} \\
+%    \cmidrule(r){1-1} \cmidrule(l){2-3}
+%    MSRE (measured) & 2580(129) & {-} \\
+%    MSRE (predicted) & 2750 & 6.6 \\
+%    OpenMC (this work) & 2771 & 7.4\\
+%    \bottomrule
+%  \end{tabular}
+%  \label{table:u233-rod-worth}
+%\end{table}
+%
+%The 3-D quarter-core \gls{MSRE} model for the reactivity insertion simulation contains only one
+%control rod, but it is equivalent to four rods due to rotational symmetry of the quarter-core model.
+%The total rod worth of the 3-D quarter-core \gls{MSRE} model using the hybrid method on Moltres is
+%9702 pcm, just under four times the worth of Rod 1 from Table \ref{table:u233-rod-worth} due to
+%rod shadowing effects that reduce the worth of adjacent control rods. Figure \ref{fig:u233-rod-worth-curve} shows the rod
+%worth values from Moltres for the 3-D quarter-core model and the monotone cubic spline interpolation
+%curve. Despite having fewer data points than the $^{235}$U rod worth analysis (Section
+%\ref{sec:3d-nts}), this interpolation curve is valid for the reactivity insertion experiment since
+%the rod heights will strictly remain within the middle
+%range where the change in rod worth is approximately linear to changes in rod height.
+%
+%The required
+%rod travel in the quarter-core model for a 24.8 pcm reactivity insertion in this linear region
+%is 0.105 inches (0.267 cm). The estimated rod travel required to induce the same reactivity
+%insertion in the actual \gls{MSRE} reactor is 0.396 inches (1.00 cm) obtained by scaling with the
+%total measured \gls{MSRE} rod worth in Table \ref{table:u233-rod-worth}. Using
+%an estimated rod speed of 1.27 cm s$^{-1}$ \cite{robertson_msre_1965}, the reactivity insertion
+%occurs over a 0.791-s ramp (reactivity insertions at this timescale are considered step reactivity
+%insertions). The rod is assumed to accelerate and decelerate instantaneously given
+%the low rod speed.
+%The reactivity insertion simulation ran with timestep sizes of 0.198 s during rod movement and
+%0.25 s after the rod reaches its final position. These timestep sizes were chosen based on
+%considerations balancing accuracy and available computational resources. The main purpose of this
+%work is the demonstration of the hybrid method for time-dependent control rod modeling; a detailed
+%time-stepping truncation error analysis is out of the scope of this work.
+%
+%\begin{figure}[t]
+%  \centering
+%  \includegraphics[width=0.8\columnwidth]{u233-rod-worth-curve}
+%  \caption{Rod worth data points and the monotone cubic spline interpolation curve from $k$-eigenvalue
+%  hybrid method simulations on the 3-D quarter-core \gls{MSRE} model.}
+%  \label{fig:u233-rod-worth-curve}
+%\end{figure}
+%
+%\subsection{Temperature Reactivity Coefficient}
+%
+%This work compares isothermal temperature reactivity coefficients calculated from $k_\text{eff}$
+%values measured at 872 K and 972 K with the $^{233}$-fueled \gls{MSRE} at operating fuel
+%concentration. This approach is identical to that
+%of the \gls{MSRE} experimental measurements \cite{engel_zero-power_1972} and the temperature
+%coefficient calculation procedure used for the $^{235}$U-fueled \gls{MSRE} in Section
+%\ref{sec:temp-coef}.
+%
+%Table \ref{table:temp-coef-u233} compares the isothermal reactivity coefficients from the experimental
+%and numerical sources. No uncertainty data is available from the relevant \gls{ORNL} report
+%\cite{engel_zero-power_1972}. However, the authors noted significantly different readings ranging
+%from 12.4 to 15.3 pcm K$^{-1}$ due to gas bubbles introduced by the salt pump. The
+%presence of gas bubbles led to smaller reactivity coefficient measurements. The numerical OpenMC
+%and hybrid models overestimate the temperature reactivity coefficient magnitude by 12 \% and 11 \%,
+%respectively, relative to the experimental data.
+%
+%\begin{table}[t]
+%  \small
+%  \centering
+%  \caption{Isothermal temperature reactivity coefficients with $^{233}$U-fueled \gls{MSRE} from
+%    \gls{MSRE} experimental data, \gls{MSRE} predicted value, and the OpenMC
+%  and Moltres models in this work.}
+%  \begin{tabular}{l S[separate-uncertainty=true,table-format=2.2(2)] S}
+%    \toprule
+%     & {Temperature coefficient [pcm K$^{-1}$]} & {Percentage discrepancy [\%]}\\
+%     \cmidrule(rl){2-2} \cmidrule(l){3-3}
+%    \gls{MSRE} experimental data & -15.3 & {-}\\
+%    \gls{MSRE} predicted value & -15.8 & -3.4 \\
+%    OpenMC (This work) & -17.14 & -12 \\
+%    Hybrid (This work) & -16.96 & -11 \\
+%    \bottomrule
+%  \end{tabular}
+%  \label{table:temp-coef-u233}
+%\end{table}
+%
+%\subsection{Reactivity Insertion Numerical Results \& Discussion}
+%
+%Figure \ref{fig:insertion-power} shows the \gls{MSRE} data and simulation results for change in
+%power output following the 24.8 pcm reactivity
+%insertion. From the experimental data, the power output rises rapidly at the start of the
+%reactivity insertion before oscillating back down below the initial power output. Power output
+%peaks again later at around $t=360$ s. Moltres reproduces the same trend in power output, but 
+%over/under-predicts the first power peak by y \%. It also over/under-predicts the subsequent power
+%oscillations.
+%Significant contributors to the discrepancies are the uncertainties in the physical properties of
+%the fuel salt, temperature reactivity coefficient, and the dynamic response of the heat exchanger.
+%
+%Figure \ref{fig:insertion-temperature} shows the maximum fuel salt and graphite temperatures, and
+%the mean salt inlet and outlet
+%temperatures following the reactivity insertion. The increase in maximum temperatures is more
+%gradual than the increase in power output as the salt and graphite take time to heat up. The salt
+%outlet temperature follows closely behind as the heated salt flows out of the core. The salt inlet
+%temperature lags behind the other temperature measurements due to the time taken for the heated
+%salt to travel through the out-of-core region. The temperature values appear to settle at higher
+%temperatures to compensate for the inserted reactivity.
+%
+%\begin{figure}[t]
+%  \centering
+%%  \includegraphics[width=0.8\columnwidth]{u233-rod-worth-curve}
+%  \caption{Change in power output following a step reactivity insertion of 24.8 pcm with the
+%  \gls{MSRE} initially at 1 MW.}
+%  \label{fig:insertion-power}
+%\end{figure}
+%
+%\begin{figure}[t]
+%  \centering
+%%  \includegraphics[width=0.8\columnwidth]{u233-rod-worth-curve}
+%  \caption{Change in maximum fuel salt and graphite temperatures, and the mean salt inlet and outlet
+%  temperatures following a step reactivity insertion of 24.8 pcm with the \gls{MSRE} initially at 1
+%  MW.}
+%  \label{fig:insertion-temperature}
+%\end{figure}
 
-For the isotopic content of uranium, I used a composition restruction strategy developed by Burke
-\cite{burke_analysis_2019}. Their analysis determined that the $^{233}$U salt consists of residual
-salt from $^{235}$U operations, a depleted uranium feed, and a stock $^{233}$U feed. The residue
-and depleted uranium feed masses, taken from Burke's work, are 1.935 kg and 0.890 kg, respectively.
-The stock feed uranium
-mass is estimated to be 35.234 kg reported on January 12 1969 \cite{thoma_chemical_1971}. The
-experiment was performed on January 16 1969 \cite{steffy_experimental_1970}. Table * shows
-the breakdown and total isotopic content of uranium in the salt. The density of $^{233}$U salt is
-2.264 g cm$^{-3}$ at 911 K \cite{thoma_chemical_1971}.
-%
-\begin{table}[t]
-  \small
-  \centering
-  \setlength\tabcolsep{4pt}
-  \caption{Uranium isotopic composition reconstruction table \cite{burke_analysis_2019}.}
-  \begin{tabular}{l S S S S S S S}
-    \toprule
-    \multirow{2}{*}{Uranium Source} & {\multirow{2}{*}{Mass [kg]}} & \multicolumn{6}{c}{Weight percentage [\%]} \\
-                                    & & {$^{232}$U} & {$^{233}$U} & {$^{234}$U} & {$^{235}$U} & {$^{236}$U} & {$^{238}$U} \\
-                                    \cmidrule(r){1-1} \cmidrule(rl){2-2} \cmidrule(l){3-8}
-    Residue & 1.935 & & & 0.343 & 32.705 & 0.483 & 66.469 \\
-    Depleted & 0.890 & & & & 0.3 & & 99.7 \\
-    Stock & 35.234 & 0.02 & 91.49 & 7.6 & 0.7 & 0.05 & 0.14 \\
-    \bottomrule
-  \end{tabular}
-  \label{table:u233-composition}
-\end{table}
-
-Due to constraints on computational resources, the reactivity insertion simulation ran a 3-D
-quarter-core \gls{MSRE} model. Refer to Figure \ref{fig:1/4-geom} for the horizontal cross section
-view of the 3-D quarter-core model. Figure * shows the side view from a vertical side of the
-quarter-core and the direction of salt flow in each major region. The regions are the salt-graphite
-lattice, the upper plenum, the out-of-core components, the downcomer, and the lower plenum. The
-model assumes a flat salt velocity profile in each region. The actual total volumetric flow rate of
-the \gls{MSRE} is 1200 gpm (gallons per minute) \cite{robertson_msre_1965}. Dividing this value by
-four for the quarter-core model and converting to SI units gives a volumetric flow rate of
-0.01893 m$^3$ s$^{-1}$.
-
-Table \ref{table:msre-flow} lists the dimensions, salt flow velocities, and salt residence
-times in each region to maintain the calculated volumetric flow rate. The salt residence times are
-consistent with \gls{MSRE} specifications \cite{robertson_msre_1965}. The out-of-core region is
-modeled as a 1-D pipe due to its lower neutronic importance other than \gls{DNP} decay and heat
-removal through the heat exchanger. The out-of-core region salt velocity is set to the same magnitude
-as the downcomer for ease of implementation with the height adjusted accordingly to maintain the
-out-of-core salt residence time. Note that the velocity in the downcomer is negative since salt
-flows downwards unlike in the other regions.
-%
-\begin{table}[t]
-  \small
-  \centering
-  \caption{Dimensions, salt flow velocities, and salt residence times in each region of the \gls{MSRE}.}
-  \begin{tabular}{l S S S S}
-    \toprule
-    Component & {Salt cross-sectional area [m$^2$]} & {Height [m]} & {Velocity [m s$^{-1}$]} & {Time [s]} \\
-    \cmidrule(r){1-1} \cmidrule(l){2-5}
-    Upper plenum & 0.42411 & 0.2539 & 0.04463 & 5.69 \\
-    Lattice      & 0.08646 & 1.6637 & 0.21891 & 7.60 \\
-    Lower plenum & 0.42614 & 0.1875 & 0.04935 & 3.80 \\
-    Out-of-core  & {-}     & 5.3081 & 0.65532 & 8.1 \\
-    Downcomer    & 0.02888 & 1.6637 & -0.65532 & 2.539 \\
-    \bottomrule
-  \end{tabular}
-  \label{table:msre-flow}
-\end{table}
-
-% Unlike with $^{235}$U salt, there is no widely recognized benchmark value
-% for the density of $^{233}$U salt. During \gls{MSRE} operation, researchers were still developing
-% reliable techniques for measuring the density of molten salt mixtures \cite{thoma_chemical_1971}.
-% \gls{MSRE} $^{233}$U fuel salt with has a lower density than $^{235}$U salt largely due to its
-% lower $^{238}$U content.
-
-The MultiApp nested iterative coupling structure for this reactivity insertion simulation is similar
-to the implementation for the rod drop simulation. Figure \ref{fig:insertion-coupling}
-provides an updated illustration of the nexted iteration coupling with temperature. The \gls{DNP}
-and temperature distributions are solved together with additive Schwarz method preconditioning in
-Moltres since they are both advective problems. The 1-D out-of-core model accounts for \gls{DNP}
-loss and heat removal outside the core. Delayed neutrons produced in this region are considered
-lost and do not contribute to the neutron flux in the core.
-
-\begin{figure}[htb!]
-  \centering
-%  \includegraphics[width=0.8\columnwidth]{rod-height}
-  \caption{The nested iteration structure coupling the $S_N$, neutron diffusion,
-  \gls{DNP}-temperature, and out-of-core
-  solvers for the rod drop simulation using the hybrid $S_N$-diffusion method.}
-  \label{fig:insertion-coupling}
-\end{figure}
-
-The reactivity insertion simulation requires a prior ``steady-state'' simulation to obtain the group
-flux, \gls{DNP}, and temperature distributions under 1 MW power operation and steady salt flow. The
-``steady-state'' simulation is a time-dependent simulation with timestep sizes of 1 s to advance
-the model towards 1 MW steady state operation. Following this, the reactivity insertion simulation
-uses the \gls{MOOSE} Restart capability to load in variable distributions for the initial conditions.
-
-\subsection{Preliminary Rod Worth Calculations}
-
-The control rod worth of the \gls{MSRE} quarter-core model with $^{233}$U fuel must be determined
-to estimate the initial rod height and rod travel necessary for the reactivity insertion
-simulation. We start this process with comparisons of Rod 1 worths between a full-core OpenMC model
-and experimental data. Table \ref{table:u233-rod-worth} compares the total rod worth of Rod 1 at
-922 K between the OpenMC model from this work
-and \gls{MSRE} data. The rod worth estimate from OpenMC agrees closely with the predicted (i.e.,
-calculated) \gls{MSRE} value. The measured \gls{MSRE} rod worth value is about 7 \% smaller than
-the predicted value. \gls{MSRE} researchers noted poor precision in their rod worth measurements
-due to ``wear and aging'' in the control rod drive assembly \cite{engel_zero-power_1972}. This may
-be responsible for some uncertainty. However, the actual rod worth is likely closer to the measured
-value than the predicted value, and the \gls{MSRE} model of this work is likely to contain as yet
-unknown biases. Nevertheless, the rod worth estimate is close enough that it should not affect the
-\gls{MSRE} power response following a fixed reactivity insertion.
-%
-\begin{table}[t]
-  \centering
-  \caption{Total rod worth of Rod 1 at 922 K from OpenMC full-core \gls{MSRE} model and \gls{MSRE}
-  experimental data.}
-  \begin{tabular}{l S[table-format=4(3)] S}
-    \toprule
-    & {Rod worth [pcm]} & {Percentage discrepancy [\%]} \\
-    \cmidrule(r){1-1} \cmidrule(l){2-3}
-    MSRE (measured) & 2580(129) & {-} \\
-    MSRE (predicted) & 2750 & 6.6 \\
-    OpenMC (this work) & 2771 & 7.4\\
-    \bottomrule
-  \end{tabular}
-  \label{table:u233-rod-worth}
-\end{table}
-
-The 3-D quarter-core \gls{MSRE} model for the reactivity insertion simulation contains only one
-control rod, but it is equivalent to four rods due to rotational symmetry of the quarter-core model.
-The total rod worth of the 3-D quarter-core \gls{MSRE} model using the hybrid method on Moltres is
-9702 pcm, just under four times the worth of Rod 1 from Table \ref{table:u233-rod-worth} due to
-rod shadowing effects that reduce the worth of adjacent control rods. Figure \ref{fig:u233-rod-worth-curve} shows the rod
-worth values from Moltres for the 3-D quarter-core model and the monotone cubic spline interpolation
-curve. Despite having fewer data points than the $^{235}$U rod worth analysis (Section
-\ref{sec:3d-nts}), this interpolation curve is valid for the reactivity insertion experiment since
-the rod heights will strictly remain within the middle
-range where the change in rod worth is approximately linear to changes in rod height.
-
-The required
-rod travel in the quarter-core model for a 24.8 pcm reactivity insertion in this linear region
-is 0.105 inches (0.267 cm). The estimated rod travel required to induce the same reactivity
-insertion in the actual \gls{MSRE} reactor is 0.396 inches (1.00 cm) obtained by scaling with the
-total measured \gls{MSRE} rod worth in Table \ref{table:u233-rod-worth}. Using
-an estimated rod speed of 1.27 cm s$^{-1}$ \cite{robertson_msre_1965}, the reactivity insertion
-occurs over a 0.791-s ramp (reactivity insertions at this timescale are considered step reactivity
-insertions). The rod is assumed to accelerate and decelerate instantaneously given
-the low rod speed.
-The reactivity insertion simulation ran with timestep sizes of 0.198 s during rod movement and
-0.25 s after the rod reaches its final position. These timestep sizes were chosen based on
-considerations balancing accuracy and available computational resources. The main purpose of this
-work is the demonstration of the hybrid method for time-dependent control rod modeling; a detailed
-time-stepping truncation error analysis is out of the scope of this work.
-
-\begin{figure}[t]
-  \centering
-  \includegraphics[width=0.8\columnwidth]{u233-rod-worth-curve}
-  \caption{Rod worth data points and the monotone cubic spline interpolation curve from $k$-eigenvalue
-  hybrid method simulations on the 3-D quarter-core \gls{MSRE} model.}
-  \label{fig:u233-rod-worth-curve}
-\end{figure}
-
-\subsection{Reactivity Insertion Numerical Results \& Discussion}
-
-Figure \ref{fig:insertion-power} shows the \gls{MSRE} data and simulation results for change in
-power output following the 24.8 pcm reactivity
-insertion. From the experimental data, the power output rises rapidly at the start of the
-reactivity insertion before oscillating back down below the initial power output. Power output
-peaks again later at around $t=360$ s. Moltres reproduces the same trend in power output, but 
-over/under-predicts the first power peak by y \%. It also over/under-predicts the subsequent power
-oscillations.
-Significant contributors to the discrepancies are the uncertainties in the physical properties of
-the fuel salt, temperature reactivity coefficient, and the dynamic response of the heat exchanger.
-
-Figure \ref{fig:insertion-temperature} shows the maximum fuel salt and graphite temperatures, and
-the mean salt inlet and outlet
-temperatures following the reactivity insertion. The increase in maximum temperatures is more
-gradual than the increase in power output as the salt and graphite take time to heat up. The salt
-outlet temperature follows closely behind as the heated salt flows out of the core. The salt inlet
-temperature lags behind the other temperature measurements due to the time taken for the heated
-salt to travel through the out-of-core region. The temperature values appear to settle at higher
-temperatures to compensate for the inserted reactivity.
-
-\begin{figure}[t]
-  \centering
-%  \includegraphics[width=0.8\columnwidth]{u233-rod-worth-curve}
-  \caption{Change in power output following a step reactivity insertion of 24.8 pcm with the
-  \gls{MSRE} initially at 1 MW.}
-  \label{fig:insertion-power}
-\end{figure}
-
-\begin{figure}[t]
-  \centering
-%  \includegraphics[width=0.8\columnwidth]{u233-rod-worth-curve}
-  \caption{Change in maximum fuel salt and graphite temperatures, and the mean salt inlet and outlet
-  temperatures following a step reactivity insertion of 24.8 pcm with the \gls{MSRE} initially at 1
-  MW.}
-  \label{fig:insertion-temperature}
-\end{figure}
+\FloatBarrier
 
@@
 
+%The hybrid $S_N$-diffusion method combines the strengths of the $S_N$ and neutron
+%diffusion methods accurately model control rods while remaining tractable on modest computational
+%resources. This chapter covers work demonstrating the hybrid $S_N$-diffusion method through two
+%time-dependent reactivity-initiated simulations. The hybrid $S_N$-diffusion solver implemented in
+%Moltres is coupled to preexisting \gls{DNP} and temperature solver capabilities to model
+%transient behavior in reactors. 
+%
+%The first time-dependent simulation is based on a zero-power \gls{MSRE} rod drop experiment. The
+%rod drop is estimated to induce a 1500 pcm reactivity withdrawal in the \gls{MSRE}. Preliminary
+%calculations determined the required initial rod height in the numerical model to replicate the
+%same reactivity withdrawal effect and eliminate rod worth discrepancies. Following an initial
+%$k$-eigenvalue simulation to obtain initial flux and \gls{DNP} distributions, the rod drop
+%simulation ran with total fission neutron production rate measurements at each timestep. The
+%integral neutron production, normalized to match the initial experimental neutron count rate,
+%showed good agreement with the experimental integral neutron count curve following the rod drop.
+%The numerical result has a maximum discrepancy of x \% at $t=16$ s. Experimental uncertainties in
+%the actual $^{235}$U content, initial rod height, rod acceleration, and the initial neutron count
+%rate may be responsible for some of the discrepancy. Snapshots of the 3-D neutron group flux
+%distributions at various timesteps show expected localized flux decreases around the control rod
+%as it moves through the core.
+%
+%The second time-dependent simulation is based on a reactivity insertion experiment on the
+%$^{233}$U-fueled \gls{MSRE}. A control rod withdrawal induces a 24.8 pcm step reactivity
+%insertion on the \gls{MSRE} initially running at 1 MW with steady salt flow. Moltres reproduces
+%the general shape of the first power peak and the subsequent power oscillations. The power peak
+%magnitude differs by y \% compared to \gls{MSRE} data. The fuel
+%salt, graphite, inlet, and outlet temperature measurements show delayed responses to the increase
+%in power output due to the time taken for the salt and graphite to heat up. The temperature values
+%eventually settle higher than the initial
+%state to compensate for the reactivity insertion.
+%
+%Overall, this work has successfully demonstrated the hybrid method for coupled time-dependent
+%reactor analyses of reactivity-initiated transients involving control rods.
+
 The hybrid $S_N$-diffusion method combines the strengths of the $S_N$ and neutron
 diffusion methods accurately model control rods while remaining tractable on modest computational
-resources. This chapter covers work demonstrating the hybrid $S_N$-diffusion method through two
-time-dependent reactivity-initiated simulations.The hybrid $S_N$-diffusion solver implemented in
+resources. This chapter covers work demonstrating the hybrid $S_N$-diffusion method through a
+time-dependent reactivity-initiated simulations. The hybrid $S_N$-diffusion solver implemented in
 Moltres is coupled to preexisting \gls{DNP} and temperature solver capabilities to model
@@
 
-The first time-dependent simulation is based on a zero-power \gls{MSRE} rod drop experiment. The
+The time-dependent simulation is based on a zero-power \gls{MSRE} rod drop experiment. The
 rod drop is estimated to induce a 1500 pcm reactivity withdrawal in the \gls{MSRE}. Preliminary
@@
 showed good agreement with the experimental integral neutron count curve following the rod drop.
-The numerical result has a maximum discrepancy of x \% at $t=16$ s. Experimental uncertainties in
-the actual $^{235}$U content, initial rod height, rod acceleration, and the initial neutron count
-rate may be responsible for some of the discrepancy. Snapshots of the 3-D neutron group flux
-distributions at various timesteps show expected localized flux decreases around the control rod
-as it moves through the core.
-
-The second time-dependent simulation is based on a reactivity insertion experiment on the
-$^{233}$U-fueled \gls{MSRE}. A control rod withdrawal induces a 24.8 pcm step reactivity
-insertion on the \gls{MSRE} initially running at 1 MW with steady salt flow. Moltres reproduces
-the general shape of the first power peak and the subsequent power oscillations. The power peak
-magnitude differs by y \% compared to \gls{MSRE} data. The fuel
-salt, graphite, inlet, and outlet temperature measurements show delayed responses to the increase
-in power output due to the time taken for the salt and graphite to heat up. The temperature values
-eventually settle higher than the initial
-state to compensate for the reactivity insertion.
-
+Convergence issues with the nested coupling structure of the hybrid method and the \gls{DNP}
+subsolver impacted the precision of several data points during the simulation.
+Future investigations could look into better coupling schemes to resolves these issues.
+Experimental uncertainties in the actual $^{235}$U concentration, initial \& final rod heights, rod
+acceleration, and the initial neutron count rate were also responsible for discrepancies relative
+to experimental data.
+%Snapshots of the 3-D neutron group flux
+%distributions at various timesteps show expected localized flux decreases around the control rod
+%as it moves through the core.
 Overall, this work has successfully demonstrated the hybrid method for coupled time-dependent
-reactor analyses of reactivity-initiated transients involving control rods.
+reactor analysis of reactivity-initiated transients involving control rods.

### 2024-12-09 presentation/1d.tex
presentation edits

@@
 \begin{frame}
-  \frametitle{Hybrid $S_N$-Diffusion Method: 1-D Neutronics Eigenvalue Simulations}
-  \textbf{1-D Neutronics Model Geometries}
+  \frametitle{1-D MSRE Neutronics Model Geometries}
   \begin{figure}[htb!]
@@
 \begin{frame}
-  \frametitle{Hybrid $S_N$-Diffusion Method: 1-D Neutronics Eigenvalue Simulations}
+  \frametitle{1-D MSRE Neutronics Modeling Approach}
   \begin{columns}
@@
       \item Material compositions derived from the MSRE design
-      \item Reduced gadolinium content in control rod to reduce rod worth from 50000 pcm to 20000 pcm
+      \item Reduced gadolinium content in control rod to 0.35 wt\%
       \item Eight neutron energy groups
@@
 \begin{frame}
-  \frametitle{Hybrid $S_N$-Diffusion Method: 1-D Neutronics Eigenvalue Simulations}
+  \frametitle{1-D MSRE Neutronics Modeling Approach}
   \textbf{1-D Neutronics Model Numerical Solvers}
@@
 \begin{frame}
-  \frametitle{Hybrid $S_N$-Diffusion Method: 1-D Neutronics Eigenvalue Simulations}
+  \frametitle{1-D MSRE Neutronics Simulation Results}
   \textbf{1-D Neutronics Model Reactivity Results}
@@
 \begin{frame}
-  \frametitle{Hybrid $S_N$-Diffusion Method: 1-D Neutronics Eigenvalue Simulations}
+  \frametitle{1-D MSRE Neutronics Simulation Results}
   \textbf{1-D Neutronics Model Control Rod Worth Results}
@@
     \column{5.5cm}
-    \begin{figure}[htb!]
-      \centering
-      \includegraphics[width=\columnwidth]{case-3a-flux}
-      \caption{Case 3a neutron group flux distributions from OpenMC-CE and OpenMC-MG.}
-      \label{fig:3a-flux}
-    \end{figure}
-    \column{5.5cm}
-    \begin{figure}[htb!]
-      \centering
-      \includegraphics[width=\columnwidth]{case-3b-flux}
-      \caption{Case 3b neutron group flux distributions from OpenMC-CE and OpenMC-MG.}
-      \label{fig:3b-flux}
-    \end{figure}
-  \end{columns}
-\end{frame}
-
-\begin{frame}
-%  \frametitle{Hybrid $S_N$-Diffusion Method: 1-D Neutronics Eigenvalue Simulations}
-  \begin{columns}
-    \column{5.5cm}
-    \textbf{Case 3a Neutron Flux Distributions}
+    \textbf{Case 3b Neutron Flux Distributions}
     \begin{itemize}
@@
       \item The hybrid method performs better than the neutron diffusion method near $x=0$ cm where
-        the correction region is situated.
-    \end{itemize}
-    \column{5.5cm}
-    \begin{figure}[htb!]
-      \centering
-      \includegraphics[width=.975\columnwidth]{case-3a-flux-diff}
-      \caption{Absolute difference in neutron group flux distributions for Case 3a from Moltres-$S_8$,
-      Moltres-diffusion, Moltres-hybrid, and Python-hybrid relative to OpenMC-MG.}
-      \label{fig:3a-flux-diff}
-    \end{figure}
-  \end{columns}
-\end{frame}
-
-\begin{frame}
-%  \frametitle{Hybrid $S_N$-Diffusion Method: 1-D Neutronics Eigenvalue Simulations}
-  \begin{columns}
-    \column{5.5cm}
-    \textbf{Case 3b Neutron Flux Distributions}
-    \begin{itemize}
-      \item The neutron diffusion method performance worsens with the control rod region introduced
-        near $x=0$.
+        the control rod is situated.
     \end{itemize}
@@
 \begin{frame}
-  \frametitle{Hybrid $S_N$-Diffusion Method: 1-D Neutronics Eigenvalue Simulations}
-  \textbf{Impact of Correction Subregion Sizes on $k$}
-  \begin{itemize}
-    \item Minimizing the correction region size is essential for the hybrid method to be
-      computationally efficient for time-dependent simulations.
-    \item $k$ varies by up to 164 pcm for Case 3a and 109 pcm for Case 3b.
-    \item The hybrid method $k$ value does not converge monotonically towards the $S_8$ method $k$ value,
-      implying other sources of discrepancies.
-  \end{itemize}
-  \begin{figure}[h]
-    \centering
-    \begin{subfigure}[b]{0.49\columnwidth}
-      \centering
-      \includegraphics[width=\columnwidth]{correction-size-a-k}
-      \caption{Case 3a}
-      \label{fig:v1-size-a-k}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[b]{0.49\columnwidth}
-      \centering
-      \includegraphics[width=\columnwidth]{correction-size-b-k}
-      \caption{Case 3b}
-      \label{fig:v1-size-b-k}
-    \end{subfigure}
-    \caption{$k_\text{eff}$ estimates from the hybrid method for Cases 3a and 3b with different
-    correction subregion sizes. The horizontal lines indicate $k_\text{eff}$ estimates from the
-    OpenMC-CE, OpenMC-MG, and $S_8$ methods.}
-    \label{fig:v1-size-k}
-  \end{figure}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Hybrid $S_N$-Diffusion Method: 1-D Neutronics Eigenvalue Simulations}
+  \frametitle{1-D MSRE Neutronics Simulation Results}
   \textbf{Impact of Correction Subregion Sizes on Rod Worth}
@@
 \begin{frame}
-  \frametitle{Hybrid $S_N$-Diffusion Method: 1-D Neutronics Eigenvalue Simulations}
-  \textbf{Relaxing the $S_N$ Convergence Tolerance}
-  \begin{itemize}
-    \item The hybrid $S_N$-diffusion method iteratively couples the neutron diffusion solver with
-      the $S_N$ subsolver.
-    \item Diffusion-based acceleration schemes for neutron transport methods typically do not
-      require the neutron transport calculations to fully converge.
-    \item Transport correction parameters tend to converge faster than scalar flux.
-    \item Relaxing the $S_N$ subsolver convergence tolerance would provide computational savings to
-      the hybrid method.
-  \end{itemize}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Hybrid $S_N$-Diffusion Method: 1-D Neutronics Eigenvalue Simulations}
+  \frametitle{1-D MSRE Neutronics Simulation Results}
   \begin{columns}
@@
     \begin{itemize}
-      \item The number of outer iterations generally increases with the $S_8$ convergence tolerance.
+      \item Transport correction parameters converge faster than scalar flux in the $S_N$ subsolver.
+      \item Relaxing the $S_N$ subsolver convergence tolerance would provide computational savings.
       \item The hybrid method exhibits superlinear ($q=1.333$) convergence in $k$ with respect to

### 2024-12-09 presentation/2d.tex
slide edits
presentation edits

@@
 \begin{frame}
-  \frametitle{Hybrid $S_N$-Diffusion Method: 2-D Neutronics Eigenvalue Simulations}
-  \textbf{2-D Neutronics Quarter-Core \& Full-Core MSRE Models}
+  \frametitle{2-D Neutronics Quarter-Core \& Full-Core MSRE Models}
   \begin{columns}
@@
 \begin{frame}
-  \frametitle{Hybrid $S_N$-Diffusion Method: 2-D Neutronics Eigenvalue Simulations}
+  \frametitle{2-D MSRE Neutronics Modeling Approach}
   \begin{columns}
@@
 \begin{frame}
-  \frametitle{Hybrid $S_N$-Diffusion Method: 2-D Neutronics Eigenvalue Simulations}
+  \frametitle{2-D MSRE Neutronics Simulation Results}
   \textbf{2-D Quarter-Core $k$ \& Rod Worth Results}
@@
 
-  The hybrid method produces accurate rod worth estimates while the neutron diffusion method
+  $\Rightarrow$ The hybrid method produces accurate rod worth estimates while the neutron diffusion method
       significantly overestimates rod worth.
@@
     \column{6cm}
-    \centerline{\textbf{Control Rod Withdrawn}}
+    \centerline{\large \textbf{Control Rod Withdrawn}}
   \begin{figure}[p]
@@
     \column{6cm}
-    \centerline{\textbf{Control Rod Inserted}}
+    \centerline{\large \textbf{Control Rod Inserted}}
   \begin{figure}[p]
@@
 \begin{frame}
-  \frametitle{Hybrid $S_N$-Diffusion Method: 2-D Neutronics Eigenvalue Simulations}
+  \frametitle{2-D MSRE Neutronics Simulation Results}
   \textbf{2-D Quarter-Core Normalized Fuel Channel Power Distribution}
@@
 \begin{frame}
-  \frametitle{Hybrid $S_N$-Diffusion Method: 2-D Neutronics Eigenvalue Simulations}
+  \frametitle{2-D MSRE Neutronics Simulation Results}
   \textbf{2-D Full-Core Rod Worth Results}
@@
 
-  The hybrid method remains effective at improving control rod worth estimates in the full-core
+  $\Rightarrow$ The hybrid method remains effective at improving control rod worth estimates in the full-core
   model.
@@
 \begin{frame}
-  \frametitle{Hybrid $S_N$-Diffusion Method: 2-D Neutronics Eigenvalue Simulations}
+  \frametitle{2-D MSRE Neutronics Simulation Results}
   \textbf{2-D Full-Core Normalized Fuel Channel Power Distribution}

### 2024-12-09 presentation/3d.tex
slide edits
presentation edits

@@
 \begin{frame}
-  \frametitle{Hybrid $S_N$-Diffusion Method: 3-D Neutronics Eigenvalue Simulations}
+  \frametitle{3-D MSRE Full-Core Model}
   \begin{columns}
     \column{5.5cm}
-  \textbf{3-D Full-Core MSRE Model}
   \begin{figure}[p]
     \centering
-    \includegraphics[width=0.85\columnwidth]{msre-picture}
+    \includegraphics[width=0.8\columnwidth]{msre-picture}
     \caption{Vertical cross section of the actual \gls{MSRE} vessel.}
@@
     \centering
-    \includegraphics[width=.95\columnwidth]{msre-geom-vert}
+    \includegraphics[width=.87\columnwidth]{msre-geom-vert}
     \caption{Vertical cross section of the 3-D numerical \gls{MSRE} model offset by 5.08 cm to show
@@
 \begin{frame}
-  \frametitle{Hybrid $S_N$-Diffusion Method: 3-D Neutronics Eigenvalue Simulations}
+  \frametitle{3-D MSRE Neutronics Modeling Approach}
+  \begin{columns}
+    \column{5.5cm}
     \textbf{3-D Full-Core MSRE Model Details}
@@
       \item Hybrid $S_6$-diffusion method
+      \item Eight neutron energy groups
       \item $^{235}$U concentration at initial criticality
       \item Uniform temperature at 911 K
-    \item Stabilization factor of $c=250$
-    \item Void constant of $\varsigma=0.5$
-    \item Uncorrected diffusion coefficient values capped at $D_g = 2.5$
     \end{itemize}
-  \vspace{.2cm}
-
-  These parameters are required for numerical stability in the air-filled control rod
-  thimble region.
     \vspace{.2cm}
@@
     report data (Serpent 2 model) \cite{fratoni_molten_2020}, and the OpenMC model in this work.
+    \column{5.5cm}
+    \begin{figure}[p]
+      \centering
+      \includegraphics[width=.9\columnwidth]{wedge-view-high-res.png}
+      \caption{3-D section view of the 3-D numerical \gls{MSRE} model showing
+      the three control rod thimbles and the fuel-graphite lattice.}
+      \label{fig:msre-geom-vert}
+    \end{figure}
+  \end{columns}
 \end{frame}
@@
 \begin{frame}
-  \frametitle{Hybrid $S_N$-Diffusion Method: 3-D Neutronics Eigenvalue Simulations}
+  \frametitle{3-D MSRE Neutronics Simulation Results}
   \textbf{MSRE at Initial Criticality}
@@
       \toprule
-       & {$k_\text{eff}$} \\
+      Source & {$k_\text{eff}$} \\
        \midrule
@@
       OpenMC (This work) & 1.01308(20) \\
-      Hybrid (This work) & 1.01955 \\
+      Hybrid (This work) & 1.01957 \\
       Diffusion (This work) & 1.01885 \\
@@
   \end{table}
+  \vspace{.2cm}
 
-  \begin{itemize}
-    \item The Serpent 2 benchmark model reports a $k_\text{eff}$ of 1.01093(3) after considering some
-      of the geometrical modifications used in the OpenMC model.
-    \item The hybrid and neutron diffusion models agree with the OpenMC model within 700 pcm.
-  \end{itemize}
+  $\Rightarrow$ The hybrid and neutron diffusion models agree with the Serpent 2 and OpenMC models within
+  700 pcm.
 \end{frame}
@@
 \begin{frame}
-  \frametitle{Hybrid $S_N$-Diffusion Method: 3-D Neutronics Eigenvalue Simulations}
+  \frametitle{3-D MSRE Neutronics Simulation Results}
   \textbf{MSRE Rod Worth Measurements}
@@
     \centering
-    \includegraphics[width=0.8\columnwidth]{rod-worth-curve}
+    \includegraphics[width=0.75\columnwidth]{rod-worth-curve}
     \caption{Reactivity inserted by Rod 1 at various rod positions relative to the full insertion.}
@@
     \caption{Total rod worth of Rod 1 when fully inserted.}
+    \small
     \begin{tabular}{l S[table-format=4(2)] S}
@@
 \begin{frame}
-  \frametitle{Hybrid $S_N$-Diffusion Method: 3-D Neutronics Eigenvalue Simulations}
+  \frametitle{3-D MSRE Neutronics Simulation Results}
   \textbf{Computational Performance}
@@
 \begin{frame}
-  \frametitle{Hybrid $S_N$-Diffusion Method: 3-D Neutronics Eigenvalue Simulations}
+  \frametitle{3-D MSRE Neutronics Simulation Results}
   \textbf{Strong Scaling Test}
@@
   The $S_N$-diffusion data transfer processes scale poorly beyond 40 nodes.
+  \begin{columns}
+  \column{11cm}
   \begin{figure}[t]

### 2024-12-09 presentation/backup.tex
presentation edits

\begin{frame}[noframenumbering]
  \frametitle{V\&V Study 1: Verification of Moltres with the CNRS Benchmark}

  Published in \textit{S.M. Park, M. Munk, "Verification of Moltres for Multiphysics Simulations of
    Fast-Spectrum Molten Salt Reactors," Annals of Nuclear Energy, vol. 173, Aug 2022.}
  \vspace{.2cm}

  \begin{columns}
    \column[t]{6.5cm}
    \textbf{Description of the CNRS Benchmark \cite{tiberga_results_2020}}
    \vspace{.2cm}

      \begin{itemize}
        \item A numerical benchmark for multiphysics software dedicated to modeling ``pool-type''
          fast-spectrum MSRs
        \item Problem Setup
          \begin{itemize}
            \item LiF-BeF$_2$-UF$_4$ molten salt
            \item Six neutron energy groups
            \item Eight precursor groups
            \item Volumetric conjugate heat sink term
            \item Incompressible flow
          \end{itemize}
        \item Consists of three phases
          \begin{itemize}
            \item Phase 0: Single-physics verification
            \item Phase 1: Steady-state coupling
            \item Phase 2: Time-dependent coupling
          \end{itemize}
      \end{itemize}
    \column[t]{3.5cm}
    \begin{figure}
      \centering
      \includegraphics[width=\columnwidth]{../images/cnrs-geometry}
      \caption{CNRS benchmark problem domain \cite{tiberga_results_2020}}
    \end{figure}
  \end{columns}
\end{frame}

\begin{frame}[noframenumbering]
  \frametitle{V\&V Study 1: Verification of Moltres with the CNRS Benchmark}
  \begin{block}{\textbf{Study Outcome}}
    Moltres is consistent with the other multiphysics software in the benchmark problems.
  \end{block}
  \begin{columns}
    \hfill
    \column[t]{4cm}
    \vspace{.3cm}
    \begin{figure}
      \centering
      \includegraphics[width=\columnwidth]{../images/full-coupled}
      \caption{Step 1.4 - Temperature distribution and flow streamlines for the fully
      coupled system \cite{park_verification_2022}.}
    \end{figure}
    \column[t]{4cm}
    \begin{figure}
      \centering
      \includegraphics[width=\columnwidth]{../images/2-1-gain-plot}
      \caption{Step 2.1 - Bode gain plot of the frequency response of the fully
      coupled system \cite{park_verification_2022}.}
    \end{figure}
    \column[t]{4cm}
    \begin{figure}
      \centering
      \includegraphics[width=\columnwidth]{../images/2-1-phase-plot}
      \caption{Step 2.1 - Bode phase plot of the frequency response of the fully
      coupled system \cite{park_verification_2022}.}
    \end{figure}
    \hfill
  \end{columns}
\end{frame}

\begin{frame}[noframenumbering]
  \frametitle{V\&V Study 2: MSRE Pump Start-up \& Coast-Down Transients}
  \textbf{DNP Distributions Under Static Conditions}
  \begin{figure}[t]
    \centering
    \begin{subfigure}[b]{0.48\columnwidth}
      \centering
      \includegraphics[width=\columnwidth]{centerline_pre}
      \caption{Normalized centerline \gls{DNP} distributions.}
      \label{fig:centerline-pre-dist}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.48\columnwidth}
      \centering
      \includegraphics[width=\columnwidth]{midplane_pre}
      \caption{Normalized midplane \gls{DNP} distributions.}
      \label{fig:midplane-pre-dist}
    \end{subfigure}
    \caption{\gls{DNP} distribution from Moltres and ratios comparing QuasiMolto and
    Moltres models under static \gls{MSRE} conditions. Maximum relative difference $\approx 0.1 \%$.}
    \label{fig:centerline-pre}
  \end{figure}
  \begin{itemize}
    \item Perfect overlap in centerline and midplane DNP distributions between Moltres and QuasiMolto.
  \end{itemize}
\end{frame}

\begin{frame}[noframenumbering]
  \frametitle{Backward-Facing Step Flow Validation Test Results}
  \begin{columns}
    \column{11.5cm}
    \begin{figure}[htb]
      \centering
      \hfill
      \begin{subfigure}[t]{0.24\columnwidth}
        \centering
        \includegraphics[width=\columnwidth]{bfs_downstream_vel}
        \caption{Normalized velocity distributions downstream of step.}
        \label{fig:bfs-downstream}
      \end{subfigure} %\hfill \\
  %    \centering
      \hfill
      \begin{subfigure}[t]{0.24\columnwidth}
        \centering
        \includegraphics[width=\columnwidth]{bfs_cf}
        \caption{Skin friction coefficient along the bottom wall.}
        \label{fig:bfs-cf}
      \end{subfigure}
      \hfill
      \begin{subfigure}[t]{0.24\columnwidth}
        \centering
        \includegraphics[width=\columnwidth]{bfs_cp}
        \caption{Skin pressure coefficient along the bottom wall.}
        \label{fig:bfs-cp}
      \end{subfigure}% \hfill \\
  %    \centering
      \begin{subfigure}[t]{0.24\columnwidth}
        \centering
        \includegraphics[width=\columnwidth]{bfs_stress}
        \caption{Normalized turbulent shear stress downstream
        of step.}
        \label{fig:bfs-stress}
      \end{subfigure}
      \caption{Comparison of backward facing step flow results against reference
      experimental data and computational data from CFL3D. $x/H$ values are normalized horizontal
      distances relative to the step.}
      \label{fig:bfs-plots}
    \end{figure}
  \end{columns}
  \begin{itemize}
    \item Minor discrepancies between numerical and experimental data typical of RANS-based turbulence models.
    \item Moltres provides better predictments along $x/H=1$ than CFL3D.
  \end{itemize}
\end{frame}

\begin{frame}[noframenumbering]
  \frametitle{Hybrid $S_N$-Diffusion Method: Theory}
  \textbf{Weak Formulation of the Multigroup SAAF $S_N$ Equations}
  \begin{gather}
    \shortintertext{Streaming term:}
    \sum^G_{g=1}\sum^{N_d}_{d=1}w_d\left(\hat{\Omega}_d\cdot\nabla\Psi^*_{g,d},\tau_g\hat{\Omega}
    \cdot\nabla\Psi_{g,d}-(1-\tau_g\Sigma_{t,g})\Psi_{g,d}\right)_\mathcal{D} \\
    \shortintertext{Collision term: }
    \sum^G_{g=1}\sum^{N_d}_{d=1}w_d\left(\Psi^*_{g,d},\Sigma_{t,g}\Psi_{g,d}\right)_\mathcal{D} \\
    \shortintertext{Scattering term:}
    \sum^G_{g=1}\sum^{N_d}_{d=1}w_d\left(\Psi^*_{g,d}+\tau_g\hat{\Omega}_d\cdot\nabla\Psi^*_{g,d},
    \sum^G_{g'=1}\sum^L_{l=0}\Sigma^{g'\rightarrow g}_{s,l}\sum^l_{m=-l}
    \frac{2l+1}{w}Y_{l,m}(\hat{\Omega}_d)\phi_{g',l,m}\right)_\mathcal{D} \\
    \shortintertext{Fission source term: }
    \sum^G_{g=1}\sum^{N_d}_{d=1}w_d\left(\Psi^*_{g,d}+\tau_g\hat{\Omega}_d\cdot\nabla\Psi^*_{g,d},
    \frac{1}{w}\frac{\chi_{p,g}(1-\beta)}{k}\sum^G_{g'=1}\nu\Sigma_{f,g'}\phi_{g'}\right)_\mathcal{D}
  \end{gather}
\end{frame}

\begin{frame}[noframenumbering]
  \frametitle{Hybrid $S_N$-Diffusion Method: Theory}
  \textbf{Weak Formulation of the Multigroup SAAF $S_N$ Equations}
  \begin{gather}
    \shortintertext{Delayed neutron source term:}
    \sum^G_{g=1}\sum^{N_d}_{d=1}w_d\left(\Psi^*_{g,d}+\tau_g\hat{\Omega}_d\cdot\nabla\Psi^*_{g,d},
    \frac{1}{w}\sum ^I_{i=1}\chi_{d,g}\lambda_i C_i\right)_\mathcal{D}
    \shortintertext{Boundary source term:}
    \begin{cases}
      \sum^G_{g=1}\sum^{N_d}_{d=1}w_d\left(\Psi^*_{g,d},
      \hat{\Omega}_d\cdot\hat{n}_b\Psi_{g,d}\right)_{\partial\mathcal{D}},
      & \hat{\Omega}\cdot\hat{n}_b>0,\vec{r}\in\partial\mathcal{D} \\
      \sum^G_{g=1}\sum^{N_d}_{d=1}w_d\left(\Psi^*_{g,d},
      \hat{\Omega}_d\cdot\hat{n}_b\Psi^\text{inc}_{g,d}\right)_{\partial\mathcal{D}},
      & \hat{\Omega}\cdot\hat{n}_b<0,\vec{r}\in\partial\mathcal{D}
    \end{cases} \label{eq:boundary-source} \\
    \shortintertext{Reflecting boundary term:}
    \begin{cases}
      \sum^G_{g=1}\sum^{N_d}_{d=1}w_d\left(\Psi^*_{g,d},
      \hat{\Omega}_d\cdot\hat{n}_b\Psi_{g,d}\right)_{\partial\mathcal{D}},
      & \hat{\Omega}\cdot\hat{n}_b>0,\vec{r}\in\partial\mathcal{D}_s \\
      \sum^G_{g=1}\sum^{N_d}_{d=1}w_d\left(\Psi^*_{g,d},
      \hat{\Omega}_d\cdot\hat{n}_b\Psi_{g,d_r}\right)_{\partial\mathcal{D}},
      & \hat{\Omega}\cdot\hat{n}_b<0,\vec{r}\in\partial\mathcal{D}_s
    \end{cases} \label{eq:reflecting-bc}
  \end{gather}
\end{frame}

\begin{frame}[noframenumbering]
  \frametitle{Hybrid $S_N$-Diffusion Method: Theory}
  \textbf{Weak Formulation of the Multigroup SAAF $S_N$ Equations}
  \begin{gather}
    \shortintertext{Void stabilization parameter \cite{wang_diffusion_2014}:}
    \tau_g =
    \begin{cases}
      \frac{1}{c\Sigma_{t,g}} \mbox{ for } ch\Sigma_{t,g} \geq \varsigma \\
      \frac{h}{\varsigma} \mbox{ for } ch\Sigma_{t,g} < \varsigma
    \end{cases}, \\
    \shortintertext{where}
    \begin{align*}
      h &= \mbox{mesh element size,} & \\
      c &= \mbox{maximum stabilization factor,} & \\
      \varsigma &= \mbox{void constant.} &
    \end{align*}
  \end{gather}
  $c=1$ and $\varsigma=0.5$ by default.
  \vspace{.2cm}

  The SAAF-$S_N$ equations require this stabilization scheme
  in near-void regions where $\Sigma_{t,g}$ is very small.
\end{frame}

\begin{frame}[noframenumbering]
%  \frametitle{Hybrid $S_N$-Diffusion Method: 1-D Neutronics Eigenvalue Simulations}
  \begin{columns}
    \column{5.5cm}
    \begin{figure}[htb!]
      \centering
      \includegraphics[width=\columnwidth]{case-3a-flux}
      \caption{Case 3a neutron group flux distributions from OpenMC-CE and OpenMC-MG.}
      \label{fig:3a-flux}
    \end{figure}
    \column{5.5cm}
    \begin{figure}[htb!]
      \centering
      \includegraphics[width=\columnwidth]{case-3b-flux}
      \caption{Case 3b neutron group flux distributions from OpenMC-CE and OpenMC-MG.}
      \label{fig:3b-flux}
    \end{figure}
  \end{columns}
\end{frame}

\begin{frame}[noframenumbering]
%  \frametitle{Hybrid $S_N$-Diffusion Method: 1-D Neutronics Eigenvalue Simulations}
  \begin{columns}
    \column{5.5cm}
    \textbf{Case 3a Neutron Flux Distributions}
    \begin{itemize}
      \item The neutron diffusion and hybrid methods fare worse than the $S_8$ method at capturing
        the oscillatory flux pattern.
      \item The hybrid method performs better than the neutron diffusion method near $x=0$ cm where
        the correction region is situated.
    \end{itemize}
    \column{5.5cm}
    \begin{figure}[htb!]
      \centering
      \includegraphics[width=.975\columnwidth]{case-3a-flux-diff}
      \caption{Absolute difference in neutron group flux distributions for Case 3a from Moltres-$S_8$,
      Moltres-diffusion, Moltres-hybrid, and Python-hybrid relative to OpenMC-MG.}
      \label{fig:3a-flux-diff}
    \end{figure}
  \end{columns}
\end{frame}

\begin{frame}[noframenumbering]
  \frametitle{Hybrid $S_N$-Diffusion Method: 1-D Neutronics Eigenvalue Simulations}
  \textbf{Impact of Correction Subregion Sizes on $k$}
  \begin{itemize}
    \item Minimizing the correction region size is essential for the hybrid method to be
      computationally efficient for time-dependent simulations.
    \item $k$ varies by up to 164 pcm for Case 3a and 109 pcm for Case 3b.
    \item The hybrid method $k$ value does not converge monotonically towards the $S_8$ method $k$ value,
      implying other sources of discrepancies.
  \end{itemize}
  \begin{figure}[h]
    \centering
    \begin{subfigure}[b]{0.49\columnwidth}
      \centering
      \includegraphics[width=\columnwidth]{correction-size-a-k}
      \caption{Case 3a}
      \label{fig:v1-size-a-k}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.49\columnwidth}
      \centering
      \includegraphics[width=\columnwidth]{correction-size-b-k}
      \caption{Case 3b}
      \label{fig:v1-size-b-k}
    \end{subfigure}
    \caption{$k_\text{eff}$ estimates from the hybrid method for Cases 3a and 3b with different
    correction subregion sizes. The horizontal lines indicate $k_\text{eff}$ estimates from the
    OpenMC-CE, OpenMC-MG, and $S_8$ methods.}
    \label{fig:v1-size-k}
  \end{figure}
\end{frame}

\begin{frame}[noframenumbering]
  \frametitle{Hybrid $S_N$-Diffusion Method: 1-D Neutronics Eigenvalue Simulations}
  \textbf{Relaxing the $S_N$ Convergence Tolerance}
  \begin{itemize}
    \item The hybrid $S_N$-diffusion method iteratively couples the neutron diffusion solver with
      the $S_N$ subsolver.
    \item Diffusion-based acceleration schemes for neutron transport methods typically do not
      require the neutron transport calculations to fully converge.
    \item Transport correction parameters tend to converge faster than scalar flux.
    \item Relaxing the $S_N$ subsolver convergence tolerance would provide computational savings to
      the hybrid method.
  \end{itemize}
\end{frame}

\begin{frame}[noframenumbering]
  \frametitle{Hybrid $S_N$-Diffusion Method: 3-D Neutronics Eigenvalue Simulations}
  \textbf{3-D Full-Core MSRE Model Details}
  \begin{itemize}
    \item Hybrid $S_6$-diffusion method
    \item $^{235}$U concentration at initial criticality
    \item Uniform temperature at 911 K
    \item Stabilization factor of $c=250$
    \item Void constant of $\varsigma=0.5$
    \item Uncorrected diffusion coefficient values capped at $D_g = 2.5$
  \end{itemize}
  \vspace{.2cm}

  These parameters are required for numerical stability in the air-filled control rod
  thimble region.
  \vspace{.2cm}

  All hybrid method results are compared with MSRE experimental data, the MSRE numerical benchmark
  report data (Serpent 2 model) \cite{fratoni_molten_2020}, and the OpenMC model in this work.
\end{frame}

\begin{frame}[noframenumbering]
  \frametitle{Hybrid $S_N$-Diffusion Method: 3-D Neutronics Eigenvalue Simulations}
  \textbf{MSRE at Initial Criticality}
  \begin{table}[htb]
    \centering
    \caption{$k_\text{eff}$ values from \gls{MSRE} experimental data, the \gls{MSRE} numerical
    benchmark \cite{fratoni_molten_2020}, and the OpenMC and Moltres models in this work.}
    \begin{tabular}{l S[table-format=1.5(2)]}
      \toprule
       & {$k_\text{eff}$} \\
       \midrule
      \gls{MSRE} experimental data & 1.00000(420) \\
      Serpent 2 (Numerical benchmark) & 1.02132(3) \\
      OpenMC (This work) & 1.01308(20) \\
      Hybrid (This work) & 1.01955 \\
      Diffusion (This work) & 1.01885 \\
      \bottomrule
    \end{tabular}
    \label{table:initial-crit}
  \end{table}

  \begin{itemize}
    \item The Serpent 2 benchmark model reports a $k_\text{eff}$ of 1.01093(3) after considering some
      of the geometrical modifications used in the OpenMC model.
    \item The hybrid and neutron diffusion models agree with the OpenMC model within 700 pcm.
  \end{itemize}
\end{frame}

\begin{frame}[noframenumbering]
  \frametitle{MSRE Rod Drop Experiment}
  \textbf{Rod Height}
  \begin{figure}[htb!]
    \centering
    \includegraphics[width=0.7\columnwidth]{rod-height}
    \caption{Evolution of rod height evaluated at each timestep within the first second of the rod
    drop simulation. The rod reaches its fully inserted height at $t=0.4046$ s.}
    \label{fig:rod-height}
  \end{figure}
\end{frame}

%\begin{frame}[noframenumbering]
%  \frametitle{Molten Salt Reactor Designs}
%  \begin{columns}
%    \column{3cm}
%    \begin{figure}
%      \centering
%      \footnotesize
%      \includegraphics[width=\textwidth]{./images/msre-photo}
%      \caption{Graphite assembly for the Molten Salt Reactor Experiment
%      \cite{ornl_first-ever_2023}.}
%    \end{figure}
%    \column{7cm}
%    \begin{table}
%      \footnotesize
%      \centering
%      \caption{Thermal-spectrum MSR designs under active development.}
%      \begin{tabular}{l l}
%        \toprule
%        Reactor & Organization \\
%        \midrule
%        Integral Molten Salt Reactor & Terrestrial Energy \\
%        TMSR-LF & CAS (China) \\
%        Compact Molten Salt Reactor & Seaborg Technologies \\
%        Copenhagen Atomics Waste Burner & Copenhagen Atomics \\
%        \bottomrule
%      \end{tabular}
%    \end{table}
%    \begin{table}
%      \footnotesize
%      \centering
%      \caption{Fast-spectrum MSR designs under active development.}
%      \begin{tabular}{l l}
%        \toprule
%        Reactor & Organization \\
%        \midrule
%        Molten Chloride Fast Reactor & TerraPower \\
%        Molten Salt Fast Reactor & CNRS (France) \\
%        Stable Salt Reactor - Wasteburner & Moltex Energy \\
%        Molten Chloride Salt Fast Reactor & Elysium Industries \\
%        \bottomrule
%      \end{tabular}
%    \end{table}
%  \end{columns}
%\end{frame}
%
%\begin{frame}[noframenumbering]
%  \frametitle{V\&V Study 1: Verification of Moltres with the CNRS Benchmark}
%  \begin{table}
%      \caption{List of software packages and their corresponding model
%      specifications for the CNRS Benchmark simulations
%      \cite{tiberga_results_2020}.}
%      \centering
%      \footnotesize
%      \begin{tabular}{p{1.8cm} p{3.3cm} p{1.6cm} p{1cm} p{1.1cm}}
%          \toprule
%          Software & Institute & Numerical method & Mesh & Neutronics model \\
%          \midrule
%          OpenFOAM & Centre national de la recherche scientifique (CNRS) & Finite volume & 200$\times$200 \newline Non-uniform & $SP_1$ \& $SP_3$ \\
%          OpenFOAM & Politecnico di Milano (PoliMi) & Finite volume & 400$\times$400 \newline Uniform & Neutron diffusion \\
%          GeN-Foam & Paul Scherrer Institute (PSI) & Finite volume & 200$\times$200 \newline Non-uniform & Neutron diffusion \\
%          PHANTOM-$S_N$ DGFlows & Delft University of Technology (TUD) & Discontinuous finite \newline element & 50$\times$50 \newline Uniform & $S_2$ \& $S_6$ \\
%          Moltres (This work) & University of Illinois at Urbana-Champaign (UIUC) & Continuous \& discontinuous finite element & 200$\times$200 \newline Uniform & Neutron diffusion \\
%          \bottomrule
%      \end{tabular}
%      \label{table:software}
%  \end{table}
%\end{frame}
%
%\begin{frame}[noframenumbering]
%  \frametitle{Turbulence Models}
%  Numerous types of turbulence models exist at various levels of fidelity. From lowest to highest
%  computational complexity:
%  \begin{itemize}
%      \item RANS-based models
%      \begin{itemize}
%          \item Eddy viscosity models
%          \begin{itemize}
%              \item Algebraic models
%              \item One-equation
%              \item Two-equation models
%          \end{itemize}
%          \item \gls{RSM}
%      \end{itemize}
%      \item \gls{DES}
%      \item \gls{LES}
%      \item \gls{DNS}
%  \end{itemize}
%\end{frame}
%
%\begin{frame}[noframenumbering]
%  \frametitle{Turbulence Models}
%  \gls{RANS}-based models are based on the RANS equations obtained by applying time-averaging on
%  the fluid flow equations:
%  \begin{gather}
%      \frac{\partial U_i}{\partial t} + U_j \frac{\partial u_i}{\partial x_j} =
%      -\frac{1}{\rho} \frac{\partial P}{\partial x_i} + \nu \nabla^2 U_i -
%      \frac{\partial \langle u_i u_j \rangle}{x_j}
%      \shortintertext{where}
%      U = \mbox{ mean component,} \qquad u = \mbox{ fluctuating component.} \nonumber
%  \end{gather}
%  Eddy viscosity models operate on the eddy viscosity hypothesis:
%  \begin{align}
%      \langle u_iu_j \rangle =& \frac{2}{3}k \delta_{ij} - \nu_T \left(
%      \frac{\partial U_i}{\partial x_j} + \frac{\partial U_j}{\partial x_i}
%      \right)
%      \shortintertext{where}
%        \nu_T =& \mbox{ eddy viscosity.} \nonumber
%  \end{align}
%  The various eddy viscosity models mainly differ in their approach to the closure problem of
%  calculating the eddy viscosity.
%\end{frame}
%
%\begin{frame}[noframenumbering]
%  \frametitle{Control Rods in MSRs}
%  \begin{table}
%    \centering
%    \footnotesize
%    \caption{List of MSR designs which contain control rods.}
%    \begin{tabular}{l l}
%      \toprule
%      Reactor & Spectrum \\
%      \midrule
%      Molten Salt Reactor Experiment & Thermal \\
%      Integral Molten Salt Reactor & Thermal \\
%      TMSR-LF1 & Thermal \\
%      Liquid Fluoride Thorium Reactor & Thermal \\
%      Compact Molten Salt Reactor* & Thermal \\
%      Stable Salt Reactor - Wasteburner* & Fast \\
%      ThorCon Reactor* & Thermal \\
%      \bottomrule
%    \end{tabular}
%  \end{table}
%  Asterisks indicate MSR designs with ``control'' rods labeled as ``shutdown'' rods.
%\end{frame}

### 2024-12-09 presentation/conclusion.tex
slide edits
presentation edits

@@
     \begin{itemize}
-      \item CNRS numerical benchmark study with Moltres
+      \item CNRS numerical benchmark study with Moltres (Dissertation)
       \item MSRE pump transient V\&V study with Moltres
@@
   \frametitle{Conclusion}
-  \begin{block}{\textbf{Hybrid $S_N$-Diffusion Method}}
+  \begin{block}{\textbf{Hybrid $\bm{S_N}$-Diffusion Method}}
     \textbf{Relevance}
     \begin{itemize}
-      \item No MSR studies exist for time-dependent MSR simulations involving reactivity-initiated
-        transients using control rods
+      \item There are no MSR simulation tools or studies that explicitly model moving control rods.
     \end{itemize}
@@
         rod regions for improved rod worth estimates
-      \item The hybrid method applies an adaptive domain decomposition technique to
-        couple the $S_N$ subsolver and the neutron diffusion solver
+      \item The hybrid method applies an adaptive boundary coupling between
+        the $S_N$ subsolver and the neutron diffusion solver to smooth flux gradients
       \item The hybrid method accurately reproduces control rod worths at only four times the
         computational cost of the neutron diffusion method
-      \item The hybrid method exhibits decent scalability on high-performance computing clusters
+      \item The hybrid method exhibits good scalability on high-performance computing clusters
     \end{itemize}

### 2024-12-09 presentation/control-rod.tex
presentation edits

@@
 \begin{frame}
-  \frametitle{Hybrid $S_N$-Diffusion Method: Motivation}
+  \frametitle{Control Rods in MSRs}
   \begin{columns}
@@
     \column[t]{5cm}
-    \textbf{Control Rods in MSRs}
     \vspace{.2cm}
@@
 \begin{frame}
-  \frametitle{Hybrid $S_N$-Diffusion Method: Motivation}
+  \frametitle{Control Rods in MSRs}
   \textbf{Control rods provide a means of controlling the fission rate in nuclear reactors.} 
@@
   \end{itemize}
-  \textbf{MSRs contain comparatively fewer control rods than most other reactor types due to:}
-  \begin{itemize}
-    \item Uniform liquid fuel burnup
-    \item Strong passive safety of the liquid fuel form
-    \item Low excess reactivity of fuel inventory
-    \item Availability of other control mechanisms
-  \end{itemize}
-  \pause
-  \textbf{Nevertheless, it is important to characterize control rod effects in all relevant
+  \textbf{Control rods are also important for licensing and safety characterization.}
+
+  $\Rightarrow$ \textbf{It is important to characterize control rod effects in all relevant
   transient scenarios.}
+%  \textbf{Many MSR designs contain comparatively fewer control rods than most other reactor types due to:}
+%  \begin{itemize}
+%    \item Uniform liquid fuel burnup
+%    \item Strong passive safety of the liquid fuel form
+%    \item Low excess reactivity of fuel inventory
+%    \item Availability of other control mechanisms
+%  \end{itemize}
+
 \end{frame}
@@
 \begin{frame}
-  \frametitle{Hybrid $S_N$-Diffusion Method: Motivation}
+  \frametitle{Control Rod Modeling}
   \textbf{Control rods induce highly anisotropic neutron fluxes and steep flux gradients in their
@@
 \begin{frame}
-  \frametitle{Hybrid $S_N$-Diffusion Method: Motivation}
-  \textbf{Control Rod Modeling in MSR Multiphysics Studies}
+  \frametitle{Control Rod Modeling in MSR Multiphysics Studies}
   \vspace{.2cm}
 
-  Common simplifications applied to control rod modeling include:
+  Common approximations applied to control rod modeling include:
   \begin{itemize}
@@
       \begin{itemize}
-        \item Neutronics mesh differs from thermal-hydraulics mesh
-        \item 1-D thermal-hydraulics model with uniform or prescribed heat deposition
+        \item Neutronics geometry differs from thermal-hydraulics geometry
         \item Static calibration factors may not be suitable for time-dependent simulations
@@
   \begin{block}{\textbf{Technical Gap}}
-    \textbf{There are no existing MSR simulation tools or studies involving moving control rods
-    which are explicitly modeled.}
+    \textbf{There are no MSR simulation tools or studies that explicitly model moving control rods.}
   \end{block}

### 2024-12-09 presentation/moltres.tex
presentation edits

\begin{frame}
  \frametitle{Moltres for MSR Multiphysics Modeling}
  \textbf{What is Moltres?}
  \begin{itemize}
    \item Moltres \cite{lindsay_introduction_2018} is an open-source, MOOSE-based
      multiphysics application for modeling MSRs
	\item \gls{MOOSE} \cite{lindsay_20_2022} is an open source finite-element framework
      that supports the development of multiphysics solvers
    \item Robust and highly scalable numerical routines from PETSc
    \item Flexible coupling capabilities provided by the MOOSE Multi-App system
%    \item Highly extensible and user-friendly interface for software development through
%      object-oriented programming paradigm in C++ and native compatibility with other MOOSE-based
%      apps
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Previous MSR Analyses with Moltres}
  \begin{columns}
    \hfill
    \column[t]{4cm}
      \textbf{\gls{MSRE}}
      \vfill
      \begin{figure}
        \centering
	    \includegraphics[width=.48\columnwidth]{../images/2d_gamma_heating_group1}
        \hfill
	    \includegraphics[width=.48\columnwidth]{../images/2d_gamma_heating_group2}
	    \caption{\footnotesize Group 1 and 2 neutron fluxes in a 2-D axisymmetric MSRE
	      model \cite{lindsay_moltres_2017}.}
      \end{figure}
    \hfill
    \column[t]{3.5cm}
      \textbf{\gls{MSFR}}
      \vfill
      \begin{figure}
        \centering
        \includegraphics[width=.8\columnwidth]{images/flow-temp-plasma}
        \caption{\footnotesize Steady-state temperature distribution and salt flow velocity fields
          in a 2-D axisymmetric MSFR model. \cite{park_advancement_2020}}
      \end{figure}
    \hfill
    \column[t]{3.5cm}
      \textbf{\gls{TAP} MSR}
      \vfill
      \begin{figure}
        \centering
        \includegraphics[width=\columnwidth]{images/tap-msr}
        \caption{\footnotesize Group 1 and 2 neutron fluxes in a 3-D TAP MSR model.
        \cite{lee_neutronics_2020}}
      \end{figure}
    \hfill
  \end{columns}
\end{frame}

\begin{frame}
  \frametitle{Moltres}
  \begin{block}{\textbf{Software Capabilities for MSR Modeling}}
    \begin{itemize}
  	  \item Multigroup neutron diffusion solver
      \item Temperature reactivity feedback by interpolating temperature-dependent group
        constant data\footnote{Group constants refer to neutron cross section data and other
        neutronics-related material properties}
      \item \gls{DNP} drift coupling with flow modeling
      \item Out-of-core \gls{DNP} decay
      \item Couples with MOOSE modules for thermal-hydraulics modeling
    \end{itemize}
  \end{block}
  \pause
  \begin{block}{\textbf{Areas of Improvement for Moltres for MSR Modeling}}
    \begin{itemize}
      \item Moltres does not currently have turbulence modeling capability for simulating turbulent
        flows
      \item Moltres cannot model highly neutron-absorbing control rods accurately
    \end{itemize}
  \end{block}
\end{frame}

### 2024-12-09 presentation/msr.tex
presentation edits

@@
-\begin{frame}
-  \frametitle{What is a Molten Salt Reactor (MSR)?}
-	\begin{itemize}
-	  \item Advanced nuclear reactor fueled by fissile
-		material dissolved in a molten salt mixture
-	  \item In most MSR designs, the fuel salt mixture doubles as the primary coolant for the
-        reactor
-	  \item Both thermal- and fast-spectrum configurations are viable
-	\end{itemize}
-	\begin{figure}
-	  \centering
-	  \includegraphics[width=.5\textwidth]{./images/msr}
-      \caption{Schematic diagram of a general channel-type MSR concept
-      \cite{u.s._doe_nuclear_energy_research_advisory_committee_technology_2002}.}
-	  \label{fig:msr}
-	\end{figure}
-\end{frame}
-
 %\begin{frame}
 %  \frametitle{What is a Molten Salt Reactor (MSR)?}
-%  \textbf{Advantages of MSRs over other reactor types}
 %	\begin{itemize}
-%    \item Robust passive safety
-%      \begin{itemize}
-%        \item strong temperature feedback
-%        \item large thermal margin to boiling
+%	  \item Advanced nuclear reactor fueled by fissile
+%		material dissolved in a molten salt mixture
+%	  \item In most MSR designs, the fuel salt mixture doubles as the primary coolant for the
+%        reactor
+%	  \item Both thermal- and fast-spectrum configurations are viable
 %	\end{itemize}
-%    \item Freeze plug drains fuel salt into containment tanks under emergency situations
-%    \item Online refueling and reprocessing can:
-%      \begin{itemize}
-%        \item reduce reactor downtime
-%        \item reduce required fissile inventory in reactor at any given time
-%        \item reduce transuranic waste produced
-%        \item allow for $^{233}$U breeding from $^{232}$Th
-%      \end{itemize}
-%    \item Viable as fast-spectrum breeder/burner reactor
-%    \item Can provide high-temperature heat for industrial heat applications
-%  \end{itemize}
+%	\begin{figure}
+%	  \centering
+%	  \includegraphics[width=.5\textwidth]{./images/msr}
+%      \caption{Schematic diagram of a general channel-type MSR concept
+%      \cite{u.s._doe_nuclear_energy_research_advisory_committee_technology_2002}.}
+%	  \label{fig:msr}
+%	\end{figure}
 %\end{frame}
@@
   \frametitle{Molten Salt Reactor Modeling \& Simulation}
-  \textbf{While modeling MSRs is not necessarily more difficult than modeling solid-fueled
-  reactors, we must adapt our software tools to accurately model phenomena unique to MSRs.}\\
+  \textbf{What is a Molten Salt Reactor (MSR)?}
+  \begin{itemize}
+	\item MSRs are advanced nuclear reactor fueled by fissile
+	  material dissolved in a molten salt mixture.
+    \item While modeling MSRs is not necessarily more difficult than modeling solid-fueled
+      reactors, we must adapt our software tools to accurately model phenomena unique to MSRs.
+  \end{itemize}
 
-  \textbf{Phenomena which pose challenges for MSR modeling}
+  \textbf{Phenomena that pose challenges for MSR modeling}
   \begin{itemize}
@@
 		\item Strong temperature feedback due to thermal expansion of liquid fuel salt
-		\item Movement of \gls{DNP} along primary coolant loop
+		\item Movement of \glspl{DNP} along primary coolant loop
         \item Advection-dominated heat transfer

### 2024-12-09 presentation/msre-transients.tex
presentation edits

@@
 \begin{frame}
-  \frametitle{V\&V Study 2: MSRE Pump Start-up \& Coast-Down Transients}
-  \textbf{Description of MSRE Pump Experiments}
+  \frametitle{Description of MSRE Pump Experiments}
   \begin{itemize}
-    \item Starting from zero-power critical states with/without forced flow, the fuel salt pump
-  was coasted down/started up
+    \item Starting from zero-power critical states under static/steady-flow conditions, the fuel salt pump
+  was started up/coasted down.
     \item Reactor was kept critical by a flux-servo controller moving the rod in response to flux
-      changes
+      changes.
   \end{itemize}
-  \begin{block}{\textbf{MSR Phenomena Involved}}
+  \begin{block}{\textbf{MSR Physics Phenomena Involved}}
     \begin{itemize}
@@
 \begin{frame}
-  \frametitle{V\&V Study 2: MSRE Pump Start-up \& Coast-Down Transients}
+  \frametitle{V\&V Study: MSRE Pump Start-up \& Coast-Down Transients}
   \begin{columns}
@@
       \begin{itemize}
-         \item Develop a V\&V study based on the MSRE pump transients
-         \item Verify Moltres against QuasiMolto in collaboration with Aaron Reynolds
-           (formerly at Oregon State University)
+         \item Develop a verification and validation (V\&V) study based on the MSRE pump transients
+         \item Verify Moltres against QuasiMolto \cite{reynolds_analysis_2023} in collaboration
+           with Dr.\ Aaron Reynolds
          \item Validate Moltres against MSRE experimental data for modeling \gls{DNP} drift
@@
         \item Ignored upper and lower plena in the geometry
-        \item Looped salt flow with a salt circulation time of 25.2 s
+%        \item Looped salt flow with a salt circulation time of 25.2 s
         \item $k$-eigenvalue calculation at every time-step ($dt=0.1$ s) in place of reactivity control
@@
 \begin{frame}
-  \frametitle{V\&V Study 2: MSRE Pump Start-up \& Coast-Down Transients}
-  \textbf{$k_\text{eff}$ Under Static \& Steady-Flow Conditions}
+  \frametitle{MSRE $k_\text{eff}$ Estimates Under Static \& Steady-Flow Conditions}
   \begin{table}[htb]
@@
 \begin{frame}
-  \frametitle{V\&V Study 2: MSRE Pump Start-up \& Coast-Down Transients}
-  \textbf{Neutron Flux Distributions Under Static Conditions}
+  \frametitle{MSRE Neutron Flux Distributions Under Static Conditions}
   \begin{figure}[h]
@@
 \begin{frame}
-  \frametitle{V\&V Study 2: MSRE Pump Start-up \& Coast-Down Transients}
-  \textbf{DNP Distributions Under Static Conditions}
-  \begin{figure}[t]
-    \centering
-    \begin{subfigure}[b]{0.48\columnwidth}
-      \centering
-      \includegraphics[width=\columnwidth]{centerline_pre}
-      \caption{Normalized centerline \gls{DNP} distributions.}
-      \label{fig:centerline-pre-dist}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[b]{0.48\columnwidth}
-      \centering
-      \includegraphics[width=\columnwidth]{midplane_pre}
-      \caption{Normalized midplane \gls{DNP} distributions.}
-      \label{fig:midplane-pre-dist}
-    \end{subfigure}
-    \caption{\gls{DNP} distribution from Moltres and ratios comparing QuasiMolto and
-    Moltres models under static \gls{MSRE} conditions. Maximum relative difference $\approx 0.1 \%$.}
-    \label{fig:centerline-pre}
-  \end{figure}
-  \begin{itemize}
-    \item Perfect overlap in centerline and midplane DNP distributions between Moltres and QuasiMolto.
-  \end{itemize}
-\end{frame}
-
-\begin{frame}
-  \frametitle{V\&V Study 2: MSRE Pump Start-up \& Coast-Down Transients}
-  \textbf{Pump Start-Up Transient Results}
+  \frametitle{MSRE Pump Start-Up Transient Results}
   \begin{columns}
@@
     \includegraphics[width=\columnwidth]{start-up-v2-reactivity}
-    \caption{Reactivity loss (relative to static no-flow conditions) during the \gls{MSRE} pump
+    \caption{Reactivity loss (relative to static conditions) during the \gls{MSRE} pump
     start-up transient.}
@@
   \column{5.5cm}
-  \textbf{Observations}
+  \textbf{Numerical Results}
   \begin{itemize}
-    \item Steep initial reactivity loss due to DNP outflow
-    \item Reactivity loss plateaus temporarily before dropping due to DNP reentry
-    \item Circulation of the excess DNP cause reactivity oscillations
-    \item Oscillations eventually dissipate due to decay of excess DNP
-    \item Excellent agreement between Moltres and QuasiMolto
+    \item Reactivity loss plateaus after excess DNPs exit.
+    \item Circulation of the excess DNPs cause reactivity oscillations that eventually dissipate.
+    \item Excellent agreement between Moltres and QuasiMolto.
+  \end{itemize}
+  \pause
+  \textbf{Discrepancies against MSRE Data}
+  \begin{itemize}
+    \item MSRE had a smaller final reactivity loss.
+    \item MSRE did not report a reactivity plateau (due to DNPs present in the upper and lower plena).
+    \item MSRE did not log significant reactivity oscillations after the first peak.
   \end{itemize}
@@
 
+%\begin{frame}
+%  \frametitle{V\&V Study 2: MSRE Pump Start-up \& Coast-Down Transients}
+%  \textbf{Pump Start-Up Transient Results}
+%  \begin{columns}
+%  \column{5.5cm}
+%  \begin{figure}[t]
+%    \centering
+%    \includegraphics[width=\columnwidth]{start-up-v2-reactivity}
+%    \caption{Reactivity loss (relative to static no-flow conditions) during the \gls{MSRE} pump
+%    start-up transient.}
+%    \label{fig:start-up-reactivity}
+%  \end{figure}
+%  \column{5.5cm}
+%  \textbf{Discrepancies against MSRE Data}
+%  \begin{itemize}
+%    \item MSRE reported a smaller final reactivity loss
+%    \item MSRE did not report a reactivity plateau (due to DNP present in the upper and lower plena)
+%    \item MSRE did not log significant reactivity oscillations after the first peak
+%  \end{itemize}
+%  \end{columns}
+%\end{frame}
+
 \begin{frame}
-  \frametitle{V\&V Study 2: MSRE Pump Start-up \& Coast-Down Transients}
-  \textbf{Pump Start-Up Transient Results}
-  \begin{columns}
-  \column{5.5cm}
+  \frametitle{Pump Coast-Down Transient Results}
   \begin{figure}[t]
     \centering
-    \includegraphics[width=\columnwidth]{start-up-v2-reactivity}
-    \caption{Reactivity loss (relative to static no-flow conditions) during the \gls{MSRE} pump
-    start-up transient.}
-    \label{fig:start-up-reactivity}
-  \end{figure}
-  \column{5.5cm}
-  \textbf{Discrepancies against MSRE Data}
-  \begin{itemize}
-    \item MSRE reported a smaller final reactivity loss
-    \item MSRE did not report a reactivity plateau (due to DNP present in the upper and lower plena)
-    \item MSRE did not log significant reactivity oscillations after the first peak
-  \end{itemize}
-  \end{columns}
-\end{frame}
-
-\begin{frame}
-  \frametitle{V\&V Study 2: MSRE Pump Start-up \& Coast-Down Transients}
-  \textbf{Pump Coast-Down Transient Results}
-  \begin{columns}
-  \column{5.5cm}
-  \begin{figure}[t]
-    \centering
-    \includegraphics[width=\columnwidth]{coast-down-v2-reactivity}
+    \includegraphics[width=.7\columnwidth]{coast-down-v2-reactivity}
     \caption{Reactivity loss (relative to static no-flow conditions) during the \gls{MSRE} pump
@@
   \end{figure}
-  \column{5.5cm}
-  \textbf{Observations}
-  \begin{itemize}
-    \item Reactivity loss remains briefly stable at the start due to flow inertia
-    \item Monotonic decline in reactivity loss over time
-  \end{itemize}
-  \end{columns}
+  Good agreement between numerical and experimental data other than the magnitude differences.
 \end{frame}
@@
 \begin{frame}
-  \frametitle{V\&V Study 2: MSRE Pump Start-up \& Coast-Down Transients}
+  \frametitle{Summary: MSRE Pump Start-up \& Coast-Down Transients}
   \begin{block}{\textbf{Study Outcome}}

### 2024-12-09 presentation/objectives.tex
slide edits
presentation edits

@@
   \begin{block}{\textbf{Overarching Goal of this Work}}
-    \textbf{Improve Moltres as a MSR simulation tool by enabling a wider range of MSR simulation
-    scenarios.}
+    Improve Moltres as a MSR simulation tool by enabling a wider range of MSR simulation
+    scenarios.
   \end{block}
@@
       \begin{itemize}
-        \item \textbf{Verify and Validate Existing Multiphysics Coupling Capabilities in Moltres}
-        \item \textbf{Implement a RANS-based Turbulence Model in Moltres}
+        \item Verify and validate existing multiphysics coupling capabilities in Moltres
+        \item Implement a RANS-based turbulence model in Moltres
       \end{itemize}
       \item \textbf{Develop a Hybrid Neutronics Method for Control Rod Modeling in Moltres}
+      \begin{itemize}
+        \item Develop and implement the hybrid $S_N$-diffusion method in Moltres
+        \item Verify the hybrid method against high-fidelity neutronics codes and MSRE
+          experimental data
+        \item Demonstrate a time-dependent reactivity-initiated simulation using the hybrid
+          method
+      \end{itemize}
     \end{enumerate}

### 2024-12-09 presentation/review.tex
presentation edits

@@
 \begin{frame}
-  \frametitle{Hybrid $S_N$-Diffusion Method: Literature Review}
-  \textbf{Transport-Correction Techniques For Neutron Diffusion-Based Solvers}
-  \vspace{.3cm}
-
+  \frametitle{Transport-Correction Techniques For Neutron Diffusion-Based Solvers}
   Techniques for augmenting the neutron diffusion method (or equivalent low-order equations) with
@@
     \item Ronen method \cite{ronen_accurate_2004, tomatis_ronen_2021, gross_comprehensive_2023}
-    \item Multilevel quasi-diffusion method \cite{goldin_quasi-diffusion_1964, anistratov_computational_2012, tamang_multilevel_2014}
+    \item Multilevel quasi-diffusion method \cite{goldin_quasi-diffusion_1964, tamang_multilevel_2014, reynolds_analysis_2023}
     \item Multischeme transport method \cite{wang_hybrid_2017}
@@
 \begin{frame}
-  \frametitle{Hybrid $S_N$-Diffusion Method: Literature Review}
-  \textbf{The Ronen Method \cite{ronen_accurate_2004, tomatis_ronen_2021, gross_comprehensive_2023}}
+  \frametitle{The Ronen Method}
   \vspace{.2cm}
@@
 \begin{frame}
-  \frametitle{Hybrid $S_N$-Diffusion Method: Literature Review}
-  \textbf{Multilevel Quasi-Diffusion Method}
-  \vspace{.2cm}
-
+  \frametitle{Multilevel Quasi-Diffusion Method}
   Gol'din originally developed the quasi-diffusion (QD) method as a transport method acceleration
@@
 \begin{frame}
-  \frametitle{Hybrid $S_N$-Diffusion Method: Literature Review}
-  \textbf{Multilevel Quasi-Diffusion Method}
-  \vspace{.2cm}
-
+  \frametitle{Multilevel Quasi-Diffusion Method}
   Tamang \& Anistratov applied the QD method to 1-D time-dependent multiphysics neutron transport
-  and heat transfer problems by iteratively coupling the gray low-order equations to the heat
+  and heat transfer problems by iteratively coupling the one-group low-order equations to the heat
   transfer and precursor equations \cite{tamang_multilevel_2014}.
@@
   \end{figure}
+  \textbf{Limitation: Computational expensive for 3-D full-core simulations.}
 \end{frame}
@@
 \begin{frame}
-  \frametitle{Hybrid $S_N$-Diffusion Method: Literature Review}
-  \textbf{Multischeme Transport Method}
-  \vspace{.2cm}
-
-  Implemented in the Griffin reactor physics application. This method involves making predetermined
+  \frametitle{Multischeme Transport Method}
+  Implemented in the Griffin reactor physics application \cite{wang_hybrid_2017}. This method
+  involves making predetermined
   domain decompositions to be treated with the $S_N$, $P_N$, or diffusion methods. Adjoining
@@
 \begin{frame}
-  \frametitle{Hybrid $S_N$-Diffusion Method: Literature Review}
-  \textbf{Hybrid Transport-Diffusion Method}
+  \frametitle{Hybrid Transport-Diffusion Method \cite{anistratov_computational_2012, stehle_hybrid_2014}}
   \vspace{.2cm}
@@
   estimates as a metric for deducing whether a region is diffusive or
-  transport-like \cite{anistratov_computational_2012, stehle_hybrid_2014}.
+  transport-like.
   \vspace{.2cm}
@@
   neighboring diffusion subdomains.
+
+  \textbf{Limitation: Complicated implementation and management of computational resources.}
 \end{frame}

### 2024-12-09 presentation/theory.tex
presentation edits

@@
 \begin{frame}
-  \frametitle{Hybrid $S_N$-Diffusion Method: Theory}
-  \textbf{Hybrid $\bm{S_N}$-Diffusion Method}
-%  \begin{block}{\textbf{Aim}}
-%    To enable accurate control rod modeling in time-dependent molten salt reactor simulations
-%  \end{block}
+  \frametitle{Hybrid $S_N$-Diffusion Method}
   \begin{block}{\textbf{Method Overview}}
@@
 \begin{frame}
-  \frametitle{Hybrid $S_N$-Diffusion Method: Theory}
+  \frametitle{Hybrid $S_N$-Diffusion Method}
   \textbf{Multigroup Discrete Ordinates $\bm{S_N}$ Neutron Transport Equations}
@@
 \begin{frame}
-  \frametitle{Hybrid $S_N$-Diffusion Method: Theory}
+  \frametitle{Hybrid $S_N$-Diffusion Method}
   \textbf{Self-Adjoint Angular Flux (SAAF) Formulation of the Multigroup $\bm{S_N}$ Equations}
@@
       \item Compatible with the efficient and scalable Hypre-BoomerAMG preconditioning
-      \item Uses a modified formulation to handle $1/\Sigma_{t,g}$ factor in void regions (similar to
+      \item Uses a modified formulation to handle $1/\Sigma_{t,g}$ factor in near-void regions (similar to
         Streamline-Upwind/Petrov Galerkin (SUPG) stabilization scheme) \cite{wang_diffusion_2014}
@@
 \begin{frame}
-  \frametitle{Hybrid $S_N$-Diffusion Method: Theory}
-  \textbf{Weak Formulation of the Multigroup SAAF $S_N$ Equations}
-  \begin{gather}
-    \shortintertext{Streaming term:}
-    \sum^G_{g=1}\sum^{N_d}_{d=1}w_d\left(\hat{\Omega}_d\cdot\nabla\Psi^*_{g,d},\tau_g\hat{\Omega}
-    \cdot\nabla\Psi_{g,d}-(1-\tau_g\Sigma_{t,g})\Psi_{g,d}\right)_\mathcal{D} \\
-    \shortintertext{Collision term: }
-    \sum^G_{g=1}\sum^{N_d}_{d=1}w_d\left(\Psi^*_{g,d},\Sigma_{t,g}\Psi_{g,d}\right)_\mathcal{D} \\
-    \shortintertext{Scattering term:}
-    \sum^G_{g=1}\sum^{N_d}_{d=1}w_d\left(\Psi^*_{g,d}+\tau_g\hat{\Omega}_d\cdot\nabla\Psi^*_{g,d},
-    \sum^G_{g'=1}\sum^L_{l=0}\Sigma^{g'\rightarrow g}_{s,l}\sum^l_{m=-l}
-    \frac{2l+1}{w}Y_{l,m}(\hat{\Omega}_d)\phi_{g',l,m}\right)_\mathcal{D} \\
-    \shortintertext{Fission source term: }
-    \sum^G_{g=1}\sum^{N_d}_{d=1}w_d\left(\Psi^*_{g,d}+\tau_g\hat{\Omega}_d\cdot\nabla\Psi^*_{g,d},
-    \frac{1}{w}\frac{\chi_{p,g}(1-\beta)}{k}\sum^G_{g'=1}\nu\Sigma_{f,g'}\phi_{g'}\right)_\mathcal{D}
-  \end{gather}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Hybrid $S_N$-Diffusion Method: Theory}
-  \textbf{Weak Formulation of the Multigroup SAAF $S_N$ Equations}
-  \begin{gather}
-    \shortintertext{Delayed neutron source term:}
-    \sum^G_{g=1}\sum^{N_d}_{d=1}w_d\left(\Psi^*_{g,d}+\tau_g\hat{\Omega}_d\cdot\nabla\Psi^*_{g,d},
-    \frac{1}{w}\sum ^I_{i=1}\chi_{d,g}\lambda_i C_i\right)_\mathcal{D}
-    \shortintertext{Boundary source term:}
-    \begin{cases}
-      \sum^G_{g=1}\sum^{N_d}_{d=1}w_d\left(\Psi^*_{g,d},
-      \hat{\Omega}_d\cdot\hat{n}_b\Psi_{g,d}\right)_{\partial\mathcal{D}},
-      & \hat{\Omega}\cdot\hat{n}_b>0,\vec{r}\in\partial\mathcal{D} \\
-      \sum^G_{g=1}\sum^{N_d}_{d=1}w_d\left(\Psi^*_{g,d},
-      \hat{\Omega}_d\cdot\hat{n}_b\Psi^\text{inc}_{g,d}\right)_{\partial\mathcal{D}},
-      & \hat{\Omega}\cdot\hat{n}_b<0,\vec{r}\in\partial\mathcal{D}
-    \end{cases} \label{eq:boundary-source} \\
-    \shortintertext{Reflecting boundary term:}
-    \begin{cases}
-      \sum^G_{g=1}\sum^{N_d}_{d=1}w_d\left(\Psi^*_{g,d},
-      \hat{\Omega}_d\cdot\hat{n}_b\Psi_{g,d}\right)_{\partial\mathcal{D}},
-      & \hat{\Omega}\cdot\hat{n}_b>0,\vec{r}\in\partial\mathcal{D}_s \\
-      \sum^G_{g=1}\sum^{N_d}_{d=1}w_d\left(\Psi^*_{g,d},
-      \hat{\Omega}_d\cdot\hat{n}_b\Psi_{g,d_r}\right)_{\partial\mathcal{D}},
-      & \hat{\Omega}\cdot\hat{n}_b<0,\vec{r}\in\partial\mathcal{D}_s
-    \end{cases} \label{eq:reflecting-bc}
-  \end{gather}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Hybrid $S_N$-Diffusion Method: Theory}
-  \textbf{Weak Formulation of the Multigroup SAAF $S_N$ Equations}
-  \begin{gather}
-    \shortintertext{Void stabilization parameter \cite{wang_diffusion_2014}:}
-    \tau_g =
-    \begin{cases}
-      \frac{1}{c\Sigma_{t,g}} \mbox{ for } ch\Sigma_{t,g} \geq \varsigma \\
-      \frac{h}{\varsigma} \mbox{ for } ch\Sigma_{t,g} < \varsigma
-    \end{cases}, \\
-    \shortintertext{where}
-    \begin{align*}
-      h &= \mbox{mesh element size,} & \\
-      c &= \mbox{maximum stabilization factor,} & \\
-      \varsigma &= \mbox{void constant.} &
-    \end{align*}
-  \end{gather}
-  $c=1$ and $\varsigma=0.5$ by default.
-  \vspace{.2cm}
-
-  The SAAF-$S_N$ equations require this stabilization scheme
-  in near-void regions where $\Sigma_{t,g}$ is very small.
-\end{frame}
-
-\begin{frame}
-  \frametitle{Hybrid $S_N$-Diffusion Method: Theory}
+  \frametitle{Hybrid $S_N$-Diffusion Method}
   \textbf{Multigroup Neutron Diffusion Equations}
@@
 \begin{frame}
-  \frametitle{Hybrid $S_N$-Diffusion Method: Theory}
+  \frametitle{Hybrid $S_N$-Diffusion Method}
   \textbf{Diffusion Correction Scheme (Implemented in Python)}
@@
 
-  In this work, I investigated two transport correction schemes. The first scheme is the diffusion
+  In this work, I investigated two transport correction schemes.
+  The first scheme is the diffusion
   correction scheme similar to Fick's first law of diffusion:
@@
 %  flux solution.
-  By replacing $D_g$ with $D^s_g$, we are effectively adding the following transport correction term
-  %
-  \begin{gather}
-    -\nabla\cdot (D^s_g-D_g)\nabla\phi_g
-  \end{gather}
-  to the neutron diffusion equations.
-  \vspace{.2cm}
+%  By replacing $D_g$ with $D^s_g$, we are effectively adding the following transport correction term
+%  %
+%  \begin{gather}
+%    -\nabla\cdot (D^s_g-D_g)\nabla\phi_g
+%  \end{gather}
+%  to the neutron diffusion equations.
+%  \vspace{.2cm}
 
@@
 \begin{frame}
-  \frametitle{Hybrid $S_N$-Diffusion Method: Theory}
+  \frametitle{Hybrid $S_N$-Diffusion Method}
   \textbf{Drift Correction Scheme (Implemented on Moltres)}
@@
   The second scheme is the drift correction scheme arising from adding a drift term
-  ($\vec{D}_g\cdot\nabla \phi_g$) to the neutron diffusion equations:
+  ($\vec{D}_g\cdot\nabla \phi_g$) \cite{wang_diffusion_2014} to the neutron diffusion equations:
   \begin{gather}
@@
 \begin{frame}
-  \frametitle{Hybrid $S_N$-Diffusion Method: Theory}
+  \frametitle{Hybrid $S_N$-Diffusion Method}
   \textbf{$S_N$ Subsolver Boundary Conditions}
@@
   the $S_N$ subproblem.
-  We may obtain
-  the best estimate by applying the $P_1$ approximation for evaluating the neutron angular flux along
-  the discrete ordinates $\hat{\Omega}_d$ of the $S_N$ angular quadrature set as follows:
+
+  The $P_1$ approximation for evaluating the neutron angular flux along
+  the discrete ordinates $\hat{\Omega}_d$ of the $S_N$ angular quadrature set is:
   \begin{align}
@@
 \begin{frame}
-  \frametitle{Hybrid $S_N$-Diffusion Method: Theory}
+  \frametitle{Hybrid $S_N$-Diffusion Method Algorithm}
   \begin{columns}
     \column{.35\textwidth}
-    \textbf{Hybrid $S_N$-Diffusion Method Algorithm}
     \vspace{.5cm}
@@
   {\small
+    \textbf{Legend}
+    \vspace{.1cm}
+
   $V_0$: Full problem domain
@@
 \begin{frame}
-  \frametitle{Hybrid $S_N$-Diffusion Method: Theory}
-  \textbf{Correction Region ($\bm{V_1}$) and Buffer Zone}
-  \begin{itemize}
-    \item Recall that the full problem domain and the correction region are defined as $V_0$
-      and $V_1$, where $V_1 \subseteq V_0$
-    \item The approximate $S_N$ boundary conditions will generally yield some deviations in the
-      flux distribution since neutron fluxes in realistic reactor systems are at least slightly
-      anisotropic throughout most of the system
-    \item However, the influence of boundary conditions on the ratio of $J$ and $\frac{d\phi}{dx}$
-      does not extend far from the boundary $\partial V_1$ in optically thick media; transport
-      correction parameters are accurate everywhere except near $\partial V_1$
-    \item There exists a buffer zone near $\partial V_1$ where transport correction parameters
-      are inaccurate
-  \end{itemize}
-%  \pause
-%  $\Rightarrow$ Approach for preliminary results: Define $V_1$ such that it is large enough to
-%  provide sufficient transport
-%  corrections and accommodate inaccurate SVDCs near $\partial V_1$. Discard inaccurate SVDCs
-%  in favor of the default $P_1$-based diffusion coefficients.
-\end{frame}
+  \frametitle{Correction Region ($V_1$) and Buffer Zone}
 
-\begin{frame}
-  \frametitle{Hybrid $S_N$-Diffusion Method: Theory}
-  \textbf{Correction Region ($\bm{V_1}$) and Buffer Zone}
+  \begin{columns}
+    \column{5.5cm}
     \begin{figure}[htb!]
       \centering
-    \includegraphics[width=.5\columnwidth]{case-3b-geometry}
+      \includegraphics[width=\columnwidth]{hybrid-illustration}
       \caption{1-D geometry for Case 3b.}
@@
     \end{figure}
+    \column{5.5cm}
+    \begin{itemize}
+      \item The approximate $S_N$ boundary conditions will yield some flux deviations near the correction
+    region boundary.
+      \item This affects transport correction parameters near the boundary.
+    \end{itemize}
+  \end{columns}
   \begin{figure}[htb!]
       \centering
-      \begin{subfigure}[t]{.49\textwidth}
+      \begin{subfigure}[t]{.46\textwidth}
           \centering
@@
       \hfill
-      \begin{subfigure}[t]{.49\textwidth}
+      \begin{subfigure}[t]{.46\textwidth}
           \centering
@@
 \begin{frame}
-  \frametitle{Hybrid $S_N$-Diffusion Method: Theory}
-  \textbf{Correction Region ($\bm{V_1}$) and Buffer Zone}
-  \vspace{.2cm}
+  \frametitle{Correction Region ($V_1$) and Buffer Zone}
 
-  A natural/intuitive criterion for the location of the buffer zone cutoff boundary
+  A natural/intuitive criterion for the location of the buffer region cutoff boundary
   would be wherever the components of the drift correction variable $\vec{D}_g$ is zero, i.e.,
@@
       \centering
-      \begin{subfigure}[t]{.49\textwidth}
+      \begin{subfigure}[t]{.46\textwidth}
           \centering
@@
 \begin{frame}
-  \frametitle{Hybrid $S_N$-Diffusion Method: Theory}
+  \frametitle{Hybrid $S_N$-Diffusion Method}
   \textbf{Numerical Implementation}

### 2024-12-09 presentation/turb-moltres.tex
presentation edits

--- 2024-12-04 presentation/turb-moltres.tex
@@
   \frametitle{Spalart-Allmaras Turbulence Model in Moltres}
-  \begin{block}{\textbf{Why the Spalart-Allmaras Model for Moltres?}}
-    A computationally efficient one-equation model that provides reasonable estimates for turbulent
-    effects in MSRs and is tractable for large problems on small computing clusters.
+  \begin{block}{\textbf{Why the Spalart-Allmaras Model \cite{spalart_one-equation_1994} for Moltres?}}
+    A computationally efficient one-equation model that provides reasonable estimates for wall-bounded
+    turbulent flows and is tractable for large problems on small computing clusters.
   \end{block}
@@
     \item Built on top of an existing incompressible Navier-Stokes model in MOOSE
-    \item Uses \gls{SUPG} stabilization scheme for numerical stability
-    \item $f_{t2}$ turbulence trip term is togglable (Negligible difference in fully turbulent flows)
+    \item Uses Streamline-Upwind/Petrov-Galerkin (SUPG) stabilization scheme for numerical stability
     \item Rotation correction scheme \cite{aupoix_extensions_2003, dacles-mariani_numericalexperimental_1995}
@@
         \item Reference results: Experimental results by Driver \& Seegmiller \cite{driver_features_1985}
-        \item Reference results: Spalart-Allmaras model results from CFL3D CFD code \cite{krist_cfl3d_1998}
+        \& Spalart-Allmaras model results from CFL3D CFD code \cite{krist_cfl3d_1998}
       \end{itemize}
@@
 \begin{frame}
-  \frametitle{Spalart-Allmaras Turbulence Model in Moltres}
-  \textbf{Turbulent Channel Flow Verification Test Results}
+  \frametitle{Turbulent Channel Flow Verification Test Results}
   \begin{figure}[htb!]
@@
     \caption{Comparison of turbulent channel flow results at Re$_\tau\approx395$ against reference
-    \gls{DNS} results \cite{moser_direct_1999}. The Moltres Spalart-Allmaras model agrees consistently with
-    the reference data.}
+    \gls{DNS} results \cite{moser_direct_1999}.}
     \label{fig:channel-verification}
   \end{figure}
+  \vspace{.1cm}
+
+  $\Rightarrow$ The Moltres Spalart-Allmaras model results are consistent with reference DNS channel flow data.
 \end{frame}
@@
 \begin{frame}
-  \frametitle{Spalart-Allmaras Turbulence Model in Moltres}
-  \textbf{Turbulent Pipe Flow Validation Test Results}
+  \frametitle{Turbulent Pipe Flow Validation Test Results}
   \begin{figure}[htb]
@@
   \end{figure}
+  \vspace{.1cm}
+
+  $\Rightarrow$ The Moltres Spalart-Allmaras model results are consistent with reference experimental pipe flow data.
 \end{frame}
@@
 \begin{frame}
-  \frametitle{Spalart-Allmaras Turbulence Model in Moltres}
-  \textbf{Backward-Facing Step Flow Validation Test Results}
+  \frametitle{Backward-Facing Step Flow Validation Test Setup}
   \begin{columns}
@@
 \begin{frame}
-  \frametitle{Spalart-Allmaras Turbulence Model in Moltres}
-  \textbf{Backward-Facing Step Flow Validation Test Results}
+  \frametitle{Backward-Facing Step Flow Validation Test Results}
   \begin{columns}
@@
     \end{table}
+    \vspace{.1cm}
+
+    $\Rightarrow$ Moltres is consistent with experimental data (within uncertainty range) while CFL3D
+    falls outside the uncertainty range.
   \end{columns}
@@
 \begin{frame}
-  \frametitle{Spalart-Allmaras Turbulence Model in Moltres}
-  \textbf{Backward-Facing Step Flow Validation Test Results}
+  \frametitle{Backward-Facing Step Flow Validation Test Results}
+  \begin{columns}
+    \column{11.5cm}
     \begin{figure}[htb]
       \centering
-    \hfill
-    \begin{subfigure}[b]{0.24\columnwidth}
+      \begin{subfigure}[t]{0.32\columnwidth}
         \centering
@@
       \end{subfigure} %\hfill \\
-%    \centering
-    \hfill
-    \begin{subfigure}[b]{0.24\columnwidth}
+      \begin{subfigure}[t]{0.32\columnwidth}
         \centering
@@
       \end{subfigure}
-    \hfill
-    \begin{subfigure}[b]{0.24\columnwidth}
-      \centering
-      \includegraphics[width=\columnwidth]{bfs_cp}
-      \caption{Skin pressure coefficient along the bottom wall.}
-      \label{fig:bfs-cp}
-    \end{subfigure}% \hfill \\
-%    \centering
-    \begin{subfigure}[b]{0.24\columnwidth}
+      \begin{subfigure}[t]{0.32\columnwidth}
         \centering
@@
     \end{figure}
+  \end{columns}
+  \begin{itemize}
+    \item Minor discrepancies between numerical and experimental data typical of RANS-based turbulence models.
+    \item Moltres provides more accurate measurements along $x/H=1$ than CFL3D.
+  \end{itemize}
 \end{frame}
@@
   \frametitle{Spalart-Allmaras Turbulence Model in Moltres}
-  \textbf{Spalart-Allmaras Model V\&V Test Summary}
+  \begin{block}{\textbf{Spalart-Allmaras Model V\&V Test Summary}}
     \begin{itemize}
       \item Good agreement between Moltres and reference solutions in the turbulent channel and pipe
-      flow tests
-    \item Good agreement between Moltres and CFL3D in the backward-facing step flow test
+        flow tests.
+      \item Good agreement between Moltres and CFL3D in the backward-facing step flow test.
       \item Some discrepancy between Moltres and experimental data in the backward-facing step flow

### 2024-12-09 presentation/turb-motivation.tex
presentation edits

\begin{frame}
  \frametitle{Motivation for Turbulence Modeling Implementation}
  \textbf{Turbulent Flows in MSRs}
  \begin{itemize}
    \item The Reynolds number of salt flow in MSRs can go up to $10^6$ in the MSFR core
      \cite{fiorina_modelling_2014}.
    \item Turbulent effects expected in MSRs
    \begin{itemize}
      \item Turbulent diffusion of heat and \gls{DNP}
      \item Flow separation and recirculation zones due to abrupt geometry changes
    \end{itemize}
    \item Temperature hotspots may induce thermal stress on structural
      components and temperature-induced reactivity effects.
  \end{itemize}
  \pause
  \begin{block}{\textbf{Area of Improvement for Moltres for MSR Modeling}}
    Moltres does not currently have turbulence modeling capability for simulating
    turbulent flows in MSRs.
  \end{block}
\end{frame}

### 2024-12-12 presentation/1d.tex
slide edits

@@
     \item OpenMC in multigroup mode (OpenMC-MG)
-    \item Standard neutron diffusion method (Moltres \& Python)
+    \item Neutron diffusion method (Moltres \& Python)
     \item $S_8$ method (Moltres \& Python)
@@
 \begin{frame}
-%  \frametitle{Hybrid $S_N$-Diffusion Method: 1-D Neutronics Eigenvalue Simulations}
-  \begin{columns}
-    \column{5.5cm}
+  \frametitle{1-D MSRE Neutronics Simulation Results}
+  \textbf{Case 3b Geometry}
+  \begin{figure}
+    \centering
+    \includegraphics[width=.7\textwidth]{hybrid-illustration}
+    \caption{Geometry of 1-D Case 3b.}
+  \end{figure}
+\end{frame}
+
+\begin{frame}
+  \frametitle{1-D MSRE Neutronics Simulation Results}
   \textbf{Case 3b Neutron Flux Distributions}
-    \begin{itemize}
-      \item The neutron diffusion and hybrid methods fare worse than the $S_8$ method at capturing
-        the oscillatory flux pattern.
-      \item The hybrid method performs better than the neutron diffusion method near $x=0$ cm where
-        the control rod is situated.
-    \end{itemize}
-    \column{5.5cm}
   \begin{figure}[htb!]
     \centering
-      \includegraphics[width=.975\columnwidth]{case-3b-flux-diff}
+    \includegraphics[width=.75\columnwidth]{images/case-3b-flux-diff-pres}
     \caption{Absolute difference in neutron group flux distributions for Case 3b from Moltres-$S_8$,
-      Moltres-diffusion, Moltres-hybrid, and Python-hybrid relative to OpenMC-MG.}
+    Moltres-diffusion, and Moltres-hybrid relative to OpenMC-MG.}
     \label{fig:3b-flux-diff}
   \end{figure}
-  \end{columns}
+%  \begin{itemize}
+%    \item The neutron diffusion and hybrid methods fare worse than the $S_8$ method at capturing
+%      the oscillatory flux pattern.
+    The hybrid method provides more accurate flux estimates than the neutron diffusion method near $x=0$ cm where
+      the control rod is situated.
+%  \end{itemize}
 \end{frame}
+
+%\begin{frame}
+%%  \frametitle{Hybrid $S_N$-Diffusion Method: 1-D Neutronics Eigenvalue Simulations}
+%  \begin{columns}
+%    \column{5.5cm}
+%    \textbf{Case 3b Neutron Flux Distributions}
+%    \begin{itemize}
+%      \item The neutron diffusion and hybrid methods fare worse than the $S_8$ method at capturing
+%        the oscillatory flux pattern.
+%      \item The hybrid method performs better than the neutron diffusion method near $x=0$ cm where
+%        the control rod is situated.
+%    \end{itemize}
+%    \column{5.5cm}
+%    \begin{figure}[htb!]
+%      \centering
+%      \includegraphics[width=.975\columnwidth]{case-3b-flux-diff}
+%      \caption{Absolute difference in neutron group flux distributions for Case 3b from Moltres-$S_8$,
+%      Moltres-diffusion, Moltres-hybrid, and Python-hybrid relative to OpenMC-MG.}
+%      \label{fig:3b-flux-diff}
+%    \end{figure}
+%  \end{columns}
+%\end{frame}
 
@@
     \item Rod worth estimates vary non-monotonically with increasing correction subregion size.
-    \item Rod worth estimates remain within 0.2 \% of the $S_8$ method rod worth.
+    \item Rod worth estimates remain within 0.2\% of the $S_8$ method rod worth.
     \item The hybrid method produces accurate rod worth estimates as long as the correction region
@@
     \caption{Percentage difference in rod worth from the hybrid method relative to OpenMC-CE for
-      Cases 3a and 3b with different correction subregion sizes. The horizontal lines indicate
-      equivalent rod worth differences from the OpenMC-MG and $S_8$ methods.}
+      Cases 3a and 3b with different correction subregion sizes.}
     \label{fig:v1-size-rho}
@@
   \begin{columns}
-    \column{7cm}
-    \textbf{Relaxing the $S_N$ Convergence Tolerance}
+    \column{6.5cm}
+    \textbf{Relaxing the $S_N$ Convergence Tolerance,} $\bm{\epsilon}_\text{tol}$
     \begin{itemize}
@@
         \toprule
-        $S_8$ subsolver tolerance, $\epsilon_\text{tol}$ & {$10^{-8}$} & {$10^{-7}$} & {$10^{-6}$} & {$10^{-5}$} & {$10^{-4}$} & {$10^{-3}$} \\
+        Tolerance value, $\epsilon_\text{tol}$ & {$10^{-8}$} & {$10^{-7}$} & {$10^{-6}$} & {$10^{-5}$} & {$10^{-4}$} & {$10^{-3}$} \\
         \midrule
-        Number of outer iterations & 3 & 3 & 3 & 2 & 2 & 1 \\
+        No.\ of outer iterations & 3 & 3 & 3 & 2 & 2 & 1 \\
         \bottomrule
@@
     \end{table}
-    \column{4cm}
+    \column{4.5cm}
     \begin{figure}[h]
@@
       \includegraphics[width=\columnwidth]{sn-tol}
-      \caption{$k_\text{eff}$ error estimates of Case 3b for a range of convergence tolerance values
-      imposed on the $S_8$ subsolver relative to the reference $k_\text{eff}$ value when
+      \caption{$k_\text{eff}$ error estimates of Case 3b for a range of $S_8$ subsolver convergence tolerance values
+      relative to the reference $k_\text{eff}$ value when
       $\epsilon_\text{tol}=10^{-8}$.}

### 2024-12-12 presentation/2d.tex
slide edits

@@
   \frametitle{2-D MSRE Neutronics Simulation Results}
-  \textbf{2-D Quarter-Core $k$ \& Rod Worth Results}
+  \textbf{2-D Quarter-Core $\bm{k}$ \& Rod Worth Results}
   \begin{table}[htb]
@@
   \frametitle{2-D MSRE Neutronics Simulation Results}
-  \textbf{2-D Quarter-Core Normalized Fuel Channel Power Distribution}
+  \textbf{2-D Quarter-Core Normalized Channel Fission Rate Distribution}
   \begin{table}[htb]
@@
   \frametitle{2-D MSRE Neutronics Simulation Results}
-  \textbf{2-D Full-Core Normalized Fuel Channel Power Distribution}
+  \textbf{2-D Full-Core Normalized Channel Fission Rate Distribution}
   \begin{table}[htb]
@@
   \begin{itemize}
-    \item Mean percentage error for the hybrid method remains consistent at 0.43 \%.
+    \item Mean percentage error for the hybrid method remains consistent at 0.43\%.
     \item Significant improvement in maximum percentage error of the channel fission rate.

### 2024-12-12 presentation/3d-transient.tex
presentation edits
slide edits

@@
 \begin{frame}
-  \frametitle{MSRE Rod Drop Experiment}
-  \textbf{Rod Height}
-  \begin{figure}[htb!]
-    \centering
-    \includegraphics[width=0.7\columnwidth]{rod-height}
-    \caption{Evolution of rod height evaluated at each timestep within the first second of the rod
-    drop simulation. The rod reaches its fully inserted height at $t=0.4046$ s.}
-    \label{fig:rod-height}
-  \end{figure}
-\end{frame}
-
-\begin{frame}
-  \frametitle{MSRE Rod Drop Experiment}
+  \frametitle{MSRE Rod Drop Modeling Approach}
   \textbf{Nested Coupling Iteration Structure for the Rod Drop Simulation}
@@
 \begin{frame}
-  \frametitle{MSRE Rod Drop Experiment}
+  \frametitle{MSRE Rod Drop Simulation Results}
   \textbf{Neutron Count Rate Following Rod Drop}
@@
       \item Convergence issues prevented the simulation from converging from $t = 0.8$ s
-      \item I raised the convergence tolerance to help the simulation continue
+      \item Raised the convergence tolerance value after $t=0.8$ s to help the simulation continue
     \end{itemize}
@@
 \begin{frame}
-  \frametitle{MSRE Rod Drop Experiment}
+  \frametitle{MSRE Rod Drop Simulation Results}
   \textbf{Integral Neutron Count Following Rod Drop}
@@
     \begin{itemize}
-      \item Moltres reproduces the expected trend in the integral neutron count rate
+      \item Moltres reproduces the expected trend in the integral neutron count rate.
       \item Slight underprediction relative to MSRE rod drop experimental data
@@
   \begin{itemize}
-    \item Poor convergence rate relative to 1-D \& 2-D modeling
+    \item Slow convergence rate relative to 1-D \& 2-D modeling
     \begin{itemize}

### 2024-12-12 presentation/3d.tex
slide edits

@@
   \end{table}
-  \vspace{.2cm}
+  \vspace{.1cm}
 
@@
     \item The hybrid method took about four times longer than the neutron diffusion method.
-    \item 30 \% of compute time spent on data transfers between the neutron diffusion and $S_N$
-      subsolvers (Reduced to 18 \% in subsequent simulations after optimizations)
+    \item 30\% of compute time spent on data transfers between the neutron diffusion and $S_N$
+      subsolvers (reduced to 18\% in subsequent simulations after optimizations).
   \end{itemize}

### 2024-12-12 presentation/backup.tex
slide edits

@@
 \begin{frame}[noframenumbering]
-  \frametitle{Backward-Facing Step Flow Validation Test Results}
+  \frametitle{Backward-Facing Step Flow Validation Test Setup}
   \begin{columns}
-    \column{11.5cm}
-    \begin{figure}[htb]
+    \column{5.5cm}
+    \begin{figure}[h]
       \centering
-      \hfill
-      \begin{subfigure}[t]{0.24\columnwidth}
+      \includegraphics[width=\columnwidth]{backstep-geom-pres}
+      \caption{Backward step geometry for the turbulent BFS flow verification test. The red box indicates
+      the region shown by the close-up view in Figure \ref{fig:bfs-mesh}.}
+      \label{fig:backstep-geom}
+    \end{figure}
+    \column{5.5cm}
+    \begin{figure}[h]
       \centering
-        \includegraphics[width=\columnwidth]{bfs_downstream_vel}
-        \caption{Normalized velocity distributions downstream of step.}
-        \label{fig:bfs-downstream}
-      \end{subfigure} %\hfill \\
-  %    \centering
-      \hfill
-      \begin{subfigure}[t]{0.24\columnwidth}
-        \centering
-        \includegraphics[width=\columnwidth]{bfs_cf}
-        \caption{Skin friction coefficient along the bottom wall.}
-        \label{fig:bfs-cf}
-      \end{subfigure}
-      \hfill
-      \begin{subfigure}[t]{0.24\columnwidth}
-        \centering
-        \includegraphics[width=\columnwidth]{bfs_cp}
-        \caption{Skin pressure coefficient along the bottom wall.}
-        \label{fig:bfs-cp}
-      \end{subfigure}% \hfill \\
-  %    \centering
-      \begin{subfigure}[t]{0.24\columnwidth}
-        \centering
-        \includegraphics[width=\columnwidth]{bfs_stress}
-        \caption{Normalized turbulent shear stress downstream
-        of step.}
-        \label{fig:bfs-stress}
-      \end{subfigure}
-      \caption{Comparison of backward facing step flow results against reference
-      experimental data and computational data from CFL3D. $x/H$ values are normalized horizontal
-      distances relative to the step.}
-      \label{fig:bfs-plots}
+      \includegraphics[width=\columnwidth]{bfs_mesh}
+      \caption{Close-up view of the mesh for the BFS flow test. The step is situated at $x=110$ m
+      with a height of $H=1$ m.}
+      \label{fig:bfs-mesh}
     \end{figure}
   \end{columns}
-  \begin{itemize}
-    \item Minor discrepancies between numerical and experimental data typical of RANS-based turbulence models.
-    \item Moltres provides better predictments along $x/H=1$ than CFL3D.
-  \end{itemize}
 \end{frame}
@@
 \begin{frame}[noframenumbering]
+  \frametitle{1-D Neutronics Model Mesh Convergence Tests}
+  \begin{figure}[t]
+    \centering
+    \begin{subfigure}[b]{0.48\columnwidth}
+      \centering
+      \includegraphics[width=\columnwidth]{diffusion-mesh-convergence-k}
+      \caption{Neutron diffusion method}
+      \label{fig:diff-mesh-k}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[b]{0.48\columnwidth}
+      \centering
+      \includegraphics[width=\columnwidth]{sn-mesh-convergence-k}
+      \caption{$S_8$ neutron transport method}
+      \label{fig:sn-mesh-k}
+    \end{subfigure}
+    \caption{Convergence of multiplication factor ($k_\text{eff}$) estimates for Case 3b across four
+      levels of mesh refinement relative to the finest mesh resolution.}
+    \label{fig:mesh-k}
+  \end{figure}
+\end{frame}
+
+\begin{frame}[noframenumbering]
 %  \frametitle{Hybrid $S_N$-Diffusion Method: 1-D Neutronics Eigenvalue Simulations}
@@
   \end{columns}
+\end{frame}
+
+\begin{frame}[noframenumbering]
+  \begin{figure}[t]
+    \centering
+    \includegraphics[width=.5\columnwidth]{case-3b-drift}
+    \caption{Multigroup drift correction ($\vec{D}_g$) $x$-component distributions from the
+    Moltres-hybrid and Moltres-$S_8$ solvers.}
+    \label{fig:3b-drift}
+  \end{figure}
 \end{frame}
@@
   \end{itemize}
+\end{frame}
+
+\begin{frame}[noframenumbering]
+  \begin{columns}
+    \column{5.5cm}
+    \begin{figure}[t]
+      \small
+      \centering
+      \includegraphics[width=.75\columnwidth]{ray-effect-12}
+      \caption{Group 1 and 2 neutron flux distributions in the hybrid $S_N$-diffusion method correction
+      region with $S_6$, $S_8$, $S_{10}$, \& $S_{12}$.}
+      \label{fig:ray-effect-12}
+    \end{figure}
+    \column{5.5cm}
+    \begin{figure}[t]
+      \small
+      \centering
+      \includegraphics[width=.75\columnwidth]{ray-effect-34}
+      \caption{Group 3 and 4 neutron flux distributions in the hybrid $S_N$-diffusion method correction
+      region with $S_6$, $S_8$, $S_{10}$, \& $S_{12}$.}
+      \label{fig:ray-effect-34}
+    \end{figure}
+  \end{columns}
 \end{frame}
@@
 \begin{frame}[noframenumbering]
+  \begin{figure}[t]
+    \centering
+    \includegraphics[width=.6\columnwidth]{3d-view-high-res}
+    \caption{3-D section view of the \gls{MSRE} model geometry with Rod 1 inserted by 4.4 inches at
+    initial criticality.}
+    \label{fig:3d-view}
+  \end{figure}
+\end{frame}
+
+\begin{frame}[noframenumbering]
+  \begin{figure}[t]
+    \centering
+    \includegraphics[width=.6\columnwidth]{ic-g1}
+    \caption{Group 1 neutron flux distribution with Rod 1 inserted by 4.4 inches at initial
+    criticality.}
+    \label{fig:ic-g1}
+  \end{figure}
+\end{frame}
+
+\begin{frame}[noframenumbering]
+  \begin{figure}[t]
+    \centering
+    \includegraphics[width=.6\columnwidth]{ic-g8}
+    \caption{Group 8 neutron flux distribution with Rod 1 inserted by 4.4 inches at initial
+    criticality.}
+    \label{fig:ic-g8}
+  \end{figure}
+\end{frame}
+
+\begin{frame}[noframenumbering]
   \frametitle{MSRE Rod Drop Experiment}
@@
     \label{fig:rod-height}
+  \end{figure}
+\end{frame}
+
+\begin{frame}[noframenumbering]
+  \frametitle{MSRE Rod Drop Simulation Interpolated Results}
+  \begin{columns}
+    \column{5.5cm}
+    \centerline{\large \textbf{Neutron Count Rate}}
+    \begin{figure}[t]
+      \centering
+      \includegraphics[width=\columnwidth]{count-rate-interp}
+      \caption{Neutron count rate during the rod drop experiment with linearly
+      interpolated data between $t=1.325$ s and $t=3.825$ s.}
+      \label{fig:count-rate}
+    \end{figure}
+    \column{5.5cm}
+    \centerline{\large \textbf{Integral Neutron Count}}
+    \begin{figure}[t]
+      \centering
+      \includegraphics[width=\columnwidth]{integral-count-interp}
+      \caption{Integral neutron count during the rod drop experiment with linearly interpolated
+      count rate data between $t=1.325$ s and $t=3.825$ s.}
+      \label{fig:count-rate}
+    \end{figure}
+  \end{columns}
+\end{frame}
+
+\begin{frame}[noframenumbering]
+  \frametitle{Rod Cusping Correction}
+  \begin{figure}[t]
+      \centering
+      \begin{subfigure}[t]{.49\textwidth}
+          \centering
+          \includegraphics[width=\textwidth]{rod-cusping-1}
+          \caption{Uncorrected volume weighting}
+          \label{fig:rod-cusping-1}
+      \end{subfigure}
+      \hfill
+      \begin{subfigure}[t]{.49\textwidth}
+          \centering
+          \includegraphics[width=\textwidth]{rod-cusping-2}
+          \caption{Corrected volume weighting}
+          \label{fig:rod-cusping-2}
+      \end{subfigure}
+      \caption{Fraction of reactivity change against control rod volume fraction in mixed mesh
+      elements with the rod inserted at mid-reactor height (27.4 to 29.4 inches) and full insertion
+      height (0.0 to 1.9 inches).}
+      \label{fig:rod-cusping}
   \end{figure}

### 2024-12-12 presentation/conclusion.tex
slide edits

@@
-%\begin{frame}
-%  \frametitle{Conclusion}
-%  \begin{block}{\textbf{Objectives}}
-%    \begin{enumerate}
-%      \item \textbf{Improve Multiphysics Capabilities in Moltres}
-%      \begin{itemize}
-%        \item \textbf{Verify and Validate Existing Multiphysics Coupling Capabilities in Moltres}
-%        \item \textbf{Implement a RANS-based Turbulence Model in Moltres}
-%      \end{itemize}
-%      \item \textbf{Develop a Hybrid Neutronics Method for Control Rod Modeling in Moltres}
-%    \end{enumerate}
-%  \end{block}
-%  \begin{block}{\textbf{Research Impact}}
-%    \begin{itemize}
-%      \item \textbf{Improve Moltres as a simulation tool for MSR multiphysics modeling}
-%      \item \textbf{Allow for a wider range of MSR simulation scenarios on Moltres involving
-%        turbulent flow and control rod movement}
-%      \item \textbf{Analysis of neutron transport effects in MSRs that are neglected by neutron
-%        diffusion methods}
-%%      \item \textbf{Publication of numerical data for MSRE control rod insertion/withdrawal
-%%        simulations}
-%    \end{itemize}
-%  \end{block}
-%\end{frame}
-
 \begin{frame}
@@
     \begin{itemize}
-      \item Existing multiphysics coupling capabilities in Moltres must be verified and validated
+      \item Existing multiphysics coupling capabilities in Moltres must be verified and validated.
     \end{itemize}
@@
       \item Moltres requires turbulence modeling capability to model turbulent salt
-        flow in MSRs
+        flow in MSRs.
     \end{itemize}
@@
       \item The Spalart-Allmaras turbulence model in Moltres is accurate for turbulent channel and
-        pipe flows
+        pipe flows.
       \item It is also consistent with other implementations when simulating the
@@
       \item The hybrid method is an iterative method that applies a $S_N$ subsolver around control
-        rod regions for improved rod worth estimates
+        rod regions for improved rod worth estimates.
       \item The hybrid method applies an adaptive boundary coupling between
-        the $S_N$ subsolver and the neutron diffusion solver to smooth flux gradients
+        the $S_N$ subsolver and the neutron diffusion solver to smooth flux gradients.
       \item The hybrid method accurately reproduces control rod worths at only four times the
-        computational cost of the neutron diffusion method
-      \item The hybrid method exhibits good scalability on high-performance computing clusters
+        computational cost of the neutron diffusion method.
+      \item The hybrid method exhibits good scalability on high-performance computing clusters.
     \end{itemize}

### 2024-12-12 presentation/review.tex
slide edits

@@
 \begin{frame}
-  \frametitle{Transport-Correction Techniques For Neutron Diffusion-Based Solvers}
+  \frametitle{Literature Review}
+  \textbf{Transport-Correction Techniques For Neutron Diffusion-Based Solvers}
+  \vspace{.2cm}
+
   Techniques for augmenting the neutron diffusion method (or equivalent low-order equations) with
@@
   \end{figure}
-  \textbf{Limitation: Computational expensive for 3-D full-core simulations.}
+  \textbf{Limitation: Computationally expensive for 3-D full-core simulations.}
 \end{frame}
@@
   neighboring diffusion subdomains.
+  \vspace{.2cm}
 
@@
 \begin{frame}
-  \frametitle{Hybrid $S_N$-Diffusion Method: Literature Review}
+  \frametitle{Literature Review}
   \begin{block}{\textbf{Summary}}

### 2024-12-12 presentation/turb-moltres.tex
slide edits

@@
+%\begin{frame}
+%  \frametitle{Spalart-Allmaras Turbulence Model in Moltres}
+%  \begin{block}{\textbf{Why the Spalart-Allmaras Model \cite{spalart_one-equation_1994} for Moltres?}}
+%    A computationally efficient one-equation model that provides reasonable estimates for wall-bounded
+%    turbulent flows and is tractable for large problems on small computing clusters.
+%  \end{block}
+%  \textbf{Spalart-Allmaras Model with Rotation Correction Scheme
+%  \cite{spalart_one-equation_1994, aupoix_extensions_2003, dacles-mariani_numericalexperimental_1995}}\\
+%  \\
+%  Governing equation for the modified dynamic viscosity, $\tilde{\mu}$:
+%  \begin{multline}
+%    \rho \frac{\partial\tilde{\mu}}{\partial t} + \rho \mathbf{u}\cdot\nabla\tilde{\mu} = \rho c_{b1}
+%    \left(1-f_{t2}\right)\left[\tilde{S}+C_\text{rot}\min(0,S-\Omega)\right]\tilde{\mu} \\ + \frac{1}{\sigma}\{\nabla\cdot\left[\left(\mu+
+%    \tilde{\mu}\right)\nabla\tilde{\mu}\right] + c_{b2}|\nabla\tilde{\mu}|^2\}
+%    - \left(c_{w1}f_w - \frac{c_{b1}}{\kappa^2}f_{t2}\right)\left(\frac{\tilde{\mu}}{d}\right)^2
+%  \end{multline}
+%  \begin{align*}
+%    \shortintertext{where}
+%      \mu_t &= \tilde{\mu}f_{v1} = \text{turbulent eddy viscosity}, \\
+%      f_{v1} &= \frac{\chi^3}{\chi^3 + c_{v1}^3}, \\
+%      \chi &= \frac{\tilde{\mu}}{\mu}.
+%  \end{align*}
+%\end{frame}
+
 \begin{frame}
@@
     \rho \frac{\partial\tilde{\mu}}{\partial t} + \rho \mathbf{u}\cdot\nabla\tilde{\mu} = \rho c_{b1}
-    \left(1-f_{t2}\right)\tilde{S}\tilde{\mu} + \frac{1}{\sigma}\{\nabla\cdot\left[\left(\mu+
-    \tilde{\mu}\right)\nabla\tilde{\mu}\right] + c_{b2}|\nabla\tilde{\mu}|^2\} \\
+    \left(1-f_{t2}\right)\left[\tilde{S}+C_\text{rot}\min(0,S-\Omega)\right]\tilde{\mu} \\ + \frac{1}{\sigma}\{\nabla\cdot\left[\left(\mu+
+    \tilde{\mu}\right)\nabla\tilde{\mu}\right] + c_{b2}|\nabla\tilde{\mu}|^2\}
     - \left(c_{w1}f_w - \frac{c_{b1}}{\kappa^2}f_{t2}\right)\left(\frac{\tilde{\mu}}{d}\right)^2
@@
   \end{align*}
-%      \tilde{S} &= \Omega + \frac{\tilde{\nu}}{\kappa^2 d^2} f_{v2}, \\
-%      f_{v2} &= 1 - \frac{\chi}{1+\chi f_{v1}}, \\
-%      \Omega &= \sqrt{2W_{ij}W_{ij}} = \text{vorticity magnitude}, \\
-%      W_{ij} &= \frac{1}{2}\left(\frac{\partial u_i}{\partial x_j} - \frac{\partial u_j}{\partial x_i}
-%      \right), \\
-%        f_w &= g\left(\frac{1 + c_{w3}^6}{g^6 + c_{w3}^6}\right)^{1/6}, \\
-%        g &= r + c_{w2}\left(r^6 - r\right), \\
-%        r &= \text{min}\left(\frac{\tilde{\nu}}{\tilde{S}\kappa^2d^2}, 10\right), \\
-%        f_{t2} &= c_{t3} \exp{\left(-c_{t4}\chi^2\right)},
-%    \end{align*}
-%  \shortintertext{and the constants are}
-%    \sigma = \frac{2}{3}, \ c_{b1} = 0.1355, \ c_{b2} = 0.622, \ \kappa = 0.41, \
-%    c_{w1} = \frac{c_{b1}}{\kappa^2} + \frac{1+c_{b2}}{\sigma}, \nonumber \\
-%    c_{w2} = 0.3, \ c_{w3} = 2, \
-%    c_{v1} = 7.1, \ c_{t3} = 1.2, \ c_{t4} = 0.5 \ \text{.} \nonumber
 \end{frame}
@@
 
-\begin{frame}
-  \frametitle{Backward-Facing Step Flow Validation Test Setup}
-  \begin{columns}
-    \column{5.5cm}
-    \begin{figure}[h]
-      \centering
-      \includegraphics[width=\columnwidth]{backstep-geom-pres}
-      \caption{Backward step geometry for the turbulent BFS flow verification test. The red box indicates
-      the region shown by the close-up view in Figure \ref{fig:bfs-mesh}.}
-      \label{fig:backstep-geom}
-    \end{figure}
-    \column{5.5cm}
-    \begin{figure}[h]
-      \centering
-      \includegraphics[width=\columnwidth]{bfs_mesh}
-      \caption{Close-up view of the mesh for the BFS flow test. The step is situated at $x=110$ m
-      with a height of $H=1$ m.}
-      \label{fig:bfs-mesh}
-    \end{figure}
-  \end{columns}
-\end{frame}
+%\begin{frame}
+%  \frametitle{Backward-Facing Step Flow Validation Test Setup}
+%  \begin{columns}
+%    \column{5.5cm}
+%    \begin{figure}[h]
+%      \centering
+%      \includegraphics[width=\columnwidth]{backstep-geom-pres}
+%      \caption{Backward step geometry for the turbulent BFS flow verification test. The red box indicates
+%      the region shown by the close-up view in Figure \ref{fig:bfs-mesh}.}
+%      \label{fig:backstep-geom}
+%    \end{figure}
+%    \column{5.5cm}
+%    \begin{figure}[h]
+%      \centering
+%      \includegraphics[width=\columnwidth]{bfs_mesh}
+%      \caption{Close-up view of the mesh for the BFS flow test. The step is situated at $x=110$ m
+%      with a height of $H=1$ m.}
+%      \label{fig:bfs-mesh}
+%    \end{figure}
+%  \end{columns}
+%\end{frame}
 

### 2025-01-03 abstract.tex
Edits before format review

@@
 This work builds on existing capabilities in Moltres, a MOOSE-based \gls{MSR} simulation software,
-to tackle these challenges and support efforts towards \gls{MSR} deployment.
+to address these challenges and support efforts towards \gls{MSR} deployment.
 
@@
 %simulations also performed well in reproducing expected trends observed in experimental data.
-A demonstration of the hybrid method for a time-dependent rod drop
-simulation also performed well in reproducing expected trends observed in experimental data.
+The hybrid method also performed well in a time-dependent rod drop simulation
+by reproducing expected trends observed in experimental data.
 Analysis of its computational performance indicated the possibility of further optimizations

### 2025-01-03 appendix.tex
Full draft + edits
Edits before format review

@@
-\chapter{Numerical Implementation of Neutronics Solvers on Python} \label{chap:implementation}
+\chapter{Numerical Implementation of Neutronics Solvers in Python} \label{chap:implementation}
 
@@
 \end{figure}
+
+\chapter{3-D \gls{MSRE} Neutron Flux Distributions}
+
+This appendix contains figures showing the group 1 and 8 neutron flux distributions in the 3-D
+\gls{MSRE} full-core model with Rod 1 inserted by 27.5 and 51 inches.
+
+\begin{figure}[p]
+  \centering
+  \includegraphics[width=\columnwidth]{halfway-g1}
+  \caption{Group 1 neutron flux distribution with Rod 1 inserted by 27.5 inches.}
+  \label{fig:halfway-g1}
+\end{figure}
+
+\begin{figure}[p]
+  \centering
+  \includegraphics[width=\columnwidth]{halfway-g8}
+  \caption{Group 8 neutron flux distribution with Rod 1 inserted by 27.5 inches.}
+  \label{fig:halfway-g8}
+\end{figure}
+
+\begin{figure}[p]
+  \centering
+  \includegraphics[width=\columnwidth]{inserted-g1}
+  \caption{Group 1 neutron flux distribution with Rod 1 inserted by 51 inches.}
+  \label{fig:inserted-g1}
+\end{figure}
+
+\begin{figure}[p]
+  \centering
+  \includegraphics[width=\columnwidth]{inserted-g8}
+  \caption{Group 8 neutron flux distribution with Rod 1 inserted by 51 inches.}
+  \label{fig:inserted-g8}
+\end{figure}

### 2025-01-03 conclusion.tex
Edits before format review

@@
 reactivity. These discrepancies could be reduced through
-neutron leakage correction at the vacuum boundaries. Remedies for this issue exist in a number of
-forms such as group-wise or matrix albedo boundary conditions.
+neutron leakage correction at the vacuum boundaries. Remedies for the $k_\text{eff}$ mismatch
+exist in a number of forms such as group-wise or matrix albedo boundary conditions.
 
 The 3-D simulations in this work uncovered significant memory usage by the $S_N$ solver, even
-with the distributed mesh feature to spread mesh and variable storage across the compute nodes.
+with the distributed mesh feature to spread mesh and variable data storage across the compute nodes.
 Each full-core simulation required at least 40 nodes with 512 GB memory each to run.
@@
 
+Time-dependent simulations in this work uncovered two issues: rod cusping effects when a moving
+control rod does not align with the mesh interfaces, and slow solution convergence rates during
+control rod motion. While this work applied an empirical technique to correct for rod cusping
+effects, Moltres would benefit from a more robust technique for the long term. Slow solution
+convergence rates could be resolved by investigating better nested solver coupling structures to
+mitigate the effects of lagged solutions and applying solution relaxation schemes.
+
 Finally, future work could demonstrate the hybrid method to its fullest potential through the

### 2025-01-03 hybrid.tex
Edits before format review

@@
 
-This chapter starts with the $S_N$ method implementation required for the hybrid $S_N$-diffusion
-method. I implemented both $S_N$ and hybrid methods using \gls{FEM} numerical solver capabilities
-available through Moltres \cite{lindsay_moltres_2017} and the \gls{MOOSE} framework. In this
-section, the derivation for the \gls{SAAF} formulation of the $S_N$ method and its
+The second-order linear \gls{SAAF} formulation of the $S_N$ neutron transport equations is
+compatible with the highly efficient and scalable HYPRE-BoomerAMG preconditioner
+\cite{hypre_hypre_2022} available through
+the \gls{MOOSE} framework. Implementing the $S_N$ method in Moltres allows it to natively interface
+with the existing neutron diffusion solver for nonlinear diffusion acceleration and for the
+subsequent hybrid method implementation.
+The derivation of the \gls{SAAF} formulation of the $S_N$ method and its
 implementation in Moltres follows the derivation developed by Wang et al.\
@@
 \begin{multline}
-  \frac{\partial}{\partial t}\left[\frac{\Psi_g(\vec{r},\hat{\Omega},t)}{v_g}\right] +
-  \hat{\Omega}\cdot\nabla\Psi_g(\vec{r},\hat{\Omega},t) + \Sigma_{t,g}
-  \Psi_g(\vec{r},\hat{\Omega},t) =
-  \sum^G_{g'=1}\int_\mathcal{S} \Sigma_s^{g'\rightarrow g}(\hat{\Omega}'\rightarrow\hat{\Omega})
+  \frac{\partial}{\partial t}\left[\frac{\Psi_g(\vec{r},\hat{\Omega},t)}{v_g(\vec{r},t)}\right] +
+  \hat{\Omega}\cdot\nabla\Psi_g(\vec{r},\hat{\Omega},t) + \Sigma_{t,g}(\vec{r},t)
+  \Psi_g(\vec{r},\hat{\Omega},t) \\
+  = \sum^G_{g'=1}\int_\mathcal{S} \Sigma_s^{g'\rightarrow g}(\vec{r}, \hat{\Omega}'\rightarrow\hat{\Omega}, t)
   \Psi_{g'}(\vec{r},\hat{\Omega}',t)d\hat{\Omega}' \\
-  + \frac{1}{4\pi}\chi_{p,g}(1-\beta)\sum^G_{g'=1} \nu\Sigma_{f,g'} \phi_{g'}(\vec{r},t)
+  + \frac{1}{4\pi}\chi_{p,g}(1-\beta)\sum^G_{g'=1} \nu\Sigma_{f,g'}(\vec{r},t) \phi_{g'}(\vec{r},t)
   + \frac{1}{4\pi}\sum^I_{i=1}\chi_{d,g}
@@
   \begin{align*}
-    \chi_{p,g} &= \mbox{prompt fission neutron spectrum in group $g$,} \\
-    \beta &= \sum^I_{i=1} \beta_i = \mbox{total delayed neutron fraction,} \\
-    \chi_{d,g} &= \mbox{delayed fission neutron spectrum in group $g$,} \\
-    \lambda_i &= \mbox{decay constant of precursor group $i$,} \\
-    C_i &= \mbox{delayed neutron precursor concentration for group $i$,} \\
-    \Psi^\text{inc}_g &= \mbox{incident surface source in group $g$,} \\
+    \chi_{p,g}(\vec{r},t) &= \mbox{prompt fission neutron spectrum in group $g$,} \\
+    \beta(\vec{r},t) &= \sum^I_{i=1} \beta_i(\vec{r},t) = \mbox{total delayed neutron fraction,} \\
+    \chi_{d,g}(\vec{r},t) &= \mbox{delayed fission neutron spectrum in group $g$,} \\
+    \lambda_i(\vec{r},t) &= \mbox{decay constant of precursor group $i$,} \\
+    C_i(\vec{r},t) &= \mbox{delayed neutron precursor concentration for group $i$,} \\
+    \Psi^\text{inc}_g(\vec{r},t) &= \mbox{incident surface source in group $g$,} \\
     \alpha^s_g &= \mbox{specular reflectivity on }\partial \mathcal{D} \mbox{ for group }g, \\
@@
 %
+The definitions of the remaining variables may be found in Eq.\ \ref{eq:bte}.
+We drop the phase space dependence notation from this point on for brevity.
+
 In order to introduce operators and facilitate subsequent mathematical derivations, we will define
@@
   \langle\bm{a},\bm{b}\rangle^{\pm} \equiv \sum^G_{g=1}\sum_{s\in\partial\mathcal{D}}\int_s
-  d\vec{r} \int_{mathcal{S}^\pm_{\hat{n}_b}}d\hat{\Omega} |\hat{\Omega}\cdot\hat{n}_b|
+  d\vec{r} \int_{\mathcal{S}^\pm_{\hat{n}_b}}d\hat{\Omega} |\hat{\Omega}\cdot\hat{n}_b|
   \sum_{i\in E_s}a_{g,i}(\hat{\Omega})b_i(\vec{r})\sum_{j\in E_s}b_{g,j}(\hat{\Omega})b_j(\vec{r}),
@@
 \begin{gather}
-  \left(\bm{a},\bm{b}\right)_\mathcal{D} \equiv \sum^G_{g=1} \int_\mathcal{D}d\vec{r}
+  \left(\bm{a},\bm{b}\right)_\mathcal{D} \equiv \int_\mathcal{D}d\vec{r}
   \sum_{i\in E_e}a_{g,i}b_i(\vec{r})\sum_{j\in E_e}b_{g,j}b_j(\vec{r}), \\
-  \left(\bm{a},\bm{b}\right)_{\partial\mathcal{D}} \equiv \sum^G_{g=1}
+  \left(\bm{a},\bm{b}\right)_{\partial\mathcal{D}} \equiv
   \sum_{s\in\partial\mathcal{D}}\int_s d\vec{r}\sum_{i\in E_s}a_{g,i}b_i(\vec{r})\sum_{j\in E_s}
@@
 involving integrals over the spatial domain to explicitly show the $S_N$ angular discretization.
+
+Eqs.\ \ref{eq:time-derivative} to \ref{eq:reflecting-bc} present the terms from the \gls{SAAF}
+$S_N$ equations in operator notation (Eq.\ \ref{eq:saaf-vt}) and their expanded forms
+with explicit summations by neutron energy group and direction of neutron travel.
 
@@
 %
-\begin{gather}
-  \left(\left(\mathbb{I}+\bm{\tau}\mathbb{L}_1\right)\bm{\Psi}^*,\mathbb{S}\bm{\Psi}\right) =
+\begin{multline}
+  \left(\left(\mathbb{I}+\bm{\tau}\mathbb{L}_1\right)\bm{\Psi}^*,\mathbb{S}\bm{\Psi}\right) = \\
   \sum^G_{g=1}\sum^{N_d}_{d=1}w_d\left(\Psi^*_{g,d}+\tau_g\hat{\Omega}_d\cdot\nabla\Psi^*_{g,d},
@@
   \frac{2l+1}{w}Y_{l,m}(\hat{\Omega}_d)\phi_{g',l,m}\right)_\mathcal{D}
-\end{gather}
+\end{multline}
 %
@@
 resulting in diffusion coefficient values tending to positive or negative infinity when the
-flux gradient is close to zero. Larger mesh sizes may help avoid this issue at the cost of
-worsening flux accuracy in regions with
+flux gradient is close to zero (see Figure \ref{fig:3b-svdc}. Larger mesh sizes may help avoid this
+issue at the cost of worsening flux accuracy in regions with
 steep, non-linear fluxes, such as near control rods. For the Ronen method, Gross et al.\
@@
 Derivations for the drift terms depend on the discretization schemes of the high- and low-order
-equations. For this work, we will adopt drift term derivations developed by Wang et al.\
-\cite{wang_diffusion_2014, wang_rattlesnake_2018}. We start by integrating Eq.\ \ref{eq:saaf-vt}
+equations. For this work, we adopt drift term derivations developed by Wang et al.\
+\cite{wang_diffusion_2014, wang_rattlesnake_2018} specifically for the \gls{SAAF} $S_N$ equations
+with void treatment. Their work uses drift terms for \gls{NDA}, while this work employs drift terms
+for the hybrid method. We start by integrating Eq.\ \ref{eq:saaf-vt}
 over the 2-D unit sphere angular domain $\mathcal{S}$ and replacing $\bm{\Psi}^*$ with isotropic
@@
 %
-\begin{figure}[htb!]
+\begin{figure}[b]
 	\centering
@@
       from $S_8$ and $S_8$-diffusion simulations. The correction subregion $V_1$ spans $x=0$ cm to
-      $x=10$ cm.}
+      $x=10$ cm. The buffer region, $V_1'$, is highlighted in yellow.}
     \label{fig:3b-drift-1}
@@
 repository\footnote{\url{https://github.com/smpark7/moltres/tree/sn-method}} as of the time of
-writing. This branch may be merged into the original upstream Moltres GitHub repository in the near
+writing. This branch will be merged into the original upstream Moltres GitHub repository in the near
 future.
@@
 residual contributions falls below a defined tolerance value. All $S_N$ simulations in this work
-employ the \gls{PJFNK} method \cite{knoll_jacobian-free_2004} with HYPRE-BoomerAMG
+employ the \gls{PJFNK} solver \cite{knoll_jacobian-free_2004} with HYPRE-BoomerAMG
 \cite{hypre_hypre_2022} for preconditioning. BoomerAMG is a parallel implementation of the
@@
 establishes a buffer region in which transport correction parameters are discarded to improve
-$S_N$-diffusion coupling and preserve smoothness of the neutron flux gradients.
+$S_N$-diffusion coupling and preserve smoothness of the neutron flux gradients. The algorithm
+operates independently in each neutron energy group since the buffer region cutoffs
+do not necessarily coincide for all energy groups. The
+correction region is pre-determined at the start of the simulation, thus requiring preliminary
+analysis by the user to confirm that the region size is appropriate for the problem.
 

### 2025-01-03 introduction.tex
Edits before format review

@@
 \cite{u.s._doe_nuclear_energy_research_advisory_committee_technology_2002}.
-Due to the high thermal expansion coefficient, \glspl{MSR} possess an inherently robust
+Due to the large fuel expansion coefficient, \glspl{MSR} possess an inherently robust
 safety feature in the strong negative fuel temperature coefficient of
 reactivity \cite{elsheikh_safety_2013}. This reactivity coefficient limits the
-maximum temperature that the reactor core would experience in an accident
-scenario such as an unprotected reactivity insertion because the subsequent
+maximum temperature that the reactor core would experience in particular accident
+scenarios, such as an unprotected reactivity insertion, because the subsequent
 rise in core temperatures induces a significant drop in reactivity,
@@
 their experience with the \gls{MSRE}, \gls{ORNL} proposed a new program for constructing and
-operating a demonstration reactor based on the \gls{MSBR} concept, which they had also developed
+operating a demonstration reactor based on the \gls{MSBR} concept
 \cite{macpherson_molten_1985}. The \gls{MSBR} is a thermal-spectrum \gls{MSR} with fertile
@@
 However, the \gls{MSBR} project was canceled prior to the demonstration stage in
-favor of the \gls{LMFBR}, which had benefited from a head start in development and more substantial
+favor of the \gls{LMFBR}, which had benefited from a longer development time and more substantial
 political backing \cite{macpherson_molten_1985}.
@@
 the \gls{GIF}
-provided new impetus for \gls{MSR} research and development. As of the end of 2022, numerous
+provided new impetus for \gls{MSR} research and development. As of the end of 2024, numerous
 \gls{MSR} designs exist at various stages of development. Leading \gls{MSR} designs, in terms of
@@
 2022. Lastly, Canada-based Terrestrial Energy is also developing its \gls{IMSR}, a small modular
-\gls{MSR} based on the \gls{MSRE}. It passed a joint technical review by the
+\gls{MSR} based on the \gls{MSRE}. It passed a joint technical review by
 Canadian and US nuclear regulators in July 2022.
@@
 Developing \gls{MSR} simulation software plays a vital role in supporting \gls{MSR}
-development. Accurate reactor modeling capabilities are vital because they
+development. Accurate reactor modeling capabilities
 accelerate reactor design and optimization by enabling quicker iteration through numerous design
@@
 the challenges of simulating \glspl{MSR} compared to solid-fueled reactors stem
-mainly from the liquid fuel form of the fuel salt \cite{diamond_phenomena_2018,
+primarily from the liquid fuel form of the fuel salt \cite{diamond_phenomena_2018,
 huff_identifying_2019}.
@@
 with excess reactivity to ensure enough fissile material is available to maintain power operation
-and last through to the next scheduled refueling. Control rods and burnable poisons cancel out the
-excess reactivity. Control rods are also helpful
+and last through to the next scheduled refueling. Control rods and burnable poisons balance the
+initial excess reactivity. Control rods are also helpful
 for ramping the power up or down during start-up, shut-down, or load-following operations. Most
@@
 control rod worth, which is measured as the relative reduction
-in the neutron multiplication factor $k_{eff}$ due to the insertion of the control rod in the
+in the neutron multiplication factor $k_\text{eff}$ due to the insertion of the control rod in the
 reactor. They are also interested in the accompanying change in flux shape near the
@@
 
-\section{Research Objectives}
+\section{Research Objectives and Outline}
 
@@
 simulation tool for multiphysics \gls{MSR} analysis that can be run on workstations as well as on
-larger computing clusters. Thus, new and existing capabilities in Moltres must
+large leadership-class computing clusters. Thus, new and existing capabilities in Moltres must
 be accurate within accepted bounds, rigorously verified, computationally efficient, and highly
@@
 
-\section{Outline}
-
 Chapter 2 presents a literature review of existing multiphysics \gls{MSR} simulation software,

### 2025-01-03 lit.tex
Edits before format review

@@
 exchanger, and secondary heat exchanger as 0D systems (lumped-parameter model)
-and substituting the 2-D fuel channel with a 3-D fuel channel which more closely
+and substituting the 2-D fuel channel with a 3-D fuel channel that more closely
 resembled the actual fuel channels in the \gls{MSRE}
@@
 in use today \cite{rodi_turbulence_2017}, operate on the eddy viscosity
-hypothesis which states that the Reynolds stresses in the \gls{RANS} equations
+hypothesis that states that the Reynolds stresses in the \gls{RANS} equations
 are given by:
@@
 \cite{wilcox_turbulence_2006}. More minor improvements are observed in modeling
-free shear and backward-facing step flows \cite{wilcox_turbulence_2006}.
+free shear and backward-facing step flows.
 
@@
 
-There is a lack of research interest in methods for improving control rod modeling in
-\glspl{MSR}.
+Research interest in methods for improving control rod modeling in \glspl{MSR} is subdued compared
+to the focus on modeling other physical interactions in \glspl{MSR}.
 The self-regulating nature and low excess reactivity of \glspl{MSR} reduce or even eliminate the
@@
 designs with no control rods include the fast-spectrum \gls{MSFR} and TerraPower's \gls{MCFR}
-\cite{terrapower_terrapower_2021}. However, \gls{MSR} power control schemes which rely heavily on
+\cite{terrapower_terrapower_2021}. However, \gls{MSR} power control schemes that rely heavily on
 these mechanisms may not be suitable for all accident scenarios. Control rod insertion provides
@@
 Few \gls{MSR} multiphysics studies explicitly include control rods in their models. For instance,
-for the \gls{MSRE} pump start-up and coast-down experiments at \gls{ORNL} which involved control
+for the \gls{MSRE} pump start-up and coast-down experiments at \gls{ORNL}, which involved control
 rod movement, most numerical studies simulate the reactivity effects of the control rods by scaling
@@
   \Sigma_f(\vec{r},E',t) =& \mbox{ macroscopic fission cross section,} \nonumber \\
-  S(\vec{r},E,\hat{\Omega}',t) =& \mbox{ external neutron source.} \nonumber
+  S(\vec{r},E,\hat{\Omega},t) =& \mbox{ external neutron source.} \nonumber
 \end{align}
@@
 
+\begin{figure}[b!]
+    \centering
+    \includegraphics[width=.4\columnwidth]{mecs-geometry}
+    \caption{Geometries of the absorber region and its vicinity in the supercell
+      (top) and the diffusion solver mesh (bottom).
+        The shaded and unshaded regions represent the homogenized absorber
+        region and the adjacent non-absorber region, respectively.
+        Retrieved from Fen et al.\ \cite{fen_modelling_1992}.}
+    \label{fig:mecs-geometry}
+\end{figure}
+
 Scherer \& Neef developed the \gls{MECS} \cite{scherer_determination_1976} to improve control rod
@@
 analytic calculation to obtain equivalent diffusion coefficients or cross sections for the absorber
-region. The diffusion coefficient formula for this analytic calculation depends on the solution
+region.
+
+The diffusion coefficient formula for this analytic calculation depends on the solution
 method of the diffusion solver. In the most widely used implementation of \gls{MECS} used in
@@
 \ref{fig:mecs-geometry}) in the CITATION code as provided by Fen et al.\ \cite{fen_modelling_1992}.
-%
-\begin{figure}[htb]
-    \centering
-    \includegraphics[width=.4\columnwidth]{mecs-geometry}
-    \caption{Geometries of the absorber region and its vicinity in the supercell
-      (top) and the diffusion solver mesh (bottom).
-        The shaded and unshaded regions represent the homogenized absorber
-        region and the adjacent non-absorber region, respectively.
-        Retrieved from Fen et al.\ \cite{fen_modelling_1992}.}
-    \label{fig:mecs-geometry}
-\end{figure}
 
@@
 nodalization of the rest of the reactor geometry \cite{ougouag_transport_2010}. \gls{MECS} is
-incompatible with reactor geometries which contain control rods that are too close to each other
+incompatible with reactor geometries that contain control rods that are too close to each other
 such that there is not enough distance in between to define neighboring nodes where
@@
     \delta J(x,E) =& J_{tr}(x,E) - J_D(x,E), \nonumber \\
+    J_D =& \mbox{ diffusion-derived neutron current,} \nonumber \\
     \Delta x_i =& \mbox{ size of $i$-th spatial interval.} \nonumber
@@
 heterogeneities in reactor geometries to reduce the computational costs of running full-core
-simulations. Equivalence techniques which reduce spatial homogenization error are also effective
+simulations. Equivalence techniques that reduce spatial homogenization error are also effective
 for accurately modeling the integral worths of control rods within fuel assemblies. The \gls{GET}
@@
 factors can be precombined with the cross sections to generate \gls{SPH}-corrected cross sections.
-\gls{GET} requires diffusion solvers, which allow for flux discontinuities at the interfaces.
+\gls{GET} requires diffusion solvers that allow for flux discontinuities at the interfaces.
 Furthermore, \gls{GET} requires more memory to store up to six discontinuity factors, one for each
@@
 
-Lastly, there is a lack of research focus into robust techniques for accurate control rod modeling
+Lastly, there is a lack of existing work into robust techniques for accurate control rod modeling
 for time-dependent \gls{MSR} analyses. Control rods induce highly anisotropic fluxes within

### 2025-01-03 main.tex
Edits before format review

@@
 \committee{Associate Professor Kathryn D. Huff, Co-Chair \\
-           Assistant Professor Madicken Munk, Co-Chair \\
+           Assistant Professor Madicken Munk, Co-Chair, Oregon State University \\
            Professor Rizwan Uddin \\
@@
 
-%\input{acks}
+\input{acks}
 

### 2025-01-03 msre.tex
Edits before format review

@@
 correction schemes for the hybrid method: the diffusion and drift correction
-schemes. This chapter demonstrates
-the hybrid method through 1-D, 2-D, and 3-D $k$-eigenvalue simulations of the \gls{MSRE}.
-This chapter employs the OpenMC Monte Carlo neutral particle transport software \cite{romano_openmc:_2015} to
-generate group constants for the hybrid method and to generate reference solutions for method verification.
+schemes. This chapter demonstrates the hybrid method through 1-D, 2-D, and 3-D $k$-eigenvalue
+simulations of the \gls{MSRE}. The work in this chapter employs the OpenMC Monte Carlo neutral
+particle transport software \cite{romano_openmc:_2015} to generate group constants for the hybrid
+method and to generate reference solutions for method verification.
 The 1-D models serve to verify the newly-implemented $S_N$ and hybrid methods
@@
 \ref{sec:nts-methods} describes and labels the different neutronics methods employed for the
-$k$-eigenvalue simulations. Section \ref{sec:test-metrics} details basic test metrics relevant to
+$k$-eigenvalue simulations. Section \ref{sec:test-metrics} details test metrics relevant to
 neutronics verification and validation. Sections \ref{sec:1d-results}, \ref{sec:2d-results}, and
@@
 
+Deterministic neutron diffusion and $S_N$ neutron transport methods require group constant data in
+the form of group-averaged macroscopic neutron cross sections, neutron speeds, and delayed neutron
+precursor data.
 Group constants relevant to the neutron diffusion, $S_N$ neutron transport, or both
-methods are:
+methods in Moltres are:
 %
 \begin{align*}
-  \Sigma_{t,g}=& \mbox{ Macroscopic total cross section in group $g$,}\\
-  \Sigma_{r,g}=& \mbox{ Macroscopic removal cross section in group $g$,}\\
-  \Sigma_s^{g'\rightarrow g}=& \mbox{ Macroscopic group-to-group scattering} \\
-  \mbox{cross section matrix,}\\
+  \Sigma_{t,g}=& \mbox{ Macroscopic total cross section in group $g$,} &\\
+  \Sigma_{r,g}=& \mbox{ Macroscopic removal cross section in group $g$,} &\\
+  \Sigma_s^{g'\rightarrow g}=& \mbox{ Macroscopic group-to-group scattering cross section matrix,} &\\
   \Sigma_{s,l}^{g'\rightarrow g}=& \mbox{ $l$-th Legendre moment of the macroscopic
-  group-to-group scattering cross section matrix,}\\
+        group-to-group scattering} & \\
+                           & \mbox{cross section matrix,} &\\
   \Sigma_{sp,l}^{g'\rightarrow g}=& \mbox{ $l$-th Legendre moment of the macroscopic
-  group-to-group scattering production} \\
-  \mbox{cross section matrix,}\\
-  D_g=& \mbox{ $P_1$-based diffusion coefficient in group $g$,}\\
-  \nu\Sigma_{f,g}=& \mbox{ Product of the average number of neutrons produced per fission and} \\
-  \mbox{the macroscopic fission cross section in group $g$,}\\
-  \chi_g=& \mbox{ Neutron fission spectrum in group $g$.}\\
+  group-to-group scattering production} & \\
+                                        & \mbox{cross section matrix,} &\\
+  D_g=& \mbox{ $P_1$-based diffusion coefficient in group $g$,} &\\
+  \nu\Sigma_{f,g}=& \mbox{ Product of the average number of neutrons produced per fission and} &\\
+                  &\mbox{the macroscopic fission cross section in group $g$,}&\\
+  \chi_g=& \mbox{ Neutron fission spectrum in group $g$.}&\\
 \end{align*}
-%
+
 OpenMC generates these group constants using its multigroup cross section generation capability
@@
 \cite{chadwick_endf/b-vii.1_2011}. All models have a uniform
-temperature of 900 K unless otherwise stated. Detailed information pertaining specifically to the
-1-D, 2-D, and 3-D OpenMC models may be found in the respective subsections below.
+temperature of 900 K unless otherwise stated. Sections \ref{sec:1d-model-setup},
+\ref{sec:2d-model-setup}, \ref{sec:3d-model-setup} provide detailed information pertaining
+specifically to the 1-D, 2-D, and 3-D model geometries, respectively.
 
@@
 
-I ran all cases on OpenMC to generate the required group constants and reference $k_\text{eff}$
-values to assess the accuracy of the deterministic methods ($S_N$, neutron diffusion, hybrid 
-methods) against OpenMC in continuous-energy (OpenMC-CE) and multigroup (OpenMC-MG) modes. Table
+Besides generating group constants for the deterministic methods, OpenMC also provides reference
+netruon flux and $k_\text{eff}$ values for all studied cases. OpenMC can run in continuous-energy
+(OpenMC-CE) and multigroup (OpenMC-MG) modes. Table
 \ref{table:var} details how OpenMC and the deterministic methods represent the position, direction
@@
   \caption{Implementation differences of neutronics methods on Moltres and Python.}
-  \begin{tabular}{p{0.14\linewidth} p{0.25\linewidth} p{0.15\linewidth} p{0.15\linewidth} p{0.17\linewidth}}
-    \toprule
-    Implementations & Spatial discretization & $S_N$ formulation & $S_N$ acceleration &
-    Hybrid transport\newline correction \\
+  \begin{tabular}{l l l}
+    \toprule
+    Implementation detail & Moltres & Python \\
     \midrule
-    Moltres & Finite element method & SAAF & Nonlinear diffusion acceleration & Drift correction \\
-    Python & Finite difference (Diffusion)\newline Diamond difference ($S_N$) & First-order & None & Diffusion correction \\
+    Spatial discretization & Finite element method & Finite difference \& Diamond difference \\
+    $S_N$ formulation & SAAF & First-order \\
+    $S_N$ acceleration & Nonlinear diffusion acceleration & None \\
+    Hybrid transport correction & Drift correction & Diffusion correction \\
     \bottomrule
@@
 
-\section{Basic Test Metrics} \label{sec:test-metrics}
+\section{Test Metrics} \label{sec:test-metrics}
 
@@
 neutronics methods.
-
 For steady-state eigenvalue calculations, all solvers directly provide the effective neutron
@@
 
-This section presents the models, results, and discussion for the 1-D test cases.
-
-\subsection{1-D Neutronics Model Setup \& Geometries}
+This section presents the models, results, and discussion for the 1-D test cases. Section
+\ref{sec:1d-model-setup} describes the model setup and geometries of the 1-D test cases.
+Section \ref{sec:1d-mesh-conv} discusses the 1-D mesh convergence tests. Lastly, Section
+\ref{sec:1d-results-sub} shows the results and discussion of the 1-D test cases.
+
+\subsection{1-D Neutronics Model Setup \& Geometries} \label{sec:1d-model-setup}
 
@@
 
-\subsection{1-D Neutronics Model Mesh Convergence Tests}
+\subsection{1-D Neutronics Model Mesh Convergence Tests} \label{sec:1d-mesh-conv}
 
@@
   \centering
-  \includegraphics[width=.95\columnwidth]{diffusion-mesh-convergence-flux}
+  \includegraphics[width=.96\columnwidth]{diffusion-mesh-convergence-flux}
   \caption{Absolute difference in group fluxes ($\phi_g$) from the neutron diffusion method for
@@
   \centering
-  \includegraphics[width=.95\columnwidth]{sn-mesh-convergence-flux}
+  \includegraphics[width=.96\columnwidth]{sn-mesh-convergence-flux}
   \caption{Absolute difference in group fluxes ($\phi_g$) from the $S_8$ neutron transport method
@@
 
-\subsection{1-D Neutronics Numerical Results \& Discussion}
+\subsection{1-D Neutronics Numerical Results \& Discussion} \label{sec:1d-results-sub}
 
@@
   \caption{Difference in reactivity $\rho$ of all neutronics methods investigated relative
-  to OpenMC-CE.}
+  to OpenMC-CE. The standard deviation refers to the uncertainty associated with the Monte Carlo
+  method in OpenMC-CE.}
   \label{fig:1d-rho}
@@
 discrepancies.
-
-\begin{figure}[htb!]
-  \centering
-  \includegraphics[width=\columnwidth]{worth}
-  \caption{Percentage difference in rod worth for Cases 2 and 3 of all neutronics methods
-  investigated relative to OpenMC-CE.}
-  \label{fig:1d-worth}
-\end{figure}
 
@@
 
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=\columnwidth]{worth}
+  \caption{Percentage difference in rod worth for Cases 2 and 3 of all neutronics methods
+  investigated relative to OpenMC-CE. All percentage difference values are consistently positive.}
+  \label{fig:1d-worth}
+\end{figure}
+
 \subsubsection{Neutron flux distribution}
@@
 
-\begin{figure}[htb!]
-  \centering
-  \includegraphics[width=\columnwidth]{case-3b-drift}
+\begin{figure}[p]
+  \centering
+  \includegraphics[width=.99\columnwidth]{case-3b-drift}
   \caption{Multigroup drift correction ($\vec{D}_g$) $x$-component distributions from the
@@
 
-\begin{figure}[htb!]
-  \centering
-  \includegraphics[width=\columnwidth]{case-3b-svdc}
+\begin{figure}[p]
+  \centering
+  \includegraphics[width=.99\columnwidth]{case-3b-svdc}
   \caption{Multigroup diffusion correction ($D^s_g$) $x$-component distributions from the
@@
 control rod worth estimates with Cases 3a and 3b by varying the subregion sizes from 10 cm to 40 cm
-at 5 cm-intervals.
+at 5 cm-intervals. All subregion sizes place the subregion boundary more than 1 neutron mean free
+path away from the control rod. The shortest measured neutron mean free paths are 4.81 cm and 5.85
+cm for group 1 neutrons in the fuel and graphite regions, respectively.
 
@@
 
-\begin{table}[htb]
+\begin{table}[h]
   \centering
@@
 \end{table}
+%
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=0.6\columnwidth]{sn-tol}
+  \caption{$k_\text{eff}$ error estimates of Case 3b for a range of convergence tolerance values
+  imposed on the $S_8$ subsolver relative to the reference $k_\text{eff}$ value when
+  $\epsilon_\text{tol}=10^{-8}$. The $q=1.333$ line represents the approximate rate of
+  convergence.}
+  \label{fig:sn-tol}
+\end{figure}
 
@@
 
-\begin{figure}[htb!]
-  \centering
-  \includegraphics[width=0.6\columnwidth]{sn-tol}
-  \caption{$k_\text{eff}$ error estimates of Case 3b for a range of convergence tolerance values
-  imposed on the $S_8$ subsolver relative to the reference $k_\text{eff}$ value when
-  $\epsilon_\text{tol}=10^{-8}$. The $q=1.333$ line represents the approximate rate of
-  convergence.}
-  \label{fig:sn-tol}
-\end{figure}
+\FloatBarrier
 
@@
 
-This section presents the models, results, and discussion of the 2-D cases.
+This section presents the models, results, and discussion of the 2-D cases. Section
+\ref{sec:2d-model-setup} describes the model setup of the 2-D cases. Section
+\ref{sec:2d-nts-results} covers the results and discussion of the 2-D cases.
 
@@
 
-\begin{figure}[htb!]
-  \centering
-  \includegraphics[width=\columnwidth]{quarter-core-geom}
-  \caption{2-D \gls{MSRE} quarter-core model based on the horizontal cross section of the actual
-  \gls{MSRE} geometry.}
-  \label{fig:1/4-geom}
-\end{figure}
-
-\begin{figure}[htb!]
-  \centering
-  \includegraphics[width=\columnwidth]{full-core-geom}
-  \caption{2-D \gls{MSRE} full-core model based on the horizontal cross section of the actual
-  \gls{MSRE} geometry.}
-  \label{fig:full-geom}
-\end{figure}
-
 \begin{table}[htb]
@@
 
-\begin{figure}[htb!]
-  \centering
-  \includegraphics[width=0.7\columnwidth]{full-core-closeup}
-  \caption{Detailed view of the control rod thimbles and sample basket in 2-D \gls{MSRE} full-core
-    model.}
-  \label{fig:full-geom-closeup}
-\end{figure}
-
 I ran neutronics simulations on the quarter- and full-core models using OpenMC under both continuous
@@
 drift correction parameters for the coupled diffusion solver.
+
+\begin{figure}[h!]
+  \centering
+  \includegraphics[width=\columnwidth]{quarter-core-geom}
+  \caption{2-D \gls{MSRE} quarter-core model based on the horizontal cross section of the actual
+  \gls{MSRE} geometry.}
+  \label{fig:1/4-geom}
+\end{figure}
+
+\begin{figure}[p]
+  \centering
+  \includegraphics[width=\columnwidth]{full-core-geom}
+  \caption{2-D \gls{MSRE} full-core model based on the horizontal cross section of the actual
+  \gls{MSRE} geometry.}
+  \label{fig:full-geom}
+\end{figure}
+
+\begin{figure}[p]
+  \centering
+  \includegraphics[width=0.7\columnwidth]{full-core-closeup}
+  \caption{Detailed view of the control rod thimbles and sample basket in 2-D \gls{MSRE} full-core
+    model.}
+  \label{fig:full-geom-closeup}
+\end{figure}
 
@@
 of error in the hybrid method (e.g., neutron leakage, heterogeneous core geometry).
-.
 The OpenMC-CE and OpenMC-MG rod worth estimates agree within two standard deviations of each other.
@@
 This section presents the models, results, and discussion of the 3-D \gls{MSRE} simulations.
+Section \ref{sec:3d-model-setup} describes the model setup for the 3-D \gls{MSRE} simulations.
+Section \ref{sec:3d-nts} covers the results and discussion of the 3-D \gls{MSRE} simulations.
 
@@
   \centering
-  \caption{List of key reactor dimensions.}
+  \caption{List of key dimensions of the 3-D \gls{MSRE} model used in this work.}
   \begin{tabular}{l S}
@@
 
-Figures * show a 3-D section view of the \gls{MSRE} model geometry and the group 1 \& 8 neutron
+Figures \ref{fig:3d-view}, \ref{fig:ic-g1}, and \ref{fig:ic-g8} show a 3-D section view of the
+\gls{MSRE} model geometry and the group 1 \& 8 neutron
 flux distributions. The salt-graphite lattice structure is visible from the group flux
@@
   \includegraphics[width=\columnwidth]{ic-g1}
-  \caption{}
+  \caption{Group 1 neutron flux distribution with Rod 1 inserted by 4.4 inches at initial
+  criticality.}
   \label{fig:ic-g1}
@@
   \includegraphics[width=\columnwidth]{ic-g8}
-  \caption{}
+  \caption{Group 8 neutron flux distribution with Rod 1 inserted by 4.4 inches at initial
+  criticality.}
   \label{fig:ic-g8}
@@
 domain where flux peaks and troughs occur. These discontinuities and negative diffusion coefficient
-values are unphysical and would lead to numerical stability issues. Therefore, the hybrid method
+values are unphysical and lead to numerical stability issues. Therefore, the hybrid method
 implementation with drift correction scheme was chosen for the remaining 1-D analyses and the
@@
 control rod worth estimates as long as the subregion size is kept consistent (i.e., same size) when
-performing control rod worth measurements. Rod worths varied by less than 0.35 \% across the
+performing control rod worth measurements, and the subregion boundary is more than 1 neutron mean
+free path away from the rod. Rod worths varied by less than 0.35 \% across the
 different subregion sizes investigated. A separate analysis showed that the $k_\text{eff}$
@@
 full-core model. The hybrid method reported rod worth error magnitudes of less than 40 pcm relative
-to OpenMC-CE, which are significant improvements over the neutron diffusion method error estimates
-ranging from 569 pcm to 1484 pcm. The hybrid method also showed significant improvements in the
-absolute mean and maximum errors in fuel channel power distributions relative to the reference
-OpenMC-CE distribution.
+to OpenMC-CE, which are significant improvements over the neutron diffusion method error estimates,
+which range from 569 pcm to 1484 pcm from OpenMC-CE. The hybrid method also showed significant
+improvements in the absolute mean and maximum errors in fuel channel power distributions relative
+to the reference OpenMC-CE distribution.
 
@@
 The simulation setups mitigated excessively large memory requirements through the distributed mesh
-capability in Moltres. Some performance optimizations may be possible to reduce data transfer times
-between the $S_N$ and diffusion solvers which currently take up about 18 \% of the total solution
-time as a result of the inter-processor communications.
+capability in Moltres. Performance optimizations on data transfer and caching routines may reduce
+data transfer times between the $S_N$ and diffusion solvers which currently take up about 18 \% of
+the total solution time as a result of the inter-processor communications.
 
 Overall, the hybrid $S_N$-diffusion method provides a significant improvement over the neutron
-diffusion method at only four times the computational expense. Control rod estimates
+diffusion method at approximately four times the computational expense. Control rod estimates
 show good agreement with reference Monte Carlo solutions and \gls{MSRE} experimental data.

### 2025-01-03 transient.tex
Edits before format review

@@
 
-This chapter builds on the foundations of Chapter \ref{chap:msre} to verify and validate the hybrid
+This chapter builds on the foundations of Chapter \ref{chap:msre} to validate the hybrid
 method for neutronics modeling in time-dependent 3-D simulations of reactivity-initiated
-transients. This chapter presents two sets of simulation results involving the hybrid method and
-other capabilities in Moltres of the \gls{MSRE} rod drop and reactivity insertion
-experiments. Beyond time-dependent modeling with the hybrid method, the simulations will also
+transients. This chapter presents simulation results from Moltres for the \gls{MSRE} rod drop
+experiment. Beyond time-dependent modeling with the hybrid method, the simulations will also
 demonstrate compatibility between the existing neutronics/thermal-hydraulics coupling
@@
 \gls{MSRE} researchers conducted a series of rod drop experiments at three different $^{235}$U
-loadings \cite{prince_zero-power_1968}. These experimented were conducted from an initially critical
+loadings \cite{prince_zero-power_1968}. These experiments were conducted from an initially critical
 and low power state with an initial neutron count rate of approximately 30,000 counts per second
@@
 authors back-calculated the critical $^{235}$U loading with Rod 1 fully withdrawn to be 65.25 kg.
-They also provided a $^{235}$U concentration reactivity coefficient of 0.223 $(\delta k/k)/\delta m
-/m)$. The excess $^{235}$U mass at initial criticality compensating for Rod 1 inserted to 46.6
+They also provided a $^{235}$U concentration reactivity coefficient of 0.223 $(\delta k/k)/(\delta m
+/m)$, where $k$ is the multiplication factor and $m$ is the $^{235}$U mass. The excess $^{235}$U
+mass at initial criticality compensating for Rod 1 inserted to 46.6
 inches can estimated through the following steps:
@@
 
-\begin{figure}[t]
+\begin{figure}[h]
   \centering
@@
 \gls{MSRE} model from Chapter
-\ref{chap:msre} at the critical $^{235}$ loading. Refer to Section \ref{sec:3d-model-setup} for
+\ref{chap:msre} at the critical $^{235}$U loading. Refer to Section \ref{sec:3d-model-setup} for
 details on the 3-D model. The initial rod height for the \gls{MSRE} model to induce a similar total
@@
 Figure \ref{fig:rod-drop-coupling} illustrates this iterative coupling structure and the data
-transfers between the solvers.
+transfers between the solvers. The \gls{DNP} solver is a preexisting capability in Moltres
+described in Section \ref{sec:moltres-physics}. For the rod drop simulation, the salt velocity and
+\gls{DNP} diffusion coefficients were set to zero, i.e., static salt condition. The simulation
+grouped the \glspl{DNP} into six \gls{DNP} decay groups automatically determined by the
+ENDF/B-VII.1 nuclear data library \cite{chadwick_endf/b-vii.1_2011} and OpenMC during group
+constant generation.
 
@@
 a failed flux-volume weighting implementation and time constraints, this work uses an empirical
-approach for rod cusping effect correction. The approach involves precomputing the $k_\text{eff}$
-with the control rod at various levels of insertion within a mesh element. These simulations use
-the simple volume weighting approach to compute a homogenized group constant value as follows:
-%
-\begin{gather}
-  \Sigma_\text{vw} = \frac{V_\text{r} \Sigma_\text{r} + V_\text{nr} \Sigma_\text{nr}}{V_\text{r} + V_\text{nr}} \label{eq:vw}
-  \shortintertext{where}
-  \begin{align*}
-    \Sigma_\text{vw} &= \mbox{ homogenized group constant in mixed mesh element,} \\
-    V_\text{r} &= \mbox{ rod volume fraction in mixed mesh element,} \\
-    \Sigma_\text{r} &= \mbox{ rod group constant value,} \\
-    V_\text{nr} &= \mbox{ non-rod volume fraction in mixed mesh element,} \\
-    \Sigma_\text{nr} &= \mbox{ non-rod group constant value.}
-  \end{align*}
-\end{gather}
-%
-Note that vertical material and mesh interfaces in the 3-D \gls{MSRE} model always align perfectly
-because the control rod moves vertically. Therefore, the mixed mesh elements occur only around the
-bottom tip of the control rod. Additionally, the volume fraction is equivalent to the height
-fraction, and all mixed mesh elements at the same height share the same volume fraction. Figure
-\ref{fig:rod-cusping-1} shows how the fraction of reactivity change changes with the control rod
-volume fraction when the control rod is at around mid-reactor and full insertion heights. The
-figure depicts the nonlinear relation between the reactivity change and volume fraction that the
-``cusping effect'' is named after. Within this small range of rod motion, the relation should be
-approximately linear.
+approach for rod cusping effect correction. The empirical approach involves precomputing the
+$k_\text{eff}$ with the control rod at various levels of insertion within a mesh element. These
+simulations apply a simple volume weighting approach to compute a homogenized group constant value
+as follows:
 
@@
 \end{figure}
+
+\begin{gather}
+  \Sigma_\text{vw} = \frac{V_\text{r} \Sigma_\text{r} + V_\text{nr} \Sigma_\text{nr}}{V_\text{r} + V_\text{nr}} \label{eq:vw}
+  \shortintertext{where}
+  \begin{align*}
+    \Sigma_\text{vw} &= \mbox{ volume-weighted group constant in mixed mesh element,} \\
+    V_\text{r} &= \mbox{ rod volume fraction in mixed mesh element,} \\
+    \Sigma_\text{r} &= \mbox{ rod group constant value,} \\
+    V_\text{nr} &= \mbox{ non-rod volume fraction in mixed mesh element,} \\
+    \Sigma_\text{nr} &= \mbox{ non-rod group constant value.}
+  \end{align*}
+\end{gather}
+
+Note that vertical material and mesh interfaces in the 3-D \gls{MSRE} model always align perfectly
+because the control rod moves vertically. Therefore, the mixed mesh elements occur only around the
+bottom tip of the control rod. Additionally, the volume fraction is equivalent to the height
+fraction, and all mixed mesh elements at the same height share the same volume fraction. Figure
+\ref{fig:rod-cusping-1} shows how the fraction of reactivity change changes with the control rod
+volume fraction when the control rod is at around mid-reactor and full insertion heights. The data
+points depict the nonlinear relation between the reactivity change and volume fraction that the
+``cusping effect'' is named after. Within this small range of rod motion, the relation should be
+approximately linear.
 
@@
 simulation managed to continue after raising the convergence tolerance value of the neutron
-diffusion subsolver and the fixed point iterations from $\epsilon_\text{tol}=10^{-8}$ to $10^{-7}$. However,
-this action led to under-converged scalar flux and neutron count results between $t=1.325$ s and
-$3.825$ s.
+diffusion subsolver and the fixed point iterations from $\epsilon_\text{tol}=10^{-8}$ to $10^{-6}$.
+However, this action led to under-converged scalar flux and neutron count results between $t=1.325$
+s and $3.825$ s.
 
@@
 count curve shape is more consistent with the experimental data, albeit seemingly in response to a
-larger reactivity withdrawal.
+larger reactivity withdrawal. Sources of experimental uncertainties listed in the previous
+paragraph may have contributed to the discrepancy between the experimental and numerical data.
 
@@
   \centering
-  \includegraphics[width=0.8\columnwidth]{count-rate}
+  \includegraphics[width=0.78\columnwidth]{count-rate}
   \caption{Neutron count rate during the rod drop experiment from Moltres. The convergence tolerance
@@
   \label{fig:count-rate}
-  \includegraphics[width=0.8\columnwidth]{integral-count}
+  \includegraphics[width=0.78\columnwidth]{integral-count}
   \caption{Integral neutron count during the rod drop experiment from \gls{MSRE} experimental data
@@
   \centering
-  \includegraphics[width=0.8\columnwidth]{count-rate-interp}
+  \includegraphics[width=0.78\columnwidth]{count-rate-interp}
   \caption{Neutron count rate during the rod drop experiment from Moltres with linearly interpolated
@@
   \label{fig:count-rate-interp}
-  \includegraphics[width=0.8\columnwidth]{integral-count-interp}
+  \includegraphics[width=0.78\columnwidth]{integral-count-interp}
   \caption{Integral neutron count during the rod drop experiment from \gls{MSRE} experimental data
@@
 solution convergence, particularly when the control rod was in motion. This limited the simulation
-from taking larger time-steps. Further work could include more sophiscated coupling schemes to
-help the solver converge faster.
+from taking larger time-steps. Further work could include rearranging the nested coupling structure
+or implementing relaxation schemes to improve solver numerical stability and convergence.
 
@@
 The hybrid $S_N$-diffusion method combines the strengths of the $S_N$ and neutron
-diffusion methods accurately model control rods while remaining tractable on modest computational
+diffusion methods to accurately model control rods while remaining tractable on modest computational
 resources. This chapter covers work demonstrating the hybrid $S_N$-diffusion method through a
-time-dependent reactivity-initiated simulations. The hybrid $S_N$-diffusion solver implemented in
+time-dependent reactivity-initiated simulation. The hybrid $S_N$-diffusion solver implemented in
 Moltres is coupled to preexisting \gls{DNP} and temperature solver capabilities to model
@@
 subsolver impacted the precision of several data points during the simulation.
-Future investigations could look into better coupling schemes to resolves these issues.
+Future investigations could look into alternative nested solver coupling structures and relaxation
+schemes to resolve these issues.
 Experimental uncertainties in the actual $^{235}$U concentration, initial \& final rod heights, rod

### 2025-01-03 turbulence.tex
Edits before format review

@@
 
-\begin{gather}
+\pagebreak
+\begin{multline}
   \rho \frac{\partial\tilde{\mu}}{\partial t} + \rho \mathbf{u}\cdot\nabla\tilde{\mu} = \rho c_{b1}
-  \left(1-f_{t2}\right)\tilde{S}\tilde{\mu} + \frac{1}{\sigma}\{\nabla\cdot\left[\left(\mu+
+  \left(1-f_{t2}\right)\left[\tilde{S}+C_\text{rot}\min(0,S-\Omega)\right]\tilde{\mu} \\ + \frac{1}{\sigma}\{\nabla\cdot\left[\left(\mu+
   \tilde{\mu}\right)\nabla\tilde{\mu}\right] + c_{b2}|\nabla\tilde{\mu}|^2\} - \left(c_{w1}f_w -
@@
   \frac{\tilde{\mu}}{d}\right)^2
+\end{multline}
+\begin{gather}
   \shortintertext{where}
@@
   c_{w2} = 0.3, \ c_{w3} = 2, \
-  c_{v1} = 7.1, \ c_{t3} = 1.2, \ c_{t4} = 0.5 \ \text{.} \nonumber
+  c_{v1} = 7.1, \ c_{t3} = 1.2, \ c_{t4} = 0.5, \ C_\text{rot} = 2 \ \text{.} \nonumber
 \end{gather}
@@
 
-\begin{figure}[p]
-  \centering
-  \includegraphics[width=0.9\columnwidth]{channel-geom}
-  \caption{Channel geometry for the turbulent channel flow verification test. The red box indicates
-  the region shown by the close-up view in Figure \ref{fig:channel-mesh}.}
-  \label{fig:channel-geom}
-%\end{figure}
-%%
-%\begin{figure}[htb!]
-  \centering
-  \includegraphics[width=\columnwidth]{channel_mesh}
-  \caption{Close-up view of the refined mesh near the inlet (left boundary) for the channel flow
-    test. The mesh for the rest of the channel follows the mesh resolution of the rightmost column
-  of elements.}
-  \label{fig:channel-mesh}
-\end{figure}
-
-\begin{table}[htb]
+\begin{table}[h]
   \centering
@@
 wall distances, and stresses from the \gls{DNS} data \cite{moser_direct_1999}
-and the Moltres Spalart-Allmaras model. The Moltres Spalart-Allmaras model is largely consistent with the reference
+and the Moltres Spalart-Allmaras model. Moltres is largely consistent with the reference
 \gls{DNS} flow data and reproduces the expected trends in the velocity and stress distributions
 across the channel and as a function of the wall distance.
+%
+\begin{figure}[h!]
+  \centering
+  \includegraphics[width=.85\columnwidth]{channel-geom}
+  \caption{Channel geometry for the turbulent channel flow verification test. The red box indicates
+  the region shown by the close-up view in Figure \ref{fig:channel-mesh}.}
+  \label{fig:channel-geom}
+  \vspace{.5cm}
+  \includegraphics[width=\columnwidth]{channel_mesh}
+  \caption{Close-up view of the refined mesh near the inlet (left boundary) for the channel flow
+    test. The mesh for the rest of the channel follows the mesh resolution of the rightmost column
+  of elements.}
+  \label{fig:channel-mesh}
+\end{figure}
 
-\begin{figure}[htb!]
+\begin{figure}[p]
   \centering
@@
 \end{figure}
+
+\FloatBarrier
 
@@
 wall distances, and stresses from the experimental data \cite{laufer_structure_1954}
-and the Moltres Spalart-Allmaras model. The Moltres Spalart-Allmaras model is largely consistent with the reference
+and the Moltres Spalart-Allmaras model. Moltres is largely consistent with the reference
 experimental flow data and reproduces the expected trends in the velocity and stress distributions
@@
 
-\begin{figure}[htb]
+\begin{figure}[p]
   \centering
@@
 
-\begin{figure}[p]
-  \centering
-  \includegraphics[width=0.9\columnwidth]{backstep-geom}
-  \caption{Backward step geometry for the turbulent \gls{BFS} flow verification test. The red box indicates
-  the region shown by the close-up view in Figure \ref{fig:bfs-mesh}.}
-  \label{fig:backstep-geom}
-%\begin{figure}[htb!]
-  \centering
-  \includegraphics[width=0.9\columnwidth]{bfs_mesh}
-  \caption{Close-up view of the mesh for the \gls{BFS} flow test. The step is situated at $x=110$ m
-  with a height of $H=1$ m.}
-  \label{fig:bfs-mesh}
-\end{figure}
-
-The Moltres model for this test is a 2-D 160 m-long channel (Figure \ref{fig:backstep-geom}.
-The main flow direction is in the
-positive $x$ direction with the inlet and outlet on the left and right ends, respectively. The
-channel is 8 m tall before the 1 m-tall step at $x=110$ m. Figure \ref{fig:bfs-mesh} shows a
-close-up view of the mesh around the step. The dimensionless wall distance parameter $y^+$ of the
-first mesh elements along wall boundaries just prior to the step and at the end of the channel are
-0.735 and 0.555, respectively, meeting the $y^+ \lesssim 1$ requirement for properly wall-resolved
-flow. Table \ref{table:bfs} lists relevant flow parameters for the turbulent \gls{BFS} flow test.
-The Moltres \gls{BFS} flow model ran as two separate simulations for the upstream turbulent flow
-profile to fully develop from $x=0$ m to $x=104$ m and the \gls{BFS} flow profile from $x=104$ m to
-$x=160$ m. The fully-developed flow profile at $x=104$ m from the first simulation served as the
-inlet flow profile for the second simulation. Both simulations ran with adaptive time-stepping
-starting from a timestep size of 0.1 s up to a maximum timestep size of 10 s with a timestep size
-growth factor of 1.1. The upstream and \gls{BFS} simulations automatically terminated at $t=212$ s
-and $t=223$ s when Moltres detected that the flow profiles reached steady-state.
-
 \begin{table}[htb]
@@
 \end{table}
+
+The Moltres model for this test is a 2-D 160 m-long channel (Figure \ref{fig:backstep-geom}.
+The main flow direction is in the
+positive $x$ direction with the inlet and outlet on the left and right ends, respectively. The
+channel is 8 m tall before the 1 m-tall step at $x=110$ m. Figure \ref{fig:bfs-mesh} shows a
+close-up view of the mesh around the step. The dimensionless wall distance parameter $y^+$ of the
+first mesh elements along wall boundaries just prior to the step and at the end of the channel are
+0.735 and 0.555, respectively, meeting the $y^+ \lesssim 1$ requirement for properly wall-resolved
+flow. Table \ref{table:bfs} lists relevant flow parameters for the turbulent \gls{BFS} flow test.
+The Moltres \gls{BFS} flow model ran as two separate simulations for the upstream turbulent flow
+profile to fully develop from $x=0$ m to $x=104$ m and the \gls{BFS} flow profile from $x=104$ m to
+$x=160$ m. The fully-developed flow profile at $x=104$ m from the first simulation served as the
+inlet flow profile for the second simulation. Both simulations ran with adaptive time-stepping
+starting from a timestep size of 0.1 s up to a maximum timestep size of 10 s with a timestep size
+growth factor of 1.1. The upstream and \gls{BFS} simulations automatically terminated at $t=212$ s
+and $t=223$ s when Moltres detected that the flow profiles reached steady-state.
+
+\begin{figure}[p]
+  \centering
+  \includegraphics[width=0.87\columnwidth]{backstep-geom}
+  \caption{Backward step geometry for the turbulent \gls{BFS} flow verification test. The red box indicates
+  the region shown by the close-up view in Figure \ref{fig:bfs-mesh}.}
+  \label{fig:backstep-geom}
+%\begin{figure}[htb!]
+  \centering
+  \includegraphics[width=0.87\columnwidth]{bfs_mesh}
+  \caption{Close-up view of the mesh for the \gls{BFS} flow test. The step is situated at $x=110$ m
+  with a height of $H=1$ m.}
+  \label{fig:bfs-mesh}
+\end{figure}
 
@@
   \hfill
-  \begin{subfigure}[b]{0.38\columnwidth}
+  \begin{subfigure}[b]{0.369\columnwidth}
     \centering
@@
   \centering
-  \begin{subfigure}[b]{0.38\columnwidth}
+  \begin{subfigure}[b]{0.369\columnwidth}
     \centering
@@
   \caption{Comparison of backward facing step flow results against reference
-  experimental data and computational data from CFL3D. $x/H$ values are normalized horizontal
+  experimental data and computational data from CFL3D. $x/H$ values are normalized
   distances relative to the step.}

### 2025-01-03 acks.tex
Acknowledgements

This dissertation marks the end of a long journey that certainly was not possible on my own.
I would first like to thank my Ph.D.\ advisors, Prof.\ Kathryn D.\ Huff and Prof.\ Madicken Munk. I
am deeply grateful to Prof.\ Huff for her technical guidance on multiphysics reactor modeling \&
simulation and best practices in software development. I am equally grateful to Prof.\ Munk for her
guidance on neutron transport methods for
this dissertation. I am also grateful to both advisors for their support and advice on time
management, scientific writing, and other matters. In many ways, they created a welcoming
and supportive environment in the Advanced Reactors and Fuel Cycles (ARFC)
research group, without which this dissertation would not have been possible. I would also like to
thank my Ph.D.\ committee members, Prof.\ Rizwan Uddin, Prof.\ Tomasz Kozlowski, and Prof.\ Paul
Fischer, for teaching me much of what I know today about reactor physics and computational fluid
dynamics and for their guidance on this dissertation.

I am thankful to current and former groupmates in ARFC---Sam Dotson, Nathan Ryan, Luke Seifert,
Olek Yardas, Zoe Richter, Nathan Glaser, Gwendolyn Chee, Amanda Bachmann,
Andrei Rykhlevskii, Anshuman Chaube, and Greg Westphal---for their friendship and their help with
code and writing reviews. I am also thankful to
Alvin Lee, Xin Zhi Tan, Jeremy Mettler, and other friends from NPRE and UIUC
for the fun times we shared and for their help whenever I needed it.

I am especially thankful to my partner, Gwendolyn Chee, for her constant support and encouragement,
which spurred me on through the toughest times.
Finally, I want to thank my family for their unwavering love and support. I would not be where I am
today without them.

I would also like to acknowledge funding support for my graduate studies from the Singapore
Nuclear Safety \& Research Initiative (SNRSI) and the Department of Nuclear, Plasma \& Radiological
Engineering (NPRE) at the University of Illinois Urbana-Champaign (UIUC).

This research used resources of the Argonne Leadership Computing Facility, a U.S.\ Department of
Energy (DOE) Office of Science user facility at Argonne National Laboratory, and is based on
research supported by the U.S.\ DOE Office of Science---Advanced Scientific Computing Research
Program, under Contract No.\ DE-AC02-06CH11357.

### 2025-01-03 presentation/backup.tex
Final presentation

@@
 \begin{frame}[noframenumbering]
+  \begin{figure}[t]
+    \centering
+    \includegraphics[width=.65\columnwidth]{halfway-g1}
+    \caption{Group 1 neutron flux distribution with Rod 1 inserted by 27.5 inches.}
+    \label{fig:halfway-g1}
+  \end{figure}
+\end{frame}
+
+\begin{frame}[noframenumbering]
+  \begin{figure}[t]
+    \centering
+    \includegraphics[width=.65\columnwidth]{halfway-g8}
+    \caption{Group 8 neutron flux distribution with Rod 1 inserted by 27.5 inches.}
+    \label{fig:halfway-g8}
+  \end{figure}
+\end{frame}
+
+\begin{frame}[noframenumbering]
+  \begin{figure}[t]
+    \centering
+    \includegraphics[width=.65\columnwidth]{inserted-g1}
+    \caption{Group 1 neutron flux distribution with Rod 1 inserted by 51 inches.}
+    \label{fig:inserted-g1}
+  \end{figure}
+\end{frame}
+
+\begin{frame}[noframenumbering]
+  \begin{figure}[t]
+    \centering
+    \includegraphics[width=.65\columnwidth]{inserted-g8}
+    \caption{Group 8 neutron flux distribution with Rod 1 inserted by 51 inches.}
+    \label{fig:inserted-g8}
+  \end{figure}
+\end{frame}
+
+\begin{frame}[noframenumbering]
   \frametitle{MSRE Rod Drop Experiment}
